### 2021-08-12 paper/sections/numerical.tex
Set up the repo
- Add all functions and scripts necessary for the merger initialisation pipeline

- Improved parameter file design

\subsection{The KETJU Code}
\drafting{Basic overview and updates, refer to previous papers}

\subsection{Initial Conditions and Simulations}

### 2021-08-12 paper/sections/results.tex
Set up the repo
- Add all functions and scripts necessary for the merger initialisation pipeline

- Improved parameter file design

\label{sec:results}

\drafting{
Figures to do
\begin{itemize}
\item Good ones
\end{itemize}
Table of galaxy properties,
and maybe another one of BH merger remnant properties.
}

%\begin{figure*}
%\includegraphics{binary_params_main}
%\includegraphics{binary_params_secondary}
%\includegraphics{binary_params_tertiary}
%\caption{
%\drafting{
%Binary orbital parameters. Dashed line for triplet outer orbit (dashes not very %visible at the moment).
%Lower panels are mergers occurring in separate galaxies.
%Note the use of $1-e$ for the last panel.
%}
%}
%\label{fig:binary_params}
%\end{figure*}

### 2021-08-12 code/cm_functions/__init__.py

@@ -1,0 +1,7 @@
+from . import cosmology
+from . import general
+from . import initialise
+from . import mathematics
+from . import literature
+from . import plotting
+from . import utils

### 2021-08-12 code/cm_functions/cosmology.py

@@ -1,0 +1,39 @@
+import numpy as np
+import scipy.constants
+
+__all__ = ['cosmology', 'time2z']
+
+
+"""define the assumed cosmology constants, taken from planck 2018"""
+cosmology = dict(
+                    h = 0.6736,
+                    omega_L = 0.6847,
+                    omega_M = 0.3153
+)
+
+
+def time2z(t, H0=100*cosmology['h'], pres=False):
+    """
+    Estimate the redshift for a corresponding cosmic time
+    from Carmeli 2008
+
+    Parameters
+    ----------
+    t: cosmic time since big bang (Gyr) -> pres = False
+    t: time before present (Gyr) -> pres = True
+    H0: Hubble constant in km/s/Mpc
+    pres: bool, t is given from today not Big Bang
+
+    Returns
+    -------
+    z: redshift
+    """
+    if pres:
+        #determine cosmic time
+        ## TODO: make this better dependent on cosmology
+        t = 13.8 - t
+    #convert t [Gyr] to s
+    t = t * scipy.constants.year * 1e9
+    H0 = H0 / (1e3 * scipy.constants.parsec) #convert km/s/Mpc -> Hz
+    z = np.sqrt(2 / (H0 * t) - 1) - 1
+    return z

### 2021-08-12 code/cm_functions/general/__init__.py

@@ -1,0 +1,2 @@
+from .general import *
+from .pygad_helper import *

### 2021-08-12 code/cm_functions/general/general.py

@@ -1,0 +1,35 @@
+import numpy as np
+import scipy.optimize, scipy.special
+
+__all__ = ['arcsec_to_kpc', 'sersic_b_param']
+
+
+def arcsec_to_kpc(dist_mod, angle_in_arcsec):
+    """
+    Convert arcseconds to kpc using a distance modulus
+    Parameters
+    ----------
+    dist_mod: distance modulus
+    angle_in_arcsec: angle to convert
+
+    Returns
+    -------
+    distance in kpc
+    """
+    return np.tan(np.radians(angle_in_arcsec/3600)) * 10**(1+dist_mod/5)/1e3
+
+
+def sersic_b_param(n):
+    """
+    Determine the b parameter in the Sersic function
+    search interval given by n=[0.5,20] -> 2n-0.33+0.009876/n
+
+    Parameters
+    ----------
+    n: sersic index
+
+    Returns
+    -------
+    b parameter
+    """
+    return scipy.optimize.toms748(lambda t: 2*scipy.special.gammainc(2*n, t)-1, 0.6, 19.9)

### 2021-08-12 code/cm_functions/general/pygad_helper.py

@@ -1,0 +1,21 @@
+import pygad.units
+
+
+__all__ = ['convert_gadget_time']
+
+def convert_gadget_time(snap, new_unit='Gyr'):
+    """
+    Helper function to convert gadget unit time to physical time
+    1 unit time ~ 0.978 Gyr
+
+    Parameters
+    ----------
+    snap: gadget snapshot object
+    new_unit: time units, 'Gyr' or 's'
+
+    Returns
+    -------
+    float(t): numeric value of time in new_unit
+    """
+    t = pygad.UnitQty(snap.time, f"({snap.gadget_units['LENGTH']})/({snap.gadget_units['VELOCITY']})").in_units_of(new_unit, subs=snap)
+    return float(t)

### 2021-08-12 code/cm_functions/initialise/__init__.py

@@ -1,0 +1,2 @@
+from .galaxy_ic_class import *
+from .mergers import *

### 2021-08-12 code/cm_functions/initialise/galaxy_ic_class.py

@@ -1,0 +1,294 @@
+import copy
+import numpy as np
+import scipy.integrate, scipy.constants
+from functools import cached_property
+import os
+
+from ..utils import param_io
+from ..literature import *
+from ..cosmology import *
+from ..general import *
+#from . import astro_functions as kfa
+#from . import literature_functions as kfl
+#from . import general as kfg
+
+
+__all__ = ['galaxy_ic_base', 'ic_general_data', 'stellar_component', 'stellar_cuspy_ic', 'stellar_cored_ic', 'dm_component', 'dm_halo_NFW', 'dm_halo_dehnen', 'smbh']
+
+class galaxy_ic_base:
+    def __init__(self, parameter_file, stars=False, dm=False, bh=False):
+        self.general = ic_general_data(parameter_file)
+        if stars:
+            if parameter_file.stellarCored:
+                self.stars = stellar_cored_ic(parameter_file, self.general)
+            else:
+                self.stars = stellar_cuspy_ic(parameter_file, self.general)
+        else:
+            self.stars = None
+        if dm:
+            if parameter_file.use_NFW:
+                self.dm = dm_halo_NFW(parameter_file, self.stars)
+            else:
+                self.dm = dm_halo_dehnen(parameter_file, self.stars)
+        else:
+            self.dm = None
+        if bh:
+            self.bh = smbh(parameter_file, self.stars)
+        else:
+            self.bh = None
+
+    def to_gadget_mass_units(self, verbose=True):
+        if verbose:
+            print('Converting mass units to gadget')
+        assert(self.general.mass_units == 'msol')
+        self.stars.particle_mass /= 1e10
+        self.stars.total_mass /= 1e10
+        self.dm.particle_mass /= 1e10
+        self.dm.peak_mass = (False, self.dm.peak_mass/1e10)
+        self.bh.mass = (False, self.bh.mass/1e10)
+        self.general.mass_units = 'gadget'
+
+    def print_masses(self):
+        print('Printing masses')
+        print('Mass unit: {:s}'.format(self.general.mass_units))
+        print('Stellar particle mass: {:.3e}'.format(self.stars.particle_mass))
+        print('Stellar total mass: {:.3e}'.format(self.stars.total_mass))
+        print('DM particle mass: {:.3e}'.format(self.dm.particle_mass))
+        print('DM total mass: {:.3e}'.format(self.dm.peak_mass))
+        print('BH mass: {:.3e}'.format(self.bh.mass))
+
+
+class ic_general_data:
+    def __init__(self, parameter_file):
[…]

### 2021-08-12 code/cm_functions/initialise/mergers.py

@@ -1,0 +1,19 @@
+__all__ = ['e_from_rperi']
+
+
+def e_from_rperi(x, a=0.320, b=1.629, c=0.176):
+    """
+    Determine eccentricity from r/Rvir using fit to Khochfar & Burkett 2006
+    Fig. 6
+
+    Parameters
+    ----------
+    x: normalised rperi values (normalised to the virial radius of the larger
+        progenitor)
+    a, b, c = shape parameters
+
+    Returns
+    -------
+    eccentricity of approach
+    """
+    return (1 + (x/a)**b)**(-c)

### 2021-08-12 code/cm_functions/literature/__init__.py

@@ -1,0 +1,5 @@
+from .bh_bulge import *
+from .density_profiles import *
+from .dm_bulge import *
+from .radial_relations import *
+from .smbh_spins import *

### 2021-08-12 code/cm_functions/literature/bh_bulge.py

@@ -1,0 +1,75 @@
+import numpy as np
+
+__all__ = ['Haring04', 'Magorrian98', 'Sahu19', 'Scott13']
+
+
+def Haring04(logMstar):
+    '''
+    define Haring+04 BH-Bulge fit function
+
+    Parameters
+    ----------
+    logMstar: log of bulge mass [Msol]
+
+    Return
+    ------
+    predicted log of bh mass [Msol]
+    '''
+    assert(logMstar < 13)
+    return 8.20 + 1.12 * (logMstar - 11)
+
+
+def Magorrian98(logMstar):
+    """
+    define Magorrian+98 BH-Bulge fit function
+
+    Parameters
+    ----------
+    logMstar = log of bulge mass [Msol]
+
+    Returns
+    -------
+    predicted log of bh mass [Msol]
+    """
+    assert(logMstar < 13)
+    return -1.79 + 0.96 * logMstar
+
+
+def Sahu19(logmstar, u='Ks'):
+    """
+    define the Sahu+19 bulge mass - BH mass relation
+    # TODO: define the other two u values
+
+    Parameters
+    ----------
+    logmstar: log of bulge (spheroidal) mass
+    u: one of Ks, scaling factor
+
+    Returns
+    -------
+    log BH Mass
+    """
+    if u == 'Ks':
+        uval = 10**(-0.06 * (logmstar-10) - 0.06)
+    else:
+        raise NotImplementedError('Only Ks currently implemented')
+    return 1.27 * (logmstar - np.log10(uval * 5e10)) + 8.41
+
+
+def Scott13(logmstar, cored=False):
+    """
+    define the Scott+13 M_bh - M_bulge scaling relations for cored and non-
+    cored galaxies
+    Parameters
+    ----------
+    logmstar: stellar mass, log Msol
+    cored: (bool), is the galaxy cored?
+
+    Returns
+    -------
+    log of BH mass
+    """
+    if cored:
+        return 0.97 * (logmstar - np.log10(3e11)) + 9.27
+    else:
+        return 2.22 * (logmstar - np.log10(3e10)) + 7.89

### 2021-08-12 code/cm_functions/literature/density_profiles.py

@@ -1,0 +1,93 @@
+import numpy as np
+import scipy.optimize
+
+
+__all__ = ['Dehnen', 'fit_Dehnen_profile', 'halfMassDehnen', 'Terzic05']
+
+
+def Dehnen(r, a, g, M):
+    '''
+    Calculate the Dehnen density profile
+
+    Parameters
+    ----------
+    r: radius [kpc]
+    a: scale radius [kpc]
+    g: shape profile
+    M: total Mass [Msol]
+
+    Returns
+    -------
+    rho: density
+    '''
+    rho = (3 - g) * M / (4 * np.pi) * a / (r**g * (r + a)**(4 - g))
+    return rho
+
+
+def fit_Dehnen_profile(radii, density, total_mass, bounds=([1, 0], [1000,3])):
+    """
+    Fit a Dehnen profile to some data using the scipy.optimize library.
+
+    Parameters
+    ----------
+    radii: radii values of the data
+    density: 3D density values of the data (not surface density)
+    total_mass: total mass in units consistent with radii and density
+    bounds: parameter limits as a tuple of two lists of the form
+            ([lower scale radius, lower gamma], [upper scale radius, upper
+            gamma])
+
+    Returns
+    -------
+    param_best: list of best-fit parameters
+    """
+    #fit the curve
+    param_best, param_cov = scipy.optimize.curve_fit(lambda r, a, g: Dehnen(r, a, g, total_mass), radii, density, bounds=bounds)
+    return param_best
+
+
+def halfMassDehnen(a, g):
+    """
+    Determine analytical half mass radius from Dehnen sphere
+    taken from Rantala+18
+
+    Parameters
+    ----------
+    a: scale radius (generally in kpc)
+    g: Dehnen gamma
+
+    Returns
+    -------
+    rhm: half mass radius [kpc]
+    re: effective radius approximate [kpc]
+    """
+    rhm = a * (2**(1/(3-g)) - 1)**(-1)
+    re = 0.75 * rhm
+    return rhm, re
+
+
+def Terzic05(r, rhob, rb, n, g, b, Re, a=100):
+    """
+    Define a density function for a cored system, with overall sersic profile
+    as taken from Terzic+05
+    # TODO: change mass at break radius to mass deficit?
+
+    Parameters
+    ----------
+    r = radius to evaluate at [kpc]
+    rhob = density at break radius
[…]

### 2021-08-12 code/cm_functions/literature/dm_bulge.py

@@ -1,0 +1,146 @@
+import numpy as np
+import scipy.optimize
+
+__all__ = ['Behroozi19', 'Girelli20', 'Moster10']
+
+
+
+###### HELPER FUNCTIONS #######
+def _moster10(h, M1, mM0, b, g):
+    """the form of the moster 10 function for dm-stellar mass relation"""
+    return 2 * mM0 * ((h / M1)**(-b) + (h / M1)**g)**(-1) * h
+
+def _moster_helper(h, s, M1, mM0, b, g):
+    """helper function for bisection method (could be an anonymous lambda
+    function, but is called a few times)"""
+    return s - _moster10(h, M1, mM0, b, g)
+
+
+###### CALLABLE FUNCTIONS #######
+
+def Behroozi19(sm, hm=[1e10, 1e15], z=0, plotting=False, numPoints=1000):
+    """
+    define the Behroozi+19 relation for stellar-halo mass.
+
+    Parameters
+    ----------
+    sm: stellar mass in Msol
+    hm: list of estimated bounds halo mass falls within
+    z: assumed redshift, for redshift dependent scaling relation
+    plotting: recover the halo mass, or return an array of values for plotting?
+    numPoints: number of points used in bisection method (and plotting) to
+               determine halo mass from stellar mass
+
+    Returns
+    -------
+    log10(h_mass): logarithm of halo mass
+    logMstar: logarithm of corresponding stellar masses (if plotting = True)
+    """
+    assert(len(hm)==2 and hm[0]<hm[1])
+    M0 = 12.069
+    Ma = 2.646
+    Mlna = 2.710
+    Mz = -0.431
+    eps0 = -1.480
+    epsa = -0.831
+    epslna = -1.351
+    epsz = 0.321
+    alpha0 = 1.899
+    alphaa = -2.901
+    alphalna = -2.413
+    alphaz = 0.332
+    beta0 = 0.502
+    betaa = -0.315
+    betaz = -0.218
+    delta0 = 0.397
+    gamma0 = -0.867
+    gammaa = -1.146
+    gammaz = -0.294
+    a = 1/(1+z)
+
+    sm = np.log10(sm)
+    logM1 = M0 + Ma * (a-1) - Mlna * np.log(a) + Mz * z
+    epsilon = eps0 + epsa*(a-1) - epslna * np.log(a) + epsz*z
+    alpha = alpha0 + alphaa * (a-1) - alphalna*np.log(a) + alphaz*z
+    beta = beta0 + betaa*(a-1) + betaz*z
[…]

### 2021-08-12 code/cm_functions/literature/radial_relations.py

@@ -1,0 +1,21 @@
+import numpy as np
+
+"""
+A collection of miscellaneous relations that are somehow dependent on a radial
+value. 
+"""
+__all__ = ['Sahu20']
+
+
+def Sahu20(logRe):
+    """
+    Define the Sahu+20 spheroidal mass - Re relation
+    Parameters
+    ----------
+    logRe: log of effective radius in kpc
+
+    Returns
+    -------
+    expected log spheroidal mass
+    """
+    return 1.08 * logRe + 10.32

### 2021-08-12 code/cm_functions/literature/smbh_spins.py

@@ -1,0 +1,6 @@
+__all__ = ['zlochower_dry_spins']
+
+zlochower_dry_spins = dict(
+                            spin_mag_a = 10.5868,
+                            spin_mag_b = 4.66884
+)

### 2021-08-12 code/cm_functions/mathematics/__init__.py

@@ -1,0 +1,1 @@
+from .array_operations import *

### 2021-08-12 code/cm_functions/mathematics/array_operations.py

@@ -1,0 +1,50 @@
+import numpy as np
+
+
+__all__ = ['find_index', 'get_histogram_bin_centres']
+
+
+def find_index(val, arr, tol=1e-7):
+    """
+    General method to find the index of a given value in a array,
+    even if that value is potentially not in the array
+    #TODO if val not in arr, ensure entire arr is NOT returned
+
+    Parameters
+    ----------
+    val: value to find
+    arr: array to search
+    tol: comparison tolerance: if min(arr-val)< tol -> val in arr
+
+    Returns
+    -------
+    idx = position of the value closest to val in arr
+
+    Raises
+    ------
+    AssertionError: val not an int or a float
+    RuntimeError: val is not in arr
+    """
+    assert(isinstance(val, float) or isinstance(val, int))
+    #determine which value in arr is closest to val and its position
+    mindiff = np.nanmin(np.abs(val-arr))
+    #if difference larger than specified tolerance, val isn't really in arr
+    if mindiff > tol:
+        raise RuntimeError('The value {:.3e} was not found within error {:.3e}!\nThe minimum difference between input <val> and elements in <arr> is {:.3e}'.format(val, tol, mindiff))
+    idx = np.where(np.abs(val-arr)-mindiff < tol)[0][0]
+    return idx
+
+
+def get_histogram_bin_centres(bins):
+    """
+    Convenience function to get the centres of some histogram bins.
+
+    Parameters
+    ----------
+    bins: array of bin edges
+
+    Returns
+    -------
+    the bin centres
+    """
+    return (bins[:-1] + bins[1:]) / 2

### 2021-08-12 code/cm_functions/plotting/__init__.py

@@ -1,0 +1,2 @@
+from .general import *
+from .specific_plots import *

### 2021-08-12 code/cm_functions/plotting/general.py

@@ -1,0 +1,53 @@
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.anchored_artists import AnchoredSizeBar
+import scipy.interpolate
+
+
+__all__ = ['draw_sizebar', 'mplColours']
+
+
+def draw_sizebar(ax, length, units, location='lower right', pad=0.1, borderpad=0.5, sep=5, frameon=False, unitconvert=None):
+    '''
+    Draw a horizontal scale bar using the mpl toolkit
+
+    Parameters
+    ----------
+    ax: axis to add the bar to
+    length: length of scale bar in data units
+    units: string stating unit name
+    location: where to place bar (standard pyplot location string)
+    pad: padding around label
+    borderpad: padding around border
+    sep: separation between label and scale bar
+    frameon: draw box around scale bar
+    unitconvert: convert units of scalebar
+
+    Returns
+    -------
+        None (draw scale bar)
+    '''
+    if unitconvert is None:
+        label = str(length)+' '+units
+    elif unitconvert == 'kilo2base':
+        label = str(length*1000)+' '+units
+    else:
+        #TODO other unit conversions
+        raise NotImplementedError('Other unit conversions yet to be implemented')
+    asb = AnchoredSizeBar(ax.transData, length, label, loc=location, pad=pad, borderpad=borderpad, sep=sep, frameon=frameon)
+    ax.add_artist(asb)
+
+
+def mplColours():
+    """
+    access the default matplotlib color palette by index
+    maximum index:
+
+    Parameters
+    ----------
+    None
+
+    Returns
+    -------
+    the colour array, to be used in plt.plot(x,y,color=THIS)
+    """
+    return plt.rcParams['axes.prop_cycle'].by_key()['color']

### 2021-08-12 code/cm_functions/plotting/specific_plots.py

@@ -1,0 +1,71 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.stats
+
+
+__all__ = ['plot_parameter_contours']
+
+
+def plot_parameter_contours(ax, fun, xvals, yvals, init_lims, data_err=None, args=(), numPoints=350, repeats=1, slope=1.2, sigma_level=3):
+    """
+    Plot 2D parameter contours from chi2 test.
+    # TODO: move chi2 calculation to mathematics submodule?
+
+    Parameters
+    ----------
+    ax: the matplotlib axis to plot to
+    fun: function to evaluate
+    xvals: x coordinates to evaluate the function at
+    yvals: y coordinates of data
+    args: other arguments to the function to be evaluated as *args
+    data_err: error in observed data
+    init_limits: tuple of 2 arrays, each array with the min/max of a parameter
+    numPoints: number of points to scan for each parameter
+    repeats: allow for reducing search interval by zeroing in this many times
+    slope: amount to decrease the search range by each repeat
+    sigma_level: number of sigma levels to plot
+
+    Returns
+    -------
+    None
+    """
+    assert(repeats > 0 and isinstance(repeats, int))
+    print('Creating parameter contours...')
+    if data_err is None:
+        data_err = 0.3*yvals+1e-4
+    for repeat in range(repeats):
+        print('  Level: {:d}'.format(repeat+1))
+        print('    Parameter 1 Limits: {:.2e} - {:.2e}'.format(init_lims[0][0], init_lims[0][1]))
+        print('    Parameter 2 Limits: {:.2e} - {:.2e}'.format(init_lims[1][0], init_lims[1][1]))
+        param1_seq = np.linspace(init_lims[0][0], init_lims[0][1], numPoints)
+        param2_seq = np.linspace(init_lims[1][0], init_lims[1][1], numPoints)
+
+        #initialise chi2 array
+        chi_array = np.full((numPoints, numPoints), np.nan)
+
+        #scan the 2D parameter space
+        for ind1, param1 in enumerate(param1_seq):
+            for ind2, param2 in enumerate(param2_seq):
+                fun_val = fun(xvals, param1, param2, *args)
+                chi_array[ind1][ind2] = kfm.chi_square(yvals, fun_val, data_err)
[…]

### 2021-08-12 code/cm_functions/utils/__init__.py

@@ -1,0 +1,3 @@
+from .param_io import *
+from .dataset_operations import *
+from .get_cl_args import *

### 2021-08-12 code/cm_functions/utils/dataset_operations.py

@@ -1,0 +1,23 @@
+import numpy as np
+import pandas as pd
+
+__all__ = ['create_error_col']
+
+
+def create_error_col(dat, col, splitat='+/-'):
+    """
+    Split a pandas dataframe column with an error into two columns
+
+    Parameters
+    ----------
+    dat: data frame object
+    col: column to split
+    splitat: the character combination to split the entry at
+
+    Return
+    ------
+    None: new data column is appended to dataframe with column name [col]_ERR
+    """
+    val_err = np.array([d.split(splitat) for d in dat.loc[:,col]], dtype='float')
+    dat.loc[:,col] = val_err[:,0]
+    dat.insert(len(dat.columns), col+'_ERR', val_err[:,1])

### 2021-08-12 code/cm_functions/utils/get_cl_args.py

@@ -1,0 +1,26 @@
+import argparse
+
+__all__ = ['argparse_for_initialise']
+
+
+def argparse_for_initialise(description='', update_help=None):
+    """
+    Get the command line arguments necessary to run the initialisation scripts,
+    as they all have a similar format with similar options.
+
+    Parameters
+    ----------
+    description: main description of program invoked with help flag
+    update_help: description of which parameters can be updated. A value
+                 of None excludes the argument
+
+    Returns
+    -------
+    parser object that can have other options added to it
+    """
+    parser = argparse.ArgumentParser(description=description, allow_abbrev=False)
+    parser.add_argument(type=str, help='path to parameter file', dest='paramFile')
+    if update_help is not None:
+        parser.add_argument('-u', '--update', dest='parameter_update', help=update_help, action='store_true', default=False)
+    parser.add_argument('-v', '--verbose', dest='verbose', action='store_true', help='verbose printing in script')
+    return parser

### 2021-08-12 code/cm_functions/utils/param_io.py

@@ -1,0 +1,108 @@
+import re
+import sys
+import os
+import importlib
+import numpy as np
+
+
+__all__ = ['read_parameters', 'write_parameters']
+
+def read_parameters(filepath, verbose=True):
+    """
+    Read parameters from a python file as if it were a module. If the invocation is:
+        param_vals = read_parameters(file)
+    then the parameter file values can be accessed as param_vals.value
+    Strings have trailing '/' characters removed, and all lists are converted
+    to class <np.ndarray>
+
+    Parameters
+    ----------
+    filepath: absolute or relative path to the parameter file
+    verbose: print the name of the file being read
+
+    Returns
+    -------
+    the imported parameter file as a module
+    """
+    name_ext = filepath.split('/')[-1]
+    sys.path.append(os.path.dirname(os.path.abspath(filepath)))
+    module_name, module_ext = name_ext.split('.')
+    if module_ext != 'py':
+        raise ValueError('Input file must have .py extension!')
+    else:
+        if verbose:
+            print('Reading parameters from: {}'.format(filepath))
+        params = importlib.import_module(module_name)
+        #remove trailing '/' characters from strings
+        for p in dir(params):
+            if p[:2] == '__':
+                continue
+            v = getattr(params, p)
+            if isinstance(v, str) and v[-1].rstrip()=='/':
+                setattr(params, p, v.rstrip('/'))
+            elif isinstance(v, list):
+                setattr(params, p, np.array(v, dtype='float64'))
+        return params
+
+
+def write_parameters(values, filepath=None, verbose=True, allow_updates=()):
+    """
+    Write parameters that have been loaded with read_parameters() to a file.
+
+    Parameters
+    ----------
+    values: the module name which contains the parameters
+    filepath: the file to save the parameters to. If None, values.__file__ is
+              used
+    verbose: print the name of the file being written to
+    allow_updates: tuple of parameter names as strings that are allowed to be
+                   updated (prevents accidental overwriting)
+    """
+    if filepath is None:
+        filepath = values.__file__
+    new_vars = False
+    with open(filepath, 'r+') as f:
+        contents = f.read()
+        for var in dir(values):
+            if var[:2] != '__':
+                value = getattr(values, var)
+                #include any potential comments
+                line = re.search(r'\b{}\b.*'.format(var), contents)
+                if line is not None:
+                    #the variable exists in the parameter file
+                    if var not in allow_updates:
+                        # but we don't want to update this value
+                        continue
[…]

### 2021-08-12 code/figure_scripts/figure_config.py

@@ -1,0 +1,21 @@
+import os
+import matplotlib
+import sys
+
+this_dir = os.path.dirname(os.path.realpath(__file__))
+#data_dir = os.path.abspath(os.path.join(this_dir, '../data/'))
+raw_data_dir = '/scratch/pjohanss/arawling/collisionless-merger/'
+processed_data_dir = '../../data/'
+figure_dir = os.path.abspath(os.path.join(this_dir, '../figures/'))
+figure_data_dir =  os.path.abspath(os.path.join(data_dir, 'figure_data/'))
+
+#allow importing common analysis scripts/modules
+#sys.path.append(os.path.abspath(os.path.join(this_dir, '../analysis_scripts')))
+
+# get the complete path for saving a figure
+def fig_path(fname):
+    return os.path.join(figure_dir, fname)
+
+# make sure we're using the same settings
+matplotlib.rcdefaults()
+matplotlib.rc_file(os.path.join(this_dir,'matplotlibrc'))

### 2021-08-12 code/initialise_scripts/bh_dm_mass.py

@@ -1,0 +1,94 @@
+import argparse
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import os
+import sys
+
+import cm_functions as cmf
+
+
+#get the command line arguments
+parser = cmf.utils.argparse_for_initialise(description='Determine the SMBH and DM Halo masses as a function of the stellar mass.', update_help='allow updates of SMBH mass, DM mass, and redshift')
+args = parser.parse_args()
+
+pfv = cmf.utils.read_parameters(args.paramFile)
+
+print('\nRunning bh_dm_mass.py\n')
+
+galaxy = cmf.initialise.galaxy_ic_base(pfv, stars=True, dm=True, bh=True)
+
+galaxy.dm.peak_mass = (args.verbose, None)
+galaxy.bh.mass = (args.verbose, None)
+
+markersz = 1.5
+linewd = 1
+
+
+#read in literature data
+mass_data = pd.read_table(galaxy.general.lit_location + '/' + pfv.massData, sep=',')
+bh_data = pd.read_table(galaxy.general.lit_location + '/' + pfv.bulgeBHData, sep=',', header=0)
+#restrict to only ETGs (exclude also S0)
+bh_data = bh_data.loc[np.logical_or(bh_data.loc[:,'Type']=='E', bh_data.loc[:,'Type']=='ES'), :]
+
+cored_galaxies = np.zeros(bh_data.shape[0], dtype='bool')
+for ind, gal in enumerate(bh_data.loc[:,'Galaxy']):
+    if gal[-1] == 'a':
+        cored_galaxies[ind] = 1
+bh_data.insert(2, 'Cored', cored_galaxies)
+
+cmf.utils.create_error_col(bh_data, 'logM*_sph')
+cmf.utils.create_error_col(bh_data, 'logMbh')
+
+fig, (ax1, ax2, ax3) = plt.subplots(1, 3, figsize=(8,4.5))
+
+#plot bulge mass distribution
[…]

### 2021-08-12 code/initialise_scripts/create_select_galaxies.sh

@@ -1,0 +1,43 @@
+#!/usr/bin/env bash
+
+if [ $# -eq 0 ]; then
+    echo "No galaxy name provided!"
+    echo "usage:"
+    echo "    ./create_select_galaxies.sh path name1 name2 ... nameN"
+    echo "path: path to parameter file directory"
+    echo "name_: parameter file name"
+    exit 1
+fi
+
+#create initial conditions for the specified galaxies
+param_dir=$1
+#add a slash if not in the path
+if [[ ${param_dir: -1} != "/" ]]
+then
+    param_dir="$param_dir/"
+fi
+
+#run the pipeline
+echo "------------------------------------"
+echo "Running the initialisation pipeline"
+echo "------------------------------------"
+
+first_arg_flag=1
+file_ext=""
+for galaxy in $@
+do
+    if [ $first_arg_flag -eq 1 ]
+    then
+        first_arg_flag=0
+        continue
+    elif [[ ${galaxy: -3} != ".py" ]]
+    then
+        file_ext=".py"
+    fi
+    sim="$param_dir$galaxy$file_ext"
+    echo "Running: $sim"
+    python ./bh_dm_mass.py  $sim -v
+    python ./galaxy_gen.py $sim -v
+    python ./ic_kinematics.py $sim -v
+    echo $'\n'
+done

### 2021-08-12 code/initialise_scripts/galaxy_gen.py

@@ -1,0 +1,109 @@
+import argparse
+import matplotlib.pyplot as plt
+import numpy as np
+import scipy.stats
+import sys
+
+import merger_ic_generator as mg
+import cm_functions as cmf
+
+
+#get the command line arguments
+parser = cmf.utils.argparse_for_initialise(description='Generate the galaxy initial conditions.', update_help='allow updates of SMBH spin and other derived quantities')
+args = parser.parse_args()
+
+pfv = cmf.utils.read_parameters(args.paramFile)
+parameters_to_update=[]
+
+print('\nRunning galaxy_gen.py\n')
+
+galaxy = cmf.initialise.galaxy_ic_base(pfv, stars=True, dm=True, bh=True)
+
+galaxy.dm.peak_mass = (args.verbose, None)
+galaxy.bh.mass = (args.verbose, None)
+#seed the random generator
+np.random.seed(galaxy.general.seed)
+
+if pfv.BH_spin_from == 'zlochower_dry':
+    bh_spin_params = cmf.literature.zlochower_dry_spins
+else:
+    raise NotImplementedError('Dry Spins only implemented currently')
+
+
+#set SMBH spin
+if isinstance(pfv.BH_spin, str):
+    #assign random spin
+    spin_mag = scipy.stats.beta.rvs(bh_spin_params['spin_mag_a'], bh_spin_params['spin_mag_b'], size=1)
+    spin_dir_theta = 2*np.pi*scipy.stats.uniform.rvs(size=1)
+    spin_dir_phi = np.arccos(2*scipy.stats.uniform.rvs(size=1) - 1)
+    galaxy.bh.spin = np.array([
+        spin_mag * np.sin(spin_dir_phi) * np.cos(spin_dir_theta),
+        spin_mag * np.sin(spin_dir_phi) * np.sin(spin_dir_theta),
+        spin_mag * np.cos(spin_dir_phi)
+    ]).flatten()
+    parameters_to_update.append('BH_spin')
+else:
+    galaxy.bh.spin = pfv.BH_spin
+
+
+#need to convert units to gadget units!!
+galaxy.to_gadget_mass_units()
+if args.verbose:
+    galaxy.print_masses()
+
+#create galaxy
+if isinstance(galaxy.stars, cmf.initialise.stellar_cored_ic):
+    #dictionary for density function
+    df_kwargs = dict(
+        rhob = galaxy.stars.core_density,
+        rb=galaxy.stars.core_radius,
+        n=galaxy.stars.sersic_index,
[…]

### 2021-08-12 code/initialise_scripts/ic_kinematics.py

@@ -1,0 +1,271 @@
+import argparse
+import numpy as np
+import matplotlib.pyplot as plt
+from mpl_toolkits.axes_grid1.inset_locator import inset_axes
+import h5py
+import scipy.stats
+import pandas as pd
+
+import cm_functions as cmf
+import pygad
+
+
+#get the command line arguments
+parser = cmf.utils.argparse_for_initialise(description='Analyse the kinematics of the initial conditions.', update_help='allow updates of kinematic quantities')
+parser.add_argument('--plotcontours', dest='plot_contours', help='plot contours of scale radius and gamma for Dehnen models to show degeneracy', action='store_true')
+parser.add_argument('--numrotations', dest='num_rots', help='number of rotations to use for averaging projected quantities', default=10)
+args = parser.parse_args()
+
+print('\nRunning ic_kinematics.py\n')
+
+pfv = cmf.utils.read_parameters(args.paramFile)
+
+galaxy = cmf.initialise.galaxy_ic_base(pfv, stars=True, dm=True, bh=True)
+galaxy.dm.total_mass = (args.verbose, None)
+galaxy.bh.mass = (args.verbose, None)
+
+#this is the "snapshot"
+icfile = galaxy.general.save_location + '/' + galaxy.general.name +'.hdf5'
+
+try:
+    virialRadius = pfv.virial_radius
+    print('Using a previously defined virial radius')
+    find_vir_rad_flag = False
+except AttributeError:
+    print('Virial radius will be searched for using maximum radius.')
+    virialRadius = galaxy.general.maximum_radius
+    find_vir_rad_flag = True
+
+markersz = 1.5
+linewd = 1
+capsize=2
+sim_col = 'tab:blue'
+fit_col = ['tab:orange', 'tab:red', 'tab:green']
+erralpha=0.9
+
+#load literature data
+bulgeBHData = pd.read_table(galaxy.general.lit_location + '/' + pfv.bulgeBHData, sep=',', header=0)
+#restrict to only ETGs (exclude also S0)
[…]

### 2021-08-12 code/initialise_scripts/literature_data/prep_cappellari.py

@@ -1,0 +1,20 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas as pd
+import ketju_functions as kf
+
+cappellari_13_a = pd.read_fwf('cappellari_13_a.txt', header=0, na_values=('--', '----'))
+cappellari_13_b = pd.read_fwf('cappellari_13_b.txt', header=0, na_values=('--', '----'))
+cappellari = cappellari_13_a.merge(cappellari_13_b, on='Galaxy', how='inner')
+cappellari.dropna(inplace=True)
+
+#print(cappellari.columns)
+
+z = 0
+mstar = (10**(cappellari.loc[:,'logML_star'].astype(float) + cappellari.loc[:,'logLum'].astype(float))).to_numpy()
+#print(mstar)
+
+peak_mass = np.full_like(mstar, np.nan)
+for ind, m in enumerate(mstar):
+    peak_mass[ind] = 10**kf.Behroozi19(m, [1e8, 1e15], z=0, plotting=False)
+print(peak_mass)

### 2021-08-12 code/initialise_scripts/merger_setup.py

@@ -1,0 +1,71 @@
+import argparse
+import numpy as np
+import matplotlib.pyplot as plt
+
+import pygad
+import merger_ic_generator as mg
+import cm_functions as cmf
+
+
+#get the command line arguments
+parser = cmf.utils.argparse_for_initialise(description='Generate the merger from two generated initial conditions.', update_help='allow updates of used virial_radius, r0, rperi, and e')
+parser.add_argument('-p', '--plot', dest='plot', help='plot the merger setup', action='store_true')
+args = parser.parse_args()
+
+pfv = cmf.utils.read_parameters(args.paramFile)
+
+print('\nRunning merger_gen.py\n')
+
+galaxy1 = mg.SnapshotSystem(pfv.file1)
+galaxy2 = mg.SnapshotSystem(pfv.file2)
+
+if 'virial' in pfv.initialSeparation:
+    pfv.virial_radius = -10
+    for i, g in enumerate((galaxy1, galaxy2), start=1):
+        gal = pygad.Snapshot(getattr(pfv, 'file{}'.format(i)))
+        gal.to_physical_units()
+        centre_guess = pygad.analysis.center_of_mass(gal.stars)
+        #refine the center estimate with shrinking sphere
+        center_ss = pygad.analysis.shrinking_sphere(gal.stars, centre_guess, 25.0)
+        this_virial_radius, virial_mass = pygad.analysis.virial_info(gal, center=center_ss)
+        if this_virial_radius > pfv.virial_radius:
+            pfv.virial_radius = this_virial_radius.view(np.ndarray)
+    r0_frac = float(pfv.initialSeparation[6:])
+    pfv.r0 = r0_frac * pfv.virial_radius
+else:
+    raise NotImplementedError
+
+if 'virial' in pfv.pericentreDistance:
+    rperi_frac = float(pfv.pericentreDistance[6:])
+    pfv.rperi = rperi_frac * pfv.virial_radius
+else:
+    raise NotImplementedError
+
+#get the eccentricity of the approach
+pfv.e = cmf.initialise.e_from_rperi(pfv.rperi/pfv.virial_radius)
+
+merger = mg.Merger(galaxy1, galaxy2, pfv.r0, pfv.rperi, e=pfv.e)
+
[…]

### 2021-08-12 code/initialise_scripts/stability.py

@@ -1,0 +1,102 @@
+import argparse
+import os
+import numpy as np
+import matplotlib.pyplot as plt
+import h5py
+import pandas as pd
+
+import cm_functions as cmf
+import pygad
+
+
+#get the command line arguments
+parser = cmf.utils.argparse_for_initialise(description='Analyse the stability of the initial conditions.')
+parser.add_argument('-q', '--quantilecut', dest='quantile_cut', help='remove data that is above this \nradial quantile', default=1)
+parser.add_argument('-p', '--percentages', dest='percentages_to_check', help='find the radius corresponding to these percentages of mass (can be specified multiple times)', action='append', default=None, type=float)
+args = parser.parse_args()
+
+print('\nRunning stability.py\n')
+
+pfv = cmf.utils.read_parameters(args.paramFile)
+#population percentages
+if args.percentages_to_check is None:
+    args.percentages_to_check = [5, 10, 30, 50, 80]
+linewd = 2
+print(args.percentages_to_check)
+
+outpath = pfv.saveLocation + '/' + pfv.galaxyName
+datapath = outpath + '/' + pfv.dataLocation
+figPath = outpath + '/' + pfv.figureLocation
+
+datfiles = []
+#get the data files and strip new line character
+with os.scandir(datapath) as f:
+    for f_ind in f:
+        if f_ind.name.endswith('.hdf5') and f_ind.name != 'ketju_bhs.hdf5':
+            datfiles.append(f_ind.path)
+datfiles.sort()
+
+#create data frame
+stability_data = pd.DataFrame(data={'Time': np.full_like(datfiles, np.nan)})
+for ind, col in enumerate(args.percentages_to_check):
+    stability_data.insert(1+ind*2, 'stars_'+str(col), np.full_like(datfiles, np.nan))
+    stability_data.insert(2+ind*2, 'dm_'+str(col), np.full_like(datfiles, np.nan))
+
+for ind, this_file in enumerate(datfiles):
+    print('Reading: ' + this_file)
+    gal = pygad.Snapshot(this_file)
+    gal.to_physical_units()
+    stability_data.loc[ind, 'Time'] = cmf.general.convert_gadget_time(gal) * 1e3
+    particle_families = []
+    if 'stars' in gal:
+        particle_families.append('stars')
+        mass_centre = pygad.analysis.shrinking_sphere(gal.stars, pygad.analysis.center_of_mass(gal.stars), 25.0)
[…]

### 2021-08-12 parameters/parameters-galaxies/mytest.py

@@ -1,0 +1,77 @@
+#----------------------------------user input
+
+#always assume units of:
+#mass: solar masses
+#distance: kpc
+#time: Gyr
+#velocity: km/s
+#unless otherwise stated
+
+
+#---- general
+simulationTime = 2.0                     #total simulation time
+randomSeed = 265200                      #random seed
+galaxyName = "test"                      #file name
+
+
+#---- file information
+saveLocation = "/Volumes/Rawlings_Storage/KETJU/initialise/"  #parent save directory
+figureLocation = "figures"         #figures saved in saveLocation/figureLocation
+dataLocation = "output"            #simulation output saved here
+litDataLocation = "literature_data"    #path to where literature data is
+
+
+#---- stellar properties
+stellarCored = 0
+logStellarMass = 11.4                    #total log10 stellar mass
+stellarParticleMass = 1e5                #mass of stellar particle
+stellarScaleRadius = 3.2                 #star Dehnen scale radius
+stellarGamma = 1.4                       #star Dehnen gamma
+maximumRadius = 8000                     #maximum galaxy radius
+minimumRadius = 1.5e-3                   #prevent tight orbits
+anisotropyRadius = 0.4                   #Osipkov Merrit anisotropy radius
+stellar_softening = 3.5e-3               #stellar softening to be used
+
+
+#---- halo properties
+use_NFW = 1                              #use NFW profile, otherwise Dehnen
+DMParticleMass = 6e7                     #mass of halo particle
+DM_softening = 0.1                       #DM softening to be used
+DM_mass_from = "Behroozi"                #which scaling relation to use
+
+
+#---- black hole properties
+BH_softening = 3e-3                      #BH softening length
+BH_spin_from = 'zlochower_dry'           #spin distribution to use
+BH_spin = [-0.76785,+0.33621,-0.0172 ]  #BH spin
+real_BH_mass = 8.94                      #measured log BH mass
+
+
+#---- literature files
+bulgeBHData = "sahu_20.txt"           #BH - bulge relation data
+massData = "sdss_z.csv"               #stellar mass distribution
+BHsigmaData = "bosch_16.txt"          #BH - sigma relation data
+fDMData = "jin_2020.dat"              #inner dm fraction data
+
+
+#----------------------------------returned values
[…]

### 2021-08-12 parameters/parameters-mergers/NGCa0524b_NGCa3348.py

@@ -1,0 +1,36 @@
+#----------------------------------user input
+
+#always assume units of:
+#mass: solar masses
+#distance: kpc
+#time: Gyr
+#velocity: km/s
+#unless otherwise stated
+
+
+#---- general
+galaxyName1 = "test1"                       #galaxy 1 name
+galaxyName2 = "test2"                       #galaxy 2 name
+
+
+#--- file location
+file1 = "/Volumes/Rawlings_Storage/KETJU/initialise/galaxies/NGCa0524b/NGCa0524b.hdf5"
+file2 = "/Volumes/Rawlings_Storage/KETJU/initialise/galaxies/NGCa3348/NGCa3348.hdf5"
+saveLocation = "/Volumes/Rawlings_Storage/KETJU/initialise/merger/"
+
+#file1 /scratch/pjohanss/arawling/collisionless_merger/stability-test/NGCa0524b/output/NGCa0524b_009.hdf5                #file to generate IC from
+#file2 /scratch/pjohanss/arawling/collisionless_merger/stability-test/NGCa3348/output/NGCa3348_009.hdf5                #file to generate IC from
+#saveLocation /scratch/pjohanss/arawling/collisionless_merger/orbit-test/        #file will be saved to saveLocation/galaxyName_1-galaxyName_2
+
+
+#---- orbital properties
+initialSeparation = "virial2"       #'touch', 'overlapXX', 'virialX' or number
+pericentreDistance = "virial1e-3"   #distance at first pericentre
+
+#----------------------------------returned values
+e = 9.99985e-01
+r0 = 1.20901e+03
+rperi = 6.04506e-01
+#----------------------
+virial_radius = 6.04506e+02
+#----------------------

### 2022-03-29 paper/sections/numerical.tex
Begin writing the IC section

@@
 
-\subsection{Initial Conditions and Simulations}
+\subsection{Initial Conditions}
+
+We create four\footnote{A fifth galaxy, model B, was created with much higher mass than the four presented here.
+As the merger system did not result in a bound SMBH binary within some $10\,\text{Gyr}$, we exclude discussion of this galaxy.} unique progenitor galaxies, motivated by analogues observed in the real Universe.
+Each galaxy is modelled as a multi-component system: a stellar component embedded within a DM halo, and a single SMBH at rest at the centre of the galaxy. 
+Two galaxies, A and D, have a stellar density profile given by the `cuspy' \citet{dehnen93} profile:
+\begin{equation}\label{eq:dehnen}
+    \rho_\star(r) = \frac{M_\star}{2\pi} \frac{a}{r(r+a)^{(4-\gamma)}}
+\end{equation}
+and two galaxies, C and E, by the `cored' \citet{terzic05} profile:
+\begin{align}\label{eq:terzic}
+    \rho_\star(r) &= \rho' \left[ 1+\left(\frac{r_b}{r} \right)^\alpha \right]^{\gamma/\alpha} \\
+    &\times \left\{ \left[ \frac{r^\alpha+r_b^\alpha}{R_\text{e}^\alpha} \right]^{-p/\alpha} \text{e}^{-b \left[ \left(r^\alpha+r_b^\alpha\right)/R_\text{e}^\alpha \right]^{1/n\alpha}} \right\}, \nonumber
+\end{align}
+where
+\begin{equation}
+    \rho' = \rho_b 2^{(p-\gamma)/\alpha} \left(\frac{r_b}{r} \right)^p \text{e}^{b\left(2^{1/\alpha r_b/R_\text{e}}\right)^{1/n}}.
+\end{equation}
+
+The stellar component, whilst having a spherically-symmetric density distribution, employs the anisotropic Osipkov-Merritt distribution function (DF), from which the phase space is Monte-Carlo sampled using a minimum of $\sim10^6$ particles with a mass of $10^5\,\text{M}_\odot$.
+The Osipkov-Merritt DF has a free parameter $r_a$, termed the anisotropy radius, which determines the degree of radial anisotropy. 
+The DF is near-ergodic in the inner regions $r<r_a$, and becomes increasingly radially-biased at larger $r$.
+It has the form:
+\begin{equation}\label{eq:om}
+    f(Q) = \frac{1}{\sqrt{8}\pi^2} \left[ \int_0^Q \dv{\Psi}{\sqrt{Q-\Psi}} \dv{^2\nu_q}{\Psi^2} + \frac{1}{\sqrt{Q}} \left(\dv{\nu_Q}{\Psi}\right)_{\Psi=0}\right],
+\end{equation}
+where
+\begin{equation*}
+    \nu_Q(r) \equiv \left(1+\frac{r^2}{r_a^2}\right) \nu(r)
+\end{equation*}
+and
+\begin{equation*}
+    Q \equiv \mathcal{E} - \frac{L^2}{2r_a^2}.
+\end{equation*}
+The parameter $r_a$ is chosen so as to reproduce observed kinematic relations, most notably the $M_\bullet\text{--}\sigma_\star$ relation (see \drafting{figure}), whilst maintaining a positive DF.
+
+Conversely, the DM halo of all galaxies is modelled as an ergodic Navarro-Frenk-White (NFW) profile \citep{navarro97}, with expected virial mass taken from the \citet{behroozi19} peak halo mass -- stellar mass relation.
+\begin{equation}\label{eq:nfw}
+    \rho_\text{DM}(r) = \frac{\rho_0}{(r/a)(1+r/a)^2}
+\end{equation}
+We do not incorporate a redshift-dependent halo mass, instead opting to model our haloes as if they were taken at $z=0$. 
+The NFW concentration parameter $c\equiv r_\text{vir}/a$, where $a$ is the scale radius, is taken from the \citet{dutton14} relation.
+To prevent the radially-divergent mass distribution associated with the NFW profile, we multiply the density profile \eqref{eq:nfw} with a sigmoid function, and demand that the ratio of the actual DM density to the theoretical NFW density satisfies
+\begin{equation*}
+    \frac{\rho_\text{true}(r)}{\rho_\text{NFW}(r)} < 10^{-5}
+\end{equation*}
+by $7\,R_\text{vir}$, with significant departure from the NFW profile already by $3\,R_\text{vir}$. 
+With this modified NFW profile, the phase space is Monte Carlo sampled using the approach of \eqref{eq:om} and a minimum of $\sim10^6$ particles with a mass of $3.0\times10^7\,\text{M}_\odot$, assuming $L=0$ and $r_a\rightarrow\infty$.
+The virial radius of the system is then re-determined numerically, to ensure it is in close agreement with the expected value from \citet{behroozi19}.
+
+The final component of the model galaxies, the SMBH, is represented as a point mass at rest at the origin. 
+The SMBH mass is taken from the $M_\text{bulge}\text{--}M_\bullet$ relation of \citet{sahu19}, and by definition lies on this relation. 
+The SMBH is also given an initial spin $\chi$ from the \citet{lousto10} distribution, so that we may also investigate GW-induced recoil kicks in our merger remnants.
+
+After creating the multicomponent ICs, the galaxies are evolved in isolation using \gadget for at least \drafting{$8\,\text{Gyr}$}, to ensure the systems are fully-relaxed and any potential effects arising from radial instabilities are mitigated.
+We then extract the evolved systems, and use these as the progenitors for all subsequent merger configurations.
+
+\begin{deluxetable*}{cccccccccc}
+    \tablecaption{
+        Properties of Progenitor Galaxies
+    }
+    \tablehead{
+        \colhead{Name} &
+        \colhead{Stellar Profile} &
+        \colhead{$r_a/\text{kpc}$} &
+        \colhead{$M_\star/10^{11}\,\text{M}_\odot$} &
+        \colhead{$N_\star$} &
+        \colhead{$M_\text{DM}/10^{13}\,\text{M}_\odot$} &
+        \colhead{$c_\text{DM}$} &
+        \colhead{$N_\text{DM}$} &
+        \colhead{$M_\bullet/10^{9}\,\text{M}_\odot$} &
+        \colhead{$R_\text{vir}/\text{kpc}$}
+    }
+    \startdata
+        A & Dehnen & 0.4 & 2.51 & $2.5\times10^6$ & 9.65 & 7.55 & $3.22\times10^6$ & 3.04 & 554
+        \\
+        C & Terzi\'c & 1.5 & 2.57 & $2.6\times10^6$ & 5.13 & 7.51 & $3.38\times10^6$ & 3.14 & 562
+        \\
+        D & Dehnen & 0.3 & 1.58 & $1.58\times10^6$ & 2.00 & 8.23 & $1.29\times10^6$ & 1.64 & 412
+        \\
+        E & Terzi\'c & 0.5 & 1.46 & $1.46\times10^6$ & 3.29 & 8.36 & $1.10\times10^6$ & 1.46 & 392
+    \enddata
+    \end{deluxetable*}
+
+\subsection{Simulations}
+
+We simulate all 6 different progenitor combinations, namely AC, AD, AE, CD, CE, DE. 
+This allows us to sample a vast range of progenitor configurations: each of equal and unequal mass, for each combination of cuspy and cored profiles.
+
+In keeping with our theme of realistic initial conditions, we set all galaxy merger orbits to an initial separation of $r_0=3\,R_\text{vir,1}$, where $R_\text{vir,1}$ is the virial radius of the larger of the two progenitors.
+The progenitor galaxies are then set on a quasi-Keplerian orbit, assuming each galaxy is a point-mass containing half its total mass. 
+The eccentricity of the initial orbit is chosen as a function of the initial pericentre distance $r_\text{peri}$, motivated from cosmological simulations, using the data presented in \citet{khochfar06}.
+By fitting a functional form to the median eccentricity value in logarithmically-spaced bins, we arrive at the relation
+\begin{equation}\label{eq:rperi-e}
+    e(r_\text{peri}/R_\text{vir,1}) = \left[ 1 + \left( \frac{r_\text{peri}/R_\text{vir,1}}{0.320} \right)^{1.629} \right]^{-0.176}
+\end{equation}
+from which we set the initial eccentricity of all our merger orbits.
+For each progenitor combination, we sample six different pericentre distances, and thus six different eccentricities, in the range $10^{-3} \leq r_\text{peri}/R_\text{vir,1} \leq 1.0$.
+
+The integration of the system is conducted in two phases: a \textit{parent phase} and a \textit{child phase}. 
+The former is done to capture the large-scale dynamics when the SMBHs are at large separation, and the latter when the SMBH interactions become of primary interest.
+
+The progenitors are evolved through the \textit{parent phase} from $t(r_0)=0$ with \gadget until the SMBHs form a bound binary at $t_\text{stop}$, with a stellar softening of $\epsilon_\star=10\,\text{pc}$ and integration tolerances \drafting{check}.
+We then initialise the \textit{child phase} by restarting from a snapshot sometime before $t_\text{stop}$.
+We apply a small perturbation to each SMBH position and velocity vector to mimic the effect of Brownian motion due to discrete sampling of the DF, thus negating the need to run many simulations from $r_0=3\,R_\text{vir,1}$.
+The perturbations are drawn from a Gaussian distribution modelling the typical offset of the SMBH phase space coordinates from the phase space coordinates of the stellar centre of mass, determined using the shrinking sphere method.
+The Gaussian distributions are characterised by $\mathcal{N}_\text{pos}\sim(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc})$, and $\mathcal{N}_\text{vel}\sim(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km\,s}^{-1})$ for each progenitor \drafting{should we show a figure of this?}.
+From each \parent, we create ten uniquely-perturbed \child runs, allowing us to explore the distribution of parameters describing both the orbital and remnant properties of each \parent merger. 
+Additionally, we lower the stellar softening to $\epsilon_\star=3.5\,\text{pc}$ and tighten the integration tolerances to \drafting{check} to optimise the run for \ketju.
+We check for energy conservation between the \parent and \child runs, and confirm that the total energy is conserved at the sub-percent level. 
+In total, we perform 360 ultra-high resolution unique \child simulations using \ketju, and use this sample for our analysis.
+
+Within a relatively short time frame from beginning the \child run, the SMBH binary hardens predominantly through three-body stellar interactions. 
+We evolve the binary to the time $t_{a_h}=t(a_h)$, and continue to when $t_\text{safe}=t(a=15\,\text{pc})$. 
+We then analytically predict the orbital evolution of the binary by determining the hardening constant \drafting{assume this was explained in the intro} $H$ and eccentricity constant $K$ by fitting the derivative of $\text{d}(1/a)/\text{d}t$ with a linear function between $t_{a_h}$ and $t_\text{safe}$.
+Using $a_0=a_h$ and three different values of $e_0$, taken to be the 5\%, 50\%, and 95\% quantiles of the eccentricity distribution with the same time interval 
+$t_{a_h} \text{--} t_\text{safe}$, we predict the expected merger timescale of the system. 
+We continue to run the \child runs until one of three conditions are met:
+\begin{enumerate}
+    \item The SMBHs merge, or
+    \item The Hubble time, $13.8\,\text{Gyr}$, is exceeded, or
+    \item The eccentricity stabilises to below the analytically-determined eccentricity estimated from the 5\%-quantile eccentricity value, and said analytical-estimate indicates that the merger will occur significantly after the Hubble time.
+\end{enumerate}
+If it is not clear if the third condition is satisfied, we continue to run the merger until the second condition is met.
+
+\begin{deluxetable*}{ccc}
+\tablecaption{
+    Properties of Merger Orbits
+}
+\tablehead{
+    \colhead{Name} &
+    \colhead{$r_\text{peri}/R_\text{vir,1}$} &
+    \colhead{$e_0$} 
+}
+\startdata
+    XY-030-0001 & $1.0\times10^{-3}$ & 0.999
+    \\
+    XY-030-0005 & $5.0\times10^{-3}$ & 0.998
+    \\
+    XY-030-0010 & $1.0\times10^{-2}$ & 0.994
+    \\
+    XY-030-0050 & $5.0\times10^{-2}$ & 0.992
+    \\
+    XY-030-0100 & $1.0\times10^{-1}$ & 0.976
+    \\
+    XY-030-1000 & $1.0\times10^{-0}$ & 0.703
+\enddata
+\end{deluxetable*}
+
+\begin{figure*}
+    \centering
+    \includegraphics{compic.png}
+    \caption{\drafting{some key IC plots}}
+    \label{fig:ics}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.5\textwidth]{initial_orbits.jpeg}
+    \caption{\drafting{Plots of the initial orbit?}}
+    \label{fig:icorbit}
+\end{figure*}

### 2022-07-26 paper/sections/numerical.tex
Small updates to paper draft

@@
 After creating the multicomponent ICs, the galaxies are evolved in isolation using \gadget for at least \drafting{$8\,\text{Gyr}$}, to ensure the systems are fully-relaxed and any potential effects arising from radial instabilities are mitigated.
-We then extract the evolved systems, and use these as the progenitors for all subsequent merger configurations.
+We then extract the evolved systems, and align each system with its reduced inertia tensor.
+We use these as the progenitor galaxies for all subsequent merger configurations.
 
@@
 
-The progenitors are evolved through the \textit{parent phase} from $t(r_0)=0$ with \gadget until the SMBHs form a bound binary at $t_\text{stop}$, with a stellar softening of $\epsilon_\star=10\,\text{pc}$ and integration tolerances \drafting{check}.
+The progenitors are evolved through the \textit{parent phase} from $t(r_0)=0$ with \gadget until the SMBHs form a bound binary at $t_\text{stop}$, with a stellar softening of $\epsilon_\star=10\,\text{pc}$ and \gadget integration tolerances \texttt{ErrTolIntAccuracy}=$0.02$ and \texttt{ErrTolForceAcc}=$0.005$.
 We then initialise the \textit{child phase} by restarting from a snapshot sometime before $t_\text{stop}$.
@@
 The perturbations are drawn from a Gaussian distribution modelling the typical offset of the SMBH phase space coordinates from the phase space coordinates of the stellar centre of mass, determined using the shrinking sphere method.
-The Gaussian distributions are characterised by $\mathcal{N}_\text{pos}\sim(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc})$, and $\mathcal{N}_\text{vel}\sim(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km\,s}^{-1})$ for each progenitor \drafting{should we show a figure of this?}.
+The Gaussian distributions are characterised by $\mathcal{N}_\text{pos}\sim(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc}^2)$, and $\mathcal{N}_\text{vel}\sim(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km}^2\text{\,s}^{-2})$ for each progenitor \drafting{should we show a figure of this?}.
 From each \parent, we create ten uniquely-perturbed \child runs, allowing us to explore the distribution of parameters describing both the orbital and remnant properties of each \parent merger. 
-Additionally, we lower the stellar softening to $\epsilon_\star=3.5\,\text{pc}$ and tighten the integration tolerances to \drafting{check} to optimise the run for \ketju.
+Additionally, we lower the stellar softening to $\epsilon_\star=3.5\,\text{pc}$ and tighten the integration tolerance \texttt{ErrTolIntAccuracy} to $0.002$ to optimise the run for \ketju.
 We check for energy conservation between the \parent and \child runs, and confirm that the total energy is conserved at the sub-percent level. 
@@
 
+\subsection{Bayesian Hierarchical Modelling}
+\drafting{\begin{enumerate}
+    \item HMC algorithm?
+    \item Description of hierarchical model
+    \item Prior distributions
+    \item forward modelling
+\end{enumerate}
+}
+
 \begin{figure*}

### 2022-07-26 paper/sections/results.tex
Small updates to paper draft

@@
 \begin{itemize}
-\item Good ones
+\item Example of forward-modelled semimajor axis and eccentricity
+\item Example of forward-modelled density profile
+\item Example of forward-modelled velocity dispersion profile
 \end{itemize}
@@
 and maybe another one of BH merger remnant properties.
+}
+
+\drafting{
+    WHAT ARE THE MAIN OBSERVATIONS WE SEE and WHAT CONCLUSIONS CAN WE DRAW?
+    \begin{enumerate}
+        \item stability of hardening rate -> semimajor axis can be robustly predicted
+        \item wide variance of eccentricity -> N-body models offer no quantitative prediction on likelihood of merger, can only comment on what merger looks like \textit{IF} it happens
+        \item variance in observables: trends between observable properties and what observers can actually make use of (don't have this yet), e.g. signatures of SMBH scouring in velocity profile?
+    \end{enumerate}
 }

### 2022-08-26 paper/sections/results.tex
Add some preliminary figures

@@
 
+\begin{figure*}
+    \includegraphics[width=0.7\textwidth]{graham_density_prior_pred_log10_Sigma_mean.png}
+    \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit.
+Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
+The observed data is overlaid.
+From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.
+TODO??
+This isn't a true prior predictive test, as the independent variable $R$ is not the $R$ of the data but some arbitrary range.
+Change it?}}
+    \label{fig:priorpred}
+\end{figure*}
+\begin{figure*}
+    \includegraphics[width=0.7\textwidth]{graham_density_pair_0.png}
+    \caption{\drafting{Corner plot of hierarchical model hyperparameters.
+Each of the hyperdistributions are modelled as Gamma distributions with shape parameters $\alpha$ and $\beta$.
+For example, the hyperdistribution on the core radius $r_\mathrm{b}$ is modelled as $r_\mathrm{b} \sim \text{Gamma}(r_{\mathrm{b},\alpha}, r_{\mathrm{b},\beta})$.
+Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well.
+}}
+    \label{fig:corner_params_1}
+\end{figure*}
+\begin{figure*}
+    \includegraphics[width=0.7\textwidth]{graham_density_pair_1.png}
+    \caption{\drafting{Same as \ref{fig:corner_params_1}}}
+    \label{fig:corner_params_2}
+\end{figure*}
+\begin{figure*}
+    \includegraphics[width=0.7\textwidth]{graham_density_posterior_pred_log10_Sigma_mean.png}
+    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit.
+Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
+The observed data is overlaid, and seen to lie well within the 50\% BCI.
+}}
+    \label{fig:postpred}
+\end{figure*}
+
 %\begin{figure*}

### 2022-08-26 paper/sections/results.tex
Begin writing a bit on HMC
- update some figures using the new analysis methods

@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{graham_density_prior_pred_log10_Sigma_mean.png}
+    \includegraphics[width=\textwidth]{{graham_density-A-C-3.0-0.05_prior_pred_log10_proj_density_mean}.png}
     \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit.
@@
 The observed data is overlaid.
-From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.
-TODO??
-This isn't a true prior predictive test, as the independent variable $R$ is not the $R$ of the data but some arbitrary range.
-Change it?}}
+From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.}}
     \label{fig:priorpred}
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{graham_density_pair_0.png}
+    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_0}.png}
     \caption{\drafting{Corner plot of hierarchical model hyperparameters.
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{graham_density_pair_1.png}
+    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
     \caption{\drafting{Same as \ref{fig:corner_params_1}}}
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{graham_density_posterior_pred_log10_Sigma_mean.png}
+    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_posterior_pred_log10_proj_density_mean}.png}
     \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit.

### 2022-08-26 paper/sections/stats.tex
Begin writing a bit on HMC
- update some figures using the new analysis methods

We perform full Bayesian Hierarchical Modelling for the projected mass density profiles of our simulated galaxies. 
The modelling is performed using the statistical software \textsc{Stan}\footnote{https://mc-stan.org}, which itself implements Hamiltonian Monte Carlo (HMC).

\subsection{Hamiltonian Monte Carlo}
An excellent overview of HMC may be found in \citet{betancourt17}, with a brief description provided here.

In numerically sampling an arbitrary target distribution, we are required to sample from the \textit{typical set}, or those regions of the target distribution which provide the dominant contribution to an expectation value. 
Simplistically, this may be achieved using any Markov generating process, such as the widely-used Metropolis-Hastings (MH) algorithm, albeit at the expense of inefficient and biased sampling. 
The MH algorithm generates from a starting position $q$ a proposal for the next position $q'$, with proposals accepted with some probability $Q(q'|q)$, where $Q$ is often chosen to be the normal distribution. 
Sampling with increased efficiency can be achieved by using information about the geometry of the typical set of the target distribution through a vector field aligned with the typical set, which may be estimated using the gradient of the target distribution. 
To align the gradient vector field with the typical set vector field, we can introduce auxillary momenta variables $p$, promoting our $D$-dimensional target distribution to a projection of a $2D$-dimensional canonical distribution in some phase space:
\begin{equation*}
    P(q,p) = P(p|q)P(q) = e^{-H(q,p)}
\end{equation*}
where $H$ is an invariant Hamiltonian function. 
The desired vector field may then be generated from the Hamiltonian $H$ using Hamilton's equations. 
This formulation also lends itself to improved sampling efficiency by decomposing the Hamiltonian into a \textit{kinetic energy} term and a \textit{potential energy} term, with the latter equivalent to the logarithm of the target distribution. 
As the Hamiltonian is phase space conserving, they may be accurately integrated using a volume-preserving integration scheme, such as symplectic integrators. 
Of these, the leapfrog scheme offers both a simple implementation and high accuracy over long integration times.

\subsection{Hierarchical Models}
\begin{itemize}
    \item Data doesn't have error, model has uncertainty
    \item Require multiple observations from a population --> samples
    \item Assume population described by a series of hyperparameters
    \item Prior distributions on those hyperparameters
    \item Forward fold through model, combine with likelihood function from data
\end{itemize}

\drafting{\begin{enumerate}
    \item HMC algorithm?
    \item Description of hierarchical model
    \item Prior distributions
    \item forward modelling
\end{enumerate}
}

### 2022-09-06 paper/sections/numerical.tex
Fix typo in Dehnen formula
Begin writing a bit on HMC
- update some figures using the new analysis methods
Switch to MNRAS template
- Discuss more about statistical implementation

- Change project name from Himmel

@@
 
+\begin{table*}
+    \centering
+    \caption{Properties of progenitor galaxies}
+    \label{tab:ics}
+    \begin{tabular}{ccccccccccc}
+        \hline
+        Name & Stellar Profile & $r_a/\text{kpc}$ & $M_\star/(10^{11}\,\Msun)$ & $N_\star$ & $M_\text{DM}/(10^{13}\,\Msun)$ & $c_\text{DM}$ & $N_\text{DM}$ & $M_\bullet/(10^{9}\,\Msun)$ & $R_\text{vir}/\mathrm{kpc}$ & $R_\mathrm{e}/\mathrm{kpc}$ \\
+        \hline
+        A & Dehnen & 0.4 & 2.51 & $2.5\times10^6$ & 9.65 & 7.55 & $3.22\times10^6$ & 3.04 & 554 & 4.40 \\
+        C & Terzi\'c & 1.5 & 2.57 & $2.6\times10^6$ & 5.13 & 7.51 & $3.38\times10^6$ & 3.14 & 562 & 4.05 \\
+        D & Dehnen & 0.3 & 1.58 & $1.58\times10^6$ & 2.00 & 8.23 & $1.29\times10^6$ & 1.64 & 412 & 3.16 \\
+        E & Terzi\'c & 0.5 & 1.46 & $1.46\times10^6$ & 3.29 & 8.36 & $1.10\times10^6$ & 1.46 & 392 & 2.79 \\
+        \hline
+    \end{tabular}
+\end{table*}
+
 \subsection{Initial Conditions}
+
+\begin{figure*}
+    \centering
+    \includegraphics{compic.pdf}
+    \caption{\drafting{Comparison of progenitor galaxies to key observables.
+\textit{Left:} The simulated galaxies have a SMBH mass determined from the \citet{sahu19} relation.
+Cored galaxies are shown as circles, and cuspy galaxies as squares.
+Note the progenitor galaxies A and C overlap.
+\textit{Centre}: The stellar velocity dispersion within $1\,R_\mathrm{e}$ is computed for the progenitor galaxies, and compared to observational data collated in \citet{bosch16}.
+\textit{Right:} The dark matter fraction within $1\,R_\mathrm{e}$ is computed for the progenitor galaxies, and compared to Schwarzschild-modelling results from \citet{jin20}.}}
+    \label{fig:ics}
+\end{figure*}
 
@@
 \begin{equation}\label{eq:dehnen}
-    \rho_\star(r) = \frac{M_\star}{2\pi} \frac{a}{r(r+a)^{(4-\gamma)}}
+    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}
 \end{equation}
@@
 The NFW concentration parameter $c\equiv r_\text{vir}/a$, where $a$ is the scale radius, is taken from the \citet{dutton14} relation.
-To prevent the radially-divergent mass distribution associated with the NFW profile, we multiply the density profile \eqref{eq:nfw} with a sigmoid function, and demand that the ratio of the actual DM density to the theoretical NFW density satisfies
+To prevent the radially-divergent mass distribution associated with the NFW profile, we multiply the density profile \autoref{eq:nfw} with a sigmoid function, and demand that the ratio of the actual DM density to the theoretical NFW density satisfies
 \begin{equation*}
@@
 by $7\,R_\text{vir}$, with significant departure from the NFW profile already by $3\,R_\text{vir}$. 
-With this modified NFW profile, the phase space is Monte Carlo sampled using the approach of \eqref{eq:om} and a minimum of $\sim10^6$ particles with a mass of $3.0\times10^7\,\text{M}_\odot$, assuming $L=0$ and $r_a\rightarrow\infty$.
+With this modified NFW profile, the phase space is Monte Carlo sampled using the approach of \autoref{eq:om} and a minimum of $\sim10^6$ particles with a mass of $3.0\times10^7\,\text{M}_\odot$, assuming $L=0$ and $r_a\rightarrow\infty$.
 The virial radius of the system is then re-determined numerically, to ensure it is in close agreement with the expected value from \citet{behroozi19}.
@@
 The final component of the model galaxies, the SMBH, is represented as a point mass at rest at the origin. 
-The SMBH mass is taken from the $M_\text{bulge}\text{--}M_\bullet$ relation of \citet{sahu19}, and by definition lies on this relation. 
+The SMBH mass is taken from the $M_\text{bulge}\text{--}M_\bullet$ relation of \citet{sahu19}, and by definition lies on this relation, shown in the left panel of \autoref{fig:ics}. 
 The SMBH is also given an initial spin $\chi$ from the \citet{lousto10} distribution, so that we may also investigate GW-induced recoil kicks in our merger remnants.
 
-After creating the multicomponent ICs, the galaxies are evolved in isolation using \gadget for at least \drafting{$8\,\text{Gyr}$}, to ensure the systems are fully-relaxed and any potential effects arising from radial instabilities are mitigated.
+We compare the progenitor systems to key constraining observables in \autoref{fig:ics}.
+In the centre panel, we overlay our progenitor galaxies on data used to fit the black hole mass - central stellar velocity dispersion relation of \citet{bosch16}. 
+The progenitor galaxies lie within the spread of the data, and have velocity dispersions indistinguishable from observational data. 
+Similarly, we find that the fraction of dark matter within $1\,R_\mathrm{e}$ is comparable to the dark matter fraction determined with Schwarzschild modelling of central and satellite galaxies in the MaNGA survey \citep{jin20}, shown in the right panel of \autoref{fig:ics}.
+Our progenitor galaxies are consistently slightly below the median relation (the median $f_\mathrm{DM}$ value is shown as crosses in stellar mass bins), but still well within the scatter of \citet{jin20}.
+We conclude that our progenitor galaxy models, whilst neglecting the inclusion of gas, plausibly represent observable massive galaxies, particularly at the mass range of $M_\star/\Msun > 10^{11}$ that we are interested in.
+
+After creating and assessing the multicomponent ICs, the galaxies are evolved in isolation using \gadget for at least \drafting{$8\,\text{Gyr}$}, to ensure the systems are fully-relaxed and any potential effects arising from radial instabilities are mitigated.
 We then extract the evolved systems, and align each system with its reduced inertia tensor.
@@
 
-\begin{deluxetable*}{cccccccccc}
-    \tablecaption{
-        Properties of Progenitor Galaxies
-    }
-    \tablehead{
-        \colhead{Name} &
-        \colhead{Stellar Profile} &
-        \colhead{$r_a/\text{kpc}$} &
-        \colhead{$M_\star/10^{11}\,\text{M}_\odot$} &
-        \colhead{$N_\star$} &
-        \colhead{$M_\text{DM}/10^{13}\,\text{M}_\odot$} &
-        \colhead{$c_\text{DM}$} &
-        \colhead{$N_\text{DM}$} &
-        \colhead{$M_\bullet/10^{9}\,\text{M}_\odot$} &
-        \colhead{$R_\text{vir}/\text{kpc}$}
-    }
-    \startdata
-        A & Dehnen & 0.4 & 2.51 & $2.5\times10^6$ & 9.65 & 7.55 & $3.22\times10^6$ & 3.04 & 554
-        \\
-        C & Terzi\'c & 1.5 & 2.57 & $2.6\times10^6$ & 5.13 & 7.51 & $3.38\times10^6$ & 3.14 & 562
-        \\
-        D & Dehnen & 0.3 & 1.58 & $1.58\times10^6$ & 2.00 & 8.23 & $1.29\times10^6$ & 1.64 & 412
-        \\
-        E & Terzi\'c & 0.5 & 1.46 & $1.46\times10^6$ & 3.29 & 8.36 & $1.10\times10^6$ & 1.46 & 392
-    \enddata
-    \end{deluxetable*}
-
 \subsection{Simulations}
 
-We simulate all 6 different progenitor combinations, namely AC, AD, AE, CD, CE, DE. 
+We simulate all six different progenitor combinations, namely AC, AD, AE, CD, CE, DE. 
 This allows us to sample a vast range of progenitor configurations: each of equal and unequal mass, for each combination of cuspy and cored profiles. 
+Herein, we refer to a particular progenitor combination as a \textit{progenitor class}.
 
@@
 from which we set the initial eccentricity of all our merger orbits.
-For each progenitor combination, we sample six different pericentre distances, and thus six different eccentricities, in the range $10^{-3} \leq r_\text{peri}/R_\text{vir,1} \leq 1.0$.
+For each progenitor class, we sample six different pericentre distances, and thus six different eccentricities, in the range $10^{-3} \leq r_\text{peri}/R_\text{vir,1} \leq 1.0$. 
+Consequently, we investigate 36 `families' of merger configurations.
 
@@
 The perturbations are drawn from a Gaussian distribution modelling the typical offset of the SMBH phase space coordinates from the phase space coordinates of the stellar centre of mass, determined using the shrinking sphere method.
-The Gaussian distributions are characterised by $\mathcal{N}_\text{pos}\sim(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc}^2)$, and $\mathcal{N}_\text{vel}\sim(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km}^2\text{\,s}^{-2})$ for each progenitor \drafting{should we show a figure of this?}.
+The Gaussian distributions are characterised by $\mathcal{N}_\text{pos}(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc}^2)$, and $\mathcal{N}_\text{vel}(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km}^2\text{\,s}^{-2})$ for each progenitor \drafting{should we show a figure of this?}.
 From each \parent, we create ten uniquely-perturbed \child runs, allowing us to explore the distribution of parameters describing both the orbital and remnant properties of each \parent merger. 
@@
 
-\begin{deluxetable*}{ccc}
-\tablecaption{
-    Properties of Merger Orbits
-}
-\tablehead{
-    \colhead{Name} &
-    \colhead{$r_\text{peri}/R_\text{vir,1}$} &
-    \colhead{$e_0$} 
-}
-\startdata
-    XY-030-0001 & $1.0\times10^{-3}$ & 0.999
-    \\
-    XY-030-0005 & $5.0\times10^{-3}$ & 0.998
-    \\
-    XY-030-0010 & $1.0\times10^{-2}$ & 0.994
-    \\
-    XY-030-0050 & $5.0\times10^{-2}$ & 0.992
-    \\
-    XY-030-0100 & $1.0\times10^{-1}$ & 0.976
-    \\
-    XY-030-1000 & $1.0\times10^{-0}$ & 0.703
-\enddata
-\end{deluxetable*}
+\begin{table}[t]
+    \centering
+    \caption{Properties of merger orbits}
+    \label{tab:mergers}
+    \begin{tabular}{ccc}
+        \hline
+        Name & $r_\text{peri}/R_\text{vir,1}$ & $e_0$ \\
+        \hline
+        XY-030-0001 & $1.0\times10^{-3}$ & 0.999 \\
+        XY-030-0005 & $5.0\times10^{-3}$ & 0.998 \\
+        XY-030-0010 & $1.0\times10^{-2}$ & 0.994 \\
+        XY-030-0050 & $5.0\times10^{-2}$ & 0.992 \\
+        XY-030-0100 & $1.0\times10^{-1}$ & 0.976 \\
+        XY-030-1000 & $1.0\times10^{0}$ & 0.703 \\
+        \hline
+    \end{tabular}
+\end{table}
 
-\subsection{Bayesian Hierarchical Modelling}
-\drafting{\begin{enumerate}
-    \item HMC algorithm?
-    \item Description of hierarchical model
-    \item Prior distributions
-    \item forward modelling
-\end{enumerate}
-}
-
-\begin{figure*}
-    \centering
-    \includegraphics{compic.png}
-    \caption{\drafting{some key IC plots}}
-    \label{fig:ics}
-\end{figure*}
-
-\begin{figure*}
+\begin{figure}
     \centering
     \includegraphics[width=0.5\textwidth]{initial_orbits.jpeg}
-    \caption{\drafting{Plots of the initial orbit?}}
+    \caption{\drafting{Plots of the initial orbit? Update this figure if so}}
     \label{fig:icorbit}
-\end{figure*}
+\end{figure}

### 2022-09-06 paper/sections/results.tex
Switch to MNRAS template
- Discuss more about statistical implementation

- Change project name from Himmel

@@
 \end{itemize}
-Table of galaxy properties,
 and maybe another one of BH merger remnant properties.
@@
 
+\drafting{ENSURE CONSISTENCY WITH SIGMA AND I. AS WE ARE USING MASS DENSITY, SHOULD BE SIGMA, NOT I, OR ASSUME A M/L RATIO OF 1. NEED TO GO THROUGH AND ENSURE THIS}
+
+\subsection{Projected Mass Density Profiles}
+We use hierarchical modelling to recover the population parameter distributions of the core-S\'ersic model \autoref{eq:coresersic} for each simulation family. 
+We enforce a sharp transition from the core to outer galaxy regions by setting $\alpha$ in \autoref{eq:coresersic} to 10.0. 
+A probabilistic graphical model detailing the connection between hyperparameters, latent parameters, and observables is given in \autoref{fig:graphicalmodel}.
+
+\begin{table}
+    \centering
+    \caption{The parameters for each hyperparameter distribution for the core-S\'ersic model, where each distribution is a gamma distribution.}
+    \label{tab:cshypers}
+    \begin{tabular}{ccc}
+        \hline
+        Hyperparameter & $\alpha$ & $\beta$ \\
+        \hline
+        $r_{\mathrm{b}, \alpha}$ & 3.0 & 2.0 \\
+        $r_{\mathrm{b}, \beta}$ & 10.0 & 8.0 \\
+        $R_{\mathrm{e}, \alpha}$ & 30.0 & 2.0 \\
+        $R_{\mathrm{e}, \beta}$ & 40.0 & 2.0 \\
+        $\tilde{\Sigma}_{\mathrm{b}, \alpha}$ & 20.0 & 4.0 \\
+        $\tilde{\Sigma}_{\mathrm{b}, \beta}$ & 20.0 & 2.0 \\
+        $\gamma_\alpha$ & 1.0 & 2.0 \\
+        $\gamma_\beta$ & 4.0 & 2.0 \\
+        $n_\alpha$ & 16.0 & 2.0 \\
+        $n_\beta$ & 4.0 & 2.0 \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}
+    \includegraphics{hm_coresersic.pdf}
+    \caption{Probabilistic graphical model for the likelihood of the core-S\'ersic model.
+Probabilistic parameters and observable quantities (hatted variables) are shown as circles, whereas deterministic quantities are represented by diamonds.
+Hyperparameters are represented by purple circles, latent parameters in blue, and observable quantities as red circles.
+The nodes indicate the distribution used in constructing the prior distributions (for latent parameters) and likelihood function (for $\hat{\Sigma}$).
+The radial box represents observed and latent variables for each radial bin for a \child, and the child box represents latent parameters for each \child simulation in a family.
+}
+    \label{fig:graphicalmodel}
+\end{figure}
+
+\subsubsection{Observable Quantities}
+The hierarchical model takes three observable quantities, denoted by hat-variables in red circles in \autoref{fig:graphicalmodel}. 
+These variables are the radius $\hat{R}$, and the mean projected mass density $\hat{\Sigma}$, and the standard deviation of the projected mass density $\hat{\sigma}_\Sigma$ measured in the radial bins centred on $\hat{R}$.
+For each \child simulation, we choose to `observe' the first snapshot where the binary semimajor axis is \textit{at most} $10\,\mathrm{pc}$. 
+In the event that there is no snapshot satisfying the above requirement (due to, for example, rapid merging of the SMBH binary in a time period shorter than the snapshot output frequency), the \child simulation is not included in the analysis.
+We enforce a minimum of six \child simulations per family for that family to be analysed with our hierarchical modelling technique. 
+For those \child simulations which pass our selection criteria, we take 30 random projections of the merger remnant, and determine the projected mass density in 50 radial bins evenly-spaced in logarithm space, with a minimum value of $0.2\,\mathrm{kpc}$ and a maximum value of $20.0\,\mathrm{kpc}$.
+This allows us to build up a distribution of projected mass densities per radial bin, from which $\hat{\Sigma}$ and $\hat{\sigma}_\Sigma$ are determined.
+
+\subsubsection{Latent Parameters for Observables}
+With the omission of the parameter $\alpha$ in the modelling process, there are five latent parameters ($r_\mathrm{b}$, $R_\mathrm{e}$, $\tilde{\Sigma}_\mathrm{b}$, $\gamma$, and $n$, the set of which we denote as $\kappa$) that describe each \child simulation in a family.
+Here, we define $\tilde{\Sigma}_\mathrm{b} \equiv \Sigma_\mathrm{b}/10^{10}$, where $\Sigma_\mathrm{b}$ is defined in \autoref{eq:coresersic2}.
+These latent parameters, depicted as non-hatted variables in blue circles in \autoref{fig:graphicalmodel}, are sampled to obtain a latent estimate on the projected mass density, $\Sigma$.
+As each latent parameter in the set $\kappa$ is strictly non-negative, we use as prior distributions for each latent parameter a gamma distribution with hyperparameters $(\kappa_\alpha, \kappa_\beta)$, which are themselves drawn from a gamma distribution.
+To ensure our hyperparameter prior distributions are weakly-informative, the hyperparameters are drawn from the distributions in \autoref{tab:cshypers} for all simulation families. 
+The sampled hyperparameter distributions are shown as corner plots in Figures \ref{fig:corner_params_1}-\ref{fig:corner_params_3}.
+
+We ensure our hyperparameter and latent parameter distributions are representative of our observed data by performing prior predictive checks, an example of which is shown for the A-C-3.0-0.05 family in \autoref{fig:priorpred}.
+Critically, both the range of values spanned by the prior distributions and the general shape of the distribution suggests that the data is well represented by the model and the prior distributions on the hyper and latent parameters.
+
+As likelihood function, we use the lognormal distribution $\mathrm{Lognorm}(\Sigma, \hat{\sigma}_\Sigma^2)$ as it a) has support on $[0,\infty)$, and b) produces normally-distributed variates in logarithm space, which is representative of the distribution of $\hat{\Sigma}$ for a given radial bin when we backwards-model and na\"ively bin our observations of $\hat{\Sigma}$ for that radial bin.
+
+\drafting{Write explicitly the posterior distribution we are sampling from.}
+
+\subsection{Latent Parameter Posterior Distributions}
+We compute the posterior distribution using \textsc{Stan}, ensuring that the sampler exits without encountering any divergences, and that all \drafting{R-hat (is a summary stat, but how to draw?
+Don't want to confuse with observed radius)} values are within the range \drafting{$[0.99, 1.01]$}.
+We first inspect how the posterior distribution appears visually by computing posterior predictive checks, an example of which is shown in \autoref{fig:postpred} for the family A-C-3.0-0.05. 
+We find that the data is well described by the 50\% BCI \drafting{introduce what BCI is} of the posterior distribution, as desired.
+
+As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
+We compare the latent parameter distributions in \autoref{fig:compareAC} for the AC merger class\footnote{The merger family A-C-3.0-1.0 did not have enough \child samples (3 from 10) from which to construct the likelihood function, and thus is omitted here.}, with line colour corresponding to the first pericentre distance of the merger family, $r_\mathrm{peri}/R_{\mathrm{vir},1}$.
+We find no convincing evidence for an influence of first pericentre distance on the resulting projected mass density profiles, with the distributions of latent parameters consistent between all families in the AC progenitor class.
+Whilst we purposely avoid point estimates on our latent parameter distributions, we report the median values here to give a sense of scale. 
+For the AC progenitor class, we find the core radius to be $\sim 1.4\,\mathrm{kpc}$, effective radius $\sim 7.2\,\mathrm{kpc}$ (recall progenitor A had $R_\mathrm{e} \sim 4\,\mathrm{kpc}$), normalising density of $\sim 3.1\times10^9\,\Msun\,\mathrm{kpc}^{-2}$, core slope index of $\sim 0.05$, and S\'ersic index of $\sim 1.8$.
+
+\drafting{
+    \begin{itemize}
+        \item corner plot description
+        \item comparison between rperi values, maybe new section?
+    \end{itemize}
+}
+
 \begin{figure*}
@@
 \end{figure*}
+
 \begin{figure*}
@@
 \end{figure*}
+
 \begin{figure*}
     \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
-    \caption{\drafting{Same as \ref{fig:corner_params_1}}}
+    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
     \label{fig:corner_params_2}
 \end{figure*}
+
+\begin{figure*}
+    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_2}.png}
+    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
+    \label{fig:corner_params_3}
+\end{figure*}
+
 \begin{figure*}
@@
 
-%\begin{figure*}
-%\includegraphics{binary_params_main}
-%\includegraphics{binary_params_secondary}
-%\includegraphics{binary_params_tertiary}
-%\caption{
-%\drafting{
-%Binary orbital parameters. Dashed line for triplet outer orbit (dashes not very %visible at the moment).
-%Lower panels are mergers occurring in separate galaxies.
-%Note the use of $1-e$ for the last panel.
-%}
-%}
-%\label{fig:binary_params}
-%\end{figure*}
+\begin{figure*}
+    \includegraphics{compare_AC_gqs.png}
+    \caption{\drafting{Comparing the latent parameters for different members of the same progenitor class.
+Here shown is those families part of the AC merger class.
+There is no significant difference in latent parameter values with increasing rperi.
+TODO do this same plot for each merger class.}}
+    \label{fig:compareAC}
+\end{figure*}

### 2022-09-06 paper/sections/stats.tex
Switch to MNRAS template
- Discuss more about statistical implementation

- Change project name from Himmel

@@
 \subsection{Hamiltonian Monte Carlo}
-An excellent overview of HMC may be found in \citet{betancourt17}, with a brief description provided here.
+An excellent overview of HMC may be found in \citet{betancourt17}, with a brief description provided here. \drafting{Mention NUTS here somewhere?}
 
@@
 To align the gradient vector field with the typical set vector field, we can introduce auxillary momenta variables $p$, promoting our $D$-dimensional target distribution to a projection of a $2D$-dimensional canonical distribution in some phase space:
-\begin{equation*}
+\begin{equation}\label{eq:hmc}
     P(q,p) = P(p|q)P(q) = e^{-H(q,p)}
-\end{equation*}
+\end{equation}
 where $H$ is an invariant Hamiltonian function. 
@@
 This formulation also lends itself to improved sampling efficiency by decomposing the Hamiltonian into a \textit{kinetic energy} term and a \textit{potential energy} term, with the latter equivalent to the logarithm of the target distribution.
-As the Hamiltonian is phase space conserving, they may be accurately integrated using a volume-preserving integration scheme, such as symplectic integrators. 
-Of these, the leapfrog scheme offers both a simple implementation and high accuracy over long integration times.
+\begin{align}
+    H(q,p) &= -\log P(q,p) \\
+    &= -\log \left( P(p|q)P(q) \right) \\
+    &= -\log P(p|q) -\log P(q)
+\end{align}
+As the Hamiltonian is phase space conserving, the Hamilton equations may be accurately integrated using a volume-preserving integration scheme, such as symplectic integrators. 
+Of these integrators, the leapfrog scheme offers both a simple implementation and high accuracy over long integration times.
 
 \subsection{Hierarchical Models}
+In general, a model, whether empirical or physical, is a mathematical description of some observed data involving one or more model parameters. 
+There are a vast number of methods, such as least-squares fitting and $\chi^2$-minimisation, that aim to recover the optimal or `best-fit' parameters for a model given some data. 
+Hierarchical models extend this concept by generalising parameters as realisations of an underlying multi-dimensional population distribution with some given shape parameters, or \textit{hyperparameters}.
+Given a sample of observations that may be reasonably assumed to be drawn from some larger population, we are able to use Bayesian analysis to recover the optimal latent parameters, constrained by prior information we have on the hyperparameters.
+
+In this work, we have 36 merger configuration families, with each family containing ten perturbed realisations of the same initial state. 
+As we only slightly perturb the SMBH phase space coordinates, each \child simulation can be thought of as a sampled observation of the general galaxy-merger configuration for its corresponding family.
+Thus, we are able to employ hierarchical modelling to recover latent parameter distributions for various observable quantities of the merger remnant.
+
+Robust Bayesian analysis relies on a well-informed choice of the (hyper-) parameter prior distributions. 
+Broadly speaking, there are three categories of prior distributions used:
+\begin{enumerate}
+    \item \textit{Uninformative} priors: no prior information is given to the prior distribution, and thus all physically-allowed values are equally likely.
+A uniform distribution over the parameter domain is used.
+    \item \textit{Weakly-informative} priors: a small amount of prior information is encoded into the prior distribution, however the distribution is chosen to have a large variance.
+This generally ensures parameter values close to the expected order of magnitude are favoured over values that are `insensibly' extreme.
+An example is a normal distribution $\mathcal{N}(\mu=0, \sigma^2=100)$ for a parameter than is expected to be in the range $[-1,1]$.
+    \item \textit{Strongly-informative} priors: a large amount of prior information is encoded into the prior distribution, with small variance.
+This can lead to severely-biased parameter estimates from Bayesian analysis due to excessive `weight' given to the prior distribution over the data.
+An example is a normal distribution $\mathcal{N}(\mu=0, \sigma^2=0.5)$ for a parameter than is expected to be in the range $[-1,1]$.
+\end{enumerate}
+In this work, we opt for weakly-informative priors: from the literature, we know the approximate range of `sensible' values that our model parameters may take, but ensure that the prior distribution variance is wide so as to not excessively weight our priors relative to our data.
+
+Additionally, we must also consider the shape of each parameter's prior distribution, as parameters are generally constrained by some physical condition. 
+For example, a normalising radius $a$ cannot take a value less than 0, so a normal distribution, which has an unbounded domain, is an inappropriate choice for the prior distribution of this variable.
+Choosing a log-normal or gamma distribution with support on $[0,\infty)$ would represent a physically-motivated choice for the prior distribution here.
+
+\subsection{Predictive Checks}
+\drafting{Check names: forward fold, push forward}
+To ensure consistency between the (hyper-) parameter prior distributions and the observed data, we perform both prior-predictive and posterior-predictive checks of our modelling process.
+
+Prior predictive checks involve sampling model parameters directly from the model prior distributions, and ensuring that the observed dependent variable can be reasonably explained by those prior distributions given the observed independent variable as input \citep{gabry19}.
+If a model is described by a set of parameters $\theta$ which are drawn from a prior distribution $P$, then the model parameters are sampled according to $\theta^\mathrm{sim} \sim P(\theta)$. 
+Simulated data $y^\mathrm{sim}$ may then be sampled from the sampling distribution $Q$ given the set $\theta^\mathrm{sim}$, so that $y^\mathrm{sim} \sim Q(y | \theta^\mathrm{sim})$.
+
+Posterior predictive checks are similar to prior predictive checks, except that the posterior distributions that are recovered from the Bayesian analysis are sampled instead of the prior distributions of the model, and are thus conditioned on the data \citep{gelman15}.
+For some data dependent $y$, new dependent data $y^\mathrm{rep}$ \textit{using the original independent data} is predicted following:
+\begin{equation}
+    P(y^\mathrm{rep} | y) = \int_\Omega P(y^\mathrm{rep} | \theta) P(\theta | y) \dd{\theta}
+\end{equation}
+
+Ideally, the the observed dataset $y$ should appear somehow typical of those datasets $r^\mathrm{rep}$ that are predicted.
+In this work, we visually classify a `good' fit as one where the data falls within the 50\% highest density interval (HDI, also referred to as the posterior density region). 
+The $\alpha$\% HDI region is defined \citep[e.g.][]{gelman15} as the region containing:
+\begin{enumerate}
+    \item $100(1-\alpha)$\% of the probability, and
+    \item the density within the HDI region is always greater than the density outside the region.
+\end{enumerate}   
+The HDI differs from a regular confidence interval in that a confidence interval does not impose condition (ii), making it unsuitable for describing distributions other than those that are unimodal and symmetric.
+
+In addition to visual inspection of the posterior predictive distribution, we wish to quantitatively assess the statistical ability of our model to capture the key features we are interested in.
+Critically, the exact form of such a test will depend on which features we define as being `important' for the model to capture.
+Typically though, a sense of how well a posterior distribution captures the first and second moments of a distribution (i.e., the mean and variance) are metrics we will be guided by.
+\begin{enumerate}
+    \item Does the posterior predictive distribution capture the \textit{location} of the data generating process?
+If our model predicts data $y^\mathrm{sim}$ significantly offset from our observed data $y$, then the model is not very useful in predicting future observations generated from the same data generating process as $y$.
+    \item Does the posterior predictive distribution capture the \textit{scale} of the data generating process? \drafting{What does this imply??}
+\end{enumerate}
+
+To answer these questions, we define a test statistic $T$ from which we determine a Bayes-equivalent of the frequentist $p$-value \citep{gelman15}.
+\begin{align}
+    p_\mathrm{Bayes} &= \mathrm{Pr}\left( T(y^\mathrm{rep}, \theta) \geq T(y, \theta)|y \right) \\
+    &= \iint I_{T(y^\mathrm{rep}, \theta) \geq T(y, \theta)|y} p\left( y^\mathrm{rep}|\theta\right) p\left( \theta|y\right) \dd{y^\mathrm{rep}} \dd{\theta}
+\end{align}
+Here, $I$ is the indicator function.
+
+\subsection{Assessing the HMC}
+\drafting{Probably cite \citet{betancourt17} here, maybe others.
+Is this section necessary, or saying `default stan exit conditions' sufficient?
+Don't want to just be rewording stan docs, but also want to be transparent}.
+
+In determining whether or not the multi-dimensional distribution described by a hierarchical model has been well sampled, a number of checks are performed by \textsc{Stan}, which we describe briefly here.
+
+\drafting{
     \begin{itemize}
-    \item Data doesn't have error, model has uncertainty
-    \item Require multiple observations from a population --> samples
-    \item Assume population described by a series of hyperparameters
-    \item Prior distributions on those hyperparameters
-    \item Forward fold through model, combine with likelihood function from data
+        \item sampler tree depth
+        \item divergences
+        \item potential energy
+        \item effective sample size
+        \item split R-hat values
     \end{itemize}
-
-\drafting{\begin{enumerate}
-    \item HMC algorithm?
-    \item Description of hierarchical model
-    \item Prior distributions
-    \item forward modelling
-\end{enumerate}
 }

### 2022-09-21 papers/paper-1/diagrams/hm_coresersic.tex
Divide paper into two papers, better division of scientific message

\documentclass[border=10pt]{standalone}
\usepackage{newtxtext, newtxmath}
\usepackage{tikz}
\usetikzlibrary{bayesnet}

\begin{document}
\begin{tikzpicture}

    % define nodes
    % define layer 1
    \node[obs] (Shat) {$\hat{\Sigma}$};
    \factor[below=of Shat] {S_f} {right:Lognorm} {} {};
    \node[det, below=of S_f] (S) {$\Sigma$};
    \node[obs, below=of S] (R) {$\hat{R}$};
    \node[obs, right=of S] (sigma) {$\hat{\sigma_\Sigma}^2$};

    % define layer 2
    \node[latent, left=of S] (Sb) {$\tilde{\Sigma}_\mathrm{b}$};
    \node[latent, below=of Sb] (Re) {$R_\mathrm{e}$};
    \node[latent, below=of Re] (rb) {$r_\mathrm{b}$};
    \node[latent, above=of Sb] (g) {$\gamma$};
    \node[latent, above=of g] (n) {$n$};

    % define layer 3
    \factor[left=of rb] {rb_f} {below:$\Gamma$} {} {};
    \node[hyper, left=of rb_f, yshift=-1.2em] (rba) {$r_{\mathrm{b},\alpha}$};
    \node[hyper, left=of rb_f, yshift=1.2em] (rbb) {$r_{\mathrm{b},\beta}$};

    \factor[left=of Re] {Re_f} {below:$\Gamma$} {} {};
    \node[hyper, left=of Re_f, yshift=-1.2em] (Rea) {$R_{\mathrm{e},\alpha}$};
    \node[hyper, left=of Re_f, yshift=1.2em] (Reb) {$R_{\mathrm{e},\beta}$};

    \factor[left=of Sb] {Sb_f} {below:$\Gamma$} {} {};
    \node[hyper, left=of Sb_f, yshift=-1.2em] (Sba) {$\tilde{\Sigma}_{\mathrm{b},\alpha}$};
    \node[hyper, left=of Sb_f, yshift=1.2em] (Sbb) {$\tilde{\Sigma}_{\mathrm{b},\beta}$};

    \factor[left=of g] {g_f} {below:$\Gamma$} {} {};
    \node[hyper, left=of g_f, yshift=-1.2em] (ga) {$\gamma_\alpha$};
    \node[hyper, left=of g_f, yshift=1.2em] (gb) {$\gamma_\beta$};

    \factor[left=of n] {n_f} {below:$\Gamma$} {} {};
    \node[hyper, left=of n_f, yshift=-1.2em] (na) {$n_\alpha$};
    \node[hyper, left=of n_f, yshift=1.2em] (nb) {$n_\beta$};

    % connect the nodes
    \factoredge {sigma, S} {S_f} {Shat};

    \edge {rb, Re, Sb, g, n, R} {S};

    \factoredge {rba, rbb} {rb_f} {rb};
    \factoredge {Rea, Reb} {Re_f} {Re};
    \factoredge {Sba, Sbb} {Sb_f} {Sb};
    \factoredge {ga, gb} {g_f} {g};
    \factoredge {na, nb} {n_f} {n};

    % plates
    \plate {rS} {(R)(S)(sigma)(Shat)} {$\forall R,\sigma_\Sigma^2$};
    \plate {pars} {(R)(S)(sigma)(rb)(Re)(Sb)(g)(n)} {$\forall$ child};

\end{tikzpicture}
\end{document}

### 2022-09-21 papers/paper-1/paper-1.tex
Divide paper into two papers, better division of scientific message

% mnras_template.tex 
%
% LaTeX template for creating an MNRAS paper
%
% v3.0 released 14 May 2015
% (version numbers match those of mnras.cls)
%
% Copyright (C) Royal Astronomical Society 2015
% Authors:
% Keith T. Smith (Royal Astronomical Society)

% Change log
%
% v3.0 May 2015
%    Renamed to match the new package name
%    Version number matches mnras.cls
%    A few minor tweaks to wording
% v1.0 September 2013
%    Beta testing only - never publicly released
%    First version: a simple (ish) template for creating an MNRAS paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Basic setup. Most papers should leave these options alone.
\documentclass[fleqn,usenatbib]{mnras}

% MNRAS is set in Times font. If you don't have this installed (most LaTeX
% installations will be fine) or prefer the old Computer Modern fonts, comment
% out the following line
\usepackage{newtxtext,newtxmath}
% Depending on your LaTeX fonts installation, you might get better results with one of these:
%\usepackage{mathptmx}
%\usepackage{txfonts}

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}

% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
\DeclareRobustCommand{\VAN}[3]{#2}
\let\VANthebibliography\thebibliography
\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}

%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%

% Only include extra packages if you really need them. Common packages are:
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%

% Please keep new commands to a minimum, and use \newcommand not \def to avoid
% overwriting existing commands. Example:
%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
\newcommand{\ketju}{\textsc{Ketju} }
\newcommand{\mstar}{\textsc{Mstar} }
\newcommand{\gadget}{\textsc{Gadget-3} }
\newcommand{\parent}{\textit{parent }}
\newcommand{\child}{\textit{child }}

\newcommand{\Msun}{\ensuremath{M_{\sun}}}

\newcommand{\dd}[1]{\ensuremath{\mathrm{d}#1}}
\newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}

\graphicspath{{figures/}}

% XXX: For highlighting draft notes, these should be removed when done
\usepackage{color}
\newcommand{\drafting}[1]{
{\leavevmode\color[RGB]{224, 77, 24}#1}
}

\newcommand{\phj}[1]{
{\leavevmode\color[RGB]{0,0,255}#1}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%

% Title of the paper, and the short title which is used in the headers.
% Keep the title short and informative.
\title[BAGGinS]{Bayesian Analysis of Gas-free Galaxies in Simulations I: Reviving Stochasticity}

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[A. Rawlings et al.]{
Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
Others,$^{1}$
\\
% List of institutions
$^{1}$
Department of Physics,
Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland
}

% These dates will be filled out by the publisher
\date{Accepted XXX. Received YYY; in original form ZZZ}

% Enter the current year, for the copyright statements etc.
\pubyear{2022}

% Don't change these lines
\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle

% Abstract of the paper
\begin{abstract}
We present the Bayesian Analysis of Gas-free Galaxies in Simulations (BAGGinS) project, show that stochasticity is ubiquitous with collisionless merger sims.
\end{abstract}

% Select between one and six entries from the list of approved keywords.
% Don't make up new ones.
\begin{keywords}
keyword1 -- keyword2 -- keyword3
\end{keywords}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%

\input{./sections/intro}

\input{./sections/numerical}

\input{./sections/stats}

\input{./sections/results}

\input{./sections/discuss}

\input{./sections/conclusions}

\textit{Software:} \ketju \citep{rantala17}, \gadget \citep{springel05}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.

\section*{Acknowledgements}

P.H.J. acknowledge the support
by the European Research Council via ERC Consolidator Grant KETJU (no. 818930).

The numerical simulations used computational resources provided by
the CSC -- IT Center for Science, Finland.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Data Availability}

\drafting{The inclusion of a Data Availability Statement is a requirement for articles published in MNRAS. Data Availability Statements provide a standardised format for readers to understand the availability of data underlying the research results described in the article. The statement may refer to original data generated in the course of the study or to third-party data analysed in the article. The statement should describe and provide means of access, where possible, by linking to the data or providing the required accession numbers for the relevant databases or DOIs.}

%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%

% The best way to enter references is to use BibTeX:

\bibliographystyle{mnras}
\bibliography{../ref} % if your bibtex file is called example.bib

% Alternatively you could enter them by hand, like this:
% This method is tedious and prone to error if you have lots of references
%\begin{thebibliography}{99}
%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
%Others S., 2012, Journal of Interesting Stuff, 17, 198
%\end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%

\appendix

\section{Some extra material}

If you want to present additional material which would interrupt the flow of the main paper,
it can be placed in an Appendix which appears after the list of references.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}
\end{document}

% End of mnras_template.tex

### 2022-09-21 papers/paper-1/sections/intro.tex
Divide paper into two papers, better division of scientific message

\section{Introduction}

\drafting{Something cool}

\drafting{
    Introduce core sersic profile here.
}
\begin{equation}\label{eq:coresersic}
    \Sigma(R) = \Sigma'\left[ 1+\left(\frac{r_\mathrm{b}}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[ -b_n \left(\frac{R^\alpha+r_\mathrm{b}^\alpha}{R_\mathrm{e}^\alpha}\right)^{1/(\alpha n)} \right],
\end{equation}
where
\begin{equation}\label{eq:coresersic2}
    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[ b_n \left( 2^{1/\alpha} r_\mathrm{b}/R_\mathrm{e} \right)\right]
\end{equation}

### 2022-09-21 papers/paper-1/sections/numerical.tex
Divide paper into two papers, better division of scientific message

\section{Numerical Simulations}

\drafting{Describe the numerical set up, code, etc}

### 2022-09-21 papers/paper-1/sections/results.tex
Divide paper into two papers, better division of scientific message

--- 2022-09-06 paper/sections/results.tex
@@
 \item Example of forward-modelled velocity dispersion profile
+\item comparison latent parameters to SAMI (n, Re) and Thomas (rb)
 \end{itemize}
@@
 As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
-We compare the latent parameter distributions in \autoref{fig:compareAC} for the AC merger class\footnote{The merger family A-C-3.0-1.0 did not have enough \child samples (3 from 10) from which to construct the likelihood function, and thus is omitted here.}, with line colour corresponding to the first pericentre distance of the merger family, $r_\mathrm{peri}/R_{\mathrm{vir},1}$.
-We find no convincing evidence for an influence of first pericentre distance on the resulting projected mass density profiles, with the distributions of latent parameters consistent between all families in the AC progenitor class.
-Whilst we purposely avoid point estimates on our latent parameter distributions, we report the median values here to give a sense of scale. 
 For the AC progenitor class, we find the core radius to be $\sim 1.4\,\mathrm{kpc}$, effective radius $\sim 7.2\,\mathrm{kpc}$ (recall progenitor A had $R_\mathrm{e} \sim 4\,\mathrm{kpc}$), normalising density of $\sim 3.1\times10^9\,\Msun\,\mathrm{kpc}^{-2}$, core slope index of $\sim 0.05$, and S\'ersic index of $\sim 1.8$.
@@
 \end{figure*}
-
-\begin{figure*}
-    \includegraphics{compare_AC_gqs.png}
-    \caption{\drafting{Comparing the latent parameters for different members of the same progenitor class.
-Here shown is those families part of the AC merger class.
-There is no significant difference in latent parameter values with increasing rperi.
-TODO do this same plot for each merger class.}}
-    \label{fig:compareAC}
-\end{figure*}

### 2022-09-21 papers/paper-2/sections/results.tex
Divide paper into two papers, better division of scientific message

\section{Results}

\drafting{Have some nice plots, groundbreaking results}

\begin{figure*}
    \includegraphics{compare_AC_gqs.png}
    \caption{\drafting{Comparing the latent parameters for different members of the same progenitor class. Here shown is those families part of the (approximately equal-mass) AC merger class. There is no significant difference in latent parameter values with increasing rperi. TODO do this same plot for each merger class.}}
    \label{fig:compareAC}
\end{figure*}

\begin{figure*}
    \includegraphics{compare_AD_gqs.pdf}
    \caption{\drafting{Same as \autoref{fig:compareAC}, but for the (unequal mass) AD merger class. For merger families with $R_\mathrm{peri}/R_{\mathrm{vir},1}\leq0.1$, the latent parameter distributions are indistinguishable. COnversely, for the A-D-3.0-1.0 family, the effective radius has its mode at lower values than the other merger class members, and its S\'ersic index peaks at slightly higher values of $n$.}}
    \label{fig:compareAD}
\end{figure*}

### 2022-10-07 papers/paper-1/paper-1.tex
Update gitignore to prevent caching paper pdf
Edits to paper 1

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\ketju}{\textsc{Ketju} }
-\newcommand{\mstar}{\textsc{Mstar} }
-\newcommand{\gadget}{\textsc{Gadget-3} }
-\newcommand{\parent}{\textit{parent }}
-\newcommand{\child}{\textit{child }}
+\newcommand{\ketju}{\textsc{Ketju}}
+\newcommand{\mstar}{\textsc{Mstar}}
+\newcommand{\gadget}{\textsc{Gadget-3}}
+\newcommand{\parent}{\textit{parent}}
+\newcommand{\child}{\textit{child}}
 
 \newcommand{\Msun}{\ensuremath{M_{\sun}}}
+\newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}
 
@@
 % Keep the title short and informative.
-\title[BAGGinS]{Bayesian Analysis of Gas-free Galaxies in Simulations I: Reviving Stochasticity}
+\title[BAGGinS-I]{Bayesian Analysis of Gas-free Galaxies in Simulations I: Reviving Stochasticity}
 
@@
 
-\section{Some extra material}
-
-If you want to present additional material which would interrupt the flow of the main paper,
-it can be placed in an Appendix which appears after the list of references.
+\input{sections/appendix}
 

### 2022-10-07 papers/paper-1/sections/appendix.tex
Edits to paper 1

--- 2022-08-26 paper/sections/results.tex
@@
-\label{sec:results}
-
-\drafting{
-Figures to do
-\begin{itemize}
-\item Example of forward-modelled semimajor axis and eccentricity
-\item Example of forward-modelled density profile
-\item Example of forward-modelled velocity dispersion profile
-\end{itemize}
-Table of galaxy properties,
-and maybe another one of BH merger remnant properties.
-}
-
-\drafting{
-    WHAT ARE THE MAIN OBSERVATIONS WE SEE and WHAT CONCLUSIONS CAN WE DRAW?
-    \begin{enumerate}
-        \item stability of hardening rate -> semimajor axis can be robustly predicted
-        \item wide variance of eccentricity -> N-body models offer no quantitative prediction on likelihood of merger, can only comment on what merger looks like \textit{IF} it happens
-        \item variance in observables: trends between observable properties and what observers can actually make use of (don't have this yet), e.g. signatures of SMBH scouring in velocity profile?
-    \end{enumerate}
-}
+\section{Prior Predictive Checks for Projected Mass Density}
+Here we show the prior predictive check for the projected mass density of the H-1.00 system, as well the corner plots of the latent parameter hyperdistributions.
 
@@
 \end{figure*}
+
 \begin{figure*}
     \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
-    \caption{\drafting{Same as \ref{fig:corner_params_1}}}
+    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
     \label{fig:corner_params_2}
 \end{figure*}
+
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_posterior_pred_log10_proj_density_mean}.png}
-    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit.
-Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
-The observed data is overlaid, and seen to lie well within the 50\% BCI.
-}}
-    \label{fig:postpred}
+    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_2}.png}
+    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
+    \label{fig:corner_params_3}
 \end{figure*}
-
-%\begin{figure*}
-%\includegraphics{binary_params_main}
-%\includegraphics{binary_params_secondary}
-%\includegraphics{binary_params_tertiary}
-%\caption{
-%\drafting{
-%Binary orbital parameters. Dashed line for triplet outer orbit (dashes not very %visible at the moment).
-%Lower panels are mergers occurring in separate galaxies.
-%Note the use of $1-e$ for the last panel.
-%}
-%}
-%\label{fig:binary_params}
-%\end{figure*}

### 2022-10-07 papers/paper-1/sections/numerical.tex
Edits to paper 1

@@
 
-\drafting{Describe the numerical set up, code, etc}
+\subsection{Initial Conditions}
+
+\begin{table*}
+    \centering
+    \caption{Parameters of Hernquist sphere galaxy models used in the sample.}
+    \label{tab:hernquist_pars}
+    \begin{tabular}{ccccccc}
+        \hline
+        Name & $m_\star / (10^5 \Msun)$ & $N_\star/10^6$ & $\varepsilon_\star$/pc & $m_\mathrm{DM} / (10^7 \Msun)$ & $N_\mathrm{DM}/10^6$ & $\varepsilon_\mathrm{DM}$/pc\\
+        \hline
+        H-1.00 & 1.0 & 1.00 & 3.5 & 3.0 & 2.54 & 200.0 \\
+        H-0.50 & 2.0 & 0.50 & 3.5 & 6.0 & 1.27 & 200.0 \\
+        H-0.25 & 4.0 & 0.25 & 3.5 & 12.0 & 0.64 & 200.0 \\
+        H-0.10 & 10.0 & 0.10 & 7.0 & 30.0 & 0.25 & 200.0 \\
+        H-0.05 & 20.0 & 0.05 & 7.0 & 60.0 & 0.13 & 200.0 \\
+        \hline
+    \end{tabular}
+\end{table*}
+
+We simulate the merger of two identical galaxies. 
+Each galaxy is modelled as a spherically-symmetric multicomponent system: a stellar component embedded within a DM halo, and a single SMBH at the centre of the galaxy. 
+The total stellar mass is set to $M_\star=10^{11}\,\Msun$, the total DM mass is set to $M_\mathrm{DM} \sim 8\times 10^{12}\,\Msun$, and the SMBH mass is set to $M_\bullet \sim 8.8\times10^8\,\Msun$.
+Five realisations of the system are generated with varying mass resolutions between the stellar (and DM) particles and the SMBH particle. 
+Both the stellar and DM components have a density profile given by an ergodic Hernquist sphere parametrised by scale radius $a$:
+\begin{equation}\label{eq:hernquist}
+    \rho = \frac{2M}{4\pi} \frac{a}{r(r+a)^3}.
+\end{equation}
+We use a value of $a_\star=4.0\,\mathrm{kpc}$ for the stellar component, and a value of $a_\mathrm{DM}=300.0\,\mathrm{kpc}$ for the DM component. 
+We Monte Carlo sample the phase space distribution function
+\begin{equation}\label{eq:eddington}
+    f(\mathcal{E}) = \frac{1}{\sqrt{8}\pi^2} \left[ \int_0^\mathcal{E} \dv{\Psi}{\sqrt{\mathcal{E}-\Psi}} \dv{^2\nu_\mathcal{E}}{\Psi^2} + \frac{1}{\sqrt{\mathcal{E}}} \left(\dv{\nu_\mathcal{E}}{\Psi}\right)_{\Psi=0}\right],
+\end{equation}
+to generate position and velocity coordinates for all stellar and DM particles.
+As mentioned, the SMBH is initialised at rest at the origin.
+Additionally, we give the SMBH a randomly-drawn value of the dimensionless spin parameter $\chi$ following \drafting{one of the Zlochower papers}, with shape parameters $\alpha=10.5868, \beta=4.66884$.
+This distribution has a mode at $\chi \sim 0.8$.
+
+We simulate the equal-mass merger between like-mass resolution systems for each mass resolution listed in \autoref{tab:hernquist_pars}.
+The progenitor galaxies in each merger are set to an initial separation of $3.0\Rvir$ on quasi-Keplerian orbit, assuming each galaxy is a point-mass containing half its total mass. 
+The initial orbital eccentricity is chosen to be almost radial ($e_0 = 0.9999$), and the first pericentre distance $r_\mathrm{peri}$ is set to $10^{-3}\Rvir$.
+
+\subsection{Simulations}
+In typical N-body simulations, one tests the sensitivity of the model by Monte Carlo sampling the phase space distribution function given by \autoref{eq:eddington} a number of times with different random seeds.
+This is done to quantify the uncertainty in the model due to the discrete sampling of the phase space of field particles, which induce a Brownian motion to the orbit of the more massive SMBH particle.
+We observe that resampling \autoref{eq:eddington} is in effect displacing the centre of mass (CoM) of the system with the respect to the SMBH.
+To avoid the need to run a large number of simulations from the large $3.0\Rvir$ initial separation, we instead \textit{displace the SMBH} with respect to the CoM of the system at a time $t_\mathrm{perturb}$ after the simulation has begun.
+This is done by measuring the Brownian-motion induced displacement in position and velocity space of the SMBH from the CoM for a galaxy in isolation.
+After identifying the displacement along each axis to be Gaussian-distributed, we use the variance of this distribution to reposition the SMBH relative to the CoM. 
+We thus `recentre' the SMBH on the CoM with a perturbation of $\mathcal{N}_\text{pos}(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc}^2)$ for each position axis, and $\mathcal{N}_\text{vel}(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km}^2\text{\,s}^{-2})$ for each velocity axis.
+In this way, we mimic the effect of Brownian motion for a single run beginning at large separation (the \parent{} run) to generate an arbitrary number of perturbed \child{} runs. 
+The perturbation is applied no later than the second-to-last pericentre passage before the SMBH particles form a bound binary system, and ensure the separation of the SMBHs at the time of the perturbation is of the order 1000 larger than the perturbation applied.
+
+The progenitors are evolved through the \textit{parent phase} from $t(r_0)=0$ with \gadget{} until the SMBHs form a bound binary at $t_\text{stop}$, with a stellar softening of $\epsilon_\star=10\,\text{pc}$ and \gadget integration tolerances \texttt{ErrTolIntAccuracy}=$0.02$ and \texttt{ErrTolForceAcc}=$0.005$.
+We then initialise the \textit{child phase} by restarting from the snapshot closest to but earlier than $t=t_\mathrm{perturb}$.
+We apply the small perturbation to each SMBH position and velocity vector in the simulation, from which we create ten uniquely-perturbed \child{} runs, allowing us to explore the distribution of parameters describing both the orbital and remnant properties of each \parent{} merger.
+Additionally, we lower the stellar softening to the values presented in \autoref{tab:hernquist_pars} and tighten the integration tolerance \texttt{ErrTolIntAccuracy} to $0.002$ to optimise the run for \ketju{}.
+We check for energy conservation between the \parent{} and \child{} runs, and confirm that the total energy is conserved at the sub-percent level. 
+In total, we perform 50 unique \child{} simulations using \ketju{}, and use this sample for our analysis.
+
+Within a relatively short time frame from beginning the \child{} run, the SMBH binary hardens predominantly through dynamical friction and three-body stellar interactions. 
+We evolve the binary to the time $t_{a_h}=t(a_h)$, and continue to when $t_\text{safe}=t(a=10\,\text{pc})$. 
+We then analytically predict the orbital evolution of the binary by determining the hardening constant $H$ in \autoref{eq:quinlan_H} and eccentricity constant $K$ in \autoref{eq:quinlan_K} by fitting the derivative of $\text{d}(1/a)/\text{d}t$ with a linear function between $t_{a_h}$ and $t_\text{safe}$.
+Using $a_0=a_h$ and three different values of $e_0$, taken to be the 5\%, 50\%, and 95\% quantiles of the eccentricity distribution within the same time interval 
+$t_{a_h} \text{--} t_\text{safe}$, we predict the expected merger timescale of the system. 
+We continue to run the \child runs until one of three conditions are met:
+\begin{enumerate}
+    \item The SMBHs merge, or
+    \item The Hubble time, $13.8\,\text{Gyr}$, is exceeded, or
+    \item The eccentricity stabilises to below the analytically-determined eccentricity estimated from the 5\%-quantile eccentricity value, and said analytical-estimate indicates that the merger will occur significantly after the Hubble time.
+\end{enumerate}
+If it is not clear if the third condition is satisfied, we continue to run the merger until the second condition is met.

### 2022-10-07 papers/paper-1/sections/results.tex
Edits to paper 1

@@
-\label{sec:results}
+\section{Results}\label{sec:results}
 
@@
 \item Example of forward-modelled semimajor axis and eccentricity
-\item Example of forward-modelled density profile
 \item Example of forward-modelled velocity dispersion profile
-\item comparison latent parameters to SAMI (n, Re) and Thomas (rb)
 \end{itemize}
@@
 The nodes indicate the distribution used in constructing the prior distributions (for latent parameters) and likelihood function (for $\hat{\Sigma}$).
-The radial box represents observed and latent variables for each radial bin for a \child, and the child box represents latent parameters for each \child simulation in a family.
+The radial box represents observed and latent variables for each radial bin for a \child{}, and the child box represents latent parameters for each \child{} simulation in a family.
 }
@@
 These variables are the radius $\hat{R}$, and the mean projected mass density $\hat{\Sigma}$, and the standard deviation of the projected mass density $\hat{\sigma}_\Sigma$ measured in the radial bins centred on $\hat{R}$.
-For each \child simulation, we choose to `observe' the first snapshot where the binary semimajor axis is \textit{at most} $10\,\mathrm{pc}$. 
-In the event that there is no snapshot satisfying the above requirement (due to, for example, rapid merging of the SMBH binary in a time period shorter than the snapshot output frequency), the \child simulation is not included in the analysis.
-We enforce a minimum of six \child simulations per family for that family to be analysed with our hierarchical modelling technique. 
-For those \child simulations which pass our selection criteria, we take 30 random projections of the merger remnant, and determine the projected mass density in 50 radial bins evenly-spaced in logarithm space, with a minimum value of $0.2\,\mathrm{kpc}$ and a maximum value of $20.0\,\mathrm{kpc}$.
+For each \child{} simulation, we choose to `observe' the first snapshot where the binary semimajor axis is \textit{at most} $10\,\mathrm{pc}$. 
+In the event that there is no snapshot satisfying the above requirement (due to, for example, rapid merging of the SMBH binary in a time period shorter than the snapshot output frequency), the \child{} simulation is not included in the analysis.
+We enforce a minimum of six \child{} simulations per family for that family to be analysed with our hierarchical modelling technique. 
+For those \child{} simulations which pass our selection criteria, we take 30 random projections of the merger remnant, and determine the projected mass density in 50 radial bins evenly-spaced in logarithm space, with a minimum value of $0.2\,\mathrm{kpc}$ and a maximum value of $20.0\,\mathrm{kpc}$.
 This allows us to build up a distribution of projected mass densities per radial bin, from which $\hat{\Sigma}$ and $\hat{\sigma}_\Sigma$ are determined.
@@
 \subsubsection{Latent Parameters for Observables}
-With the omission of the parameter $\alpha$ in the modelling process, there are five latent parameters ($r_\mathrm{b}$, $R_\mathrm{e}$, $\tilde{\Sigma}_\mathrm{b}$, $\gamma$, and $n$, the set of which we denote as $\kappa$) that describe each \child simulation in a family.
+With the omission of the parameter $\alpha$ in the modelling process, there are five latent parameters ($r_\mathrm{b}$, $R_\mathrm{e}$, $\tilde{\Sigma}_\mathrm{b}$, $\gamma$, and $n$, the set of which we denote as $\kappa$) that describe each \child{} simulation in a family.
 Here, we define $\tilde{\Sigma}_\mathrm{b} \equiv \Sigma_\mathrm{b}/10^{10}$, where $\Sigma_\mathrm{b}$ is defined in \autoref{eq:coresersic2}.
@@
 
-We ensure our hyperparameter and latent parameter distributions are representative of our observed data by performing prior predictive checks, an example of which is shown for the A-C-3.0-0.05 family in \autoref{fig:priorpred}.
+We ensure our hyperparameter and latent parameter distributions are representative of our observed data by performing prior predictive checks, an example of which is shown for the H-1.00 family in \autoref{fig:priorpred}.
 Critically, both the range of values spanned by the prior distributions and the general shape of the distribution suggests that the data is well represented by the model and the prior distributions on the hyper and latent parameters.
@@
 
+\begin{figure*}
+    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_posterior_pred_log10_proj_density_mean}.png}
+    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit.
+Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
+The observed data is overlaid, and seen to lie well within the 50\% BCI.
+}}
+    \label{fig:postpred}
+\end{figure*}
+
 \subsection{Latent Parameter Posterior Distributions}
@@
 Don't want to confuse with observed radius)} values are within the range \drafting{$[0.99, 1.01]$}.
-We first inspect how the posterior distribution appears visually by computing posterior predictive checks, an example of which is shown in \autoref{fig:postpred} for the family A-C-3.0-0.05. 
+We first inspect how the posterior distribution appears visually by computing posterior predictive checks, an example of which is shown in \autoref{fig:postpred} for the family H-1.00. 
 We find that the data is well described by the 50\% BCI \drafting{introduce what BCI is} of the posterior distribution, as desired.
@@
 
-\begin{figure*}
-    \includegraphics[width=\textwidth]{{graham_density-A-C-3.0-0.05_prior_pred_log10_proj_density_mean}.png}
-    \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit.
-Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
-The observed data is overlaid.
-From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.}}
-    \label{fig:priorpred}
-\end{figure*}
+\begin{figure}
+    \includegraphics{compare_binaries.pdf}
+    \caption{\drafting{Show an example of the scatter we get from perturbations.
+These are plotted with t=0 corresponding to when the child run started.
+Best way to show this, or should a time-shift be done?
+This way at least we see clearly that different sims become bound at different times.}}
+    \label{fig:ketju_scatter}
+\end{figure}
 
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_0}.png}
-    \caption{\drafting{Corner plot of hierarchical model hyperparameters.
-Each of the hyperdistributions are modelled as Gamma distributions with shape parameters $\alpha$ and $\beta$.
-For example, the hyperdistribution on the core radius $r_\mathrm{b}$ is modelled as $r_\mathrm{b} \sim \text{Gamma}(r_{\mathrm{b},\alpha}, r_{\mathrm{b},\beta})$.
-Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well.
-}}
-    \label{fig:corner_params_1}
-\end{figure*}
+\subsection{Binary Properties}
+After considering the robustness of the projected density profile to SMBH Brownian motion, we turn our attention to a similar analysis of the orbital parameters of the SMBH binary, namely the semimajor axis $a$ and the eccentricity $e$.
+It is instructive to first consider the raw data of a particular simulation family, shown in \autoref{fig:ketju_scatter} for the case of H-1.00. 
+Immediately apparent is the scatter in eccentricity between different realisations, with some runs (e.g.
+006) having $e<0.2$ for the majority of the integration time, whilst other realisations (e.g.
+002) have $e>0.9$ for the same time period.
+As discussed in \drafting{section}, the efficiency of GW as a hardening mechanism is directly related to the binary eccentricity, and is highly non-linear. 
+This is observed in those \child{} realisations that merged, and the timescale within which they merged.
+As seen in \autoref{fig:ketju_scatter}, a higher eccentricity leads to more rapid BH binary mergers as opposed to a lower eccentricity. 
+Out of the 50 simulations that we ran, we observe \drafting{XXX} systems that resulted in SMBH binary coalescence.
 
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
-    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
-    \label{fig:corner_params_2}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_2}.png}
-    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
-    \label{fig:corner_params_3}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_posterior_pred_log10_proj_density_mean}.png}
-    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit.
-Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
-The observed data is overlaid, and seen to lie well within the 50\% BCI.
-}}
-    \label{fig:postpred}
-\end{figure*}
+Also of note is the rapid oscillation in eccentricity between $\sim100$--$300\,\mathrm{Myr}$. 
+This is contrasted by a relatively smooth evolution in semimajor axis, indicating that the SMBH binary orbit is highly sensitive to the surrounding stellar mass distribution.

### 2022-10-11 papers/paper-1/paper-1.tex
Update figures to reflect hernquist runs

@@
 
-\textit{Software:} \ketju \citep{rantala17}, \gadget \citep{springel05}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.
+\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel05}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.
 

### 2022-10-11 papers/paper-1/sections/results.tex
Update figures to reflect hernquist runs

@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_posterior_pred_log10_proj_density_mean}.png}
-    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit.
+    \includegraphics[width=0.7\textwidth]{{graham_density-H-H-3.0-0.001_posterior_pred_log10_proj_density_mean}.png}
+    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution.
 Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
@@
 
+\begin{figure*}
+    \includegraphics{compare_HH_gqs_res.png}
+    \caption{\drafting{Comparison of core sersic fit for different mass resolutions.
+Note the scatter does not seem to decrease with increasing mass resolution, however the distribution modes seem to converge towards something.
+y-axes should read PDF}}
+    \label{fig:cs_mass_compare}
+\end{figure*}
+
 \subsection{Binary Properties}

### 2022-10-14 code/analysis_scripts/hierarchical_models/compare_graham_density.py
- Updates to hierarchical modelling for better comparisons

@@ -1,22 +1,28 @@
 import argparse
-from cmath import log
 import os.path
 import numpy as np
 import matplotlib.pyplot as plt
-from matplotlib.gridspec import GridSpec
+from matplotlib import rcParams
 import yaml
 
 import cm_functions as cmf
 
-
 parser = argparse.ArgumentParser(description="Compare Graham density fits of different mergers", allow_abbrev=False)
 parser.add_argument(help=".yaml file of stan files to compare", type=str, dest="files")
 parser.add_argument(help="family to compare", type=str, dest="fam")
-parser.add_argument("-c", "--colour-var", type=str, choices=["rperi", "res", "generic"], default="rperi", dest="cvar", help="variable to colour samples by")
+parser.add_argument("-c", "--colour-var", type=str, choices=["rperi", "res", "generic"], default="generic", dest="cvar", help="variable to colour samples by")
+parser.add_argument("-P", "--Publish", action="store_true", dest="publish", help="use publishing format")
 parser.add_argument("-v", "--verbosity", type=str, choices=cmf.VERBOSITY, dest="verbose", default="INFO", help="verbosity level")
 args = parser.parse_args()
 
 SL = cmf.ScriptLogger("script", console_level=args.verbose)
+
+if args.publish:
+    cmf.plotting.set_publishing_style()
+    full_figsize = rcParams["figure.figsize"]
+    full_figsize[0] *= 2
+else:
+    full_figsize = None
 
 figname_base = f"hierarchical_models/density/compare"
 stan_model_file = "stan/graham.stan"
@@ -47,42 +53,38 @@ except KeyError:
     raise
 SL.logger.debug(f"Colour values are {colour_var}")
 
-fig = plt.figure()
-ax = []
-gs = GridSpec(2, 6, figure=fig, top=0.95)
-for i in range(3):
-    ax.append(fig.add_subplot(gs[0, 2*i:2*(i+1)]))
-for i in range(2):
-    ax.append(fig.add_subplot(gs[1, 2*i+1:2*(i+1)+1]))
-ax = np.array(ax)
[…]

### 2022-10-14 code/analysis_scripts/hierarchical_models/graham_density.py

@@ -2,12 +2,12 @@ import argparse
 from datetime import datetime
 import os.path
 import numpy as np
+import scipy.optimize
 import h5py
+from matplotlib import rcParams
 import matplotlib.pyplot as plt
-from matplotlib.gridspec import GridSpec
 from arviz.labels import MapLabeller
 import cm_functions as cmf
-
 
 
 parser = argparse.ArgumentParser(description="Run stan model for Core-Sersic model.", allow_abbrev=False)
@@ -16,10 +16,19 @@ parser.add_argument(type=str, help="path to analysis parameter file", dest="apf"
 parser.add_argument("-p", "--prior", help="Plot for prior", action="store_true", dest="prior")
 parser.add_argument("-l", "--load", type=str, help="Load previous stan file", dest="load_file", default=None)
 parser.add_argument("-r", "--r", type=int, dest="random_sample", default=None, help="randomly sample x observations from each population member")
+parser.add_argument("-c", "--compare", help="Compare to naive statistics", action="store_true", dest="compare")
+parser.add_argument("-P", "--Publish", action="store_true", dest="publish", help="use publishing format")
 parser.add_argument("-v", "--verbosity", type=str, choices=cmf.VERBOSITY, dest="verbose", default="INFO", help="verbosity level")
 args = parser.parse_args()
 
 SL = cmf.ScriptLogger("script", console_level=args.verbose)
+
+if args.publish:
+    cmf.plotting.set_publishing_style()
+    full_figsize = rcParams["figure.figsize"]
+    full_figsize[0] *= 2
+else:
+    full_figsize = None
 
 HMQ_files = cmf.utils.get_files_in_dir(args.dir)
 with h5py.File(HMQ_files[0], mode="r") as f:
@@ -87,7 +96,7 @@ if args.prior:
     graham_model.sample_prior(data=stan_data, sample_kwargs=analysis_params["stan"]["sample_kwargs"])
 
     # prior predictive check
-    fig, ax = plt.subplots(1,1)
+    fig, ax = plt.subplots(1,1, figsize=full_figsize)
     ax.set_ylim(-1, 15.1)
     ax.set_xlabel("R/kpc")
[…]

### 2022-10-14 code/cm_functions/analysis/analysis_classes/StanModel.py

@@ -1,6 +1,7 @@
 import os
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib import rcParams
 from datetime import datetime
 import cmdstanpy
 import arviz as az
@@ -400,7 +401,7 @@ class StanModel:
         return self.generated_quantities.stan_variable(gq)
     
 
-    def parameter_plot(self, var_names, figsize=(9.0, 6.75), labeller=None, levels=[50, 90, 95, 99]):
+    def parameter_plot(self, var_names, figsize=None, labeller=None, levels=[50, 90, 95, 99]):
         """
         Plot key pair plots and diagnostics of a stan likelihood model.
 
@@ -413,6 +414,10 @@ class StanModel:
         levels : list, optional
             HDI intervals to plot, by default [50, 90, 95, 99]
         """
+        # TODO choose good figsize always, also labeller sometimes still not working...
+        if figsize is None:
+            max_dim = max(rcParams["figure.figsize"])
+            figsize = (max_dim, max_dim)
         assert len(var_names) > 1, "Pair plot requires at least two variables!"
         if len(var_names) > 4:
             _logger.logger.warning("Corner plots with more than 4 variables may not correctly map the labels given by the labeller!")
@@ -430,8 +435,8 @@ class StanModel:
 
         # plot pair
         levels = [l/100 for l in levels]
-        ax = az.plot_pair(self._fit_for_az, var_names=var_names, kind="scatter", marginals=True, scatter_kwargs={"marker":".", "markeredgecolor":"k", "markeredgewidth":0.5, "alpha":0.2}, figsize=figsize, labeller=labeller)
-        az.plot_pair(self._fit_for_az, var_names=var_names, kind="kde", divergences=True, ax=ax, point_estimate="mode", marginals=True, kde_kwargs={"contour_kwargs":{"linewidths":0.5}, "hdi_probs":levels}, point_estimate_marker_kwargs={"marker":""}, labeller=labeller)
+        ax = az.plot_pair(self._fit_for_az, var_names=var_names, kind="scatter", marginals=True, scatter_kwargs={"marker":".", "markeredgecolor":"k", "markeredgewidth":0.5, "alpha":0.2}, figsize=figsize, labeller=labeller, textsize=rcParams["font.size"])
[…]

### 2022-10-14 code/cm_functions/literature/density_profiles.py

@@ -4,7 +4,7 @@ import pygad
 from ..general import sersic_b_param
 
 
-__all__ = ["Dehnen", "fit_Dehnen_profile", "halfMassDehnen", "Terzic05", "fit_Terzic05_profile"]
+__all__ = ["Dehnen", "fit_Dehnen_profile", "halfMassDehnen", "Terzic05", "fit_Terzic05_profile", "core_Sersic_profile"]
 
 
 def Dehnen(r, a, g, M):
@@ -77,6 +77,23 @@ def halfMassDehnen(a, g):
     return rhm, re
 
 
+def sersic_b_param_approx(n):
+    """
+    Approximate value of the Sersic b parameter
+
+    Parameters
+    ----------
+    n : _type_
+        _description_
+
+    Returns
+    -------
+    float
+        value of the b parameter
+    """
+    #assert n > 0.5 and n < 10
+    return 2*n - 0.333333333 + 0.009876/n
+
 def Terzic05(r, rhob, rb, n, g, Re, b=None, a=100, mode="own"):
     """
     Define a density function for a cored system, with overall sersic profile
@@ -116,7 +133,7 @@ def Terzic05(r, rhob, rb, n, g, Re, b=None, a=100, mode="own"):
             b = sersic_b_param(n)
         else:
             #use an approximation when fitting
-            b = 2*n - 0.33 + 0.009876/n
+            b = sersic_b_param_approx(n)
     if isinstance(Re, pygad.UnitArr):
         Re = Re.view(np.ndarray)
     p = 1.0 - 0.6097/n + 0.05563/n**2
@@ -145,3 +162,36 @@ def fit_Terzic05_profile(r, density, Re, p0=[1e2, 1, 4, 1, 1], **kwargs):
     popt, param_cov = scipy.optimize.curve_fit(f, r, density, bounds=bounds, p0=p0, **kwargs)
     param_best = {"rhob": popt[0], "rb": popt[1], "n":popt[2], "g": popt[3], "a":popt[4]}
     return param_best
+
+
+def core_Sersic_profile(r, Re, rb, Ib, n, gamma, alpha=10.0):
+    """
+    Core Sersic profile from Graham 2003. Note this is a projected (2D) profile!
+    Note that no unit-consistency checks are done.
+
+    Parameters
+    ----------
+    r : array-like
+        radial values to determine profile for
+    Re : float
+        effective radius
+    rb : float
+        core radius
+    Ib : float
[…]

### 2022-10-14 code/cm_functions/plotting/general.py

@@ -1,13 +1,14 @@
 from collections import OrderedDict
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.gridspec import GridSpec
 from mpl_toolkits.axes_grid1.anchored_artists import AnchoredSizeBar
 from matplotlib import colors
 import itertools
 from ..env_config import _cmlogger
 
 
-__all__ = ["draw_sizebar", "create_normed_colours", "mplColours", "mplLines", "mplChars", "shade_bool_regions"]
+__all__ = ["draw_sizebar", "create_normed_colours", "mplColours", "mplLines", "mplChars", "shade_bool_regions", "create_odd_number_subplots"]
 
 _logger = _cmlogger.copy(__file__)
 
@@ -169,4 +170,41 @@ def shade_bool_regions(ax, xdata, mask, **kwargs):
     regions = [(group[0], group[-1]) for group in (list(group) for key, group in itertools.groupby(range(len(mask)), key=mask.__getitem__) if key)]
     for region in regions:
         ax.axvspan(xdata[region[0]], xdata[region[1]], **kwargs)
+
+
+def create_odd_number_subplots(nrow, ncol, fkwargs={}, gskwargs={}):
+    """
+    Create a set of subplots where the final row has an odd number of panels 
+    (1 less than the previous rows).
+
+    Parameters
+    ----------
+    nrow : int
+        number of rows
+    ncol : int
+        number of columns for all but the last row, which will have 1-ncol 
+        columns
+    fkwargs : dict, optional
+        kwargs dict to be parsed to plt.figure(), by default {}
+    gskwargs : dict, optional
+        kwargs dict to be parsed to GridSpec(), by default {}
+
+    Returns
+    -------
+    _type_
+        _description_
+    """
+    fig = plt.figure(**fkwargs)
+    ax = []
+    if "top" not in gskwargs:
+        gskwargs["top"] = 0.95
+    gs = GridSpec(nrow, 2*ncol, figure=fig, **gskwargs)
+    # "normal" rows
+    for j in range(nrow-1):
+        for i in range(ncol):
+            ax.append(fig.add_subplot(gs[j, 2*i:2*(i+1)]))
+    # final row, with odd number of panels
+    for i in range(ncol-1):
+        ax.append(fig.add_subplot(gs[-1, 2*i+1:2*(i+1)+1]))
+    return fig, np.array(ax)
 

### 2022-10-14 papers/paper-1/sections/results.tex
Update results section

@@
 \item Example of forward-modelled semimajor axis and eccentricity
-\item Example of forward-modelled velocity dispersion profile
 \end{itemize}
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-H-H-3.0-0.001_posterior_pred_log10_proj_density_mean}.png}
+    \includegraphics{{graham_density-H-H-3.0-0.001_posterior_pred_log10_proj_density_mean}.pdf}
     \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution.
@@
     \label{fig:postpred}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics{{graham_density-H-H-3.0-0.001_latentqty_compare}.pdf}
+    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats.}}
+    \label{fig:cs_compare_naive}
 \end{figure*}
@@
 As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
-For the AC progenitor class, we find the core radius to be $\sim 1.4\,\mathrm{kpc}$, effective radius $\sim 7.2\,\mathrm{kpc}$ (recall progenitor A had $R_\mathrm{e} \sim 4\,\mathrm{kpc}$), normalising density of $\sim 3.1\times10^9\,\Msun\,\mathrm{kpc}^{-2}$, core slope index of $\sim 0.05$, and S\'ersic index of $\sim 1.8$.
+We show an example of these latent parameter distributions in \autoref{fig:cs_compare_naive}, where the PDF of the latent parameter distributions is given in blue. 
+For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive Gaussian statistics.
+To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the mean and standard deviation of the sample. 
+The mean value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
+The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates: this is seen most clearly for the latent parameter $n$, whose Gaussian error estimate implies that by $2\sigma$, there is a non-zero probability of $n$ assuming a value less than 0, which is unphysical.
 
-\drafting{
+We show the latent parameter distributions for each mass resolution family in \autoref{fig:cs_mass_compare}, and compare the median and IQRs \drafting{ensure this acronym explained somewhere} in \autoref{fig:cs_mass_compare_2}.
+We find first that the spread in the latent parameters (as quantified through their IQR) does not dramatically differ between mass resolutions.
+Secondly, we observe that not all latent parameters in the core-S\"ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
+For other latent parameters, such as the core radius $r_\mathrm{b}$, there is a clear convergence of the median value with increasing mass resolution.
+\drafting{More particles --> core better resolved.
+Discuss here or in discussion?}
+
+Whilst we are unable to attribute the error in \autoref{fig:cs_mass_compare} to a particular phenomenon, we may at least rule out that the spread is due to post-processing analysis introduced from differing observation projections of the merger remnant.
+We re-determine the mean and standard deviation of the logarithm of the projected density profile using a subset of the total 30 projections we have available to us for the H-1.00 merger family.
+Repeating this data-reduction process four times, we then rerun the HMC algorithm for each subsample.
+We compare the latent parameter distributions for each subsample to the latent parameter distributions for the H-1.00 family determined using the full projection sample, and find:
+\begin{enumerate}
+    \item The error in the latent parameters is the consistent when using a subsample of projections, or the full sample, implying
+    \item The error in the latent parameters is independent of the projection direction.
+\end{enumerate}
+Consequently, we may attribute the uncertainty in the latent parameter distributions to one of:
 \begin{itemize}
-        \item corner plot description
-        \item comparison between rperi values, maybe new section?
+    \item Intrinsic scatter inherent to the physical system, or 
+    \item Numerical noise arising from a source independent of mass resolution, or 
+    \item A combination of the two.
 \end{itemize}
-}
+
+\begin{figure*}
+    \includegraphics{compare_HH_gqs_res.pdf}
+    \caption{\drafting{Comparison of core sersic fit for different mass resolutions.
+Note the scatter does not seem to decrease with increasing mass resolution, however the distribution modes seem to converge towards something.}}
+    \label{fig:cs_mass_compare}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics{compare_HH_gqs_res_2d.pdf}
+    \caption{\drafting{Same as \autoref{fig:cs_mass_compare}, but plotted as a function of mass resolution.
+Points correspond to the median estimate for the latent parameters, and error bars to the IQR.}}
+    \label{fig:cs_mass_compare_2}
+\end{figure*}
 
@@
 \end{figure}
-
-\begin{figure*}
-    \includegraphics{compare_HH_gqs_res.png}
-    \caption{\drafting{Comparison of core sersic fit for different mass resolutions.
-Note the scatter does not seem to decrease with increasing mass resolution, however the distribution modes seem to converge towards something.
-y-axes should read PDF}}
-    \label{fig:cs_mass_compare}
-\end{figure*}
 

### 2022-10-14 papers/paper-1/sections/stats.tex
Divide paper into two papers, better division of scientific message
Update results section

--- 2022-09-06 paper/sections/stats.tex
@@
+\section{Statistical Analysis}
+
 We perform full Bayesian Hierarchical Modelling for the projected mass density profiles of our simulated galaxies. 
@@
 
+\drafting{Now using LOO method. Which section should this go in? Relation to HDI?}
 In addition to visual inspection of the posterior predictive distribution, we wish to quantitatively assess the statistical ability of our model to capture the key features we are interested in.

### 2022-10-23 papers/paper-1/sections/appendix.tex
Update figures to reflect hernquist runs
Update results section
Small updates to paper for distribution to Portegies-Zwart

@@
 \begin{figure*}
-    \includegraphics[width=\textwidth]{{graham_density-A-C-3.0-0.05_prior_pred_log10_proj_density_mean}.png}
+    \includegraphics{{graham_density-H-H-3.0-0.001_gqs_prior}.pdf}
+    \caption{\drafting{Prior predictive distributions of latent parameters. Note the mode of the distributions roughly coincide with what one would expect from the literature.}}
+    \label{fig:prior_latent}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics{{graham_density-H-H-3.0-0.001_prior_pred_log10_proj_density_mean}.pdf}
     \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit.
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_0}.png}
+    \includegraphics{{graham_density-H-H-3.0-0.001_pair_0}.pdf}
     \caption{\drafting{Corner plot of hierarchical model hyperparameters.
@@
 Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well.
+The contours correspond to the 50\%, 90\%, 95\%, and 99\% BCIs.
 }}
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
+    \includegraphics{{graham_density-H-H-3.0-0.001_pair_1}.pdf}
     \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
@@
 \begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_2}.png}
+    \includegraphics{{graham_density-H-H-3.0-0.001_pair_2}.pdf}
     \caption{\drafting{Same as \autoref{fig:corner_params_1}}}

### 2022-10-23 papers/paper-1/sections/intro.tex
Edits to paper 1
Small updates to paper for distribution to Portegies-Zwart

@@
 
-\drafting{Something cool}
+\drafting{There was a paper by \citet{nasim20} about ``defeating stochasticity''.
+They use some questionable methods to extrapolate some orders of magnitude to say that at physical resolution (i.e.
+solar mass particles) there is no stochasticity in binary evolution.
+I'm of the opinion this is not the case.
+The title of this paper is sort of a rebuke to their work.}
 
 \drafting{
-    Introduce core sersic profile here.
+    Introduce necessary formulae
 }
@@
 \end{equation}
+
+\begin{equation}\label{eq:quinlan_H}
+    \dv{}{t}\left( \frac{1}{a} \right) = H \frac{G\rho}{\sigma}
+\end{equation}
+
+\begin{equation}\label{eq:quinlan_K}
+    \dv{e}{t} = -K a^{-1} \dv{a}{t}
+\end{equation}
+
+\begin{equation}\label{eq:kepler_L}
+    L = \mu\sqrt{GMa \left(1-e^2\right)}
+\end{equation}
+
+\begin{equation}\label{eq:influence_radius}
+    r_\mathrm{infl} = r_\star(<2 M)
+\end{equation}
+
+\begin{equation}\label{eq:a_hard}
+    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{infl}}{4},
+\end{equation}
+where $q=M_{\bullet,2}/M_{\bullet,1} < 1$.

### 2022-10-23 papers/paper-1/sections/results.tex
Small updates to paper for distribution to Portegies-Zwart

@@
 
-\drafting{ENSURE CONSISTENCY WITH SIGMA AND I. AS WE ARE USING MASS DENSITY, SHOULD BE SIGMA, NOT I, OR ASSUME A M/L RATIO OF 1. NEED TO GO THROUGH AND ENSURE THIS}
+\drafting{ENSURE CONSISTENCY WITH BCI AND HDI. SAME THING, JUST DIFFERENT NAME}
 
@@
     \includegraphics{{graham_density-H-H-3.0-0.001_latentqty_compare}.pdf}
-    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats.}}
+    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats. Note error bars on black points are smaller than the point itself if not visible}}
     \label{fig:cs_compare_naive}
@@
 For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive Gaussian statistics.
-To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the mean and standard deviation of the sample. 
-The mean value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
-The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates: this is seen most clearly for the latent parameter $n$, whose Gaussian error estimate implies that by $2\sigma$, there is a non-zero probability of $n$ assuming a value less than 0, which is unphysical.
+To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the median and IQR of the sample. 
+The median value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
+The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates, with some parameters (e.g.
+$n$) converging to the boundary value we limit the Least Squares method to (for $n$, the bounds are $[0,20]$).
+We try successively relaxing and tightening the boundary values of each parameter to obtain a `physically-realistic' value for each optimal latent parameter value, with no success. 
+Henceforth, we abandon any attempt to use a Least Squares minimisation to recover the optimal latent parameter values, returning instead to our Bayesian approach.
 
@@
 We find first that the spread in the latent parameters (as quantified through their IQR) does not dramatically differ between mass resolutions.
-Secondly, we observe that not all latent parameters in the core-S\"ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
+Secondly, we observe that not all latent parameters in the core-S\'ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
 For other latent parameters, such as the core radius $r_\mathrm{b}$, there is a clear convergence of the median value with increasing mass resolution.
@@
 This is contrasted by a relatively smooth evolution in semimajor axis, indicating that the SMBH binary orbit is highly sensitive to the surrounding stellar mass distribution.
+
+\drafting{
+    QUESTIONS FOR SIMON:\\
+    \begin{itemize}
+        \item Hierarchical models rely on the ability to model a set of observations as belonging to a \textit{population}.
+After the chaos talk, can a collection of BH-perturbed systems be described still as belonging to a population?
+I think yes (stellar mass distribution unchanged), but would be keen to hear thoughts.
+        \item In this paper we're perturbing the phase space coordinates of the BHs as opposed to a single field particle (we did this as another sensitivity test).
+Let's consider the case of perturbing a field particle.
+I find it still somehow curious that such a small displacement (a fraction of a softening length) still propagates the perturbation to such large scales, considering we are using a tree code for the particles not-near the BHs.
+I would have (na\"ively) thought that round off error from the tree algorithm would dominate the perturbation of a single particle (our code is deterministic given a specific hardware configuration, i.e.
+running the same ICs on the same CPU count will give identical results).
+Thoughts on this?
+    \end{itemize}
+}

### 2023-02-21 papers/paper-1/diagrams/hm_coresersic.tex
Small updates in preparation for SPZ meeting

@@
     \node[latent, above=of g] (n) {$n$};
+    \node[latent, above=of n] (a) {$a$};
 
     % define layer 3
-    \factor[left=of rb] {rb_f} {below:$\Gamma$} {} {};
-    \node[hyper, left=of rb_f, yshift=-1.2em] (rba) {$r_{\mathrm{b},\alpha}$};
-    \node[hyper, left=of rb_f, yshift=1.2em] (rbb) {$r_{\mathrm{b},\beta}$};
+    \factor[left=of rb] {rb_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of rb_f, yshift=-1.2em] (rbs) {$r_{\mathrm{b},\sigma}$};
+    \node[hyper, left=of rb_f, yshift=1.2em] (rbm) {$r_{\mathrm{b},\mu}$};
 
-    \factor[left=of Re] {Re_f} {below:$\Gamma$} {} {};
-    \node[hyper, left=of Re_f, yshift=-1.2em] (Rea) {$R_{\mathrm{e},\alpha}$};
-    \node[hyper, left=of Re_f, yshift=1.2em] (Reb) {$R_{\mathrm{e},\beta}$};
+    \factor[left=of Re] {Re_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of Re_f, yshift=-1.2em] (Res) {$R_{\mathrm{e},\sigma}$};
+    \node[hyper, left=of Re_f, yshift=1.2em] (Rem) {$R_{\mathrm{e},\mu}$};
 
-    \factor[left=of Sb] {Sb_f} {below:$\Gamma$} {} {};
-    \node[hyper, left=of Sb_f, yshift=-1.2em] (Sba) {$\tilde{\Sigma}_{\mathrm{b},\alpha}$};
-    \node[hyper, left=of Sb_f, yshift=1.2em] (Sbb) {$\tilde{\Sigma}_{\mathrm{b},\beta}$};
+    \factor[left=of Sb] {Sb_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of Sb_f, yshift=-1.2em] (Sbs) {$\tilde{\Sigma}_{\mathrm{b},\sigma}$};
+    \node[hyper, left=of Sb_f, yshift=1.2em] (Sbm) {$\tilde{\Sigma}_{\mathrm{b},\mu}$};
 
-    \factor[left=of g] {g_f} {below:$\Gamma$} {} {};
-    \node[hyper, left=of g_f, yshift=-1.2em] (ga) {$\gamma_\alpha$};
-    \node[hyper, left=of g_f, yshift=1.2em] (gb) {$\gamma_\beta$};
+    \factor[left=of g] {g_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of g_f, yshift=-1.2em] (gs) {$\gamma_\sigma$};
+    \node[hyper, left=of g_f, yshift=1.2em] (gm) {$\gamma_\mu$};
 
-    \factor[left=of n] {n_f} {below:$\Gamma$} {} {};
-    \node[hyper, left=of n_f, yshift=-1.2em] (na) {$n_\alpha$};
-    \node[hyper, left=of n_f, yshift=1.2em] (nb) {$n_\beta$};
+    \factor[left=of n] {n_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of n_f, yshift=-1.2em] (ns) {$n_\sigma$};
+    \node[hyper, left=of n_f, yshift=1.2em] (nm) {$n_\mu$};
+
+    \factor[left=of a] {a_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of a_f, yshift=-1.2em] (as) {$a_\sigma$};
+    \node[hyper, left=of a_f, yshift=1.2em] (am) {$a_\mu$};
 
@@
 
-    \edge {rb, Re, Sb, g, n, R} {S};
+    \edge {rb, Re, Sb, g, n, a, R} {S};
 
-    \factoredge {rba, rbb} {rb_f} {rb};
-    \factoredge {Rea, Reb} {Re_f} {Re};
-    \factoredge {Sba, Sbb} {Sb_f} {Sb};
-    \factoredge {ga, gb} {g_f} {g};
-    \factoredge {na, nb} {n_f} {n};
+    \factoredge {rbs, rbm} {rb_f} {rb};
+    \factoredge {Res, Rem} {Re_f} {Re};
+    \factoredge {Sbs, Sbm} {Sb_f} {Sb};
+    \factoredge {gs, gm} {g_f} {g};
+    \factoredge {ns, nm} {n_f} {n};
+    \factoredge {as, am} {a_f} {a};
 
@@
     \plate {rS} {(R)(S)(sigma)(Shat)} {$\forall R,\sigma_\Sigma^2$};
-    \plate {pars} {(R)(S)(sigma)(rb)(Re)(Sb)(g)(n)} {$\forall$ child};
+    \plate {pars} {(R)(S)(sigma)(rb)(Re)(Sb)(g)(n)(a)} {$\forall$ child};
 

### 2023-02-21 papers/paper-1/sections/results.tex
Small updates in preparation for SPZ meeting

@@
 We use hierarchical modelling to recover the population parameter distributions of the core-S\'ersic model \autoref{eq:coresersic} for each simulation family. 
-We enforce a sharp transition from the core to outer galaxy regions by setting $\alpha$ in \autoref{eq:coresersic} to 10.0. 
 A probabilistic graphical model detailing the connection between hyperparameters, latent parameters, and observables is given in \autoref{fig:graphicalmodel}.
@@
         \hline
-        Hyperparameter & $\alpha$ & $\beta$ \\
+        Hyperparameter & $\mu$ & $\sigma$ \\
         \hline
-        $r_{\mathrm{b}, \alpha}$ & 3.0 & 2.0 \\
-        $r_{\mathrm{b}, \beta}$ & 10.0 & 8.0 \\
-        $R_{\mathrm{e}, \alpha}$ & 30.0 & 2.0 \\
-        $R_{\mathrm{e}, \beta}$ & 40.0 & 2.0 \\
-        $\tilde{\Sigma}_{\mathrm{b}, \alpha}$ & 20.0 & 4.0 \\
-        $\tilde{\Sigma}_{\mathrm{b}, \beta}$ & 20.0 & 2.0 \\
-        $\gamma_\alpha$ & 1.0 & 2.0 \\
-        $\gamma_\beta$ & 4.0 & 2.0 \\
-        $n_\alpha$ & 16.0 & 2.0 \\
-        $n_\beta$ & 4.0 & 2.0 \\
+        $r_{\mathrm{b}, \mu}$ & 0.0 & 2.0 \\
+        $r_{\mathrm{b}, \sigma}$ & 0.0 & 1.0 \\
+        $R_{\mathrm{e}, \mu}$ & 0.0 & 14.0 \\
+        $R_{\mathrm{e}, \sigma}$ & 0.0 & 6.0 \\
+        $\tilde{\Sigma}_{\mathrm{b}, \mu}$ & 0.0 & 20.0 \\
+        $\tilde{\Sigma}_{\mathrm{b}, \sigma}$ & 0.0 & 10.0 \\
+        $\gamma_\mu$ & 0.0 & 0.5 \\
+        $\gamma_\sigma$ & 0.0 & 1.0 \\
+        $n_\mu$ & 0.0 & 8.0 \\
+        $n_\sigma$ & 0.0 & 5.0 \\
+        $a_\mu$ & 0.0 & 20.0 \\
+        $a_\sigma$ & 0.0 & 10.0 \\
         \hline
@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_posterior_pred_log10_proj_density_mean}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_posterior_pred_log10_proj_density_mean}.pdf}
     \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution.
@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_latentqty_compare}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_latentqty_compare}.pdf}
     \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats. Note error bars on black points are smaller than the point itself if not visible}}

### 2023-02-21 papers/paper-1/sections/stats.tex
Small updates in preparation for SPZ meeting

@@
 The modelling is performed using the statistical software \textsc{Stan}\footnote{https://mc-stan.org}, which itself implements Hamiltonian Monte Carlo (HMC).
-
-\subsection{Hamiltonian Monte Carlo}
-An excellent overview of HMC may be found in \citet{betancourt17}, with a brief description provided here. \drafting{Mention NUTS here somewhere?}
-
-In numerically sampling an arbitrary target distribution, we are required to sample from the \textit{typical set}, or those regions of the target distribution which provide the dominant contribution to an expectation value.
-Simplistically, this may be achieved using any Markov generating process, such as the widely-used Metropolis-Hastings (MH) algorithm, albeit at the expense of inefficient and biased sampling. 
-The MH algorithm generates from a starting position $q$ a proposal for the next position $q'$, with proposals accepted with some probability $Q(q'|q)$, where $Q$ is often chosen to be the normal distribution.
-Sampling with increased efficiency can be achieved by using information about the geometry of the typical set of the target distribution through a vector field aligned with the typical set, which may be estimated using the gradient of the target distribution.
-To align the gradient vector field with the typical set vector field, we can introduce auxillary momenta variables $p$, promoting our $D$-dimensional target distribution to a projection of a $2D$-dimensional canonical distribution in some phase space:
-\begin{equation}\label{eq:hmc}
-    P(q,p) = P(p|q)P(q) = e^{-H(q,p)}
-\end{equation}
-where $H$ is an invariant Hamiltonian function. 
-The desired vector field may then be generated from the Hamiltonian $H$ using Hamilton's equations. 
-This formulation also lends itself to improved sampling efficiency by decomposing the Hamiltonian into a \textit{kinetic energy} term and a \textit{potential energy} term, with the latter equivalent to the logarithm of the target distribution.
-\begin{align}
-    H(q,p) &= -\log P(q,p) \\
-    &= -\log \left( P(p|q)P(q) \right) \\
-    &= -\log P(p|q) -\log P(q)
-\end{align}
-As the Hamiltonian is phase space conserving, the Hamilton equations may be accurately integrated using a volume-preserving integration scheme, such as symplectic integrators. 
-Of these integrators, the leapfrog scheme offers both a simple implementation and high accuracy over long integration times.
 
@@
 
-To answer these questions, we define a test statistic $T$ from which we determine a Bayes-equivalent of the frequentist $p$-value \citep{gelman15}.
-\begin{align}
-    p_\mathrm{Bayes} &= \mathrm{Pr}\left( T(y^\mathrm{rep}, \theta) \geq T(y, \theta)|y \right) \\
-    &= \iint I_{T(y^\mathrm{rep}, \theta) \geq T(y, \theta)|y} p\left( y^\mathrm{rep}|\theta\right) p\left( \theta|y\right) \dd{y^\mathrm{rep}} \dd{\theta}
-\end{align}
-Here, $I$ is the indicator function.
+%To answer these questions, we define a test statistic $T$ from which we determine a Bayes-equivalent of the frequentist $p$-value \citep{gelman15}.
+%\begin{align}
+%    p_\mathrm{Bayes} &= \mathrm{Pr}\left( T(y^\mathrm{rep}, \theta) \geq T(y, \theta)|y \right) \\
+%    &= \iint I_{T(y^\mathrm{rep}, \theta) \geq T(y, \theta)|y} p\left( y^\mathrm{rep}|\theta\right) p\left( \theta|y\right) \dd{y^\mathrm{rep}} \dd{\theta}
+%\end{align}
+%Here, $I$ is the indicator function.
 

### 2023-03-01 papers/paper-1/sections/appendix.tex
Small updates in preparation for SPZ meeting
Start adding small section on hardening rate

@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_gqs_prior}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_gqs_prior}.pdf}
     \caption{\drafting{Prior predictive distributions of latent parameters. Note the mode of the distributions roughly coincide with what one would expect from the literature.}}
@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_prior_pred_log10_proj_density_mean}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_prior_pred_log10_proj_density_mean}.pdf}
     \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit.
-Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
+Darker shaded regions correspond to greater Highest Density Intervals (HDIs).
 The observed data is overlaid.
@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_pair_0}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_pair_0}.pdf}
     \caption{\drafting{Corner plot of hierarchical model hyperparameters.
@@
 Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well.
-The contours correspond to the 50\%, 90\%, 95\%, and 99\% BCIs.
+The contours correspond to the 50\%, 90\%, 95\%, and 99\% HDIs.
 }}
@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_pair_1}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_pair_1}.pdf}
     \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
@@
 \begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001_pair_2}.pdf}
+    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_pair_2}.pdf}
     \caption{\drafting{Same as \autoref{fig:corner_params_1}}}

### 2023-03-01 papers/paper-1/sections/results.tex
Start adding small section on hardening rate

@@
 
-\drafting{ENSURE CONSISTENCY WITH BCI AND HDI. SAME THING, JUST DIFFERENT NAME}
-
 \subsection{Projected Mass Density Profiles}
@@
     \centering
-    \caption{The parameters for each hyperparameter distribution for the core-S\'ersic model, where each distribution is a gamma distribution.}
+    \caption{The parameters for each hyperparameter distribution for the core-S\'ersic model, where each distribution is a normal distribution.}
     \label{tab:cshypers}
@@
         $R_{\mathrm{e}, \sigma}$ & 0.0 & 6.0 \\
-        $\tilde{\Sigma}_{\mathrm{b}, \mu}$ & 0.0 & 20.0 \\
-        $\tilde{\Sigma}_{\mathrm{b}, \sigma}$ & 0.0 & 10.0 \\
+        $\log_{10}\Sigma_{\mathrm{b}, \mu}$ & 0.0 & 20.0 \\
+        $\log_{10}\Sigma_{\mathrm{b}, \sigma}$ & 0.0 & 10.0 \\
         $\gamma_\mu$ & 0.0 & 0.5 \\
@@
 \subsubsection{Observable Quantities}
-The hierarchical model takes three observable quantities, denoted by hat-variables in red circles in \autoref{fig:graphicalmodel}. 
+The hierarchical model takes three observable quantities, denoted by hat-variables in circles in \autoref{fig:graphicalmodel}. 
 These variables are the radius $\hat{R}$, and the mean projected mass density $\hat{\Sigma}$, and the standard deviation of the projected mass density $\hat{\sigma}_\Sigma$ measured in the radial bins centred on $\hat{R}$.
@@
 \subsubsection{Latent Parameters for Observables}
-With the omission of the parameter $\alpha$ in the modelling process, there are five latent parameters ($r_\mathrm{b}$, $R_\mathrm{e}$, $\tilde{\Sigma}_\mathrm{b}$, $\gamma$, and $n$, the set of which we denote as $\kappa$) that describe each \child{} simulation in a family.
-Here, we define $\tilde{\Sigma}_\mathrm{b} \equiv \Sigma_\mathrm{b}/10^{10}$, where $\Sigma_\mathrm{b}$ is defined in \autoref{eq:coresersic2}.
-These latent parameters, depicted as non-hatted variables in blue circles in \autoref{fig:graphicalmodel}, are sampled to obtain a latent estimate on the projected mass density, $\Sigma$.
-As each latent parameter in the set $\kappa$ is strictly non-negative, we use as prior distributions for each latent parameter a gamma distribution with hyperparameters $(\kappa_\alpha, \kappa_\beta)$, which are themselves drawn from a gamma distribution.
+There are six latent parameters ($r_\mathrm{b}$, $R_\mathrm{e}$, $\log_{10}\Sigma_\mathrm{b}$, $\gamma$, $n$, and $\alpha$, the set of which we denote as $\kappa$) that describe each \child{} simulation in a family.
+These latent parameters, depicted as non-hatted variables in circles in \autoref{fig:graphicalmodel}, are sampled to obtain a latent estimate on the projected mass density, $\Sigma$.
+The latent parameters in the set $\kappa$ are strictly non-negative, with the exception of $\log_{10}\Sigma_\mathrm{b}$. 
+For these five positive-constrained latent parameters, we model the latent distribution as a half normal distribution. 
+Each location hyperparameter for each of these latent parameters is also drawn from a truncated normal distribution, as a negative expectation value may be excluded on physical grounds.
+For $\log_{10}\Sigma_\mathrm{b}$, we assume a normal distribution, with a location hyperparameter centred at $\log_{10}\Sigma_{\mathrm{b}, \mu}=10$. 
+For all scale hyperparameters, we use a half normal distribution (as the scale of a distribution must of course always be positive).
 To ensure our hyperparameter prior distributions are weakly-informative, the hyperparameters are drawn from the distributions in \autoref{tab:cshypers} for all simulation families. 
 The sampled hyperparameter distributions are shown as corner plots in Figures \ref{fig:corner_params_1}-\ref{fig:corner_params_3}.
+\drafting{Corner plots maybe not so informative here, do we really expect correlations?
+Maybe 1D distributions better for the hyperpriors.}
 
@@
 Critically, both the range of values spanned by the prior distributions and the general shape of the distribution suggests that the data is well represented by the model and the prior distributions on the hyper and latent parameters.
+\drafting{Not yet, but will be.
+Some issue with the sampler.}
 
-As likelihood function, we use the lognormal distribution $\mathrm{Lognorm}(\Sigma, \hat{\sigma}_\Sigma^2)$ as it a) has support on $[0,\infty)$, and b) produces normally-distributed variates in logarithm space, which is representative of the distribution of $\hat{\Sigma}$ for a given radial bin when we backwards-model and na\"ively bin our observations of $\hat{\Sigma}$ for that radial bin.
+For each radial bin, we backwards-model and na\"ively bin our thirty observations of $\log_{10}\hat{\Sigma}$ for that radial bin to assess the likely distribution.
+We find that the different projections of the mass density profile in a given radial bin are well-described by a normal distribution in logarithm space. 
+Thus, as likelihood function, we use the normal distribution $\mathrm{Lognorm}(\Sigma, \hat{\sigma}_\Sigma^2)$.
 
@@
     \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution.
-Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
-The observed data is overlaid, and seen to lie well within the 50\% BCI.
+Darker shaded regions correspond to greater Highest Density Intervals (HDIs).
+The observed data is overlaid, and seen to lie well within the 50\% HDIs.
 }}
@@
 We first inspect how the posterior distribution appears visually by computing posterior predictive checks, an example of which is shown in \autoref{fig:postpred} for the family H-1.00. 
-We find that the data is well described by the 50\% BCI \drafting{introduce what BCI is} of the posterior distribution, as desired.
+We find that the data is well described by the 50\% HDI \drafting{introduce what HDI is} of the posterior distribution, as desired.
 
-As part of the hierarchical modelling sampling process, we obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined. 
+As part of the hierarchical modelling sampling process, we can obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined by marginalising out the other parameters.
 We show an example of these latent parameter distributions in \autoref{fig:cs_compare_naive}, where the PDF of the latent parameter distributions is given in blue. 
-For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive Gaussian statistics.
+For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive $\chi^2$ minimisation techniques.
 To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the median and IQR of the sample. 
 The median value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
-The Gaussian statistics fail to capture the correct estimates for the latent parameter estimates, with some parameters (e.g.
+The $\chi^2$ minimisation fails to capture the correct estimates for the latent parameter estimates, with some parameters (e.g.
 $n$) converging to the boundary value we limit the Least Squares method to (for $n$, the bounds are $[0,20]$).
@@
 
+\subsection{Hardening Rate}
+After considering the robustness of the projected density profile to SMBH Brownian motion, we turn our attention to a similar analysis of the hardening rate of the SMBH binary. 
+The hardening rate $H$ in \autoref{eq:quinlan_H} connects the evolution of the inverse semimajor axis of the binary system to the surrounding stellar mass distribution. 
+To extract the data for our hierarchical model for hardening rate, we take for each \child{} simulation the semimajor axis values between $a=a_\mathrm{h}$ and $a/\mathrm{pc}=10$.
+We then fit the differential equation give in \autoref{eq:quinlan_H} to obtain a parameter estimate of the combined quantity $HG\rho/\sigma$.
+We fit for the combined property in preference to just $H$ as the temporal-resolution of the semimajor axis $a$ is \drafting{check} $10^4$ times more frequent than the temporal resolution of the snapshots, requiring linear (or higher order) interpolation to be done between snapshots to obtain a continuous function of $G\rho/\sigma$ that can be mapped to the temporal resolution of $a$.
+To avoid these assumptions, we estimate the combined quantity $HG\rho/\sigma$.
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{quinlan_posterior_pred_inv_a.png}
+    \caption{\drafting{Example plot. This is an old one and needs to be updated, but gives us the idea of what we can talk about}}
+    \label{fig:quinlan_H}
+\end{figure}
+
+\subsection{Binary Properties}
 \begin{figure}
@@
 \end{figure}
-
-\subsection{Binary Properties}
-After considering the robustness of the projected density profile to SMBH Brownian motion, we turn our attention to a similar analysis of the orbital parameters of the SMBH binary, namely the semimajor axis $a$ and the eccentricity $e$.
+\drafting{INtro here} orbital parameters of the SMBH binary, namely the semimajor axis $a$ and the eccentricity $e$. 
 It is instructive to first consider the raw data of a particular simulation family, shown in \autoref{fig:ketju_scatter} for the case of H-1.00. 

### 2023-03-01 papers/paper-1/sections/stats.tex
Start adding small section on hardening rate

@@
 For example, a normalising radius $a$ cannot take a value less than 0, so a normal distribution, which has an unbounded domain, is an inappropriate choice for the prior distribution of this variable.
-Choosing a log-normal or gamma distribution with support on $[0,\infty)$ would represent a physically-motivated choice for the prior distribution here.
+Choosing a truncated normal or gamma distribution with support on $[0,\infty)$ would represent a physically-motivated choice for the prior distribution here.
 

### 2023-03-11 papers/paper-1/sections/appendix.tex
Start to edit paper 1 to reflect project direction change

--- 2022-10-07 papers/paper-1/sections/appendix.tex
@@
 Here we show the prior predictive check for the projected mass density of the H-1.00 system, as well the corner plots of the latent parameter hyperdistributions.
-
-\begin{figure*}
-    \includegraphics[width=\textwidth]{{graham_density-A-C-3.0-0.05_prior_pred_log10_proj_density_mean}.png}
-    \caption{\drafting{Prior predictive estimate for the core-S\'ersic fit.
-Darker shaded regions correspond to higher Bayesian Credible Intervals (BCIs).
-The observed data is overlaid.
-From the prior predictive estimate, it is seen that the model and the assumed parameter hyperdistributions are able to describe the observed data, in that the observed data lies within the range allowed by the prior predictive estimate.}}
-    \label{fig:priorpred}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_0}.png}
-    \caption{\drafting{Corner plot of hierarchical model hyperparameters.
-Each of the hyperdistributions are modelled as Gamma distributions with shape parameters $\alpha$ and $\beta$.
-For example, the hyperdistribution on the core radius $r_\mathrm{b}$ is modelled as $r_\mathrm{b} \sim \text{Gamma}(r_{\mathrm{b},\alpha}, r_{\mathrm{b},\beta})$.
-Critically, the hyperdistributions are unimodal with highest-density regions located some distance from the distribution edges, indicating the HMC sampling has converged well.
-}}
-    \label{fig:corner_params_1}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_1}.png}
-    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
-    \label{fig:corner_params_2}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics[width=0.7\textwidth]{{graham_density-A-C-3.0-0.05_pair_2}.png}
-    \caption{\drafting{Same as \autoref{fig:corner_params_1}}}
-    \label{fig:corner_params_3}
-\end{figure*}

### 2023-03-11 papers/paper-1/sections/numerical.tex
Start adding small section on hardening rate
Start to edit paper 1 to reflect project direction change

--- 2022-09-21 papers/paper-1/sections/numerical.tex
@@
 
-\drafting{Describe the numerical set up, code, etc}
+\subsection{Initial Conditions}
+
+\begin{table*}
+    \centering
+    \caption{Parameters of Hernquist sphere galaxy models used in determining the binary hardening convergence.}
+    \label{tab:hernquist_pars}
+    \begin{tabular}{cccccccc}
+        \hline
+        Name & $m_\star / (10^5 \Msun)$ & $N_\star/10^6$ & $\varepsilon_\star$/pc & $m_\mathrm{DM} / (10^7 \Msun)$ & $N_\mathrm{DM}/10^6$ & $\varepsilon_\mathrm{DM}$/pc & $M_\bullet/m_\star$\\
+        \hline
+        H-1.000 & 0.5 & 5.02 & 5.0 & 1.0 & 4.54 & 300.0 & $6.0\times 10^4$ \\
+        H-0.500 & 1.0 & 2.51 & 5.0 & 2.0 & 2.27 & 300.0 & $3.0\times 10^4$ \\
+        H-0.250 & 2.0 & 1.25 & 5.0 & 4.0 & 1.13 & 300.0 & $1.5\times 10^4$ \\
+        H-0.100 & 5.0 & 0.50 & 5.0 & 10.0 & 0.45 & 300.0 & $6.0\times 10^3$ \\
+        H-0.050 & 10.0 & 0.25 & 10.0 & 20.0 & 0.23 & 600.0 & $3.0\times 10^3$ \\
+        H-0.010 & 50.0 & 0.05 & 10.0 & 100.0 & 0.04 & 600.0 & $6.0\times 10^2$ \\
+        H-0.005 & 100.0 & 0.025 & 20.0 & 200.0 & 0.02 & 1000.0 & $3.0\times 10^1$ \\
+        H-0.001 & 500.0 & 0.005 & 20.0 & 1000.0 & 0.004 & 1000.0 & $6.0\times 10^1$ \\
+        \hline
+    \end{tabular}
+\end{table*}
+
+We simulate the merger of two identical galaxies. 
+Each galaxy is modelled as a spherically-symmetric multicomponent system: a stellar component embedded within a DM halo, and a single SMBH at the centre of the galaxy. 
+The total stellar mass is set to $M_\star=2.5\times10^{11}\,\Msun$, the total DM mass is set to $M_\mathrm{DM} \sim 4.9\times 10^{13}\,\Msun$, and the SMBH mass is set to $M_\bullet \sim 3.0\times10^9\,\Msun$.
+The total DM mass and BH mass are determined as a function of total stellar mass from the relations of \citet{behroozi19} and \citet{sahu19}, respectively. 
+Both the stellar and DM components have a density profile given by an ergodic Hernquist sphere parametrised by scale radius $a$:
+\begin{equation}\label{eq:hernquist}
+    \rho(r) = \frac{2M}{4\pi} \frac{a}{r(r+a)^3}.
+\end{equation}
+We use a value of $a_\star=3.0\,\mathrm{kpc}$ for the stellar component, and a value of $a_\mathrm{DM}=300.0\,\mathrm{kpc}$ for the DM component. 
+We Monte Carlo sample the phase space distribution function
+\begin{equation}\label{eq:eddington}
+    f(\mathcal{E}) = \frac{1}{\sqrt{8}\pi^2} \left[ \int_0^\mathcal{E} \dv{\Psi}{\sqrt{\mathcal{E}-\Psi}} \dv{^2\nu_\mathcal{E}}{\Psi^2} + \frac{1}{\sqrt{\mathcal{E}}} \left(\dv{\nu_\mathcal{E}}{\Psi}\right)_{\Psi=0}\right],
+\end{equation}
+to generate position and velocity coordinates for all stellar and DM particles.
+As mentioned, the SMBH is initialised at rest at the origin.
+Additionally, we give the SMBH a randomly-drawn value of the dimensionless spin parameter $\chi$ following \drafting{one of the Zlochower papers}, with shape parameters $\alpha=10.5868, \beta=4.66884$.
+This distribution has a mode at $\chi \sim 0.8$. 
+We test eight mass resolutions between the SMBH mass and the stellar particle mass, with four realisations generated for each mass resolution by independently Monte Carlo sampling the phase space distribution function.
+
+We simulate the equal-mass merger between like-mass resolution systems for each mass resolution listed in \autoref{tab:hernquist_pars} for all combination of galaxy pairs (i.e., ten mergers per mass resolution).
+The progenitor galaxies in each merger are set to an initial separation of $20\,\mathrm{kpc}$ on quasi-Keplerian orbit, assuming each galaxy is a point-mass containing half its total mass. 
+The initial orbital eccentricity is chosen to be almost radial ($e_0 = 0.99$), and the first pericentre distance $r_\mathrm{peri}$ is set to $\sim 8\,\mathrm{kpc}$.
+
+\subsection{Simulations}
+
+The progenitors are evolved with \ketju{}.
+Within a relatively short time frame, the SMBH binary hardens predominantly through dynamical friction and three-body stellar interactions. 
+We evolve the binary to the time $t_{a_h}=t(a_h)$, and continue to at least when $t_\text{safe}=t(a=10\,\text{pc})$. 
+As we do not require the gravitational wave dominated regime to be reached for our analysis, we may terminate the runs before the SMBH binary merges.

### 2023-03-11 papers/paper-1/sections/results.tex
Start to edit paper 1 to reflect project direction change

@@
     \begin{enumerate}
+        \item compare to \citet{nasim20}
         \item stability of hardening rate -> semimajor axis can be robustly predicted
         \item wide variance of eccentricity -> N-body models offer no quantitative prediction on likelihood of merger, can only comment on what merger looks like \textit{IF} it happens
-        \item variance in observables: trends between observable properties and what observers can actually make use of (don't have this yet), e.g. signatures of SMBH scouring in velocity profile?
+        \item sample posterior to determine merger timescale
     \end{enumerate}
@@
 
-\subsection{Projected Mass Density Profiles}
-We use hierarchical modelling to recover the population parameter distributions of the core-S\'ersic model \autoref{eq:coresersic} for each simulation family. 
-A probabilistic graphical model detailing the connection between hyperparameters, latent parameters, and observables is given in \autoref{fig:graphicalmodel}.
+\subsection{Comparison to Previous Work}
+\begin{enumerate}
+    \item \citet{nasim20} show convergence of eccentricity to follow $\sigma_\mathrm{ecc} \propto 1/\sqrt{N_\star}$
+    \item Perform same analysis
+    \item Results disagree $\rightarrow$ because of highly radial orbit?? 
+\end{enumerate}
 
-\begin{table}
-    \centering
-    \caption{The parameters for each hyperparameter distribution for the core-S\'ersic model, where each distribution is a normal distribution.}
-    \label{tab:cshypers}
-    \begin{tabular}{ccc}
-        \hline
-        Hyperparameter & $\mu$ & $\sigma$ \\
-        \hline
-        $r_{\mathrm{b}, \mu}$ & 0.0 & 2.0 \\
-        $r_{\mathrm{b}, \sigma}$ & 0.0 & 1.0 \\
-        $R_{\mathrm{e}, \mu}$ & 0.0 & 14.0 \\
-        $R_{\mathrm{e}, \sigma}$ & 0.0 & 6.0 \\
-        $\log_{10}\Sigma_{\mathrm{b}, \mu}$ & 0.0 & 20.0 \\
-        $\log_{10}\Sigma_{\mathrm{b}, \sigma}$ & 0.0 & 10.0 \\
-        $\gamma_\mu$ & 0.0 & 0.5 \\
-        $\gamma_\sigma$ & 0.0 & 1.0 \\
-        $n_\mu$ & 0.0 & 8.0 \\
-        $n_\sigma$ & 0.0 & 5.0 \\
-        $a_\mu$ & 0.0 & 20.0 \\
-        $a_\sigma$ & 0.0 & 10.0 \\
-        \hline
-    \end{tabular}
-\end{table}
+\subsection{Hardening Rate}
+\begin{enumerate}
+    \item Introduce HM for hardening rate
+    \item Corner plots of latent parameters, forward folded PPC
+    \item Discussion uncertainty on $HG\rho/\sigma$
+\end{enumerate}
 
-\begin{figure}
-    \includegraphics{hm_coresersic.pdf}
-    \caption{Probabilistic graphical model for the likelihood of the core-S\'ersic model.
-Probabilistic parameters and observable quantities (hatted variables) are shown as circles, whereas deterministic quantities are represented by diamonds.
-Hyperparameters are represented by purple circles, latent parameters in blue, and observable quantities as red circles.
-The nodes indicate the distribution used in constructing the prior distributions (for latent parameters) and likelihood function (for $\hat{\Sigma}$).
-The radial box represents observed and latent variables for each radial bin for a \child{}, and the child box represents latent parameters for each \child{} simulation in a family.
-}
-    \label{fig:graphicalmodel}
-\end{figure}
-
-\subsubsection{Observable Quantities}
-The hierarchical model takes three observable quantities, denoted by hat-variables in circles in \autoref{fig:graphicalmodel}. 
-These variables are the radius $\hat{R}$, and the mean projected mass density $\hat{\Sigma}$, and the standard deviation of the projected mass density $\hat{\sigma}_\Sigma$ measured in the radial bins centred on $\hat{R}$.
-For each \child{} simulation, we choose to `observe' the first snapshot where the binary semimajor axis is \textit{at most} $10\,\mathrm{pc}$. 
-In the event that there is no snapshot satisfying the above requirement (due to, for example, rapid merging of the SMBH binary in a time period shorter than the snapshot output frequency), the \child{} simulation is not included in the analysis.
-We enforce a minimum of six \child{} simulations per family for that family to be analysed with our hierarchical modelling technique. 
-For those \child{} simulations which pass our selection criteria, we take 30 random projections of the merger remnant, and determine the projected mass density in 50 radial bins evenly-spaced in logarithm space, with a minimum value of $0.2\,\mathrm{kpc}$ and a maximum value of $20.0\,\mathrm{kpc}$.
-This allows us to build up a distribution of projected mass densities per radial bin, from which $\hat{\Sigma}$ and $\hat{\sigma}_\Sigma$ are determined.
-
-\subsubsection{Latent Parameters for Observables}
-There are six latent parameters ($r_\mathrm{b}$, $R_\mathrm{e}$, $\log_{10}\Sigma_\mathrm{b}$, $\gamma$, $n$, and $\alpha$, the set of which we denote as $\kappa$) that describe each \child{} simulation in a family.
-These latent parameters, depicted as non-hatted variables in circles in \autoref{fig:graphicalmodel}, are sampled to obtain a latent estimate on the projected mass density, $\Sigma$.
-The latent parameters in the set $\kappa$ are strictly non-negative, with the exception of $\log_{10}\Sigma_\mathrm{b}$. 
+We enforce a minimum of six simulations per mass resolution for that mass resolution to be analysed with our hierarchical modelling technique. 
 For these five positive-constrained latent parameters, we model the latent distribution as a half normal distribution. 
@@
 To ensure our hyperparameter prior distributions are weakly-informative, the hyperparameters are drawn from the distributions in \autoref{tab:cshypers} for all simulation families. 
-The sampled hyperparameter distributions are shown as corner plots in Figures \ref{fig:corner_params_1}-\ref{fig:corner_params_3}.
-\drafting{Corner plots maybe not so informative here, do we really expect correlations?
-Maybe 1D distributions better for the hyperpriors.}
+The sampled hyperparameter distributions are shown as corner plots in Figures
 
@@
 
-For each radial bin, we backwards-model and na\"ively bin our thirty observations of $\log_{10}\hat{\Sigma}$ for that radial bin to assess the likely distribution.
-We find that the different projections of the mass density profile in a given radial bin are well-described by a normal distribution in logarithm space. 
-Thus, as likelihood function, we use the normal distribution $\mathrm{Lognorm}(\Sigma, \hat{\sigma}_\Sigma^2)$.
-
-\drafting{Write explicitly the posterior distribution we are sampling from.}
-
-\begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_posterior_pred_log10_proj_density_mean}.pdf}
-    \caption{\drafting{Posterior predictive estimate for the core-S\'ersic fit for highest (fiducial) mass resolution.
-Darker shaded regions correspond to greater Highest Density Intervals (HDIs).
-The observed data is overlaid, and seen to lie well within the 50\% HDIs.
-}}
-    \label{fig:postpred}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics{{graham_density-H-H-3.0-0.001-hierarchy_latentqty_compare}.pdf}
-    \caption{\drafting{Latent params, and compare to naive estimates using Gaussian stats. Note error bars on black points are smaller than the point itself if not visible}}
-    \label{fig:cs_compare_naive}
-\end{figure*}
-
-\subsection{Latent Parameter Posterior Distributions}
 We compute the posterior distribution using \textsc{Stan}, ensuring that the sampler exits without encountering any divergences, and that all \drafting{R-hat (is a summary stat, but how to draw?
@@
 As part of the hierarchical modelling sampling process, we can obtain in addition to the posterior distribution on $\Sigma(R)$ the latent parameter distributions from which $\Sigma(R)$ is determined by marginalising out the other parameters.
-We show an example of these latent parameter distributions in \autoref{fig:cs_compare_naive}, where the PDF of the latent parameter distributions is given in blue. 
-For comparison, we also show the estimated uncertainty on each latent parameter determined using na\"ive $\chi^2$ minimisation techniques.
-To do this, we fit each density profile in logarithm space using a Least-Squares method, and then determine the median and IQR of the sample. 
-The median value and spread for each latent parameter using this na\"ive method is shown above the corresponding posterior distribution estimate in \autoref{fig:cs_compare_naive}.
-The $\chi^2$ minimisation fails to capture the correct estimates for the latent parameter estimates, with some parameters (e.g.
-$n$) converging to the boundary value we limit the Least Squares method to (for $n$, the bounds are $[0,20]$).
-We try successively relaxing and tightening the boundary values of each parameter to obtain a `physically-realistic' value for each optimal latent parameter value, with no success. 
-Henceforth, we abandon any attempt to use a Least Squares minimisation to recover the optimal latent parameter values, returning instead to our Bayesian approach.
 
-We show the latent parameter distributions for each mass resolution family in \autoref{fig:cs_mass_compare}, and compare the median and IQRs \drafting{ensure this acronym explained somewhere} in \autoref{fig:cs_mass_compare_2}.
-We find first that the spread in the latent parameters (as quantified through their IQR) does not dramatically differ between mass resolutions.
-Secondly, we observe that not all latent parameters in the core-S\'ersic model display asymptotic convergence, and that in some cases, such as for the effective radius $R_\mathrm{e}$, all tested mass resolutions produce estimates that are consistent with each other.
-For other latent parameters, such as the core radius $r_\mathrm{b}$, there is a clear convergence of the median value with increasing mass resolution.
-\drafting{More particles --> core better resolved.
-Discuss here or in discussion?}
-
-Whilst we are unable to attribute the error in \autoref{fig:cs_mass_compare} to a particular phenomenon, we may at least rule out that the spread is due to post-processing analysis introduced from differing observation projections of the merger remnant.
-We re-determine the mean and standard deviation of the logarithm of the projected density profile using a subset of the total 30 projections we have available to us for the H-1.00 merger family.
-Repeating this data-reduction process four times, we then rerun the HMC algorithm for each subsample.
-We compare the latent parameter distributions for each subsample to the latent parameter distributions for the H-1.00 family determined using the full projection sample, and find:
-\begin{enumerate}
-    \item The error in the latent parameters is the consistent when using a subsample of projections, or the full sample, implying
-    \item The error in the latent parameters is independent of the projection direction.
-\end{enumerate}
-Consequently, we may attribute the uncertainty in the latent parameter distributions to one of:
-\begin{itemize}
-    \item Intrinsic scatter inherent to the physical system, or 
-    \item Numerical noise arising from a source independent of mass resolution, or 
-    \item A combination of the two.
-\end{itemize}
-
-\begin{figure*}
-    \includegraphics{compare_HH_gqs_res.pdf}
-    \caption{\drafting{Comparison of core sersic fit for different mass resolutions.
-Note the scatter does not seem to decrease with increasing mass resolution, however the distribution modes seem to converge towards something.}}
-    \label{fig:cs_mass_compare}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics{compare_HH_gqs_res_2d.pdf}
-    \caption{\drafting{Same as \autoref{fig:cs_mass_compare}, but plotted as a function of mass resolution.
-Points correspond to the median estimate for the latent parameters, and error bars to the IQR.}}
-    \label{fig:cs_mass_compare_2}
-\end{figure*}
-
-\subsection{Hardening Rate}
 After considering the robustness of the projected density profile to SMBH Brownian motion, we turn our attention to a similar analysis of the hardening rate of the SMBH binary. 
 The hardening rate $H$ in \autoref{eq:quinlan_H} connects the evolution of the inverse semimajor axis of the binary system to the surrounding stellar mass distribution. 
-To extract the data for our hierarchical model for hardening rate, we take for each \child{} simulation the semimajor axis values between $a=a_\mathrm{h}$ and $a/\mathrm{pc}=10$.
+To extract the data for our hierarchical model for hardening rate, we take for each simulation the semimajor axis values between $a=a_\mathrm{h}$ and $a/\mathrm{pc}=10$.
 We then fit the differential equation give in \autoref{eq:quinlan_H} to obtain a parameter estimate of the combined quantity $HG\rho/\sigma$.
@@
 \subsection{Binary Properties}
+\begin{enumerate}
+    \item introduce HM for binary quantities 
+    \item corner plots $a$, $e$, discuss generacy
+    \item forward folded PPC for $l$
+    \item demonstrate $e$ cannot be robustly predicted (e.g. multimodal marginal distribution)
+\end{enumerate}
+
 \begin{figure}
@@
 As discussed in \drafting{section}, the efficiency of GW as a hardening mechanism is directly related to the binary eccentricity, and is highly non-linear.
-This is observed in those \child{} realisations that merged, and the timescale within which they merged.
+
 As seen in \autoref{fig:ketju_scatter}, a higher eccentricity leads to more rapid BH binary mergers as opposed to a lower eccentricity. 
@@
 
-\drafting{
-    QUESTIONS FOR SIMON:\\
-    \begin{itemize}
-        \item Hierarchical models rely on the ability to model a set of observations as belonging to a \textit{population}.
-After the chaos talk, can a collection of BH-perturbed systems be described still as belonging to a population?
-I think yes (stellar mass distribution unchanged), but would be keen to hear thoughts.
-        \item In this paper we're perturbing the phase space coordinates of the BHs as opposed to a single field particle (we did this as another sensitivity test).
-Let's consider the case of perturbing a field particle.
-I find it still somehow curious that such a small displacement (a fraction of a softening length) still propagates the perturbation to such large scales, considering we are using a tree code for the particles not-near the BHs.
-I would have (na\"ively) thought that round off error from the tree algorithm would dominate the perturbation of a single particle (our code is deterministic given a specific hardware configuration, i.e.
-running the same ICs on the same CPU count will give identical results).
-Thoughts on this?
-    \end{itemize}
-}
+\subsection{Merger Timescale}
+\begin{enumerate}
+    \item Sample from marginal distributions $HG\rho/\sigma$, $a_\mathrm{h}$, $e_\mathrm{h}$
+    \item Plug in to Quinlan-Peter analytical formula (write this explicitly)
+    \item Posterior estimate on $t_\mathrm{merge}$
+    \item discuss uncertainty
+\end{enumerate}

### 2023-04-19 papers/paper-eccentricity/paper-eccentricity.tex
Set up directory for eccentricity paper

--- 2022-10-11 papers/paper-1/paper-1.tex
@@
 \newcommand{\gadget}{\textsc{Gadget-3}}
-\newcommand{\parent}{\textit{parent}}
-\newcommand{\child}{\textit{child}}
 
@@
 % Keep the title short and informative.
-\title[BAGGinS-I]{Bayesian Analysis of Gas-free Galaxies in Simulations I: Reviving Stochasticity}
+\title[Reviving Stochasticity]{Reviving stochasticity: eccentricity convergence depends on black hole binary orbit}
 
@@
 \begin{abstract}
-We present the Bayesian Analysis of Gas-free Galaxies in Simulations (BAGGinS) project, show that stochasticity is ubiquitous with collisionless merger sims.
+We present some stuff about eccentricity convergence in Nbody sims
 \end{abstract}
@@
 \input{./sections/numerical}
-
-\input{./sections/stats}
 
@@
 
-\input{sections/appendix}
+%\input{sections/appendix}
 

### 2023-04-19 papers/paper-eccentricity/sections/numerical.tex
Set up directory for eccentricity paper

\section{Numerical Simulations}

\begin{itemize}
    \item describe toy model
    \item describe galaxy ICs: 0.5 Dehnen spheres
    \item describe merger orbit: low and high e
\end{itemize}

### 2023-04-19 papers/paper-eccentricity/sections/results.tex
Set up directory for eccentricity paper

\section{Results}

\subsection{Eccentricity scatter in a toy model}
\begin{itemize}
    \item Show that brownian motion affects impact parameter of the final passage before BHs bound --> eccentricity set
\end{itemize}

\subsection{Eccentricity scatter in low-e orbits}
\begin{itemize}
    \item Show that impact parameters not sensitive
\end{itemize}

\subsection{Eccentricity scatter in high-e orbits}
\begin{itemize}
    \item show that impact parameters sensitive
\end{itemize}

### 2023-04-21 papers/paper-eccentricity/sections/results.tex
Add some ideas and drafting to the paper

@@
 \end{itemize}
+
+\drafting{Nasim result takes the mean of the eccentricity over some time period, and then determines the standard deviation across the means of different realisations of a merger system.
+We probably need to show this too for `consistency', however we could also introduce a simple hierarchical model here that allows us to do something similar more robustly.}
 
@@
 \end{itemize}
+
+\begin{figure*}
+    \includegraphics{orbit_pars}
+    \caption{\drafting{Orbital parameters of the different merger systems, colour coded according to the initial orbital eccentricity}}
+    \label{fig:orbitpars}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics{nasim_style_plot}
+    \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
+    \label{fig:nasim}
+\end{figure*}
+
+\begin{figure*}
+    \includegraphics{hyperparameter}
+    \caption{\drafting{Hyperparameters of a simple hierarchical model of the binary angular momentum and energy at time when binary is hard.
+To show convergence, we could compare the the distributions of $\sigma_{e_h}$ directly between resolutions}}
+    \label{fig:hyperparameter}
+\end{figure*}

### 2023-04-28 papers/paper-eccentricity/sections/results.tex
Updates to paper ideas
Start writing about the toy model

@@
 
-\subsection{Eccentricity scatter in a toy model}
 \begin{itemize}
@@
     \includegraphics{hyperparameter}
-    \caption{\drafting{Hyperparameters of a simple hierarchical model of the binary angular momentum and energy at time when binary is hard.
+    \caption{\drafting{Hyperparameters of a hierarchical model of the binary hardening from a time when the binary is hard.
 To show convergence, we could compare the the distributions of $\sigma_{e_h}$ directly between resolutions}}
@@
 \end{figure*}
+
+\subsection{Explaining the eccentricity scatter with a toy model}
+
+We can gain insight into the large binary eccentricity scatter observed in the
+runs with high galactic merger orbit eccentricities by considering a simplified
+model that captures the essential features of how the SMBH binary becomes bound
+and settles on an orbit with a given eccentricity.
+The components we include in this model are the two SMBHs,
+a smooth background potential, and dynamical friction.
+For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
+to lie in the $xz$-plane.
+In this case the relative motion of the BHs can be treated as a one-body problem,
+with the equation of motion
+\begin{equation}
+\ddot{\vb{x}} = -\frac{2 G M_\bullet}{|\vb{x}|^3} \vb{x} + \vb{a}_\mathrm{bg} + \vb{a}_\mathrm{DF},
+\end{equation}
+where $\vb{x}$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
+$\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
+and $\vb{a}_\mathrm{DF}$ the acceleration due to dynamical friction.
+
+The background galaxy is modelled as a with constant density spheroidal potential,
+causing the acceleration \drafting{(e.g.\ B\&T)}
+\begin{equation}
+\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y,A_z) \vb{x},
+\end{equation}
+where $\rho$ is the stellar density,
+and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
+\begin{equation}
+A_x = A_y  = 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]
+\end{equation}
+\begin{equation}
+A_z = \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+        \left[ \frac{1}{1-e_\mathrm{s}^2} 
+        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
+\end{equation}
+With eccentricity values of $e_\mathrm{s} \approx 0.8\textnormal{--}0.9$,
+this spheroidal potential matches fairly well with the central potential in the galaxy during the
+BH binary binding phase, apart from the contributions from the BHs themselves and the groups of stars bound to them.
+\drafting{(Figure showing this?)}
+
+For dynamical friction, we use the \drafting{Chandrasekhar ref} formula assuming
+a Maxwellian distribution with a constant velocity dispersion $\sigma$.
+This formula gives the acceleration of a single BH due to dynamical friction as
+\begin{equation}\label{eq:chandra_df}
+\vb{a}_{1,\mathrm{DF}} = - 4 \pi G^2 M_\bullet \rho \ln{\Lambda} 
+                        (\operatorname{erf}(X) - 2 \pi^{-1/2} X e^{-X^2}) \frac{\vb{v}_{1}}{|\vb{v}_{1}|^3},
+\end{equation}
+where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
+$X = |\vb{v}_1|/\sqrt{2}\sigma$,
+and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 10$.
+To account for differences between the constant density background of stars
+assumed here and the more complex system that is present in the simulations
+we additionally multiply the acceleration given by Eq. \eqref{eq:chandra_df}
+with a scaling and cut-off function
+\begin{equation}
+f(r) = \frac{\alpha}{1 + \exp[(r_\mathrm{c}-r)/d_\mathrm{c}]},
+\end{equation}
+which serves to remove the effect of dynamical friction when the BH binary has become
+hard in addition to allowing the strength of dynamical friction to be scaled to
+achieve similar behaviour as observed in the simulations.
+We set the scaling factor $\alpha=0.5$, and use cut-off scales $r_\mathrm{c}=2 a_\mathrm{h}$
+and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
+with $a_\mathrm{h}$ being the semimajor-axis of a hard BH binary.
+The total dynamical friction term is then
+\begin{equation}
+\vb{a}_{\mathrm{DF}} = 2 f(|\vb{x}|) \vb{a}_{1,\mathrm{DF}}.
+\end{equation}
+
+\drafting{
+Description of the model
+\begin{itemize}
+\item scaling of the model?
+\item initial condition: specify impact parameter, initial velocity and separation. Like the radial trajectories in the sims.
+\item solving numerically
+\end{itemize}
+
+Findings
+\begin{itemize}
+\item Final $e$ varies with impact parameter for given $v_0$, two minima on small scales at $b$s that are likely to occur in simulations. 
+\item location and depth of minima varies with $v_0$, for some values possible to span the full range of $e$ for different impact params.
+\end{itemize}
+
+Interpretation:
+Impact param differences seeded by the brownian motion in sims,
+but few pc differences small enough that you can't measure them from large scale merger orbits anyway.
+So the scatter is physically relevant.
+
+}

### 2023-04-28 papers/paper-eccentricity/scripts/toy_model.py

@@ -5,14 +5,12 @@ import matplotlib.pyplot as plt
 from scipy.integrate import solve_ivp
 from scipy.special import erf
 from scipy.interpolate import interp1d
-from scipy.optimize import minimize
+from scipy.optimize import minimize_scalar
 
 # units Gyr, 1e10 Msol, kpc
 G = 44900
 km_per_s = 1.02201216 #kpc/Gyr
 
-BH_rel_vel0 = 800*km_per_s # reference relative vel of BHs
-#BH_rel_vel0 = 760*km_per_s
 
 class System:
     
@@ -22,7 +20,7 @@ class System:
                  stellar_sigma=310*km_per_s, df_fudge_factor=0.5,
                  ):
         
-        self.M_BH = 3e-1 # single BH mass
+        self.M_BH = M_BH # single BH mass
          
         e2s = e_spheroid**2
         A1 = (1-e2s)/e2s*(1/(1-e2s) - 1/(2*e_spheroid)*np.log((1+e_spheroid)/(1-e_spheroid)))
@@ -34,11 +32,14 @@ class System:
         self.df_fudge_factor = df_fudge_factor
         self.logL = 10
 
+        self.a_hard = G * M_BH / (8 * stellar_sigma**2)
+        self.r_infl = np.cbrt(2*M_BH/(rho*(4/3*np.pi)))
+
 
 # The actual potential in the sims is fairly similar to a spheroidal potential
 # with e~0.7-0.9
     def ellipsoid_accel(self, x):
-        return -2*np.pi*G*self.rho * self.Avec * x
+        return -2 * np.pi * G * self.rho * self.Avec * x
 
 
     def dynamical_friction(self, x, v):
@@ -59,7 +60,9 @@ class System:
 
     def df_decoupling_factor(self, x):
         r = np.linalg.norm(x,axis=0)
-        return 1/(1+np.exp(-(r-0.02)/0.005))
+        cutoff_point = self.a_hard * 2
+        cutoff_scale = self.a_hard * 0.5
+        return 1/(1+np.exp(-(r-cutoff_point)/cutoff_scale))
 
 
 # x is the relative distance vector between BHs
@@ -76,9 +79,8 @@ class System:
             return np.append(y[2:], dvdt, axis=0)
 
         def min_E(t,y):
-            min_a = 2e-2
             E = self.orbital_energy(y[:2], y[2:])
-            return E + G*self.M_BH/min_a
[…]

### 2023-05-08 papers/paper-eccentricity/sections/results.tex
Add some toy model details to draft

@@
 with the equation of motion
-\begin{equation}
+\begin{equation} \label{eq:toy_model_eom}
 \ddot{\vb{x}} = -\frac{2 G M_\bullet}{|\vb{x}|^3} \vb{x} + \vb{a}_\mathrm{bg} + \vb{a}_\mathrm{DF},
@@
 \begin{equation}
-f(r) = \frac{\alpha}{1 + \exp[(r_\mathrm{c}-r)/d_\mathrm{c}]},
+f(a) = \frac{\alpha}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]},
 \end{equation}
@@
 achieve similar behaviour as observed in the simulations.
-We set the scaling factor $\alpha=0.5$, and use cut-off scales $r_\mathrm{c}=2 a_\mathrm{h}$
+Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f=\alpha$.
+We set the scaling factor $\alpha=0.5$, and use cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
 and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
@@
 \begin{equation}
-\vb{a}_{\mathrm{DF}} = 2 f(|\vb{x}|) \vb{a}_{1,\mathrm{DF}}.
+\vb{a}_{\mathrm{DF}} = 2 f(a) \vb{a}_{1,\mathrm{DF}}.
 \end{equation}
 
+After specifying the initial position and velocity, as well as the other parameters of the model,
+the equation of motion can be solved using standard numerical integration techniques.
+We use the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{scipy}
+for this purpose.
+To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
+we specify the initial conditions confined to the $xz$-plane as 
+\begin{gather}
+\vb{x}_0 = (r_0, b)\\
+\dot{\vb{x}}_0 = (-v_0, 0).
+\end{gather}
+We fix $r_0 = 100\,\mathrm{pc}$,
+and use model parameters
+$\rho = 400 \,\Msun\,\mathrm{pc}^{-3}$,
+$\sigma=200\,\mathrm{km\,s^{-1}}$,
+$M_\bullet=10^8\,\Msun$,
+and
+$e_\mathrm{s}=0.85$,
+which match fairly well with the central properties of the merging galaxy when
+the SMBHs are becoming bound in the $e=0.99$ merger orbit simulations.
+\drafting{(Check the params)}
+
 \drafting{
-Description of the model
+Figures to do:
 \begin{itemize}
-\item scaling of the model?
-\item initial condition: specify impact parameter, initial velocity and separation. Like the radial trajectories in the sims.
-\item solving numerically
+\item $b$-$e$ curve at a few different $v_0$ values, shows minima in $e$.
+      Also plot data points from high-e simulations, we can hopefully match the curve shape at least roughly.
+      Possibly also deflection angle plot.
+\item Plot of the minimum possible eccentricity at different initial velocities, and the corresponding $b$, deflection angles.
 \end{itemize}
+}
 
+\drafting{
 Findings
@@
 \item location and depth of minima varies with $v_0$, for some values possible to span the full range of $e$ for different impact params.
+\item Minimum $e$ for deflection angles $\sim 60^\circ, 120^\circ$.
 \end{itemize}
+}
 
+\drafting{
 Interpretation:
-Impact param differences seeded by the brownian motion in sims,
+\begin{itemize}
+\item Realistic merger orbits generally result in linearly plunging orbits just before binary formation,
+        with fairly small impact parameters.
+\item Often the relative velocities are in the range that leads to deflections around $90^\circ$
+      (some reason for this from energy considerations, since $b_{90}=GM_\mathrm{tot}/v^2$?),
+      small changes in impact param around that point hit the eccentricity minima in this model.
+\item Impact param differences seeded by the Brownian motion in sims,
       but few pc differences small enough that you can't measure them from large scale merger orbits anyway.
-So the scatter is physically relevant.
-
+\item Conclusion: the scatter is physically relevant, not pure numerics.
+\end{itemize}
 }

### 2023-05-08 papers/paper-eccentricity/scripts/toy_model.py

@@ -5,7 +5,7 @@ import matplotlib.pyplot as plt
 from scipy.integrate import solve_ivp
 from scipy.special import erf
 from scipy.interpolate import interp1d
-from scipy.optimize import minimize_scalar
+from scipy.optimize import minimize_scalar, shgo
 
 # units Gyr, 1e10 Msol, kpc
 G = 44900
@@ -58,18 +58,19 @@ class System:
 
         return 2*df_single * self.df_fudge_factor
 
-    def df_decoupling_factor(self, x):
-        r = np.linalg.norm(x,axis=0)
+    def df_decoupling_factor(self, x, v):
+        a = self.semimajor_axis(x,v)
+        a[a<0] = np.inf
         cutoff_point = self.a_hard * 2
         cutoff_scale = self.a_hard * 0.5
-        return 1/(1+np.exp(-(r-cutoff_point)/cutoff_scale))
+        return 1/(1+np.exp(-(a-cutoff_point)/cutoff_scale))
 
 
 # x is the relative distance vector between BHs
     def accel(self, x, v):
         return (-G * 2 * self.M_BH/np.linalg.norm(x,axis=0)**3 * x 
                 + 2 * self.ellipsoid_accel(x/2)
-                + self.dynamical_friction(x,v)*self.df_decoupling_factor(x)
+                + self.dynamical_friction(x,v)*self.df_decoupling_factor(x,v)
                 )
         
 
@@ -113,7 +114,7 @@ def compute_res(args):
     sys, b, v, r = args
     x0 = [r,b]
     v0 = [-v,0]
-    tmax = 200 * r/v
+    tmax = 100 * r/v
     ts = np.linspace(0,tmax, 500)
 
     x,v,t = sys.integrate(x0,v0,ts)
@@ -135,11 +136,13 @@ def main_plots():
     M_BH_ref = 1e-2 # single BH mass
     rho_ref = 40 # stellar density
     sigma_ref = 200*km_per_s # stellar velocity dispersion
-    gal_e = 0.9
+    gal_e = 0.89
 
-    v0_per_sigma = 2.5 # initial BH velocity / stellar sigma
-    r0_per_rinfl = 3 # initial BH separation/single BH influence radius
-    bmin_per_r0,bmax_per_r0 = 1e-3, 2e-1 # impact parameter limits / initial separation
+    v0_per_sigma = 2.3 # initial BH velocity / stellar sigma
+    r0_per_rinfl = 1 # initial BH separation/single BH influence radius
[…]

### 2023-05-09 papers/paper-1/diagrams/hm_coresersic.tex

@@
     % define layer 1
-    \node[obs] (Shat) {$\hat{\Sigma}$};
-    \factor[below=of Shat] {S_f} {right:Lognorm} {} {};
-    \node[det, below=of S_f] (S) {$\Sigma$};
+    \node[obs] (Shat) {$\log_{10}\hat{\Sigma}$};
+    \factor[below=of Shat] {S_f} {right:Norm} {} {};
+    \node[det, below=of S_f] (S) {$\log_{10}\Sigma$};
     \node[obs, below=of S] (R) {$\hat{R}$};
@@
     % define layer 2
-    \node[latent, left=of S] (Sb) {$\tilde{\Sigma}_\mathrm{b}$};
+    \node[latent, left=of S] (Sb) {$\log_{10}{\Sigma}_\mathrm{b}$};
     \node[latent, below=of Sb] (Re) {$R_\mathrm{e}$};
@@
     \node[latent, above=of g] (n) {$n$};
-    \node[latent, above=of n] (a) {$a$};
+    \node[latent, above=of n] (a) {$\alpha$};
 
     % define layer 3
-    \factor[left=of rb] {rb_f} {below:$\mathcal{N}$} {} {};
+    \factor[left=of rb, xshift=-0.5em] {rb_f} {below:$\mathcal{N}_\mathrm{T}$} {} {};
     \node[hyper, left=of rb_f, yshift=-1.2em] (rbs) {$r_{\mathrm{b},\sigma}$};
@@
 
-    \factor[left=of Re] {Re_f} {below:$\mathcal{N}$} {} {};
+    \factor[left=of Re, xshift=-0.5em] {Re_f} {below:$\mathcal{N}_\mathrm{T}$} {} {};
     \node[hyper, left=of Re_f, yshift=-1.2em] (Res) {$R_{\mathrm{e},\sigma}$};
@@
 
-    \factor[left=of Sb] {Sb_f} {below:$\mathcal{N}$} {} {};
-    \node[hyper, left=of Sb_f, yshift=-1.2em] (Sbs) {$\tilde{\Sigma}_{\mathrm{b},\sigma}$};
-    \node[hyper, left=of Sb_f, yshift=1.2em] (Sbm) {$\tilde{\Sigma}_{\mathrm{b},\mu}$};
+    \factor[left=of Sb, xshift=-0.5em] {Sb_f} {below:$\mathcal{N}$} {} {};
+    \node[hyper, left=of Sb_f, yshift=-2.5em, xshift=-1em] (Sbs) {$\log_{10} \Sigma_{\mathrm{b},\sigma}$};
+    \node[hyper, left=of Sb_f, yshift=2.5em, xshift=-1em] (Sbm) {$\log_{10}  \Sigma_{\mathrm{b},\mu}$};
 
-    \factor[left=of g] {g_f} {below:$\mathcal{N}$} {} {};
+    \factor[left=of g, xshift=-0.5em] {g_f} {below:$\mathcal{N}_\mathrm{T}$} {} {};
     \node[hyper, left=of g_f, yshift=-1.2em] (gs) {$\gamma_\sigma$};
@@
 
-    \factor[left=of n] {n_f} {below:$\mathcal{N}$} {} {};
+    \factor[left=of n, xshift=-0.5em] {n_f} {below:$\mathcal{N}_\mathrm{T}$} {} {};
     \node[hyper, left=of n_f, yshift=-1.2em] (ns) {$n_\sigma$};
@@
 
-    \factor[left=of a] {a_f} {below:$\mathcal{N}$} {} {};
-    \node[hyper, left=of a_f, yshift=-1.2em] (as) {$a_\sigma$};
-    \node[hyper, left=of a_f, yshift=1.2em] (am) {$a_\mu$};
+    \factor[left=of a, xshift=-0.5em] {a_f} {below:$\mathcal{N}_\mathrm{T}$} {} {};
+    \node[hyper, left=of a_f, yshift=-1.2em] (as) {$\alpha_\sigma$};
+    \node[hyper, left=of a_f, yshift=1.2em] (am) {$\alpha_\mu$};
 
@@
     % plates
-    \plate {rS} {(R)(S)(sigma)(Shat)} {$\forall R,\sigma_\Sigma^2$};
+    \plate {rS} {(R)(S)(sigma)(Shat)} {$\forall \hat{R},\hat{\sigma}_\Sigma^2$};
     \plate {pars} {(R)(S)(sigma)(rb)(Re)(Sb)(g)(n)(a)} {$\forall$ child};

### 2023-05-09 papers/paper-1/sections/results.tex

@@
 \end{enumerate}
+We begin our analysis by confirming the previous result of \citet{nasim20}, that \textit{intra-}simulation eccentricity variance decreases as a function of mass resolution between the SMBH and stellar particles.
+For each simulation at a given mass resolution, we take the variance of the eccentricity $\sigma^2_e$ across five orbital periods centred on when the SMBH binary becomes hard, $a=a_\mathrm{h}$. 
+We then determine the mean of the variances, and take the square-root of this value, to determine our expected intra-simulation eccentricity spread $\bar{\sigma}_e$.
+We fit the same functional form as \citet{nasim20} between $\bar{\sigma}_e$ and $N_\star$, namely:
+\begin{equation}
+    \bar{\sigma}_e(N_\star;k) = \frac{k}{\sqrt{N_\star + k^2}},
+\end{equation}
+and find the fit is optimised with a choice of $k=5.31$. 
+This is shown in \autoref{fig:nasim}.
+
+\begin{figure}
+    \includegraphics[width=0.5\textwidth]{nasim_plot.pdf}
+    \caption{\drafting{Intra-simulation binary ecccentricity standard deviation as a function of mass resolution, where the variation is seen to decrease as $\propto 1/\sqrt{N_\star}$.
+Error bars indicate the $1-\sigma$ spread in $\bar{\sigma}_e$ (note the error bars pertain to the $y$-axis, and not the $x$-axis.)}}
+    \label{fig:nasim}
+\end{figure}
 

### 2023-05-09 papers/paper-1/sections/stats.tex

@@
 
-In this work, we have 36 merger configuration families, with each family containing ten perturbed realisations of the same initial state. 
-As we only slightly perturb the SMBH phase space coordinates, each \child simulation can be thought of as a sampled observation of the general galaxy-merger configuration for its corresponding family.
+In this work, we have eight mass resolution families, with each family containing ten realisations of the same initial state. 
 Thus, we are able to employ hierarchical modelling to recover latent parameter distributions for various observable quantities of the merger remnant.

### 2023-05-09 papers/paper-2/sections/numerical.tex
Edits to paper 1

--- 2022-09-06 paper/sections/numerical.tex
@@
 \begin{equation}\label{eq:om}
-    f(Q) = \frac{1}{\sqrt{8}\pi^2} \left[ \int_0^Q \dv{\Psi}{\sqrt{Q-\Psi}} \dv{^2\nu_q}{\Psi^2} + \frac{1}{\sqrt{Q}} \left(\dv{\nu_Q}{\Psi}\right)_{\Psi=0}\right],
+    f(Q) = \frac{1}{\sqrt{8}\pi^2} \left[ \int_0^Q \dv{\Psi}{\sqrt{Q-\Psi}} \dv{^2\nu_Q}{\Psi^2} + \frac{1}{\sqrt{Q}} \left(\dv{\nu_Q}{\Psi}\right)_{\Psi=0}\right],
 \end{equation}
@@
 
+In typical N-body simulations, one tests the sensitivity of the model by Monte Carlo sampling the phase space distribution function given by \autoref{eq:eddington} a number of times with different random seeds.
+This is done to quantify the uncertainty in the model due to the discrete sampling of the phase space of field particles, which induce a Brownian motion to the orbit of the more massive SMBH particle.
+We observe that resampling \autoref{eq:eddington} is in effect displacing the centre of mass (CoM) of the system with the respect to the SMBH.
+To avoid the need to run a large number of simulations from the large $3.0\Rvir$ initial separation, we instead \textit{displace the SMBH} with respect to the CoM of the system at a time $t_\mathrm{perturb}$ after the simulation has begun.
+This is done by measuring the Brownian-motion induced displacement in position and velocity space of the SMBH from the CoM for a galaxy in isolation.
+After identifying the displacement along each axis to be Gaussian-distributed, we use the variance of this distribution to reposition the SMBH relative to the CoM. 
+We thus `recentre' the SMBH on the CoM with a perturbation of $\mathcal{N}_\text{pos}(\mu=0\,\text{pc}, \sigma^2=100\,\text{pc}^2)$ for each position axis, and $\mathcal{N}_\text{vel}(\mu=0\,\text{km\,s}^{-1}, \sigma^2=100\,\text{km}^2\text{\,s}^{-2})$ for each velocity axis.
+In this way, we mimic the effect of Brownian motion for a single run beginning at large separation (the \parent{} run) to generate an arbitrary number of perturbed \child{} runs. 
+The perturbation is applied no later than the second-to-last pericentre passage before the SMBH particles form a bound binary system, and ensure the separation of the SMBHs at the time of the perturbation is of the order 1000 larger than the perturbation applied.
+
 The progenitors are evolved through the \textit{parent phase} from $t(r_0)=0$ with \gadget until the SMBHs form a bound binary at $t_\text{stop}$, with a stellar softening of $\epsilon_\star=10\,\text{pc}$ and \gadget integration tolerances \texttt{ErrTolIntAccuracy}=$0.02$ and \texttt{ErrTolForceAcc}=$0.005$.

### 2023-05-09 papers/paper-eccentricity/sections/discuss.tex

\section{Discussion}

\drafting{
    An interesting point to discuss is the origin of our eccentricity scatter. Early tests of the Gualandris setup seems to indicate that low eccentricity mergers do not demonstrate orbital flips, and are more converged. How to distinguish if the flip is the cause of the eccentricity scatter, or just correlated with eccentricity scatter?
}

### 2023-05-09 papers/paper-eccentricity/sections/numerical.tex

@@
 \end{itemize}
+
+\subsection{Galaxy ICs}
+We construct our galaxy models to match those of \citet{gualandris22}.
+Each galaxy is represented as a stellar-only Dehnen sphere \citep{dehnen93} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+\begin{equation}\label{eq:dehnen}
+    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
+\end{equation}
+The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
+We test \drafting{X} different mass resolutions by sampling the stellar phase space distribution with varying number of stellar particles: \drafting{list resolutions here}.
+
+\subsection{Merger ICs}
+We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
+We follow the same orbital configurations as \citet{gualandris22}, with one additional orbit. 
+The chosen eccentricities $e$ are 0.5, 0.7, 0.9, and 0.99.
+We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
+An example of each orbital configuration is shown in \autoref{fig:orbits}.
+For each orbital configuration, we run \drafting{X} realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space.
+
+\begin{figure*}
+    \includegraphics{orbits}
+    \caption{\drafting{Orbits of the different merger systems, colour coded according to the initial orbital eccentricity.
+Note that the duration shown for each orbit is not currently consistent: probably should cut the plot to when the binary semimajor axis reaches some set value}}
+    \label{fig:orbits}
+\end{figure*}

### 2023-05-09 papers/paper-eccentricity/sections/numerical.tex

@@
-\section{Numerical Simulations}
+\section{Numerical Methods}
 
@@
 
-\subsection{Galaxy ICs}
+\subsection{Galaxy ICs}\label{ssec:gics}
 We construct our galaxy models to match those of \citet{gualandris22}.
@@
 For each orbital configuration, we run \drafting{X} realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
+Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
+Additionally, we run the $e=0.9$ and $e=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
+
+\subsection{Assessing Eccentricity Uncertainty}
+In our analysis, we wish to determine the population spread in eccentricity from a finite set of observations.
+We assess the uncertainty in eccentricity using two methods: an inverse modelling approach and a forward modelling approach.
+
+The first method follows \citet{nasim20}, and is an inverse problem. 
+For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over a \drafting{X Myr} time window after the SMBH binary has become hard.
+The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt13}
+\begin{equation*}
+    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{h}}{4},
+\end{equation*}
+where $r_\mathrm{h}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
+Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means.
+
+The second method is a forward modelling method, where we use Bayesian analysis to construct a simple hierarchical model. 
+We construct the hierarchical model using \textsc{Stan}, which is an implementation of Hamiltonian Monte Carlo.
+Our likelihood function is constructed using the analytical hardening relations \citep{quinlan96,sesana06}:
+\begin{equation}\label{eq:quinlan_H}
+    \dv{}{t}\left( \frac{1}{a} \right) = H \frac{G\rho}{\sigma},
+\end{equation}
+and
+\begin{equation}\label{eq:quinlan_K}
+    \dv{e}{t} = -K a^{-1} \dv{a}{t}.
+\end{equation}
+In these equations, $H$ and $K$ are dimensionless constants describing the evolution rate of the semimajor axis and eccentricity of the SMBH binary, respectively, and $\rho$ and $\sigma$ are the stellar density and velocity dispersion within the binary influence radius, respectively.
+We wish to determine the uncertainty on the free parameters $H' = HG\rho/\sigma$ and $K$, in addition to the initial conditions for the differential equation system, $a_\mathrm{h}$ and $e_\mathrm{h}$, at a time $t_\mathrm{h}$.
+Let us denote the set of parameters $[H', K, a_\mathrm{h}, e_\mathrm{h}]$ as $\theta$.
+To determine the uncertainty in the parameter set $\theta$, we solve the differential equation system analytically for a time period after the SMBH binary becomes hard where the effect of GW emission can be ignored: here taken to be $t_\mathrm{end} = t(a=10\,\mathrm{pc})$.
+Defining a time translation $t' = t - t_\mathrm{h}$, \autoref{eq:quinlan_H} becomes:
+\begin{equation}\label{eq:quinlan_H2}
+    \frac{1}{a(t')} \simeq H't' + \frac{1}{a_\mathrm{h}},
+\end{equation}
+and \autoref{eq:quinlan_K} becomes:
+\begin{equation}\label{eq:quinlan_K2}
+    e(t') \simeq K \ln\left[ \frac{a_\mathrm{h}}{a(t')} \right] + e_\mathrm{h}.
+\end{equation}
+The parallel to the first method now becomes apparent: in a given set of simulations, each simulation will have its own value of $\theta$, however each realisation of $\theta$ is itself a draw from a common population distribution that we wish to understand.
+If we assume that $e_\mathrm{h}$ is normally distributed with some mean $\mu_{e_\mathrm{h}}$ and standard deviation $\sigma_{e_\mathrm{h}}$, then the distribution of $\sigma_{e_\mathrm{h}}$ is precisely the uncertainty in the standard deviation of the binary eccentricity at the time the binary becomes hard.
+The primary benefit to using this second, forward modelling approach, is that we obtain an uncertainty estimate on $\sigma_{e_\mathrm{h}}$, which is not obtainable when using the first, inverse modelling approach.
+
+A probabilistic graphical model is shown in \drafting{FIGURE}.
 

### 2023-05-12 code/analysis_scripts/compare_binaries.py
Nasim eccentricity plot routine
- setting publish matploblibrc file changes default figure extension to pdf

@@ -45,7 +45,7 @@ if args.orbits:
 myr = ketjugw.units.yr * 1e6
 kpc = ketjugw.units.pc * 1e3
 cols = cmf.plotting.mplColours()
-linestyles = list(cmf.plotting.mplLines().values())
+linestyles = cmf.plotting.mplLines()
 num_dirs = len(ketju_dirs)
 SL.logger.debug(f"We will be plotting {num_dirs} different families...")
 

### 2023-05-12 code/analysis_scripts/nasim_eccentricity.py

@@ -1,6 +1,9 @@
 import argparse
+import os.path
 import numpy as np
+import scipy.optimize
 import matplotlib.pyplot as plt
+from matplotlib import matplotlib_fname
 import cm_functions as cmf
 
 
@@ -39,29 +42,64 @@ ax.set_xscale("log")
 ax.set_yscale("log")
 
 # list to store the values to plot
-x = []
+e_ini = []
+mass_res = []
 sigma_e = []
+sim_count = 0
 
 
 for d in hmq_dirs:
     # we can hack into the Kepler HM classes to extract the data
-    # TODO may need to fix place holder figname parameter
+    HMQ_files = cmf.utils.get_files_in_dir(d)
     km = cmf.analysis.KeplerModelHierarchy("", "", "")
-    km.extract_data(d, analysis_params)
+    km.extract_data(HMQ_files, analysis_params)
     sigma_e.append(np.nanstd(
         [np.nanmean(ecc) for ecc in km.obs["e"]]
     ))
-    if args.groups == "e":
-        x.append(km.obs["e_ini"])
-    else:
-        min_bh_mass = min(min(km.obs["mass1"]), min(km.obs["mass2"]))
-        try:
-            assert np.allclose(np.diff([m for m in km.obs["star_mass"]]))
-        except AssertionError:
-            SL.logger.exception(f"Non-unique stellar masses! A fair comparison cannot be made. Stellar masses are {km.obs['star_mass']}", exc_info=True)
-            raise
-        x.append(min_bh_mass/km.obs["star_mass"][0])
+    e_ini.extend(km.obs["e_ini"])
+    min_bh_mass = min(min(km.obs["mass1"]), min(km.obs["mass2"]))
+    try:
+        assert np.allclose(np.diff(np.concatenate(km.obs["star_mass"])), np.zeros(km.num_groups-1))
+    except AssertionError:
+        SL.logger.exception(f"Non-unique stellar masses! A fair comparison cannot be made. Stellar masses are {km.obs['star_mass']}", exc_info=True)
+        raise
+    mass_res.append(min_bh_mass/km.obs["star_mass"][0])
+    sim_count += km.num_groups
 
-ax.scatter(x, sigma_e, lw=0.5, ec="k")
+e_ini = np.concatenate(e_ini)
+mass_res = np.concatenate(mass_res)
+
[…]

### 2023-05-12 code/cm_functions/analysis/analysis_classes/KeplerModels.py

@@ -29,8 +29,8 @@ class _KeplerModelBase(StanModel_1D):
 
         Parameters
         ----------
-        dir : str
-            directory where hierarchical model data (from simulations) is
+        dir : list
+            list of paths to hierarchical model data files
         pars : dict
             analysis parameters
         """
@@ -39,7 +39,7 @@ class _KeplerModelBase(StanModel_1D):
         for f in dir:
             _logger.logger.info(f"Loading file: {f}")
             hmq = HMQuantitiesData.load_from_file(f)
-            obs["star_mass"] = hmq.particle_masses["stars"]
+            obs["star_mass"].append([hmq.particle_masses["stars"]])
             try:
                 idx = hmq.get_idx_in_vec(np.nanmedian(hmq.hardening_radius), hmq.semimajor_axis)
             except ValueError:

### 2023-05-12 code/cm_functions/env_config.py

@@ -35,7 +35,7 @@ if user_params["logging"]["file_level"] not in ["", " "]:
 try:
     assert fig_ext in ("pdf", "png")
 except AssertionError:
-    _cmlogger.logger.exception("Invalid figure format given in env_params.json! Reverting to default format: png")
+    _cmlogger.logger.exception(f"Invalid figure format {fig_ext} given in env_params.yml! Reverting to default format: png")
     fig_ext = "png"
 
 # make the figure directory
@@ -48,11 +48,11 @@ rc_file(os.path.join(this_dir, "plotting/matplotlibrc"))
 # get the git hash
 # only set git hash if in the collisionless-merger-sample repo
 if "collisionless-merger-sample" in os.getcwd():
-    # the standard git describe command, save git hash to json file for use 
+    # the standard git describe command, save git hash to yml file for use 
     # of cm_functions outside the collisionless-merger-sample repo
     git_hash = subprocess.run(["git", "describe", "--always", "--long", "--all"], check=True, capture_output=True).stdout.decode().rstrip("\n")
     internal_params["git_hash"] = git_hash
-    # make updates to env_params.json file
+    # make updates to env_params.yml file
     with open(env_params_file, "w") as f:
         yaml.safe_dump_all([user_params, internal_params], f, explicit_end=True, explicit_start=True)
 else:

### 2023-05-12 code/cm_functions/plotting/general.py

@@ -1,4 +1,3 @@
-from collections import OrderedDict
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib.gridspec import GridSpec
@@ -120,25 +119,26 @@ def mplLines(regular=5, loose=10, dense=1):
 
     Returns
     -------
-    : collections.OrderedDict
-        linestyles that can be accessed by keyword or .items() notation
+    : list
+        linestyles
     """
-    return OrderedDict([
+    d = dict([
             ("solid",                    (0, ())),
             ("dotted",                   (0, (1,1))),
             ("dashed",                   (0, (5,regular))),
             ("dashdotted",               (0, (3,regular,1,regular))),
             ("dashdotdotted",            (0, (3,regular,1,regular,1,regular))),
 
-            ("densely dashed",           (0, (5,dense))),
-            ("densely dashdotted",       (0, (3,dense,1,dense))),
-            ("densely dashdotdotted",    (0, (3,dense,1,dense,1,dense))),
+            ("densely-dashed",           (0, (5,dense))),
+            ("densely-dashdotted",       (0, (3,dense,1,dense))),
+            ("densely-dashdotdotted",    (0, (3,dense,1,dense,1,dense))),
 
-            ("loosely dotted",           (0, (1,loose))),
-            ("loosely dashed",           (0, (5,loose))),
-            ("loosely dashdotted",       (0, (3,loose,1,loose))),
-            ("loosely dashdotdotted",    (0,(3,loose,1,loose,1,loose))),
+            ("loosely-dotted",           (0, (1,loose))),
+            ("loosely-dashed",           (0, (5,loose))),
+            ("loosely-dashdotted",       (0, (3,loose,1,loose))),
+            ("loosely-dashdotdotted",    (0,(3,loose,1,loose,1,loose))),
     ])
+    return list(d.values())
 
 
 def mplChars():
@@ -211,7 +211,7 @@ def create_odd_number_subplots(nrow, ncol, fkwargs={}, gskwargs={}):
     return fig, np.array(ax)
 
 
-def nice_log10_scale(ax):
+def nice_log10_scale(ax, axis="y"):
     """
     Ensure nice log scaling on an axis
 
@@ -220,8 +220,15 @@ def nice_log10_scale(ax):
     ax : matplotlib.axes._subplots.AxesSubplot
         axis to plot to
     """
[…]

### 2023-05-12 code/cm_functions/plotting/plot_utils.py

@@ -28,6 +28,8 @@ def savefig(fname, fig=None, save_kwargs={}, force_ext=False):
     """
     if fig is None:
         fig = gcf()
+    if publishing_style_set:
+        fig_ext = "pdf"
     f = inspect.stack()[-1] # get outermost caller on stack
     now = datetime.now()
     if fig_ext == "png":
@@ -81,6 +83,8 @@ def set_publishing_style():
     Set a custom matplolibrc file that is designed specifically for 
     publishing-style plots. Mainly changes figure size and axis label size.
     """
+    global publishing_style_set
+    publishing_style_set = True
     rcdefaults()
     rc_file(os.path.join(os.path.dirname(os.path.realpath(__file__)), "matplotlibrc_publish"))
     _logger.logger.info("Publishing Matplotlib style set")

### 2023-05-12 papers/paper-eccentricity/sections/numerical.tex
-add figures and small amount of text to results

@@
 The first method follows \citet{nasim20}, and is an inverse problem. 
-For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over a \drafting{X Myr} time window after the SMBH binary has become hard.
+For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
 The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt13}

### 2023-05-12 papers/paper-eccentricity/sections/results.tex
Nasim eccentricity plot routine
- setting publish matploblibrc file changes default figure extension to pdf
-add figures and small amount of text to results

@@
 \end{itemize}
+\begin{figure*}
+    \includegraphics{nasim_scatter_e0-990.pdf}
+    \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
+    \label{fig:nasim_e90}
+\end{figure*}
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:nasim_e90}. 
+We find that the scaling of $\sigma_e$ is better described by a relation proportional to $1/N_\star$, as opposed to the $1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
+We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
 
@@
 \begin{figure*}
-    \includegraphics{nasim_style_plot}
-    \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
-    \label{fig:nasim}
-\end{figure*}
-
-\begin{figure*}
     \includegraphics{hyperparameter}
@@
 \end{figure*}
+
+\begin{figure}
+    \includegraphics{deflect_e-0900}
+    \caption{\drafting{Theta-e plane for $e_0=0.90$ orbit}}
+    \label{fig:deflect_e90}
+\end{figure}
+
+\begin{figure}
+    \includegraphics{deflect_e-0990}
+    \caption{\drafting{Theta-e plane for $e_0=0.99$ orbit}}
+    \label{fig:deflect_e99}
+\end{figure}
 

### 2023-05-12 code/cm_functions/analysis/analysis_classes/KeplerModels.py
fix bug in KeplerModels that resulted in an extra group being added if a data extraction process failed

@@ -39,7 +39,6 @@ class _KeplerModelBase(StanModel_1D):
         for f in dir:
             _logger.logger.info(f"Loading file: {f}")
             hmq = HMQuantitiesData.load_from_file(f)
-            obs["star_mass"].append([hmq.particle_masses["stars"]])
             try:
                 idx = hmq.get_idx_in_vec(np.nanmedian(hmq.hardening_radius), hmq.semimajor_axis)
             except ValueError:
@@ -73,6 +72,7 @@ class _KeplerModelBase(StanModel_1D):
                 obs["e_ini"].append([hmq.initial_galaxy_orbit["e0"]])
             except AttributeError:
                 _logger.logger.warning(f"File {f} is missing the 'initial_galaxy_orbit' attribute. Ideally, re-run HMQ extraction process.")
+            obs["star_mass"].append([hmq.particle_masses["stars"]])
             i += 1
         self.obs = obs
 

### 2023-05-15 papers/paper-eccentricity/paper-eccentricity.tex
Add some ideas and drafting to the paper
Start writing about the toy model
- add to results section

@@
 \usepackage{amsmath}	% Advanced maths commands
-\usepackage{amssymb}	% Extra maths symbols
+%\usepackage{amssymb}	% Extra maths symbols
 
@@
 \newcommand{\mstar}{\textsc{Mstar}}
-\newcommand{\gadget}{\textsc{Gadget-3}}
+\newcommand{\gadget}{\textsc{Gadget-4}}
 
-\newcommand{\Msun}{\ensuremath{M_{\sun}}}
+\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}
 \newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}
@@
 \newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}
+\newcommand{\vb}[1]{\ensuremath{\mathbf{#1}}}
 
@@
 
-\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel05}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.
+\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel22}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.
 

### 2023-05-15 papers/paper-eccentricity/sections/numerical.tex
- add to results section

@@
 We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
-We follow the same orbital configurations as \citet{gualandris22}, with one additional orbit. 
-The chosen eccentricities $e$ are 0.5, 0.7, 0.9, and 0.99.
+We follow the same orbital configurations as \citet{gualandris22}, with three additional orbit. 
+The chosen eccentricities $e$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
 We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
 An example of each orbital configuration is shown in \autoref{fig:orbits}.
-For each orbital configuration, we run \drafting{X} realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
+For each orbital configuration, we run ten realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
 Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.

### 2023-05-15 papers/paper-eccentricity/sections/results.tex
fix bug in KeplerModels that resulted in an extra group being added if a data extraction process failed
- add to results section

@@
 \end{itemize}
+\begin{figure*}
+    \includegraphics{nasim_scatter_e0-900.pdf}
+    \caption{\drafting{Inverse modelling convergence of eccentricity scatter as a function of mass resolution for $e_0=0.90$. NOTE AXIS LABEL SHOULD BE MASS RESOLUTION NOT PARTICLE NUMBER HERE}}
+    \label{fig:nasim_e90}
+\end{figure*}
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:nasim_e99}. 
+Immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
+The peculiar values of $\sigma_e$ arises due to the assumed Gaussian distribution of the eccentricity between simulations.
+Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at each resolution which has a value of $e_\mathrm{h}<0.4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by removing these two points we recover the $1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim20}.
+However, these low eccentricity values are a valid measurement from the simulation, and highlight the weakness in the inverse modelling approach to determine the spread in eccentricity. 
+The cause of these low eccentricity values is discussed in \autoref{ssec:toy}.
+\drafting{Run Bayesian model, more robust?}
 
@@
     \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
-    \label{fig:nasim_e90}
+    \label{fig:nasim_e99}
 \end{figure*}
-We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:nasim_e90}. 
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:nasim_e99}. 
 We find that the scaling of $\sigma_e$ is better described by a relation proportional to $1/N_\star$, as opposed to the $1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
 We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
+\drafting{TODO: figure comparing different eccentricity scatters}
 
@@
 
-\subsection{Explaining the eccentricity scatter with a toy model}
+\subsection{Explaining the eccentricity scatter with a toy model}\label{ssec:toy}
 

### 2023-05-15 code/analysis_scripts/nasim_eccentricity.py
Nasim plot can mask out eccentricity values
KeplerModels raise error if minimum number of groups not extracted, need to implement for other HM classes

Publishing matplotlibrc has ticks on all axes

@@ -3,9 +3,7 @@ import os.path
 import numpy as np
 import scipy.optimize
 import matplotlib.pyplot as plt
-from matplotlib import matplotlib_fname
 import cm_functions as cmf
-
 
 
 parser = argparse.ArgumentParser(description="Perform the eccentricity dispersion calculation in Nasim et al. 2020", allow_abbrev=False)
@@ -15,6 +13,7 @@ parser.add_argument("-s", "--save", action="store_true", dest="save", help="save
 parser.add_argument("-P", "--Publish", action="store_true", dest="publish", help="use publishing format")
 parser.add_argument("-d", "--dir", type=str, action="append", default=[], dest="extra_dirs", help="other directories of HMQ files to compare")
 parser.add_argument("-g", "--groups", choices=["e", "res"], help="Data groups", default="e", dest="groups")
+parser.add_argument("-c", "--cut", type=float, help="eccentricity cut in std.devs.", default=None, dest="ecut")
 parser.add_argument("-v", "--verbosity", type=str, default="INFO", choices=cmf.VERBOSITY, dest="verbosity", help="set verbosity level")
 args = parser.parse_args()
 
@@ -35,28 +34,30 @@ if args.extra_dirs:
 # read in the analysis parameters
 analysis_params = cmf.utils.read_parameters(args.apf)
 
-fig, ax = plt.subplots(1,1)
-ax.set_xlabel(r"$N_\star$")
-ax.set_ylabel(r"$\sigma_e$")
-ax.set_xscale("log")
-ax.set_yscale("log")
-
 # list to store the values to plot
 e_ini = []
 mass_res = []
 sigma_e = []
-sim_count = 0
+sigma_e_cut = []
+suite_count = 0
 
 
 for d in hmq_dirs:
+    SL.logger.debug(f"Reading from directory: {d}")
     # we can hack into the Kepler HM classes to extract the data
     HMQ_files = cmf.utils.get_files_in_dir(d)
     km = cmf.analysis.KeplerModelHierarchy("", "", "")
     km.extract_data(HMQ_files, analysis_params)
-    sigma_e.append(np.nanstd(
-        [np.nanmean(ecc) for ecc in km.obs["e"]]
-    ))
-    e_ini.extend(km.obs["e_ini"])
[…]

### 2023-05-15 code/cm_functions/_backend/Loggers.py

@@ -22,7 +22,7 @@ class InternalLogger:
             redirect all warnings to the logger, by default True
         """
         self.logger = logging.getLogger(name)
-        self._console_level = console_level
+        self._console_level = self._ensure_valid_level(console_level)
         self._capture_warnings = capture_warnings
         self._base_name = self.logger.name
         self._is_copy = False
@@ -35,10 +35,12 @@ class InternalLogger:
         self._logfile = None
         self._file_level = None
         logging.captureWarnings(self._capture_warnings)
-    
+
+
     @property
     def name(self):
         return self.logger.name
+
 
     def _set_handler(self, handler, level, fmt):
         """
@@ -56,7 +58,8 @@ class InternalLogger:
         handler.setLevel(getattr(logging, level))
         handler.setFormatter(logging.Formatter(fmt))
         self.logger.addHandler(handler)
-    
+
+
     def _ensure_valid_level(self, v):
         """
         Helper function to ensure a valid logging level is parsed
@@ -80,7 +83,8 @@ class InternalLogger:
         if v not in ["NOTSET", "DEBUG", "INFO", "WARNING", "ERROR", "CRITICAL"]:
             raise ValueError(f"Logging level <{v}> invalid!")
         return v
-    
+
+
     def add_file_handler(self, logfile, file_level="ERROR"):
         """
         Add a file handler to the logger
@@ -93,12 +97,13 @@ class InternalLogger:
             logging level, by default "ERROR"
         """
         self._logfile = logfile
-        self._file_level = file_level
+        self._file_level = self._ensure_valid_level(file_level)
         if not self._is_copy:
             with open(logfile, "a") as f:
                 f.write(f"{datetime.now()}: parent logger <{self.name}> created\n")
         self._f_handler = logging.FileHandler(filename=logfile)
         self._set_handler(self._f_handler, file_level, fmt=self._fh_format)
+
 
     def copy(self, n=None):
         """
@@ -124,6 +129,8 @@ class InternalLogger:
         if self._logfile is not None:
             c.add_file_handler(self._logfile, self._file_level)
         return c
+
+
 
 
 class ScriptLogger(InternalLogger):

### 2023-05-15 code/cm_functions/analysis/analysis_classes/KeplerModels.py

@@ -37,7 +37,7 @@ class _KeplerModelBase(StanModel_1D):
         obs = {"angmom":[], "energy":[], "a":[], "e":[], "mass1":[], "mass2":[], "star_mass":[], "e_ini":[]}
         i = 0
         for f in dir:
-            _logger.logger.info(f"Loading file: {f}")
+            _logger.logger.debug(f"Loading file: {f}")
             hmq = HMQuantitiesData.load_from_file(f)
             try:
                 idx = hmq.get_idx_in_vec(np.nanmedian(hmq.hardening_radius), hmq.semimajor_axis)
@@ -74,6 +74,11 @@ class _KeplerModelBase(StanModel_1D):
                 _logger.logger.warning(f"File {f} is missing the 'initial_galaxy_orbit' attribute. Ideally, re-run HMQ extraction process.")
             obs["star_mass"].append([hmq.particle_masses["stars"]])
             i += 1
+        try:
+            assert i >= pars["stan"]["min_num_samples"]
+        except AssertionError:
+            _logger.logger.exception(f"We require no less than {pars['stan']['min_num_samples']} samples groups, but have only {i}!", exc_info=True)
+            raise
         self.obs = obs
 
         # some transformations we need

### 2023-05-15 code/cm_functions/analysis/analysis_classes/StanModel.py

@@ -298,6 +298,9 @@ class _StanModel:
             except AssertionError:
                 _logger.logger.exception(f"Observation {obs_name} has already been collapsed! Cannot collapse again!", exc_info=True)
                 raise
+            except IndexError:
+                _logger.logger.exception(f"Error collapsing {obs_name}, {self.obs[obs_name]}")
+                raise
         try:
             assert len(np.unique(dim)) == 1
         except AssertionError:

### 2023-05-16 code/analysis_scripts/analysis_submit.sh
Set up config for figure scripts in paper directory
- data files for paper figures

Nasim convergence plots for paper

Deflection angle and nasim convergence scripts can save data to a directory as pickle files

New class  that preserves the boolean state of a setting, used to ascertain if publishing style set for figures

@@ -11,21 +11,10 @@ module purge
 module restore py393
 
 
-#python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-099/dehnen_250K_mergers -m mc
+#python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-090/dehnen_4M_mergers -m mc
 
-#python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-099/dehnen_500K_mergers -m mc
+python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-099/dehnen_4M_mergers -m mc
 
-python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-090/dehnen_250K_mergers -m mc
-
-python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-090/dehnen_1M_mergers -m mc
-
-#python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-095/dehnen_2M_mergers -m mc
-
[…]

### 2023-05-16 code/analysis_scripts/deflection_angle.py

@@ -14,6 +14,7 @@ parser.add_argument("-o", "--orbits", type=int, dest="orbits", help="number of o
 parser.add_argument("-l", "--label", type=str, dest="label", help="labelling method", choices=["e", "res"], default=None)
 parser.add_argument("-P", "--Publish", action="store_true", dest="publish", help="use publishing format")
 parser.add_argument("-s", "--save", action="store_true", dest="save", help="save figure")
+parser.add_argument("-sd", "--save-data", dest="save_data", type=str, default=None, help="directory to save data to")
 parser.add_argument("-v", "--verbosity", type=str, choices=cmf.VERBOSITY, dest="verbosity", default="INFO", help="verbosity level")
 args = parser.parse_args()
 
@@ -41,9 +42,16 @@ else:
 
 
 # arrays to hold data
-thetas =  np.array([])
-median_eccs = np.array([])
-iqr_eccs = np.array([[],[]])
+data = dict(
+    thetas = [],
+    median_eccs = [],
+    low_iqr_ecc = [],
+    high_iqr_ecc = [],
+    e_ini = [],
+    mass_res = [],
+    threshold_angle = args.angle
+)
+
 
 for j, datdir in enumerate(data_dirs):
     SL.logger.info(f"Reading from directory: {datdir}")
@@ -65,6 +73,10 @@ for j, datdir in enumerate(data_dirs):
             else:
                 labels[j] = cmf.general.represent_numeric_in_scientific(hmq.mass_resolution())
                 legend_title = r"$M_\bullet/m_\star$"
+        
+        # save some data already
+        data["e_ini"].append(hmq.initial_galaxy_orbit["e0"])
+        data["mass_res"].append(hmq.mass_resolution())
 
         thetas[i], theta_idx = cmf.analysis.first_major_deflection_angle(hmq.prebound_deflection_angles, angle_defl)
         if theta_idx is None:
@@ -81,6 +93,12 @@ for j, datdir in enumerate(data_dirs):
         m, iqr = cmf.mathematics.quantiles_relative_to_median(hmq.eccentricity[period_idxs[0]:period_idxs[1]])
         median_eccs[i] = m
         iqr_eccs[0,i], iqr_eccs[1,i] = iqr
+    # save data
+    data["thetas"].extend(thetas*180/np.pi)
[…]

### 2023-05-16 code/analysis_scripts/nasim_eccentricity.py

@@ -10,6 +10,7 @@ parser = argparse.ArgumentParser(description="Perform the eccentricity dispersio
 parser.add_argument(type=str, help="path to directory of HMQ files", dest="path")
 parser.add_argument(type=str, help="path to analysis parameter file", dest="apf")
 parser.add_argument("-s", "--save", action="store_true", dest="save", help="save figure")
+parser.add_argument("-sd", "--save-data", dest="save_data", type=str, default=None, help="directory to save data to")
 parser.add_argument("-P", "--Publish", action="store_true", dest="publish", help="use publishing format")
 parser.add_argument("-d", "--dir", type=str, action="append", default=[], dest="extra_dirs", help="other directories of HMQ files to compare")
 parser.add_argument("-g", "--groups", choices=["e", "res"], help="Data groups", default="e", dest="groups")
@@ -52,6 +53,8 @@ for d in hmq_dirs:
     sigma_e.append(np.nanstd(eccs))
     if args.ecut is not None:
         sigma_e_cut.append(np.nanstd(eccs[np.abs(eccs-np.nanmean(eccs))<args.ecut*np.nanstd(eccs)]))
+    else:
+        sigma_e_cut.append(np.nan)
     try:
         assert np.allclose(np.diff(np.concatenate(km.obs["e_ini"])), np.zeros(km.num_groups-1))
     except AssertionError:
@@ -112,5 +115,15 @@ else:
 
 if args.save:
     cmf.plotting.savefig(os.path.join(cmf.FIGDIR, f"merger/nasim_scatter_{fig_prefix}.png"))
+
+if args.save_data is not None:
+    data = dict(
+        mass_res = mass_res,
+        e_ini = e_ini,
+        sigma_e = sigma_e,
+        sigma_e_cut = sigma_e_cut,
+        sigma_e_cut_threshold = args.ecut
+    )
+    cmf.utils.save_data(data, os.path.join(args.save_data, f"nasim_scatter_{fig_prefix}.pickle"))
 
 plt.show()

### 2023-05-16 code/cm_functions/_backend/States.py

@@ -1,0 +1,25 @@
+__all__ = [ "PublishingState"]
+
+class _State:
+    """
+    Simple object to store the state of a setting
+    """
+    def __init__(self) -> None:
+        self._setting_set = False
+    
+    def turn_on(self):
+        self._setting_set = True
+    
+    def turn_off(self):
+        self._setting_set = False
+    
+    def is_set(self):
+        return self._setting_set
+
+
+class PublishingState(_State):
+    def __init__(self) -> None:
+        """
+        Simple object to store the state of a plotting-style setting
+        """
+        super().__init__()

### 2023-05-16 code/cm_functions/analysis/analysis_classes/HMQuantitiesData.py

@@ -297,8 +297,8 @@ class HMQuantitiesData(HDF5Base):
 
         Returns
         -------
-        _type_
-            _description_
+        : float
+            mass resolution
         """
         if "stars" not in self.particle_masses:
             _logger.logger.error(f"Key 'stars' not present in 'particle_masses', trying 'dm' instead...")

### 2023-05-16 code/cm_functions/plotting/plot_utils.py

@@ -4,12 +4,15 @@ from matplotlib import rc_file, rcdefaults, rcParams
 from matplotlib.pyplot import gcf
 from PIL import Image
 import os.path
+from .._backend.States import PublishingState
 from ..env_config import _cmlogger, git_hash, username, date_format, fig_ext
 
 __all__ = ["savefig", "get_meta", "set_publishing_style", "get_figure_size"]
 
 _logger = _cmlogger.copy(__file__)
 
+# state information
+_PS = PublishingState()
 
 def savefig(fname, fig=None, save_kwargs={}, force_ext=False):
     """
@@ -28,11 +31,13 @@ def savefig(fname, fig=None, save_kwargs={}, force_ext=False):
     """
     if fig is None:
         fig = gcf()
-    if publishing_style_set:
-        fig_ext = "pdf"
+    if _PS.is_set():
+        _fig_ext = "pdf"
+    else:
+        _fig_ext = fig_ext
     f = inspect.stack()[-1] # get outermost caller on stack
     now = datetime.now()
-    if fig_ext == "png":
+    if _fig_ext == "png":
         now = now.strftime(date_format)
     # ensure things are deterministic
     try:
@@ -51,7 +56,7 @@ def savefig(fname, fig=None, save_kwargs={}, force_ext=False):
             # protect against cases where no extension is specified, and the 
             # file name has a "." in it
             _fname = fname_name if fname_ext in (".png", ".pdf") else fname
-            figname = f"{_fname}.{fig_ext}"
+            figname = f"{_fname}.{_fig_ext}"
             fig.savefig(figname, metadata=meta_data, **save_kwargs)
     finally:
         del f
@@ -83,8 +88,7 @@ def set_publishing_style():
     Set a custom matplolibrc file that is designed specifically for 
     publishing-style plots. Mainly changes figure size and axis label size.
     """
-    global publishing_style_set
-    publishing_style_set = True
+    _PS.turn_on()
     rcdefaults()
     rc_file(os.path.join(os.path.dirname(os.path.realpath(__file__)), "matplotlibrc_publish"))
     _logger.logger.info("Publishing Matplotlib style set")

### 2023-05-16 code/cm_functions/utils/data_handling.py

@@ -37,7 +37,7 @@ def save_data(data, filename, protocol=pickle.HIGHEST_PROTOCOL):
     assert(filename.endswith(".pickle"))
     with open(filename, 'wb') as f:
         pickle.dump(data, f, protocol=protocol)
-    _logger.logger.info(f"File {f} saved")
+    _logger.logger.info(f"File {filename} saved")
 
 
 def load_data(filename):

### 2023-05-16 papers/paper-eccentricity/scripts/eccentricity_scatter.py

@@ -1,0 +1,61 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pickle
+import cm_functions as cmf
+import figure_config
+
+
+#------------------------------------------------------------
+# plot of convergence in sigma_e across resolutions
+#------------------------------------------------------------
+
+# read in data
+data_e090 = cmf.utils.load_data(figure_config.data_path("nasim_scatter_e0-0.900.pickle"))
+data_e099 = cmf.utils.load_data(figure_config.data_path("nasim_scatter_e0-0.990.pickle"))
+
+# set up the figure
+fig, ax = plt.subplots(1,1)
+ax.set_xlabel(r"$M_\bullet/m_\star$")
+ax.set_xscale("log")
+ax.set_ylabel(r"$\sigma_e$")
+ax.set_yscale("log")
+marker_kwargs = {"edgecolor":"k", "lw":0.5}
+cmapper, sm = cmf.plotting.create_normed_colours(5e-3, 0.2, cmap="BuPu", normalisation="LogNorm")
+markers = cmf.plotting.mplChars()
+
+
+# plot data
+for i, (d, lab) in enumerate(zip((data_e090, data_e099), (r"$e_0=0.90$", r"$e_0=0.99$"))):
+    sc = ax.scatter(d["mass_res"], d["sigma_e"], label=f"{lab}", c=cmapper(1-d["e_ini"]), marker=markers[0], zorder=10, **marker_kwargs)
+    ax.scatter(d["mass_res"], d["sigma_e_cut"], label=f"{lab} ($<{d['sigma_e_cut_threshold']:.1f}\\sigma$)", c=cmapper(1-d["e_ini"]), marker=markers[1], s=sc.get_sizes()*1.5, zorder=1, **marker_kwargs)
+
+# add the sqrt{resolution} scaling line
+mass_res_seq = np.geomspace(2e3, 5e4, 10)
+ax.plot(mass_res_seq, 3*mass_res_seq**-0.5, c="k", label=r"$\propto 1/\sqrt{M_\bullet/m_\star}$")
+
+# final touch ups
+cmf.plotting.nice_log10_scale(ax, "xy")
+ax.legend(fontsize="small")
+
+# save
+cmf.plotting.savefig(figure_config.fig_path("convergence.pdf"), force_ext=True)
+plt.close(fig)
+
+
+#------------------------------------------------------------
+# plot of sigma_e across initial eccentricities
[…]

### 2023-05-16 papers/paper-eccentricity/scripts/figure_config.py

@@ -1,0 +1,19 @@
+import os
+from matplotlib import rc_file, rcdefaults
+
+this_dir = os.path.dirname(os.path.realpath(__file__))
+data_dir = os.path.abspath(os.path.join(this_dir, 'data/'))
+figure_dir = os.path.abspath(os.path.join(this_dir, '../figures/'))
+
+# get the complete path for saving a figure
+def fig_path(fname):
+    return os.path.join(figure_dir, fname)
+
+# get the complete path to a saved data file
+def data_path(dname):
+    return os.path.join(data_dir, dname)
+
+# make sure were using the same settings
+rcdefaults()
+rc_file(os.path.join(this_dir,'matplotlibrc_publish'))
+

### 2023-05-16 papers/paper-eccentricity/sections/numerical.tex
Orbit plot for paper
New parameter files for updated Hernquist runs

New organisation of plotting classes, with update to GradientPlot methods

@@
     \includegraphics{orbits}
-    \caption{\drafting{Orbits of the different merger systems, colour coded according to the initial orbital eccentricity.
-Note that the duration shown for each orbit is not currently consistent: probably should cut the plot to when the binary semimajor axis reaches some set value}}
-    \label{fig:orbits}
+    \caption{\drafting{Orbit of a SMBH in one of the e=0.99 mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Line coloured according to time.
+Point A shows a time just before the first hard scattering, and point B a time just after.
+The circle marker indicates the final position of the BH before the BHs become bound.}}
+    \label{fig:orbit}
 \end{figure*}

### 2023-05-16 code/analysis_scripts/compare_binaries.py

@@ -42,11 +42,12 @@ if args.orbits:
     ax2[1].set_xlabel("x/kpc")
     ax2[1].set_xlabel("y/kpc")
 
-myr = ketjugw.units.yr * 1e6
-kpc = ketjugw.units.pc * 1e3
+myr = cmf.general.units.Myr
+kpc = cmf.general.units.kpc
 cols = cmf.plotting.mplColours()
 linestyles = cmf.plotting.mplLines()
 num_dirs = len(ketju_dirs)
+total_sim_count = 0
 SL.logger.debug(f"We will be plotting {num_dirs} different families...")
 
 for j, d in enumerate(ketju_dirs):
@@ -78,11 +79,13 @@ for j, d in enumerate(ketju_dirs):
                     ax2[0].plot((bh.x[:,0]-op["x_CM"][:,0])/kpc, (bh.x[:,2]-op["x_CM"][:,2])/kpc, alpha=0.7, c=cols[j])
                     ax2[1].plot((bh.x[:,0]-op["x_CM"][:,0])/kpc, (bh.x[:,1]-op["x_CM"][:,1])/kpc, c=cols[j], alpha=0.7)
         line_count += 1
+        total_sim_count += 1
     if not bound_bhs_present:
         SL.logger.warning(f"No bound BHs present in {d}")
 
 try:
-    ax[0].legend(loc="upper right", **legend_kwargs)
+    if total_sim_count < 10:
+        ax[0].legend(loc="upper right", **legend_kwargs)
     ax[0].set_xscale("log")
     if args.save:
         now = datetime.now().strftime("%Y%m%d_%H%M%S")

### 2023-05-16 code/cm_functions/analysis/analysis_classes/GrahamModels.py

@@ -69,7 +69,7 @@ class _GrahamModelBase(StanModel_2D):
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             plotting axis
         """
         fig, ax = plt.subplots(2,3, figsize=figsize)
@@ -117,7 +117,7 @@ class GrahamModelSimple(_GrahamModelBase):
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             plotting axis
         """
         # latent parameter plots (corners, chains, etc)
@@ -156,7 +156,7 @@ class GrahamModelHierarchy(_GrahamModelBase):
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             plotting axis
         """
         # hyper parameter plots (corners, chains, etc)

### 2023-05-16 code/cm_functions/analysis/analysis_classes/KeplerModels.py

@@ -116,7 +116,7 @@ class _KeplerModelBase(StanModel_1D):
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             plotting axis
         """
         fig, ax = plt.subplots(1, 2, figsize=figsize)

### 2023-05-16 code/cm_functions/analysis/analysis_classes/PQModels.py

@@ -68,7 +68,7 @@ class _PQModelBase(StanModel_2D):
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             plotting axis
         """
         fig, ax = plt.subplots(1, 2, figsize=figsize)

### 2023-05-16 code/cm_functions/analysis/analysis_classes/QuinlanModels.py

@@ -85,7 +85,7 @@ class _QuinlanModelBase(StanModel_2D):
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             plotting axis
         """
         fig, ax = plt.subplots(2, 2, figsize=figsize)

### 2023-05-16 code/cm_functions/analysis/analysis_classes/StanModel.py

@@ -522,7 +522,7 @@ class _StanModel:
 
         Returns
         -------
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             corner plot
         """
         if levels is None:
@@ -614,7 +614,7 @@ class _StanModel:
         ----------
         gq : list
             variables to plot
-        ax : matplotlib.axes._subplots.AxesSubplot or np.ndarray of, optional
+        ax : matplotlib.axes.Axes or np.ndarray of, optional
             axes object to plot to, by default None
         xlabels : list, optional
             labels for the x-axis, by default None
@@ -623,7 +623,7 @@ class _StanModel:
 
         Returns
         -------
-        matplotlib.axes._subplots.AxesSubplot
+        matplotlib.axes.Axes
             plotting axes
         """
         if ax is None:
@@ -831,7 +831,7 @@ class StanModel_2D(_StanModel):
              None
         levels : list, optional
             HDI intervals to plot, by default None
-        ax : matplotlib.axes._subplots.AxesSubplot, optional
+        ax : matplotlib.axes.Axes, optional
             axis to plot to, by default None (creates new instance)
         collapsed : bool, optional
             plotting collapsed observations?

### 2023-05-16 code/cm_functions/initialise/MergerIC.py

@@ -203,6 +203,8 @@ class MergerIC:
             raise
         mg.write_hdf5_ic_file(filename=file_name, system=merger, save_plots=False)
         _logger.logger.info(f"Merger IC file written to {file_name}")
+        # copy parameter file to simulation directory
+        shutil.copyfile(self.paramfile, os.path.join(self.save_location, os.path.basename(self.paramfile)))
         # save parameters
         self.write_calculated_parameters()
         # print some velocity information about merger

### 2023-05-16 code/cm_functions/literature/LiteratureTables.py

@@ -115,7 +115,7 @@ class LiteratureTables:
         ----------
         var : str
             column to histogram
-        ax : matplotlib.axes._subplots.AxesSubplot, optional
+        ax : matplotlib.axes.Axes, optional
             axes to plot to, by default None
         label : str, optional
             legend label, by default None
@@ -124,7 +124,7 @@ class LiteratureTables:
 
         Returns
         -------
-        matplotlib.axes._subplots.AxesSubplot
+        matplotlib.axes.Axes
             plotting axis
         """
         if ax is None:
@@ -148,7 +148,7 @@ class LiteratureTables:
             quantile to plot
         var : str
             column to histogram
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             axes to plot to
         xaxis : bool, optional
             plot on the x-axis?, by default True
@@ -181,7 +181,7 @@ class LiteratureTables:
         yerr : str, list, tuple, optional
             column for y-errors, or list/tuple of strings for non-symmetric 
             errors, by default None
-        ax : matplotlib.axes._subplots.AxesSubplot, optional
+        ax : matplotlib.axes.Axes, optional
             axes to plot to, by default None
         scatter_kwargs : dict, optional
             scatter parameters parsed to scatter() or errorbar(), by default {}
@@ -192,7 +192,7 @@ class LiteratureTables:
 
         Returns
         -------
-        matplotlib.axes._subplots.AxesSubplot
+        matplotlib.axes.Axes
             plotting axis
         """
         if ax is None:

### 2023-05-16 code/cm_functions/plotting/PlotClasses.py

@@ -1,0 +1,200 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib import colors
+from ..env_config import _cmlogger
+
+
+__all__ = ["GradientLinePlot", "GradientScatterPlot"]
+
+
+_logger = _cmlogger.copy(__file__)
+
+
+
+class _GradientPlot:
+    """
+    
+    """
+    def __init__(self, ax, cmap="cividis", plot_kwargs={}):
+        """
+        Class to create pyplot plots with a colour gradient. The colour 
+        gradient is consistent between all lines/points in the figure. This is 
+        done by storing the data first, and then only plotting the data when 
+        explicitly called.
+
+        Parameters
+        ----------
+        ax : matplotlib.axes.Axes
+            axis to plot to
+        cmap : str, optional
+            pyplot colour map name, by default "cividis"
+        plot_kwargs : dict, optional
+            arguments to be parsed to either plt.plot() or plt.scatter(), by 
+            default {}
+        """
+        self.ax = ax
+        self.data_count = 0
+        self.all_x = []
+        self.all_y = []
+        self.all_c = []
+        self.all_label = []
+        self.cmap= getattr(plt.cm, cmap)
+        self.all_marker = []
+        self.all_pks = [plot_kwargs]
+        self.norm = [0,1]
+
+
+    def __len__(self):
+        return self.data_count
+
+
+    def add_data(self, x, y, c, label=None, marker="o", plot_kwargs={}):
+        """
+        Add a dataset to the plot (note that the data is just stored here for 
+        future use).
+
+        Parameters
+        ----------
+        x : np.ndarray
+            x data
+        y : np.ndarray
+            y data
+        c : np.ndarray
+            data to map colours to
+        label : _type_, optional
+            label of plot, by default None
+        marker : str, optional
+            end marker, by default "o"
+        plot_kwargs : dict, optional
+            dict of other parameters to parse to pyplot.plot() or pyplot.scatter
+            (), by default {}
+        """
+        try:
+            dat_len = [len(v) for v in (x,y,c)]
+            assert np.allclose(np.diff(dat_len), (0,0))
+        except AssertionError:
+            _logger.logger.exception(f"Input data must all be of the same length! Currently has lengths: {dat_len}", exc_info=True)
[…]

### 2023-05-16 code/cm_functions/plotting/__init__.py

@@ -1,3 +1,4 @@
 from .general import *
 from .specific_plots import *
 from .plot_utils import *
+from .PlotClasses import GradientLinePlot, GradientScatterPlot

### 2023-05-16 code/cm_functions/plotting/general.py

@@ -18,7 +18,7 @@ def draw_sizebar(ax, length, units, location="lower right", pad=0.1, borderpad=0
 
     Parameters
     ----------
-    ax : matplotlib.axes._subplots.AxesSubplot
+    ax : matplotlib.axes.Axes
         axis to add the bar to
     length : float
         length of scale bar in data units
@@ -159,7 +159,7 @@ def shade_bool_regions(ax, xdata, mask, **kwargs):
 
     Parameters
     ----------
-    ax : matplotlib.axes._subplots.AxesSubplot
+    ax : matplotlib.axes.Axes
         axis to plot to
     xdata : np.ndarray
         x data values
@@ -217,7 +217,7 @@ def nice_log10_scale(ax, axis="y"):
 
     Parameters
     ----------
-    ax : matplotlib.axes._subplots.AxesSubplot
+    ax : matplotlib.axes.Axes
         axis to plot to
     """
     if "y" in axis:

### 2023-05-16 code/cm_functions/plotting/specific_plots.py

@@ -1,4 +1,3 @@
-import warnings
 import numpy as np
 from scipy.interpolate import interp1d
 import matplotlib.pyplot as plt
@@ -12,7 +11,7 @@ from ..general import convert_gadget_time, units
 from ..env_config import _cmlogger
 
 
-__all__ = ["plot_galaxies_with_pygad", "GradientLinePlot", "GradientScatterPlot", "binary_param_plot", "twin_axes_from_samples", "voronoi_plot", "seaborn_jointplot_cbar", "draw_unit_sphere", "heatmap", "annotate_heatmap"]
+__all__ = ["plot_galaxies_with_pygad", "binary_param_plot", "twin_axes_from_samples", "voronoi_plot", "seaborn_jointplot_cbar", "draw_unit_sphere", "heatmap", "annotate_heatmap"]
 
 _logger = _cmlogger.copy(__file__)
 
@@ -50,7 +49,7 @@ def plot_galaxies_with_pygad(snap, return_ims=False, orientate=None, figax=None,
     fig : matplotlib.figure.Figure
         pyplot figure object
     ax : np.ndarray
-        array of matplotlib.axes._subplots.AxesSubplot instances
+        array of matplotlib.axes.Axes instances
     ims : list, optional
         list of images from pygad plotting routine
     """
@@ -94,163 +93,6 @@ def plot_galaxies_with_pygad(snap, return_ims=False, orientate=None, figax=None,
         return fig, ax
 
 
-class GradientPlot:
-    """
-    
-    """
-    def __init__(self, ax, cmap="viridis", plot_kwargs={}):
-        """
-        Class to create pyplot plots with a colour gradient. The colour 
-        gradient is consistent between all lines/points in the figure. This is 
-        done by storing the data first, and then only plotting the data when 
-        explicitly called.
-
-        Parameters
-        ----------
-        ax : matplotlib.axes._subplots.AxesSubplot
-            axis to plot to
-        cmap : str, optional
-            pyplot colour map name, by default "viridis"
-        plot_kwargs : dict, optional
-            arguments to be parsed to either plt.plot() or plt.scatter(), by 
-            default {}
-        """
-        self.ax = ax
-        self.data_count = 0
-        self.all_x = []
-        self.all_y = []
-        self.all_c = []
-        self.all_label = []
-        self.cmap= getattr(plt.cm, cmap)
-        self.all_marker = []
-        self.all_pks = [plot_kwargs]
-        self.norm = [0,1]
-    
[…]

### 2023-05-16 code/cm_functions/visualisation/animation.py

@@ -36,7 +36,7 @@ class SMBHtrajectory:
         ----------
         bhdata : str
             path to the ketju_bhs.hdf5 to read in
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             axis to plot to
         centre : int, optional
             which BH to centre on, by default 1
@@ -135,7 +135,7 @@ class OverviewAnimation:
             list of ordered snapshots to plot (note no reordering is done)
         fig : matplotlib.figure.Figure
             figure object
-        ax : matplotlib.axes._subplots.AxesSubplot
+        ax : matplotlib.axes.Axes
             axis to plot to
         centre : str, optional
             centring of animation, options are:

### 2023-05-16 papers/paper-eccentricity/scripts/eccentricity_scatter.py

@@ -1,6 +1,5 @@
 import numpy as np
 import matplotlib.pyplot as plt
-import pickle
 import cm_functions as cmf
 import figure_config
 

### 2023-05-16 papers/paper-eccentricity/scripts/orbit_plot.py

@@ -1,0 +1,78 @@
+import os.path
+import matplotlib.pyplot as plt
+import cm_functions as cmf
+import figure_config
+
+main_path = "/scratch/pjohanss/arawling/collisionless_merger/mergers/eccentricity_study/e-099/2M"
+data_dirs = [
+    os.path.join(main_path, "D_2M_a-D_2M_a-3.720-0.028"),
+    os.path.join(main_path, "D_2M_a-D_2M_b-3.720-0.028"),
+    os.path.join(main_path, "D_2M_a-D_2M_c-3.720-0.028"),
+]
+
+# turn off for fast plotting, line will be a single colour
+use_gradient_line = True
+
+# some shortcuts
+kpc = cmf.general.units.kpc
+Myr = cmf.general.units.Myr
+min_time_colour = 22
+A_idx = 31400
+B_idx = A_idx + 400
+
+
+# set up figure
+fig, ax = plt.subplots(1,1)
+ax.set_xlabel(r"$x/\mathrm{kpc}$")
+ax.set_ylabel(r"$z/\mathrm{kpc}$")
+axins = ax.inset_axes([0.1, 0.4, 0.5, 0.5])
+
+
+for dd in data_dirs:
+    ketjufile = cmf.utils.get_ketjubhs_in_dir(dd)[0]
+    bh1, bh2 = cmf.analysis.get_binary_before_bound(ketjufile)
+    bh1, bh2 = cmf.analysis.move_to_centre_of_mass(bh1, bh2)
+    # ensure the plotted BH starts in the upper right corner
+    if bh1.x[0,2] > 0:
+        bh = bh1
+    else:
+        bh = bh2
+    if use_gradient_line:
+        glp = cmf.plotting.GradientLinePlot(ax=ax, cmap="cividis")
+        glp.add_data(bh.x[1:,0]/kpc, bh.x[1:,2]/kpc, c=bh.t[1:]/Myr)
+        glp.plot(logcolour=False, vmin=min_time_colour)
+        glp.plot(logcolour=False, ax=axins, vmin=min_time_colour)
+        glp.add_cbar(ax, label=r"$t/\mathrm{Myr}$", extend="min")
+    else:
+        ax.plot(bh.x[1:,0]/kpc, bh.x[1:,2]/kpc)
+        axins.plot(bh.x[1:,0]/kpc, bh.x[1:,2]/kpc)
+    # annotate points just before and just after the first hard scatter
+    axins.annotate(
[…]

### 2023-05-16 papers/paper-eccentricity/paper-eccentricity.tex
Paper edits

@@
 % Keep the title short and informative.
-\title[Reviving Stochasticity]{Reviving stochasticity: eccentricity convergence depends on black hole binary orbit}
+\title[Reviving Stochasticity]{Reviving stochasticity: SMBH-binary eccentricity is inherently random}
 
@@
 Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
+Matias Mannerkoski,$^{1}$
 Others,$^{1}$
@@
 
-\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel22}, \textsc{Stan} \citep{carpenter17}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}, Arviz \citep{arviz}.
+\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel22}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}.
 

### 2023-05-16 papers/paper-eccentricity/sections/bayesian_methods.tex
Paper edits

--- 2023-05-12 papers/paper-eccentricity/sections/numerical.tex
@@
-\section{Numerical Methods}
-
-\begin{itemize}
-    \item describe toy model
-    \item describe galaxy ICs: 0.5 Dehnen spheres
-    \item describe merger orbit: low and high e
-\end{itemize}
-
-\subsection{Galaxy ICs}\label{ssec:gics}
-We construct our galaxy models to match those of \citet{gualandris22}.
-Each galaxy is represented as a stellar-only Dehnen sphere \citep{dehnen93} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
-\begin{equation}\label{eq:dehnen}
-    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
-\end{equation}
-The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
-We test \drafting{X} different mass resolutions by sampling the stellar phase space distribution with varying number of stellar particles: \drafting{list resolutions here}.
-
-\subsection{Merger ICs}
-We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
-We follow the same orbital configurations as \citet{gualandris22}, with one additional orbit. 
-The chosen eccentricities $e$ are 0.5, 0.7, 0.9, and 0.99.
-We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
-An example of each orbital configuration is shown in \autoref{fig:orbits}.
-For each orbital configuration, we run \drafting{X} realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
-Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
-Additionally, we run the $e=0.9$ and $e=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
-
-\subsection{Assessing Eccentricity Uncertainty}
-In our analysis, we wish to determine the population spread in eccentricity from a finite set of observations.
-We assess the uncertainty in eccentricity using two methods: an inverse modelling approach and a forward modelling approach.
-
-The first method follows \citet{nasim20}, and is an inverse problem. 
-For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
-The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt13}
-\begin{equation*}
-    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{h}}{4},
-\end{equation*}
-where $r_\mathrm{h}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
-Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means.
-
 The second method is a forward modelling method, where we use Bayesian analysis to construct a simple hierarchical model. 
@@
 A probabilistic graphical model is shown in \drafting{FIGURE}.
-
-\begin{figure*}
-    \includegraphics{orbits}
-    \caption{\drafting{Orbits of the different merger systems, colour coded according to the initial orbital eccentricity.
-Note that the duration shown for each orbit is not currently consistent: probably should cut the plot to when the binary semimajor axis reaches some set value}}
-    \label{fig:orbits}
-\end{figure*}

### 2023-05-16 papers/paper-eccentricity/sections/numerical.tex
Paper edits

--- 2023-05-15 papers/paper-eccentricity/sections/numerical.tex
@@
 We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
-An example of each orbital configuration is shown in \autoref{fig:orbits}.
 For each orbital configuration, we run ten realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
 Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
-Additionally, we run the $e=0.9$ and $e=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
+Additionally, we run the $e_0=0.9$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
 
@@
 The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt13}
-\begin{equation*}
+\begin{equation}\label{eq:ahard}
     a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{h}}{4},
-\end{equation*}
+\end{equation}
 where $r_\mathrm{h}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
 Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means.
-
-The second method is a forward modelling method, where we use Bayesian analysis to construct a simple hierarchical model. 
-We construct the hierarchical model using \textsc{Stan}, which is an implementation of Hamiltonian Monte Carlo.
-Our likelihood function is constructed using the analytical hardening relations \citep{quinlan96,sesana06}:
-\begin{equation}\label{eq:quinlan_H}
-    \dv{}{t}\left( \frac{1}{a} \right) = H \frac{G\rho}{\sigma},
-\end{equation}
-and
-\begin{equation}\label{eq:quinlan_K}
-    \dv{e}{t} = -K a^{-1} \dv{a}{t}.
-\end{equation}
-In these equations, $H$ and $K$ are dimensionless constants describing the evolution rate of the semimajor axis and eccentricity of the SMBH binary, respectively, and $\rho$ and $\sigma$ are the stellar density and velocity dispersion within the binary influence radius, respectively.
-We wish to determine the uncertainty on the free parameters $H' = HG\rho/\sigma$ and $K$, in addition to the initial conditions for the differential equation system, $a_\mathrm{h}$ and $e_\mathrm{h}$, at a time $t_\mathrm{h}$.
-Let us denote the set of parameters $[H', K, a_\mathrm{h}, e_\mathrm{h}]$ as $\theta$.
-To determine the uncertainty in the parameter set $\theta$, we solve the differential equation system analytically for a time period after the SMBH binary becomes hard where the effect of GW emission can be ignored: here taken to be $t_\mathrm{end} = t(a=10\,\mathrm{pc})$.
-Defining a time translation $t' = t - t_\mathrm{h}$, \autoref{eq:quinlan_H} becomes:
-\begin{equation}\label{eq:quinlan_H2}
-    \frac{1}{a(t')} \simeq H't' + \frac{1}{a_\mathrm{h}},
-\end{equation}
-and \autoref{eq:quinlan_K} becomes:
-\begin{equation}\label{eq:quinlan_K2}
-    e(t') \simeq K \ln\left[ \frac{a_\mathrm{h}}{a(t')} \right] + e_\mathrm{h}.
-\end{equation}
-The parallel to the first method now becomes apparent: in a given set of simulations, each simulation will have its own value of $\theta$, however each realisation of $\theta$ is itself a draw from a common population distribution that we wish to understand.
-If we assume that $e_\mathrm{h}$ is normally distributed with some mean $\mu_{e_\mathrm{h}}$ and standard deviation $\sigma_{e_\mathrm{h}}$, then the distribution of $\sigma_{e_\mathrm{h}}$ is precisely the uncertainty in the standard deviation of the binary eccentricity at the time the binary becomes hard.
-The primary benefit to using this second, forward modelling approach, is that we obtain an uncertainty estimate on $\sigma_{e_\mathrm{h}}$, which is not obtainable when using the first, inverse modelling approach.
-
-A probabilistic graphical model is shown in \drafting{FIGURE}.
-
-\begin{figure*}
-    \includegraphics{orbits}
-    \caption{\drafting{Orbits of the different merger systems, colour coded according to the initial orbital eccentricity.
-Note that the duration shown for each orbit is not currently consistent: probably should cut the plot to when the binary semimajor axis reaches some set value}}
-    \label{fig:orbits}
-\end{figure*}

### 2023-05-16 papers/paper-eccentricity/sections/results.tex
Nasim plot can mask out eccentricity values
KeplerModels raise error if minimum number of groups not extracted, need to implement for other HM classes

Publishing matplotlibrc has ticks on all axes
Set up config for figure scripts in paper directory
- data files for paper figures

Nasim convergence plots for paper

Deflection angle and nasim convergence scripts can save data to a directory as pickle files

New class  that preserves the boolean state of a setting, used to ascertain if publishing style set for figures
Paper edits

--- 2023-05-12 papers/paper-eccentricity/sections/results.tex
@@
 
-\drafting{Nasim result takes the mean of the eccentricity over some time period, and then determines the standard deviation across the means of different realisations of a merger system.
-We probably need to show this too for `consistency', however we could also introduce a simple hierarchical model here that allows us to do something similar more robustly.}
+\begin{figure}
+    \includegraphics{convergence}
+    \caption{\drafting{Convergence of eccentricity scatter as a function of mass resolution for $e_0=0.90$ and $e_0=0.99$. }}
+    \label{fig:convergence}
+\end{figure}
 
@@
 \end{itemize}
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with purple markers. 
+Taking the full set of ten simulations for this set (circle markers), immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
+The peculiar values of $\sigma_e$ arises due to the assumed Gaussian distribution of the eccentricity between simulations.
+Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at each resolution which has a value of $e_\mathrm{h}<0.4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim20} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
+However, the eccentricity values beyond $2.5\sigma$ from the mean are each a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups.
+The cause of these low eccentricity values is discussed in \autoref{ssec:toy}.
 
@@
 \end{itemize}
-\begin{figure*}
-    \includegraphics{nasim_scatter_e0-990.pdf}
-    \caption{\drafting{Show a plot like this where the scatter in eccentricity is determined as the standard deviation across simulations}}
-    \label{fig:nasim_e90}
-\end{figure*}
-We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:nasim_e90}. 
-We find that the scaling of $\sigma_e$ is better described by a relation proportional to $1/N_\star$, as opposed to the $1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:convergence} with blue markers, where the marker symbols are consistent with those for the $e_0=0.90$ suite.
+We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{M_\bullet/m_\star}\propto1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
 We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
+This observation is highlighted in \autoref{fig:nasim_e_comp}.
+As the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar06}, understanding if the origin of the scatter in eccentricity is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
 
-\begin{figure*}
-    \includegraphics{orbit_pars}
-    \caption{\drafting{Orbital parameters of the different merger systems, colour coded according to the initial orbital eccentricity}}
-    \label{fig:orbitpars}
-\end{figure*}
-
-\begin{figure*}
-    \includegraphics{hyperparameter}
-    \caption{\drafting{Hyperparameters of a hierarchical model of the binary hardening from a time when the binary is hard.
-To show convergence, we could compare the the distributions of $\sigma_{e_h}$ directly between resolutions}}
-    \label{fig:hyperparameter}
-\end{figure*}
+\begin{figure}
+    \includegraphics{eccentricities.pdf}
+    \caption{\drafting{Inverse modelling of eccentricity scatter as a function of initial eccentricity for $M_\bullet/m_\star=2\times10^4$. Combine as a panel into a the convergence figure?}}
+    \label{fig:nasim_e_comp}
+\end{figure}
 
@@
     \includegraphics{deflect_e-0900}
-    \caption{\drafting{Theta-e plane for $e_0=0.90$ orbit}}
+    \caption{\drafting{Theta-e plane for $e_0=0.90$ orbit.
+Update this confetti colour scheme.
+This figure and the one for $e_0=0.99$ will be combined into a single plot overlaid on the tow model results}}
     \label{fig:deflect_e90}
@@
 which match fairly well with the central properties of the merging galaxy when
-the SMBHs are becoming bound in the $e=0.99$ merger orbit simulations.
+the SMBHs are becoming bound in the $e_0=0.99$ merger orbit simulations.
 \drafting{(Check the params)}
+
+\drafting{This right place for this section?}
+To investigate the effect of the deflection angle, and by extension the impact parameter, on the final SMBH binary eccentricity, we determine the deflection angle at each pericentre passage before the SMBHs become bound \citep{binney08}:
+\begin{equation}
+    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right)
+\end{equation}
+where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
+We set as the first `hard-scattering' angle the first angle that exceeds an orbit-dependent threshold: $\theta_{\mathrm{defl,min}}=90$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45$ for $e_0=0.90$ \drafting{need to insert degrees sign, but command not working?}.
+To quantify the eccentricity of the resulting SMBH binary, we determine when the binary becomes hard (\autoref{eq:ahard}), and calculate the median and interquartile range (IQR) over a maximum of ten orbits about the time when $t=t(a=a_\mathrm{h})$.
+These points are shown in \drafting{FIGURE}, where the IQR is generally smaller than the symbol marker used in the plot.
+
+\begin{figure}
+    \includegraphics{orbit}
+    \caption{\drafting{Orbit of a SMBH in one of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Line coloured according to time.
+Point A shows a time just before the first hard scattering, and point B a time just after.
+The circle marker indicates the final position of the BH before the BHs become bound.}}
+    \label{fig:orbit}
+\end{figure}
 

### 2023-05-17 papers/paper-eccentricity/sections/results.tex
Work on toy model scripts

@@
 and
-$e_\mathrm{s}=0.85$,
-which match fairly well with the central properties of the merging galaxy when
-the SMBHs are becoming bound in the $e_0=0.99$ merger orbit simulations.
+$e_\mathrm{s}=0.9$,
+which are similar to the central properties of the merging galaxy when
+the SMBHs are becoming bound in the $e_0=0.90$ and $e_0=0.99$ merger orbit simulations.
 \drafting{(Check the params)}
@@
 \begin{itemize}
-\item $b$-$e$ curve at a few different $v_0$ values, shows minima in $e$.
+\item $\theta$-$e$ curve at a few different $v_0$ values, shows minima in $e$.
       Also plot data points from high-e simulations, we can hopefully match the curve shape at least roughly.
-      Possibly also deflection angle plot.
-\item Plot of the minimum possible eccentricity at different initial velocities, and the corresponding $b$, deflection angles.
 \end{itemize}

### 2023-05-17 papers/paper-eccentricity/scripts/potential_plot.py

@@ -1,0 +1,26 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pickle
+
+with open('data/hernquist_merger_potential.pkl', 'rb') as f:
+    data = pickle.load(f)
+
+X, Y = data['X'], data['Y']
+pot = data['pots'][8]
+
+print(np.min(pot), np.max(pot))
+phi = np.radians(-35)
+plt.contour((np.cos(phi)*X - np.sin(phi)*Y), (np.cos(phi)*Y + np.sin(phi)*X), pot,
+            colors='tab:blue', levels=np.min(pot)+np.linspace(1.4e10,4e10,20), linestyles='-')
+
+e_spheroid = 0.8
+e2s = e_spheroid**2
+A1 = (1-e2s)/e2s*(1/(1-e2s) - 1/(2*e_spheroid)*np.log((1+e_spheroid)/(1-e_spheroid)))
+A3 = 2*(1-e2s)/e2s*(1/(2*e_spheroid)*np.log((1+e_spheroid)/(1-e_spheroid)) - 1)
+
+plt.contour(X,Y, A3*X**2 + A1*Y**2, colors='tab:orange', levels=np.linspace(0,1,10), linestyles='--')
+plt.gca().set_aspect('equal')
+plt.xlabel('x/kpc')
+plt.ylabel('y/kpc')
+
+plt.show()

### 2023-05-17 papers/paper-eccentricity/scripts/toy_model.py

@@ -1,5 +1,8 @@
 import itertools
 import multiprocessing
+import pickle
+import sys
+
 import numpy as np
 import matplotlib.pyplot as plt
 from scipy.integrate import solve_ivp
@@ -114,35 +117,186 @@ def compute_res(args):
     sys, b, v, r = args
     x0 = [r,b]
     v0 = [-v,0]
-    tmax = 100 * r/v
-    ts = np.linspace(0,tmax, 500)
+    tmax = 150 * r/v
+    ts = np.concatenate(([0], np.linspace(0.5*r/v,1.5*r/v, 50), np.linspace(1.5*r/v,tmax, 500)[1:]))
 
     x,v,t = sys.integrate(x0,v0,ts)
     return x,v,t
 
-def compute_deflection_angle(x,v,rmin):
+def compute_deflection_angle(sys,x,v):
     r = np.linalg.norm(x,axis=0)
     rdot = np.sum(x*v,axis=0)/r
-    i = np.nonzero((rdot>0)&(r>rmin))[0][0]
-    return np.pi + np.arctan2(v[1,i],v[0,i])
+    i = np.nonzero(rdot>0)[0][0]
+    E = sys.orbital_energy(x[:,i],v[:,i])
+    L = np.cross(x[:,i], v[:,i])
+    return 2*np.arctan(G*2*sys.M_BH/(L*np.sqrt(2*E)))
 
-def main_plots():
+def compute_argument_of_periapsis(sys, x, v):
+    r = np.linalg.norm(x,axis=0)
+    rdot = np.sum(x*v,axis=0)/r
+    i = np.nonzero(rdot>0)[0][0]
+    h = np.cross(x[:,i], v[:,i])
+    evec = np.cross(v[:,i], [0,0,h])[:2]/(2*G*sys.M_BH) - x[:,i]/r[i]
+    return np.arctan2(evec[1],evec[0])
+    
+
+
+def calculate_b_theta_e_curves(sys, r0, v0s, bs):
+    Bs, V0s = np.meshgrid(bs, v0s)
+    efin = []
+    deflection_angle = []
+    n_tot = np.prod(Bs.shape)
+    with multiprocessing.Pool() as pool:
+        for i, (x,v,t) in enumerate(pool.imap(compute_res,
+                                              zip(itertools.repeat(sys),
+                                                  Bs.ravel(),
+                                                  V0s.ravel(),
+                                                  itertools.repeat(r0))
[…]

### 2023-05-17 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -1,0 +1,139 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pickle
+from scipy.interpolate import griddata
+
+from figure_config import fig_path, data_path
+
+degree_format_str = '${x:.0f}\degree$'
+
+def load_data(fname):
+    with open(fname, 'rb') as f:
+        return pickle.load(f)
+
+def b_v_e_map(data):
+    plt.figure()
+    plt.pcolormesh(data['bs'], data['v0s'], data['e'], vmin=0, vmax=1, shading='gouraud')
+    plt.colorbar().set_label('$e$')
+    plt.xlabel('$b/\mathrm{pc}$')
+    plt.ylabel('$v_0/\mathrm{km\,s^{-1}}$')
+
+def b_v_th_map(data):
+    plt.figure()
+    plt.pcolormesh(data['bs'], data['v0s'], np.degrees(data['theta']), vmin=20, vmax=180, shading='gouraud')
+    plt.colorbar().set_label(r'$\theta/\degree$')
+    plt.xlabel('$b/\mathrm{pc}$')
+    plt.ylabel('$v_0/\mathrm{km\,s^{-1}}$')
+
+def th_v_e_map(data):
+    th,v = np.meshgrid(np.linspace(0,np.pi, 200), data['v0s']) 
+    _,v_data = np.meshgrid(data['bs'], data['v0s'])
+    e = griddata((data['theta'].ravel(), v_data.ravel()), data['e'].ravel(), (th, v))
+
+    plt.figure()
+    plt.pcolormesh(np.degrees(th), v, e, vmin=0, vmax=1, shading='gouraud')
+    plt.colorbar().set_label('$e$')
+    plt.xlabel(r'$\theta$')
+    plt.gca().xaxis.set_major_formatter(degree_format_str)
+    plt.ylabel('$v_0/\mathrm{km\,s^{-1}}$')
+
+def min_e_plot(data):
+    min_e_ind = np.argmin(data['e'], axis=-1)
+    fig, axes = plt.subplots(3,1, sharex='col')
+    
+    axes[0].plot(data['v0s'], data['e'][np.indices(min_e_ind.shape)[0],min_e_ind])
+    axes[0].set_ylabel(r'$e_\mathrm{min}$')
+
+    axes[1].plot(data['v0s'], np.degrees(data['theta'][np.indices(min_e_ind.shape)[0],min_e_ind]))
[…]

### 2023-05-18 papers/paper-eccentricity/paper-eccentricity.tex
More writing, flesh out some other sections

@@
 
-\textit{Software:} \ketju{} \citep{rantala17}, \gadget{} \citep{springel22}, NumPy \citep{numpy}, SciPy \citep{scipy}, Matplotlib \citep{matplotlib}, pygad \citep{rottgers20}.
+\textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2022}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
 

### 2023-05-18 papers/paper-eccentricity/sections/discuss.tex
More writing, flesh out some other sections

@@
 }
+
+We present in \autoref{sec:results} evidence for a tight correlation between the first hard-scattering angle $\theta_\mathrm{defl}$ and the hard SMBH binary eccentricity. 
+With the inclusion of the toy model, which provides a resolution-free model of the SMBH scattering process, we may draw two key interpretations of the results.
+
+First, the numerical simulations performed with \ketju{} are deterministic for a given set of initial conditions, given the same computing architecture and numerical parameters (e.g.
+the number of computing cores used).
+The scatter we observe in $\theta_\mathrm{defl}$ is thus a manifestation of Brownian motion induced by the phase space discretisation and other numerical errors, such as floating point round-off errors, denoted $\sigma_\mathrm{sim}$.
+Thus, the numerical simulations present an instance where there is zero uncertainty in the initial conditions of the system, however there is error in the numerical evolution of the system.
+
+Second, in the real Universe there is no numerical error affecting the system evolution.
+There is however uncertainty in the initial conditions of the system, whereby the chance of two galaxy merger events having the exact same initial configuration is practically zero.
+We denote this uncertainty as $\sigma_\mathrm{true}$.
+Intuitively, we expect that $\sigma_\mathrm{sim} > \sigma_\mathrm{real}$: in other words, $\sigma_\mathrm{sim}$ provides an upper estimate on the uncertainty we can expect from galaxy mergers in the real Universe.
+In particular, processes that perturb the impact parameter associated with the hard scattering of a SMBH pair will manifest as an effect in the resulting binary eccentricity.

### 2023-05-18 papers/paper-eccentricity/sections/numerical.tex
More writing, flesh out some other sections

@@
 
-\begin{itemize}
-    \item describe toy model
-    \item describe galaxy ICs: 0.5 Dehnen spheres
-    \item describe merger orbit: low and high e
-\end{itemize}
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017}, implemented in \gadget{}.
+For details about the implementation of \ketju{}, we point the reader to \drafting{Mannerkoski in prep}.
 
 \subsection{Galaxy ICs}\label{ssec:gics}
-We construct our galaxy models to match those of \citet{gualandris22}.
-Each galaxy is represented as a stellar-only Dehnen sphere \citep{dehnen93} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+We construct our galaxy models to match those of \citet{gualandris2022}.
+Each galaxy is represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
 \begin{equation}\label{eq:dehnen}
@@
 The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
-We test \drafting{X} different mass resolutions by sampling the stellar phase space distribution with varying number of stellar particles: \drafting{list resolutions here}.
+We test five different mass resolutions by sampling the stellar phase space distribution with varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
 
@@
 We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
-We follow the same orbital configurations as \citet{gualandris22}, with three additional orbit. 
+We follow the same orbital configurations as \citet{gualandris2022}, with three additional orbit. 
 The chosen eccentricities $e$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
-We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris22}.
+We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022}.
 For each orbital configuration, we run ten realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
@@
 
-The first method follows \citet{nasim20}, and is an inverse problem. 
+The first method follows \citet{nasim2020}, and is an inverse problem. 
 For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
-The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt13}
+The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2013}
 \begin{equation}\label{eq:ahard}

### 2023-05-18 papers/paper-eccentricity/sections/results.tex
More writing, flesh out some other sections

--- 2023-05-16 papers/paper-eccentricity/sections/results.tex
@@
-\section{Results}
+\section{Results}\label{sec:results}
 
-\begin{itemize}
-    \item Show that brownian motion affects impact parameter of the final passage before BHs bound --> eccentricity set
-\end{itemize}
+We present our findings first for the SMBH binary eccentricity scatter for initially low-eccentricity orbits (\autoref{ssec:lowe}), and then present the findings for initially high-eccentricity orbits (\autoref{ssec:highe}).
+We explain the results using an analytical representation in \autoref{ssec:toy}.
 
@@
 
-\subsection{Eccentricity scatter in low-e orbits}
-\begin{itemize}
-    \item Show that impact parameters not sensitive
-\end{itemize}
+\subsection{Eccentricity scatter in low-e orbits}\label{ssec:lowe}
+
 We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with purple markers. 
@@
 Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at each resolution which has a value of $e_\mathrm{h}<0.4$.
-This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim20} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
 However, the eccentricity values beyond $2.5\sigma$ from the mean are each a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups.
@@
 
-\subsection{Eccentricity scatter in high-e orbits}
-\begin{itemize}
-    \item show that impact parameters sensitive
-\end{itemize}
+\subsection{Eccentricity scatter in high-e orbits}\label{ssec:highe}
+
 We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:convergence} with blue markers, where the marker symbols are consistent with those for the $e_0=0.90$ suite.
-We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{M_\bullet/m_\star}\propto1/\sqrt{N_\star}$ relation described in \citet{nasim20}.
+We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{M_\bullet/m_\star}\propto1/\sqrt{N_\star}$ relation described in \citet{nasim2020}.
 We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
 This observation is highlighted in \autoref{fig:nasim_e_comp}.
-As the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar06}, understanding if the origin of the scatter in eccentricity is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
+As the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar2006}, understanding if the origin of the scatter in eccentricity is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
 
@@
 the equation of motion can be solved using standard numerical integration techniques.
-We use the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{scipy}
+We use the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}
 for this purpose.
@@
 \drafting{This right place for this section?}
-To investigate the effect of the deflection angle, and by extension the impact parameter, on the final SMBH binary eccentricity, we determine the deflection angle at each pericentre passage before the SMBHs become bound \citep{binney08}:
+To investigate the effect of the deflection angle, and by extension the impact parameter, on the final SMBH binary eccentricity, we determine the deflection angle at each pericentre passage before the SMBHs become bound \citep{binney2008}:
 \begin{equation}
@@
 where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
-We set as the first `hard-scattering' angle the first angle that exceeds an orbit-dependent threshold: $\theta_{\mathrm{defl,min}}=90$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45$ for $e_0=0.90$ \drafting{need to insert degrees sign, but command not working?}.
+We set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first angle that exceeds an orbit-dependent threshold: $\theta_{\mathrm{defl,min}}=90$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45$ for $e_0=0.90$ \drafting{need to insert degrees sign, but command not working?}.
+An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
 To quantify the eccentricity of the resulting SMBH binary, we determine when the binary becomes hard (\autoref{eq:ahard}), and calculate the median and interquartile range (IQR) over a maximum of ten orbits about the time when $t=t(a=a_\mathrm{h})$.
@@
     \includegraphics{orbit}
-    \caption{\drafting{Orbit of a SMBH in one of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
-Line coloured according to time.
-Point A shows a time just before the first hard scattering, and point B a time just after.
-The circle marker indicates the final position of the BH before the BHs become bound.}}
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two realisations of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Each line is coloured according to time of the simulation, however the colour for $t<22\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
+In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
+The circle marker indicates the final position of the SMBH before the SMBHs form a bound binary.}
     \label{fig:orbit}

### 2023-05-18 papers/fix_zotero.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+sed -i .bak "s/\\\\textbackslash /\\\\/g" $1
+sed -i .bak "s/\\\\aa p/\\\\aap/g" $1
+rm $1.bak

### 2023-05-19 papers/paper-eccentricity/sections/discuss.tex
Small edits to discussion

@@
 Thus, the numerical simulations present an instance where there is zero uncertainty in the initial conditions of the system, however there is error in the numerical evolution of the system.
+We determine the positional displacement between corresponding SMBHs across the simulations at a given mass resolution for the $e_0=0.90$ and $e_0-0.99$ orbits by moving the SMBH binary system to the centre of mass frame, and setting the SMBH positions to coincide with the origin at the time of the simulation.
+Computing the standard deviation in the positional displacements $\sigma_\mathrm{displ}$ between simulation $i$ and simulation $j\neq i$ for all times $t<t_\mathrm{bound}$, we find the SMBH positions vary of the order $\sim10\,\mathrm{pc}$ at the time of the hard scattering, which translates to an uncertainty in the deflection angle.
 
@@
 We denote this uncertainty as $\sigma_\mathrm{true}$.
-Intuitively, we expect that $\sigma_\mathrm{sim} > \sigma_\mathrm{real}$: in other words, $\sigma_\mathrm{sim}$ provides an upper estimate on the uncertainty we can expect from galaxy mergers in the real Universe.
+For the mergers of massive galaxies of the types modelled in this work, an uncertainty in the initial state of $\sigma_\mathrm{true} \sim \sigma_\mathrm{sim}$ is not unexpected, given the influence radius of a $10^8\,\Msun$ SMBH in a stellar background with velocity dispersion of the order $200\,\mathrm{km}\,\mathrm{s}^{-1}$ is $\sim 10\,\mathrm{pc}$ \drafting{From MBW, find better reference}.
 In particular, processes that perturb the impact parameter associated with the hard scattering of a SMBH pair will manifest as an effect in the resulting binary eccentricity.

### 2023-05-19 papers/paper-eccentricity/sections/intro.tex

\section{Introduction}

Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies. 
As galaxies dynamically interact, so too are their supermassive black holes expected to interact in a three-phase merger process.
The first phase, dynamical friction, acts to bring the SMBHs from kpc-scale separations to pc-scale separations, after which the SMBHs form a bound binary system, which can be characterised by an orbit with a semimajor axis $a$ and eccentricity $e$.
Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution, and undergoes two types of interactions: hierarchical interactions and chaotic interactions.
Hierarchical interactions occur when a perturbing third body is far from from the SMBH binary, $a_\mathrm{bin} < a_\mathrm{perturber}$, and the energy of the system does not change, however the angular momentum does.
Hierarchical interactions drive changes in the SMBH binary eccentricity. 
Conversely, chaotic interactions between the SMBH binary and a perturber remove orbital energy from the system when the semimajor axis of the perturber is comparable to the semimajor axis of the SMBH binary, $a_\mathrm{bin}\simeq a_\mathrm{perturber}$.
The angular momentum of the system also varies during the chaotic interactions.
Chaotic interactions drive the SMBH binary to sub-pc separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.

It is well known that the efficiency by which GW emission can drive a SMBH binary to coalescence is highly sensitive to the eccentricity of the binary system \citep{peters1964}, with the merger timescale of the SMBH binary being particularly sensitive. 
Additionally, the complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays (PTAs) and the upcoming Laser Interferometer Space Antenna (LISA) mission.
The requirement to understand the faithfulness to which SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the GW observational community aims to detect.

Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in simulations is an artefact of poor phase space discretisation in simulations, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.99$, stochasticity in the final SMBH binary eccentricity is present. 
This paper is organised \drafting{as follows: X, X, X.}
\drafting{Add relevant references.}

### 2023-05-20 papers/paper-eccentricity/sections/numerical.tex
Small edits to discussion

@@
 In our analysis, we wish to determine the population spread in eccentricity from a finite set of observations.
-We assess the uncertainty in eccentricity using two methods: an inverse modelling approach and a forward modelling approach.
-
-The first method follows \citet{nasim2020}, and is an inverse problem. 
+We assess the uncertainty in eccentricity using an inverse modelling approach following \citet{nasim2020}. 
 For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.

### 2023-05-22 papers/paper-eccentricity/sections/discuss.tex

--- 2023-05-18 papers/paper-eccentricity/sections/discuss.tex
@@
 Thus, the numerical simulations present an instance where there is zero uncertainty in the initial conditions of the system, however there is error in the numerical evolution of the system.
+We determine the positional displacement between corresponding SMBHs across the simulations at a given mass resolution for the $e_0=0.90$ and $e_0=0.99$ orbits by moving the SMBH binary system to the centre of mass frame, and setting the SMBH positions to coincide with the origin at the beginning of the simulation.
+Computing the standard deviation in the positional displacements for corresponding SMBHs $\sigma_\mathrm{displ}$ between simulation $i$ and simulation $j\neq i$ for all times $t<t_\mathrm{bound}$, we find the SMBH positions vary of the order $\sim10\,\mathrm{pc}$ at the time of the hard scattering, which translates to an uncertainty in the deflection angle.
 
@@
 We denote this uncertainty as $\sigma_\mathrm{true}$.
-Intuitively, we expect that $\sigma_\mathrm{sim} > \sigma_\mathrm{real}$: in other words, $\sigma_\mathrm{sim}$ provides an upper estimate on the uncertainty we can expect from galaxy mergers in the real Universe.
-In particular, processes that perturb the impact parameter associated with the hard scattering of a SMBH pair will manifest as an effect in the resulting binary eccentricity.
+For the mergers of massive galaxies of the types modelled in this work, an uncertainty in the initial state of $\sigma_\mathrm{true} \sim \sigma_\mathrm{sim}$ is not unexpected, given the influence radius of a $10^8\,\Msun$ SMBH in a stellar background with velocity dispersion of the order $200\,\mathrm{km}\,\mathrm{s}^{-1}$ is $\sim 10\,\mathrm{pc}$ \drafting{From MBW, find better reference}.
+
+Thus, any process that perturbs the impact parameter associated with the hard scattering angle will result in an eccentricity sampled from the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve, subject to some small scatter in $e$ resulting from the hierarchical interactions of distant perturbers, or by exchanges in angular momentum with the stellar cluster immediately surrounding the SMBH.
+Critically however, the expected value of eccentricity is a direct consequence of the hard scattering angle, and is not an independent random process, and can be reliably predicted.
+We caution that there is not necessarily a one-to-one mapping of the initial merger eccentricity $e_0$ and the hard binary eccentricity $e_\mathrm{h}$, unless the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve is uniform.
+As such a uniform function is not associated with the radial ($e_0>0.99$) mergers expected to occur in reality, we expect that for a given initial merger orbit $e_0$ a range of possible $e_\mathrm{h}$ would be observed.
+The converse also holds: inference about merging SMBH binary eccentricity does not necessarily provide a mapping back to the initial merger eccentricity unless the history of the system is known, which is unable to be achieved in a practical sense in reality.

### 2023-05-22 papers/paper-eccentricity/sections/results.tex
Add toy model figures with sim data

@@
 
-\begin{figure}
-    \includegraphics{deflect_e-0900}
-    \caption{\drafting{Theta-e plane for $e_0=0.90$ orbit.
-Update this confetti colour scheme.
-This figure and the one for $e_0=0.99$ will be combined into a single plot overlaid on the tow model results}}
-    \label{fig:deflect_e90}
-\end{figure}
-
-\begin{figure}
-    \includegraphics{deflect_e-0990}
-    \caption{\drafting{Theta-e plane for $e_0=0.99$ orbit}}
-    \label{fig:deflect_e99}
-\end{figure}
+\begin{figure*}
+    \includegraphics{theta_e_sim_and_model_090}
+    \includegraphics{theta_e_sim_and_model_099}
+    \caption{
+    Scattering deflection angle $\theta$ and the resulting binary eccentricity
+    in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
+    compared to reasonably well fitting model curves (see \autoref{ssec:toy}).
+    The model curve for the $e_0=0.90$ case has been shifted left by $12\degr$
+    to better match the simulation data.
+    }
+    \label{fig:theta_e}
+\end{figure*}
 
@@
 \end{gather}
-We fix $r_0 = 100\,\mathrm{pc}$,
+We fix $r_0 = 25\,\mathrm{pc}$,
 and use model parameters
-$\rho = 400 \,\Msun\,\mathrm{pc}^{-3}$,
+$\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
 $\sigma=200\,\mathrm{km\,s^{-1}}$,
+and
 $M_\bullet=10^8\,\Msun$,
-and
-$e_\mathrm{s}=0.9$,
-which are similar to the central properties of the merging galaxy when
+which are fairly similar to the central properties of the merging galaxy when
 the SMBHs are becoming bound in the $e_0=0.90$ and $e_0=0.99$ merger orbit simulations.
-\drafting{(Check the params)}
+Based on the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.90$ case,
+and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
+The spheroid eccentricity parameter is not easily measurable from the simulations,
+but by testing different values we find that $e_\mathrm{s}=0.905$ gives behaviour
+that closely resembles the simulation results.
+We compute the resulting eccentricities for a range of impact parameters for the two initial velocities.
 
-\drafting{This right place for this section?}
-To investigate the effect of the deflection angle, and by extension the impact parameter, on the final SMBH binary eccentricity, we determine the deflection angle at each pericentre passage before the SMBHs become bound \citep{binney2008}:
+To compare the model against the simulations,
+we use the deflection angle in the first pericentre passage instead of the impact parameter,
+as it can be more robustly measured from the simulation data.
+The deflection angle is given by \citep{binney2008}:
 \begin{equation}
@@
 \end{equation}
-where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
-We set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first angle that exceeds an orbit-dependent threshold: $\theta_{\mathrm{defl,min}}=90$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45$ for $e_0=0.90$ \drafting{need to insert degrees sign, but command not working?}.
+where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH 
+binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
+To measure this angle from the simulations,
+we set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first 
+angle that exceeds an orbit-dependent threshold:
+$\theta_{\mathrm{defl,min}}=90\degr$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45\degr$ for $e_0=0.90$.
+This corresponds to the second or third actual pericentre passage in the simulations.
+In the model calculations we simply use the first actual pericentre for this calculation.
+
+The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
+To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$.
+Otherwise the model appears to give a very good match to the simulation results,
+and the required slight shift could be due to the galactic potential being slightly tilted compared to the
+BH trajectories due to the complex dynamics at this stage.
+In the $e_0=0.99$ case we do not shift the model curve, although here a shift of a few degrees might give a better
+fit to the location of the minimum at $\theta \approx 100\degr$.
+In this case we can also see that the simulation data has more scatter around
+the model curve in the $\theta \gtrsim 120\degr$ region,
+although the model does appear to capture the mean behaviour fairly well.
+The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
+which might be related to the large scatter in the simulation data.
+
+\drafting{(Move these bits to the right place)}
 An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
@@
 \end{figure}
-
-\drafting{
-Figures to do:
-\begin{itemize}
-\item $\theta$-$e$ curve at a few different $v_0$ values, shows minima in $e$.
-      Also plot data points from high-e simulations, we can hopefully match the curve shape at least roughly.
-\end{itemize}
-}
 

### 2023-05-22 papers/paper-eccentricity/scripts/toy_model.py

@@ -230,17 +230,17 @@ def parameter_space_scan_g05_conf2():
     
 def parameter_space_scan_g05_conf3():
 # params matching gamma=0.5 e=0.99 runs fairly well
-    for gal_e, df_fudge in itertools.product([0.85,0.9],[0.3]):
+    for gal_e, df_fudge in itertools.product([0.9],[0.5]):
         sys = System(M_BH=1e-2,
-                     rho=20,
+                     rho=30,
                      e_spheroid=gal_e,
-                     stellar_sigma=225*km_per_s,
+                     stellar_sigma=200*km_per_s,
                      df_fudge_factor=df_fudge)
 
-        r0 = sys.r_infl
+        r0 = 0.025
 
         v0s = np.linspace(400, 650, 40) * km_per_s
-        bs = np.linspace(0.01, 25, 250) * 1e-3
+        bs = np.linspace(0.01, 25, 200) * 1e-3
         
         theta, efin = calculate_b_theta_e_curves(sys, r0, v0s, bs)
         
@@ -250,24 +250,33 @@ def parameter_space_scan_g05_conf3():
         with open(f'data/g05_3_b_v_scan_es_{gal_e:.2f}_df_{df_fudge:.1f}.pkl', 'wb') as f:
             pickle.dump(res, f, protocol=-1)
 
+# main plot data generated with this function
 def high_res_well_fitting_models():
-    #e=0.9 mergers are well fitted by this system, and the parameters are pretty similar to the sims as well
+    #e=0.9, 0.99 mergers are well described by this system, and the parameters are pretty similar to the sims as well
     sys = System(M_BH=1e-2,
-                 rho=40,
-                 e_spheroid=0.9,
+                 rho=30,
+                 e_spheroid=0.905,
                  stellar_sigma=200*km_per_s,
                  df_fudge_factor=0.5)
 
-    r0 = sys.r_infl
+    r0 = 25e-3
 
-    v0s = [470 * km_per_s]
+    v0s = [450 * km_per_s] # vel from simulations
     bs = np.geomspace(0.1, 20, 500) * 1e-3
-    
-    theta, efin = calculate_b_theta_e_curves(sys, r0, v0s, bs)
-    res090 = dict(theta=theta[0], e=efin[0], b=bs)
 
[…]

### 2023-05-22 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -1,3 +1,4 @@
+import itertools
 import numpy as np
 import matplotlib
 import matplotlib.pyplot as plt
@@ -5,6 +6,7 @@ import pickle
 from scipy.interpolate import griddata
 
 from figure_config import fig_path, data_path
+from Plotter import Plotter
 
 degree_format_str = '${x:.0f}\degree$'
 
@@ -68,11 +70,11 @@ def th_e_curves(data):
     plt.xlabel(r'$\theta$')
     plt.gca().xaxis.set_major_formatter(degree_format_str)
     
-def plot_specific_th_e(data, v0s, simdata=None):
+def plot_specific_th_e(data, v0s, simdata=None, shift=0):
     plt.figure()
     for v0 in v0s:
         i = np.argmin(abs(np.array(data['v0s'])-v0))
-        plt.plot(np.degrees(data['theta'][i]), data['e'][i], label=f"$v_0 = {data['v0s'][i]:3.0f}"r'/\mathrm{km\,s^{-1}}$')
+        plt.plot(np.degrees(data['theta'][i]) + shift, data['e'][i], label=f"$v_0 = {data['v0s'][i]:3.0f}"r'/\mathrm{km\,s^{-1}}$')
 
 
     if simdata is not None:
@@ -106,13 +108,70 @@ def hr_well_fitting_models_plot(curve_shift=None):
 
 
     sim_data = load_data(data_path('deflection_angles_e0-0.900.pickle')) 
-    plot_sim_data_th_e(plt.gca(), {'1':sim_data090})
+    plot_sim_data_th_e(plt.gca(), {'1':sim_data})
+
+    plt.figure()
+    plt.ylabel('$e$')
+    plt.xlabel(r'$\theta$')
+    plt.gca().xaxis.set_major_formatter(degree_format_str)
+
+    d = load_data(data_path('well_fitting_e_0.99_model_curve.pkl'))
+    l, = plt.plot(np.degrees(d['theta'])+curve_shift, d['e'], color='tab:orange')
+    plt.plot(np.degrees(d['theta']), d['e'], alpha=0.5, color=l.get_color())
+
+    sim_data = load_data(data_path('deflection_angles_e0-0.990.pickle')) 
+    plot_sim_data_th_e(plt.gca(), {'1':sim_data})
+
+def paper_plots():
+    plotter = Plotter()
+
+    plt.figure()
+    plt.ylabel('$e$')
[…]

### 2023-05-22 papers/paper-eccentricity/paper-eccentricity.tex
Update paper to ApJL format
- move section text to main .tex document instead of separated .tex files

@@
-% mnras_template.tex 
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
+%% Based on file 'sample631.tex'
+%%
+%% Modified 2021 March
+%%
+%% This is a sample manuscript marked up using the
+%% AASTeX v6.31 LaTeX 2e macros.
+%%
+%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
+%% (Copyright 2000-2015).  See the classfile for details.
 
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
+%% AASTeX requires revtex4-1.cls and other external packages such as
+%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
+%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
+%% AASTeX v6+ still uses v4.1. All of these external packages should 
+%% already be present in the modern TeX distributions but not always.
+%% For example, revtex4.1 seems to be missing in the linux version of
+%% TexLive 2020. One should be able to get all packages from www.ctan.org.
+%% In particular, revtex v4.1 can be found at 
+%% https://www.ctan.org/pkg/revtex4-1.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
-\documentclass[fleqn,usenatbib]{mnras}
+%% The first piece of markup in an AASTeX v6.x document is the \documentclass
+%% command. LaTeX will ignore any data that comes before this command. The 
+%% documentclass can take an optional argument to modify the output style.
+%% The command below calls the preprint style which will produce a tightly 
+%% typeset, one-column, single-spaced document.  It is the default and thus
+%% does not need to be explicitly stated.
+%%
+%% using aastex version 6.3
+\documentclass[twocolumn]{aastex631}
 
-% MNRAS is set in Times font. If you don't have this installed (most LaTeX
-% installations will be fine) or prefer the old Computer Modern fonts, comment
-% out the following line
-\usepackage{newtxtext,newtxmath}
-% Depending on your LaTeX fonts installation, you might get better results with one of these:
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
+%% The default is a single spaced, 10 point font, single spaced article.
+%% There are 5 other style options available via an optional argument. They
+%% can be invoked like this:
+%%
+%% \documentclass[arguments]{aastex631}
+%% 
+%% where the layout options are:
+%%
+%%  twocolumn   : two text columns, 10 point font, single spaced article.
+%%                This is the most compact and represent the final published
+%%                derived PDF copy of the accepted manuscript from the publisher
+%%  manuscript  : one text column, 12 point font, double spaced article.
+%%  preprint    : one text column, 12 point font, single spaced article.  
+%%  preprint2   : two text columns, 12 point font, single spaced article.
+%%  modern      : a stylish, single text column, 12 point font, article with
+%% 		  wider left and right margins. This uses the Daniel
+%% 		  Foreman-Mackey and David Hogg design.
+%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
+%%                but now that abstracts are required this is obsolete for
+%%                AAS Journals. Authors might need it for other reasons. DO NOT
+%%                use \begin{abstract} and \end{abstract} with this style.
+%%
+%% Note that you can submit to the AAS Journals in any of these 6 styles.
+%%
+%% There are other optional arguments one can invoke to allow other stylistic
+%% actions. The available options are:
+%%
+%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
+%%   tighten      : Makes baselineskip slightly smaller, only works with 
+%%                  the twocolumn substyle.
+%%   times        : uses times font instead of the default
+%%   linenumbers  : turn on lineno package.
+%%   trackchanges : required to see the revision mark up and print its output
+%%   longauthor   : Do not use the more compressed footnote style (default) for 
+%%                  the author/collaboration/affiliations. Instead print all
+%%                  affiliation information after each name. Creates a much 
+%%                  longer author list but may be desirable for short 
+%%                  author papers.
+%% twocolappendix : make 2 column appendix.
+%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
+%%                  for dual anonymous review.
+%%
+%% these can be used in any combination, e.g.
+%%
+%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
+%%
+%% AASTeX v6.* now includes \hyperref support. While we have built in specific
+%% defaults into the classfile you can manually override them with the
+%% \hypersetup command. For example,
+%%
+%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
+%%
+%% will change the color of the internal links to red, the links to the
+%% bibliography to green, the file links to cyan, and the external links to
+%% magenta. Additional information on \hyperref options can be found here:
+%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
+%%
+%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
+%% to improve the accessibility of the compiled pdf file.
+%%
+%% If you want to create your own macros, you can do so
+%% using \newcommand. Your macros should appear before
+%% the \begin{document} command.
+%%
+\usepackage{amsmath}
+\usepackage{amssymb}
 
-% Use vector fonts, so it zooms properly in on-screen viewing software
-% Don't change these lines unless you know what you are doing
-\usepackage[T1]{fontenc}
+\newcommand{\vdag}{(v)^\dagger}
+\newcommand\aastex{AAS\TeX}
+\newcommand\latex{La\TeX}
 
-% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
-% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
-\DeclareRobustCommand{\VAN}[3]{#2}
-\let\VANthebibliography\thebibliography
-\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}
-
-%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
-
-% Only include extra packages if you really need them. Common packages are:
-\usepackage{graphicx}	% Including figure files
-\usepackage{amsmath}	% Advanced maths commands
-%\usepackage{amssymb}	% Extra maths symbols
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
 \newcommand{\ketju}{\textsc{Ketju}}
@@
 
-\graphicspath{{figures/}}
-
-% XXX: For highlighting draft notes, these should be removed when done
-\usepackage{color}
+% drafting macros
 \newcommand{\drafting}[1]{
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Reintroduced the \received and \accepted commands from AASTeX v5.2
+%\received{March 1, 2021}
+%\revised{April 1, 2021}
+%\accepted{\today}
 
-%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+%% Command to document which AAS Journal the manuscript was submitted to.
+%% Adds "Submitted to " the argument.
+%\submitjournal{PSJ}
 
-% Title of the paper, and the short title which is used in the headers.
-% Keep the title short and informative.
-\title[Reviving Stochasticity]{Reviving stochasticity: SMBH-binary eccentricity is inherently random}
+%% For manuscript that include authors in collaborations, AASTeX v6.31
+%% builds on the \collaboration command to allow greater freedom to 
+%% keep the traditional author+affiliation information but only show
+%% subsets. The \collaboration command now must appear AFTER the group
+%% of authors in the collaboration and it takes TWO arguments. The last
+%% is still the collaboration identifier. The text given in this
+%% argument is what will be shown in the manuscript. The first argument
+%% is the number of author above the \collaboration command to show with
+%% the collaboration text. If there are authors that are not part of any
+%% collaboration the \nocollaboration command is used. This command takes
+%% one argument which is also the number of authors above to show. A
+%% dashed line is shown to indicate no collaboration. This example manuscript
+%% shows how these commands work to display specific set of authors 
+%% on the front page.
+%%
+%% For manuscript without any need to use \collaboration the 
+%% \AuthorCollaborationLimit command from v6.2 can still be used to 
+%% show a subset of authors.
+%
+%\AuthorCollaborationLimit=2
+%
+%% will only show Schwarz & Muench on the front page of the manuscript
+%% (assuming the \collaboration and \nocollaboration commands are
+%% commented out).
+%%
+%% Note that all of the author will be shown in the published article.
+%% This feature is meant to be used prior to acceptance to make the
+%% front end of a long author article more manageable. Please do not use
+%% this functionality for manuscripts with less than 20 authors. Conversely,
+%% please do use this when the number of authors exceeds 40.
+%%
+%% Use \allauthors at the manuscript end to show the full author list.
+%% This command should only be used with \AuthorCollaborationLimit is used.
 
-% The list of authors, and the short list which is used in the headers.
-% If you need two or more lines of authors, add an extra line using \newauthor
-\author[A. Rawlings et al.]{
-Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
-Matias Mannerkoski,$^{1}$
-Others,$^{1}$
-\\
-% List of institutions
-$^{1}$
-Department of Physics,
-Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland
+%% The following command can be used to set the latex table counters.  It
+%% is needed in this document because it uses a mix of latex tabular and
+%% AASTeX deluxetables.  In general it should not be needed.
+%\setcounter{table}{1}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%
+%% The following section outlines numerous optional output that
+%% can be displayed in the front matter or as running meta-data.
+%%
+%% If you wish, you may supply running head information, although
+%% this information may be modified by the editorial offices.
+\shorttitle{Reviving Stochasticity}
+\shortauthors{A. Rawlings et al.}
+%%
+%% You can add a light gray and diagonal water-mark to the first page 
+%% with this command:
+%% \watermark{text}
+%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
+%% long you can control the water-mark size with:
+%% \setwatermarkfontsize{dimension}
+%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
+%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\graphicspath{{./}{figures/}}
+%% This is the end of the preamble.  Indicate the beginning of the
+%% manuscript itself with \begin{document}.
+
+\begin{document}
+
+\title{Reviving stochasticity: SMBH-binary eccentricity is inherently random}
+
+%% LaTeX will automatically break titles if they run longer than
+%% one line. However, you may use \\ to force a line break if
+%% you desire. In v6.31 you can include a footnote in the title.
+
+%% A significant change from earlier AASTEX versions is in the structure for 
+%% calling author and affiliations. The change was necessary to implement 
+%% auto-indexing of affiliations which prior was a manual process that could 
+%% easily be tedious in large author manuscripts.
+%%
+%% The \author command is the same as before except it now takes an optional
+%% argument which is the 16 digit ORCID. The syntax is:
+%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
+%%
+%% This will hyperlink the author name to the author's ORCID page. Note that
+%% during compilation, LaTeX will do some limited checking of the format of
+%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
+%% present or in the LaTeX pathway, the OrcID icon will appear next to
+%% the authors name.
+%%
+%% Use \affiliation for affiliation information. The old \affil is now aliased
+%% to \affiliation. AASTeX v6.31 will automatically index these in the header.
+%% When a duplicate is found its index will be the same as its previous entry.
+%%
+%% Note that \altaffilmark and \altaffiltext have been removed and thus 
+%% can not be used to document secondary affiliations. If they are used latex
+%% will issue a specific error message and quit. Please use multiple 
+%% \affiliation calls for to document more than one affiliation.
+%%
+%% The new \altaffiliation can be used to indicate some secondary information
+%% such as fellowships. This command produces a non-numeric footnote that is
+%% set away from the numeric \affiliation footnotes.  NOTE that if an
+%% \altaffiliation command is used it must come BEFORE the \affiliation call,
+%% right after the \author command, in order to place the footnotes in
+%% the proper location.
+%%
+%% Use \email to set provide email addresses. Each \email will appear on its
+%% own line so you can put multiple email address in one \email call. A new
+%% \correspondingauthor command is available in V6.31 to identify the
+%% corresponding author of the manuscript. It is the author's responsibility
+%% to make sure this name is also in the author list.
+%%
+%% While authors can be grouped inside the same \author and \affiliation
+%% commands it is better to have a single author for each. This allows for
+%% one to exploit all the new benefits and should make book-keeping easier.
+%%
+%% If done correctly the peer review system will be able to
+%% automatically put the author and affiliation information from the manuscript
+%% and save the corresponding author the trouble of entering it by hand.
+
+%\correspondingauthor{August Muench}
+%\email{greg.schwarz@aas.org, gus.muench@aas.org}
+
+\author[0000-0003-1807-6321]{Alexander Rawlings}
+\affiliation{Department of Physics,
+Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland}
+
+\author[0000-0001-5721-9335]{Matias Mannerkoski}
+\affiliation{Department of Physics,
+Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland}
+
+\author{Others}
+
+\correspondingauthor{Alexander Rawlings}
+\email{alexander.rawlings@helsinki.fi}
+
+%% Note that the \and command from previous versions of AASTeX is now
+%% depreciated in this version as it is no longer necessary. AASTeX 
+%% automatically takes care of all commas and "and"s between authors names.
+
+%% AASTeX 6.31 has the new \collaboration and \nocollaboration commands to
+%% provide the collaboration status of a group of authors. These commands 
+%% can be used either before or after the list of corresponding authors. The
+%% argument for \collaboration is the collaboration identifier. Authors are
+%% encouraged to surround collaboration identifiers with ()s. The 
+%% \nocollaboration command takes no argument and exists to indicate that
+%% the nearby authors are not part of surrounding collaborations.
+
+%% Mark off the abstract in the ``abstract'' environment. 
+\begin{abstract}
+
+We present some stuff about eccentricity convergence in Nbody sims. Limited to 250 words.
+
+\end{abstract}
+
+%% Keywords should appear after the \end{abstract} command. 
+%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
+%% https://astrothesaurus.org
+%% You will be asked to selected these concepts during the submission process
+%% but this old "keyword" functionality is maintained in case authors want
+%% to include these concepts in their preprints.
+\keywords{Edit keywords}
+
+%% From the front matter, we move on to the body of the paper.
+%% Sections are demarcated by \section and \subsection, respectively.
+%% Observe the use of the LaTeX \label
+%% command after the \subsection to give a symbolic KEY to the
+%% subsection for cross-referencing in a \ref command.
+%% You can use LaTeX's \ref and \label commands to keep track of
+%% cross-references to sections, equations, tables, and figures.
+%% That way, if you change the order of any elements, LaTeX will
+%% automatically renumber them.
+%%
+%% We recommend that authors also use the natbib \citep
+%% and \citet commands to identify citations.  The citations are
+%% tied to the reference list via symbolic KEYs. The KEY corresponds
+%% to the KEY in the \bibitem in the reference list below.
+
+%-------------------------- MAIN TEXT STARTS HERE --------------------------%
+% INTRODUCTION
+\section{Introduction}
+Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies. 
+As galaxies dynamically interact, so too are their supermassive black holes expected to interact in a three-phase merger process.
+The first phase, dynamical friction, acts to bring the SMBHs from kpc-scale separations to pc-scale separations, after which the SMBHs form a bound binary system, which can be characterised by an orbit with a semimajor axis $a$ and eccentricity $e$.
+Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution, and undergoes two types of interactions: hierarchical interactions and chaotic interactions.
+Hierarchical interactions occur when a perturbing third body is far from from the SMBH binary, $a_\mathrm{bin} < a_\mathrm{perturber}$, and the energy of the system does not change, however the angular momentum does.
+Hierarchical interactions drive changes in the SMBH binary eccentricity. 
+Conversely, chaotic interactions between the SMBH binary and a perturber remove orbital energy from the system when the semimajor axis of the perturber is comparable to the semimajor axis of the SMBH binary, $a_\mathrm{bin}\simeq a_\mathrm{perturber}$.
+The angular momentum of the system also varies during the chaotic interactions.
+Chaotic interactions drive the SMBH binary to sub-pc separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
+
+It is well known that the efficiency by which GW emission can drive a SMBH binary to coalescence is highly sensitive to the eccentricity of the binary system \citep{peters1964}, with the merger timescale of the SMBH binary being particularly sensitive.
+Additionally, the complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays (PTAs) and the upcoming Laser Interferometer Space Antenna (LISA) mission.
+The requirement to understand the faithfulness to which SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the GW observational community aims to detect.
+
+Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in simulations is an artefact of poor phase space discretisation in simulations, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
+In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
+Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.99$, stochasticity in the final SMBH binary eccentricity is present. 
+This paper is organised \drafting{as follows: X, X, X.}
+\drafting{Add relevant references.}
+
+% NUMERICAL
+\section{Numerical Methods}
+
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017}, implemented in \gadget{}.
+For details about the implementation of \ketju{}, we point the reader to \drafting{Mannerkoski in prep}.
+
+\subsection{Galaxy ICs}\label{ssec:gics}
+We construct our galaxy models to match those of \citet{gualandris2022}.
+Each galaxy is represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+\begin{equation}\label{eq:dehnen}
+    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
+\end{equation}
+The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
+We test five different mass resolutions by sampling the stellar phase space distribution with varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
+
+\subsection{Merger ICs}
+We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
+We follow the same orbital configurations as \citet{gualandris2022}, with three additional orbit. 
+The chosen eccentricities $e$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
+We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022}.
+For each orbital configuration, we run ten realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
+Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
+Additionally, we run the $e_0=0.9$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
+
+\subsection{Assessing Eccentricity Uncertainty}
+In our analysis, we wish to determine the population spread in eccentricity from a finite set of observations.
+We assess the uncertainty in eccentricity using an inverse modelling approach following \citet{nasim2020}. 
+For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
+The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2013}
+\begin{equation}\label{eq:ahard}
+    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{h}}{4},
+\end{equation}
+where $r_\mathrm{h}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
+Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means.
+
+% RESULTS
+\section{Results}\label{sec:results}
+
+We present our findings first for the SMBH binary eccentricity scatter for initially low-eccentricity orbits (\autoref{ssec:lowe}), and then present the findings for initially high-eccentricity orbits (\autoref{ssec:highe}).
+We explain the results using an analytical representation in \autoref{ssec:toy}.
+
+\begin{figure}
+    \includegraphics{convergence}
+    \caption{\drafting{Convergence of eccentricity scatter as a function of mass resolution for $e_0=0.90$ and $e_0=0.99$. }}
+    \label{fig:convergence}
+\end{figure}
+
+\subsection{Eccentricity scatter in low-e orbits}\label{ssec:lowe}
+
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with purple markers. 
+Taking the full set of ten simulations for this set (circle markers), immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
+The peculiar values of $\sigma_e$ arises due to the assumed Gaussian distribution of the eccentricity between simulations.
+Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at each resolution which has a value of $e_\mathrm{h}<0.4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
+However, the eccentricity values beyond $2.5\sigma$ from the mean are each a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups.
+The cause of these low eccentricity values is discussed in \autoref{ssec:toy}.
+
+\subsection{Eccentricity scatter in high-e orbits}\label{ssec:highe}
+
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:convergence} with blue markers, where the marker symbols are consistent with those for the $e_0=0.90$ suite.
+We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{M_\bullet/m_\star}\propto1/\sqrt{N_\star}$ relation described in \citet{nasim2020}.
+We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
+This observation is highlighted in \autoref{fig:nasim_e_comp}.
+As the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar2006}, understanding if the origin of the scatter in eccentricity is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
+
+\begin{figure}
+    \includegraphics{eccentricities.pdf}
+    \caption{\drafting{Inverse modelling of eccentricity scatter as a function of initial eccentricity for $M_\bullet/m_\star=2\times10^4$. Combine as a panel into a the convergence figure?}}
+    \label{fig:nasim_e_comp}
+\end{figure}
+
+\begin{figure*}
+    \includegraphics{theta_e_sim_and_model_090}
+    \includegraphics{theta_e_sim_and_model_099}
+    \caption{
+    Scattering deflection angle $\theta$ and the resulting binary eccentricity
+    in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
+    compared to reasonably well fitting model curves (see \autoref{ssec:toy}).
+    The model curve for the $e_0=0.90$ case has been shifted left by $12\degr$
+    to better match the simulation data.
+    }
+    \label{fig:theta_e}
+\end{figure*}
+
+\subsection{Explaining the eccentricity scatter with a toy model}\label{ssec:toy}
+
+We can gain insight into the large binary eccentricity scatter observed in the
+runs with high galactic merger orbit eccentricities by considering a simplified
+model that captures the essential features of how the SMBH binary becomes bound
+and settles on an orbit with a given eccentricity.
+The components we include in this model are the two SMBHs,
+a smooth background potential, and dynamical friction.
+For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
+to lie in the $xz$-plane.
+In this case the relative motion of the BHs can be treated as a one-body problem,
+with the equation of motion
+\begin{equation} \label{eq:toy_model_eom}
+\ddot{\vb{x}} = -\frac{2 G M_\bullet}{|\vb{x}|^3} \vb{x} + \vb{a}_\mathrm{bg} + \vb{a}_\mathrm{DF},
+\end{equation}
+where $\vb{x}$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
+$\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
+and $\vb{a}_\mathrm{DF}$ the acceleration due to dynamical friction.
+
+The background galaxy is modelled as a with constant density spheroidal potential,
+causing the acceleration \drafting{(e.g.\ B\&T)}
+\begin{equation}
+\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y,A_z) \vb{x},
+\end{equation}
+where $\rho$ is the stellar density,
+and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
+\begin{equation}
+A_x = A_y  = 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]
+\end{equation}
+\begin{equation}
+A_z = \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+        \left[ \frac{1}{1-e_\mathrm{s}^2} 
+        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
+\end{equation}
+With eccentricity values of $e_\mathrm{s} \approx 0.8\textnormal{--}0.9$,
+this spheroidal potential matches fairly well with the central potential in the galaxy during the
+BH binary binding phase, apart from the contributions from the BHs themselves and the groups of stars bound to them.
+\drafting{(Figure showing this?)}
+
+For dynamical friction, we use the \drafting{Chandrasekhar ref} formula assuming
+a Maxwellian distribution with a constant velocity dispersion $\sigma$.
+This formula gives the acceleration of a single BH due to dynamical friction as
+\begin{equation}\label{eq:chandra_df}
+\vb{a}_{1,\mathrm{DF}} = - 4 \pi G^2 M_\bullet \rho \ln{\Lambda} 
+                        (\operatorname{erf}(X) - 2 \pi^{-1/2} X e^{-X^2}) \frac{\vb{v}_{1}}{|\vb{v}_{1}|^3},
+\end{equation}
+where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
+$X = |\vb{v}_1|/\sqrt{2}\sigma$,
+and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 10$.
+To account for differences between the constant density background of stars
+assumed here and the more complex system that is present in the simulations
+we additionally multiply the acceleration given by Eq. \eqref{eq:chandra_df}
+with a scaling and cut-off function
+\begin{equation}
+f(a) = \frac{\alpha}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]},
+\end{equation}
+which serves to remove the effect of dynamical friction when the BH binary has become
+hard in addition to allowing the strength of dynamical friction to be scaled to
+achieve similar behaviour as observed in the simulations.
+Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f=\alpha$.
+We set the scaling factor $\alpha=0.5$, and use cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
+and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
+with $a_\mathrm{h}$ being the semimajor-axis of a hard BH binary.
+The total dynamical friction term is then
+\begin{equation}
+\vb{a}_{\mathrm{DF}} = 2 f(a) \vb{a}_{1,\mathrm{DF}}.
+\end{equation}
+
+After specifying the initial position and velocity, as well as the other parameters of the model,
+the equation of motion can be solved using standard numerical integration techniques.
+We use the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}
+for this purpose.
+To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
+we specify the initial conditions confined to the $xz$-plane as 
+\begin{gather}
+\vb{x}_0 = (r_0, b)\\
+\dot{\vb{x}}_0 = (-v_0, 0).
+\end{gather}
+We fix $r_0 = 25\,\mathrm{pc}$,
+and use model parameters
+$\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
+$\sigma=200\,\mathrm{km\,s^{-1}}$,
+and
+$M_\bullet=10^8\,\Msun$,
+which are fairly similar to the central properties of the merging galaxy when
+the SMBHs are becoming bound in the $e_0=0.90$ and $e_0=0.99$ merger orbit simulations.
+Based on the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.90$ case,
+and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
+The spheroid eccentricity parameter is not easily measurable from the simulations,
+but by testing different values we find that $e_\mathrm{s}=0.905$ gives behaviour
+that closely resembles the simulation results.
+We compute the resulting eccentricities for a range of impact parameters for the two initial velocities.
+
+To compare the model against the simulations,
+we use the deflection angle in the first pericentre passage instead of the impact parameter,
+as it can be more robustly measured from the simulation data.
+The deflection angle is given by \citep{binney2008}:
+\begin{equation}
+    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right)
+\end{equation}
+where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH 
+binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
+To measure this angle from the simulations,
+we set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first 
+angle that exceeds an orbit-dependent threshold:
+$\theta_{\mathrm{defl,min}}=90\degr$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45\degr$ for $e_0=0.90$.
+This corresponds to the second or third actual pericentre passage in the simulations.
+In the model calculations we simply use the first actual pericentre for this calculation.
+
+The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
+To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$.
+Otherwise the model appears to give a very good match to the simulation results,
+and the required slight shift could be due to the galactic potential being slightly tilted compared to the
+BH trajectories due to the complex dynamics at this stage.
+In the $e_0=0.99$ case we do not shift the model curve, although here a shift of a few degrees might give a better
+fit to the location of the minimum at $\theta \approx 100\degr$.
+In this case we can also see that the simulation data has more scatter around
+the model curve in the $\theta \gtrsim 120\degr$ region,
+although the model does appear to capture the mean behaviour fairly well.
+The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
+which might be related to the large scatter in the simulation data.
+
+\drafting{(Move these bits to the right place)}
+An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
+To quantify the eccentricity of the resulting SMBH binary, we determine when the binary becomes hard (\autoref{eq:ahard}), and calculate the median and interquartile range (IQR) over a maximum of ten orbits about the time when $t=t(a=a_\mathrm{h})$.
+These points are shown in \drafting{FIGURE}, where the IQR is generally smaller than the symbol marker used in the plot.
+
+\begin{figure}
+    \includegraphics{orbit}
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two realisations of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Each line is coloured according to time of the simulation, however the colour for $t<22\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
+In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
+The circle marker indicates the final position of the SMBH before the SMBHs form a bound binary.}
+    \label{fig:orbit}
+\end{figure}
+
+\drafting{
+Findings
+\begin{itemize}
+\item Final $e$ varies with impact parameter for given $v_0$, two minima on small scales at $b$s that are likely to occur in simulations. 
+\item location and depth of minima varies with $v_0$, for some values possible to span the full range of $e$ for different impact params.
+\item Minimum $e$ for deflection angles $\sim 60^\circ, 120^\circ$.
+\end{itemize}
 }
 
-% These dates will be filled out by the publisher
-\date{Accepted XXX. Received YYY; in original form ZZZ}
+\drafting{
+Interpretation:
+\begin{itemize}
+\item Realistic merger orbits generally result in linearly plunging orbits just before binary formation,
+        with fairly small impact parameters.
+\item Often the relative velocities are in the range that leads to deflections around $90^\circ$
+      (some reason for this from energy considerations, since $b_{90}=GM_\mathrm{tot}/v^2$?),
+      small changes in impact param around that point hit the eccentricity minima in this model.
+\item Impact param differences seeded by the Brownian motion in sims,
+      but few pc differences small enough that you can't measure them from large scale merger orbits anyway.
+\item Conclusion: the scatter is physically relevant, not pure numerics.
+\end{itemize}
+}
 
-% Enter the current year, for the copyright statements etc.
-\pubyear{2022}
+% DISCUSSION
+\section{Discussion}
 
-% Don't change these lines
-\begin{document}
-\label{firstpage}
-\pagerange{\pageref{firstpage}--\pageref{lastpage}}
-\maketitle
+\drafting{
+    An interesting point to discuss is the origin of our eccentricity scatter.
+Early tests of the Gualandris setup seems to indicate that low eccentricity mergers do not demonstrate orbital flips, and are more converged.
+How to distinguish if the flip is the cause of the eccentricity scatter, or just correlated with eccentricity scatter?
+}
 
-% Abstract of the paper
-\begin{abstract}
-We present some stuff about eccentricity convergence in Nbody sims
-\end{abstract}
+We present in \autoref{sec:results} evidence for a tight correlation between the first hard-scattering angle $\theta_\mathrm{defl}$ and the hard SMBH binary eccentricity. 
+With the inclusion of the toy model, which provides a resolution-free model of the SMBH scattering process, we may draw two key interpretations of the results.
 
-% Select between one and six entries from the list of approved keywords.
-% Don't make up new ones.
-\begin{keywords}
-keyword1 -- keyword2 -- keyword3
-\end{keywords}
+First, the numerical simulations performed with \ketju{} are deterministic for a given set of initial conditions, given the same computing architecture and numerical parameters (e.g.
+the number of computing cores used).
+The scatter we observe in $\theta_\mathrm{defl}$ is thus a manifestation of Brownian motion induced by the phase space discretisation and other numerical errors, such as floating point round-off errors, denoted $\sigma_\mathrm{sim}$.
+Thus, the numerical simulations present an instance where there is zero uncertainty in the initial conditions of the system, however there is error in the numerical evolution of the system.
+We determine the positional displacement between corresponding SMBHs across the simulations at a given mass resolution for the $e_0=0.90$ and $e_0=0.99$ orbits by moving the SMBH binary system to the centre of mass frame, and setting the SMBH positions to coincide with the origin at the beginning of the simulation.
+Computing the standard deviation in the positional displacements for corresponding SMBHs $\sigma_\mathrm{displ}$ between simulation $i$ and simulation $j\neq i$ for all times $t<t_\mathrm{bound}$, we find the SMBH positions vary of the order $\sim10\,\mathrm{pc}$ at the time of the hard scattering, which translates to an uncertainty in the deflection angle.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+Second, in the real Universe there is no numerical error affecting the system evolution.
+There is however uncertainty in the initial conditions of the system, whereby the chance of two galaxy merger events having the exact same initial configuration is practically zero.
+We denote this uncertainty as $\sigma_\mathrm{true}$.
+For the mergers of massive galaxies of the types modelled in this work, an uncertainty in the initial state of $\sigma_\mathrm{true} \sim \sigma_\mathrm{sim}$ is not unexpected, given the influence radius of a $10^8\,\Msun$ SMBH in a stellar background with velocity dispersion of the order $200\,\mathrm{km}\,\mathrm{s}^{-1}$ is $\sim 10\,\mathrm{pc}$ \drafting{From MBW, find better reference}.
 
-%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+Thus, any process that perturbs the impact parameter associated with the hard scattering angle will result in an eccentricity sampled from the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve, subject to some small scatter in $e$ resulting from the hierarchical interactions of distant perturbers, or by exchanges in angular momentum with the stellar cluster immediately surrounding the SMBH.
+Critically however, the expected value of eccentricity is a direct consequence of the hard scattering angle, and is not an independent random process, and can be reliably predicted.
+We caution that there is not necessarily a one-to-one mapping of the initial merger eccentricity $e_0$ and the hard binary eccentricity $e_\mathrm{h}$, unless the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve is uniform.
+As such a uniform function is not associated with the radial ($e_0>0.99$) mergers expected to occur in reality, we expect that for a given initial merger orbit $e_0$ a range of possible $e_\mathrm{h}$ would be observed.
+The converse also holds: inference about merging SMBH binary eccentricity does not necessarily provide a mapping back to the initial merger eccentricity unless the history of the system is known, which is unable to be achieved in a practical sense in reality.
 
-\input{./sections/intro}
+% CONCLUSIONS
+\section{Conclusions}
 
-\input{./sections/numerical}
-
-\input{./sections/results}
-
-\input{./sections/discuss}
-
-\input{./sections/conclusions}
-
+% FINAL BITS
 \textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2022}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Data Availability}
+%% For this sample we use BibTeX plus aasjournals.bst to generate the
+%% the bibliography. The sample631.bib file was populated from ADS. To
+%% get the citations to show in the compiled file do the following:
+%%
+%% pdflatex sample631.tex
+%% bibtext sample631
+%% pdflatex sample631.tex
+%% pdflatex sample631.tex
 
-\drafting{The inclusion of a Data Availability Statement is a requirement for articles published in MNRAS.
-Data Availability Statements provide a standardised format for readers to understand the availability of data underlying the research results described in the article.
-The statement may refer to original data generated in the course of the study or to third-party data analysed in the article.
-The statement should describe and provide means of access, where possible, by linking to the data or providing the required accession numbers for the relevant databases or DOIs.}
+\bibliography{../ref}{}
+\bibliographystyle{aasjournal}
 
-%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
+%% This command is needed to show the entire author+affiliation list when
+%% the collaboration and author truncation commands are used.  It has to
+%% go at the end of the manuscript.
+%\allauthors
 
-% The best way to enter references is to use BibTeX:
+%% Include this line if you are using the \added, \replaced, \deleted
+%% commands to see a summary list of all changes at the end of the article.
+%\listofchanges
 
-\bibliographystyle{mnras}
-\bibliography{../ref} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-%\begin{thebibliography}{99}
-%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-%Others S., 2012, Journal of Interesting Stuff, 17, 198
-%\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
-
-\appendix
-
-%\input{sections/appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% Don't change these lines
-\bsp	% typesetting comment
-\label{lastpage}
 \end{document}
 
-% End of mnras_template.tex
+% End of file `sample631.tex'.

### 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
Update fig 4, move some sim related text to a new subsection

@@
 
+\subsection{Binary binding process and the scatter in eccentricity}
+
+\drafting{
+Explain the deflection angle and eccentricity stuff seen in the simulations here,
+then the next section just focuses on the toy model.
+}
+
+\drafting{(These paragraphs just copypasted, were mixed in the toy model discussion. Rewrite as appropriate)}
+
+An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
+To quantify the eccentricity of the resulting SMBH binary, we determine when the binary becomes hard (\autoref{eq:ahard}), and calculate the median and interquartile range (IQR) over a maximum of ten orbits about the time when $t=t(a=a_\mathrm{h})$.
+These points are shown in \drafting{FIGURE}, where the IQR is generally smaller than the symbol marker used in the plot.
+
+We use the deflection angle in the first pericentre passage instead of the impact parameter,
+as it can be more robustly measured from the simulation data.
+The deflection angle is given by \citep{binney2008}:
+\begin{equation}
+    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right)
+\end{equation}
+where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH 
+binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
+To measure this angle from the simulations,
+we set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first 
+angle that exceeds an orbit-dependent threshold:
+$\theta_{\mathrm{defl,min}}=90\degr$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45\degr$ for $e_0=0.90$.
+This corresponds to the second or third actual pericentre passage in the simulations.
+
+\begin{figure}
+    \includegraphics{orbit}
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two realisations of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Each line is coloured according to time of the simulation, however the colour for $t<22\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
+In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
+The circle marker indicates the final position of the SMBH before the SMBHs form a bound binary.}
+    \label{fig:orbit}
+\end{figure}
+
 \begin{figure*}
-    \includegraphics{theta_e_sim_and_model_090}
-    \includegraphics{theta_e_sim_and_model_099}
+    \centering
+    \includegraphics{theta_e_sim_and_model}
     \caption{
@@
     in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
-    compared to reasonably well fitting model curves (see \autoref{ssec:toy}).
-    The model curve for the $e_0=0.90$ case has been shifted left by $12\degr$
-    to better match the simulation data.
+    compared to reasonably well fitting model curves.
+    The solid lines show the results from the toy model with $e_\mathrm{s}=0.905$,
+    while the semitransparent lines show the results for  $e_\mathrm{s}=0.900$ and $e_\mathrm{s}=0.910$.
+    The model curves have been shifted left by $12\degr$ and $6\degr$
+    to better match the simulation data, with the shift also indicated by arrows.
     }
@@
 
-\subsection{Explaining the eccentricity scatter with a toy model}\label{ssec:toy}
+\subsection{Reproducing the eccentricity scatter with a toy model}\label{ssec:toy}
 
@@
 \end{equation}
-With eccentricity values of $e_\mathrm{s} \approx 0.8\textnormal{--}0.9$,
-this spheroidal potential matches fairly well with the central potential in the galaxy during the
-BH binary binding phase, apart from the contributions from the BHs themselves and the groups of stars bound to them.
-\drafting{(Figure showing this?)}
+With eccentricity values of $e_\mathrm{s} \approx 0.9$,
+this spheroidal potential matches the central potential in the galaxy during the
+BH binary binding phase reasonably well,
+apart from the contributions from the BHs themselves and the groups of stars bound to them.
 
@@
 that closely resembles the simulation results.
-We compute the resulting eccentricities for a range of impact parameters for the two initial velocities.
-
-To compare the model against the simulations,
-we use the deflection angle in the first pericentre passage instead of the impact parameter,
-as it can be more robustly measured from the simulation data.
-The deflection angle is given by \citep{binney2008}:
-\begin{equation}
-    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right)
-\end{equation}
-where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH 
-binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
-To measure this angle from the simulations,
-we set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first 
-angle that exceeds an orbit-dependent threshold:
-$\theta_{\mathrm{defl,min}}=90\degr$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45\degr$ for $e_0=0.90$.
-This corresponds to the second or third actual pericentre passage in the simulations.
-In the model calculations we simply use the first actual pericentre for this calculation.
+We compute the resulting eccentricities for a range of impact parameters up to $b=20\mathrm{pc}$
+for the two initial velocities.
+In the model, a $90\degr$ deflection during the first pericenter passage
+corresponds to initial impact parameters of $b\approx 5.8\,\mathrm{pc}$ and
+$b\approx 3.3\,\mathrm{pc}$ for the $e_0=0.90$ and $e_0=0.99$ cases, respectively.
 
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
-To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$.
+In addition to the reference $e_\mathrm{s}=0.905$ model, the figure also shows the model results for
+$e_\mathrm{s}=0.900$ and $e_\mathrm{s}=0.910$, as particularly the $e_0=0.99$ case is quite sensitive to this parameter.
+To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
+while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
 Otherwise the model appears to give a very good match to the simulation results,
-and the required slight shift could be due to the galactic potential being slightly tilted compared to the
-BH trajectories due to the complex dynamics at this stage.
-In the $e_0=0.99$ case we do not shift the model curve, although here a shift of a few degrees might give a better
-fit to the location of the minimum at $\theta \approx 100\degr$.
-In this case we can also see that the simulation data has more scatter around
+and the required slight shift could be due to the galactic potential being slightly tilted in the simulations 
+compared to the BH trajectories due to the complex dynamics at this stage.
+In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
 the model curve in the $\theta \gtrsim 120\degr$ region,
@@
 which might be related to the large scatter in the simulation data.
-
-\drafting{(Move these bits to the right place)}
-An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
-To quantify the eccentricity of the resulting SMBH binary, we determine when the binary becomes hard (\autoref{eq:ahard}), and calculate the median and interquartile range (IQR) over a maximum of ten orbits about the time when $t=t(a=a_\mathrm{h})$.
-These points are shown in \drafting{FIGURE}, where the IQR is generally smaller than the symbol marker used in the plot.
-
-\begin{figure}
-    \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two realisations of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
-Each line is coloured according to time of the simulation, however the colour for $t<22\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
-In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
-The circle marker indicates the final position of the SMBH before the SMBHs form a bound binary.}
-    \label{fig:orbit}
-\end{figure}
-
-\drafting{
-Findings
-\begin{itemize}
-\item Final $e$ varies with impact parameter for given $v_0$, two minima on small scales at $b$s that are likely to occur in simulations. 
-\item location and depth of minima varies with $v_0$, for some values possible to span the full range of $e$ for different impact params.
-\item Minimum $e$ for deflection angles $\sim 60^\circ, 120^\circ$.
-\end{itemize}
-}
-
-\drafting{
-Interpretation:
-\begin{itemize}
-\item Realistic merger orbits generally result in linearly plunging orbits just before binary formation,
-        with fairly small impact parameters.
-\item Often the relative velocities are in the range that leads to deflections around $90^\circ$
-      (some reason for this from energy considerations, since $b_{90}=GM_\mathrm{tot}/v^2$?),
-      small changes in impact param around that point hit the eccentricity minima in this model.
-\item Impact param differences seeded by the Brownian motion in sims,
-      but few pc differences small enough that you can't measure them from large scale merger orbits anyway.
-\item Conclusion: the scatter is physically relevant, not pure numerics.
-\end{itemize}
-}
+Overall we can however see that this simple model correctly captures the behaviour of eccentricity seen in the simulations.
 

### 2023-05-23 papers/paper-eccentricity/scripts/toy_model.py
Update fig 4, move some sim related text to a new subsection

@@ -252,29 +252,31 @@ def parameter_space_scan_g05_conf3():
 
 # main plot data generated with this function
 def high_res_well_fitting_models():
-    #e=0.9, 0.99 mergers are well described by this system, and the parameters are pretty similar to the sims as well
-    sys = System(M_BH=1e-2,
-                 rho=30,
-                 e_spheroid=0.905,
-                 stellar_sigma=200*km_per_s,
-                 df_fudge_factor=0.5)
-
     r0 = 25e-3
-
-    v0s = [450 * km_per_s] # vel from simulations
     bs = np.geomspace(0.1, 20, 500) * 1e-3
+    e_spheroids = [0.895, 0.9, 0.905, 0.91]
 
-    theta, efin = calculate_b_theta_e_curves(sys, r0, v0s, bs)
+    v0s = [450 * km_per_s, 560 * km_per_s]
+    res090 = []
+    res099 = []
+    for e_s in e_spheroids:
+    #e=0.9, 0.99 mergers are well fitted by this system, and the parameters are pretty similar to the sims as well
+        sys = System(M_BH=1e-2,
+                     rho=30,
+                     e_spheroid=e_s,
+                     stellar_sigma=200*km_per_s,
+                     df_fudge_factor=0.5)
 
-    res090 = dict(theta=theta[0], e=efin[0], b=bs)
+        theta, efin = calculate_b_theta_e_curves(sys, r0, v0s, bs)
+        res090.append(dict(theta=theta[0], e=efin[0], b=bs))
+        res099.append(dict(theta=theta[1], e=efin[1], b=bs))
+
     with open(f'data/well_fitting_e_0.90_model_curve.pkl', 'wb') as f:
-        pickle.dump(res090, f, protocol=-1)
+        pickle.dump(dict(e_spheroids=e_spheroids, res=res090), f, protocol=-1)
 
-    v0s = [560 * km_per_s]
-    theta, efin = calculate_b_theta_e_curves(sys, r0, v0s, bs)
-    res099 = dict(theta=theta[0], e=efin[0], b=bs)
     with open(f'data/well_fitting_e_0.99_model_curve.pkl', 'wb') as f:
-        pickle.dump(res099, f, protocol=-1)
+        pickle.dump(dict(e_spheroids=e_spheroids, res=res099), f, protocol=-1)
+
 
 
 
@@ -288,15 +290,15 @@ def test_plots():
 # the individual free parameters:
[…]

### 2023-05-23 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -3,12 +3,12 @@ import numpy as np
 import matplotlib
 import matplotlib.pyplot as plt
 import pickle
-from scipy.interpolate import griddata
+from scipy.interpolate import griddata, interp1d
 
 from figure_config import fig_path, data_path
 from Plotter import Plotter
 
-degree_format_str = '${x:.0f}\degree$'
+degree_format_str = '{x:.0f}°'
 
 def load_data(fname):
     with open(fname, 'rb') as f:
@@ -95,47 +95,43 @@ def plot_sim_data_th_e(ax, simdata):
                    )
 
 
-def hr_well_fitting_models_plot(curve_shift=None):
-
-    plt.figure()
-    plt.ylabel('$e$')
-    plt.xlabel(r'$\theta$')
-    plt.gca().xaxis.set_major_formatter(degree_format_str)
-
-    d = load_data(data_path('well_fitting_e_0.90_model_curve.pkl'))
-    l, = plt.plot(np.degrees(d['theta'])+curve_shift, d['e'], color='tab:orange')
-    plt.plot(np.degrees(d['theta']), d['e'], alpha=0.5, color=l.get_color())
-
-
-    sim_data = load_data(data_path('deflection_angles_e0-0.900.pickle')) 
-    plot_sim_data_th_e(plt.gca(), {'1':sim_data})
-
-    plt.figure()
-    plt.ylabel('$e$')
-    plt.xlabel(r'$\theta$')
-    plt.gca().xaxis.set_major_formatter(degree_format_str)
-
-    d = load_data(data_path('well_fitting_e_0.99_model_curve.pkl'))
-    l, = plt.plot(np.degrees(d['theta'])+curve_shift, d['e'], color='tab:orange')
-    plt.plot(np.degrees(d['theta']), d['e'], alpha=0.5, color=l.get_color())
-
-    sim_data = load_data(data_path('deflection_angles_e0-0.990.pickle')) 
-    plot_sim_data_th_e(plt.gca(), {'1':sim_data})
-
 def paper_plots():
     plotter = Plotter()
 
-    plt.figure()
-    plt.ylabel('$e$')
-    plt.xlabel(r'$\theta$')
-    plt.gca().xaxis.set_major_formatter(degree_format_str)
-    plt.ylim(0,1)
-    plt.xlim(20,165)
+    fig, axes = plt.subplots(1,2,sharey='row')
+    fig.set_figwidth(2*fig.get_figwidth())
 
[…]

### 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
Text edits

@@
 Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution, and undergoes two types of interactions: hierarchical interactions and chaotic interactions.
-Hierarchical interactions occur when a perturbing third body is far from from the SMBH binary, $a_\mathrm{bin} < a_\mathrm{perturber}$, and the energy of the system does not change, however the angular momentum does.
+Hierarchical interactions occur when a perturbing third body, or a collection of bodies, is far from the SMBH binary, $a_\mathrm{bin} < a_\mathrm{perturber}$, and the energy of the system does not change, however the angular momentum does.
 Hierarchical interactions drive changes in the SMBH binary eccentricity. 
@@
 Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.99$, stochasticity in the final SMBH binary eccentricity is present. 
-This paper is organised \drafting{as follows: X, X, X.}
 \drafting{Add relevant references.}
@@
 
-We present our findings first for the SMBH binary eccentricity scatter for initially low-eccentricity orbits (\autoref{ssec:lowe}), and then present the findings for initially high-eccentricity orbits (\autoref{ssec:highe}).
-We explain the results using an analytical representation in \autoref{ssec:toy}.
+\begin{figure*}
+    \includegraphics{eccentricities.pdf}
+    \caption{Equivalent eccentricity $1-e$ as a function of time $t'$ for a reprsentative set of simulations with $e_0=0.90$ (left) and $e_0 = 0.90$ (right).
+\drafting{Shift data so $t'$ is defined such that $t'=0$ when $t=t_h$}.
+The dashed line indicates the collective mean of the equivalent eccentricities averaged over the span of the respective line.
+The grey area corresponds to $1\sigma$ scatter in the data \drafting{change to $2.5\sigma$}.
+Note the asymmetry in the grey area is due to the logarithmic scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.}
+    \label{fig:eccentricities}
+\end{figure*}
 
@@
 
-We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with purple markers. 
-Taking the full set of ten simulations for this set (circle markers), immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
-The peculiar values of $\sigma_e$ arises due to the assumed Gaussian distribution of the eccentricity between simulations.
-Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at each resolution which has a value of $e_\mathrm{h}<0.4$.
-This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
+An representative set of realisations for the $e_0=0.90$ orbit is shown in the left panel of \autoref{fig:eccentricities}.
+We define a shifted time $t'$ such that $t'(0) = t(a=a_\mathrm{h})$, and the data is truncated to a maximum of five orbits centred about $t'(0)$. 
+One simulation is seen to lie beyond $2.5\sigma$ of the mean $1-e$ line.
+
+We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with circle and square markers. 
+Taking for each mass resolution the full set of ten simulations (circle markers), immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
+The peculiar values of $\sigma_e$ arise due to the assumed Gaussian distribution of the eccentricity between simulations.
+Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at both $5\times10^3$ and $1\times10^4$ resolutions which has a value of $e_\mathrm{h}<0.4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity (i.e., taking those lines within the grey region of the left panel of \autoref{fig:eccentricities}) we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
 However, the eccentricity values beyond $2.5\sigma$ from the mean are each a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups.
@@
 
-We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit in \autoref{fig:convergence} with blue markers, where the marker symbols are consistent with those for the $e_0=0.90$ suite.
+Similar to the $e_0=0.90$ case, we show a representative set of realisation for the $e_0=0.99$ orbit in the right panel of \autoref{fig:eccentricities}.
+The dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit is shown in \autoref{fig:convergence} with triangle and diamond markers. 
 We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{M_\bullet/m_\star}\propto1/\sqrt{N_\star}$ relation described in \citet{nasim2020}.
 We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
-This observation is highlighted in \autoref{fig:nasim_e_comp}.
 As the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar2006}, understanding if the origin of the scatter in eccentricity is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
-
-\begin{figure}
-    \includegraphics{eccentricities.pdf}
-    \caption{\drafting{Inverse modelling of eccentricity scatter as a function of initial eccentricity for $M_\bullet/m_\star=2\times10^4$. Combine as a panel into a the convergence figure?}}
-    \label{fig:nasim_e_comp}
-\end{figure}
 
@@
 
-An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
-To quantify the eccentricity of the resulting SMBH binary, we determine when the binary becomes hard (\autoref{eq:ahard}), and calculate the median and interquartile range (IQR) over a maximum of ten orbits about the time when $t=t(a=a_\mathrm{h})$.
-These points are shown in \drafting{FIGURE}, where the IQR is generally smaller than the symbol marker used in the plot.
-
-We use the deflection angle in the first pericentre passage instead of the impact parameter,
-as it can be more robustly measured from the simulation data.
-The deflection angle is given by \citep{binney2008}:
+To understand the observed scatter in eccentricity, we investigate the binary binding process in each simulation. 
+Whilst the two SMBHs are unbound, they undergo a number of hyperbolic orbits influenced by the galactic merger potential.
+At each pericentre passage, the two SMBHs scatter off each other with some deflection angle $\theta_\mathrm{defl}$, defined:
 \begin{equation}
@@
 where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH 
-binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage.
-To measure this angle from the simulations,
-we set as the first `hard-scattering' angle $\theta_\mathrm{defl}$ the first 
-angle that exceeds an orbit-dependent threshold:
+binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage \citep{binney2008}.
+The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory, however is able to be more robustly measured than $b$ from the simulation data.
+We set as the first `hard-scattering' angle the first 
+angle $\theta_\mathrm{defl}$ that exceeds an orbit-dependent threshold:
 $\theta_{\mathrm{defl,min}}=90\degr$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45\degr$ for $e_0=0.90$.
-This corresponds to the second or third actual pericentre passage in the simulations.
+In general, this corresponds to the second or third actual pericentre passage of the SMBHs in the simulations.
+An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
+We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, which we aim to explain using a simple toy model.
 
@@
 The background galaxy is modelled as a with constant density spheroidal potential,
-causing the acceleration \drafting{(e.g.\ B\&T)}
+causing the acceleration \citep[e.g.][]{binney2008}
 \begin{equation}
@@
 % DISCUSSION
-\section{Discussion}
+\section{Discussion and Conclusions}
 
@@
 The converse also holds: inference about merging SMBH binary eccentricity does not necessarily provide a mapping back to the initial merger eccentricity unless the history of the system is known, which is unable to be achieved in a practical sense in reality.
-
-% CONCLUSIONS
-\section{Conclusions}
 

### 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
Some edits to the toy model description

--- 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
@@
 \usepackage{amssymb}
+\usepackage{bm} % italic bold math symbols
 
@@
 \newcommand{\ketju}{\textsc{Ketju}}
-\newcommand{\mstar}{\textsc{Mstar}}
+\newcommand{\mstar}{\textsc{MSTAR}}
 \newcommand{\gadget}{\textsc{Gadget-4}}
@@
 \newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}
-\newcommand{\vb}[1]{\ensuremath{\mathbf{#1}}}
+\newcommand{\vb}[1]{\ensuremath{\bm{#1}}} %vectors
 
@@
 
-\subsection{Reproducing the eccentricity scatter with a toy model}\label{ssec:toy}
+\subsection{Reproducing the eccentricity behavior with a toy model}\label{ssec:toy}
 
-We can gain insight into the large binary eccentricity scatter observed in the
-runs with high galactic merger orbit eccentricities by considering a simplified
-model that captures the essential features of how the SMBH binary becomes bound
-and settles on an orbit with a given eccentricity.
-The components we include in this model are the two SMBHs,
-a smooth background potential, and dynamical friction.
+To gain insight into the origin of the dependence of binary eccentricity on the 
+deflection angle $\theta$ during the binary formation,
+we have constructed a simple toy model that can qualitatively reproduce the observed
+behavior from the simulations.
+This model includes only the essential components relevant to this process,
+which are the asymmetric background stellar potential,
+dynamical friction from the stellar background,
+and naturally the SMBHs themselves.
+
+\drafting{(The technical details of this toy model might be moved to an appendix if space is an issue.)}
+
 For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
-to lie in the $xz$-plane.
+to lie in the $xy$-plane.
 In this case the relative motion of the BHs can be treated as a one-body problem,
@@
 \end{equation}
-where $\vb{x}$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
+where $\vb{x}=(x,y)$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
 $\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
@@
 
-The background galaxy is modelled as a with constant density spheroidal potential,
-causing the acceleration \drafting{(e.g.\ B\&T)}
+The background galaxy is modeled as a constant density spheroidal potential,
+causing the acceleration \citep[e.g.][]{binney2008}
 \begin{equation}
-\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y,A_z) \vb{x},
+\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y) \vb{x},
 \end{equation}
@@
 \begin{equation}
-A_x = A_y  = 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
-    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]
+A_x = 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right],
 \end{equation}
+and
 \begin{equation}
-A_z = \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+A_y = \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
         \left[ \frac{1}{1-e_\mathrm{s}^2} 
@@
 \end{equation}
+Here we have chosen the long axis of the spheroid to be aligned with the $x$-axis.
 With eccentricity values of $e_\mathrm{s} \approx 0.9$,
 this spheroidal potential matches the central potential in the galaxy during the
-BH binary binding phase reasonably well,
+SMBH binary binding phase reasonably well,
 apart from the contributions from the BHs themselves and the groups of stars bound to them.
 
-For dynamical friction, we use the \drafting{Chandrasekhar ref} formula assuming
-a Maxwellian distribution with a constant velocity dispersion $\sigma$.
+For dynamical friction, we use the \citet{1943ApJ....97..255C} formula assuming
+a Maxwellian distribution with a constant velocity dispersion $\sigma$ \citep[e.g.][]{binney2008}.
 This formula gives the acceleration of a single BH due to dynamical friction as
@@
 assumed here and the more complex system that is present in the simulations
-we additionally multiply the acceleration given by Eq. \eqref{eq:chandra_df}
+we additionally multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
 with a scaling and cut-off function
@@
 \end{equation}
-which serves to remove the effect of dynamical friction when the BH binary has become
+which serves to remove the effect of dynamical friction when the SMBH binary has become
 hard in addition to allowing the strength of dynamical friction to be scaled to
-achieve similar behaviour as observed in the simulations.
+achieve similar behavior as observed in the simulations, since in the simulations the
+stellar background is much more complex.
 Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f=\alpha$.
@@
 
-After specifying the initial position and velocity, as well as the other parameters of the model,
-the equation of motion can be solved using standard numerical integration techniques.
-We use the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}
-for this purpose.
 To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
-we specify the initial conditions confined to the $xz$-plane as 
+we specify the initial conditions confined to the $xy$-plane as 
 \begin{gather}
@@
 The spheroid eccentricity parameter is not easily measurable from the simulations,
-but by testing different values we find that $e_\mathrm{s}=0.905$ gives behaviour
-that closely resembles the simulation results.
-We compute the resulting eccentricities for a range of impact parameters up to $b=20\mathrm{pc}$
-for the two initial velocities.
-In the model, a $90\degr$ deflection during the first pericenter passage
+and we therefore perform calculations with different values of $e_\mathrm{s}$.
+We compute the resulting eccentricities for a range of impact parameters up to $b=20\,\mathrm{pc}$
+for the two initial velocities,
+by solving the equation of motion \eqref{eq:toy_model_eom} until the binary has become hard using
+the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}.
+The range of impact parameters covers most of the relevant parameter space,
+as a $90\degr$ deflection during the first pericenter passage
 corresponds to initial impact parameters of $b\approx 5.8\,\mathrm{pc}$ and
@@
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
-In addition to the reference $e_\mathrm{s}=0.905$ model, the figure also shows the model results for
-$e_\mathrm{s}=0.900$ and $e_\mathrm{s}=0.910$, as particularly the $e_0=0.99$ case is quite sensitive to this parameter.
 To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
@@
 the model curve in the $\theta \gtrsim 120\degr$ region,
-although the model does appear to capture the mean behaviour fairly well.
-The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
+although the model does appear to capture the mean behavior fairly well.
+The behavior of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
 which might be related to the large scatter in the simulation data.
-Overall we can however see that this simple model correctly captures the behaviour of eccentricity seen in the simulations.
+Overall we can however see that this simple model correctly captures the behavior of eccentricity seen in the simulations.
+On the other hand, if the model is run with low values of $e_\mathrm{s}$,
+the binary always reaches extremely high eccentricities with little dependence on the deflection angle,
+indicating that the asymmetry of the galactic potential during the SMBH binary formation is the cause for the observed eccentricity behavior.
 

### 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
Update eccentricity, orbit plots in paper

--- 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
@@
     \includegraphics{eccentricities.pdf}
-    \caption{Equivalent eccentricity $1-e$ as a function of time $t'$ for a reprsentative set of simulations with $e_0=0.90$ (left) and $e_0 = 0.90$ (right).
-\drafting{Shift data so $t'$ is defined such that $t'=0$ when $t=t_h$}.
-The dashed line indicates the collective mean of the equivalent eccentricities averaged over the span of the respective line.
-The grey area corresponds to $1\sigma$ scatter in the data \drafting{change to $2.5\sigma$}.
-Note the asymmetry in the grey area is due to the logarithmic scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.}
+    \caption{Top panel shows the equivalent \drafting{(better term?
+Reciprocal?)} eccentricity $1-e$ as a function of time $t$ for select representative simulations from the $e_0 = 0.90$ set (solid lines) and $e_0 = 0.99$ set (dashed lines).
+Time is shown on a logarithmic scale to highlight the variability in eccentricity prior to the SMBH binary becoming bound (vertical black lines).
+The equivalent eccentricity for a full set of a given mass resolution is shown in the bottom left panel for the $e_0 = 0.90$ suite and in the bottom right panel for the $e_0 = 0.99$ suite as a function of $t' = t-t(a=a_\mathrm{h})$ for five orbits centred about $t'=0$.
+The dashed line in the bottom panels indicates the collective mean of the equivalent eccentricities averaged over the span of the respective line.
+The blue area corresponds to the $2.5\sigma$ scatter in the data.
+Note the asymmetry in the grey area is due to the logarithmic scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.
+A zoom inset in the bottom right figure shows the tight bunching of $1-e$ for this particular simulation set.
+In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $1-e$ greater than $2.5\sigma$ from the collective mean.}
     \label{fig:eccentricities}
@@
 
-An representative set of realisations for the $e_0=0.90$ orbit is shown in the left panel of \autoref{fig:eccentricities}.
+An representative set of realisations for the $e_0=0.90$ orbit is shown in the bottom left panel of \autoref{fig:eccentricities}.
 We define a shifted time $t'$ such that $t'(0) = t(a=a_\mathrm{h})$, and the data is truncated to a maximum of five orbits centred about $t'(0)$. 
@@
 
-Similar to the $e_0=0.90$ case, we show a representative set of realisation for the $e_0=0.99$ orbit in the right panel of \autoref{fig:eccentricities}.
+Similar to the $e_0=0.90$ case, we show a representative set of realisation for the $e_0=0.99$ orbit in the bottom right panel of \autoref{fig:eccentricities}.
 The dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit is shown in \autoref{fig:convergence} with triangle and diamond markers. 
@@
 \subsection{Binary binding process and the scatter in eccentricity}
-
-\drafting{
-Explain the deflection angle and eccentricity stuff seen in the simulations here,
-then the next section just focuses on the toy model.
-}
-
-\drafting{(These paragraphs just copypasted, were mixed in the toy model discussion. Rewrite as appropriate)}
 
@@
 An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
-We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, which we aim to explain using a simple toy model.
+We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, which we explain using a simple toy model.
 
-\begin{figure}
-    \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two realisations of the $e_0=0.99$ mergers, which by symmetry of the equal mass system is the same as the second BH orbit reflected.
-Each line is coloured according to time of the simulation, however the colour for $t<22\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.49\textwidth]{orbit-e90}
+    \includegraphics[width=0.49\textwidth]{orbit-e99}
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Each line is coloured according to time of the simulation, however the colour for $t<31\,\mathrm{Myr}$ ($t<22\,\mathrm{Myr}$) is constant for visual clarity of the later evolution in the left (right) panel.
 In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
@@
     \label{fig:orbit}
-\end{figure}
+\end{figure*}
 
@@
     \caption{
-    Scattering deflection angle $\theta$ and the resulting binary eccentricity
+    Scattering deflection angle $\theta$ \drafting{update to $\theta_\mathrm{defl}$} and the resulting binary eccentricity
     in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,

### 2023-05-24 papers/paper-eccentricity/paper-eccentricity.tex
Some tweaks to equations

--- 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
-\drafting{(The technical details of this toy model might be moved to an appendix if space is an issue.)}
-
 For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
@@
 and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
-\begin{equation}
-A_x = 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
-    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right],
-\end{equation}
-and
-\begin{equation}
-A_y = \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+\begin{align}
+A_x &= 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]\\
+A_y &= \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
         \left[ \frac{1}{1-e_\mathrm{s}^2} 
         - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
-\end{equation}
+\end{align}
 Here we have chosen the long axis of the spheroid to be aligned with the $x$-axis.
@@
 \begin{equation}\label{eq:chandra_df}
-\vb{a}_{1,\mathrm{DF}} = - 4 \pi G^2 M_\bullet \rho \ln{\Lambda} 
-                        (\operatorname{erf}(X) - 2 \pi^{-1/2} X e^{-X^2}) \frac{\vb{v}_{1}}{|\vb{v}_{1}|^3},
+\begin{aligned}
+\vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
+                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
+\end{aligned}
 \end{equation}
@@
 we specify the initial conditions confined to the $xy$-plane as 
-\begin{gather}
-\vb{x}_0 = (r_0, b)\\
-\dot{\vb{x}}_0 = (-v_0, 0).
-\end{gather}
+\begin{align}
+\vb{x}_0 &= (r_0, b)\\
+\dot{\vb{x}}_0 &= (-v_0, 0).
+\end{align}
 We fix $r_0 = 25\,\mathrm{pc}$,
@@
 which might be related to the large scatter in the simulation data.
+
 Overall we can however see that this simple model correctly captures the behavior of eccentricity seen in the simulations.

### 2023-05-24 papers/paper-eccentricity/paper-eccentricity.tex

--- 2023-05-23 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
-\subsection{Reproducing the eccentricity scatter with a toy model}\label{ssec:toy}
+\subsection{Reproducing the eccentricity behavior with a toy model}\label{ssec:toy}
 
-We can gain insight into the large binary eccentricity scatter observed in the
-runs with high galactic merger orbit eccentricities by considering a simplified
-model that captures the essential features of how the SMBH binary becomes bound
-and settles on an orbit with a given eccentricity.
-The components we include in this model are the two SMBHs,
-a smooth background potential, and dynamical friction.
+To gain insight into the origin of the dependence of binary eccentricity on the 
+deflection angle $\theta$ during the binary formation,
+we have constructed a simple toy model that can qualitatively reproduce the observed
+behavior from the simulations.
+This model includes only the essential components relevant to this process,
+which are the asymmetric background stellar potential,
+dynamical friction from the stellar background,
+and naturally the SMBHs themselves.
+
 For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
-to lie in the $xz$-plane.
+to lie in the $xy$-plane.
 In this case the relative motion of the BHs can be treated as a one-body problem,
@@
 
-The background galaxy is modelled as a with constant density spheroidal potential,
+The background galaxy is modeled as a constant density spheroidal potential,
 causing the acceleration \citep[e.g.][]{binney2008}
 \begin{equation}
-\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y,A_z) \vb{x},
+\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y) \vb{x},
 \end{equation}
@@
 and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
-\begin{equation}
-A_x = A_y  = 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
-    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]
-\end{equation}
-\begin{equation}
-A_z = \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+\begin{align}
+A_x &= 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]\\
+A_y &= \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
         \left[ \frac{1}{1-e_\mathrm{s}^2} 
         - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
-\end{equation}
+\end{align}
+Here we have chosen the long axis of the spheroid to be aligned with the $x$-axis.
 With eccentricity values of $e_\mathrm{s} \approx 0.9$,
 this spheroidal potential matches the central potential in the galaxy during the
-BH binary binding phase reasonably well,
+SMBH binary binding phase reasonably well,
 apart from the contributions from the BHs themselves and the groups of stars bound to them.
 
-For dynamical friction, we use the \drafting{Chandrasekhar ref} formula assuming
-a Maxwellian distribution with a constant velocity dispersion $\sigma$.
+For dynamical friction, we use the \citet{1943ApJ....97..255C} formula assuming
+a Maxwellian distribution with a constant velocity dispersion $\sigma$ \citep[e.g.][]{binney2008}.
 This formula gives the acceleration of a single BH due to dynamical friction as
 \begin{equation}\label{eq:chandra_df}
-\vb{a}_{1,\mathrm{DF}} = - 4 \pi G^2 M_\bullet \rho \ln{\Lambda} 
-                        (\operatorname{erf}(X) - 2 \pi^{-1/2} X e^{-X^2}) \frac{\vb{v}_{1}}{|\vb{v}_{1}|^3},
+\begin{aligned}
+\vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
+                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
+\end{aligned}
 \end{equation}
@@
 
-After specifying the initial position and velocity, as well as the other parameters of the model,
-the equation of motion can be solved using standard numerical integration techniques.
-We use the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}
-for this purpose.
 To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
-we specify the initial conditions confined to the $xz$-plane as 
-\begin{gather}
-\vb{x}_0 = (r_0, b)\\
-\dot{\vb{x}}_0 = (-v_0, 0).
-\end{gather}
+we specify the initial conditions confined to the $xy$-plane as 
+\begin{align}
+\vb{x}_0 &= (r_0, b)\\
+\dot{\vb{x}}_0 &= (-v_0, 0).
+\end{align}
 We fix $r_0 = 25\,\mathrm{pc}$,
@@
 the model curve in the $\theta \gtrsim 120\degr$ region,
-although the model does appear to capture the mean behaviour fairly well.
-The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
+although the model does appear to capture the mean behavior fairly well.
+The behavior of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
 which might be related to the large scatter in the simulation data.
-Overall we can however see that this simple model correctly captures the behaviour of eccentricity seen in the simulations.
+
+Overall we can however see that this simple model correctly captures the behavior of eccentricity seen in the simulations.
+On the other hand, if the model is run with low values of $e_\mathrm{s}$,
+the binary always reaches extremely high eccentricities with little dependence on the deflection angle,
+indicating that the asymmetry of the galactic potential during the SMBH binary formation is the cause for the observed eccentricity behavior.
 

### 2023-05-24 papers/paper-eccentricity/paper-eccentricity.tex
Change fig 4 axis label

@@
     \caption{
-    Scattering deflection angle $\theta$ \drafting{update to $\theta_\mathrm{defl}$} and the resulting binary eccentricity
+    Scattering deflection angle $\theta_\mathrm{defl}$ and the resulting binary eccentricity
     in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,

### 2023-05-24 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -104,7 +104,7 @@ def paper_plots():
     axes[0].set_ylabel('$e$')
     axes[0].set_ylim(0,1)
     for ax in axes:
-        ax.set_xlabel(r'$\theta$')
+        ax.set_xlabel(r'$\theta_\mathrm{defl}$')
         ax.xaxis.set_major_formatter(degree_format_str)
         ax.set_xlim(20,170)
         ax.set_xticks(np.arange(30,180,30))

### 2023-05-25 papers/paper-eccentricity/paper-eccentricity.tex
Add figure showing toy model orbits

@@
 
+\begin{figure*}
+    \centering
+    \includegraphics{sample_model_orbits}
+    \caption{
+    Top:
+    Sample orbits computed with the toy model for potentials with different degrees of asymmetry
+    given by the spheroid eccentricity $e_\mathrm{s}$, with otherwise identical initial conditions.
+    The dashed lines show the isopotential contours of the model.
+    Only the early part of the evolution is shown for clarity.
+    Bottom:
+    The orbital eccentricity $e$ of the model orbits.
+    The vertical line marks the end of the period shown in the top panels. 
+    Note the non-linear scale to allow showing the high-eccentricities reached by
+    the $e_\mathrm{s}=0.2$ model.
+    }
+    \label{fig:model_orbits}
+\end{figure*}
+
 To gain insight into the origin of the dependence of binary eccentricity on the 
@@
 Overall we can however see that this simple model correctly captures the behavior of eccentricity seen in the simulations.
-On the other hand, if the model is run with low values of $e_\mathrm{s}$,
-the binary always reaches extremely high eccentricities with little dependence on the deflection angle,
-indicating that the asymmetry of the galactic potential during the SMBH binary formation is the cause for the observed eccentricity behavior.
+We can therefore use the model to understand the cause of the complicated dependence of the binary eccentricity
+on the deflection angle $\theta_\mathrm{defl}$.
+\autoref{fig:model_orbits} shows sets of orbits computed with the model for two different potential shapes,
+with the parameters otherwise as in the model that matches the $e_0=0.90$ simulation results.
+In the $e_\mathrm{s}=0.9$ case the shape of the orbit after the initial scattering depends strongly on the deflection angle,
+as the asymmetric potential causes torque on the SMBHs that are not moving along the main axes of the potential.
+On the other hand, in the more symmetric potential with $e_\mathrm{s}=0.2$,
+The orbit shape is not very sensitive to the deflection angle due to the symmetry of the system,
+and the binary always reaches extremely high eccentricities.
 

### 2023-05-25 papers/paper-eccentricity/scripts/figure_config.py

@@ -1,5 +1,6 @@
 import os
-from matplotlib import rc_file, rcdefaults
+import matplotlib as mpl
+import numpy as np
 from Plotter import Plotter
 
 this_dir = os.path.dirname(os.path.realpath(__file__))
@@ -15,8 +16,54 @@ def data_path(dname):
     return os.path.join(data_dir, dname)
 
 # make sure were using the same settings
-rcdefaults()
-rc_file(os.path.join(this_dir,'matplotlibrc_publish'))
+mpl.rcdefaults()
+mpl.rc_file(os.path.join(this_dir,'matplotlibrc_publish'))
+
+def _make_color_cycle():
+    """
+    Construct the colour palette: a reordered, 10-element version of the 
+    `twilight` colour map.
+    """
+    cols_f = lambda x: mpl.cm.twilight(mpl.colors.Normalize(-1,7)(x))
+    #cols = [cols_f(i)  for p in zip(range(1,5), range(6,10)) for i in p]
+    #cols += [cols_f(0), cols_f(5)]
+    cols = cols_f([i for i in range(0,6,1)])
+
+    return mpl.cycler(color=cols)
+
+color_cycle = _make_color_cycle()
+marker_cycle = mpl.cycler(marker=["o", "s", "^", "D", "v", "*", "p", "h", "X", "P"])
+
+mpl.rcParams['axes.prop_cycle'] = color_cycle
 
 # set up a custom plotting object
 plotter = Plotter()
+
+
+
+class EccentricityScale(mpl.scale.FuncScale):
+    """
+    A non-linear scale for bound binary eccentricity plots.
+    Approximately linear for values below ~0.6, then increasingly non-linear
+    to expand the range between 0.9 and 1.
+    """
+    name = 'eccentricity'
+    def __init__(self, axis):
+        self.fac = 4.5 
+        def forward(x):
+            with np.errstate(divide="ignore", invalid="ignore"): 
+                res = 1-10**(-np.arctanh(x)/self.fac)
+                return np.where(x>1, 1000, res) # make invalid values be mapped outside the plot
+        def inverse(x):
+            with np.errstate(divide="ignore", invalid="ignore"): 
+                return np.tanh(-self.fac*np.log10((1-x)))
+
+        super().__init__(axis, (forward, inverse))
+
+    def set_default_locators_and_formatters(self, axis):
[…]

### 2023-05-25 papers/paper-eccentricity/scripts/toy_model.py

@@ -277,6 +277,34 @@ def high_res_well_fitting_models():
     with open(f'data/well_fitting_e_0.99_model_curve.pkl', 'wb') as f:
         pickle.dump(dict(e_spheroids=e_spheroids, res=res099), f, protocol=-1)
 
+def gen_orbit_plot_data():
+# Some orbits for paper for the e0=0.9 well fitting config + ~spherical system
+    # for comparison
+    r0 = 25e-3
+    v0 = 450 * km_per_s
+    bs = np.array([3.2e-3, 6e-3, 9.8e-3])
+    for e_s in [0.2, 0.9]:
+        sys = System(M_BH=1e-2,
+                     rho=30,
+                     e_spheroid=e_s,
+                     stellar_sigma=200*km_per_s,
+                     df_fudge_factor=0.5)
+        res = []
+        for b in bs:
+            x = [r0,b]
+            v = [-v0,0]
+            tmax = 150 * r0/v0
+            ts = np.linspace(0,tmax, 10000)
+
+            x,v,t = sys.integrate(x,v,ts)
+            e = sys.eccentricity(x,v)
+            th = compute_deflection_angle(sys,x,v)
+            res.append(dict(x=x,v=v,e=e,t=t,th=th,b=b))
+            print(e_s, b, 'done')
+
+        with open(f'data/sample_orbits_e_s_{e_s:.1f}.pkl', 'wb') as f:
+            pickle.dump(res, f)
+
 
 
 
@@ -294,7 +322,7 @@ def test_plots():
     sigma_ref = 200*km_per_s # stellar velocity dispersion
     gal_e = 0.905
 
-    v0_per_sigma = 5.6/2 # initial BH velocity / stellar sigma
+    v0_per_sigma = 4.5/2 # initial BH velocity / stellar sigma
     r0_per_rinfl = .5 # initial BH separation/single BH influence radius
     #v0_per_sigma = 2.2 # initial BH velocity / stellar sigma
     #r0_per_rinfl = 2 # initial BH separation/single BH influence radius
@@ -326,7 +354,8 @@ def test_plots():
     #r0 = r0_per_rinfl * sys.r_infl
     r0 = 25e-3
     print(r0*1e3, v0/km_per_s, 2*sys.a_hard*1e3)
-    bs = np.linspace(bmin_per_r0, bmax_per_r0, 30)*r0
+    bs = np.linspace(bmin_per_r0, bmax_per_r0, 15)*r0
+
 
 
     efin = []
[…]

### 2023-05-25 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -5,8 +5,7 @@ import matplotlib.pyplot as plt
 import pickle
 from scipy.interpolate import griddata, interp1d
 
-from figure_config import fig_path, data_path
-from Plotter import Plotter
+from figure_config import fig_path, data_path, marker_cycle, color_cycle
 
 degree_format_str = '{x:.0f}°'
 
@@ -95,11 +94,9 @@ def plot_sim_data_th_e(ax, simdata):
                    )
 
 
-def paper_plots():
-    plotter = Plotter()
+def paper_th_e_curve_plot():
 
-    fig, axes = plt.subplots(1,2,sharey='row')
-    fig.set_figwidth(2*fig.get_figwidth())
+    fig, axes = plt.subplots(1,2,sharey='row', figsize=(6,3.))
 
     axes[0].set_ylabel('$e$')
     axes[0].set_ylim(0,1)
@@ -111,18 +108,12 @@ def paper_plots():
 
     data = load_data(data_path('well_fitting_e_0.90_model_curve.pkl'))
     curve_shift = -12
+    color = next(iter(color_cycle[-1:]))['color'] 
+
     for e_s, d in zip(data['e_spheroids'], data['res']):
         if e_s <0.9: continue
         a = 1 if e_s == 0.905 else 0.5
-        l, = axes[0].plot(np.degrees(d['theta'])+curve_shift, d['e'], color='tab:blue', alpha=a)
-    #axb = axes[0].secondary_xaxis('top', 
-    #            functions=(interp1d(np.degrees(d['theta']), d['b']*1e3,
-    #                                bounds_error=False, fill_value='extrapolate'),
-    #                       interp1d(d['b']*1e3, np.degrees(d['theta']),
-    #                                bounds_error=False, fill_value='extrapolate')))
-    #axb.set_xlabel('$b/\mathrm{pc}$')
-    #axb.set_xticks([1,2,3,5,10,20])
-    #axes[0].tick_params(top=False)
+        l, = axes[0].plot(np.degrees(d['theta'])+curve_shift, d['e'], color=color, alpha=a)
     print("b90 for e_0=0.90", np.interp(90,np.degrees(d['theta'][::-1]),d['b'][::-1]*1e3))
 
 
@@ -131,51 +122,94 @@ def paper_plots():
                   facecolor='k')
     axes[0].text(22,0.13,"Model shift")
 
-    axes[0].text(22,0.95,"$e_0=0.90$")
-
[…]

### 2023-06-02 papers/paper-eccentricity/paper-eccentricity.tex
Draft edits

@@
 \section{Introduction}
-Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies. 
-As galaxies dynamically interact, so too are their supermassive black holes expected to interact in a three-phase merger process.
-The first phase, dynamical friction, acts to bring the SMBHs from kpc-scale separations to pc-scale separations, after which the SMBHs form a bound binary system, which can be characterised by an orbit with a semimajor axis $a$ and eccentricity $e$.
-Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution, and undergoes two types of interactions: hierarchical interactions and chaotic interactions.
+Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies \citep[e.g.][]{kormendy2013}. 
+As galaxies dynamically interact, so too are their supermassive black holes expected to interact in a three-phase merger process \citep{begelman1980}.
+The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kpc-scale separations to pc-scale separations, after which the SMBHs form a bound binary system, which can be characterised by an orbit with a semimajor axis $a$ and eccentricity $e$.
+Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980}, and undergoes two types of interactions: hierarchical interactions and chaotic interactions.
 Hierarchical interactions occur when a perturbing third body, or a collection of bodies, is far from the SMBH binary, $a_\mathrm{bin} < a_\mathrm{perturber}$, and the energy of the system does not change, however the angular momentum does.
@@
 
+\begin{figure*}
+    \includegraphics{eccentricities.pdf}
+    \caption{Top panel shows the eccentricity $e$ (note the non-linear scale) as a function of time $t$ for select representative simulations from the $e_0 = 0.90$ set (lines with circle markers) and $e_0 = 0.99$ set (lines with square markers).
+Time is shown on a logarithmic scale to highlight the variability in eccentricity prior to the SMBH binary becoming bound (vertical black lines).
+The eccentricity for a full set of a given mass resolution is shown in the bottom left panel for the $e_0 = 0.90$ suite and in the bottom right panel for the $e_0 = 0.99$ suite as a function of $t' = t-t(a=a_\mathrm{h})$ for five orbit periods centred about $t'=0$.
+The dashed black line in the bottom panels indicates the collective mean of the equivalent eccentricities averaged over the span of the respective line.
+The blue area corresponds to the $2.5\sigma$ scatter in the data.
+Note the asymmetry in the blue area is due to the scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.
+In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean.\drafting{Add legend to top panel}.}
+    \label{fig:eccentricities}
+\end{figure*}
+
 We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017}, implemented in \gadget{}.
@@
 We follow the same orbital configurations as \citet{gualandris2022}, with three additional orbit. 
-The chosen eccentricities $e$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
+The chosen eccentricities $e_0$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
 We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022}.
@@
 where $r_\mathrm{h}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
-Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means.
+Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means, denoted $\sigma_e$.
 
@@
 
-\begin{figure*}
-    \includegraphics{eccentricities.pdf}
-    \caption{Top panel shows the equivalent \drafting{(better term?
-Reciprocal?)} eccentricity $1-e$ as a function of time $t$ for select representative simulations from the $e_0 = 0.90$ set (solid lines) and $e_0 = 0.99$ set (dashed lines).
-Time is shown on a logarithmic scale to highlight the variability in eccentricity prior to the SMBH binary becoming bound (vertical black lines).
-The equivalent eccentricity for a full set of a given mass resolution is shown in the bottom left panel for the $e_0 = 0.90$ suite and in the bottom right panel for the $e_0 = 0.99$ suite as a function of $t' = t-t(a=a_\mathrm{h})$ for five orbits centred about $t'=0$.
-The dashed line in the bottom panels indicates the collective mean of the equivalent eccentricities averaged over the span of the respective line.
-The blue area corresponds to the $2.5\sigma$ scatter in the data.
-Note the asymmetry in the grey area is due to the logarithmic scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.
-A zoom inset in the bottom right figure shows the tight bunching of $1-e$ for this particular simulation set.
-In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $1-e$ greater than $2.5\sigma$ from the collective mean.}
-    \label{fig:eccentricities}
-\end{figure*}
-
 \begin{figure}
     \includegraphics{convergence}
-    \caption{\drafting{Convergence of eccentricity scatter as a function of mass resolution for $e_0=0.90$ and $e_0=0.99$. }}
+    \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the total number of stellar particles $2N_\star$ for initial eccentricities $e_0=0.90$ and $e_0=0.99$.
+The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ suite after applying a $2.5\sigma$-cut to the simulation data.
+For the $e_0=0.99$ suite, a steeper scaling than the $1/\sqrt{N_\star}$ is obtained after applying a similar $2.5\sigma$-cut to the data.
+\drafting{Change top axis label to 2N.}}
     \label{fig:convergence}
@@
 
-An representative set of realisations for the $e_0=0.90$ orbit is shown in the bottom left panel of \autoref{fig:eccentricities}.
+A representative set of realisations for the $e_0=0.90$ orbit is shown in the bottom left panel of \autoref{fig:eccentricities}.
 We define a shifted time $t'$ such that $t'(0) = t(a=a_\mathrm{h})$, and the data is truncated to a maximum of five orbits centred about $t'(0)$. 
-One simulation is seen to lie beyond $2.5\sigma$ of the mean $1-e$ line.
+One simulation is seen to lie beyond $2.5\sigma$ of the mean eccentricity.
 
@@
 Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at both $5\times10^3$ and $1\times10^4$ resolutions which has a value of $e_\mathrm{h}<0.4$.
-This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity (i.e., taking those lines within the grey region of the left panel of \autoref{fig:eccentricities}) we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity (i.e., taking those lines within the blue region of the left panel of \autoref{fig:eccentricities}) we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
 However, the eccentricity values beyond $2.5\sigma$ from the mean are each a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups.
@@
 The dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit is shown in \autoref{fig:convergence} with triangle and diamond markers. 
-We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{M_\bullet/m_\star}\propto1/\sqrt{N_\star}$ relation described in \citet{nasim2020}.
+We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{N_\star}$ relation described in \citet{nasim2020}.
 We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
@@
 The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory, however is able to be more robustly measured than $b$ from the simulation data.
-We set as the first `hard-scattering' angle the first 
+We set as the first deflection angle the first 
 angle $\theta_\mathrm{defl}$ that exceeds an orbit-dependent threshold:
-$\theta_{\mathrm{defl,min}}=90\degr$ for $e_0=0.99$ and $\theta_{\mathrm{defl,min}}=45\degr$ for $e_0=0.90$.
+$\theta_{\mathrm{defl,min}}=90\degr$ for the $e_0=0.99$ suite and $\theta_{\mathrm{defl,min}}=45\degr$ for the $e_0=0.90$ suite.
 In general, this corresponds to the second or third actual pericentre passage of the SMBHs in the simulations.
-An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.99$ orbit in \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
+An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.90$ ($e_0=0.99$) orbit in the left (right) panel of \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
 We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, which we explain using a simple toy model.
@@
     \includegraphics[width=0.49\textwidth]{orbit-e99}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t=t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t_\mathrm{bound}+\epsilon$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+The term $\epsilon$ allows for the SMBH trajectory to be seen for some small time after the two SMBHs have become bound.
 Each line is coloured according to time of the simulation, however the colour for $t<31\,\mathrm{Myr}$ ($t<22\,\mathrm{Myr}$) is constant for visual clarity of the later evolution in the left (right) panel.
@@
     The model curves have been shifted left by $12\degr$ and $6\degr$
-    to better match the simulation data, with the shift also indicated by arrows.
+    to better match the simulation data, with the shift also indicated by arrows. \drafting{Include marginal histograms.}
     }
@@
 To gain insight into the origin of the dependence of binary eccentricity on the 
-deflection angle $\theta$ during the binary formation,
+deflection angle $\_\mathrm{defl}$ during the binary formation,
 we have constructed a simple toy model that can qualitatively reproduce the observed
@@
 
-For dynamical friction, we use the \citet{1943ApJ....97..255C} formula assuming
-a Maxwellian distribution with a constant velocity dispersion $\sigma$ \citep[e.g.][]{binney2008}.
+For dynamical friction, we use the \citet{chandrasekhar1943} formula assuming
+a Maxwellian distribution with a constant velocity dispersion $\sigma_\mathrm{M}$ \citep[e.g.][]{binney2008}.
 This formula gives the acceleration of a single BH due to dynamical friction as
@@
 where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
-$X = |\vb{v}_1|/\sqrt{2}\sigma$,
+$X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
 and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 10$.
@@
 stellar background is much more complex.
-Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f=\alpha$.
+Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f(a)=\alpha$.
 We set the scaling factor $\alpha=0.5$, and use cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
@@
 $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
-$\sigma=200\,\mathrm{km\,s^{-1}}$,
+$\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}$,
 and
@@
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
+We observe a modest decrease in deflection angle $\theta_\mathrm{defl}$ with increasing mass resolution for both the $e_0=0.90$ and $e_0=0.99$ cases. 
 To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
@@
 In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
-the model curve in the $\theta \gtrsim 120\degr$ region,
+the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
 although the model does appear to capture the mean behavior fairly well.
@@
 
-\drafting{
-    An interesting point to discuss is the origin of our eccentricity scatter.
-Early tests of the Gualandris setup seems to indicate that low eccentricity mergers do not demonstrate orbital flips, and are more converged.
-How to distinguish if the flip is the cause of the eccentricity scatter, or just correlated with eccentricity scatter?
-}
-
-We present in \autoref{sec:results} evidence for a tight correlation between the first hard-scattering angle $\theta_\mathrm{defl}$ and the hard SMBH binary eccentricity. 
+We present in \autoref{sec:results} evidence for a tight correlation between the deflection angle $\theta_\mathrm{defl}$ and the hard SMBH binary eccentricity $e_\mathrm{h}$. 
 With the inclusion of the toy model, which provides a resolution-free model of the SMBH scattering process, we may draw two key interpretations of the results.
@@
 the number of computing cores used).
-The scatter we observe in $\theta_\mathrm{defl}$ is thus a manifestation of Brownian motion induced by the phase space discretisation and other numerical errors, such as floating point round-off errors, denoted $\sigma_\mathrm{sim}$.
-Thus, the numerical simulations present an instance where there is zero uncertainty in the initial conditions of the system, however there is error in the numerical evolution of the system.
+The scatter we observe in $\theta_\mathrm{defl}$ is thus a manifestation of Brownian motion induced by the phase space discretisation and other numerical errors, such as floating point round-off errors, denoted $\tau_\mathrm{sim}$.
+Thus, the numerical simulations present an instance where there is zero uncertainty in a particular realisation of the system initial conditions, however there is error in the numerical evolution of the system.
 We determine the positional displacement between corresponding SMBHs across the simulations at a given mass resolution for the $e_0=0.90$ and $e_0=0.99$ orbits by moving the SMBH binary system to the centre of mass frame, and setting the SMBH positions to coincide with the origin at the beginning of the simulation.
@@
 There is however uncertainty in the initial conditions of the system, whereby the chance of two galaxy merger events having the exact same initial configuration is practically zero.
-We denote this uncertainty as $\sigma_\mathrm{true}$.
-For the mergers of massive galaxies of the types modelled in this work, an uncertainty in the initial state of $\sigma_\mathrm{true} \sim \sigma_\mathrm{sim}$ is not unexpected, given the influence radius of a $10^8\,\Msun$ SMBH in a stellar background with velocity dispersion of the order $200\,\mathrm{km}\,\mathrm{s}^{-1}$ is $\sim 10\,\mathrm{pc}$ \drafting{From MBW, find better reference}.
+We denote this uncertainty as $\tau_\mathrm{true}$.
+For the mergers of massive galaxies of the types modelled in this work, an uncertainty in the initial state of $\tau_\mathrm{true} \sim \tau_\mathrm{sim}$ is not unexpected, given the influence radius of a $10^8\,\Msun$ SMBH in a stellar background with velocity dispersion of the order $200\,\mathrm{km}\,\mathrm{s}^{-1}$ is $\sim 10\,\mathrm{pc}$ \citep[e.g.][]{binney2008}.
 
-Thus, any process that perturbs the impact parameter associated with the hard scattering angle will result in an eccentricity sampled from the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve, subject to some small scatter in $e$ resulting from the hierarchical interactions of distant perturbers, or by exchanges in angular momentum with the stellar cluster immediately surrounding the SMBH.
+Thus, any process that perturbs the impact parameter associated with the hard scattering angle will result in an eccentricity sampled from the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve, subject to some small scatter in $e_\mathrm{h}$ resulting from the hierarchical interactions of distant perturbers, or by exchanges in angular momentum with the stellar cluster immediately surrounding the SMBH.
 Critically however, the expected value of eccentricity is a direct consequence of the hard scattering angle, and is not an independent random process, and can be reliably predicted.
+Whilst we see a modest decrease in the scatter in $\theta_\mathrm{defl}$ with increasing mass resolution, we propose that more relevant to the reduced scatter in eccentricity is the domain space in $\theta_\mathrm{defl}$ about which the simulation data clusters for a given value of $e_0$.
+From \autoref{fig:theta_e}, we see that the simulation data happens to cluster about a relatively stable section of the $\theta_\mathrm{defl}\mathrm{-}e_\mathrm{h}$ curve for the $e_0=0.90$ suite.
+Conversely, for the $e_0=0.99$ simulations, such a stable section in $\theta_\mathrm{defl}\mathrm{-}e_\mathrm{h}$ space is not present, leading to an increased eccentricity scatter compared to the $e_0=0.90$ case.
+
 We caution that there is not necessarily a one-to-one mapping of the initial merger eccentricity $e_0$ and the hard binary eccentricity $e_\mathrm{h}$, unless the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve is uniform.
-As such a uniform function is not associated with the radial ($e_0>0.99$) mergers expected to occur in reality, we expect that for a given initial merger orbit $e_0$ a range of possible $e_\mathrm{h}$ would be observed.
+As such a uniform function is not associated with the radial ($e_0>0.99$) mergers expected to occur in reality, we expect that for a given initial merger orbit $e_0$ a range of possible $e_\mathrm{h}$ values would be observed.
 The converse also holds: inference about merging SMBH binary eccentricity does not necessarily provide a mapping back to the initial merger eccentricity unless the history of the system is known, which is unable to be achieved in a practical sense in reality.
+
+In conclusion, we have shown the importance of the hard scattering angle $\theta_\mathrm{defl}$ on the resulting hard binary eccentricity $e_\mathrm{h}$. 
+By using a resolution-free toy model of the SMBH scattering process, we demonstrate that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations.
+The results presented here justify further investigation into quantifying the expected range of hard binary eccentricities for merger timescale predictions for future gravitational wave observation missions.
 
@@
 the CSC -- IT Center for Science, Finland.
+
+We list here the roles and contributions of the authors according to the Contributor Roles Taxonomy (CRediT\footnote{https://credit.niso.org}).
+\textbf{AR}: Conceptualisation, Investigation, Formal analysis, Data curation, Writing -- original draft.
+\textbf{MM}: Conceptualisation, Formal analysis, Writing -- original draft.
 

### 2023-06-02 papers/paper-eccentricity/paper-eccentricity.tex
- add legend to timeseries plot

@@
 Note the asymmetry in the blue area is due to the scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.
-In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean.\drafting{Add legend to top panel}.}
+In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean.}
     \label{fig:eccentricities}
@@
 The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ suite after applying a $2.5\sigma$-cut to the simulation data.
-For the $e_0=0.99$ suite, a steeper scaling than the $1/\sqrt{N_\star}$ is obtained after applying a similar $2.5\sigma$-cut to the data.
-\drafting{Change top axis label to 2N.}}
+For the $e_0=0.99$ suite, a steeper scaling than the $1/\sqrt{N_\star}$ is obtained after applying a similar $2.5\sigma$-cut to the data.}
     \label{fig:convergence}
@@
     The model curves have been shifted left by $12\degr$ and $6\degr$
-    to better match the simulation data, with the shift also indicated by arrows. \drafting{Include marginal histograms.}
+    to better match the simulation data, with the shift also indicated by arrows.
+    The marginal histograms show kernel density estimates to the simulation data.
     }

### 2023-06-02 papers/paper-eccentricity/scripts/eccentricity_timeseries.py

@@ -1,6 +1,6 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from matplotlib import rcParams
+from matplotlib.lines import Line2D
 import ketjugw
 import cm_functions as cmf
 import figure_config
@@ -60,8 +60,15 @@ for suite, marker in zip((e90_data_raw, e99_data_raw), figure_config.marker_cycl
         bh1, bh2, merged = cmf.analysis.get_bound_binary(kf)
         op = ketjugw.orbital_parameters(bh1, bh2)
         axdict["A"].plot(op["t"]/cmf.general.units.Myr, op["e_t"], marker=marker, markevery=[-1], zorder=10)
+axdict["A"].legend(
+                handles = [
+                    Line2D([0], [0], marker="o", label=r"$e_0=0.90$", c="w", mec="k", lw=0.5),
+                    Line2D([0], [0], marker="s", label=r"$e_0=0.99$", c="w", mec="k", lw=0.5)
+                ]
+)
 
 
+# plot data about t=t_h in separate panels
 for axi in (axdict["B"], axdict["C"]): 
     axi.set_xlabel(r"$t'/\mathrm{Myr}$")
     axi.set_ylabel(r"$e$")

### 2023-06-05 papers/paper-eccentricity/paper-eccentricity.tex
Updates to discussion

@@
     \includegraphics{eccentricities.pdf}
-    \caption{Top panel shows the eccentricity $e$ (note the non-linear scale) as a function of time $t$ for select representative simulations from the $e_0 = 0.90$ set (lines with circle markers) and $e_0 = 0.99$ set (lines with square markers).
-Time is shown on a logarithmic scale to highlight the variability in eccentricity prior to the SMBH binary becoming bound (vertical black lines).
-The eccentricity for a full set of a given mass resolution is shown in the bottom left panel for the $e_0 = 0.90$ suite and in the bottom right panel for the $e_0 = 0.99$ suite as a function of $t' = t-t(a=a_\mathrm{h})$ for five orbit periods centred about $t'=0$.
-The dashed black line in the bottom panels indicates the collective mean of the equivalent eccentricities averaged over the span of the respective line.
-The blue area corresponds to the $2.5\sigma$ scatter in the data.
-Note the asymmetry in the blue area is due to the scale of the $y$-axis: the standard deviation is by construct symmetric about the mean.
-In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean.}
+    \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for select representative simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
+The downwards-pointing arrows indicate the mean time when the SMBH binaries becomes hard for the two simulation sets.
+In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean.
+In the $e_0=0.90$ set, two simulations have a SMBH binary merger, as indicated by the rapid circularisation of the eccentricity.}
     \label{fig:eccentricities}
@@
 We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
-We follow the same orbital configurations as \citet{gualandris2022}, with three additional orbit. 
-The chosen eccentricities $e_0$ are 0.5, 0.7, 0.9, 0.95, 0.97, and 0.99.
+The chosen eccentricities $e_0$ are 0.90 and 0.99.
 We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022}.
@@
 Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
-Additionally, we run the $e_0=0.9$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
+Additionally, we run the $e_0=0.90$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
 
@@
 
-A representative set of realisations for the $e_0=0.90$ orbit is shown in the bottom left panel of \autoref{fig:eccentricities}.
-We define a shifted time $t'$ such that $t'(0) = t(a=a_\mathrm{h})$, and the data is truncated to a maximum of five orbits centred about $t'(0)$. 
-One simulation is seen to lie beyond $2.5\sigma$ of the mean eccentricity.
+A representative set of realisations for the $e_0=0.90$ orbit is shown with blue lines in \autoref{fig:eccentricities}.
+We define a shifted time $t'$ such that $t'(0) = t_\mathrm{bound}$. 
+One simulation is seen to lie significantly beyond the mean eccentricity, with $e_\mathrm{h}\sim 0.2$.
 
@@
 
-Similar to the $e_0=0.90$ case, we show a representative set of realisation for the $e_0=0.99$ orbit in the bottom right panel of \autoref{fig:eccentricities}.
+Similar to the $e_0=0.90$ case, we show a representative set of realisation for the $e_0=0.99$ orbit with orange lines in \autoref{fig:eccentricities}.
+Again, one simulation is seen to have $e_\mathrm{h} \sim 0.1$, significantly offset from the mean value of $e_\mathrm{h}$ for this simulation set.
+
 The dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit is shown in \autoref{fig:convergence} with triangle and diamond markers. 
@@
     \centering
-    \includegraphics[width=0.49\textwidth]{orbit-e90}
-    \includegraphics[width=0.49\textwidth]{orbit-e99}
+    \includegraphics{orbit}
     \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t_\mathrm{bound}+\epsilon$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
 The term $\epsilon$ allows for the SMBH trajectory to be seen for some small time after the two SMBHs have become bound.
-Each line is coloured according to time of the simulation, however the colour for $t<31\,\mathrm{Myr}$ ($t<22\,\mathrm{Myr}$) is constant for visual clarity of the later evolution in the left (right) panel.
+Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
 In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
@@
 
-We present in \autoref{sec:results} evidence for a tight correlation between the deflection angle $\theta_\mathrm{defl}$ and the hard SMBH binary eccentricity $e_\mathrm{h}$. 
-With the inclusion of the toy model, which provides a resolution-free model of the SMBH scattering process, we may draw two key interpretations of the results.
+We present in \autoref{sec:results} evidence that variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
+The inclusion of the toy model allows us to investigate the relationship between $\theta_\mathrm{defl}$ and $e_\mathrm{h}$ in the infinite-mass resolution limit, around which the quantities measured from the simulation demonstrate minimal scatter.
+As shown in the marginal kernel density estimates of \autoref{fig:theta_e}, the variation in eccentricity can span almost the full domain space of $[0,1]$.
 
-First, the numerical simulations performed with \ketju{} are deterministic for a given set of initial conditions, given the same computing architecture and numerical parameters (e.g.
-the number of computing cores used).
-The scatter we observe in $\theta_\mathrm{defl}$ is thus a manifestation of Brownian motion induced by the phase space discretisation and other numerical errors, such as floating point round-off errors, denoted $\tau_\mathrm{sim}$.
-Thus, the numerical simulations present an instance where there is zero uncertainty in a particular realisation of the system initial conditions, however there is error in the numerical evolution of the system.
-We determine the positional displacement between corresponding SMBHs across the simulations at a given mass resolution for the $e_0=0.90$ and $e_0=0.99$ orbits by moving the SMBH binary system to the centre of mass frame, and setting the SMBH positions to coincide with the origin at the beginning of the simulation.
-Computing the standard deviation in the positional displacements for corresponding SMBHs $\sigma_\mathrm{displ}$ between simulation $i$ and simulation $j\neq i$ for all times $t<t_\mathrm{bound}$, we find the SMBH positions vary of the order $\sim10\,\mathrm{pc}$ at the time of the hard scattering, which translates to an uncertainty in the deflection angle.
+We turn our discussion to the cause of variation in $\theta_\mathrm{defl}$ (and thus the impact parameter $b$) in first simulations, and then in reality.
 
-Second, in the real Universe there is no numerical error affecting the system evolution.
-There is however uncertainty in the initial conditions of the system, whereby the chance of two galaxy merger events having the exact same initial configuration is practically zero.
-We denote this uncertainty as $\tau_\mathrm{true}$.
-For the mergers of massive galaxies of the types modelled in this work, an uncertainty in the initial state of $\tau_\mathrm{true} \sim \tau_\mathrm{sim}$ is not unexpected, given the influence radius of a $10^8\,\Msun$ SMBH in a stellar background with velocity dispersion of the order $200\,\mathrm{km}\,\mathrm{s}^{-1}$ is $\sim 10\,\mathrm{pc}$ \citep[e.g.][]{binney2008}.
+In numerical simulations with a discretised phase space, the two primary causes of variation in $\theta_\mathrm{defl}$ are Brownian motion of the SMBH binary, and initial particle sampling variation in the merger IC, both of which have been shown to be more pronounced in lower mass resolution simulations compared to higher mass resolution simulations.
+By allowing the SMBH binary to wander further from its expected location in phase space, a wider range of impact parameters, and thus deflection angles, may be sampled.
+This consequently maps to a different sampling of the eccentricity space, however due to the non-linear mapping (as shown in \autoref{fig:theta_e}), the eccentricity domain space sampled can vary between $\sigma_e \sim 0.01$ to $\sigma_e>0.1$, for the shown system configurations.
+Different systems, with different galaxy density profiles or merger orbits, may have an even more pronounced value of $\sigma_e$.
 
-Thus, any process that perturbs the impact parameter associated with the hard scattering angle will result in an eccentricity sampled from the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve, subject to some small scatter in $e_\mathrm{h}$ resulting from the hierarchical interactions of distant perturbers, or by exchanges in angular momentum with the stellar cluster immediately surrounding the SMBH.
-Critically however, the expected value of eccentricity is a direct consequence of the hard scattering angle, and is not an independent random process, and can be reliably predicted.
-Whilst we see a modest decrease in the scatter in $\theta_\mathrm{defl}$ with increasing mass resolution, we propose that more relevant to the reduced scatter in eccentricity is the domain space in $\theta_\mathrm{defl}$ about which the simulation data clusters for a given value of $e_0$.
-From \autoref{fig:theta_e}, we see that the simulation data happens to cluster about a relatively stable section of the $\theta_\mathrm{defl}\mathrm{-}e_\mathrm{h}$ curve for the $e_0=0.90$ suite.
-Conversely, for the $e_0=0.99$ simulations, such a stable section in $\theta_\mathrm{defl}\mathrm{-}e_\mathrm{h}$ space is not present, leading to an increased eccentricity scatter compared to the $e_0=0.90$ case.
+Conversely, in the real universe, variation in the impact parameter of a similar magnitude to the presented simulations, but as a consequence of different processes, is expected due to the lack of system reproducibility.
+As with many physical systems, random variations in the system state demand the use of statistics to ascertain a most probable configuration or outcome. 
+\drafting{Discussion about Sgr $A^*$ proper motion and oscillation scales? Or better example?}.
 
-We caution that there is not necessarily a one-to-one mapping of the initial merger eccentricity $e_0$ and the hard binary eccentricity $e_\mathrm{h}$, unless the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ curve is uniform.
-As such a uniform function is not associated with the radial ($e_0>0.99$) mergers expected to occur in reality, we expect that for a given initial merger orbit $e_0$ a range of possible $e_\mathrm{h}$ values would be observed.
-The converse also holds: inference about merging SMBH binary eccentricity does not necessarily provide a mapping back to the initial merger eccentricity unless the history of the system is known, which is unable to be achieved in a practical sense in reality.
+The results presented in this work are significant in constraining observational quantities that are correlated with binary eccentricity: most notably SMBH binary merger timescale. 
+As the relevant scales of the impact parameter that lead to the deflection angle are typically of the order of a few parsecs, and are thus generally unresolvable in gravitationally-softened simulations, numerical simulations sample instead a non-deterministic distribution of potential impact parameters.
+Building on from work by \citet{portegieszwart2018}, microscopic perturbations can grow to large scale, observable effects in a gravitational system over relatively short timescales, suggesting that even at physical mass resolution, chaotic dynamics still lead to a non-deterministic sampling of the impact parameter domain.
+From \autoref{fig:theta_e}, different initial orbital geometries also influence the deflection-angle domain space sampled, with not all merger geometries necessarily able to sample the full eccentricity domain (neglecting minor scatter in $e_\mathrm{h}$ due to sampling variation compared to the analytical prediction).
+Importantly however, the sampling variation in $e_\mathrm{h}$ by the numerical simulations are largely independent of mass resolution, indicating that even at moderate mass resolutions of $M_\bullet/m_\star \sim 10^3$, the eccentricity distribution can be reliably sampled.
 
-In conclusion, we have shown the importance of the hard scattering angle $\theta_\mathrm{defl}$ on the resulting hard binary eccentricity $e_\mathrm{h}$. 
+Finally, we discuss the caveats of our conclusions.
+\begin{enumerate}
+\item The systems we test in this work are exceptionally idealistic: spherical stellar bulges without dark matter (DM) on a merger orbit with very small initial separation. 
+We do not however expect the conclusions concerning the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ mapping to be significantly impacted by expanding our models to more realistic gravitational systems, as DM acts to provide a dynamical friction force over separation scales much larger than that of the impact parameter.
+Qualitatively, the scatter in eccentricity presented in this work is consistent with the eccentricity scatter in simulations including DM. 
+\item The mergers presented in this work are all of equal mass galaxies, leading to very dramatic deformations of the potential that deviate from spherical symmetry. 
+In the instance of minor mergers where the galaxy merger remnant is not as triaxial as the major merger case \drafting{reference needed}, we do not expect to see as sensitive dependence of eccentricity on deflection angle compared to the major merger case (refer to \autoref{fig:model_orbits}).
+\end{enumerate}
+
+In conclusion, we have shown the importance of the deflection angle $\theta_\mathrm{defl}$ on the resulting hard binary eccentricity $e_\mathrm{h}$. 
 By using a resolution-free toy model of the SMBH scattering process, we demonstrate that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations.
-The results presented here justify further investigation into quantifying the expected range of hard binary eccentricities for merger timescale predictions for future gravitational wave observation missions.
+The results presented here justify extending the investigation to more realistic galaxy merger scenarios, to quantify the expected range of hard binary eccentricities for merger timescale predictions for future gravitational wave observation missions.
 

### 2023-06-06 papers/paper-eccentricity/paper-eccentricity.tex
Update GradientPlot classes to improve efficiency
- minor updates to paper plots

@@
 \begin{figure*}
+    \centering
     \includegraphics{eccentricities.pdf}
@@
     \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to $t_\mathrm{bound}+\epsilon$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
-The term $\epsilon$ allows for the SMBH trajectory to be seen for some small time after the two SMBHs have become bound.
-Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity of the later evolution.
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity of the deflection-phase of the evolution.
 In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after.
-The circle marker indicates the final position of the SMBH before the SMBHs form a bound binary.}
+The circle marker indicates the position of the SMBH just before the SMBHs form a bound binary.}
     \label{fig:orbit}

### 2023-06-06 code/cm_functions/plotting/PlotClasses.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 from matplotlib import colors
+from matplotlib.collections import LineCollection
 from ..env_config import _cmlogger
 
 
@@ -15,7 +16,7 @@ class _GradientPlot:
     """
     
     """
-    def __init__(self, ax, cmap="cividis", plot_kwargs={}):
+    def __init__(self, ax, cmap="cividis"):
         """
         Class to create pyplot plots with a colour gradient. The colour 
         gradient is consistent between all lines/points in the figure. This is 
@@ -48,7 +49,6 @@ class _GradientPlot:
                 _logger.logger.exception(f"cmap `{cmap}` not present in matplotlib defaults, nor is registered as a custom map!", exc_info=True)
                 raise
         self.all_marker = []
-        self.all_pks = [plot_kwargs]
         self.norm = [0,1]
         self._default_markerkwargs = {"ec":"k", "lw":0.5}
 
@@ -67,7 +67,7 @@ class _GradientPlot:
         return min([min(c) for c in self.all_c])
 
 
-    def add_data(self, x, y, c, label=None, marker="o", plot_kwargs={}):
+    def add_data(self, x, y, c, label=None, marker="o"):
         """
         Add a dataset to the plot (note that the data is just stored here for 
         future use).
@@ -84,9 +84,6 @@ class _GradientPlot:
             label of plot, by default None
         marker : str, optional
             end marker, by default "o"
-        plot_kwargs : dict, optional
-            dict of other parameters to parse to pyplot.plot() or pyplot.scatter
-            (), by default {}
         """
         try:
             dat_len = [len(v) for v in (x,y,c)]
@@ -99,7 +96,6 @@ class _GradientPlot:
         self.all_c.append(c)
         self.all_label.append(label)
         self.all_marker.append(marker)
-        self.all_pks.append(plot_kwargs)
         self.data_count += 1
 
 
@@ -174,14 +170,19 @@ class GradientLinePlot(_GradientPlot):
     """
     Apply the _GradientPlot class for pyplot line plots
     """
-    def __init__(self, ax, cmap="viridis", plot_kwargs={}):
-        super().__init__(ax, cmap=cmap, plot_kwargs=plot_kwargs)
+    def __init__(self, ax, cmap="viridis"):
+        super().__init__(ax, cmap=cmap)
 
 
[…]

### 2023-06-06 papers/paper-eccentricity/scripts/eccentricity_timeseries.py

@@ -1,6 +1,5 @@
 import numpy as np
 import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
 import ketjugw
 import cm_functions as cmf
 import figure_config
@@ -38,6 +37,10 @@ for suite, label in zip((e90_data_raw, e99_data_raw), (r"$e_0=0.90$", r"$e_0=0.9
         tmin = min(tmin, t[start_idx])
         t = np.concatenate(([t[0]-t[1]], t)) 
         e = np.concatenate(([1], op["e_t"]))
+        if op["e_t"][-1] < 0.1:
+            print("Adding an artificial point at e=0 for visual appeal...")
+            t = np.concatenate((t, [2*t[-1]-t[-2]]))
+            e = np.concatenate((e, [0]))
         # skip first two points to deal with log scale
         l = ax.plot(t[start_idx:], e[start_idx:], zorder=10, c=col, label=(label if i==0 else ""), ls="-")
         col = l[-1].get_color()

### 2023-06-06 papers/paper-eccentricity/scripts/figure_config.py

@@ -41,7 +41,7 @@ mpl.colormaps.register(cmap=custom_diverging_cmap)
 mpl.colormaps.register(cmap=custom_diverging_cmap.reversed(), name="custom_diverging_r")
 
 custom_colors = custom_cmap(np.linspace(0,1,8))
-custom_colors_shuffled = custom_colors[[0, 4, 1, 5, 2, 6, 3, 7]]
+custom_colors_shuffled = custom_colors[[1, 4, 0, 5, 2, 6, 3, 7]]
 
 color_cycle = mpl.cycler(color=custom_colors)
 color_cycle_shuffled = mpl.cycler(color=custom_colors_shuffled)

### 2023-06-06 papers/paper-eccentricity/scripts/orbit_plot.py

@@ -11,10 +11,11 @@ print_data = False
 # turn off for fast plotting, line will be a single colour
 use_gradient_line = True
 
-
+idx_stride = 10
 min_time_colour = -2
 max_time_colour = -min_time_colour
 arrowprops = {"arrowstyle":"wedge", "fc":"k", "ec":"w", "lw":0.1}
+gradline_kwargs = {"ls":"-", "capstyle":"round"}
 
 # set up figure
 fig, ax = plt.subplots(1,2, figsize=(6,3), sharex="all", sharey="all")
@@ -50,7 +51,6 @@ for j in range(2):
         ]
 
         # some shortcuts
-        idx_stride = 10
         A_idx = [4500, 4280]
         B_idx = [A_idx[0]+35, A_idx[1]+30]
         extra_bound_bit = [1000, 2000]
@@ -79,9 +79,8 @@ for j in range(2):
         ]
 
         # some shortcuts
-        idx_stride = 10
         A_idx = [3150, 3140]
-        B_idx = [A_idx[0]+35, A_idx[1]+35]
+        B_idx = [A_idx[0]+29, A_idx[1]+35]
         extra_bound_bit = [1500, 800]
 
 
@@ -114,7 +113,7 @@ for j in range(2):
         t = bh.t[::idx_stride] / cmf.general.units.Myr
         t = t - t[bound_idx_strided[i]]
         if use_gradient_line:
-            glp.add_data(x, y, c=t, plot_kwargs={"ls":"-"})
+            glp.add_data(x, y, c=t)
         else:
             ax[j].plot(x,y, markevery=[bound_idx_strided[i]], marker="o")
             axins.plot(x,y, markevery=[bound_idx_strided[i]], marker="o")
@@ -151,7 +150,7 @@ for j in range(2):
             axins.annotate(
                         "B", 
                         (x[B_idx[i]], y[B_idx[i]]),
-                        xytext=(5, -7) if i==0 else (2.7, -12),
+                        xytext=(0, 25) if i==0 else (2, -16),
                         textcoords="offset points",
                         arrowprops=arrowprops,
                         horizontalalignment="left",
@@ -165,9 +164,9 @@ for j in range(2):
         axins.set_title(f"$\\theta_\mathrm{{defl}}={ang:.1f}\degree$", fontsize="small")
 
     if use_gradient_line:
[…]

### 2023-06-06 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -140,7 +140,7 @@ def paper_th_e_curve_plot():
     color = None
     for e_s, d in zip(data['e_spheroids'], data['res']):
         if e_s <0.9: continue
-        a = 1 if e_s == 0.905 else 0.5
+        a = 0.8 if e_s == 0.905 else 0.3
         l, = axdict["D"].plot(np.degrees(d['theta'])+curve_shift, d['e'], color=color, alpha=a)
         color = l.get_color()
     print("b90 for e_0=0.90", np.interp(90,np.degrees(d['theta'][::-1]),d['b'][::-1]*1e3))
@@ -159,7 +159,7 @@ def paper_th_e_curve_plot():
     color=None
     for e_s, d in zip(data['e_spheroids'], data['res']):
         if e_s < 0.9: continue
-        a = 1 if e_s == 0.905 else 0.5
+        a = 0.8 if e_s == 0.905 else 0.3
         l, = axdict["E"].plot(np.degrees(d['theta'])+curve_shift, d['e'], alpha=a, color=color)
         color = l.get_color()
 
@@ -295,6 +295,6 @@ def paper_orbit_plot():
 
 
 paper_th_e_curve_plot()
-#paper_orbit_plot()
+paper_orbit_plot()
 
 #plt.show()

### 2023-06-06 papers/paper-eccentricity/paper-eccentricity.tex
- minor changes to discussion

@@
 As with many physical systems, random variations in the system state demand the use of statistics to ascertain a most probable configuration or outcome. 
-\drafting{Discussion about Sgr $A^*$ proper motion and oscillation scales? Or better example?}.
+Work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks. 
+Assuming these processes are at play in other galaxies (particularly elliptical galaxies constructed through major mergers), deviations of a few parsecs of the SMBH from the host galaxy photo-centre are expected and would influence the impact parameter two merging SMBHs would experience.
 

### 2023-06-07 papers/paper-eccentricity/paper-eccentricity.tex
Paper edits

--- 2023-06-06 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
+\author[0000-0001-8741-8263]{Peter H. Johansson}
+\affiliation{Department of Physics,
+Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland}
+
 \author{Others}
@@
 
-We present some stuff about eccentricity convergence in Nbody sims. Limited to 250 words.
+\drafting{
+    Key results:
+    \begin{enumerate}
+        \item Eccentricity of hard binary depends on deflection angle 
+        \item Mapping $\theta_\mathrm{defl} \rightarrow e_\mathrm{h}$ relatively independent of resolution, however low res explores wider domain space 
+        \item Explored eccentricity space can be full $[0,1]$ range for certain systems 
+        \item Exact values of $e$ from simulations meaningless, need to understand the distribution of 
+        \item Current resolutions appropriate for exploring statistics of eccentricity
+    \end{enumerate}
+    Limited to 250 words.
+}
 
@@
 Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies \citep[e.g.][]{kormendy2013}. 
-As galaxies dynamically interact, so too are their supermassive black holes expected to interact in a three-phase merger process \citep{begelman1980}.
-The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kpc-scale separations to pc-scale separations, after which the SMBHs form a bound binary system, which can be characterised by an orbit with a semimajor axis $a$ and eccentricity $e$.
-Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980}, and undergoes two types of interactions: hierarchical interactions and chaotic interactions.
-Hierarchical interactions occur when a perturbing third body, or a collection of bodies, is far from the SMBH binary, $a_\mathrm{bin} < a_\mathrm{perturber}$, and the energy of the system does not change, however the angular momentum does.
-Hierarchical interactions drive changes in the SMBH binary eccentricity. 
-Conversely, chaotic interactions between the SMBH binary and a perturber remove orbital energy from the system when the semimajor axis of the perturber is comparable to the semimajor axis of the SMBH binary, $a_\mathrm{bin}\simeq a_\mathrm{perturber}$.
-The angular momentum of the system also varies during the chaotic interactions.
-Chaotic interactions drive the SMBH binary to sub-pc separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
+As galaxies dynamically interact, so too are their SMBHs expected to interact in a three-phase merger process \citep{begelman1980}.
+The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary system with a semimajor axis $a$ and eccentricity $e$.
+Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980, hills1983, quinlan1996}.
+Chaotic interactions drive the SMBH binary to sub-parsec separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
 
-It is well known that the efficiency by which GW emission can drive a SMBH binary to coalescence is highly sensitive to the eccentricity of the binary system \citep{peters1964}, with the merger timescale of the SMBH binary being particularly sensitive.
-Additionally, the complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays (PTAs) and the upcoming Laser Interferometer Space Antenna (LISA) mission.
-The requirement to understand the faithfulness to which SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the GW observational community aims to detect.
+It is well known that the efficiency by which GW emission can remove orbital energy and angular momentum from a binary system is highly sensitive to the binary eccentricity \citep{peters1964}, with the binary merger timescale being particularly sensitive.
+The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
+Understanding how faithfully SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the observational GW community aims to detect.
 
-Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in simulations is an artefact of poor phase space discretisation in simulations, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
+Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
 In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
-Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.99$, stochasticity in the final SMBH binary eccentricity is present. 
-\drafting{Add relevant references.}
+Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final SMBH binary eccentricity is present.
 
@@
 For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
-The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2013}
+The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}
 \begin{equation}\label{eq:ahard}
-    a_\mathrm{h} = \frac{q}{1+q} \frac{r_\mathrm{h}}{4},
+    a_\mathrm{h} = \frac{q}{(1+q)^2} \frac{r_\mathrm{m}}{4},
 \end{equation}
-where $r_\mathrm{h}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
+where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
 Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means, denoted $\sigma_e$.
@@
 At each pericentre passage, the two SMBHs scatter off each other with some deflection angle $\theta_\mathrm{defl}$, defined:
-\begin{equation}
+\begin{equation}\label{eq:theta}
     \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right)
@@
 
-We present in \autoref{sec:results} evidence that variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
+We show in \autoref{sec:results} that variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
 The inclusion of the toy model allows us to investigate the relationship between $\theta_\mathrm{defl}$ and $e_\mathrm{h}$ in the infinite-mass resolution limit, around which the quantities measured from the simulation demonstrate minimal scatter.
-As shown in the marginal kernel density estimates of \autoref{fig:theta_e}, the variation in eccentricity can span almost the full domain space of $[0,1]$.
+As shown by the simulation points and the marginal kernel density estimates (KDEs) of \autoref{fig:theta_e}, the variation in eccentricity can span almost the full domain space of $[0,1]$.
 
-We turn our discussion to the cause of variation in $\theta_\mathrm{defl}$ (and thus the impact parameter $b$) in first simulations, and then in reality.
+We turn our discussion to the cause of variation in the impact parameter $b$ (and thus by \autoref{eq:theta}, $\theta_\mathrm{defl}$) first in simulations, and then in reality.
 
-In numerical simulations with a discretised phase space, the two primary causes of variation in $\theta_\mathrm{defl}$ are Brownian motion of the SMBH binary, and initial particle sampling variation in the merger IC, both of which have been shown to be more pronounced in lower mass resolution simulations compared to higher mass resolution simulations.
+In numerical simulations with a discretised phase space, the two primary causes of variation in $b$ are Brownian motion of the SMBH binary \citep{merritt2001}, and initial particle sampling variation in the merger IC, both of which have been shown to be more pronounced in lower mass resolution simulations compared to higher mass resolution simulations.
 By allowing the SMBH binary to wander further from its expected location in phase space, a wider range of impact parameters, and thus deflection angles, may be sampled.
@@
 
-Conversely, in the real universe, variation in the impact parameter of a similar magnitude to the presented simulations, but as a consequence of different processes, is expected due to the lack of system reproducibility.
+Conversely, in the real universe, variation in the impact parameter of a similar magnitude to the presented simulations but as a consequence of different processes is expected due to the lack of system reproducibility.
 As with many physical systems, random variations in the system state demand the use of statistics to ascertain a most probable configuration or outcome. 
-\drafting{Discussion about Sgr $A^*$ proper motion and oscillation scales? Or better example?}.
+Work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks. 
+Assuming these processes are at play in other galaxies (particularly elliptical galaxies constructed through major mergers \citep{naab2006}), deviations of a few parsecs of the SMBH from the host galaxy photo-centre are expected and would influence the impact parameter two merging SMBHs would experience.
 
@@
 \item The mergers presented in this work are all of equal mass galaxies, leading to very dramatic deformations of the potential that deviate from spherical symmetry. 
-In the instance of minor mergers where the galaxy merger remnant is not as triaxial as the major merger case \drafting{reference needed}, we do not expect to see as sensitive dependence of eccentricity on deflection angle compared to the major merger case (refer to \autoref{fig:model_orbits}).
+In the instance of minor mergers where the galaxy merger remnant is not as triaxial as the major merger case \citep{khan2012}, we do not expect to see as sensitive dependence of eccentricity on deflection angle compared to the major merger case (refer to \autoref{fig:model_orbits}).
 \end{enumerate}
@@
 By using a resolution-free toy model of the SMBH scattering process, we demonstrate that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations.
-The results presented here justify extending the investigation to more realistic galaxy merger scenarios, to quantify the expected range of hard binary eccentricities for merger timescale predictions for future gravitational wave observation missions.
+The results presented here justify extending the investigation to more realistic galaxy merger scenarios, to quantify the expected range of hard binary eccentricities critical for merger timescale predictions for future gravitational wave observation missions.
 
@@
 \textbf{MM}: Conceptualisation, Formal analysis, Writing -- original draft.
+\textbf{PHJ}: Supervision, Writing -- review \& editing.
 

### 2023-06-09 papers/paper-eccentricity/paper-eccentricity.tex
Remove DF fudge factor from toy model, other edits to toy model section

@@
     Top:
-    Sample orbits computed with the toy model for potentials with different degrees of asymmetry
-    given by the spheroid eccentricity $e_\mathrm{s}$, with otherwise identical initial conditions.
+    Sample orbits computed with the toy model for different background potential eccentricities $e_\mathrm{s}$,
+    with otherwise identical initial conditions.
     The dashed lines show the isopotential contours of the model.
@@
     Bottom:
-    The orbital eccentricity $e$ of the model orbits.
+    The orbital eccentricity $e$ of the model orbits (note the non-linear scale).
     The vertical line marks the end of the period shown in the top panels. 
-    Note the non-linear scale to allow showing the high-eccentricities reached by
-    the $e_\mathrm{s}=0.2$ model.
     }
@@
 
-To gain insight into the origin of the dependence of binary eccentricity on the 
-deflection angle $\_\mathrm{defl}$ during the binary formation,
-we have constructed a simple toy model that can qualitatively reproduce the observed
+To understand the dependency of the binary eccentricity on the 
+deflection angle $\theta_\mathrm{defl}$ during the binary formation,
+we have constructed a simple toy model that can reproduce the observed
 behavior from the simulations.
@@
 
-For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
+For simplicity, we consider only equal-mass mergers, and take the BH orbits
 to lie in the $xy$-plane.
@@
 \end{equation}
-where $\vb{x}=(x,y)$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
+where $\vb{x}=(x,y)$ is the separation vector of the BHs,
+$M_\bullet = 10^8\,\Msun$ is the mass of a single BH,
 $\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
@@
 The background galaxy is modeled as a constant density spheroidal potential,
-causing the acceleration \citep[e.g.][]{binney2008}
+with its long axis aligned along the $x$-axis.
+The acceleration from this potential is \citep[e.g.][]{binney2008}
 \begin{equation}
@@
 \end{equation}
-where $\rho$ is the stellar density,
-and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
+where $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
 \begin{align}
@@
 \end{align}
-Here we have chosen the long axis of the spheroid to be aligned with the $x$-axis.
-With eccentricity values of $e_\mathrm{s} \approx 0.9$,
-this spheroidal potential matches the central potential in the galaxy during the
-SMBH binary binding phase reasonably well,
-apart from the contributions from the BHs themselves and the groups of stars bound to them.
+The stellar density is set to the value $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
+which approximately matches the values seen in the simulations when SMBH binary is becoming bound.
 
@@
 \vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
-                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
+                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1}.
 \end{aligned}
 \end{equation}
-where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
-$X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
-and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 10$.
-To account for differences between the constant density background of stars
-assumed here and the more complex system that is present in the simulations
-we additionally multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
-with a scaling and cut-off function
+Here $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH, 
+and $X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
+where the stellar velocity dispersion is taken to have the value
+$\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}$
+to approximately match the simulations.
+The Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
+which is appropriate for the size of the simulated systems.
+
+The effect of dynamical friction weakens as the BH binary orbit shrinks,
+and to account for this we multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
+with the smooth cut-off function
 \begin{equation}
-f(a) = \frac{\alpha}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]},
+f(a) = \frac{1}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]}.
 \end{equation}
-which serves to remove the effect of dynamical friction when the SMBH binary has become
-hard in addition to allowing the strength of dynamical friction to be scaled to
-achieve similar behavior as observed in the simulations, since in the simulations the
-stellar background is much more complex.
-Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f(a)=\alpha$.
-We set the scaling factor $\alpha=0.5$, and use cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
+Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f(a)=1$.
+We use the cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
 and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
-with $a_\mathrm{h}$ being the semimajor-axis of a hard BH binary.
+with $a_\mathrm{h}$ defined by equation \eqref{eq:ahard}.
 The total dynamical friction term is then
@@
 
-To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
-we specify the initial conditions confined to the $xy$-plane as 
+To mimic the nearly linearly plunging orbits before the scattering event seen in \autoref{fig:orbit},
+we specify the initial conditions as 
 \begin{align}
-\vb{x}_0 &= (r_0, b)\\
+\vb{x}_0 &= (25\,\mathrm{pc}, b)\\
 \dot{\vb{x}}_0 &= (-v_0, 0).
 \end{align}
-We fix $r_0 = 25\,\mathrm{pc}$,
-and use model parameters
-$\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
-$\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}$,
-and
-$M_\bullet=10^8\,\Msun$,
-which are fairly similar to the central properties of the merging galaxy when
-the SMBHs are becoming bound in the $e_0=0.90$ and $e_0=0.99$ merger orbit simulations.
 Based on the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.90$ case,
@@
 The spheroid eccentricity parameter is not easily measurable from the simulations,
-and we therefore perform calculations with different values of $e_\mathrm{s}$.
+and we therefore perform calculations with different values of $e_\mathrm{s}$,
+finding a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
 We compute the resulting eccentricities for a range of impact parameters up to $b=20\,\mathrm{pc}$
@@
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
-We observe a modest decrease in deflection angle $\theta_\mathrm{defl}$ with increasing mass resolution for both the $e_0=0.90$ and $e_0=0.99$ cases. 
 To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,

### 2023-06-09 papers/paper-eccentricity/scripts/toy_model.py

@@ -32,8 +32,14 @@ class System:
         self.rho = rho
 
         self.stellar_sigma = stellar_sigma
-        self.df_fudge_factor = df_fudge_factor
-        self.logL = 10
+# XXX: The fudge factor actually just changes logL, so it would be better to
+# directly specify logL.
+# Originally they were separate factors, with logL having a fixed values.
+# But I noticed this pretty obvious thing only when all the calculations for the paper were done, 
+# so it was best not to touch the interface of the code.
+# But for future applications this should be fixed to take logL directly, or
+# better yet, to calculate it based on the BH mass, sigma and size of the system.
+        self.logL = 10 * df_fudge_factor
 
         self.a_hard = G * M_BH / (8 * stellar_sigma**2)
         self.r_infl = np.cbrt(2*M_BH/(rho*(4/3*np.pi)))
@@ -59,7 +65,7 @@ class System:
                           * (erf(X) - 2*X/np.sqrt(np.pi)*np.exp(-X**2))
                           * v_single/v_single_norm**3)
 
-        return 2*df_single * self.df_fudge_factor
+        return 2*df_single
 
     def df_decoupling_factor(self, x, v):
         a = self.semimajor_axis(x,v)
@@ -425,7 +431,7 @@ if __name__ == '__main__':
             parameter_space_scan_g05_conf3()
     else:
         #high_res_well_fitting_models()
-        gen_orbit_plot_data()
-        #test_plots()
+        #gen_orbit_plot_data()
+        test_plots()
 
 

### 2023-06-09 papers/paper-eccentricity/paper-eccentricity.tex
Edits to numerical and results sections
- some minor edits to discussion

--- 2023-06-07 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
-\author{Others}
+\author[0000-0002-7314-2558]{Thorsten Naab}
+\affiliation{Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1,
+D-85748 Garching, Germany}
 
@@
 % NUMERICAL
-\section{Numerical Methods}
+\section{Numerical Simulations}
 
@@
     \includegraphics{eccentricities.pdf}
-    \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for select representative simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
-The downwards-pointing arrows indicate the mean time when the SMBH binaries becomes hard for the two simulation sets.
-In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean.
-In the $e_0=0.90$ set, two simulations have a SMBH binary merger, as indicated by the rapid circularisation of the eccentricity.}
+    \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for the highest mass resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
+    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard for the two simulation sets. 
+    In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a comparatively low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean. 
+    In the $e_0=0.90$ set, seven simulations have a SMBH binary merger, as indicated by the rapid circularisation of the eccentricity.}
     \label{fig:eccentricities}
@@
 
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017}, implemented in \gadget{}.
-For details about the implementation of \ketju{}, we point the reader to \drafting{Mannerkoski in prep}.
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017, rantala2020,mannerkoski2023}, implemented in \gadget{}.
+For details about the implementation of \ketju{}, we point the reader to \citet{mannerkoski2023}.
 
-\subsection{Galaxy ICs}\label{ssec:gics}
-We construct our galaxy models to match those of \citet{gualandris2022}.
-Each galaxy is represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+We construct our galaxy models to represent the nuclear bulge of a gas-devoid elliptical galaxy, matching those models of \citet{nasim2020}.
+Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
 \begin{equation}\label{eq:dehnen}
@@
 The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
-We test five different mass resolutions by sampling the stellar phase space distribution with varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
+We test five different mass resolutions by sampling the stellar phase space distribution with a varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
 
-\subsection{Merger ICs}
-We construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
+We then construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
 The chosen eccentricities $e_0$ are 0.90 and 0.99.
@@
 Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
-Additionally, we run the $e_0=0.90$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described in \autoref{ssec:gics}.
-
-\subsection{Assessing Eccentricity Uncertainty}
-In our analysis, we wish to determine the population spread in eccentricity from a finite set of observations.
-We assess the uncertainty in eccentricity using an inverse modelling approach following \citet{nasim2020}. 
-For each simulation in a given set, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
-The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}
-\begin{equation}\label{eq:ahard}
-    a_\mathrm{h} = \frac{q}{(1+q)^2} \frac{r_\mathrm{m}}{4},
-\end{equation}
-where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q = \{q \in \mathbb{R} | 0<q \leq 1\}$ is the mass ratio between the SMBHs.
-Under the assumption that each calculated mean eccentricity is a sample of some common distribution, we may (assuming that the distribution is Gaussian) obtain an estimate of the standard deviation of this distribution by taking the standard deviation of the collection of sample eccentricity means, denoted $\sigma_e$.
+Additionally, we run the $e_0=0.90$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described above.
 
@@
 
-\subsection{Eccentricity scatter in low-e orbits}\label{ssec:lowe}
+\subsection{\drafting{Eccentricity scatter in simulations}}
 
-A representative set of realisations for the $e_0=0.90$ orbit is shown with blue lines in \autoref{fig:eccentricities}.
-We define a shifted time $t'$ such that $t'(0) = t_\mathrm{bound}$. 
-One simulation is seen to lie significantly beyond the mean eccentricity, with $e_\mathrm{h}\sim 0.2$.
+The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets.
+At least one simulation is seen to lie significantly beyond the mean eccentricity in each set, with $e_\mathrm{h} \sim 0.8$ for the $e_0=0.90$ set and $e_\mathrm{h}\sim 0.2$ for the $e_0=0.99$ set. 
+The median time when each simulation set forms a hard binary is similar between the two sets, as indicated by the arrows in \autoref{fig:eccentricities}.
+The $e_0=0.90$ set also demonstrates seven SMBH-binary mergers, seen as a rapid circularisation in eccentricity in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
 
-We show the dependence of eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with circle and square markers. 
-Taking for each mass resolution the full set of ten simulations (circle markers), immediately apparent is the lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
+To characterise the scatter in eccentricity, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
+The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}
+\begin{equation}\label{eq:ahard}
+    a_\mathrm{h} = \frac{q}{(1+q)^2} \frac{r_\mathrm{m}}{4} = \frac{r_\mathrm{m}}{16},
+\end{equation}
+where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$, giving rise to the second equality in \autoref{eq:ahard}).
+Assuming the distribution of mean hard-binary eccentricities is Gaussian, we characterise the inter-simulation eccentricity scatter as the standard deviation in mean values of $e_\mathrm{h}$, denoted as $\sigma_e$.
+
+We show the dependence of inter-simulation eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with circle markers, and for the $e_0=0.99$ orbit using square markers.
+The full set of ten simulations per mass resolution is shown with unfilled markers, and those simulations satisfying a $2.5\sigma$-cut in eccentricity with filled markers.
+
+For the $e_0=0.90$ orbit, using the full set of simulations results in a lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
 The peculiar values of $\sigma_e$ arise due to the assumed Gaussian distribution of the eccentricity between simulations.
-Inspecting the orbital parameters for all simulations that form the $e_0=0.90$ suite, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at both $5\times10^3$ and $1\times10^4$ resolutions which has a value of $e_\mathrm{h}<0.4$.
-This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and by applying a $2.5\sigma$-cut in eccentricity (i.e., taking those lines within the blue region of the left panel of \autoref{fig:eccentricities}) we recover the $1/\sqrt{M_\bullet/m_\star}\propto 1/\sqrt{N_\star}$ scaling of $\sigma_e$ as in \citet{nasim2020} (square markers), apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
-However, the eccentricity values beyond $2.5\sigma$ from the mean are each a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups.
+Inspecting the orbital parameters for all simulations from the $e_0=0.90$ set, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at both $5\times10^3$ and $1\times10^4$ resolutions which has a value of $e_\mathrm{h}<0.4$.
+This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and with the $2.5\sigma$-cut data we recover the $1/\sqrt{N_\star}$ scaling of $\sigma_e$, confirming the results of \citet{nasim2020}, apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
+The variation in eccentricity also clearly decreases for the cleaned $e_0=0.99$ orbit simulations, albeit with a steeper gradient than the $e_0=0.90$ set. 
+Again, a single simulation at the highest mass resolution for the $e_0=0.99$ set is responsible for the inflated value of $\sigma_e$ in the full data set for this resolution.
+
+Critically, the eccentricity values beyond $2.5\sigma$ from the mean that are removed in our cleaned samples are each still a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups with a highly-radial initial orbit.
 The cause of these low eccentricity values is discussed in \autoref{ssec:toy}.
-
-\subsection{Eccentricity scatter in high-e orbits}\label{ssec:highe}
-
-Similar to the $e_0=0.90$ case, we show a representative set of realisation for the $e_0=0.99$ orbit with orange lines in \autoref{fig:eccentricities}.
-Again, one simulation is seen to have $e_\mathrm{h} \sim 0.1$, significantly offset from the mean value of $e_\mathrm{h}$ for this simulation set.
-
-The dependence of eccentricity standard deviation on mass resolution for the $e_0=0.99$ orbit is shown in \autoref{fig:convergence} with triangle and diamond markers. 
-We find that the scaling of $\sigma_e$ is steeper than the $1/\sqrt{N_\star}$ relation described in \citet{nasim2020}.
-We attribute this dramatic decrease in $\sigma_e$ to an inflated value of $\sigma_e$ at low mass resolution: there is increased scatter in eccentricity at low mass resolutions for very radial initial orbits of the SMBH binary.
-As the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar2006}, understanding if the origin of the scatter in eccentricity is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
 
@@
 As shown by the simulation points and the marginal kernel density estimates (KDEs) of \autoref{fig:theta_e}, the variation in eccentricity can span almost the full domain space of $[0,1]$.
+Galaxy mergers which have a more radial initial orbit $e_0$ are expected to demonstrate increased variance in eccentricity \citep{gualandris2022} \drafting{other refs needed here}.
+However, as the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar2006}, understanding if the origin of the eccentricity scatter is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
+We thus turn our discussion to the cause of variation in the impact parameter $b$ (and hence by \autoref{eq:theta}, $\theta_\mathrm{defl}$) first in simulations, and then in reality, as a predictor of the eccentricity variance.
+\drafting{Mention that we see qualitatively similar results for more realistic density profiles (e.g. the Hernquist runs).}
 
-We turn our discussion to the cause of variation in the impact parameter $b$ (and thus by \autoref{eq:theta}, $\theta_\mathrm{defl}$) first in simulations, and then in reality.
-
-In numerical simulations with a discretised phase space, the two primary causes of variation in $b$ are Brownian motion of the SMBH binary \citep{merritt2001}, and initial particle sampling variation in the merger IC, both of which have been shown to be more pronounced in lower mass resolution simulations compared to higher mass resolution simulations.
+In numerical simulations with a discretised phase space, the two primary causes of variation in $b$ are initial particle sampling variation in the merger IC and Brownian motion of the SMBH binary \citep{merritt2001}, both of which have been shown to be more pronounced in lower mass resolution simulations compared to higher mass resolution simulations.
 By allowing the SMBH binary to wander further from its expected location in phase space, a wider range of impact parameters, and thus deflection angles, may be sampled.
@@
 % FINAL BITS
-\textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2022}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
-
 \section*{Acknowledgements}
@@
 \textbf{PHJ}: Supervision, Writing -- review \& editing.
+
+\textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2022}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
 

### 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
Some more edits to toy model section

--- 2023-06-09 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
-To understand the dependency of the binary eccentricity on the 
-deflection angle $\theta_\mathrm{defl}$ during the binary formation,
-we have constructed a simple toy model that can reproduce the observed
-behavior from the simulations.
-This model includes only the essential components relevant to this process,
-which are the asymmetric background stellar potential,
-dynamical friction from the stellar background,
-and naturally the SMBHs themselves.
-
-For simplicity, we consider only equal-mass mergers, and take the BH orbits
-to lie in the $xy$-plane.
-In this case the relative motion of the BHs can be treated as a one-body problem,
-with the equation of motion
+The dependency of the binary eccentricity on the 
+deflection angle $\theta_\mathrm{defl}$ during the binary formation
+can be reproduced using a simple toy model,
+which includes only the essential components relevant to this process.
+Taking the orbit to lie in the $xy$-plane,
+the relative motion of the two equal-mass SMBHs in this model is follows the 
+equation of motion
 \begin{equation} \label{eq:toy_model_eom}
@@
 \end{equation}
-where $\vb{x}=(x,y)$ is the separation vector of the BHs,
-$M_\bullet = 10^8\,\Msun$ is the mass of a single BH,
-$\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
+where $\vb{x}=(x,y)$ is the separation vector of the SMBHs,
+$M_\bullet = 10^8\,\Msun$ is the mass of a single SMBH,
+$\vb{a}_\mathrm{bg}$ the acceleration due to the asymmetric background potential,
 and $\vb{a}_\mathrm{DF}$ the acceleration due to dynamical friction.
 
-The background galaxy is modeled as a constant density spheroidal potential,
-with its long axis aligned along the $x$-axis.
-The acceleration from this potential is \citep[e.g.][]{binney2008}
+The stellar background is modeled as a constant density spheroidal potential
+\citep[e.g.][]{binney2008}
 \begin{equation}
-\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y) \vb{x},
+\Phi_\mathrm{bg}(\vb{x}) = \pi G \rho (A_x x^2 + A_y y^2),
 \end{equation}
-where $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
+with the acceleration given by 
+\begin{equation}
+\vb{a}_\mathrm{bg}(\vb{x}) = - \nabla \Phi_\mathrm{bg}(\vb{x}).
+\end{equation}
+The $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
 \begin{align}
@@
         \left[ \frac{1}{1-e_\mathrm{s}^2} 
-        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
+        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right],
 \end{align}
-The stellar density is set to the value $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
+with the long axis aligned along the $x$-axis.
+The stellar density is set to $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
 which approximately matches the values seen in the simulations when SMBH binary is becoming bound.
 
-For dynamical friction, we use the \citet{chandrasekhar1943} formula assuming
-a Maxwellian distribution with a constant velocity dispersion $\sigma_\mathrm{M}$ \citep[e.g.][]{binney2008}.
+Dynamical friction is modeled using the \citet{chandrasekhar1943} formula assuming
+a Maxwellian distribution with a constant velocity dispersion
+${\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}}$,
+set based on the value measured from the simulations.
 This formula gives the acceleration of a single BH due to dynamical friction as
+\citep[e.g.][]{binney2008}
 \begin{equation}\label{eq:chandra_df}
@@
 Here $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH, 
-and $X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
-where the stellar velocity dispersion is taken to have the value
-$\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}$
-to approximately match the simulations.
-The Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
+$X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
+and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
 which is appropriate for the size of the simulated systems.
 
-The effect of dynamical friction weakens as the BH binary orbit shrinks,
-and to account for this we multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
+The effect of dynamical friction weakens as the SMBH binary orbit shrinks.
+To account for this, when the binary is bound 
+we multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
 with the smooth cut-off function
@@
 \end{equation}
-Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f(a)=1$.
-We use the cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
+Here $a$ is the semi-major axis of a bound binary,
+and the cut-off scales are $a_\mathrm{c}=2 a_\mathrm{h}$
 and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
@@
 \end{align}
-Based on the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.90$ case,
+Based on the simulations, we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
+typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
 and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
 The spheroid eccentricity parameter is not easily measurable from the simulations,
-and we therefore perform calculations with different values of $e_\mathrm{s}$,
+due to stellar components that remain bound to the SMBHs and obscure the background
+potential relevant for the dynamics,
+as well as due to the potential in the simulations being constantly evolving.
+We therefore perform calculations with different values of $e_\mathrm{s}$,
 finding a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
+
 We compute the resulting eccentricities for a range of impact parameters up to $b=20\,\mathrm{pc}$
@@
 the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}.
-The range of impact parameters covers most of the relevant parameter space,
-as a $90\degr$ deflection during the first pericenter passage
+For reference, a $90\degr$ deflection during the first pericenter passage
 corresponds to initial impact parameters of $b\approx 5.8\,\mathrm{pc}$ and
@@
 while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
-Otherwise the model appears to give a very good match to the simulation results,
-and the required slight shift could be due to the galactic potential being slightly tilted in the simulations 
-compared to the BH trajectories due to the complex dynamics at this stage.
+These shifts are likely required due to the rotation of the stellar component tilting the background potential
+relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case
+due to the less radial merger orbit.
 In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
@@
 On the other hand, in the more symmetric potential with $e_\mathrm{s}=0.2$,
-The orbit shape is not very sensitive to the deflection angle due to the symmetry of the system,
+the orbit shape is not very sensitive to the deflection angle due to the symmetry of the system,
 and the binary always reaches extremely high eccentricities.

### 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex

--- 2023-06-09 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
-To gain insight into the origin of the dependence of binary eccentricity on the 
-deflection angle $\_\mathrm{defl}$ during the binary formation,
-we have constructed a simple toy model that can qualitatively reproduce the observed
-behavior from the simulations.
-This model includes only the essential components relevant to this process,
-which are the asymmetric background stellar potential,
-dynamical friction from the stellar background,
-and naturally the SMBHs themselves.
-
-For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
-to lie in the $xy$-plane.
-In this case the relative motion of the BHs can be treated as a one-body problem,
-with the equation of motion
+The dependency of the binary eccentricity on the 
+deflection angle $\theta_\mathrm{defl}$ during the binary formation
+can be reproduced using a simple toy model,
+which includes only the essential components relevant to this process.
+Taking the orbit to lie in the $xy$-plane,
+the relative motion of the two equal-mass SMBHs in this model is follows the 
+equation of motion
 \begin{equation} \label{eq:toy_model_eom}
@@
 \end{equation}
-where $\vb{x}=(x,y)$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
-$\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
+where $\vb{x}=(x,y)$ is the separation vector of the SMBHs,
+$M_\bullet = 10^8\,\Msun$ is the mass of a single SMBH,
+$\vb{a}_\mathrm{bg}$ the acceleration due to the asymmetric background potential,
 and $\vb{a}_\mathrm{DF}$ the acceleration due to dynamical friction.
 
-The background galaxy is modeled as a constant density spheroidal potential,
-causing the acceleration \citep[e.g.][]{binney2008}
+The stellar background is modeled as a constant density spheroidal potential
+\citep[e.g.][]{binney2008}
 \begin{equation}
-\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y) \vb{x},
+\Phi_\mathrm{bg}(\vb{x}) = \pi G \rho (A_x x^2 + A_y y^2),
 \end{equation}
-where $\rho$ is the stellar density,
-and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
+with the acceleration given by 
+\begin{equation}
+\vb{a}_\mathrm{bg}(\vb{x}) = - \nabla \Phi_\mathrm{bg}(\vb{x}).
+\end{equation}
+The $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
 \begin{align}
@@
         \left[ \frac{1}{1-e_\mathrm{s}^2} 
-        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
+        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right],
 \end{align}
-Here we have chosen the long axis of the spheroid to be aligned with the $x$-axis.
-With eccentricity values of $e_\mathrm{s} \approx 0.9$,
-this spheroidal potential matches the central potential in the galaxy during the
-SMBH binary binding phase reasonably well,
-apart from the contributions from the BHs themselves and the groups of stars bound to them.
+with the long axis aligned along the $x$-axis.
+The stellar density is set to $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
+which approximately matches the values seen in the simulations when SMBH binary is becoming bound.
 
-For dynamical friction, we use the \citet{chandrasekhar1943} formula assuming
-a Maxwellian distribution with a constant velocity dispersion $\sigma_\mathrm{M}$ \citep[e.g.][]{binney2008}.
+Dynamical friction is modeled using the \citet{chandrasekhar1943} formula assuming
+a Maxwellian distribution with a constant velocity dispersion
+${\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}}$,
+set based on the value measured from the simulations.
 This formula gives the acceleration of a single BH due to dynamical friction as
+\citep[e.g.][]{binney2008}
 \begin{equation}\label{eq:chandra_df}
@@
 \vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
-                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
+                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1}.
 \end{aligned}
 \end{equation}
-where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
+Here $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH, 
 $X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
-and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 10$.
-To account for differences between the constant density background of stars
-assumed here and the more complex system that is present in the simulations
-we additionally multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
-with a scaling and cut-off function
+and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
+which is appropriate for the size of the simulated systems.
+
+The effect of dynamical friction weakens as the SMBH binary orbit shrinks.
+To account for this, when the binary is bound 
+we multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
+with the smooth cut-off function
 \begin{equation}
-f(a) = \frac{\alpha}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]},
+f(a) = \frac{1}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]}.
 \end{equation}
-which serves to remove the effect of dynamical friction when the SMBH binary has become
-hard in addition to allowing the strength of dynamical friction to be scaled to
-achieve similar behavior as observed in the simulations, since in the simulations the
-stellar background is much more complex.
-Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f(a)=\alpha$.
-We set the scaling factor $\alpha=0.5$, and use cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
+Here $a$ is the semi-major axis of a bound binary,
+and the cut-off scales are $a_\mathrm{c}=2 a_\mathrm{h}$
 and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
-with $a_\mathrm{h}$ being the semimajor-axis of a hard BH binary.
+with $a_\mathrm{h}$ defined by equation \eqref{eq:ahard}.
 The total dynamical friction term is then
@@
 
-To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
-we specify the initial conditions confined to the $xy$-plane as 
+To mimic the nearly linearly plunging orbits before the scattering event seen in \autoref{fig:orbit},
+we specify the initial conditions as 
 \begin{align}
-\vb{x}_0 &= (r_0, b)\\
+\vb{x}_0 &= (25\,\mathrm{pc}, b)\\
 \dot{\vb{x}}_0 &= (-v_0, 0).
 \end{align}
-We fix $r_0 = 25\,\mathrm{pc}$,
-and use model parameters
-$\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
-$\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}$,
-and
-$M_\bullet=10^8\,\Msun$,
-which are fairly similar to the central properties of the merging galaxy when
-the SMBHs are becoming bound in the $e_0=0.90$ and $e_0=0.99$ merger orbit simulations.
-Based on the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.90$ case,
+Based on the simulations, we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
+typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
 and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
 The spheroid eccentricity parameter is not easily measurable from the simulations,
-and we therefore perform calculations with different values of $e_\mathrm{s}$.
+due to stellar components that remain bound to the SMBHs and obscure the background
+potential relevant for the dynamics,
+as well as due to the potential in the simulations being constantly evolving.
+We therefore perform calculations with different values of $e_\mathrm{s}$,
+finding a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
+
 We compute the resulting eccentricities for a range of impact parameters up to $b=20\,\mathrm{pc}$
@@
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
-We observe a modest decrease in deflection angle $\theta_\mathrm{defl}$ with increasing mass resolution for both the $e_0=0.90$ and $e_0=0.99$ cases. 
 To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
 while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
-Otherwise the model appears to give a very good match to the simulation results,
-and the required slight shift could be due to the galactic potential being slightly tilted in the simulations 
-compared to the BH trajectories due to the complex dynamics at this stage.
+These shifts are likely required due to the rotation of the stellar component tilting the background potential
+relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case
+due to the less radial merger orbit.
 In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around

### 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
Edit sec 3.3 some more

@@
 and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
-The spheroid eccentricity parameter is not easily measurable from the simulations,
+The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measurable from the simulations,
 due to stellar components that remain bound to the SMBHs and obscure the background
@@
 as well as due to the potential in the simulations being constantly evolving.
-We therefore perform calculations with different values of $e_\mathrm{s}$,
-finding a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
+However, only sufficiently large values of $e_\mathrm{s}$ allow for low binary
+eccentricities to be produced,
+as is shown by \autoref{fig:model_orbits}.
+Performing calculations with different values of $e_\mathrm{s}$,
+we found a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
 
@@
 
-Overall we can however see that this simple model correctly captures the behavior of eccentricity seen in the simulations.
-We can therefore use the model to understand the cause of the complicated dependence of the binary eccentricity
-on the deflection angle $\theta_\mathrm{defl}$.
-\autoref{fig:model_orbits} shows sets of orbits computed with the model for two different potential shapes,
-with the parameters otherwise as in the model that matches the $e_0=0.90$ simulation results.
-In the $e_\mathrm{s}=0.9$ case the shape of the orbit after the initial scattering depends strongly on the deflection angle,
-as the asymmetric potential causes torque on the SMBHs that are not moving along the main axes of the potential.
-On the other hand, in the more symmetric potential with $e_\mathrm{s}=0.2$,
-the orbit shape is not very sensitive to the deflection angle due to the symmetry of the system,
-and the binary always reaches extremely high eccentricities.
+Otherwise the behavior of the simulation data is captured well,
+with the model correctly producing the two main eccentricity minima,
+as well as the $e\approx 1$ region between them.
+The minima occur for trajectories where the torque from the background potential together with the dynamical friction
+causes the SMBHs to loop around into a nearly circular orbit,
+while highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along the 
+the main axes of the potential, as is seen in \autoref{fig:model_orbits}.
 

### 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
- minor edits to numerics section

--- 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
@@
     In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a comparatively low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean. 
-    In the $e_0=0.90$ set, seven simulations have a SMBH binary merger, as indicated by the rapid circularisation of the eccentricity.}
+    In the $e_0=0.90$ set, seven simulations have a SMBH binary merger within $50\,\mathrm{Myr}$, as indicated by the rapid circularisation of the eccentricity.}
     \label{fig:eccentricities}
@@
 
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017, rantala2020,mannerkoski2023}, implemented in \gadget{}.
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017, rantala2020,mannerkoski2023}, implemented in \gadget{} \citep{springel2021}.
 For details about the implementation of \ketju{}, we point the reader to \citet{mannerkoski2023}.
@@
 We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022}.
-For each orbital configuration, we run ten realisations, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space. 
-Each orbital geometry is simulated using our fiducial particle resolution of $M_\bullet/m_\star = 10^4$.
-Additionally, we run the $e_0=0.90$ and $e_0=0.99$ orbits at varying mass resolutions, using the galaxy ICs at the mass resolutions described above.
+For each orbital configuration, we run ten realisations for each mass resolution, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space.
 
@@
     \includegraphics{convergence}
-    \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the total number of stellar particles $2N_\star$ for initial eccentricities $e_0=0.90$ and $e_0=0.99$.
+    \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the number of stellar particles per progenitor galaxy $N_\star$ for initial eccentricities $e_0=0.90$ and $e_0=0.99$.
     The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ suite after applying a $2.5\sigma$-cut to the simulation data. 
@@
 The median time when each simulation set forms a hard binary is similar between the two sets, as indicated by the arrows in \autoref{fig:eccentricities}.
-The $e_0=0.90$ set also demonstrates seven SMBH-binary mergers, seen as a rapid circularisation in eccentricity in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
+The $e_0=0.90$ set also demonstrates seven SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid circularisation in eccentricity in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
 
@@
     \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is the same as the second BH orbit reflected.
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about the $x=-z$ line.
     Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity of the deflection-phase of the evolution.
@@
     Top:
-    Sample orbits computed with the toy model for different background potential eccentricities $e_\mathrm{s}$,
-    with otherwise identical initial conditions.
+    Sample orbits computed with the toy model for potentials with different degrees of asymmetry
+    given by the spheroid eccentricity $e_\mathrm{s}$, with otherwise identical initial conditions.
     The dashed lines show the isopotential contours of the model.
@@
     Bottom:
-    The orbital eccentricity $e$ of the model orbits (note the non-linear scale).
+    The orbital eccentricity $e$ of the model orbits.
     The vertical line marks the end of the period shown in the top panels. 
+    Note the non-linear scale to allow showing the high-eccentricities reached by
+    the $e_\mathrm{s}=0.2$ model.
     }
@@
 
-The dependency of the binary eccentricity on the 
-deflection angle $\theta_\mathrm{defl}$ during the binary formation
-can be reproduced using a simple toy model,
-which includes only the essential components relevant to this process.
-Taking the orbit to lie in the $xy$-plane,
-the relative motion of the two equal-mass SMBHs in this model is follows the 
-equation of motion
+To gain insight into the origin of the dependence of binary eccentricity on the 
+deflection angle $\theta_\mathrm{defl}$ during the binary formation,
+we have constructed a simple toy model that can qualitatively reproduce the observed
+behavior from the simulations.
+This model includes only the essential components relevant to this process,
+which are the asymmetric background stellar potential,
+dynamical friction from the stellar background,
+and naturally the SMBHs themselves.
+
+For simplicity, we consider only symmetrical 1:1 mergers, and take the BH orbits
+to lie in the $xy$-plane.
+In this case the relative motion of the BHs can be treated as a one-body problem,
+with the equation of motion
 \begin{equation} \label{eq:toy_model_eom}
@@
 \end{equation}
-where $\vb{x}=(x,y)$ is the separation vector of the SMBHs,
-$M_\bullet = 10^8\,\Msun$ is the mass of a single SMBH,
-$\vb{a}_\mathrm{bg}$ the acceleration due to the asymmetric background potential,
+where $\vb{x}=(x,y)$ is the separation vector of the BHs, $M_\bullet$ the mass of a single BH,
+$\vb{a}_\mathrm{bg}$ the acceleration due to the background potential,
 and $\vb{a}_\mathrm{DF}$ the acceleration due to dynamical friction.
 
-The stellar background is modeled as a constant density spheroidal potential
-\citep[e.g.][]{binney2008}
+The background galaxy is modeled as a constant density spheroidal potential,
+causing the acceleration \citep[e.g.][]{binney2008}
 \begin{equation}
-\Phi_\mathrm{bg}(\vb{x}) = \pi G \rho (A_x x^2 + A_y y^2),
+\vb{a}_\mathrm{bg} = - 2 \pi G \rho \operatorname{diag}(A_x,A_y) \vb{x},
 \end{equation}
-with the acceleration given by 
-\begin{equation}
-\vb{a}_\mathrm{bg}(\vb{x}) = - \nabla \Phi_\mathrm{bg}(\vb{x}).
-\end{equation}
-The $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
+where $\rho$ is the stellar density,
+and the $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
 \begin{align}
@@
         \left[ \frac{1}{1-e_\mathrm{s}^2} 
-        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right],
+        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right].
 \end{align}
-with the long axis aligned along the $x$-axis.
-The stellar density is set to $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
-which approximately matches the values seen in the simulations when SMBH binary is becoming bound.
+Here we have chosen the long axis of the spheroid to be aligned with the $x$-axis.
+With eccentricity values of $e_\mathrm{s} \approx 0.9$,
+this spheroidal potential matches the central potential in the galaxy during the
+SMBH binary binding phase reasonably well,
+apart from the contributions from the BHs themselves and the groups of stars bound to them.
 
-Dynamical friction is modeled using the \citet{chandrasekhar1943} formula assuming
-a Maxwellian distribution with a constant velocity dispersion
-${\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}}$,
-set based on the value measured from the simulations.
+For dynamical friction, we use the \citet{chandrasekhar1943} formula assuming
+a Maxwellian distribution with a constant velocity dispersion $\sigma_\mathrm{M}$ \citep[e.g.][]{binney2008}.
 This formula gives the acceleration of a single BH due to dynamical friction as
-\citep[e.g.][]{binney2008}
 \begin{equation}\label{eq:chandra_df}
@@
 \vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
-                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1}.
+                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
 \end{aligned}
 \end{equation}
-Here $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH, 
+where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
 $X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
-and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
-which is appropriate for the size of the simulated systems.
-
-The effect of dynamical friction weakens as the SMBH binary orbit shrinks.
-To account for this, when the binary is bound 
-we multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
-with the smooth cut-off function
+and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 10$.
+To account for differences between the constant density background of stars
+assumed here and the more complex system that is present in the simulations
+we additionally multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
+with a scaling and cut-off function
 \begin{equation}
-f(a) = \frac{1}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]}.
+f(a) = \frac{\alpha}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]},
 \end{equation}
-Here $a$ is the semi-major axis of a bound binary,
-and the cut-off scales are $a_\mathrm{c}=2 a_\mathrm{h}$
+which serves to remove the effect of dynamical friction when the SMBH binary has become
+hard in addition to allowing the strength of dynamical friction to be scaled to
+achieve similar behavior as observed in the simulations, since in the simulations the
+stellar background is much more complex.
+Here $a$ is the semi-major axis of a bound binary, and for unbound binaries we set $f(a)=\alpha$.
+We set the scaling factor $\alpha=0.5$, and use cut-off scales $a_\mathrm{c}=2 a_\mathrm{h}$
 and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
-with $a_\mathrm{h}$ defined by equation \eqref{eq:ahard}.
+with $a_\mathrm{h}$ being the semimajor-axis of a hard BH binary.
 The total dynamical friction term is then
@@
 
-To mimic the nearly linearly plunging orbits before the scattering event seen in \autoref{fig:orbit},
-we specify the initial conditions as 
+To mimic the nearly linearly plunging orbit in the high-eccentricity merger orbit simulations,
+we specify the initial conditions confined to the $xy$-plane as 
 \begin{align}
-\vb{x}_0 &= (25\,\mathrm{pc}, b)\\
+\vb{x}_0 &= (r_0, b)\\
 \dot{\vb{x}}_0 &= (-v_0, 0).
 \end{align}
-Based on the simulations, we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
-typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
+We fix $r_0 = 25\,\mathrm{pc}$,
+and use model parameters
+$\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
+$\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}$,
+and
+$M_\bullet=10^8\,\Msun$,
+which are fairly similar to the central properties of the merging galaxy when
+the SMBHs are becoming bound in the $e_0=0.90$ and $e_0=0.99$ merger orbit simulations.
+Based on the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.90$ case,
 and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
 The spheroid eccentricity parameter is not easily measurable from the simulations,
-due to stellar components that remain bound to the SMBHs and obscure the background
-potential relevant for the dynamics,
-as well as due to the potential in the simulations being constantly evolving.
-We therefore perform calculations with different values of $e_\mathrm{s}$,
-finding a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
-
+and we therefore perform calculations with different values of $e_\mathrm{s}$.
 We compute the resulting eccentricities for a range of impact parameters up to $b=20\,\mathrm{pc}$
@@
 the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}.
-For reference, a $90\degr$ deflection during the first pericenter passage
+The range of impact parameters covers most of the relevant parameter space,
+as a $90\degr$ deflection during the first pericenter passage
 corresponds to initial impact parameters of $b\approx 5.8\,\mathrm{pc}$ and
@@
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
+We observe a modest decrease in deflection angle $\theta_\mathrm{defl}$ with increasing mass resolution for both the $e_0=0.90$ and $e_0=0.99$ cases. 
 To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
 while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
-These shifts are likely required due to the rotation of the stellar component tilting the background potential
-relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case
-due to the less radial merger orbit.
+Otherwise the model appears to give a very good match to the simulation results,
+and the required slight shift could be due to the galactic potential being slightly tilted in the simulations 
+compared to the BH trajectories due to the complex dynamics at this stage.
 In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
@@
 On the other hand, in the more symmetric potential with $e_\mathrm{s}=0.2$,
-the orbit shape is not very sensitive to the deflection angle due to the symmetry of the system,
+The orbit shape is not very sensitive to the deflection angle due to the symmetry of the system,
 and the binary always reaches extremely high eccentricities.
@@
 
-\textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2022}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
+\textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2021}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
 

### 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
Rewrite discussion

--- 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
@@
 causes the SMBHs to loop around into a nearly circular orbit,
-while highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along the 
+while highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along 
 the main axes of the potential, as is seen in \autoref{fig:model_orbits}.
@@
 
-We show in \autoref{sec:results} that variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
-The inclusion of the toy model allows us to investigate the relationship between $\theta_\mathrm{defl}$ and $e_\mathrm{h}$ in the infinite-mass resolution limit, around which the quantities measured from the simulation demonstrate minimal scatter.
-As shown by the simulation points and the marginal kernel density estimates (KDEs) of \autoref{fig:theta_e}, the variation in eccentricity can span almost the full domain space of $[0,1]$.
-Galaxy mergers which have a more radial initial orbit $e_0$ are expected to demonstrate increased variance in eccentricity \citep{gualandris2022} \drafting{other refs needed here}.
-However, as the majority of major galaxy mergers in the real Universe are expected to be radial \citep[e.g.][]{khochfar2006}, understanding if the origin of the eccentricity scatter is a numerical or a physical effect is of paramount importance, and will have direct implications for the predictions of SMBH merger rates by ongoing and future GW detection missions.
-We thus turn our discussion to the cause of variation in the impact parameter $b$ (and hence by \autoref{eq:theta}, $\theta_\mathrm{defl}$) first in simulations, and then in reality, as a predictor of the eccentricity variance.
-\drafting{Mention that we see qualitatively similar results for more realistic density profiles (e.g. the Hernquist runs).}
+Our results show that the variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
+Together with the simple toy model that reproduces well the behavior seen in the simulations,
+this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic mergers on highly radial orbits,
+at least in major mergers where the SMBHs have nearly equal masses,
+and not an error due to insufficient numerical resolution.
 
-In numerical simulations with a discretised phase space, the two primary causes of variation in $b$ are initial particle sampling variation in the merger IC and Brownian motion of the SMBH binary \citep{merritt2001}, both of which have been shown to be more pronounced in lower mass resolution simulations compared to higher mass resolution simulations.
-By allowing the SMBH binary to wander further from its expected location in phase space, a wider range of impact parameters, and thus deflection angles, may be sampled.
-This consequently maps to a different sampling of the eccentricity space, however due to the non-linear mapping (as shown in \autoref{fig:theta_e}), the eccentricity domain space sampled can vary between $\sigma_e \sim 0.01$ to $\sigma_e>0.1$, for the shown system configurations.
-Different systems, with different galaxy density profiles or merger orbits, may have an even more pronounced value of $\sigma_e$.
+As shown by both the simulations and the toy model, in these mergers the eccentricity of the SMBH binary can span nearly the full possible range between 0 and 1,
+depending on comparatively tiny differences on the scale of a parsec in the exact realization of the galactic merger orbit.
+On the other hand, in minor mergers where there is less asymmetry in the stellar background when the SMBHs become bound,
+the scatter in eccentricity is likely to be low even in the case of a radial merger orbit, since the system is no longer sensitive to the exact value of the deflection angle,
+or the impact parameter, as can be seen from the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
 
-Conversely, in the real universe, variation in the impact parameter of a similar magnitude to the presented simulations but as a consequence of different processes is expected due to the lack of system reproducibility.
-As with many physical systems, random variations in the system state demand the use of statistics to ascertain a most probable configuration or outcome. 
-Work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks. 
-Assuming these processes are at play in other galaxies (particularly elliptical galaxies constructed through major mergers \citep{naab2006}), deviations of a few parsecs of the SMBH from the host galaxy photo-centre are expected and would influence the impact parameter two merging SMBHs would experience.
+The simulation setup used in this work is highly idealized,
+and covers only a small part of the parameter space of merger configurations since we only consider fairly high-eccentricity galactic merger orbits.
+However, similar behavior can be expected to occur also in more realistic models,
+since the relevant dynamics occurs within few hundred parsecs from the center of the merged galaxy,
+where we do not expect any significant  effects from dark matter or the outer parts of the galaxy, which are not included in our simulations.
+Lower eccentricity merger orbits can be expected to show less scatter in the binary eccentricity due to the lack of a hard scattering event that is sensitive to slight perturbations in the merger orbit,
+and this is also seen in simulation results \citep{gualandris2022}.
+However, most major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006},
+and the dynamical friction from the dark matter halo would generally drive even lower eccentricity merger orbits to highly radial final plunges.
 
-The results presented in this work are significant in constraining observational quantities that are correlated with binary eccentricity: most notably SMBH binary merger timescale. 
-As the relevant scales of the impact parameter that lead to the deflection angle are typically of the order of a few parsecs, and are thus generally unresolvable in gravitationally-softened simulations, numerical simulations sample instead a non-deterministic distribution of potential impact parameters.
-Building on from work by \citet{portegieszwart2018}, microscopic perturbations can grow to large scale, observable effects in a gravitational system over relatively short timescales, suggesting that even at physical mass resolution, chaotic dynamics still lead to a non-deterministic sampling of the impact parameter domain.
-From \autoref{fig:theta_e}, different initial orbital geometries also influence the deflection-angle domain space sampled, with not all merger geometries necessarily able to sample the full eccentricity domain (neglecting minor scatter in $e_\mathrm{h}$ due to sampling variation compared to the analytical prediction).
-Importantly however, the sampling variation in $e_\mathrm{h}$ by the numerical simulations are largely independent of mass resolution, indicating that even at moderate mass resolutions of $M_\bullet/m_\star \sim 10^3$, the eccentricity distribution can be reliably sampled.
+In numerical simulations with a discretised phase space, the two primary causes of variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001}.
+Both of these effects scale as $1/\sqrt{N_\star}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
+is approximately linear.
+The deviation from this scaling seen in \autoref{fig:convergence}, particularly when not excluding outlying data points, is then simply explained by the fact that the eccentricity is not globally a linear function of the impact parameter.
 
-Finally, we discuss the caveats of our conclusions.
-\begin{enumerate}
-\item The systems we test in this work are exceptionally idealistic: spherical stellar bulges without dark matter (DM) on a merger orbit with very small initial separation. 
-We do not however expect the conclusions concerning the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ mapping to be significantly impacted by expanding our models to more realistic gravitational systems, as DM acts to provide a dynamical friction force over separation scales much larger than that of the impact parameter.
-Qualitatively, the scatter in eccentricity presented in this work is consistent with the eccentricity scatter in simulations including DM. 
-\item The mergers presented in this work are all of equal mass galaxies, leading to very dramatic deformations of the potential that deviate from spherical symmetry. 
-In the instance of minor mergers where the galaxy merger remnant is not as triaxial as the major merger case \citep{khan2012}, we do not expect to see as sensitive dependence of eccentricity on deflection angle compared to the major merger case (refer to \autoref{fig:model_orbits}).
-\end{enumerate}
+While the scatter in the impact parameter between different numerical realizations of the same merger is due to relatively low number of particles compared to real galaxies,
+we expect that uncertainty of a similar magnitude is also present in real systems due various mechanisms, such as perturbations from their environment.
+The SMBHs are likely also not located exactly at the center of mass of the galaxy,
+as e.g.\ work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
+It is therefore evident that it is not possible to give an exact prediction for the SMBH binary eccentricity produced by a given merger,
+and instead we must focus on predicting the distribution of eccentricities that can be produced by a given merger configuration.
+Such predictions are crucial for constraining observational quantities that are correlated with binary eccentricity,
+most notably SMBH binary merger timescale and the spectrum of GWs at nanohertz frequencies observable with PTAs \drafting{(refs)}.
+
+Predicting the eccentricity distribution requires knowing the distribution of impact parameters that can result from essentially identical mergers
+as well as the mapping between the impact parameter, or deflection angle, and the final eccentricity.
+Since the relation between the impact parameter and the eccentricity does not appear to significantly depend on the resolution,
+simulations at moderate mass resolutions of $M_\bullet/m_\star \sim 10^3$ can be used
+to map out the relation by performing a large enough number of merger simulations,
+possibly augmented by more sophisticated versions of the toy model presented here.
+The distribution of impact parameters in real systems is a much more difficult problem to tackle,
+since in simulations numerical resolution effects are likely dominant.
+In any case, already from the results shown here it is clear that the distribution has a complex structure.
 
 In conclusion, we have shown the importance of the deflection angle $\theta_\mathrm{defl}$ on the resulting hard binary eccentricity $e_\mathrm{h}$. 
-By using a resolution-free toy model of the SMBH scattering process, we demonstrate that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations.
-The results presented here justify extending the investigation to more realistic galaxy merger scenarios, to quantify the expected range of hard binary eccentricities critical for merger timescale predictions for future gravitational wave observation missions.
+By using a resolution-free toy model of the SMBH scattering process,
+we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations,
+but is rather due to the physical sensitivity of the system to small changes in the merger orbit,
+which can be caused by physical mechanisms in addition to numerical discretisation effects.
+The results presented here justify extending the investigation to more realistic galaxy merger scenarios,
+to quantify the expected range of hard binary eccentricities critical for merger timescale predictions for future gravitational wave observation missions.
 

### 2023-06-12 papers/paper-eccentricity/paper-eccentricity.tex
Discussion about Nasim orbit flips, still drafting

@@
 the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}.
-For reference, a $90\degr$ deflection during the first pericenter passage
+For reference, a $90\degr$ deflection during the first pericentre passage
 corresponds to initial impact parameters of $b\approx 5.8\,\mathrm{pc}$ and
@@
 As shown by both the simulations and the toy model, in these mergers the eccentricity of the SMBH binary can span nearly the full possible range between 0 and 1,
-depending on comparatively tiny differences on the scale of a parsec in the exact realization of the galactic merger orbit.
-On the other hand, in minor mergers where there is less asymmetry in the stellar background when the SMBHs become bound,
-the scatter in eccentricity is likely to be low even in the case of a radial merger orbit, since the system is no longer sensitive to the exact value of the deflection angle,
+depending on comparatively tiny differences on the scale of a parsec in the particular realization of the galactic merger orbit.
+\drafting{
+    This observation is at odds with \citet{nasim2021}, where binaries that formed at high eccentricity tended to proceed to higher eccentricities.
+    However, the tension can be explained if the simulations of \citet{nasim2021} lie in the linear region of the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ parameter space (where most of the simulations in this work lie), coupled with the low-number statistics of \citet{nasim2021}.
+    This would suggest that the potential of the merger remnant, as opposed to the rotation of the SMBH binary relative to the stellar background, is the dominant predictor of hard binary eccentricity.
+}
+
+Compared to major mergers, in minor mergers where there is less asymmetry in the stellar background when the SMBHs become bound,
+the scatter in eccentricity is likely to be low even in the case of a radial merger orbit, since the system is less sensitive to the exact value of the deflection angle,
 or the impact parameter, as can be seen from the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
@@
 and covers only a small part of the parameter space of merger configurations since we only consider fairly high-eccentricity galactic merger orbits.
-However, similar behavior can be expected to occur also in more realistic models,
-since the relevant dynamics occurs within few hundred parsecs from the center of the merged galaxy,
+However, similar behaviour can be expected to occur also in more realistic models,
+since the relevant dynamics occurs within a few hundred parsecs from the centre of the merged galaxy,
 where we do not expect any significant  effects from dark matter or the outer parts of the galaxy, which are not included in our simulations.
@@
 and this is also seen in simulation results \citep{gualandris2022}.
-However, most major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006},
-and the dynamical friction from the dark matter halo would generally drive even lower eccentricity merger orbits to highly radial final plunges.
+However, most major mergers are expected to occur on highly radial orbits based in cosmological simulations \citep[e.g.][]{khochfar2006},
+and the dynamical friction from the dark matter halo would generally drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
 
@@
 we expect that uncertainty of a similar magnitude is also present in real systems due various mechanisms, such as perturbations from their environment.
-The SMBHs are likely also not located exactly at the center of mass of the galaxy,
+The SMBHs are likely also not located exactly at the centre of mass of the galaxy,
 as e.g.\ work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
@@
 In conclusion, we have shown the importance of the deflection angle $\theta_\mathrm{defl}$ on the resulting hard binary eccentricity $e_\mathrm{h}$. 
-By using a resolution-free toy model of the SMBH scattering process,
+By using a simple, resolution-free toy model of the SMBH scattering process,
 we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations,

### 2023-06-13 papers/paper-eccentricity/paper-eccentricity.tex
- small edits to numerical section

@@
 For details about the implementation of \ketju{}, we point the reader to \citet{mannerkoski2023}.
+We use the same code configuration (including post-Newtonian corrections to order 3.5, fast multiple method with second order multipoles, and hierarchical time integration) as \citet{mannerkoski2023}.
 
-We construct our galaxy models to represent the nuclear bulge of a gas-devoid elliptical galaxy, matching those models of \citet{nasim2020}.
+We construct our galaxy models to represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
 Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
@@
 The chosen eccentricities $e_0$ are 0.90 and 0.99.
-We ensure that the radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022}.
-For each orbital configuration, we run ten realisations for each mass resolution, to account for Brownian motion effects of the SMBH binary caused by the discretised phase space.
+The radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022} for the $e_0=0.90$ orbit.
+For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space.
 
@@
     \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the number of stellar particles per progenitor galaxy $N_\star$ for initial eccentricities $e_0=0.90$ and $e_0=0.99$.
-    The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ suite after applying a $2.5\sigma$-cut to the simulation data. 
-    For the $e_0=0.99$ suite, a steeper scaling than the $1/\sqrt{N_\star}$ is obtained after applying a similar $2.5\sigma$-cut to the data.}
+    The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ mergers after applying a $2.5\sigma$-cut to the simulation data. 
+    For the $e_0=0.99$ mergers, a steeper scaling than the $1/\sqrt{N_\star}$ is obtained after applying a similar $2.5\sigma$-cut to the data.}
     \label{fig:convergence}
@@
 
-\subsection{\drafting{Eccentricity scatter in simulations}}
+\subsection{Eccentricity scatter in simulations}
 
@@
 Critically, the eccentricity values beyond $2.5\sigma$ from the mean that are removed in our cleaned samples are each still a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups with a highly-radial initial orbit.
-The cause of these low eccentricity values is discussed in \autoref{ssec:toy}.
 
@@
 binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage \citep{binney2008}.
-The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory, however is able to be more robustly measured than $b$ from the simulation data.
-We set as the first deflection angle the first 
-angle $\theta_\mathrm{defl}$ that exceeds an orbit-dependent threshold:
-$\theta_{\mathrm{defl,min}}=90\degr$ for the $e_0=0.99$ suite and $\theta_{\mathrm{defl,min}}=45\degr$ for the $e_0=0.90$ suite.
+The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory, however $\theta_\mathrm{defl}$ is able to be more robustly measured than $b$ from the simulation data.
+\drafting{
+    Each pericentre passage is associated with a deflection angle, however we find that it is the first extreme deflection angle that is important.
+    To quantify extreme, we set as $\theta_\mathrm{defl}$ the first 
+angle that exceeds an orbit-dependent threshold:
+$\theta_{\mathrm{defl,min}}=90\degr$ for the $e_0=0.99$ mergers and $\theta_{\mathrm{defl,min}}=45\degr$ for the $e_0=0.90$ mergers.
+}
 In general, this corresponds to the second or third actual pericentre passage of the SMBHs in the simulations.
 An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.90$ ($e_0=0.99$) orbit in the left (right) panel of \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
-We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, which we explain using a simple toy model.
+We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, shown in \autoref{fig:theta_e}.
 
@@
 
-\subsection{Reproducing the eccentricity behavior with a toy model}\label{ssec:toy}
+\subsection{Reproducing the eccentricity behaviour with a toy model}\label{ssec:toy}
 
@@
 the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
-although the model does appear to capture the mean behavior fairly well.
-The behavior of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
+although the model does appear to capture the mean behaviour fairly well.
+The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
 which might be related to the large scatter in the simulation data.
 
-Otherwise the behavior of the simulation data is captured well,
+Otherwise the behaviour of the simulation data is captured well,
 with the model correctly producing the two main eccentricity minima,
@@
 Our results show that the variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
-Together with the simple toy model that reproduces well the behavior seen in the simulations,
+Together with the simple toy model that reproduces well the behaviour seen in the simulations,
 this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic mergers on highly radial orbits,
@@
 As shown by both the simulations and the toy model, in these mergers the eccentricity of the SMBH binary can span nearly the full possible range between 0 and 1,
-depending on comparatively tiny differences on the scale of a parsec in the particular realization of the galactic merger orbit.
-\drafting{
-    This observation is at odds with \citet{nasim2021}, where binaries that formed at high eccentricity tended to proceed to higher eccentricities.
-    However, the tension can be explained if the simulations of \citet{nasim2021} lie in the linear region of the $\theta_\mathrm{defl}$--$e_\mathrm{h}$ parameter space (where most of the simulations in this work lie), coupled with the low-number statistics of \citet{nasim2021}.
-    This would suggest that the potential of the merger remnant, as opposed to the rotation of the SMBH binary relative to the stellar background, is the dominant predictor of hard binary eccentricity.
-}
-
+depending on comparatively tiny differences on the scale of a parsec in the particular realisation of the galactic merger orbit.
 Compared to major mergers, in minor mergers where there is less asymmetry in the stellar background when the SMBHs become bound,
@@
 
-The simulation setup used in this work is highly idealized,
+The simulation setup used in this work is highly idealised,
 and covers only a small part of the parameter space of merger configurations since we only consider fairly high-eccentricity galactic merger orbits.
@@
 and this is also seen in simulation results \citep{gualandris2022}.
-However, most major mergers are expected to occur on highly radial orbits based in cosmological simulations \citep[e.g.][]{khochfar2006},
+However, most major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006},
 and the dynamical friction from the dark matter halo would generally drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
 
-In numerical simulations with a discretised phase space, the two primary causes of variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001}.
+In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001}.
 Both of these effects scale as $1/\sqrt{N_\star}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
@@
 
-While the scatter in the impact parameter between different numerical realizations of the same merger is due to relatively low number of particles compared to real galaxies,
+While the scatter in the impact parameter between different numerical realisations of the same merger is due to relatively low number of particles compared to real galaxies,
 we expect that uncertainty of a similar magnitude is also present in real systems due various mechanisms, such as perturbations from their environment.
@@
 
-P.H.J. acknowledge the support
+A.R. acknowledges the support by the University of Helsinki Research Foundation.
+P.H.J. acknowledges the support
 by the European Research Council via ERC Consolidator Grant KETJU (no. 818930).

### 2023-06-20 papers/paper-eccentricity/paper-eccentricity.tex
Intro half done

@@
 \section{Introduction}
-Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies \citep[e.g.][]{kormendy2013}. 
-As galaxies dynamically interact, so too are their SMBHs expected to interact in a three-phase merger process \citep{begelman1980}.
-The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary system with a semimajor axis $a$ and eccentricity $e$.
-Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980, hills1983, quinlan1996}.
-Chaotic interactions drive the SMBH binary to sub-parsec separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
+Supermassive black holes (SMBHs) are believed to reside at the centres of all massive galaxies \citep[e.g.][]{kormendy2013}.
+In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{Volonteri2003,naab2017} their SMBHs 
+are also expected to interact in a three-phase merger process \citep{begelman1980}.
 
-It is well known that the efficiency by which GW emission can remove orbital energy and angular momentum from a binary system is highly sensitive to the binary eccentricity \citep{peters1964}, with the binary merger timescale being particularly sensitive.
+Firstly, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Merritt2013book}.
+In the second phase, the
+SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar
+distribution \citep{hills1980, hills1983, quinlan1996}. Finally, at small subparsec separations gravitational wave (GW) emission
+becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
+SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019,Mannerkoski2021,Mannerkoski2022}, where the strength of GW emission strongly depends on the
+binary eccentricity \citep{peters1963,peters1964}.
+
 The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
@@
 
+%The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary system with a semimajor axis $a$ and eccentricity $e$.
+%Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980, hills1983, quinlan1996}.
+%Chaotic interactions drive the SMBH binary to sub-parsec separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
+
+%It is well known that the efficiency by which GW emission can remove orbital energy and angular momentum from a binary system is highly sensitive to the binary eccentricity \citep{peters1964}, with the binary merger timescale being particularly sensitive.
+
+%The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
+%Understanding how faithfully SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the observational GW community aims to detect.
+
 Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
+
 In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.

### 2023-06-20 papers/paper-eccentricity/paper-eccentricity.tex
First version of Intro done

@@
 
-Firstly, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Merritt2013book}.
+Firstly, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
 In the second phase, the
@@
 becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019,Mannerkoski2021,Mannerkoski2022}, where the strength of GW emission strongly depends on the
+SMBHs to coalescence \citep[e.g.][]{Mannerkoski2021,Mannerkoski2022}, where the strength of the GW emission strongly depends on the
 binary eccentricity \citep{peters1963,peters1964}.
 
-The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
-Understanding how faithfully SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the observational GW community aims to detect.
+The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{Berentzen2009,Khan2011,Mannerkoski2019}
+to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{Lentati2015,Verbiest2016,Arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{Amaro-Seoane2023} space mission.
+Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate
+that the observational GW community aims to detect.
+
+The SMBH binary merging process and its dependence on eccentricity has been extensively studied using collisionsless merger simulations \citep[e.g.][]{Berentzen2009,Vasiliev2015,gualandris2017,gualandris2022}.
+In a recent study \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is in fact an artefact of poor phase space
+discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
+
+In this Letter, based on simulations run with the KETJU code \citep{mannerkoski2023}, we find that the SMBH binary eccentricity is reliably predictable only for particular
+initial conditions of the galaxy merger system, and does not in general hold.
+Indeed, we find that for realistic galaxy merger orbits \citep{khochfar2006} where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
+SMBH binary eccentricity is always present.
+
+%Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
 
@@
 
-Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
+%Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
 
-In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
-Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final SMBH binary eccentricity is present.
+%In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
+%Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final SMBH binary eccentricity is present.
 

### 2023-06-20 papers/paper-eccentricity/paper-eccentricity.tex
First version of Intro done

@@
 
-In this Letter, based on simulations run with the KETJU code \citep{mannerkoski2023}, we find that the SMBH binary eccentricity is reliably predictable only for particular
+In this Letter, based on simulations run with the KETJU code, we find that the SMBH binary eccentricity is reliably predictable only for particular
 initial conditions of the galaxy merger system, and does not in general hold.
-Indeed, we find that for realistic galaxy merger orbits \citep{khochfar2006} where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
+Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
 SMBH binary eccentricity is always present.

### 2023-06-20 papers/paper-eccentricity/paper-eccentricity.tex
First version of Intro done, minor additions

@@
 
-The SMBH binary merging process and its dependence on eccentricity has been extensively studied using collisionsless merger simulations \citep[e.g.][]{Berentzen2009,Vasiliev2015,gualandris2017,gualandris2022}.
-In a recent study \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is in fact an artefact of poor phase space
+The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{Berentzen2009,Vasiliev2015,Bortolas2016,gualandris2017,gualandris2022}.
+Specifically, in a recent study \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is in fact an artefact of poor phase space
 discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.

### 2023-06-20 papers/paper-eccentricity/paper-eccentricity.tex
First version of Intro done, minor additions

@@
 
-The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{Berentzen2009,Khan2011,Mannerkoski2019}
+The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{Berentzen2009,Khan2011,Dosopoulou2017,Mannerkoski2019}
 to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{Lentati2015,Verbiest2016,Arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{Amaro-Seoane2023} space mission.

### 2023-06-20 papers/paper-eccentricity/paper-eccentricity.tex
- small edits to intro, incorporate TN comments on intro

@@
 
-Firstly, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
+Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
 In the second phase, the
@@
 becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep[e.g.][]{Mannerkoski2021,Mannerkoski2022}, where the strength of the GW emission strongly depends on the
+SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019}, where the strength of the GW emission strongly depends on the
 binary eccentricity \citep{peters1963,peters1964}.
@@
 to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{Lentati2015,Verbiest2016,Arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{Amaro-Seoane2023} space mission.
-Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate
+Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus critical in constraining the SMBH merger rate
 that the observational GW community aims to detect.
@@
 The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{Berentzen2009,Vasiliev2015,Bortolas2016,gualandris2017,gualandris2022}.
-Specifically, in a recent study \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is in fact an artefact of poor phase space
+Recently, \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
 discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
 
-In this Letter, based on simulations run with the KETJU code, we find that the SMBH binary eccentricity is reliably predictable only for particular
+In this Letter, we find that the SMBH binary eccentricity is reliably predictable only for particular
 initial conditions of the galaxy merger system, and does not in general hold.
 Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
-SMBH binary eccentricity is always present.
+SMBH binary eccentricity is physical and unavoidable.
 
@@
 \textbf{PHJ}: Supervision, Writing -- review \& editing.
+\textbf{TN}:  Writing -- review \& editing.
 

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Introduction checked

@@
 SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar
-distribution \citep{hills1980, hills1983, quinlan1996}. Finally, at small subparsec separations gravitational wave (GW) emission
+distribution \citep{hills1980, hills1983, quinlan1996,Rantala2018}. Finally, at small subparsec separations gravitational wave (GW) emission
 becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019}, where the strength of the GW emission strongly depends on the
+SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019,Mannerkoski2022}, where the strength of the GW emission strongly depends on the
 binary eccentricity \citep{peters1963,peters1964}.

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Section 2 checked

@@
 
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017, rantala2020,mannerkoski2023}, implemented in \gadget{} \citep{springel2021}.
-For details about the implementation of \ketju{}, we point the reader to \citet{mannerkoski2023}.
-We use the same code configuration (including post-Newtonian corrections to order 3.5, fast multiple method with second order multipoles, and hierarchical time integration) as \citet{mannerkoski2023}.
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{mannerkoski2023,rantala2017}, in which the dynamics of
+SMBH and stars in a small region around them are integrated with an algorithmically regularised integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
+is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
+symmetric interactions and manifest momentum conservation. KETJU also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Mora2004}.
 
-We construct our galaxy models to represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
+Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
 Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
@@
 \end{equation}
-The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
-We test five different mass resolutions by sampling the stellar phase space distribution with a varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
+The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
+We test six different mass resolutions %by sampling the stellar phase space distribution
+with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
 
-We then construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
-The chosen eccentricities $e_0$ are 0.90 and 0.99.
+We then construct isolated merger initial conditions by placing two galaxies on two different elliptical orbits with eccentricities of $e_0=0.90$ and $e_0=0.99$, at a fixed
+initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
 The radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022} for the $e_0=0.90$ orbit.

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Initial check of Sec 3.1. Final check to be done when the new plots are in place.

@@
     \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for the highest mass resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
-    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard for the two simulation sets. 
+    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard. 
     In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a comparatively low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean. 
@@
 
-The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets.
+The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets, where $t_\mathrm{bound}$ is the time when the SMBH binaries become bound.
 At least one simulation is seen to lie significantly beyond the mean eccentricity in each set, with $e_\mathrm{h} \sim 0.8$ for the $e_0=0.90$ set and $e_\mathrm{h}\sim 0.2$ for the $e_0=0.99$ set. 
@@
 where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$, giving rise to the second equality in \autoref{eq:ahard}).
-Assuming the distribution of mean hard-binary eccentricities is Gaussian, we characterise the inter-simulation eccentricity scatter as the standard deviation in mean values of $e_\mathrm{h}$, denoted as $\sigma_e$.
+% Assuming the distribution of mean hard-binary eccentricities is Gaussian,
+We characterise the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
 
@@
     \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about the $x=-z$ line.
-    Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity of the deflection-phase of the evolution.
+    Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity.
+% of the deflection-phase of the evolution.
     In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after. 

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Initial check of Sec 3.2

@@
 \end{equation}
-where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$, giving rise to the second equality in \autoref{eq:ahard}).
+where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$). 
 % Assuming the distribution of mean hard-binary eccentricities is Gaussian,
@@
 binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage \citep{binney2008}.
-The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory, however $\theta_\mathrm{defl}$ is able to be more robustly measured than $b$ from the simulation data.
-\drafting{
-    Each pericentre passage is associated with a deflection angle, however we find that it is the first extreme deflection angle that is important.
-    To quantify extreme, we set as $\theta_\mathrm{defl}$ the first 
-angle that exceeds an orbit-dependent threshold:
-$\theta_{\mathrm{defl,min}}=90\degr$ for the $e_0=0.99$ mergers and $\theta_{\mathrm{defl,min}}=45\degr$ for the $e_0=0.90$ mergers.
-}
+The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory. However as $\theta_\mathrm{defl}$ can be more robustly measured from the simulation data, we instead
+opt using it for our analysis.
+
+Each pericentre passage is associated with a deflection angle, however we find that it is the first extreme deflection angle, associated with a hard scattering event, that is important.
+ To quantify an extreme deflection angle, we set as $\theta_\mathrm{defl}$ the first 
+angle that exceeds an orbit-dependent threshold of $\theta_{\mathrm{defl,min}}=90\degr$ for the $e_0=0.99$ mergers and $\theta_{\mathrm{defl,min}}=45\degr$ for the $e_0=0.90$ mergers.
+
 In general, this corresponds to the second or third actual pericentre passage of the SMBHs in the simulations.
 An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.90$ ($e_0=0.99$) orbit in the left (right) panel of \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
-We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, shown in \autoref{fig:theta_e}.
+We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, as shown in \autoref{fig:theta_e}.
 

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Initial check of Sec 3.3

@@
 Taking the orbit to lie in the $xy$-plane,
-the relative motion of the two equal-mass SMBHs in this model is follows the 
+the relative motion of the two equal-mass SMBHs in this model follows the 
 equation of motion
@@
 
-The stellar background is modeled as a constant density spheroidal potential
+The stellar background can be modeled as a constant density spheroidal potential
 \citep[e.g.][]{binney2008}
@@
 \begin{align}
-A_x &= 2 \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
+A_x &= 2 \left(\frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2}\right) 
     \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]\\
@@
 The stellar density is set to $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
-which approximately matches the values seen in the simulations when SMBH binary is becoming bound.
+which approximately matches the values seen in the simulations when the SMBH binary is becoming bound.
 
-Dynamical friction is modeled using the \citet{chandrasekhar1943} formula assuming
+The dynamical friction of a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
 a Maxwellian distribution with a constant velocity dispersion
 ${\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}}$,
-set based on the value measured from the simulations.
-This formula gives the acceleration of a single BH due to dynamical friction as
-\citep[e.g.][]{binney2008}
+set based on the value measured from the simulations \citep[e.g.][]{binney2008}:
+%This formula gives the acceleration of a single BH due to dynamical friction as
+%\citep[e.g.][]{binney2008}
 \begin{equation}\label{eq:chandra_df}
@@
 and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
-which is appropriate for the size of the simulated systems.
+which is derived from the size of the simulated systems.
 
@@
 \end{align}
-Based on the simulations, we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
-typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
-and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
-The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measurable from the simulations,
-due to stellar components that remain bound to the SMBHs and obscure the background
+In order to match the typical SMBH relative velocity seen at this separation in
+the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ for the $e_0=0.90$ case, and
+ $v_0 = 560\,\mathrm{km\,s^{-1}}$ for the the $e_0=0.99$ case.
+%we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
+%typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
+%and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
+The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
+due to stellar components that remain tightly bound to the SMBHs and obscure the background
 potential relevant for the dynamics,
-as well as due to the potential in the simulations being constantly evolving.
+as well as due to fact that the potential in the simulations is constantly evolving.
 However, only sufficiently large values of $e_\mathrm{s}$ allow for low binary
@@
 Performing calculations with different values of $e_\mathrm{s}$,
-we found a good fit to simulation results for $e_\mathrm{s}\approx 0.9$.
+we found a good fit to the simulation results for $e_\mathrm{s}\approx 0.9$.
 
@@
 while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
-These shifts are likely required due to the rotation of the stellar component tilting the background potential
+These shifts are required due to the rotation of the stellar component tilting the background potential
 relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
First version of aabstract
- update some plots

--- 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
@@
 
+\drafting{
+    We present a study of collisionless merger simulations using the recently released \ketju{} code to investigate the stochasticity in the supermassive black hole (SMBH) binary eccentricity at the time a hard binary forms.
+    We find a mapping between the hard binary eccentricity and the deflection angle the SMBHs undergo prior to becoming bound that is independent of the ratio between SMBH mass and stellar particle mass.
+    Decreasing the number of stellar particles does however sample a wider domain of deflection angles due to Brownian motion compared to increased stellar particle counts; some merger geometries explore the full $[0,1]$ domain of eccentricity as a result.
+    We argue that the SMBH binary eccentricity is not a deterministic function of the initial merger eccentricity, but rather should be understood as a distribution of possible outcomes, which are also most likely present in the observable universe.
+    We conclude that mass resolutions currently used to explore the SMBH binary merger process, including their merger timescales, are sufficient to capture the observable population.
+}
+
 \end{abstract}
@@
 \section{Introduction}
-Supermassive black holes (SMBHs) are believed to reside at the centres of all massive galaxies \citep[e.g.][]{kormendy2013}.
-In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{Volonteri2003,naab2017} their SMBHs 
-are also expected to interact in a three-phase merger process \citep{begelman1980}.
+Supermassive black holes (SMBHs) are believed to reside at the centres of almost all massive galaxies \citep[e.g.][]{kormendy2013}. 
+As galaxies dynamically interact, so too are their SMBHs expected to interact in a three-phase merger process \citep{begelman1980}.
+The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary system with a semimajor axis $a$ and eccentricity $e$.
+Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980, hills1983, quinlan1996}.
+Chaotic interactions drive the SMBH binary to sub-parsec separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
 
-Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
-In the second phase, the
-SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar
-distribution \citep{hills1980, hills1983, quinlan1996,Rantala2018}. Finally, at small subparsec separations gravitational wave (GW) emission
-becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019,Mannerkoski2022}, where the strength of the GW emission strongly depends on the
-binary eccentricity \citep{peters1963,peters1964}.
+It is well known that the efficiency by which GW emission can remove orbital energy and angular momentum from a binary system is highly sensitive to the binary eccentricity \citep{peters1964}, with the binary merger timescale being particularly sensitive.
+The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
+Understanding how faithfully SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the observational GW community aims to detect.
 
-The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{Berentzen2009,Khan2011,Dosopoulou2017,Mannerkoski2019}
-to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{Lentati2015,Verbiest2016,Arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{Amaro-Seoane2023} space mission.
-Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus critical in constraining the SMBH merger rate
-that the observational GW community aims to detect.
-
-The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{Berentzen2009,Vasiliev2015,Bortolas2016,gualandris2017,gualandris2022}.
-Recently, \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
-discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
-
-In this Letter, we find that the SMBH binary eccentricity is reliably predictable only for particular
-initial conditions of the galaxy merger system, and does not in general hold.
-Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
-SMBH binary eccentricity is physical and unavoidable.
-
-%Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
-
-%The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary system with a semimajor axis $a$ and eccentricity $e$.
-%Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980, hills1983, quinlan1996}.
-%Chaotic interactions drive the SMBH binary to sub-parsec separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
-
-%It is well known that the efficiency by which GW emission can remove orbital energy and angular momentum from a binary system is highly sensitive to the binary eccentricity \citep{peters1964}, with the binary merger timescale being particularly sensitive.
-
-%The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
-%Understanding how faithfully SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the observational GW community aims to detect.
-
-%Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
-
-%In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
-%Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final SMBH binary eccentricity is present.
+Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
+In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
+Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final SMBH binary eccentricity is present.
 
@@
     \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for the highest mass resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
-    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard. 
+    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard for the two simulation sets. 
     In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a comparatively low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean. 
@@
 
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{mannerkoski2023,rantala2017}, in which the dynamics of
-SMBH and stars in a small region around them are integrated with an algorithmically regularised integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
-is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
-symmetric interactions and manifest momentum conservation. KETJU also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Mora2004}.
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{rantala2017, rantala2020,mannerkoski2023}, implemented in \gadget{} \citep{springel2021}.
+For details about the implementation of \ketju{}, we point the reader to \citet{mannerkoski2023}.
+We use the same code configuration (including post-Newtonian corrections to order 3.5, fast multiple method with second order multipoles, and hierarchical time integration) as \citet{mannerkoski2023}.
 
-Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
+We construct our galaxy models to represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
 Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
@@
 \end{equation}
-The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
-We test six different mass resolutions %by sampling the stellar phase space distribution
-with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
+The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed with zero initial velocity.
+We test five different mass resolutions by sampling the stellar phase space distribution with a varying number of stellar particles: $N_\star = \{2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
 
-We then construct isolated merger initial conditions by placing two galaxies on two different elliptical orbits with eccentricities of $e_0=0.90$ and $e_0=0.99$, at a fixed
-initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
+We then construct isolated merger initial conditions by placing two galaxies on an elliptical orbit with varying values of eccentricity, at a fixed initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
+The chosen eccentricities $e_0$ are 0.90 and 0.99.
 The radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022} for the $e_0=0.90$ orbit.
@@
 
-The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets, where $t_\mathrm{bound}$ is the time when the SMBH binaries become bound.
+The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets.
 At least one simulation is seen to lie significantly beyond the mean eccentricity in each set, with $e_\mathrm{h} \sim 0.8$ for the $e_0=0.90$ set and $e_\mathrm{h}\sim 0.2$ for the $e_0=0.99$ set. 
@@
 where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$, giving rise to the second equality in \autoref{eq:ahard}).
-% Assuming the distribution of mean hard-binary eccentricities is Gaussian,
-We characterise the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
+Assuming the distribution of mean hard-binary eccentricities is Gaussian, we characterise the inter-simulation eccentricity scatter as the standard deviation in mean values of $e_\mathrm{h}$, denoted as $\sigma_e$.
 
@@
     \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about the $x=-z$ line.
-    Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity.
-% of the deflection-phase of the evolution.
+    Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity of the deflection-phase of the evolution.
     In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after. 
@@
 \textbf{PHJ}: Supervision, Writing -- review \& editing.
-\textbf{TN}:  Writing -- review \& editing.
 

### 2023-06-21 code/analysis_scripts/analysis_submit.sh

@@ -11,7 +11,7 @@ module purge
 module restore py393
 
 
-python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-090/dehnen_250K_mergers -m mc
+python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-099/dehnen_4M_mergers -m mc
 
 #python extract_HM_quantities.py /users/arawling/projects/collisionless-merger-sample/parameters/parameters-analysis/HMQcubes.yml /users/arawling/projects/collisionless-merger-sample/parameters/parameters-mergers/eccentricity-study/e-099/dehnen_2M_mergers -m mc
 

### 2023-06-21 papers/paper-eccentricity/scripts/toy_model_plots.py

@@ -92,6 +92,12 @@ def plot_sim_data_th_e(ax, simdata):
                    zorder=3
                    )
 
+def add_artificial_end_to_curve(d, t_shift=0):
+    d['theta'] = np.insert(d['theta'], 0, np.pi-t_shift)
+    d['e'] = np.append(d['e'], 1)
+    d['b'] = np.insert(d['b'], 0, 0)
+    return d
+
 
 def paper_th_e_curve_plot():
 
@@ -133,7 +139,7 @@ def paper_th_e_curve_plot():
         axdict[k].xaxis.set_major_formatter(degree_format_str)
     for k in "ABDE":
         axdict[k].set_xlim(20,180)
-        axdict[k].set_xticks(np.arange(30,180,30))
+        axdict[k].set_xticks(np.arange(30,200,30))
 
     data = load_data(data_path('well_fitting_e_0.90_model_curve.pkl'))
     curve_shift = -12
@@ -142,7 +148,8 @@ def paper_th_e_curve_plot():
     for e_s, d in zip(data['e_spheroids'], data['res']):
         if e_s <0.9: continue
         a = 0.8 if e_s == 0.905 else 0.3
-        l, = axdict["D"].plot(np.degrees(d['theta'])+curve_shift, d['e'], color=color, alpha=a)
+        d_plot = add_artificial_end_to_curve(d, curve_shift)
+        l, = axdict["D"].plot(np.degrees(d_plot['theta'])+curve_shift, d_plot['e'], color=color, alpha=a)
         color = l.get_color()
     print("b90 for e_0=0.90", np.interp(90,np.degrees(d['theta'][::-1]),d['b'][::-1]*1e3))
 
@@ -161,7 +168,8 @@ def paper_th_e_curve_plot():
     for e_s, d in zip(data['e_spheroids'], data['res']):
         if e_s < 0.9: continue
         a = 0.8 if e_s == 0.905 else 0.3
-        l, = axdict["E"].plot(np.degrees(d['theta'])+curve_shift, d['e'], alpha=a, color=color)
+        d_plot = add_artificial_end_to_curve(d, curve_shift)
+        l, = axdict["E"].plot(np.degrees(d_plot['theta'])+curve_shift, d_plot['e'], alpha=a, color=color)
         color = l.get_color()
 
 

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex

--- 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
@@
     Limited to 250 words.
+}
+
+\drafting{
+    We present a study of collisionless merger simulations using the recently released \ketju{} code to investigate the stochasticity in the supermassive black hole (SMBH) binary eccentricity at the time a hard binary forms.
+    We find a mapping between the hard binary eccentricity and the deflection angle the SMBHs undergo prior to becoming bound that is independent of the ratio between SMBH mass and stellar particle mass.
+    Decreasing the number of stellar particles does however sample a wider domain of deflection angles due to Brownian motion compared to increased stellar particle counts; some merger geometries explore the full $[0,1]$ domain of eccentricity as a result.
+    We argue that the SMBH binary eccentricity is not a deterministic function of the initial merger eccentricity, but rather should be understood as a distribution of possible outcomes, which are also most likely present in the observable universe.
+    We conclude that mass resolutions currently used to explore the SMBH binary merger process, including their merger timescales, are sufficient to capture the observable population.
 }

### 2023-06-21 papers/paper-eccentricity/paper-eccentricity.tex
Changes in text to reflect figure updates, also remove 2.5 sigma stuff

@@
 
-\drafting{
-    Key results:
-    \begin{enumerate}
-        \item Eccentricity of hard binary depends on deflection angle 
-        \item Mapping $\theta_\mathrm{defl} \rightarrow e_\mathrm{h}$ relatively independent of resolution, however low res explores wider domain space 
-        \item Explored eccentricity space can be full $[0,1]$ range for certain systems 
-        \item Exact values of $e$ from simulations meaningless, need to understand the distribution of 
-        \item Current resolutions appropriate for exploring statistics of eccentricity
-    \end{enumerate}
-    Limited to 250 words.
-}
+%\drafting{
+%    Key results:
+%    \begin{enumerate}
+%        \item Eccentricity of hard binary depends on deflection angle 
+%        \item Mapping $\theta_\mathrm{defl} \rightarrow e_\mathrm{h}$ relatively independent of resolution, however low res explores wider domain space 
+%        \item Explored eccentricity space can be full $[0,1]$ range for certain systems 
+%        \item Exact values of $e$ from simulations meaningless, need to understand the distribution of 
+%        \item Current resolutions appropriate for exploring statistics of eccentricity
+%    \end{enumerate}
+%    Limited to 250 words.
+%}
 
@@
     Decreasing the number of stellar particles does however sample a wider domain of deflection angles due to Brownian motion compared to increased stellar particle counts; some merger geometries explore the full $[0,1]$ domain of eccentricity as a result.
-    We argue that the SMBH binary eccentricity is not a deterministic function of the initial merger eccentricity, but rather should be understood as a distribution of possible outcomes, which are also most likely present in the observable universe.
-    We conclude that mass resolutions currently used to explore the SMBH binary merger process, including their merger timescales, are sufficient to capture the observable population.
+    We argue that the SMBH binary eccentricity is not a deterministic function of the initial merger eccentricity, but rather should be understood as a distribution of possible outcomes, and is also most likely present in the observable universe.
+    We conclude that mass resolutions currently used to explore the SMBH binary merger process, including their merger timescales, are sufficient to capture the observable population, and that stochasticity in the SMBH binary eccentricity is unavoidable.
 }
@@
     \includegraphics{eccentricities.pdf}
-    \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for the highest mass resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
+    \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for the $M_\bullet/m_\star=5000$ resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
     The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard. 
-    In both the shown $e_0 = 0.90$ and $e_0 = 0.99$ sets, one simulation has a comparatively low eccentricity, displacing it to a value of $e$ greater than $2.5\sigma$ from the collective mean. 
-    In the $e_0=0.90$ set, seven simulations have a SMBH binary merger within $50\,\mathrm{Myr}$, as indicated by the rapid circularisation of the eccentricity.}
+    In the $e_0=0.90$ set, six simulations have a SMBH binary merger within $50\,\mathrm{Myr}$, shown by the rapid orbit circularisation.}
     \label{fig:eccentricities}
@@
 is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
-symmetric interactions and manifest momentum conservation. KETJU also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Mora2004}.
+symmetric interactions and manifest momentum conservation. \ketju{} also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Mora2004}.
 
@@
     \includegraphics{convergence}
-    \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the number of stellar particles per progenitor galaxy $N_\star$ for initial eccentricities $e_0=0.90$ and $e_0=0.99$.
-    The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ mergers after applying a $2.5\sigma$-cut to the simulation data. 
-    For the $e_0=0.99$ mergers, a steeper scaling than the $1/\sqrt{N_\star}$ is obtained after applying a similar $2.5\sigma$-cut to the data.}
+    \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the total number of stellar particles $N_{\star, \mathrm{tot}}$. 
+    The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ mergers. 
+    For the $e_0=0.99$ mergers, $\sigma_e$ does not decrease until $M_\bullet/m_\star \geq 20000$.}
     \label{fig:convergence}
@@
 
-The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets, where $t_\mathrm{bound}$ is the time when the SMBH binaries become bound.
-At least one simulation is seen to lie significantly beyond the mean eccentricity in each set, with $e_\mathrm{h} \sim 0.8$ for the $e_0=0.90$ set and $e_\mathrm{h}\sim 0.2$ for the $e_0=0.99$ set. 
+The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets, where $t_\mathrm{bound}$ is the time when the binary orbital energy $\mathcal{E}$ becomes and remains negative.
 The median time when each simulation set forms a hard binary is similar between the two sets, as indicated by the arrows in \autoref{fig:eccentricities}.
-The $e_0=0.90$ set also demonstrates seven SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid circularisation in eccentricity in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
+The $e_0=0.90$ set also demonstrates six SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid orbit circularisation in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
+The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $[0,1]$.
+Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence.
 
@@
 where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$). 
-% Assuming the distribution of mean hard-binary eccentricities is Gaussian,
 We characterise the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
@@
 We show the dependence of inter-simulation eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with circle markers, and for the $e_0=0.99$ orbit using square markers.
-The full set of ten simulations per mass resolution is shown with unfilled markers, and those simulations satisfying a $2.5\sigma$-cut in eccentricity with filled markers.
 
-For the $e_0=0.90$ orbit, using the full set of simulations results in a lack of convergence of $\sigma_e$, with $5\times10^3$ and $1\times10^4$ mass resolutions demonstrating an increased scatter compared to the lowest mass resolution of $2.5\times10^3$.
-The peculiar values of $\sigma_e$ arise due to the assumed Gaussian distribution of the eccentricity between simulations.
-Inspecting the orbital parameters for all simulations from the $e_0=0.90$ set, it is found that for resolutions of $5\times10^3$ and $1\times10^4$ the eccentricity clusters around a value of $e_\mathrm{h}\simeq0.90$, except for one simulation at both $5\times10^3$ and $1\times10^4$ resolutions which has a value of $e_\mathrm{h}<0.4$.
-This one outlier for each of the two resolutions artificially inflates the measured eccentricity variance, and with the $2.5\sigma$-cut data we recover the $1/\sqrt{N_\star}$ scaling of $\sigma_e$, confirming the results of \citet{nasim2020}, apart from the highest mass resolution set of $M_\bullet/m_\star=4\times10^4$.
-The variation in eccentricity also clearly decreases for the cleaned $e_0=0.99$ orbit simulations, albeit with a steeper gradient than the $e_0=0.90$ set. 
-Again, a single simulation at the highest mass resolution for the $e_0=0.99$ set is responsible for the inflated value of $\sigma_e$ in the full data set for this resolution.
-
-Critically, the eccentricity values beyond $2.5\sigma$ from the mean that are removed in our cleaned samples are each still a valid measurement from the simulation, and highlight the presence of stochasticity in eccentricity for these idealised simulation set-ups with a highly-radial initial orbit.
+For the $e_0=0.90$ orbit, the convergence of $\sigma_e$ scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, where $N_{\star,\mathrm{tot}}$ is the total number of stellar particles in the merger, in agreement with the results of \citet{nasim2020}.
+The variation in eccentricity does not show the same scaling for the $e_0=0.99$ orbit as the $e_0=0.99$ orbit. 
+For mass resolutions $M_\bullet/m_\star<10000$, the value of $\sigma_e$ is almost constant, before significantly dropping at higher mass resolutions. \drafting{Add about 4M runs when done}
 
@@
 Each pericentre passage is associated with a deflection angle, however we find that it is the first extreme deflection angle, associated with a hard scattering event, that is important.
- To quantify an extreme deflection angle, we set as $\theta_\mathrm{defl}$ the first 
-angle that exceeds an orbit-dependent threshold of $\theta_{\mathrm{defl,min}}=90\degr$ for the $e_0=0.99$ mergers and $\theta_{\mathrm{defl,min}}=45\degr$ for the $e_0=0.90$ mergers.
+To quantify an extreme deflection angle, we set as $\theta_\mathrm{defl}$ the first angle that exceeds $\theta_{\mathrm{defl,min}}=30\degr$. \drafting{Is this detail necessary?}
 
 In general, this corresponds to the second or third actual pericentre passage of the SMBHs in the simulations.
-An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.90$ ($e_0=0.99$) orbit in the left (right) panel of \autoref{fig:orbit}, which occurs between the points A and B in the SMBH binary orbit in each inset panel.
+An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.90$ ($e_0=0.99$) orbit in the left (right) panel of \autoref{fig:orbit}, which occurs between the arrow heads along the SMBH trajectory in each inset panel.
 We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, as shown in \autoref{fig:theta_e}.
@@
 % of the deflection-phase of the evolution.
-    In each inset panel, point A shows a time just before the first hard scattering $\theta_\mathrm{defl}$, and point B a time just after. 
-    The circle marker indicates the position of the SMBH just before the SMBHs form a bound binary.}
+    In each inset panel, $\theta_\mathrm{defl}$ is the deflection between the arrow heads.
+    The circle marker indicates the SMBH position just before a bound binary is formed.}
     \label{fig:orbit}
@@
 In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001}.
-Both of these effects scale as $1/\sqrt{N_\star}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
+Both of these effects scale as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
 is approximately linear.

### 2023-06-22 papers/paper-eccentricity/paper-eccentricity.tex
Adjust toy model parameters for better fit

@@
     compared to reasonably well fitting model curves.
-    The solid lines show the results from the toy model with $e_\mathrm{s}=0.905$,
-    while the semitransparent lines show the results for  $e_\mathrm{s}=0.900$ and $e_\mathrm{s}=0.910$.
+    The solid lines show the results from the toy model with $e_\mathrm{s}=0.91$,
+    while the semitransparent lines show the results for  $e_\mathrm{s}=0.90$ and $e_\mathrm{s}=0.92$.
     The model curves have been shifted left by $12\degr$ and $6\degr$
@@
 The dynamical friction of a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
-a Maxwellian distribution with a constant velocity dispersion
-${\sigma_\mathrm{M}=200\,\mathrm{km\,s^{-1}}}$,
-set based on the value measured from the simulations \citep[e.g.][]{binney2008}:
-%This formula gives the acceleration of a single BH due to dynamical friction as
-%\citep[e.g.][]{binney2008}
+a Maxwellian distribution with a constant velocity dispersion $\sigma_\star$ \citep[e.g.][]{binney2008}:
 \begin{equation}\label{eq:chandra_df}
@@
 \vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
-                        &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1}.
+                         &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
 \end{aligned}
 \end{equation}
-Here $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH, 
-$X = |\vb{v}_1|/\sqrt{2}\sigma_\mathrm{M}$,
-and the Coulomb logarithm is taken to have the value $\ln{\Lambda} = 5$,
-which is derived from the size of the simulated systems.
+where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
+and $X = |\vb{v}_1|/\sqrt{2}\sigma_\star$.
+We set ${\sigma_\star=200\,\mathrm{km\,s^{-1}}}$ based on the value measured from the simulations.
+The value of the Coulomb logarithm is expected to be $\ln{\Lambda}\sim 4\textnormal{--}5$
+based on the size of the stellar system,
+and we find $\ln{\Lambda} = 4.7$ to give results that agree well with the simulations.
 
@@
  $v_0 = 560\,\mathrm{km\,s^{-1}}$ for the the $e_0=0.99$ case.
-%we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ to match the
-%typical SMBH relative velocity at this separation in the $e_0=0.90$ case,
-%and $v_0 = 560\,\mathrm{km\,s^{-1}}$ to match the $e_0=0.99$ case.
 The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
@@
 the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
-although the model does appear to capture the mean behaviour fairly well.
-The behaviour of the model in this region is also quite sensitive to the exact value of $e_\mathrm{s}$,
+although the model does appear to capture the mean behaviour fairly well, apart from the largest values of
+$\theta_\mathrm{defl}$.
+The behaviour of the model in this region is also quite sensitive to the values of $e_\mathrm{s}$
+and $\ln{\Lambda}$,
 which might be related to the large scatter in the simulation data.

### 2023-06-26 papers/paper-eccentricity/paper-eccentricity.tex
Different abstract versions from Overleaf

@@
 
+We study the stochasticity of supermassive black hole (SMBH) binary eccentricity in $N$-body simulations of equal-mass galaxy mergers using the \ketju{} code.
+In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a highly non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
+This mapping between the deflection angle and the binary eccentricity is captured using a simple toy model, indicating that it is driven by the interplay between an asymmetric stellar potential and dynamical friction acting on the SMBHs, with no apparent resolution dependence.
+Due to the non-linearity of this mapping, in certain merger configurations small, parsec-scale variations in the merger orbit can result in binary eccentricities varying in nearly the full possible range between 0 and 1.
+By increasing the number of stellar particles, the reduction in Brownian motion means a smaller range of deflection angles is sampled, thus allowing the binary eccentricity produced by a given merger orbit to converge towards a definite value.
+We conclude that the sensitivity of the eccentricity to small-scale perturbations of the merger orbit means that the binary eccentricity is effectively a random quantity when considering realistic galaxy mergers, however we stress that the distribution of these effectively random eccentricities can be constrained using even moderate resolution simulations.
+
 \drafting{
@@
     We conclude that mass resolutions currently used to explore the SMBH binary merger process, including their merger timescales, are sufficient to capture the observable population, and that stochasticity in the SMBH binary eccentricity is unavoidable.
+}
+
+\drafting{Another version:}
+
+\drafting{
+We study the stochasticity of the eccentricity of SMBH binaries formed in $N$-body simulations equal-mass galactic mergers.
+In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a highly non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
+This mapping between the deflection angle and binary eccentricity is reproduced with a simple toy model, showing that it is caused by an interplay between the asymmetric stellar potential and dynamical friction acting on the SMBHs.
+There is also no apparent resolution dependency in this mapping, with lower particle count simulations simply sampling a wider range of deflection angles due to Brownian motion.
+Due to the non-linearity of this mapping, in certain merger configurations small, parsec-scale variations in the merger orbit can result in eccentricities varying in nearly the full possible range between 0 and 1.
+Higher particle counts reduce the Brownian motion induced variation in the merger orbit in idealised simulations, allowing the eccentricity produced by a given merger orbit to converge towards a definite value.
+However, we argue that the sensitivity of the eccentricity on the small-scale perturbations of the merger orbit means that the binary eccentricity is an effectively random quantity when considering realistic galaxy mergers, but also that the distribution of these effectively random eccentricities can be constrained using even moderate resolution simulations.
+}
+
+\drafting{Third version:}
+
+\drafting{
+We study the stochasticity of the eccentricity of SMBH binaries formed in $N$-body simulations of equal-mass galactic mergers, run with the \ketju{} code.
+In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a highly non-linear function of the final hard scattering, as defined by the deflection angle.
+The
+mapping between the deflection angle and the binary eccentricity can be reproduced with a simple toy model, showing that it is caused by an interplay between the asymmetric stellar potential and dynamical friction acting on the SMBHs, with no apparent resolution dependency.
+At lower particle resolutions the increase in Brownian motion result in a wider sampling of the deflection angles, whereas at higher particle counts a given merger orbit converge towards a more definite eccentricity value.
+Due to the non-linearity of this mapping, for certain merger configurations small, parsec-scale variations in the merger orbit can result in final SMBH eccentricities that nearly cover the full $e=[0,1]$ eccentricity range.
+We conclude that the sensitivity of the eccentricity on the small-scale perturbations of the merger orbit means that the binary eccentricity is effectively a random quantity when considering realistic galaxy mergers, however we stress that the distribution of these effectively random eccentricities can be constrained using even moderate resolution simulations.
 }
@@
 
-Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
+Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scales down to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
 In the second phase, the
@@
 initial conditions of the galaxy merger system, and does not in general hold.
-Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
+Indeed, we find that for realistic galaxy merger orbits where the SMBHs undergo a final near-radial plunge before becoming bound, stochasticity in the final
+%initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
 SMBH binary eccentricity is physical and unavoidable.
@@
 is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
-symmetric interactions and manifest momentum conservation. \ketju{} also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Mora2004}.
+symmetric interactions and manifest momentum conservation. \ketju{} also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Blanchet2014}.
 
 Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
-Each galaxy is thus represented as a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+Each galaxy consists of a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
 \begin{equation}\label{eq:dehnen}
@@
 The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
-We test six different mass resolutions %by sampling the stellar phase space distribution
-with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$.
+We test six different mass resolutions 
+with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$, corresponding to $M_\bullet/m_\star= 1000 \textnormal{--} 40000$.
+\drafting{See if we include even higher res runs.}
 
@@
 For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space. 
+Interactions between stellar particles are softened with a softening length of $\epsilon = 2.5\,\mathrm{pc}$.
 
@@
 
-The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$ for the highest mass resolution sets, where $t_\mathrm{bound}$ is the time when the binary orbital energy $\mathcal{E}$ becomes and remains negative.
+The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$, where $t_\mathrm{bound}$ is the time when the binary orbital energy $E$ becomes permanently negative.
 The median time when each simulation set forms a hard binary is similar between the two sets, as indicated by the arrows in \autoref{fig:eccentricities}.
@@
 The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $[0,1]$.
-Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence.
+Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence during the $50\,\mathrm{Myr}$ timespan.
 
@@
 
-We show the dependence of inter-simulation eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with circle markers, and for the $e_0=0.99$ orbit using square markers.
+We show the dependence of the inter-simulation eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with blue circle markers, and for the $e_0=0.99$ orbit using orange square markers.
 
 For the $e_0=0.90$ orbit, the convergence of $\sigma_e$ scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, where $N_{\star,\mathrm{tot}}$ is the total number of stellar particles in the merger, in agreement with the results of \citet{nasim2020}.
-The variation in eccentricity does not show the same scaling for the $e_0=0.99$ orbit as the $e_0=0.99$ orbit. 
+The variation in eccentricity does not show the same scaling for the $e_0=0.99$ orbit as the $e_0=0.90$ orbit. 
 For mass resolutions $M_\bullet/m_\star<10000$, the value of $\sigma_e$ is almost constant, before significantly dropping at higher mass resolutions. \drafting{Add about 4M runs when done}
@@
 % of the deflection-phase of the evolution.
-    In each inset panel, $\theta_\mathrm{defl}$ is the deflection between the arrow heads.
+    In each inset panel, $\theta_\mathrm{defl}$ is the deflection at the pericentre between the arrows.
     The circle marker indicates the SMBH position just before a bound binary is formed.}
@@
     in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
-    compared to reasonably well fitting model curves.
+    compared to fitted model curves.
     The solid lines show the results from the toy model with $e_\mathrm{s}=0.91$,
     while the semitransparent lines show the results for  $e_\mathrm{s}=0.90$ and $e_\mathrm{s}=0.92$.
-    The model curves have been shifted left by $12\degr$ and $6\degr$
-    to better match the simulation data, with the shift also indicated by arrows.
-    The marginal histograms show kernel density estimates to the simulation data.
+    The $e_0=0.90$ model curves have been shifted left by $14\degr$
+    to better match the simulation data, with the shift also indicated by an arrow.
+    The marginal histograms show kernel density estimates of the simulation data.
+    The inset panel for the $e_0=0.90$ data shows a zoom-in with $\theta_\mathrm{defl}=[52\degr, 73\degr]$ and $e_\mathrm{h}=[0.96, 1.00]$ for clarity.
     }
@@
     The dashed lines show the isopotential contours of the model.
-    Only the early part of the evolution is shown for clarity.
+    Only the early part of the orbital evolution is shown for clarity.
     Bottom:
@@
 $\vb{a}_\mathrm{bg}$ the acceleration due to the asymmetric background potential,
-and $\vb{a}_\mathrm{DF}$ the acceleration due to dynamical friction.
+and $\vb{a}_\mathrm{DF}$ is the acceleration due to dynamical friction.
 
-The stellar background can be modeled as a constant density spheroidal potential
+The stellar background is modeled as a constant density spheroidal potential
 \citep[e.g.][]{binney2008}
@@
 
-The dynamical friction of a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
+The dynamical friction acting on a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
 a Maxwellian distribution with a constant velocity dispersion $\sigma_\star$ \citep[e.g.][]{binney2008}:
@@
 based on the size of the stellar system,
-and we find $\ln{\Lambda} = 4.7$ to give results that agree well with the simulations.
+and we find that $\ln{\Lambda} = 4.7$ gives results that agree well with the simulations.
 
@@
 To account for this, when the binary is bound 
-we multiply the acceleration given by Eq.~\eqref{eq:chandra_df}
+we multiply the acceleration given by equation~\eqref{eq:chandra_df}
 with the smooth cut-off function
@@
 and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
-with $a_\mathrm{h}$ defined by equation \eqref{eq:ahard}.
+with $a_\mathrm{h}$ defined by equation~\eqref{eq:ahard}.
 The total dynamical friction term is then
@@
  $v_0 = 560\,\mathrm{km\,s^{-1}}$ for the the $e_0=0.99$ case.
+
 The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
@@
 The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
-To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $12\degr$,
-while in the $e_0=0.99$ case a smaller shift of $6\degr$ is used.
-These shifts are required due to the rotation of the stellar component tilting the background potential
-relative to the SMBH trajectories in the simulations, with the effect being stronger in the $e_0=0.90$ case
+To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $14\degr$.
+This shift is required to account for the rotation of the stellar component tilting the background potential
+relative to the SMBH trajectories in the simulations, with the effect being only evident in the $e_0=0.90$ case
 due to the less radial merger orbit.
@@
 the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
-although the model does appear to capture the mean behaviour fairly well, apart from the largest values of
+although the simple model does appear to capture the mean behaviour relatively well, apart from the largest values of
 $\theta_\mathrm{defl}$.
@@
 and $\ln{\Lambda}$,
-which might be related to the large scatter in the simulation data.
+which might be related to the large scatter seen in the simulation data.
 
-Otherwise the behaviour of the simulation data is captured well,
+However, in general the behaviour of the simulation data is captured well,
 with the model correctly producing the two main eccentricity minima,
@@
 
-Our results show that the variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
+We find that the variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
 Together with the simple toy model that reproduces well the behaviour seen in the simulations,
-this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic mergers on highly radial orbits,
-at least in major mergers where the SMBHs have nearly equal masses,
-and not an error due to insufficient numerical resolution.
+this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic major mergers on highly radial orbits, and not an error due to insufficient numerical resolution.
+%at least in major mergers where the SMBHs have nearly equal masses,
+%and not an error due to insufficient numerical resolution.
 
-As shown by both the simulations and the toy model, in these mergers the eccentricity of the SMBH binary can span nearly the full possible range between 0 and 1,
-depending on comparatively tiny differences on the scale of a parsec in the particular realisation of the galactic merger orbit.
-Compared to major mergers, in minor mergers where there is less asymmetry in the stellar background when the SMBHs become bound,
-the scatter in eccentricity is likely to be low even in the case of a radial merger orbit, since the system is less sensitive to the exact value of the deflection angle,
-or the impact parameter, as can be seen from the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
+As shown by both the simulations and the toy model, the eccentricity of the SMBH binary can span nearly the full possible range between $e=[0,1]$, depending on comparatively small differences on the scale of a few parsecs in the particular realisation of the galactic merger orbit.
+On the other hand in minor mergers where the stellar background is less asymmetric when the SMBHs become bound,
+the scatter in eccentricity is likely to be relatively low even in the case of a radial merger orbit, since the system is less sensitive to the exact value of the deflection angle,
+as can be seen from the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
 
-The simulation setup used in this work is highly idealised,
-and covers only a small part of the parameter space of merger configurations since we only consider fairly high-eccentricity galactic merger orbits.
+Lower eccentricity merger orbits can be expected to show less scatter in the binary eccentricity due to the lack of a hard scattering event that is sensitive to slight perturbations in the merger orbit, which was also seen in the study by \citet{gualandris2022}.
+However, major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006}.
+In addition, the dynamical friction from the dark matter halo would in general drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
+
+The simulation setup used in this work is idealised,
+and covers only a small part of the parameter space of merger configurations since we only consider relatively high-eccentricity galactic merger orbits.
 However, similar behaviour can be expected to occur also in more realistic models,
 since the relevant dynamics occurs within a few hundred parsecs from the centre of the merged galaxy,
-where we do not expect any significant  effects from dark matter or the outer parts of the galaxy, which are not included in our simulations.
-Lower eccentricity merger orbits can be expected to show less scatter in the binary eccentricity due to the lack of a hard scattering event that is sensitive to slight perturbations in the merger orbit,
-and this is also seen in simulation results \citep{gualandris2022}.
-However, most major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006},
-and the dynamical friction from the dark matter halo would generally drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
+where we do not expect any significant effects from dark matter or the outer parts of the galaxy, which are not included in our present simulations.
 
-In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001}.
+In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001,Bortolas2016}.
 Both of these effects scale as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
 is approximately linear.
-The deviation from this scaling seen in \autoref{fig:convergence}, particularly when not excluding outlying data points, is then simply explained by the fact that the eccentricity is not globally a linear function of the impact parameter.
+The deviation from this scaling seen in \autoref{fig:convergence} can then be explained by the fact that the eccentricity is not globally a linear function of the impact parameter.
 
 While the scatter in the impact parameter between different numerical realisations of the same merger is due to relatively low number of particles compared to real galaxies,
-we expect that uncertainty of a similar magnitude is also present in real systems due various mechanisms, such as perturbations from their environment.
-The SMBHs are likely also not located exactly at the centre of mass of the galaxy,
-as e.g.\ work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
-It is therefore evident that it is not possible to give an exact prediction for the SMBH binary eccentricity produced by a given merger,
+we expect that uncertainty of a similar magnitude is also present in real systems due to various mechanisms, such as perturbations from their stellar environment.
+In addition, the SMBHs are also not necesssarily located exactly at the centre of mass of the galaxy,
+as e.g.\ work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by up to $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
+Thus, it is not possible to give an exact prediction for the SMBH binary eccentricity produced by a given merger,
 and instead we must focus on predicting the distribution of eccentricities that can be produced by a given merger configuration.
-Such predictions are crucial for constraining observational quantities that are correlated with binary eccentricity,
-most notably SMBH binary merger timescale and the spectrum of GWs at nanohertz frequencies observable with PTAs \drafting{(refs)}.
 
-Predicting the eccentricity distribution requires knowing the distribution of impact parameters that can result from essentially identical mergers
+Predicting the eccentricity distribution requires the knowledge of the distribution of impact parameters that can result from essentially identical mergers
 as well as the mapping between the impact parameter, or deflection angle, and the final eccentricity.
@@
 simulations at moderate mass resolutions of $M_\bullet/m_\star \sim 10^3$ can be used
-to map out the relation by performing a large enough number of merger simulations,
+to map out the relation by performing a large number of merger simulations,
 possibly augmented by more sophisticated versions of the toy model presented here.
@@
 since in simulations numerical resolution effects are likely dominant.
-In any case, already from the results shown here it is clear that the distribution has a complex structure.
+%In any case, already from the results shown here it is clear that the distribution has a complex structure.
 
@@
 The results presented here justify extending the investigation to more realistic galaxy merger scenarios,
-to quantify the expected range of hard binary eccentricities critical for merger timescale predictions for future gravitational wave observation missions.
+in order to quantify the expected range of hard binary eccentricities, which is critical for merger timescale predictions for future gravitational wave observation missions.
 
@@
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
-P.H.J. acknowledges the support
-by the European Research Council via ERC Consolidator Grant KETJU (no. 818930).
+A.R., M.M. and P.H.J. acknowledge the support
+by the European Research Council via ERC Consolidator Grant KETJU (no. 818930) and the support of the Academy of Finland grant 339127.
+TN acknowledges support from the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany’s Excellence Strategy - EXC-2094 - 390783311 from the DFG Cluster of Excellence ``ORIGINS''.
 
@@
 
-\textit{Software:} \ketju{} \citep{rantala2017}, \gadget{} \citep{springel2021}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
+\textit{Software:} \ketju{} \citep{mannerkoski2023,rantala2017}, \gadget{} \citep{springel2021}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
 
@@
 
-\bibliography{../ref}{}
+\bibliography{ref}{}
 \bibliographystyle{aasjournal}

### 2023-06-26 papers/paper-eccentricity/paper-eccentricity.tex
Update references in bib file

@@
 
-\author[0000-0001-5721-9335]{Matias Mannerkoski}
+\author[0000-0001-5721-9335]{Matias mannerkoski}
 \affiliation{Department of Physics,
@@
 Supermassive black holes (SMBHs) are believed to reside at the centres of all massive galaxies \citep[e.g.][]{kormendy2013}.
-In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{Volonteri2003,naab2017} their SMBHs 
+In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{volonteri2003,naab2017} their SMBHs 
 are also expected to interact in a three-phase merger process \citep{begelman1980}.
 
-Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scales down to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{Milosavljevic2001,Merritt2013book}.
+Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scales down to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{milosavljevic2001,merritt2013}.
 In the second phase, the
 SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar
-distribution \citep{hills1980, hills1983, quinlan1996,Rantala2018}. Finally, at small subparsec separations gravitational wave (GW) emission
+distribution \citep{hills1980, hills1983, quinlan1996,rantala2018}. Finally, at small subparsec separations gravitational wave (GW) emission
 becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep[e.g.][]{Mannerkoski2019,Mannerkoski2022}, where the strength of the GW emission strongly depends on the
+SMBHs to coalescence \citep[e.g.][]{mannerkoski2019,mannerkoski2022}, where the strength of the GW emission strongly depends on the
 binary eccentricity \citep{peters1963,peters1964}.
 
-The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{Berentzen2009,Khan2011,Dosopoulou2017,Mannerkoski2019}
-to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{Lentati2015,Verbiest2016,Arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{Amaro-Seoane2023} space mission.
+The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{berentzen2009,khan2011,dosopoulou2017,mannerkoski2019}
+to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{lentati2015,verbiest2016,arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2023} space mission.
 Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus critical in constraining the SMBH merger rate
@@
 
-The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{Berentzen2009,Vasiliev2015,Bortolas2016,gualandris2017,gualandris2022}.
+The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{berentzen2009,vasiliev2015,bortolas2016,gualandris2017,gualandris2022}.
 Recently, \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
@@
 is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
-symmetric interactions and manifest momentum conservation. \ketju{} also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{Blanchet2014}.
+symmetric interactions and manifest momentum conservation. \ketju{} also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{blanchet2014}.
 
@@
 
-In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001,Bortolas2016}.
+In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001,bortolas2016}.
 Both of these effects scale as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$

### 2023-07-21 papers/paper-eccentricity/paper-eccentricity.tex

@@
 
-\title{Reviving stochasticity: SMBH-binary eccentricity is inherently random}
+\title{Reviving stochasticity: uncertainty in SMBH binary eccentricity is unavoidable}
 
@@
 
-\author[0000-0001-5721-9335]{Matias mannerkoski}
+\author[0000-0001-5721-9335]{Matias Mannerkoski}
 \affiliation{Department of Physics,
@@
 \begin{abstract}
-
-%\drafting{
-%    Key results:
-%    \begin{enumerate}
-%        \item Eccentricity of hard binary depends on deflection angle 
-%        \item Mapping $\theta_\mathrm{defl} \rightarrow e_\mathrm{h}$ relatively independent of resolution, however low res explores wider domain space 
-%        \item Explored eccentricity space can be full $[0,1]$ range for certain systems 
-%        \item Exact values of $e$ from simulations meaningless, need to understand the distribution of 
-%        \item Current resolutions appropriate for exploring statistics of eccentricity
-%    \end{enumerate}
-%    Limited to 250 words.
-%}
-
-We study the stochasticity of supermassive black hole (SMBH) binary eccentricity in $N$-body simulations of equal-mass galaxy mergers using the \ketju{} code.
-In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a highly non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
-This mapping between the deflection angle and the binary eccentricity is captured using a simple toy model, indicating that it is driven by the interplay between an asymmetric stellar potential and dynamical friction acting on the SMBHs, with no apparent resolution dependence.
-Due to the non-linearity of this mapping, in certain merger configurations small, parsec-scale variations in the merger orbit can result in binary eccentricities varying in nearly the full possible range between 0 and 1.
-By increasing the number of stellar particles, the reduction in Brownian motion means a smaller range of deflection angles is sampled, thus allowing the binary eccentricity produced by a given merger orbit to converge towards a definite value.
-We conclude that the sensitivity of the eccentricity to small-scale perturbations of the merger orbit means that the binary eccentricity is effectively a random quantity when considering realistic galaxy mergers, however we stress that the distribution of these effectively random eccentricities can be constrained using even moderate resolution simulations.
-
-\drafting{
-    We present a study of collisionless merger simulations using the recently released \ketju{} code to investigate the stochasticity in the supermassive black hole (SMBH) binary eccentricity at the time a hard binary forms.
-    We find a mapping between the hard binary eccentricity and the deflection angle the SMBHs undergo prior to becoming bound that is independent of the ratio between SMBH mass and stellar particle mass.
-    Decreasing the number of stellar particles does however sample a wider domain of deflection angles due to Brownian motion compared to increased stellar particle counts; some merger geometries explore the full $[0,1]$ domain of eccentricity as a result.
-    We argue that the SMBH binary eccentricity is not a deterministic function of the initial merger eccentricity, but rather should be understood as a distribution of possible outcomes, and is also most likely present in the observable universe.
-    We conclude that mass resolutions currently used to explore the SMBH binary merger process, including their merger timescales, are sufficient to capture the observable population, and that stochasticity in the SMBH binary eccentricity is unavoidable.
-}
-
-\drafting{Another version:}
-
-\drafting{
-We study the stochasticity of the eccentricity of SMBH binaries formed in $N$-body simulations equal-mass galactic mergers.
-In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a highly non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
-This mapping between the deflection angle and binary eccentricity is reproduced with a simple toy model, showing that it is caused by an interplay between the asymmetric stellar potential and dynamical friction acting on the SMBHs.
-There is also no apparent resolution dependency in this mapping, with lower particle count simulations simply sampling a wider range of deflection angles due to Brownian motion.
-Due to the non-linearity of this mapping, in certain merger configurations small, parsec-scale variations in the merger orbit can result in eccentricities varying in nearly the full possible range between 0 and 1.
-Higher particle counts reduce the Brownian motion induced variation in the merger orbit in idealised simulations, allowing the eccentricity produced by a given merger orbit to converge towards a definite value.
-However, we argue that the sensitivity of the eccentricity on the small-scale perturbations of the merger orbit means that the binary eccentricity is an effectively random quantity when considering realistic galaxy mergers, but also that the distribution of these effectively random eccentricities can be constrained using even moderate resolution simulations.
-}
-
-\drafting{Third version:}
-
-\drafting{
-We study the stochasticity of the eccentricity of SMBH binaries formed in $N$-body simulations of equal-mass galactic mergers, run with the \ketju{} code.
-In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a highly non-linear function of the final hard scattering, as defined by the deflection angle.
-The
-mapping between the deflection angle and the binary eccentricity can be reproduced with a simple toy model, showing that it is caused by an interplay between the asymmetric stellar potential and dynamical friction acting on the SMBHs, with no apparent resolution dependency.
-At lower particle resolutions the increase in Brownian motion result in a wider sampling of the deflection angles, whereas at higher particle counts a given merger orbit converge towards a more definite eccentricity value.
-Due to the non-linearity of this mapping, for certain merger configurations small, parsec-scale variations in the merger orbit can result in final SMBH eccentricities that nearly cover the full $e=[0,1]$ eccentricity range.
-We conclude that the sensitivity of the eccentricity on the small-scale perturbations of the merger orbit means that the binary eccentricity is effectively a random quantity when considering realistic galaxy mergers, however we stress that the distribution of these effectively random eccentricities can be constrained using even moderate resolution simulations.
-}
+We study supermassive black hole (SMBH) binary eccentricity of equal-mass galaxy mergers in $N$-body simulations with the \ketju{} code, which combines the \gadget{} fast multipole gravity solver with accurate regularized integration and post-Newtonian corrections around SMBHs.
+In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
+This mapping between the deflection angle and the binary eccentricity has no apparent resolution dependence in our simulations spanning the resolution range of $1\times10^5$--$8\times10^6$ particles per galaxy.
+The mapping is also captured using a simple model with an analytic potential, indicating that it is driven by the interplay between a smooth asymmetric stellar background potential and dynamical friction acting on the SMBHs.
+Due to the non-linearity of this mapping, in certain merger configurations small, parsec-scale variations in the merger orbit can result in binary eccentricities varying in nearly the full possible range between $e=0$ and $e=1$.
+In idealized simulations, such variations are caused by finite resolution effects,
+and convergence of the binary eccentricity can be achieved with increasing resolution.
+However, in real galaxies, other mechanisms such as nuclear gas and substructure that perturb the merger orbit are likely to be significant enough for the binary eccentricity to be effectively random.
+Our results indicate that the distribution of these effectively random eccentricities can be studied using even moderate resolution simulations.
 
@@
 %% to include these concepts in their preprints.
-\keywords{Edit keywords}
+%\keywords{Galaxy mergers (608); Supermassive black holes (1663); Eccentricity (441); N-body simulations (1083); }
 
@@
 \section{Introduction}
-Supermassive black holes (SMBHs) are believed to reside at the centres of all massive galaxies \citep[e.g.][]{kormendy2013}.
+Supermassive black holes (SMBHs) are believed to reside at the centers of all massive galaxies \citep[e.g.][]{kormendy2013}.
 In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{volonteri2003,naab2017} their SMBHs 
@@
 
-Firstly, dynamical friction \citep{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scales down to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{milosavljevic2001,merritt2013}.
+Firstly, dynamical friction \citep[DF,][]{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scales down to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{milosavljevic2001,merritt2013}.
 In the second phase, the
@@
 becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep[e.g.][]{mannerkoski2019,mannerkoski2022}, where the strength of the GW emission strongly depends on the
-binary eccentricity \citep{peters1963,peters1964}.
+SMBHs to coalescence, where the strength of the GW emission strongly depends on the
+binary eccentricity \citep{peters1963,peters1964,mannerkoski2019}.
 
 The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{berentzen2009,khan2011,dosopoulou2017,mannerkoski2019}
-to provide quantitative predictions for observational programmes such as ground-based pulsar timing arrays \citep[PTAs, e.g.][]{lentati2015,verbiest2016,arzoumanian2020} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2023} space mission.
+to provide quantitative predictions for ongoing observational programs such as ground-based pulsar timing arrays \citep[PTAs,][]{agazie2023,2023Antoniadis,2023Xu,zic2023}
+and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2023}.
 Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus critical in constraining the SMBH merger rate
@@
 The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{berentzen2009,vasiliev2015,bortolas2016,gualandris2017,gualandris2022}.
-Recently, \citet{nasim2020} have argued that the scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
-discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
+Recently, \citet{nasim2020} have argued that the substantial scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
+discretization, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
 
-In this Letter, we find that the SMBH binary eccentricity is reliably predictable only for particular
-initial conditions of the galaxy merger system, and does not in general hold.
-Indeed, we find that for realistic galaxy merger orbits where the SMBHs undergo a final near-radial plunge before becoming bound, stochasticity in the final
-%initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final
-SMBH binary eccentricity is physical and unavoidable.
+In this Letter, we find that for realistic galaxy merger orbits the scatter in the final SMBH binary eccentricity is due to the 
+physical sensitivity of the final eccentricity to small perturbations of the final  nearly radial plunging trajectory of the SMBHs before they become bound, even in the infinite resolution limit.
+As a result, we argue that the stochasticity of the binary eccentricity is an unavoidable physical feature of realistic galaxy mergers, at least in the equal-mass case.
 
-%Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
-
-%The first phase, dynamical friction \citep{chandrasekhar1943}, acts to bring the SMBHs from kiloparsec-scale separations to parsec-scale separations, after which the SMBHs form a bound binary system with a semimajor axis $a$ and eccentricity $e$.
-%Following the binding of the SMBHs, the SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar distribution \citep{hills1980, hills1983, quinlan1996}.
-%Chaotic interactions drive the SMBH binary to sub-parsec separation: a regime where gravitational wave (GW) emission becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, resulting in the coalescence of the two SMBHs into a single SMBH.
-
-%It is well known that the efficiency by which GW emission can remove orbital energy and angular momentum from a binary system is highly sensitive to the binary eccentricity \citep{peters1964}, with the binary merger timescale being particularly sensitive.
-
-%The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques to understand, and to provide quantitative predictions for observational programmes such as pulsar timing arrays \citep[PTAs, e.g.][]{babak2016,falxa2023} and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2007} mission.
-%Understanding how faithfully SMBH binary eccentricity is captured in simulations is thus a critical piece of the puzzle in constraining the SMBH merger rate that the observational GW community aims to detect.
-
-%Previous work by \citet{nasim2020} have argued that scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses, $M_\bullet \gg m_\star$, SMBH binary eccentricity is a reliably predictable quantity.
-
-%In this work, we find that SMBH binary eccentricity is reliably predictable only for particular initial conditions of the galaxy merger system, and does not in general hold.
-%Indeed, we find that for realistic galaxy merger orbits where the initial orbital eccentricity of the system is $e_0>0.90$, stochasticity in the final SMBH binary eccentricity is present.
+\begin{figure*}
+    \centering
+    \includegraphics{eccentricities.pdf}
+    \caption{Binary eccentricity $e$ (note the non-linear scale) and shifted time $t'=t-t_\mathrm{bound}$ for the $M_\bullet/m_\star=5000$ resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
+    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard. 
+    In the $e_0=0.90$ set, six simulations have a binary merger within $50\,\mathrm{Myr}$.}
+    \label{fig:eccentricities}
+\end{figure*}
 
@@
 
-\begin{figure*}
-    \centering
-    \includegraphics{eccentricities.pdf}
-    \caption{Eccentricity $e$ (note the non-linear scale) as a function of shifted time $t'=t-t_\mathrm{bound}$ for the $M_\bullet/m_\star=5000$ resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
-    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard. 
-    In the $e_0=0.90$ set, six simulations have a SMBH binary merger within $50\,\mathrm{Myr}$, shown by the rapid orbit circularisation.}
-    \label{fig:eccentricities}
-\end{figure*}
-
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of \ketju{} \citep{mannerkoski2023,rantala2017}, in which the dynamics of
-SMBH and stars in a small region around them are integrated with an algorithmically regularised integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
+We construct a number of idealized galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017}.
+The dynamics of
+SMBHs, and stars in a small region around them, are integrated with an algorithmically regularized integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
 is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
-symmetric interactions and manifest momentum conservation. \ketju{} also includes Post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{blanchet2014}.
+symmetric interactions and manifest momentum conservation. \ketju{} also includes post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{blanchet2014}.
 
 Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
-Each galaxy consists of a stellar-only Dehnen sphere \citep{dehnen1993} with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+Each galaxy consists of a stellar-only \citet{dehnen1993} sphere with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
 \begin{equation}\label{eq:dehnen}
@@
 \end{equation}
-The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
-We test six different mass resolutions 
-with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6\}$, corresponding to $M_\bullet/m_\star= 1000 \textnormal{--} 40000$.
-\drafting{See if we include even higher res runs.}
+The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the center of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
+We test seven different mass resolutions 
+with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6, 8.0\times10^6\}$, corresponding to $M_\bullet/m_\star= 1000 \textnormal{--} 80000$.
 
 We then construct isolated merger initial conditions by placing two galaxies on two different elliptical orbits with eccentricities of $e_0=0.90$ and $e_0=0.99$, at a fixed
-initial separation of $D=3.72\,\mathrm{kpc}$ and semimajor axis $a_0=2.79\,\mathrm{kpc}$.
-The radial and tangential velocities of the initial merger setup are consistent with the values reported in \citet{gualandris2022} for the $e_0=0.90$ orbit.
-For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space. 
-Interactions between stellar particles are softened with a softening length of $\epsilon = 2.5\,\mathrm{pc}$.
+initial separation of $D=3.72\,\mathrm{kpc}$ and a semimajor axis of $a_0=2.79\,\mathrm{kpc}$.
+The $e_0=0.90$ merger orbit matches that used by \citet{nasim2020}, with the radial and tangential velocities being consistent with the values reported in \citet{gualandris2022}.
+For each orbital configuration, we run ten realizations for each mass resolution, to account for stochasticity caused by the discretized phase space. 
+Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
 
@@
 The median time when each simulation set forms a hard binary is similar between the two sets, as indicated by the arrows in \autoref{fig:eccentricities}.
-The $e_0=0.90$ set also demonstrates six SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid orbit circularisation in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
-The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $[0,1]$.
+The $e_0=0.90$ set also demonstrates six SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid orbit circularization in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
+The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $e=[0,1]$.
 Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence during the $50\,\mathrm{Myr}$ timespan.
 
-To characterise the scatter in eccentricity, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard.
-The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}
+To characterize the scatter in eccentricity, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centered on the orbit within which the SMBH binary has become hard.
+The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}:
 \begin{equation}\label{eq:ahard}
@@
 where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$). 
-We characterise the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
+We characterize the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
 
-We show the dependence of the inter-simulation eccentricity standard deviation on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with blue circle markers, and for the $e_0=0.99$ orbit using orange square markers.
+We show the dependence of $\sigma_e$ on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with blue circle markers, and for the $e_0=0.99$ orbit using orange square markers.
 
-For the $e_0=0.90$ orbit, the convergence of $\sigma_e$ scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, where $N_{\star,\mathrm{tot}}$ is the total number of stellar particles in the merger, in agreement with the results of \citet{nasim2020}.
+For the $e_0=0.90$ orbit, the convergence of $\sigma_e$ scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, where $N_{\star,\mathrm{tot}}$ is the total number of stellar particles in the merger.
+The values of $\sigma_e$ we report are quantitatively similar to the values found by \citet{nasim2020} for the same system. 
 The variation in eccentricity does not show the same scaling for the $e_0=0.99$ orbit as the $e_0=0.90$ orbit. 
-For mass resolutions $M_\bullet/m_\star<10000$, the value of $\sigma_e$ is almost constant, before significantly dropping at higher mass resolutions. \drafting{Add about 4M runs when done}
+For mass resolutions $M_\bullet/m_\star \leq 10000$, the value of $\sigma_e$ is almost constant, before significantly dropping at higher mass resolutions.
+
+\begin{figure*}
+    \centering
+    \includegraphics{orbit}
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realizations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about $x=-z$.
+    The line gradient shows the shifted time $t'=t-t_\mathrm{bound}$, however the coloring for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity. 
+    In each inset panel, $\theta_\mathrm{defl}$ is the deflection angle at the pericenter between the arrows.
+    The circle marker indicates the SMBH position when a bound binary is formed.}
+    \label{fig:orbit}
+\end{figure*}
 
@@
 To understand the observed scatter in eccentricity, we investigate the binary binding process in each simulation. 
-Whilst the two SMBHs are unbound, they undergo a number of hyperbolic orbits influenced by the galactic merger potential.
-At each pericentre passage, the two SMBHs scatter off each other with some deflection angle $\theta_\mathrm{defl}$, defined:
+Before the SMBHs are bound, the primary influence on the motion is from the galactic merger potential; two realizations for both merger orbits are shown in \autoref{fig:orbit}.
+The initial stages of the orbit show indiscernible variation between realizations, however after a particularly strong interaction between the SMBHs during a pericenter passage, the orbits deviate between realizations (inset panels, \autoref{fig:orbit}), and evolve to different binary eccentricities.
+
+To quantify the strength of the hard-scattering process that randomizes the SMBH orbit, we measure the effective two-body deflection angle, defined as:
 \begin{equation}\label{eq:theta}
-    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right)
+    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right) = 2\arctan\left( \frac{b_{90}}{b} \right)
 \end{equation}
-where $M=M_{\bullet,1}+M_{\bullet,2}$, and $L$ is the magnitude of the SMBH 
-binary angular momentum vector and $E$ the binary orbital energy at the time of the pericentre passage \citep{binney2008}.
-The deflection angle is a function of the impact parameter $b$ of the SMBH trajectory. However as $\theta_\mathrm{defl}$ can be more robustly measured from the simulation data, we instead
-opt using it for our analysis.
+where $M=M_{\bullet,1}+M_{\bullet,2}$, $L$ is the magnitude of the SMBH 
+system angular momentum vector, $E$ the orbital energy at the time of the pericenter passage, $b$ the effective impact parameter, and $b_{90}$ the $90\degr$ deflection radius \citep{binney2008}.
 
-Each pericentre passage is associated with a deflection angle, however we find that it is the first extreme deflection angle, associated with a hard scattering event, that is important.
-To quantify an extreme deflection angle, we set as $\theta_\mathrm{defl}$ the first angle that exceeds $\theta_{\mathrm{defl,min}}=30\degr$. \drafting{Is this detail necessary?}
-
-In general, this corresponds to the second or third actual pericentre passage of the SMBHs in the simulations.
-An example of $\theta_\mathrm{defl}$ is shown for two realisations of the $e_0=0.90$ ($e_0=0.99$) orbit in the left (right) panel of \autoref{fig:orbit}, which occurs between the arrow heads along the SMBH trajectory in each inset panel.
-We observe evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, as shown in \autoref{fig:theta_e}.
-
-\begin{figure*}
-    \centering
-    \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about the $x=-z$ line.
-    Each line is coloured according to the shifted time $t'=t-t_\mathrm{bound}$ of the simulation, however the colours for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity.
-% of the deflection-phase of the evolution.
-    In each inset panel, $\theta_\mathrm{defl}$ is the deflection at the pericentre between the arrows.
-    The circle marker indicates the SMBH position just before a bound binary is formed.}
-    \label{fig:orbit}
-\end{figure*}
+We observe clear evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, as shown with the data points in \autoref{fig:theta_e}.
 
@@
     \caption{
-    Scattering deflection angle $\theta_\mathrm{defl}$ and the resulting binary eccentricity
+    Deflection angle $\theta_\mathrm{defl}$ and resulting binary eccentricity
     in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
     compared to fitted model curves.
-    The solid lines show the results from the toy model with $e_\mathrm{s}=0.91$,
+    The solid lines show the results from the analytic model with $e_\mathrm{s}=0.91$,
     while the semitransparent lines show the results for  $e_\mathrm{s}=0.90$ and $e_\mathrm{s}=0.92$.
     The $e_0=0.90$ model curves have been shifted left by $14\degr$
-    to better match the simulation data, with the shift also indicated by an arrow.
+    to better match the simulation data, with the shift indicated by an arrow.
     The marginal histograms show kernel density estimates of the simulation data.
-    The inset panel for the $e_0=0.90$ data shows a zoom-in with $\theta_\mathrm{defl}=[52\degr, 73\degr]$ and $e_\mathrm{h}=[0.96, 1.00]$ for clarity.
+    The inset panel for the $e_0=0.90$ data shows a zoom-in with $\theta_\mathrm{defl}=[52\degr, 73\degr]$ and $e_\mathrm{h}=[0.96, 1.00]$.
     }
@@
 
-\subsection{Reproducing the eccentricity behaviour with a toy model}\label{ssec:toy}
-
-\begin{figure*}
-    \centering
-    \includegraphics{sample_model_orbits}
-    \caption{
-    Top:
-    Sample orbits computed with the toy model for different background potential eccentricities $e_\mathrm{s}$,
-    with otherwise identical initial conditions.
-    The dashed lines show the isopotential contours of the model.
-    Only the early part of the orbital evolution is shown for clarity.
-    Bottom:
-    The orbital eccentricity $e$ of the model orbits (note the non-linear scale).
-    The vertical line marks the end of the period shown in the top panels. 
-    }
-    \label{fig:model_orbits}
-\end{figure*}
+\subsection{Reproducing the eccentricity behavior with a simple model}\label{ssec:toy}
 
@@
 deflection angle $\theta_\mathrm{defl}$ during the binary formation
-can be reproduced using a simple toy model,
+can be reproduced using a simple analytic model,
 which includes only the essential components relevant to this process.
@@
 $\vb{a}_\mathrm{bg}$ the acceleration due to the asymmetric background potential,
-and $\vb{a}_\mathrm{DF}$ is the acceleration due to dynamical friction.
+and $\vb{a}_\mathrm{DF}$ is the acceleration due to DF.
 
 The stellar background is modeled as a constant density spheroidal potential
-\citep[e.g.][]{binney2008}
+\citep[e.g.][]{binney2008}:
 \begin{equation}
@@
 
-The dynamical friction acting on a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
+\begin{figure*}
+    \centering
+    \includegraphics{sample_model_orbits}
+    \caption{
+    Top:
+    Sample orbits computed with the analytic model for different background potential eccentricities $e_\mathrm{s}$,
+    with otherwise identical initial conditions.
+    The dashed lines show the isopotential contours of the model.
+    Only the early part of the orbital evolution is shown for clarity.
+    Bottom:
+    Orbital eccentricity $e$ of the model orbits (note the non-linear scale).
+    The vertical line marks the end of the period shown in the top panels. 
+    }
+    \label{fig:model_orbits}
+\end{figure*}
+
+The DF acting on a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
 a Maxwellian distribution with a constant velocity dispersion $\sigma_\star$ \citep[e.g.][]{binney2008}:
@@
 where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
-and $X = |\vb{v}_1|/\sqrt{2}\sigma_\star$.
+and $X = |\vb{v}_1|/\left(\sqrt{2}\sigma_\star\right)$.
 We set ${\sigma_\star=200\,\mathrm{km\,s^{-1}}}$ based on the value measured from the simulations.
-The value of the Coulomb logarithm is expected to be $\ln{\Lambda}\sim 4\textnormal{--}5$
+The value of the Coulomb logarithm is expected to be in the range $\ln{\Lambda}\sim 4\textnormal{--}5$
 based on the size of the stellar system,
@@
 
-The effect of dynamical friction weakens as the SMBH binary orbit shrinks.
+The effect of DF weakens as the SMBH binary orbit shrinks.
 To account for this, when the binary is bound 
@@
 \end{equation}
-Here $a$ is the semi-major axis of a bound binary,
+Here $a$ is the semimajor axis of a bound binary,
 and the cut-off scales are $a_\mathrm{c}=2 a_\mathrm{h}$
@@
 with $a_\mathrm{h}$ defined by equation~\eqref{eq:ahard}.
-The total dynamical friction term is then
+The total DF term is then
 \begin{equation}
@@
 The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
-due to stellar components that remain tightly bound to the SMBHs and obscure the background
+due to the presence of stellar components that remain tightly bound to the SMBHs and obscure the background
 potential relevant for the dynamics,
-as well as due to fact that the potential in the simulations is constantly evolving.
+in combination with a constantly evolving potential.
 However, only sufficiently large values of $e_\mathrm{s}$ allow for low binary
@@
 the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}.
-For reference, a $90\degr$ deflection during the first pericentre passage
-corresponds to initial impact parameters of $b\approx 5.8\,\mathrm{pc}$ and
-$b\approx 3.3\,\mathrm{pc}$ for the $e_0=0.90$ and $e_0=0.99$ cases, respectively.
+For reference, the $90\degr$ deflection radii are $b_{90}\approx 5.8\,\mathrm{pc}$ and
+$b_{90}\approx 3.3\,\mathrm{pc}$ for the $e_0=0.90$ and $e_0=0.99$ cases, respectively.
 
@@
 To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $14\degr$.
-This shift is required to account for the rotation of the stellar component tilting the background potential
+This shift is required likely due to the rotation of the stellar component tilting the background potential
 relative to the SMBH trajectories in the simulations, with the effect being only evident in the $e_0=0.90$ case
@@
 the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
-although the simple model does appear to capture the mean behaviour relatively well, apart from the largest values of
+although the simple model does appear to capture the mean behavior relatively well, apart from the very largest values of
 $\theta_\mathrm{defl}$.
-The behaviour of the model in this region is also quite sensitive to the values of $e_\mathrm{s}$
+The behaviour of the model in this region is also sensitive to the exact values of $e_\mathrm{s}$
 and $\ln{\Lambda}$,
-which might be related to the large scatter seen in the simulation data.
+which might also be related to the large scatter seen in the simulation data.
 
-However, in general the behaviour of the simulation data is captured well,
+However, in general the behavior of the simulation data is captured well,
 with the model correctly producing the two main eccentricity minima,
 as well as the $e\approx 1$ region between them.
-The minima occur for trajectories where the torque from the background potential together with the dynamical friction
+The minima occur for trajectories where the torque from the background potential together with the DF
 causes the SMBHs to loop around into a nearly circular orbit,
 while highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along 
-the main axes of the potential, as is seen in \autoref{fig:model_orbits}.
+the main axes of the potential, as shown in \autoref{fig:model_orbits}.
 
@@
 We find that the variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
-Together with the simple toy model that reproduces well the behaviour seen in the simulations,
-this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic major mergers on highly radial orbits, and not an error due to insufficient numerical resolution.
-%at least in major mergers where the SMBHs have nearly equal masses,
-%and not an error due to insufficient numerical resolution.
+Together with the simple analytic model that reproduces well the behavior seen in the simulations,
+this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic major mergers on highly radial orbits, and not an error caused by insufficient numerical resolution.
 
-As shown by both the simulations and the toy model, the eccentricity of the SMBH binary can span nearly the full possible range between $e=[0,1]$, depending on comparatively small differences on the scale of a few parsecs in the particular realisation of the galactic merger orbit.
-On the other hand in minor mergers where the stellar background is less asymmetric when the SMBHs become bound,
+As shown by both the simulations and the analytic model, the eccentricity of the SMBH binary can span nearly the full possible range between $e=[0,1]$, depending on comparatively small, parsec-scale differences in the particular realization of the galactic merger orbit.
+On the other hand, in minor mergers where the stellar background is less asymmetric when the SMBHs become bound,
 the scatter in eccentricity is likely to be relatively low even in the case of a radial merger orbit, since the system is less sensitive to the exact value of the deflection angle,
-as can be seen from the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
+as seen in the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
 
@@
 However, major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006}.
-In addition, the dynamical friction from the dark matter halo would in general drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
+In addition, the DF from the dark matter halo would in general drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
 
-The simulation setup used in this work is idealised,
-and covers only a small part of the parameter space of merger configurations since we only consider relatively high-eccentricity galactic merger orbits.
-However, similar behaviour can be expected to occur also in more realistic models,
-since the relevant dynamics occurs within a few hundred parsecs from the centre of the merged galaxy,
+The simulation setup used in this work is idealized, and covers only a small part of the parameter space of merger configurations.
+However, similar behavior can be expected to occur also in more realistic models,
+since the relevant dynamics occurs within a few hundred parsecs from the center of the merged galaxy,
 where we do not expect any significant effects from dark matter or the outer parts of the galaxy, which are not included in our present simulations.
 
-In numerical simulations with a discretised phase space, the two primary causes of random variation in the impact parameter are the initial particle sampling variation in the merger initial conditions and the Brownian motion of the SMBH binary \citep{merritt2001,bortolas2016}.
-Both of these effects scale as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
+The variation of the deflection angle is caused by the random variation of the impact parameter $b$ through equation~\eqref{eq:theta}.
+In numerical simulations with a discretized phase space, the primary cause of random variation in $b$ is the Brownian motion of the SMBHs before they form a bound binary \citep{merritt2001,bortolas2016}.
+This effect scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
 is approximately linear.
@@
 
-While the scatter in the impact parameter between different numerical realisations of the same merger is due to relatively low number of particles compared to real galaxies,
-we expect that uncertainty of a similar magnitude is also present in real systems due to various mechanisms, such as perturbations from their stellar environment.
-In addition, the SMBHs are also not necesssarily located exactly at the centre of mass of the galaxy,
-as e.g.\ work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photo-centre by up to $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
+While the scatter in the impact parameter between different numerical realizations of the same merger is due to relatively low number of particles compared to real galaxies,
+we expect that uncertainty of a similar magnitude is also present in real systems due to various mechanisms, such as perturbations from gas and substructure in their stellar environment.
+In addition, the SMBHs are also not necesssarily located exactly at the center of mass of the galaxy.
+For example, the work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photocenter by up to $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
 Thus, it is not possible to give an exact prediction for the SMBH binary eccentricity produced by a given merger,
@@
 to map out the relation by performing a large number of merger simulations,
-possibly augmented by more sophisticated versions of the toy model presented here.
-The distribution of impact parameters in real systems is a much more difficult problem to tackle,
+possibly augmented by more sophisticated versions of the analytic model presented here.
+The distribution of impact parameters in real systems is however a much more difficult problem to tackle,
 since in simulations numerical resolution effects are likely dominant.
-%In any case, already from the results shown here it is clear that the distribution has a complex structure.
 
-In conclusion, we have shown the importance of the deflection angle $\theta_\mathrm{defl}$ on the resulting hard binary eccentricity $e_\mathrm{h}$. 
-By using a simple, resolution-free toy model of the SMBH scattering process,
-we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations,
+In conclusion, we have shown the importance of the deflection angle $\theta_\mathrm{defl}$, and thus the impact parameter $b$, on the resulting hard binary eccentricity $e_\mathrm{h}$. 
+By using a simple, resolution-free analytic model of the SMBH scattering process,
+we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretization effects in galaxy merger simulations,
 but is rather due to the physical sensitivity of the system to small changes in the merger orbit,
-which can be caused by physical mechanisms in addition to numerical discretisation effects.
+which can be caused by physical mechanisms in addition to numerical discretization effects.
 The results presented here justify extending the investigation to more realistic galaxy merger scenarios,
-in order to quantify the expected range of hard binary eccentricities, which is critical for merger timescale predictions for future gravitational wave observation missions.
+in order to quantify the expected range of hard binary eccentricities, which is critical for predictions for current and future gravitational wave observation missions.
 
 % FINAL BITS
-\section*{Acknowledgements}
+\begin{acknowledgments} 
+\modulolinenumbers[100]
 
+We thank Shihong Liao, Dimitrios Irodotou, and Francesco Paolo Rizzuto for interesting discussions.
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
@@
 
-We list here the roles and contributions of the authors according to the Contributor Roles Taxonomy (CRediT\footnote{https://credit.niso.org}). 
-\textbf{AR}: Conceptualisation, Investigation, Formal analysis, Data curation, Writing -- original draft. 
-\textbf{MM}: Conceptualisation, Formal analysis, Writing -- original draft. 
+We list here the roles and contributions of the authors according to the Contributor Roles Taxonomy (\href{https://credit.niso.org}{CRediT}). 
+\textbf{AR}: Conceptualization, Investigation, Formal analysis, Data curation, Writing -- original draft. 
+\textbf{MM}: Conceptualization, Formal analysis, Writing -- original draft. 
 \textbf{PHJ}: Supervision, Writing -- review \& editing.
 \textbf{TN}:  Writing -- review \& editing.
+\end{acknowledgments}
 
-\textit{Software:} \ketju{} \citep{mannerkoski2023,rantala2017}, \gadget{} \citep{springel2021}, NumPy \citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007}, pygad \citep{rottgers2020}.
+\software{
+\ketju{} \citep{mannerkoski2023,rantala2017},
+\gadget{} \citep{springel2021},
+NumPy \citep{harris2020},
+SciPy \citep{virtanen2020},
+Matplotlib \citep{hunter2007},
+pygad \citep{rottgers2020}.
+}
 
@@
 
-\bibliography{ref}{}
+\bibliography{../ref}{}
 \bibliographystyle{aasjournal}

### 2023-09-20 papers/paper-eccentricity/paper-eccentricity.tex
Final version of Reviving Stochasticity paper

@@
-%% Based on file 'sample631.tex'
-%%
-%% Modified 2021 March
-%%
-%% This is a sample manuscript marked up using the
-%% AASTeX v6.31 LaTeX 2e macros.
-%%
-%% AASTeX is now based on Alexey Vikhlinin's emulateapj.cls 
-%% (Copyright 2000-2015).  See the classfile for details.
+% mnras_template.tex 
+%
+% LaTeX template for creating an MNRAS paper
+%
+% v3.0 released 14 May 2015
+% (version numbers match those of mnras.cls)
+%
+% Copyright (C) Royal Astronomical Society 2015
+% Authors:
+% Keith T. Smith (Royal Astronomical Society)
 
-%% AASTeX requires revtex4-1.cls and other external packages such as
-%% latexsym, graphicx, amssymb, longtable, and epsf.  Note that as of 
-%% Oct 2020, APS now uses revtex4.2e for its journals but remember that 
-%% AASTeX v6+ still uses v4.1. All of these external packages should 
-%% already be present in the modern TeX distributions but not always.
-%% For example, revtex4.1 seems to be missing in the linux version of
-%% TexLive 2020. One should be able to get all packages from www.ctan.org.
-%% In particular, revtex v4.1 can be found at 
-%% https://www.ctan.org/pkg/revtex4-1.
+% Change log
+%
+% v3.0 May 2015
+%    Renamed to match the new package name
+%    Version number matches mnras.cls
+%    A few minor tweaks to wording
+% v1.0 September 2013
+%    Beta testing only - never publicly released
+%    First version: a simple (ish) template for creating an MNRAS paper
 
-%% The first piece of markup in an AASTeX v6.x document is the \documentclass
-%% command. LaTeX will ignore any data that comes before this command. The 
-%% documentclass can take an optional argument to modify the output style.
-%% The command below calls the preprint style which will produce a tightly 
-%% typeset, one-column, single-spaced document.  It is the default and thus
-%% does not need to be explicitly stated.
-%%
-%% using aastex version 6.3
-\documentclass[twocolumn]{aastex631}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Basic setup. Most papers should leave these options alone.
+\documentclass[fleqn,usenatbib]{mnras}
 
-%% The default is a single spaced, 10 point font, single spaced article.
-%% There are 5 other style options available via an optional argument. They
-%% can be invoked like this:
-%%
-%% \documentclass[arguments]{aastex631}
-%% 
-%% where the layout options are:
-%%
-%%  twocolumn   : two text columns, 10 point font, single spaced article.
-%%                This is the most compact and represent the final published
-%%                derived PDF copy of the accepted manuscript from the publisher
-%%  manuscript  : one text column, 12 point font, double spaced article.
-%%  preprint    : one text column, 12 point font, single spaced article.  
-%%  preprint2   : two text columns, 12 point font, single spaced article.
-%%  modern      : a stylish, single text column, 12 point font, article with
-%% 		  wider left and right margins. This uses the Daniel
-%% 		  Foreman-Mackey and David Hogg design.
-%%  RNAAS       : Supresses an abstract. Originally for RNAAS manuscripts 
-%%                but now that abstracts are required this is obsolete for
-%%                AAS Journals. Authors might need it for other reasons. DO NOT
-%%                use \begin{abstract} and \end{abstract} with this style.
-%%
-%% Note that you can submit to the AAS Journals in any of these 6 styles.
-%%
-%% There are other optional arguments one can invoke to allow other stylistic
-%% actions. The available options are:
-%%
-%%   astrosymb    : Loads Astrosymb font and define \astrocommands. 
-%%   tighten      : Makes baselineskip slightly smaller, only works with 
-%%                  the twocolumn substyle.
-%%   times        : uses times font instead of the default
-%%   linenumbers  : turn on lineno package.
-%%   trackchanges : required to see the revision mark up and print its output
-%%   longauthor   : Do not use the more compressed footnote style (default) for 
-%%                  the author/collaboration/affiliations. Instead print all
-%%                  affiliation information after each name. Creates a much 
-%%                  longer author list but may be desirable for short 
-%%                  author papers.
-%% twocolappendix : make 2 column appendix.
-%%   anonymous    : Do not show the authors, affiliations and acknowledgments 
-%%                  for dual anonymous review.
-%%
-%% these can be used in any combination, e.g.
-%%
-%% \documentclass[twocolumn,linenumbers,trackchanges]{aastex631}
-%%
-%% AASTeX v6.* now includes \hyperref support. While we have built in specific
-%% defaults into the classfile you can manually override them with the
-%% \hypersetup command. For example,
-%%
-%% \hypersetup{linkcolor=red,citecolor=green,filecolor=cyan,urlcolor=magenta}
-%%
-%% will change the color of the internal links to red, the links to the
-%% bibliography to green, the file links to cyan, and the external links to
-%% magenta. Additional information on \hyperref options can be found here:
-%% https://www.tug.org/applications/hyperref/manual.html#x1-40003
-%%
-%% Note that in v6.3 "bookmarks" has been changed to "true" in hyperref
-%% to improve the accessibility of the compiled pdf file.
-%%
-%% If you want to create your own macros, you can do so
-%% using \newcommand. Your macros should appear before
-%% the \begin{document} command.
-%%
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{bm} % italic bold math symbols
+% MNRAS is set in Times font. If you don't have this installed (most LaTeX
+% installations will be fine) or prefer the old Computer Modern fonts, comment
+% out the following line
+\usepackage{newtxtext,newtxmath}
+% Depending on your LaTeX fonts installation, you might get better results with one of these:
+%\usepackage{mathptmx}
+%\usepackage{txfonts}
 
-\newcommand{\vdag}{(v)^\dagger}
-\newcommand\aastex{AAS\TeX}
-\newcommand\latex{La\TeX}
+% Use vector fonts, so it zooms properly in on-screen viewing software
+% Don't change these lines unless you know what you are doing
+\usepackage[T1]{fontenc}
 
+% Allow "Thomas van Noord" and "Simon de Laguarde" and alike to be sorted by "N" and "L" etc. in the bibliography.
+% Write the name in the bibliography as "\VAN{Noord}{Van}{van} Noord, Thomas"
+\DeclareRobustCommand{\VAN}[3]{#2}
+\let\VANthebibliography\thebibliography
+\def\thebibliography{\DeclareRobustCommand{\VAN}[3]{##3}\VANthebibliography}
+
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+
+% Only include extra packages if you really need them. Common packages are:
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+%\usepackage{amssymb}	% Extra maths symbols
+\usepackage{bm}         % Bold math fonts
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
+
+% Please keep new commands to a minimum, and use \newcommand not \def to avoid
+% overwriting existing commands. Example:
+%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
 \newcommand{\ketju}{\textsc{Ketju}}
-\newcommand{\mstar}{\textsc{MSTAR}}
-\newcommand{\gadget}{\textsc{Gadget-4}}
+\newcommand{\mstar}{\textsc{mstar}}
+\newcommand{\gadget}{\textsc{gadget-4}}
 
-\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}
-\newcommand{\Rvir}{\ensuremath{R_\mathrm{vir}}}
+\newcommand{\Msun}{\ensuremath{M_{\sun}}}
 
@@
 
-% drafting macros
-\newcommand{\drafting}[1]{
-{\leavevmode\color[RGB]{224, 77, 24}#1}
+\graphicspath{{figures/}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+
+% Title of the paper, and the short title which is used in the headers.
+% Keep the title short and informative.
+\title[Reviving Stochasticity]{Reviving stochasticity: uncertainty in SMBH binary eccentricity is unavoidable}
+
+% The list of authors, and the short list which is used in the headers.
+% If you need two or more lines of authors, add an extra line using \newauthor
+\author[A. Rawlings et al.]{
+Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
+Matias Mannerkoksi,$^{1}$
+Peter H. Johansson,$^{1}$
+Thorsten Naab$^{2}$
+\vspace*{0.1cm}\\%
+% List of institutions
+$^{1}$
+Department of Physics,
+Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland
+\\%
+$^{2}$
+Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1, D-85748 Garching, Germany
 }
 
-\newcommand{\phj}[1]{
-{\leavevmode\color[RGB]{0,0,255}#1}
-}
+% These dates will be filled out by the publisher
+\date{Accepted XXX. Received YYY; in original form ZZZ}
 
-%% Reintroduced the \received and \accepted commands from AASTeX v5.2
-%\received{March 1, 2021}
-%\revised{April 1, 2021}
-%\accepted{\today}
+% Enter the current year, for the copyright statements etc.
+\pubyear{2023}
 
-%% Command to document which AAS Journal the manuscript was submitted to.
-%% Adds "Submitted to " the argument.
-%\submitjournal{PSJ}
+% Don't change these lines
+\begin{document}
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
 
-%% For manuscript that include authors in collaborations, AASTeX v6.31
-%% builds on the \collaboration command to allow greater freedom to 
-%% keep the traditional author+affiliation information but only show
-%% subsets. The \collaboration command now must appear AFTER the group
-%% of authors in the collaboration and it takes TWO arguments. The last
-%% is still the collaboration identifier. The text given in this
-%% argument is what will be shown in the manuscript. The first argument
-%% is the number of author above the \collaboration command to show with
-%% the collaboration text. If there are authors that are not part of any
-%% collaboration the \nocollaboration command is used. This command takes
-%% one argument which is also the number of authors above to show. A
-%% dashed line is shown to indicate no collaboration. This example manuscript
-%% shows how these commands work to display specific set of authors 
-%% on the front page.
-%%
-%% For manuscript without any need to use \collaboration the 
-%% \AuthorCollaborationLimit command from v6.2 can still be used to 
-%% show a subset of authors.
-%
-%\AuthorCollaborationLimit=2
-%
-%% will only show Schwarz & Muench on the front page of the manuscript
-%% (assuming the \collaboration and \nocollaboration commands are
-%% commented out).
-%%
-%% Note that all of the author will be shown in the published article.
-%% This feature is meant to be used prior to acceptance to make the
-%% front end of a long author article more manageable. Please do not use
-%% this functionality for manuscripts with less than 20 authors. Conversely,
-%% please do use this when the number of authors exceeds 40.
-%%
-%% Use \allauthors at the manuscript end to show the full author list.
-%% This command should only be used with \AuthorCollaborationLimit is used.
-
-%% The following command can be used to set the latex table counters.  It
-%% is needed in this document because it uses a mix of latex tabular and
-%% AASTeX deluxetables.  In general it should not be needed.
-%\setcounter{table}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%
-%% The following section outlines numerous optional output that
-%% can be displayed in the front matter or as running meta-data.
-%%
-%% If you wish, you may supply running head information, although
-%% this information may be modified by the editorial offices.
-\shorttitle{Reviving Stochasticity}
-\shortauthors{A. Rawlings et al.}
-%%
-%% You can add a light gray and diagonal water-mark to the first page 
-%% with this command:
-%% \watermark{text}
-%% where "text", e.g. DRAFT, is the text to appear.  If the text is 
-%% long you can control the water-mark size with:
-%% \setwatermarkfontsize{dimension}
-%% where dimension is any recognized LaTeX dimension, e.g. pt, in, etc.
-%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\graphicspath{{./}{figures/}}
-%% This is the end of the preamble.  Indicate the beginning of the
-%% manuscript itself with \begin{document}.
-
-\begin{document}
-
-\title{Reviving stochasticity: uncertainty in SMBH binary eccentricity is unavoidable}
-
-%% LaTeX will automatically break titles if they run longer than
-%% one line. However, you may use \\ to force a line break if
-%% you desire. In v6.31 you can include a footnote in the title.
-
-%% A significant change from earlier AASTEX versions is in the structure for 
-%% calling author and affiliations. The change was necessary to implement 
-%% auto-indexing of affiliations which prior was a manual process that could 
-%% easily be tedious in large author manuscripts.
-%%
-%% The \author command is the same as before except it now takes an optional
-%% argument which is the 16 digit ORCID. The syntax is:
-%% \author[xxxx-xxxx-xxxx-xxxx]{Author Name}
-%%
-%% This will hyperlink the author name to the author's ORCID page. Note that
-%% during compilation, LaTeX will do some limited checking of the format of
-%% the ID to make sure it is valid. If the "orcid-ID.png" image file is 
-%% present or in the LaTeX pathway, the OrcID icon will appear next to
-%% the authors name.
-%%
-%% Use \affiliation for affiliation information. The old \affil is now aliased
-%% to \affiliation. AASTeX v6.31 will automatically index these in the header.
-%% When a duplicate is found its index will be the same as its previous entry.
-%%
-%% Note that \altaffilmark and \altaffiltext have been removed and thus 
-%% can not be used to document secondary affiliations. If they are used latex
-%% will issue a specific error message and quit. Please use multiple 
-%% \affiliation calls for to document more than one affiliation.
-%%
-%% The new \altaffiliation can be used to indicate some secondary information
-%% such as fellowships. This command produces a non-numeric footnote that is
-%% set away from the numeric \affiliation footnotes.  NOTE that if an
-%% \altaffiliation command is used it must come BEFORE the \affiliation call,
-%% right after the \author command, in order to place the footnotes in
-%% the proper location.
-%%
-%% Use \email to set provide email addresses. Each \email will appear on its
-%% own line so you can put multiple email address in one \email call. A new
-%% \correspondingauthor command is available in V6.31 to identify the
-%% corresponding author of the manuscript. It is the author's responsibility
-%% to make sure this name is also in the author list.
-%%
-%% While authors can be grouped inside the same \author and \affiliation
-%% commands it is better to have a single author for each. This allows for
-%% one to exploit all the new benefits and should make book-keeping easier.
-%%
-%% If done correctly the peer review system will be able to
-%% automatically put the author and affiliation information from the manuscript
-%% and save the corresponding author the trouble of entering it by hand.
-
-%\correspondingauthor{August Muench}
-%\email{greg.schwarz@aas.org, gus.muench@aas.org}
-
-\author[0000-0003-1807-6321]{Alexander Rawlings}
-\affiliation{Department of Physics,
-Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland}
-
-\author[0000-0001-5721-9335]{Matias Mannerkoski}
-\affiliation{Department of Physics,
-Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland}
-
-\author[0000-0001-8741-8263]{Peter H. Johansson}
-\affiliation{Department of Physics,
-Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland}
-
-\author[0000-0002-7314-2558]{Thorsten Naab}
-\affiliation{Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1,
-D-85748 Garching, Germany}
-
-\correspondingauthor{Alexander Rawlings}
-\email{alexander.rawlings@helsinki.fi}
-
-%% Note that the \and command from previous versions of AASTeX is now
-%% depreciated in this version as it is no longer necessary. AASTeX 
-%% automatically takes care of all commas and "and"s between authors names.
-
-%% AASTeX 6.31 has the new \collaboration and \nocollaboration commands to
-%% provide the collaboration status of a group of authors. These commands 
-%% can be used either before or after the list of corresponding authors. The
-%% argument for \collaboration is the collaboration identifier. Authors are
-%% encouraged to surround collaboration identifiers with ()s. The 
-%% \nocollaboration command takes no argument and exists to indicate that
-%% the nearby authors are not part of surrounding collaborations.
-
-%% Mark off the abstract in the ``abstract'' environment. 
+% Abstract of the paper
 \begin{abstract}
-We study supermassive black hole (SMBH) binary eccentricity of equal-mass galaxy mergers in $N$-body simulations with the \ketju{} code, which combines the \gadget{} fast multipole gravity solver with accurate regularized integration and post-Newtonian corrections around SMBHs.
-In simulations with realistic, high eccentricity galactic merger orbits, the binary eccentricity is found to be a non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
+We study supermassive black hole (SMBH) binary eccentricity of equal-mass galaxy mergers in $N$-body simulations with the \ketju{} code, which combines the \gadget{} fast multipole gravity solver with accurate regularised integration and post-Newtonian corrections around SMBHs.
+In simulations with realistic, high eccentricity galactic merger orbits, the hard binary eccentricity is found to be a non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
 This mapping between the deflection angle and the binary eccentricity has no apparent resolution dependence in our simulations spanning the resolution range of $1\times10^5$--$8\times10^6$ particles per galaxy.
 The mapping is also captured using a simple model with an analytic potential, indicating that it is driven by the interplay between a smooth asymmetric stellar background potential and dynamical friction acting on the SMBHs.
-Due to the non-linearity of this mapping, in certain merger configurations small, parsec-scale variations in the merger orbit can result in binary eccentricities varying in nearly the full possible range between $e=0$ and $e=1$.
-In idealized simulations, such variations are caused by finite resolution effects,
+Due to the non-linearity of this mapping, in eccentric major merger configurations small, parsec-scale variations in the merger orbit can result in binary eccentricities varying in nearly the full possible range between $e=0$ and $e=1$.
+In idealised simulations, such variations are caused by finite resolution effects,
 and convergence of the binary eccentricity can be achieved with increasing resolution.
@@
 Our results indicate that the distribution of these effectively random eccentricities can be studied using even moderate resolution simulations.
-
 \end{abstract}
 
-%% Keywords should appear after the \end{abstract} command. 
-%% The AAS Journals now uses Unified Astronomy Thesaurus concepts:
-%% https://astrothesaurus.org
-%% You will be asked to selected these concepts during the submission process
-%% but this old "keyword" functionality is maintained in case authors want
-%% to include these concepts in their preprints.
-%\keywords{Galaxy mergers (608); Supermassive black holes (1663); Eccentricity (441); N-body simulations (1083); }
+% Select between one and six entries from the list of approved keywords.
+% Don't make up new ones.
+\begin{keywords}
+black hole physics -- galaxies: kinematics and dynamics -- methods: numerical -- software: simulations
+\end{keywords}
 
-%% From the front matter, we move on to the body of the paper.
-%% Sections are demarcated by \section and \subsection, respectively.
-%% Observe the use of the LaTeX \label
-%% command after the \subsection to give a symbolic KEY to the
-%% subsection for cross-referencing in a \ref command.
-%% You can use LaTeX's \ref and \label commands to keep track of
-%% cross-references to sections, equations, tables, and figures.
-%% That way, if you change the order of any elements, LaTeX will
-%% automatically renumber them.
-%%
-%% We recommend that authors also use the natbib \citep
-%% and \citet commands to identify citations.  The citations are
-%% tied to the reference list via symbolic KEYs. The KEY corresponds
-%% to the KEY in the \bibitem in the reference list below.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-%-------------------------- MAIN TEXT STARTS HERE --------------------------%
+%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+
 % INTRODUCTION
 \section{Introduction}
-Supermassive black holes (SMBHs) are believed to reside at the centers of all massive galaxies \citep[e.g.][]{kormendy2013}.
+Supermassive black holes (SMBHs) are believed to reside at the centres of all massive galaxies \citep[e.g.][]{kormendy2013}.
 In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{volonteri2003,naab2017} their SMBHs 
@@
 becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence, where the strength of the GW emission strongly depends on the
-binary eccentricity \citep{peters1963,peters1964,mannerkoski2019}.
+SMBHs to coalescence \citep{peters1963,peters1964}.
 
 The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{berentzen2009,khan2011,dosopoulou2017,mannerkoski2019}
-to provide quantitative predictions for ongoing observational programs such as ground-based pulsar timing arrays \citep[PTAs,][]{agazie2023,2023Antoniadis,2023Xu,zic2023}
+to provide quantitative predictions for ongoing observational programmes such as ground-based pulsar timing arrays \citep[PTAs,][]{agazie2023,2023Antoniadis,2023Xu,zic2023}
 and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2023}.
-Understanding how faithfully SMBH binary eccentricity is captured in numerical simulations is thus critical in constraining the SMBH merger rate
-that the observational GW community aims to detect.
+In particular, the eccentricity of the binary significantly affects the GW emission, with higher eccentricities resulting both in faster mergers as well as changes in the emitted GW spectrum \citep{enoki2007, huerta2015, taylor2016}.
+Understanding SMBH binary eccentricity and how faithfully it is captured in numerical simulations is thus critical for predicting and interpreting observations done with instruments such as PTAs and LISA.
 
-The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionsless merger simulations \citep[e.g.][]{berentzen2009,vasiliev2015,bortolas2016,gualandris2017,gualandris2022}.
+The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionless merger simulations \citep[e.g.][]{berentzen2009,vasiliev2015,bortolas2016,gualandris2017,gualandris2022}.
 Recently, \citet{nasim2020} have argued that the substantial scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
-discretization, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
+discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
 
-In this Letter, we find that for realistic galaxy merger orbits the scatter in the final SMBH binary eccentricity is due to the 
+In this paper, we find that for realistic galaxy merger orbits the scatter in the final SMBH binary eccentricity is due to the 
 physical sensitivity of the final eccentricity to small perturbations of the final  nearly radial plunging trajectory of the SMBHs before they become bound, even in the infinite resolution limit.
 As a result, we argue that the stochasticity of the binary eccentricity is an unavoidable physical feature of realistic galaxy mergers, at least in the equal-mass case.
+
+This paper is structured as follows. In Section \ref{sec:num_sims} we briefly discuss the main features
+of the \ketju{} code and our numerical simulations.
+In Section \ref{sec:results} we study the eccentricity scatter in our simulations and also present a simple model that captures the observed scatter.
+We discuss our results and their implications in Section \ref{sec:discussion} and finally present our conclusions in Section \ref{sec:conclusions}.
 
@@
     \caption{Binary eccentricity $e$ (note the non-linear scale) and shifted time $t'=t-t_\mathrm{bound}$ for the $M_\bullet/m_\star=5000$ resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
-    The downwards-pointing arrows indicate the median time when the SMBH binaries becomes hard. 
-    In the $e_0=0.90$ set, six simulations have a binary merger within $50\,\mathrm{Myr}$.}
+    The circle markers indicate the time when a hard binary formed, with the preceding semitransparent lines indicating the transition period when the hard binary is still forming, and the solid lines the hard binary evolution.
+    During the transition period, the Keplerian definition of binary motion is affected by the intervening stellar mass, creating the oscillatory artefacts in the semitransparent lines.
+    In the $e_0=0.90$ simulation set, six simulations have a binary merger within $50\,\mathrm{Myr}$.}
     \label{fig:eccentricities}
@@
 \section{Numerical Simulations}
+\label{sec:num_sims}
 
-We construct a number of idealized galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017}.
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017}.
 The dynamics of
-SMBHs, and stars in a small region around them, are integrated with an algorithmically regularized integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
+SMBHs, and stars in a small region around them, are integrated with an algorithmically regularised integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
 is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
@@
 
-Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, matching the models of \citet{nasim2020}.
+Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, and are designed to match exactly the models of \citet{nasim2020}.
 Each galaxy consists of a stellar-only \citet{dehnen1993} sphere with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
@@
 \end{equation}
-The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the center of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed.
+The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed\footnote{We set the length and mass scales of the system to match one of those presented in \citet{gualandris2022}.
+However, as the performed simulations are gravity-only, the system may be transformed to other mass and length scales with the same $M_\mathrm{tot}/a$ ratio without affecting the results.}.
+Our galaxies lie on the observed $M_\bullet\text{--}\sigma$ relation presented in \citet{vandenbosch2016}.
 We test seven different mass resolutions 
@@
 The $e_0=0.90$ merger orbit matches that used by \citet{nasim2020}, with the radial and tangential velocities being consistent with the values reported in \citet{gualandris2022}.
-For each orbital configuration, we run ten realizations for each mass resolution, to account for stochasticity caused by the discretized phase space. 
-Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
+For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space. 
+Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
 
@@
     \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the total number of stellar particles $N_{\star, \mathrm{tot}}$. 
+    The results from \citet{nasim2020} are shown as teal triangles for comparison.
     The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ mergers. 
@@
 The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$, where $t_\mathrm{bound}$ is the time when the binary orbital energy $E$ becomes permanently negative.
-The median time when each simulation set forms a hard binary is similar between the two sets, as indicated by the arrows in \autoref{fig:eccentricities}.
-The $e_0=0.90$ set also demonstrates six SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid orbit circularization in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
-The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $e=[0,1]$.
-Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence during the $50\,\mathrm{Myr}$ timespan.
-
-To characterize the scatter in eccentricity, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centered on the orbit within which the SMBH binary has become hard.
+At the time $t_\mathrm{bound}$ the SMBH binary is not yet isolated due to the large quantity of intervening stellar mass, which causes the Keplerian definition of eccentricity to oscillate and not instantaneously match exactly the actual SMBH binary orbit.
+In \autoref{fig:eccentricities} we indicate the time of hard binary formation by circle markers; from this point the Keplerian eccentricity appropriately describes the binary orbit.
 The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}:
@@
 where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$).
-We characterize the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
+Before a hard binary forms, both the $e_0=0.90$ and $e_0=0.99$ sets initially demonstrate an overall decrease in eccentricity (semitransparent lines in \autoref{fig:eccentricities}) until $t'\sim1\,\mathrm{Myr}$.
+It should be noted that the precise value of the eccentricity in this regime is not robust due to the SMBHs orbiting in a potential still largely influenced by the stellar background, however the overall trend is consistent with the binary orbital geometry.
+After $t'\sim1\,\mathrm{Myr}$, some of the $e_0=0.99$ realisations continue to have a decreasing eccentricity, and all of the $e_0=0.90$ realisations have an increasing eccentricity.
+We discuss the mechanism for this in \autoref{ssec:toy}.
+
+The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $e=[0,1]$.
+As a test we also perform two sets of ten runs using the $10^6$ particle resolution set up, but reduce the SMBH mass to $5.0\times10^7\,\Msun$ and $1.0\times10^7\,\Msun$, and observe the same qualitative spread in eccentricity as in the $M_\bullet=10^8\,\Msun$ case.
+Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence during the $50\,\mathrm{Myr}$ timespan.
+The $e_0=0.90$ set demonstrates six SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid orbit circularisation in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
+
+To characterise the scatter in eccentricity, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard (equation \eqref{eq:ahard}).
+We characterise the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
 
@@
 For the $e_0=0.90$ orbit, the convergence of $\sigma_e$ scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, where $N_{\star,\mathrm{tot}}$ is the total number of stellar particles in the merger.
-The values of $\sigma_e$ we report are quantitatively similar to the values found by \citet{nasim2020} for the same system. 
+The values of $\sigma_e$ we report are quantitatively similar to the values found by \citet{nasim2020} for the same system, as can be seen by the teal triangles in \autoref{fig:convergence}.
 The variation in eccentricity does not show the same scaling for the $e_0=0.99$ orbit as the $e_0=0.90$ orbit. 
@@
     \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realizations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about $x=-z$.
-    The line gradient shows the shifted time $t'=t-t_\mathrm{bound}$, however the coloring for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity. 
-    In each inset panel, $\theta_\mathrm{defl}$ is the deflection angle at the pericenter between the arrows.
+    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about $x=-z$.
+    The line gradient shows the shifted time $t'=t-t_\mathrm{bound}$, however the colouring for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity. 
+    In each inset panel, $\theta_\mathrm{defl}$ is the deflection angle at the pericentre between the arrows.
     The circle marker indicates the SMBH position when a bound binary is formed.}
@@
 
-\subsection{Binary binding process and the scatter in eccentricity}
+\subsection{Binary binding process and the scatter in eccentricity}\label{ssec:binding}
 
 To understand the observed scatter in eccentricity, we investigate the binary binding process in each simulation. 
-Before the SMBHs are bound, the primary influence on the motion is from the galactic merger potential; two realizations for both merger orbits are shown in \autoref{fig:orbit}.
-The initial stages of the orbit show indiscernible variation between realizations, however after a particularly strong interaction between the SMBHs during a pericenter passage, the orbits deviate between realizations (inset panels, \autoref{fig:orbit}), and evolve to different binary eccentricities.
+Before the SMBHs are bound, the primary influence on the motion is from the galactic merger potential; two realisations for both merger orbits are shown in \autoref{fig:orbit}.
+The initial stages of the orbit show indiscernible variation between realisations, however after a particularly strong interaction between the SMBHs during a pericentre passage, the orbits deviate between realisations (inset panels, \autoref{fig:orbit}), and evolve to different binary eccentricities.
 
-To quantify the strength of the hard-scattering process that randomizes the SMBH orbit, we measure the effective two-body deflection angle, defined as:
+To quantify the strength of the hard-scattering process that randomises the SMBH orbit, we measure the effective two-body deflection angle, defined as:
 \begin{equation}\label{eq:theta}
@@
 where $M=M_{\bullet,1}+M_{\bullet,2}$, $L$ is the magnitude of the SMBH 
-system angular momentum vector, $E$ the orbital energy at the time of the pericenter passage, $b$ the effective impact parameter, and $b_{90}$ the $90\degr$ deflection radius \citep{binney2008}.
+system angular momentum vector, $E$ the orbital energy at the time of the pericentre passage, $b$ the effective impact parameter, and $b_{90}$ the $90\degr$ deflection radius \citep{binney2008}.
 
 We observe clear evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, as shown with the data points in \autoref{fig:theta_e}.
+We also observe a similar relationship in our two test simulation sets with the reduced SMBH masses.
 
@@
     \caption{
-    Deflection angle $\theta_\mathrm{defl}$ and resulting binary eccentricity
+    Deflection angle $\theta_\mathrm{defl}$ and resulting hard binary eccentricity
     in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
@@
 
-\subsection{Reproducing the eccentricity behavior with a simple model}\label{ssec:toy}
+\subsection{Reproducing the eccentricity behaviour  with a simple model}\label{ssec:toy}
 
@@
 
-The stellar background is modeled as a constant density spheroidal potential
+The stellar background is modelled as a constant density spheroidal potential
 \citep[e.g.][]{binney2008}:
@@
 
-The DF acting on a single BH is modeled using the \citet{chandrasekhar1943} formula assuming
+The DF acting on a single BH is modelled using the \citet{chandrasekhar1943} formula assuming
 a Maxwellian distribution with a constant velocity dispersion $\sigma_\star$ \citep[e.g.][]{binney2008}:
 \begin{equation}\label{eq:chandra_df}
-\begin{aligned}
-\vb{a}_{1,\mathrm{DF}} = &-\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}\\
-                         &\times\left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
-\end{aligned}
+\vb{a}_{1,\mathrm{DF}} = -\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}
+                         \left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
 \end{equation}
 where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
-and $X = |\vb{v}_1|/\left(\sqrt{2}\sigma_\star\right)$.
+and $X = |\vb{v}_1|/(\sqrt{2}\sigma_\star)$.
 We set ${\sigma_\star=200\,\mathrm{km\,s^{-1}}}$ based on the value measured from the simulations.
@@
 the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
-although the simple model does appear to capture the mean behavior relatively well, apart from the very largest values of
+although the simple model does appear to capture the mean behaviour  relatively well, apart from the very largest values of
 $\theta_\mathrm{defl}$.
@@
 
-However, in general the behavior of the simulation data is captured well,
+However, in general the behaviour  of the simulation data is captured well,
 with the model correctly producing the two main eccentricity minima,
 as well as the $e\approx 1$ region between them.
-The minima occur for trajectories where the torque from the background potential together with the DF
-causes the SMBHs to loop around into a nearly circular orbit,
-while highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along 
-the main axes of the potential, as shown in \autoref{fig:model_orbits}.
+The predicted minima occur for trajectories where the torque from the background potential together with the DF
+causes the SMBHs to loop around into a nearly circular orbit, resulting in the  low eccentricities seen in some of the $e_0=0.99$ simulation data.
+Conversely, highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along 
+the main axes of the potential.
+These mechanisms are contrasted in \autoref{fig:model_orbits}.
 
 % DISCUSSION
-\section{Discussion and Conclusions}
+\section{Discussion}
+\label{sec:discussion}
 
-We find that the variation in eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the variation in the deflection angle $\theta_\mathrm{defl}$, independent of the simulation mass resolution.
-Together with the simple analytic model that reproduces well the behavior seen in the simulations,
-this demonstrates that the sensitivity of the SMBH binary eccentricity to slight changes in the merger orbit is a physical feature of galactic major mergers on highly radial orbits, and not an error caused by insufficient numerical resolution.
-
-As shown by both the simulations and the analytic model, the eccentricity of the SMBH binary can span nearly the full possible range between $e=[0,1]$, depending on comparatively small, parsec-scale differences in the particular realization of the galactic merger orbit.
+As shown by both the simulations and the analytic model, the eccentricity of the hard SMBH binary in the studied merger configurations can span nearly the full possible range in the interval $e_\mathrm{h}=[0,1]$, depending on comparatively small, parsec-scale differences in the particular realisation of the galactic merger orbit.
 On the other hand, in minor mergers where the stellar background is less asymmetric when the SMBHs become bound,
@@
 
-Lower eccentricity merger orbits can be expected to show less scatter in the binary eccentricity due to the lack of a hard scattering event that is sensitive to slight perturbations in the merger orbit, which was also seen in the study by \citet{gualandris2022}.
+Merger orbits with a lower initial eccentricity $e_0$ than the values in this work can be expected to show less scatter in the binary eccentricity due to the lack of a hard scattering event that is sensitive to slight perturbations in the merger orbit, which was also seen in the study by \citet{gualandris2022}.
 However, major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006}.
@@
 
-The simulation setup used in this work is idealized, and covers only a small part of the parameter space of merger configurations.
-However, similar behavior can be expected to occur also in more realistic models,
-since the relevant dynamics occurs within a few hundred parsecs from the center of the merged galaxy,
+The simulation setup used in this work is idealised, and covers only a small part of the parameter space of merger configurations.
+However, similar behaviour  can be expected to occur also in more realistic models,
+since the relevant dynamics occurs within a few hundred parsecs from the centre of the merged galaxy,
 where we do not expect any significant effects from dark matter or the outer parts of the galaxy, which are not included in our present simulations.
+Preliminary results using steeper $(\gamma \gtrsim 1)$, more observationally-motivated galaxy density profiles indicate that similar behaviour occurs also in more realistic merger systems.
 
 The variation of the deflection angle is caused by the random variation of the impact parameter $b$ through equation~\eqref{eq:theta}.
-In numerical simulations with a discretized phase space, the primary cause of random variation in $b$ is the Brownian motion of the SMBHs before they form a bound binary \citep{merritt2001,bortolas2016}.
+In numerical simulations with a discretised phase space, the primary cause of random variation in $b$ is the Brownian motion of the SMBHs before they form a bound binary \citep{merritt2001,bortolas2016}.
 This effect scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
@@
 The deviation from this scaling seen in \autoref{fig:convergence} can then be explained by the fact that the eccentricity is not globally a linear function of the impact parameter. 
+The non-linear mapping also explains why GW-induced mergers were present within a short ($<50\,\mathrm{Myr}$) timeframe for the $e_0=0.90$ mergers and not the $e_0=0.99$ mergers in \autoref{fig:eccentricities}, and highlights the difficulty in predicting the binary eccentricity $e_\mathrm{h}$ from the initial merger orbit.
 
-While the scatter in the impact parameter between different numerical realizations of the same merger is due to relatively low number of particles compared to real galaxies,
-we expect that uncertainty of a similar magnitude is also present in real systems due to various mechanisms, such as perturbations from gas and substructure in their stellar environment.
-In addition, the SMBHs are also not necesssarily located exactly at the center of mass of the galaxy.
-For example, the work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photocenter by up to $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
+While the scatter in the impact parameter between different numerical realisations of the same merger is due to relatively low number of particles compared to real galaxies,
+we expect that uncertainty of a similar magnitude is also present in real systems due to various mechanisms, such as perturbations of the SMBH motion from gas and substructure in their stellar environment, or larger-scale perturbations of the galactic merger orbit.
+In addition, the SMBHs are also not necesssarily located exactly at the centre of mass of the galaxy.
+For example, the work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photocentre by up to $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
 Thus, it is not possible to give an exact prediction for the SMBH binary eccentricity produced by a given merger,
@@
 
-In conclusion, we have shown the importance of the deflection angle $\theta_\mathrm{defl}$, and thus the impact parameter $b$, on the resulting hard binary eccentricity $e_\mathrm{h}$. 
+Even without such extensive studies on the full distribution of SMBH binary eccentricities, the present results suggest that the distribution is likely to be broad and contain a significant number of binaries at high eccentricities.
+The possibility of a large fraction of high binary eccentricities has implications in particular for studies of the expected gravitational wave background (GWB) signal measured by PTAs, as in particular binaries with $e_\mathrm{h}\gtrsim 0.9$ can retain significant eccentricities while their GW emission is within the observable frequency band.
+This could be observable in the shape of the GWB spectrum \citep{kelley2017}, and would also affect the likelihood of detecting individual SMBH binaries with PTAs \citep{kelley2018}.
+It is interesting to note that \citet{bi2023} obtained a distribution of eccentricities consistent with the expectation of a significant fraction of SMBH binaries at high eccentricities based on analysis of the recent PTA observations.
+
+In the present work, the influences of rotation and SMBH orbital flips are not considered. 
+A counter-rotating stellar background with respect to the SMBH binary angular momentum has a tendency to increase the binary eccentricity, whereas a corotating stellar background tends to circularise the binary \citep{sesana2011, holley-bockelmann2015}.
+The SMBH angular momentum vector can however undergo reversals as a result of gravitational torques following pericentre passages of the binary \citep[][termed an orbital flip]{rantala2019} an example of which is shown in the left inset of of the left panel of \autoref{fig:orbit}.
+The chaotic reversals of the binary angular momentum act to randomise the relative rotation of the stellar background, leading in general to complex eccentricity evolution \citep{nasim2021}.
+A result of a sustained co- or counter-rotating stellar background could be an evolution of the binary eccentricity after $e_\mathrm{h}$ to a bimodal distribution for a given merger configuration, if the stellar scattering interactions were strong enough.
+Whilst these mechanisms are unlikely to alter much our findings concerning the hard binary eccentricity, they may obfuscate predictions for SMBH binary eccentricity prior to the merger, which is critical for GW detection missions.
+
+In contrast to the gas-free and fairly low density systems studied here, in higher density or gas-rich systems the binary eccentricity can also evolve significantly after the binary has become hard but before it has entered the GW emission dominated regime.
+In particular, interactions with a circumbinary accretion disc can lead to rapid evolution of the eccentricity if the accretion rate is high.
+Studies of prograde coplanar circumbinary discs have found that at almost all eccentricities the binaries evolve towards an eccentricity of $e\approx 0.5$ \citep{zrake2021,dorazio2021}, whereas retrograde discs lead to the eccentricity always increasing \citep{tiede2023}.
+However, these studies are generally limited to eccentricities below $e<0.9$.
+Our results indicate the need for extending such studies to higher binary eccentricities, as well as to more general configurations such as polar discs which become increasingly likely at high eccentricities \citep[e.g.][]{martin2017}.
+
+% CONCLUSION
+\section{Conclusions}
+\label{sec:conclusions}
+In conclusion, we find that the variation in the hard binary eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the deflection angle $\theta_\mathrm{defl}$, and thus the impact parameter $b$.
 By using a simple, resolution-free analytic model of the SMBH scattering process,
-we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretization effects in galaxy merger simulations,
+we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations,
 but is rather due to the physical sensitivity of the system to small changes in the merger orbit,
-which can be caused by physical mechanisms in addition to numerical discretization effects.
+which can be caused by physical mechanisms in addition to numerical discretisation effects.
 The results presented here justify extending the investigation to more realistic galaxy merger scenarios,
@@
 % FINAL BITS
-\begin{acknowledgments} 
-\modulolinenumbers[100]
-
-We thank Shihong Liao, Dimitrios Irodotou, and Francesco Paolo Rizzuto for interesting discussions.
+\section*{acknowledgments}
+We thank the anonymous referee for their comments which contributed to improving this work.
+We also thank Shihong Liao, Dimitrios Irodotou, and Francesco Paolo Rizzuto for interesting discussions.
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
@@
 The numerical simulations used computational resources provided by
-the CSC -- IT Center for Science, Finland.
+the CSC -- IT centre for Science, Finland.
 
+\section*{Author contributions}
 We list here the roles and contributions of the authors according to the Contributor Roles Taxonomy (\href{https://credit.niso.org}{CRediT}). 
-\textbf{AR}: Conceptualization, Investigation, Formal analysis, Data curation, Writing -- original draft. 
-\textbf{MM}: Conceptualization, Formal analysis, Writing -- original draft. 
+\textbf{AR}: Conceptualisation, Investigation, Formal analysis, Data curation, Writing -- original draft. 
+\textbf{MM}: Conceptualisation, Formal analysis, Writing -- original draft. 
 \textbf{PHJ}: Supervision, Writing -- review \& editing.
 \textbf{TN}:  Writing -- review \& editing.
-\end{acknowledgments}
 
-\software{
+\section*{Software}
 \ketju{} \citep{mannerkoski2023,rantala2017},
@@
 pygad \citep{rottgers2020}.
-}
 
-%% For this sample we use BibTeX plus aasjournals.bst to generate the
-%% the bibliography. The sample631.bib file was populated from ADS. To
-%% get the citations to show in the compiled file do the following:
-%%
-%% pdflatex sample631.tex
-%% bibtext sample631
-%% pdflatex sample631.tex
-%% pdflatex sample631.tex
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Data Availability}
+The data underlying this article will be shared on reasonable request to the corresponding author.
 
-\bibliography{../ref}{}
-\bibliographystyle{aasjournal}
+%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
 
-%% This command is needed to show the entire author+affiliation list when
-%% the collaboration and author truncation commands are used.  It has to
-%% go at the end of the manuscript.
-%\allauthors
+% The best way to enter references is to use BibTeX:
 
-%% Include this line if you are using the \added, \replaced, \deleted
-%% commands to see a summary list of all changes at the end of the article.
-%\listofchanges
+\bibliographystyle{mnras}
+\bibliography{ref} % if your bibtex file is called example.bib
 
+% Alternatively you could enter them by hand, like this:
+% This method is tedious and prone to error if you have lots of references
+%\begin{thebibliography}{99}
+%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
+%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
+%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
+%Others S., 2012, Journal of Interesting Stuff, 17, 198
+%\end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
+
+%\appendix
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Don't change these lines
+\bsp	% typesetting comment
+\label{lastpage}
 \end{document}
 
-% End of file `sample631.tex'.
+% End of mnras_template.tex

### 2023-09-22 papers/paper-corekick/paper-corekick.tex
Set up core-kick paper draft

--- 2023-09-20 papers/paper-eccentricity/paper-eccentricity.tex
@@
 %%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
+% drafting macros
+\usepackage{xcolor}
+\newcommand{\drafting}[1]{
+{\leavevmode\color[RGB]{224, 77, 24}#1}
+}
 
@@
 % Keep the title short and informative.
-\title[Reviving Stochasticity]{Reviving stochasticity: uncertainty in SMBH binary eccentricity is unavoidable}
+\title[Cores from kicks]{Cores from kicks: constraints for observations of gravitational recoil in elliptical galaxies}
 
@@
 Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
-Matias Mannerkoksi,$^{1}$
-Peter H. Johansson,$^{1}$
-Thorsten Naab$^{2}$
+et al.
 \vspace*{0.1cm}\\%
@@
 \begin{abstract}
-We study supermassive black hole (SMBH) binary eccentricity of equal-mass galaxy mergers in $N$-body simulations with the \ketju{} code, which combines the \gadget{} fast multipole gravity solver with accurate regularised integration and post-Newtonian corrections around SMBHs.
-In simulations with realistic, high eccentricity galactic merger orbits, the hard binary eccentricity is found to be a non-linear function of the deflection angle in the SMBH orbit during the final, nearly radial close encounter between the SMBHs before they form a bound binary.
-This mapping between the deflection angle and the binary eccentricity has no apparent resolution dependence in our simulations spanning the resolution range of $1\times10^5$--$8\times10^6$ particles per galaxy.
-The mapping is also captured using a simple model with an analytic potential, indicating that it is driven by the interplay between a smooth asymmetric stellar background potential and dynamical friction acting on the SMBHs.
-Due to the non-linearity of this mapping, in eccentric major merger configurations small, parsec-scale variations in the merger orbit can result in binary eccentricities varying in nearly the full possible range between $e=0$ and $e=1$.
-In idealised simulations, such variations are caused by finite resolution effects,
-and convergence of the binary eccentricity can be achieved with increasing resolution.
-However, in real galaxies, other mechanisms such as nuclear gas and substructure that perturb the merger orbit are likely to be significant enough for the binary eccentricity to be effectively random.
-Our results indicate that the distribution of these effectively random eccentricities can be studied using even moderate resolution simulations.
+\drafting{Investigate the relationship between core size and kick velocity}
 \end{abstract}
@@
 \section{Introduction}
-Supermassive black holes (SMBHs) are believed to reside at the centres of all massive galaxies \citep[e.g.][]{kormendy2013}.
-In the $\Lambda$CDM model as galaxies grow through gas accretion and mergers \citep[e.g.][]{volonteri2003,naab2017} their SMBHs 
-are also expected to interact in a three-phase merger process \citep{begelman1980}.
-
-Firstly, dynamical friction \citep[DF,][]{chandrasekhar1943} acts to bring the SMBHs from kiloparsec-scales down to parsec-scale separations, after which the SMBHs form a bound binary with a semimajor axis $a$ and eccentricity $e$ \citep[e.g.][]{milosavljevic2001,merritt2013}.
-In the second phase, the
-SMBH binary separation is reduced through sequential slingshot encounters with the surrounding stellar
-distribution \citep{hills1980, hills1983, quinlan1996,rantala2018}. Finally, at small subparsec separations gravitational wave (GW) emission
-becomes the dominant mechanism by which the SMBH binary can lose its remaining orbital energy and angular momentum, thus driving the
-SMBHs to coalescence \citep{peters1963,peters1964}.
-
-The complex nature of SMBH coalescence in a galaxy merger setting necessitates the use of numerical techniques \citep[e.g.][]{berentzen2009,khan2011,dosopoulou2017,mannerkoski2019}
-to provide quantitative predictions for ongoing observational programmes such as ground-based pulsar timing arrays \citep[PTAs,][]{agazie2023,2023Antoniadis,2023Xu,zic2023}
-and the upcoming Laser Interferometer Space Antenna \citep[LISA, e.g.][]{amaro-seoane2023}.
-In particular, the eccentricity of the binary significantly affects the GW emission, with higher eccentricities resulting both in faster mergers as well as changes in the emitted GW spectrum \citep{enoki2007, huerta2015, taylor2016}.
-Understanding SMBH binary eccentricity and how faithfully it is captured in numerical simulations is thus critical for predicting and interpreting observations done with instruments such as PTAs and LISA.
-
-The SMBH binary merging process and its dependence on eccentricity and resolution has been extensively studied using collisionless merger simulations \citep[e.g.][]{berentzen2009,vasiliev2015,bortolas2016,gualandris2017,gualandris2022}.
-Recently, \citet{nasim2020} have argued that the substantial scatter in SMBH binary eccentricity observed in gas-free merger simulations is an artefact of poor phase space
-discretisation, and that in the real Universe where SMBH masses are far greater than stellar masses $(M_\bullet \gg m_\star)$, SMBH binary eccentricity is a reliably predictable quantity.
-
-In this paper, we find that for realistic galaxy merger orbits the scatter in the final SMBH binary eccentricity is due to the 
-physical sensitivity of the final eccentricity to small perturbations of the final  nearly radial plunging trajectory of the SMBHs before they become bound, even in the infinite resolution limit.
-As a result, we argue that the stochasticity of the binary eccentricity is an unavoidable physical feature of realistic galaxy mergers, at least in the equal-mass case.
-
-This paper is structured as follows. In Section \ref{sec:num_sims} we briefly discuss the main features
-of the \ketju{} code and our numerical simulations.
-In Section \ref{sec:results} we study the eccentricity scatter in our simulations and also present a simple model that captures the observed scatter.
-We discuss our results and their implications in Section \ref{sec:discussion} and finally present our conclusions in Section \ref{sec:conclusions}.
-
-\begin{figure*}
-    \centering
-    \includegraphics{eccentricities.pdf}
-    \caption{Binary eccentricity $e$ (note the non-linear scale) and shifted time $t'=t-t_\mathrm{bound}$ for the $M_\bullet/m_\star=5000$ resolution simulations from the $e_0 = 0.90$ set (blue lines) and $e_0 = 0.99$ set (orange lines).
-    The circle markers indicate the time when a hard binary formed, with the preceding semitransparent lines indicating the transition period when the hard binary is still forming, and the solid lines the hard binary evolution.
-    During the transition period, the Keplerian definition of binary motion is affected by the intervening stellar mass, creating the oscillatory artefacts in the semitransparent lines.
-    In the $e_0=0.90$ simulation set, six simulations have a binary merger within $50\,\mathrm{Myr}$.}
-    \label{fig:eccentricities}
-\end{figure*}
+\drafting{Some general intro here}
 
@@
 
+\drafting{Taken from eccentricity paper, leave here as a guide for a similar description about Ketju}
 We construct a number of idealised galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017}.
@@
 
-Our galaxy models represent the nuclear bulge of a gas-devoid elliptical galaxy, and are designed to match exactly the models of \citet{nasim2020}.
-Each galaxy consists of a stellar-only \citet{dehnen1993} sphere with shape parameter $\gamma=0.5$ and scale radius $a=186\,\mathrm{pc}$, where the Dehnen profile is given by:
+\drafting{Rantala 2018 ICs}
 \begin{equation}\label{eq:dehnen}
@@
 \end{equation}
-The total stellar mass is $M_\star=10^{10}\,\Msun$, and at the centre of the model galaxy a SMBH of mass $M_\bullet=10^8\,\Msun$ is placed\footnote{We set the length and mass scales of the system to match one of those presented in \citet{gualandris2022}.
-However, as the performed simulations are gravity-only, the system may be transformed to other mass and length scales with the same $M_\mathrm{tot}/a$ ratio without affecting the results.}.
-Our galaxies lie on the observed $M_\bullet\text{--}\sigma$ relation presented in \citet{vandenbosch2016}.
-We test seven different mass resolutions 
-with a varying number of stellar particles: $N_\star = \{1.0\times10^5,2.5\times10^5, 5.0\times10^5, 1.0\times10^6, 2.0\times10^6, 4.0\times10^6, 8.0\times10^6\}$, corresponding to $M_\bullet/m_\star= 1000 \textnormal{--} 80000$.
 
-We then construct isolated merger initial conditions by placing two galaxies on two different elliptical orbits with eccentricities of $e_0=0.90$ and $e_0=0.99$, at a fixed
-initial separation of $D=3.72\,\mathrm{kpc}$ and a semimajor axis of $a_0=2.79\,\mathrm{kpc}$.
-The $e_0=0.90$ merger orbit matches that used by \citet{nasim2020}, with the radial and tangential velocities being consistent with the values reported in \citet{gualandris2022}.
-For each orbital configuration, we run ten realisations for each mass resolution, to account for stochasticity caused by the discretised phase space. 
-Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
+\drafting{We then construct isolated merger initial conditions}. 
+Interactions between stellar particles are softened with a softening length of \drafting{$\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.}
 
@@
 \section{Results}\label{sec:results}
-
-\begin{figure}
-    \includegraphics{convergence}
-    \caption{Convergence of the eccentricity scatter $\sigma_e$ as a function of mass resolution $M_\bullet/m_\star$ and the total number of stellar particles $N_{\star, \mathrm{tot}}$. 
-    The results from \citet{nasim2020} are shown as teal triangles for comparison.
-    The expected $1/\sqrt{N_\star}$ scaling is recovered for the $e_0 = 0.90$ mergers. 
-    For the $e_0=0.99$ mergers, $\sigma_e$ does not decrease until $M_\bullet/m_\star \geq 20000$.}
-    \label{fig:convergence}
-\end{figure}
-
-\subsection{Eccentricity scatter in simulations}
-
-The SMBH binary eccentricity for both the $e_0=0.90$ (blue) and $e_0=0.99$ (orange) orbits are shown in \autoref{fig:eccentricities} as a function of shifted time $t' = t - t_\mathrm{bound}$, where $t_\mathrm{bound}$ is the time when the binary orbital energy $E$ becomes permanently negative.
-At the time $t_\mathrm{bound}$ the SMBH binary is not yet isolated due to the large quantity of intervening stellar mass, which causes the Keplerian definition of eccentricity to oscillate and not instantaneously match exactly the actual SMBH binary orbit.
-In \autoref{fig:eccentricities} we indicate the time of hard binary formation by circle markers; from this point the Keplerian eccentricity appropriately describes the binary orbit.
-The binary hardening radius $a_\mathrm{h}$ is defined as \citep[e.g.][]{merritt2006}:
-\begin{equation}\label{eq:ahard}
-    a_\mathrm{h} = \frac{q}{(1+q)^2} \frac{r_\mathrm{m}}{4} = \frac{r_\mathrm{m}}{16},
-\end{equation}
-where $r_\mathrm{m}=r(m<2M_{\bullet,1})$ is the influence radius and $q$ is the mass ratio between the SMBHs (for our simulations $q=1.0$).
-Before a hard binary forms, both the $e_0=0.90$ and $e_0=0.99$ sets initially demonstrate an overall decrease in eccentricity (semitransparent lines in \autoref{fig:eccentricities}) until $t'\sim1\,\mathrm{Myr}$.
-It should be noted that the precise value of the eccentricity in this regime is not robust due to the SMBHs orbiting in a potential still largely influenced by the stellar background, however the overall trend is consistent with the binary orbital geometry.
-After $t'\sim1\,\mathrm{Myr}$, some of the $e_0=0.99$ realisations continue to have a decreasing eccentricity, and all of the $e_0=0.90$ realisations have an increasing eccentricity.
-We discuss the mechanism for this in \autoref{ssec:toy}.
-
-The $e_0=0.99$ runs show a wide variation in eccentricity, where $e$ spans almost the entire domain range $e=[0,1]$.
-As a test we also perform two sets of ten runs using the $10^6$ particle resolution set up, but reduce the SMBH mass to $5.0\times10^7\,\Msun$ and $1.0\times10^7\,\Msun$, and observe the same qualitative spread in eccentricity as in the $M_\bullet=10^8\,\Msun$ case.
-Even though the $e_0=0.99$ runs have a higher initial merger eccentricity than the $e_0=0.90$ runs, none of the SMBH binaries in the shown $e_0=0.99$ set obtain high enough eccentricities to undergo GW-induced coalescence during the $50\,\mathrm{Myr}$ timespan.
-The $e_0=0.90$ set demonstrates six SMBH-binary mergers within $50\,\mathrm{Myr}$ of forming a bound binary, seen as a rapid orbit circularisation in \autoref{fig:eccentricities}, which is captured self-consistently using \ketju{}.
-
-To characterise the scatter in eccentricity, we determine the mean eccentricity $e_\mathrm{h}$ over five orbital periods centred on the orbit within which the SMBH binary has become hard (equation \eqref{eq:ahard}).
-We characterise the inter-simulation eccentricity scatter in the mean values of $e_\mathrm{h}$ with the standard deviation, denoted as $\sigma_e$.
-
-We show the dependence of $\sigma_e$ on mass resolution for the $e_0=0.90$ orbit in \autoref{fig:convergence} with blue circle markers, and for the $e_0=0.99$ orbit using orange square markers.
-
-For the $e_0=0.90$ orbit, the convergence of $\sigma_e$ scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, where $N_{\star,\mathrm{tot}}$ is the total number of stellar particles in the merger.
-The values of $\sigma_e$ we report are quantitatively similar to the values found by \citet{nasim2020} for the same system, as can be seen by the teal triangles in \autoref{fig:convergence}.
-The variation in eccentricity does not show the same scaling for the $e_0=0.99$ orbit as the $e_0=0.90$ orbit. 
-For mass resolutions $M_\bullet/m_\star \leq 10000$, the value of $\sigma_e$ is almost constant, before significantly dropping at higher mass resolutions.
-
-\begin{figure*}
-    \centering
-    \includegraphics{orbit}
-    \caption{Orbits from $t=0\,\mathrm{Myr}$ to a time shortly after $t_\mathrm{bound}$ of a single SMBH in two representative realisations of the $e_0=0.90$ mergers (left) and the $e_0=0.99$ mergers (right), which by symmetry of the equal mass system is a reflection of the second BH orbit about $x=-z$.
-    The line gradient shows the shifted time $t'=t-t_\mathrm{bound}$, however the colouring for $|t'|>2\,\mathrm{Myr}$ is constant for visual clarity. 
-    In each inset panel, $\theta_\mathrm{defl}$ is the deflection angle at the pericentre between the arrows.
-    The circle marker indicates the SMBH position when a bound binary is formed.}
-    \label{fig:orbit}
-\end{figure*}
-
-\subsection{Binary binding process and the scatter in eccentricity}\label{ssec:binding}
-
-To understand the observed scatter in eccentricity, we investigate the binary binding process in each simulation. 
-Before the SMBHs are bound, the primary influence on the motion is from the galactic merger potential; two realisations for both merger orbits are shown in \autoref{fig:orbit}.
-The initial stages of the orbit show indiscernible variation between realisations, however after a particularly strong interaction between the SMBHs during a pericentre passage, the orbits deviate between realisations (inset panels, \autoref{fig:orbit}), and evolve to different binary eccentricities.
-
-To quantify the strength of the hard-scattering process that randomises the SMBH orbit, we measure the effective two-body deflection angle, defined as:
-\begin{equation}\label{eq:theta}
-    \theta_\mathrm{defl} = 2 \arctan \left(\frac{GM}{L\sqrt{2E}} \right) = 2\arctan\left( \frac{b_{90}}{b} \right)
-\end{equation}
-where $M=M_{\bullet,1}+M_{\bullet,2}$, $L$ is the magnitude of the SMBH 
-system angular momentum vector, $E$ the orbital energy at the time of the pericentre passage, $b$ the effective impact parameter, and $b_{90}$ the $90\degr$ deflection radius \citep{binney2008}.
-
-We observe clear evidence for a relationship between the deflection angle $\theta_\mathrm{defl}$ and the resulting eccentricity at the time the SMBH binary becomes hard, as shown with the data points in \autoref{fig:theta_e}.
-We also observe a similar relationship in our two test simulation sets with the reduced SMBH masses.
-
-\begin{figure*}
-    \centering
-    \includegraphics{theta_e_sim_and_model}
-    \caption{
-    Deflection angle $\theta_\mathrm{defl}$ and resulting hard binary eccentricity
-    in the $e_0=0.90$ (left) and $e_0=0.99$ (right) simulations,
-    compared to fitted model curves.
-    The solid lines show the results from the analytic model with $e_\mathrm{s}=0.91$,
-    while the semitransparent lines show the results for  $e_\mathrm{s}=0.90$ and $e_\mathrm{s}=0.92$.
-    The $e_0=0.90$ model curves have been shifted left by $14\degr$
-    to better match the simulation data, with the shift indicated by an arrow.
-    The marginal histograms show kernel density estimates of the simulation data.
-    The inset panel for the $e_0=0.90$ data shows a zoom-in with $\theta_\mathrm{defl}=[52\degr, 73\degr]$ and $e_\mathrm{h}=[0.96, 1.00]$.
+\drafting{
+    Things to show:
+    \begin{enumerate}
+        \item DAG for hierarchical model
+        \item Density profiles 
+        \item Scatter plot $v_k\mathrm{--}r_b$ with best fit function
+        \item Relation $v_k\mathrm{--}r_b$ model comparison (elpd?)
+        \item MC-sampled distribution of $r_b$
+        \item $h_4$ maps
+    \end{enumerate}
 }
-    \label{fig:theta_e}
-\end{figure*}
-
-\subsection{Reproducing the eccentricity behaviour  with a simple model}\label{ssec:toy}
-
-The dependency of the binary eccentricity on the 
-deflection angle $\theta_\mathrm{defl}$ during the binary formation
-can be reproduced using a simple analytic model,
-which includes only the essential components relevant to this process.
-Taking the orbit to lie in the $xy$-plane,
-the relative motion of the two equal-mass SMBHs in this model follows the 
-equation of motion
-\begin{equation} \label{eq:toy_model_eom}
-\ddot{\vb{x}} = -\frac{2 G M_\bullet}{|\vb{x}|^3} \vb{x} + \vb{a}_\mathrm{bg} + \vb{a}_\mathrm{DF},
-\end{equation}
-where $\vb{x}=(x,y)$ is the separation vector of the SMBHs,
-$M_\bullet = 10^8\,\Msun$ is the mass of a single SMBH,
-$\vb{a}_\mathrm{bg}$ the acceleration due to the asymmetric background potential,
-and $\vb{a}_\mathrm{DF}$ is the acceleration due to DF.
-
-The stellar background is modelled as a constant density spheroidal potential
-\citep[e.g.][]{binney2008}:
-\begin{equation}
-\Phi_\mathrm{bg}(\vb{x}) = \pi G \rho (A_x x^2 + A_y y^2),
-\end{equation}
-with the acceleration given by 
-\begin{equation}
-\vb{a}_\mathrm{bg}(\vb{x}) = - \nabla \Phi_\mathrm{bg}(\vb{x}).
-\end{equation}
-The $A$ coefficients are related to the eccentricity $e_\mathrm{s}$ of the spheroid as
-\begin{align}
-A_x &= 2 \left(\frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2}\right) 
-    \left[\frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right) - 1 \right]\\
-A_y &= \frac{1-e_\mathrm{s}^2}{e_\mathrm{s}^2} 
-        \left[ \frac{1}{1-e_\mathrm{s}^2} 
-        - \frac{1}{2 e_\mathrm{s}} \ln\left(\frac{1+e_\mathrm{s}}{1-e_\mathrm{s}}\right)\right],
-\end{align}
-with the long axis aligned along the $x$-axis.
-The stellar density is set to $\rho = 300 \,\Msun\,\mathrm{pc}^{-3}$,
-which approximately matches the values seen in the simulations when the SMBH binary is becoming bound.
-
-\begin{figure*}
-    \centering
-    \includegraphics{sample_model_orbits}
-    \caption{
-    Top:
-    Sample orbits computed with the analytic model for different background potential eccentricities $e_\mathrm{s}$,
-    with otherwise identical initial conditions.
-    The dashed lines show the isopotential contours of the model.
-    Only the early part of the orbital evolution is shown for clarity.
-    Bottom:
-    Orbital eccentricity $e$ of the model orbits (note the non-linear scale).
-    The vertical line marks the end of the period shown in the top panels. 
-    }
-    \label{fig:model_orbits}
-\end{figure*}
-
-The DF acting on a single BH is modelled using the \citet{chandrasekhar1943} formula assuming
-a Maxwellian distribution with a constant velocity dispersion $\sigma_\star$ \citep[e.g.][]{binney2008}:
-\begin{equation}\label{eq:chandra_df}
-\vb{a}_{1,\mathrm{DF}} = -\frac{4 \pi G^2 M_\bullet \rho \ln{\Lambda}}{|\vb{v}_{1}|^3}
-                         \left(\operatorname{erf}(X) - \frac{2X}{\sqrt{\pi}} \exp(-X^2)\right)\vb{v}_{1},
-\end{equation}
-where $\vb{v}_1 = \dot{\vb{x}}/2$ is the velocity of a single BH,
-and $X = |\vb{v}_1|/(\sqrt{2}\sigma_\star)$.
-We set ${\sigma_\star=200\,\mathrm{km\,s^{-1}}}$ based on the value measured from the simulations.
-The value of the Coulomb logarithm is expected to be in the range $\ln{\Lambda}\sim 4\textnormal{--}5$
-based on the size of the stellar system,
-and we find that $\ln{\Lambda} = 4.7$ gives results that agree well with the simulations.
-
-The effect of DF weakens as the SMBH binary orbit shrinks.
-To account for this, when the binary is bound 
-we multiply the acceleration given by equation~\eqref{eq:chandra_df}
-with the smooth cut-off function
-\begin{equation}
-f(a) = \frac{1}{1 + \exp[(a_\mathrm{c}-a)/d_\mathrm{c}]}.
-\end{equation}
-Here $a$ is the semimajor axis of a bound binary,
-and the cut-off scales are $a_\mathrm{c}=2 a_\mathrm{h}$
-and $d_\mathrm{c} = 0.5 a_\mathrm{h}$,
-with $a_\mathrm{h}$ defined by equation~\eqref{eq:ahard}.
-The total DF term is then
-\begin{equation}
-\vb{a}_{\mathrm{DF}} = 2 f(a) \vb{a}_{1,\mathrm{DF}}.
-\end{equation}
-
-To mimic the nearly linearly plunging orbits before the scattering event seen in \autoref{fig:orbit},
-we specify the initial conditions as 
-\begin{align}
-\vb{x}_0 &= (25\,\mathrm{pc}, b)\\
-\dot{\vb{x}}_0 &= (-v_0, 0).
-\end{align}
-In order to match the typical SMBH relative velocity seen at this separation in
-the simulations we set $v_0 = 450\,\mathrm{km\,s^{-1}}$ for the $e_0=0.90$ case, and
- $v_0 = 560\,\mathrm{km\,s^{-1}}$ for the the $e_0=0.99$ case.
-
-The spheroid eccentricity parameter $e_\mathrm{s}$ is not easily measured from the simulations,
-due to the presence of stellar components that remain tightly bound to the SMBHs and obscure the background
-potential relevant for the dynamics,
-in combination with a constantly evolving potential.
-However, only sufficiently large values of $e_\mathrm{s}$ allow for low binary
-eccentricities to be produced,
-as is shown by \autoref{fig:model_orbits}.
-Performing calculations with different values of $e_\mathrm{s}$,
-we found a good fit to the simulation results for $e_\mathrm{s}\approx 0.9$.
-
-We compute the resulting eccentricities for a range of impact parameters up to $b=20\,\mathrm{pc}$
-for the two initial velocities,
-by solving the equation of motion \eqref{eq:toy_model_eom} until the binary has become hard using
-the error controlled 8th order Runge--Kutta method DOP853 included in the SciPy library \citep{virtanen2020}.
-For reference, the $90\degr$ deflection radii are $b_{90}\approx 5.8\,\mathrm{pc}$ and
-$b_{90}\approx 3.3\,\mathrm{pc}$ for the $e_0=0.90$ and $e_0=0.99$ cases, respectively.
-
-The resulting model curves are shown together with the simulation data in \autoref{fig:theta_e}.
-To correctly match the data in the $e_0=0.90$ case, the model curve has been shifted to the left by $14\degr$.
-This shift is required likely due to the rotation of the stellar component tilting the background potential
-relative to the SMBH trajectories in the simulations, with the effect being only evident in the $e_0=0.90$ case
-due to the less radial merger orbit.
-In the $e_0=0.99$ case we can also see that the simulation data has significant scatter around
-the model curve in the $\theta_\mathrm{defl} \gtrsim 120\degr$ region,
-although the simple model does appear to capture the mean behaviour  relatively well, apart from the very largest values of
-$\theta_\mathrm{defl}$.
-The behaviour of the model in this region is also sensitive to the exact values of $e_\mathrm{s}$
-and $\ln{\Lambda}$,
-which might also be related to the large scatter seen in the simulation data.
-
-However, in general the behaviour  of the simulation data is captured well,
-with the model correctly producing the two main eccentricity minima,
-as well as the $e\approx 1$ region between them.
-The predicted minima occur for trajectories where the torque from the background potential together with the DF
-causes the SMBHs to loop around into a nearly circular orbit, resulting in the  low eccentricities seen in some of the $e_0=0.99$ simulation data.
-Conversely, highly eccentric binaries are produced when the binary is trapped into nearly radial oscillations along 
-the main axes of the potential.
-These mechanisms are contrasted in \autoref{fig:model_orbits}.
 
@@
 
-As shown by both the simulations and the analytic model, the eccentricity of the hard SMBH binary in the studied merger configurations can span nearly the full possible range in the interval $e_\mathrm{h}=[0,1]$, depending on comparatively small, parsec-scale differences in the particular realisation of the galactic merger orbit.
-On the other hand, in minor mergers where the stellar background is less asymmetric when the SMBHs become bound,
-the scatter in eccentricity is likely to be relatively low even in the case of a radial merger orbit, since the system is less sensitive to the exact value of the deflection angle,
-as seen in the $e_\mathrm{s}=0.2$ case in \autoref{fig:model_orbits}.
-
-Merger orbits with a lower initial eccentricity $e_0$ than the values in this work can be expected to show less scatter in the binary eccentricity due to the lack of a hard scattering event that is sensitive to slight perturbations in the merger orbit, which was also seen in the study by \citet{gualandris2022}.
-However, major mergers are expected to occur on highly radial orbits based on cosmological simulations \citep[e.g.][]{khochfar2006}.
-In addition, the DF from the dark matter halo would in general drive even lower eccentricity merger orbits to highly radial final plunges when the initial separation of the galaxies is large enough.
-
-The simulation setup used in this work is idealised, and covers only a small part of the parameter space of merger configurations.
-However, similar behaviour  can be expected to occur also in more realistic models,
-since the relevant dynamics occurs within a few hundred parsecs from the centre of the merged galaxy,
-where we do not expect any significant effects from dark matter or the outer parts of the galaxy, which are not included in our present simulations.
-Preliminary results using steeper $(\gamma \gtrsim 1)$, more observationally-motivated galaxy density profiles indicate that similar behaviour occurs also in more realistic merger systems.
-
-The variation of the deflection angle is caused by the random variation of the impact parameter $b$ through equation~\eqref{eq:theta}.
-In numerical simulations with a discretised phase space, the primary cause of random variation in $b$ is the Brownian motion of the SMBHs before they form a bound binary \citep{merritt2001,bortolas2016}.
-This effect scales as $1/\sqrt{N_{\star,\mathrm{tot}}}$, which can explain the observed scaling of $\sigma_e$ \citep{nasim2020} when the system falls into the region of parameter space where the relation between $e$ and $b$
-is approximately linear.
-The deviation from this scaling seen in \autoref{fig:convergence} can then be explained by the fact that the eccentricity is not globally a linear function of the impact parameter. 
-The non-linear mapping also explains why GW-induced mergers were present within a short ($<50\,\mathrm{Myr}$) timeframe for the $e_0=0.90$ mergers and not the $e_0=0.99$ mergers in \autoref{fig:eccentricities}, and highlights the difficulty in predicting the binary eccentricity $e_\mathrm{h}$ from the initial merger orbit.
-
-While the scatter in the impact parameter between different numerical realisations of the same merger is due to relatively low number of particles compared to real galaxies,
-we expect that uncertainty of a similar magnitude is also present in real systems due to various mechanisms, such as perturbations of the SMBH motion from gas and substructure in their stellar environment, or larger-scale perturbations of the galactic merger orbit.
-In addition, the SMBHs are also not necesssarily located exactly at the centre of mass of the galaxy.
-For example, the work by \citet{batcheldor2010} has suggested that the SMBH in M87 may be displaced from the galactic photocentre by up to $\sim7\,\mathrm{pc}$ due to jet acceleration or gravitational recoil kicks.
-Thus, it is not possible to give an exact prediction for the SMBH binary eccentricity produced by a given merger,
-and instead we must focus on predicting the distribution of eccentricities that can be produced by a given merger configuration.
-
-Predicting the eccentricity distribution requires the knowledge of the distribution of impact parameters that can result from essentially identical mergers
-as well as the mapping between the impact parameter, or deflection angle, and the final eccentricity.
-Since the relation between the impact parameter and the eccentricity does not appear to significantly depend on the resolution,
-simulations at moderate mass resolutions of $M_\bullet/m_\star \sim 10^3$ can be used
-to map out the relation by performing a large number of merger simulations,
-possibly augmented by more sophisticated versions of the analytic model presented here.
-The distribution of impact parameters in real systems is however a much more difficult problem to tackle,
-since in simulations numerical resolution effects are likely dominant.
-
-Even without such extensive studies on the full distribution of SMBH binary eccentricities, the present results suggest that the distribution is likely to be broad and contain a significant number of binaries at high eccentricities.
-The possibility of a large fraction of high binary eccentricities has implications in particular for studies of the expected gravitational wave background (GWB) signal measured by PTAs, as in particular binaries with $e_\mathrm{h}\gtrsim 0.9$ can retain significant eccentricities while their GW emission is within the observable frequency band.
-This could be observable in the shape of the GWB spectrum \citep{kelley2017}, and would also affect the likelihood of detecting individual SMBH binaries with PTAs \citep{kelley2018}.
-It is interesting to note that \citet{bi2023} obtained a distribution of eccentricities consistent with the expectation of a significant fraction of SMBH binaries at high eccentricities based on analysis of the recent PTA observations.
-
-In the present work, the influences of rotation and SMBH orbital flips are not considered. 
-A counter-rotating stellar background with respect to the SMBH binary angular momentum has a tendency to increase the binary eccentricity, whereas a corotating stellar background tends to circularise the binary \citep{sesana2011, holley-bockelmann2015}.
-The SMBH angular momentum vector can however undergo reversals as a result of gravitational torques following pericentre passages of the binary \citep[][termed an orbital flip]{rantala2019} an example of which is shown in the left inset of of the left panel of \autoref{fig:orbit}.
-The chaotic reversals of the binary angular momentum act to randomise the relative rotation of the stellar background, leading in general to complex eccentricity evolution \citep{nasim2021}.
-A result of a sustained co- or counter-rotating stellar background could be an evolution of the binary eccentricity after $e_\mathrm{h}$ to a bimodal distribution for a given merger configuration, if the stellar scattering interactions were strong enough.
-Whilst these mechanisms are unlikely to alter much our findings concerning the hard binary eccentricity, they may obfuscate predictions for SMBH binary eccentricity prior to the merger, which is critical for GW detection missions.
-
-In contrast to the gas-free and fairly low density systems studied here, in higher density or gas-rich systems the binary eccentricity can also evolve significantly after the binary has become hard but before it has entered the GW emission dominated regime.
-In particular, interactions with a circumbinary accretion disc can lead to rapid evolution of the eccentricity if the accretion rate is high.
-Studies of prograde coplanar circumbinary discs have found that at almost all eccentricities the binaries evolve towards an eccentricity of $e\approx 0.5$ \citep{zrake2021,dorazio2021}, whereas retrograde discs lead to the eccentricity always increasing \citep{tiede2023}.
-However, these studies are generally limited to eccentricities below $e<0.9$.
-Our results indicate the need for extending such studies to higher binary eccentricities, as well as to more general configurations such as polar discs which become increasingly likely at high eccentricities \citep[e.g.][]{martin2017}.
-
 % CONCLUSION
@@
 \label{sec:conclusions}
-In conclusion, we find that the variation in the hard binary eccentricity $e_\mathrm{h}$ for a given system configuration is tightly correlated with the deflection angle $\theta_\mathrm{defl}$, and thus the impact parameter $b$.
-By using a simple, resolution-free analytic model of the SMBH scattering process,
-we have demonstrated that uncertainty in SMBH binary eccentricity is not caused solely by discretisation effects in galaxy merger simulations,
-but is rather due to the physical sensitivity of the system to small changes in the merger orbit,
-which can be caused by physical mechanisms in addition to numerical discretisation effects.
-The results presented here justify extending the investigation to more realistic galaxy merger scenarios,
-in order to quantify the expected range of hard binary eccentricities, which is critical for predictions for current and future gravitational wave observation missions.
 
@@
 \section*{acknowledgments}
-We thank the anonymous referee for their comments which contributed to improving this work.
-We also thank Shihong Liao, Dimitrios Irodotou, and Francesco Paolo Rizzuto for interesting discussions.
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
-A.R., M.M. and P.H.J. acknowledge the support
+A.R., acknowledge the support
 by the European Research Council via ERC Consolidator Grant KETJU (no. 818930) and the support of the Academy of Finland grant 339127.
-TN acknowledges support from the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany’s Excellence Strategy - EXC-2094 - 390783311 from the DFG Cluster of Excellence ``ORIGINS''.
 
@@
 We list here the roles and contributions of the authors according to the Contributor Roles Taxonomy (\href{https://credit.niso.org}{CRediT}). 
-\textbf{AR}: Conceptualisation, Investigation, Formal analysis, Data curation, Writing -- original draft. 
-\textbf{MM}: Conceptualisation, Formal analysis, Writing -- original draft. 
-\textbf{PHJ}: Supervision, Writing -- review \& editing.
-\textbf{TN}:  Writing -- review \& editing.
+\textbf{AR}:
 

### 2023-09-22 papers/paper-corekick/paper-corekick.tex
Add text about simulation details

@@
 
-\newcommand{\Msun}{\ensuremath{M_{\sun}}}
+\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}
+\newcommand{\kmps}{\ensuremath{\mathrm{km}\,\mathrm{s}^{-1}}}
 
@@
 \drafting{Rantala 2018 ICs}
+We model the merger of two gas-poor elliptical galaxies using collisionless merger simulations.
+Our galaxy initial conditions are chosen to match the model IC-3 presented in \citet{rantala2019}. 
+The galaxy is thus represented as a multicomponent sphere, including a stellar component of total mass $M_\star\sim 1.38\times10^{11}\,\Msun$ embedded within a DM \drafting{ensure abbreviation stated} component of total mass $M_\mathrm{DM}=2.5\times10^{13}\,\Msun$, and at the centre a SMBH with mass $M_\bullet=2.93\times10^{9}\,\Msun$.
+The stellar and DM components each follow a \citet{dehnen1993} profile with shape parameter $\gamma=1$, and scale radius $a_\star=3.9\,\mathrm{kpc}$ and $a_\mathrm{DM}=245\,\mathrm{kpc}$, respectively\footnote{A Dehnen profile with $\gamma=1$ is commonly referred to as a Hernquist profile.}.
+The density profile is given by:
 \begin{equation}\label{eq:dehnen}
@@
 \end{equation}
+The mass of a stellar particle is set to $m_\star=5\times10^4\,\Msun$, and the mass of a DM particle is set to $m_\mathrm{DM}=5\times10^6\,\Msun$.
 
-\drafting{We then construct isolated merger initial conditions}. 
-Interactions between stellar particles are softened with a softening length of \drafting{$\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.}
+We create six independently Monte Carlo sampled galaxy ICs, and merge each combination of galaxies on a radial orbit.
+The merger orbit has an initial separation of $D=30\,\mathrm{kpc}$, a first pericentre distance of $r_\mathrm{peri}=2\,\mathrm{kpc}$, and an initial eccentricity of $e_0=0.97$.
+We select a merger combination that results in a rapid coalescence of the BH binary, noting that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
+We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $v_\mathrm{kick}$ prescribed, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+The upper limit on $v_\mathrm{kick}$ is chosen to match the escape velocity of the centre of the merger remnant.
+We additionally run one simulation with a recoil kick above the escape velocity, with $v_\mathrm{kick}=2000\,\kmps$.
+Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
+Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
+
+\drafting{Atte: mergers of unequal mass ratio?}
 
@@
         \item Density profiles 
-        \item Scatter plot $v_k\mathrm{--}r_b$ with best fit function
-        \item Relation $v_k\mathrm{--}r_b$ model comparison (elpd?)
+        \item Scatter plot $v_k\mathrm{-}r_b$ with best fit function
+        \item Relation $v_k\mathrm{-}r_b$ model comparison (elpd?)
         \item MC-sampled distribution of $r_b$

### 2023-10-03 papers/paper-corekick/paper-corekick.tex
Place-holder figures for core paper

@@
 
-\drafting{Atte: mergers of unequal mass ratio?}
+\drafting{Atte's runs of mergers of unequal mass ratio?}
+
+\drafting{Kicks randomly orientated?}
 
@@
 }
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{dag}
+    \caption{Directed acyclic graph of Bayesian hierarchical factor model.}
+    \label{fig:dag}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{density}
+    \caption{Density profiles with errors for select representative runs.}
+    \label{fig:density}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{vk_rb}
+    \caption{Relation and fit between core radius and kick velocity.}
+    \label{fig:vkrb}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{rb_pdf}
+    \caption{Probability density function of core radius, where kick velocity is MC sampled from Zlochower relation, and pumped through our fitted model in \autoref{fig:vrkb}.}
+    \label{fig:rb_pdf}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{h4}
+    \caption{$h_4$ kinematic maps of pre-merger and select remnants, possible to observe signature of kick in here? (Thorsten believes yes...)}
+    \label{fig:h4}
+\end{figure*}
 

### 2024-01-26 papers/paper-corekick/paper-corekick.tex
Start putting figures into paper

@@
 %\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
-\newcommand{\ketju}{\textsc{Ketju}}
-\newcommand{\mstar}{\textsc{mstar}}
-\newcommand{\gadget}{\textsc{gadget-4}}
+\newcommand{\ketju}{\textsc{Ketju}}                           % code name 
+\newcommand{\mstar}{\textsc{mstar}}                           % code name
+\newcommand{\gadget}{\textsc{gadget-4}}                       % code name
 
-\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}
-\newcommand{\kmps}{\ensuremath{\mathrm{km}\,\mathrm{s}^{-1}}}
+\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}            % Solar Mass
+\newcommand{\kmps}{\ensuremath{\mathrm{km}\,\mathrm{s}^{-1}} }% km/s formatting
+\newcommand{\Reff}{\ensuremath{R_\mathrm{e}}}                 % effective radius
 
-\newcommand{\dd}[1]{\ensuremath{\mathrm{d}#1}}
-\newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}
+\newcommand{\dd}[1]{\ensuremath{\mathrm{d}#1}}                % integral dx
+\newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}     % derivative
 \newcommand{\vb}[1]{\ensuremath{\bm{#1}}}                     %vectors
@@
 
+\drafting{
+    Main points to hammer home:
+    \begin{enumerate}
+        \item Parameters associated with the core (most importantly the core size) have a dependence on the kick velocity, whereas large-scale parameters (e.g. effective radius) do not.
+        \item Increasing kick velocity induces a decrease in the overall $\langle h_4 \rangle$ profile
+    \end{enumerate}
+}
+
 \begin{figure*}
@@
     \centering
-    \includegraphics[width=\textwidth]{vk_rb}
-    \caption{Relation and fit between core radius and kick velocity.}
+    \includegraphics[width=\textwidth]{rb-kick}
+    \caption{
+        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{b,0}$, as a function of kick velocity $v_\mathrm{kick}$.
+        The core size distributions are shown as violin plots, with the median core size indicated by a white circle, and the interquartile range by the box. 
+        Larger kick velocities are correlated with larger core sizes.
+Additionally, a greater spread in the distribution of core sizes over different viewing projections of the merger remnant is associated with larger kick velocities.}
     \label{fig:vkrb}
@@
     \centering
+    \includegraphics[width=\textwidth]{IFU_v0000}
+    \caption{
+        Mock integral field unit kinematic maps of the $v_\mathrm{kick}=0\,\kmps$ remnant, plotted out to $0.5\,\Reff$. 
+        Note the unordered rotation $V$, and the centrally-concentrated velocity dispersion $\sigma$.
+        The majority of the merger remnant within this aperture has $h_4>0$, with a small concentration of $h_4<0$ in an annulus about the central regions.
+    }
+    \label{fig:IFU0}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{IFU_v0960}
+    \caption{
+        Mock integral field unit kinematic maps of the $v_\mathrm{kick}=960\,\kmps$ remnant, again plotted out to $0.5\,\Reff$. 
+        Contrasting to the $v_\mathrm{kick}=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$ \drafting{check this}.
+    }
+    \label{fig:IFU960}
+\end{figure*}
+
+\begin{figure*}
+    \centering
     \includegraphics[width=\textwidth]{h4}
-    \caption{$h_4$ kinematic maps of pre-merger and select remnants, possible to observe signature of kick in here? (Thorsten believes yes...)}
+    \caption{
+        Plot of the flux-weighted value of $h_4$ as a function of mean bin distance $R$ (scaled to the half mass radius) for each merger remnant simulated.
+        The lines are coloured by the kick velocity. 
+        Immediately apparent is the decrease in $\langle h_4 \rangle$ with increasing $v_\mathrm{kick}$.
+        All merger remnants have a global minimum in $\langle h_4 \rangle$ at $R/\Reff\simeq0.2$.
+    }
     \label{fig:h4}
@@
 
-%\appendix
+\appendix
+\section{Core-Se\'rsic fit parameter estimates}
+\drafting{Let's put the violin plots for all core fit parameters in the appendix}.
 

### 2024-01-30 papers/paper-corekick/diagrams/dag.tex

\documentclass[border=10pt]{standalone}
\usepackage{newtxtext, newtxmath}
\usepackage{tikz}
\usetikzlibrary{bayesnet}

\begin{document}
\begin{tikzpicture}

    %%% define layer 1 %%%
    % density
    \node[obs] (dens) {$\log_{10}\Sigma$};
    % density distribution
    \factor[above=of dens] {dens_D} {right:$\mathcal{N}_\mathrm{likelihood}$} {} {};
    % calculated density
    \node[det, above=of dens_D] (dens_hat) {$\log_{10}\hat{\Sigma}$}; 
    % radius
    \node[obs, above=of dens_hat] (R) {$R$};
    % density error
    \node[latent, right=of dens_hat, yshift=-3em, xshift=-3em] (dens_err) {$\sigma_{\log_{10}\Sigma}$};

    % connect
    \factoredge{dens_hat, dens_err} {dens_D} {dens};
    \edge {R} {dens_hat};
    % add the plate
    \plate{level1} {(R)(dens_hat)(dens_err)(dens_D)(dens)} {$\forall\; R$};

    %%% define layer 2 %%%
    % break density
    \node[latent, left=of dens_hat, yshift=10em] (dens_b) {$\log_{10}\Sigma_\mathrm{b}$};
    % inner core slope
    \node[latent, below=of dens_b] (g) {$\gamma$};
    % sersic index
    \node[latent, below=of g] (n) {$n$};
    % transition index
    \node[latent, below=of n] (a) {$\alpha$};
    % core radius
    \node[latent, below=of a] (rb) {$r_\mathrm{b}$};
    % effective radius
    \node[latent, below=of rb] (Re) {$R_\mathrm{e}$};

    % connect
    \edge{dens_b, g, n, a, rb, Re} {dens_hat};
    % add the plate
    \plate{level2} {(dens_b)(g)(n)(a)(rb)(Re)(level1)} {$\forall\; \mathrm{projections}$};

    %%% define layer 3 %%%
    % break radius hypers
    \factor[left=of dens_b] {dens_b_D} {below:$\mathcal{N}$} {} {};
    \node[hyper, left=of dens_b_D, yshift=2em] (dens_b_mu) {$\mu_{\log_{10}\Sigma_\mathrm{b}}$};
    \node[hyper, below=of dens_b_mu, yshift=2em] (dens_b_err) {$\sigma_{\log_{10}\Sigma_\mathrm{b}}$};
    \factoredge{dens_b_mu, dens_b_err} {dens_b_D} {dens_b};

    % inner slope hypers
    \factor[left=of g, xshift=-1.5em] {g_D} {below:$\mathrm{Exp}_\mathrm{T}$} {} {};
    \node[hyper, left=of g_D] (g_lam) {$\lambda_\gamma$};
    \factoredge{g_lam} {g_D} {g};

    % sersic index hypers
    \factor[left=of n, xshift=-1.5em] {n_D} {below:$\mathcal{N}_\mathrm{T}$} {} {};
    \node[hyper, left=of n_D, yshift=1em] (n_mu) {$\mu_{n}$};
    \node[hyper, below=of n_mu, yshift=2em] (n_err) {$\sigma_{n}$};
    \factoredge{n_mu, n_err} {n_D} {n};

    % transition index hypers
    \factor[left=of a, xshift=-1.5em] {a_D} {below:$\mathrm{Ray}$} {} {};
    \node[hyper, left=of a_D] (a_sig) {$\sigma_\alpha$};
    \factoredge{a_sig} {a_D} {a};

    % core radius hypers
    \factor[left=of rb, xshift=-1.5em] {rb_D} {below:$\mathrm{Ray}_\mathrm{T}$} {} {};
    \node[hyper, left=of rb_D] (rb_sig) {$\sigma_{r_\mathrm{b}}$};
    \factoredge{rb_sig} {rb_D} {rb};

    % effective radius hypers
    \factor[left=of Re, xshift=-1.5em] {Re_D} {below:$\mathrm{Ray}_\mathrm{T}$} {} {};
    \node[hyper, left=of Re_D] (Re_sig) {$\sigma_{R_\mathrm{e}}$};
    \factoredge{Re_sig} {Re_D} {Re};

    % surface density error hypers
    \factor[above=of dens_err, yshift=4em] {dens_err_D} {right:$\mathcal{N}_\mathrm{T}$} {} {};
    \node[hyper, above=of dens_err_D, yshift=6em] (dens_err_err) {$\sigma_{\sigma_{\log_{10}\Sigma}}$};
    \node[hyper, left=of dens_err_err] (dens_err_mu) {$\mu_{\sigma_{\log_{10}\Sigma}}$};
    \factoredge{dens_err_mu, dens_err_err} {dens_err_D} {dens_err};

\end{tikzpicture}
\end{document}

### 2024-01-30 papers/paper-corekick/paper-corekick.tex

@@
 \newcommand{\Reff}{\ensuremath{R_\mathrm{e}}}                 % effective radius
+\newcommand{\rb}{\ensuremath{r_\mathrm{b}}}                   % core radius
+\newcommand{\vk}{\ensuremath{v_\mathrm{kick}}}                % kick velocity
 
@@
 \newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}     % derivative
-\newcommand{\vb}[1]{\ensuremath{\bm{#1}}}                     %vectors
+\newcommand{\vb}[1]{\ensuremath{\bm{#1}}}                     % vectors
 
@@
 We select a merger combination that results in a rapid coalescence of the BH binary, noting that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
-We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $v_\mathrm{kick}$ prescribed, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
-The upper limit on $v_\mathrm{kick}$ is chosen to match the escape velocity of the centre of the merger remnant.
-We additionally run one simulation with a recoil kick above the escape velocity, with $v_\mathrm{kick}=2000\,\kmps$.
+We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant.
+We additionally run one simulation with a recoil kick above the escape velocity, with $\vk=2000\,\kmps$.
 Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
+Those simulations where the maximum displacement of the kicked SMBH exceeds \drafting{$30\,\mathrm{kpc}$ check} are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
+In practice, this limits our sample of analysed simulations to those with $\vk\leq 960\kmps$, thus totalling seventeen simulations in our analysis.
 Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
@@
 
-\begin{figure*}
+\subsection{Density Profiles}
+We fit the six-parameter core-Se\'rsic profile to the projection of each simulated merger remnants:
+\begin{equation}\label{eq:cs}
+    \Sigma(R) = \Sigma' \left[1 + \left(\frac{\rb}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[-b \left(\frac{R^\alpha + \rb^\alpha}{\Reff^\alpha}\right)^{(1/\alpha n)}\right]
+\end{equation}
+where
+\begin{equation}
+    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[b\left(2^{1/a} \frac{\rb}{\Reff}\right)^{1/n}\right].
+\end{equation}
+Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius, $\alpha$ is the profile transition index, and $n$ is the Se\'rsic index.
+We refer to the collective vector of these parameters as $\vb{\theta}$.
+We view the simulated merger remnant from fifteen different angles, and use a Bayesian hierarchical model (HM) to fit the model parameters.
+As each of the projections are unordered and exchangeable (i.e., there is no distinguishing label for the projections), and each projection is of the same merger remnant, a HM naturally lends itself.
+The idea behind the HM is straightforward: for each projection, we infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible vectors $\vb{\theta}$.
+Hence, we are effectively inferring the global hyperparameters $\vb{\theta}^\mathrm{hyp}$ that describe the latent parameters $\vb{\theta}$ that we are interested in.
+This idea is demonstrated in the directed acyclic graph (DAG) for the model, shown in \autoref{fig:dag}.
+
+We use weakly-constrained priors for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these priors are not necessarily Gaussian (given that distributions of positive-constrained quantities, like radius, should not be described by a distribution with support on the full real space).
+The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
+We use the Hamiltonian Monte Carlo code \textsc{Stan} with four chains each of 2000 posterior sample draws (excluding 1000 burn-in draws) to fit the model parameters, ensuring that:
+\begin{enumerate}
+    \item the number of diverging draws is less than \drafting{5\%},
+    \item the effective draws per transition exceeds 0.001,
+    \item the sampler transitions of HMC potential energy is in excess of 0.3, and
+    \item the value $\hat{R}$ (comparison of between chain and within chain estimates) is less than 1.05.
+\end{enumerate}
+Example corner plots of the hyperparameter and latent parameters can be found in \autoref{sec:app_fit}.
+\drafting{Do we need specifics of MC parameters and exit conditions?}
+
+\begin{table}
+    \caption{Hyperparameter distributions for core-Se\'rsic model.}
+    \label{tab:hyper}
+    \begin{tabular}{llc}
+        \hline
+        Hyperparameter & Distribution & Truncation \\
+        \hline
+        $\mu_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(10, 1)$ & None \\
+        $\sigma_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(0, 0.05)$ & $x>0$ \\
+        $\lambda_\gamma$ & $\mathrm{Exp}(10)$ & None \\
+        $\mu_n$ & $\mathcal{N}(4,2)$ & $0 < x \leq 15$ \\
+        $\sigma_n$ & $\mathcal{N}(0,2)$ & $x>0$ \\
+        $\sigma_\alpha$ & $\mathcal{N}(0, 20)$ & $x>0$ \\
+        $\sigma_{r_\mathrm{b}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
+        $\sigma_{R_\mathrm{e}}$ & $\mathcal{N}(0, 20)$ & $x>0$ \\
+        $\mu_{\sigma_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
+        $\sigma_{\sigma_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
+        \hline
+    \end{tabular}
+\end{table}
+
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{dag}
-    \caption{Directed acyclic graph of Bayesian hierarchical factor model.}
+    \includegraphics[width=0.4\textwidth]{dag}
+    \caption{Directed acyclic graph of the core-Se\'rsic model of projected mass density within the Bayesian hierarchical framework.
+Single-line circle nodes represent fit parameters, double-line circles represent measured quantities (the data), and diamond nodes represent deterministic quantities.
+The particular distribution connecting nodes is written below the corresponding black square, with a subscript `T' indicating a truncated distribution, and the subscript `likelihood' indicating the likelihood function.
+The $R$ box indicates variables fit for each radial bin, and the projection box indicates variables specific to each projection realisation.
+Note that we additionally use $\hat{\Sigma}$ to distinguish the calculated value of surface density from the measured value $\Sigma$.}
     \label{fig:dag}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{density}
+    \includegraphics[width=0.4\textwidth]{density}
     \caption{Density profiles with errors for select representative runs.}
     \label{fig:density}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{rb-kick}
+    \includegraphics[width=0.4\textwidth]{rb-kick}
     \caption{
-        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{b,0}$, as a function of kick velocity $v_\mathrm{kick}$.
-        The core size distributions are shown as violin plots, with the median core size indicated by a white circle, and the interquartile range by the box. 
+        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{b,0}$, as a function of kick velocity $\vk$.
+        The core size distributions are shown as violin plots, with the median core size indicated by the central mark.
         Larger kick velocities are correlated with larger core sizes.
@@
     \label{fig:vkrb}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{rb_pdf}
-    \caption{Probability density function of core radius, where kick velocity is MC sampled from Zlochower relation, and pumped through our fitted model in \autoref{fig:vrkb}.}
+    \includegraphics[width=0.4\textwidth]{rb_pdf}
+    \caption{Probability density function of core radius, where kick velocity is MC sampled from Zlochower relation, and pumped through our fitted model in \autoref{fig:vkrb}.}
     \label{fig:rb_pdf}
-\end{figure*}
+\end{figure}
+
+\subsection{Integral field unit kinematics}
+\drafting{We do some IFU stuff}
+We observe a trend of the Voronoi bins with $h_4<0$ extending to larger radii for remnants with larger recoil kicks than for those remnants with smaller recoil kicks.
+To quantify this recoil velocity dependence, we calculate a radially-varying mass-weighted $h_4$ parameter analogous to the observational spin parameter $\lambda$:
+\begin{equation}
+    \langle h_4 \rangle = \frac{\langle R h_4 \rangle}{\langle R \sqrt{V^2+\sigma^2}\rangle},
+\end{equation}
+where the $\langle\rangle$ brackets indicate a mass-weighted average (thus equivalent to a flux weighted average in the case of constant mass-to-light ratio).
 
@@
     \caption{
-        Mock integral field unit kinematic maps of the $v_\mathrm{kick}=0\,\kmps$ remnant, plotted out to $0.5\,\Reff$. 
+        Mock integral field unit kinematic maps of the $\vk=0\,\kmps$ remnant, plotted out to $0.5\,\Reff$. 
         Note the unordered rotation $V$, and the centrally-concentrated velocity dispersion $\sigma$.
@@
     \caption{
-        Mock integral field unit kinematic maps of the $v_\mathrm{kick}=960\,\kmps$ remnant, again plotted out to $0.5\,\Reff$. 
-        Contrasting to the $v_\mathrm{kick}=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$ \drafting{check this}.
+        Mock integral field unit kinematic maps of the $\vk=960\,\kmps$ remnant, again plotted out to $0.5\,\Reff$. 
+        Contrasting to the $\vk=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$ \drafting{check this}. 
     }
@@
 
-\begin{figure*}
+\begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{h4}
+    \includegraphics[width=0.4\textwidth]{h4}
     \caption{
@@
         The lines are coloured by the kick velocity. 
-        Immediately apparent is the decrease in $\langle h_4 \rangle$ with increasing $v_\mathrm{kick}$.
+        Immediately apparent is the decrease in $\langle h_4 \rangle$ with increasing $\vk$.
         All merger remnants have a global minimum in $\langle h_4 \rangle$ at $R/\Reff\simeq0.2$.
@@
     \label{fig:h4}
-\end{figure*}
+\end{figure}
 
@@
 \appendix
-\section{Core-Se\'rsic fit parameter estimates}
-\drafting{Let's put the violin plots for all core fit parameters in the appendix}.
+\section{Core-Se\'rsic fit parameter estimates}\label{sec:app_fit}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{all-kick}
+    \caption{
+        Violin plots of the six parameters in the core-Se\'sic model (from top left to bottom right) as a function of the kick velocity: the core radius normalised to the pre-merger core radius $\rb/r_{\mathrm{b},0}$, the density at the core radius $\log_{10}\Sigma_\mathrm{b}$, the core slope $\gamma$, the effective radius $\Reff$, the profile transition index $\alpha$, and the Se\'rsic index $n$.
+    }
+    \label{fig:csparams}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{corner_latent_960}
+    \caption{
+        Corner plot of the latent parameters in the core-Se\'rsic model for the $\vk=960\kmps$ instance.
+        Contours indicate, from yellow to dark blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs), with sample draws beyond the 99\% HDI shown as individual points. 
+        All distributions are unimodal with little cross-correlation between the latent variables.
+    }
+\end{figure*}
 

### 2024-01-31 papers/paper-corekick/paper-corekick.tex
Start writing about the IFU stuff

@@
 We select a merger combination that results in a rapid coalescence of the BH binary, noting that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
-We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.
+Additionally, the merger remnant displays almost no triaxiality, as seen in the $\vk=0\,\kmps$ case in \autoref{sec:app_triax}.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
 The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant.
@@
 Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
-Those simulations where the maximum displacement of the kicked SMBH exceeds \drafting{$30\,\mathrm{kpc}$ check} are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
-In practice, this limits our sample of analysed simulations to those with $\vk\leq 960\kmps$, thus totalling seventeen simulations in our analysis.
+Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
+In practice, this limits our sample of analysed simulations to those with $\vk\leq 900\,\kmps$, thus totalling sixteen simulations in our analysis.
 Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
@@
 The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
-We use the Hamiltonian Monte Carlo code \textsc{Stan} with four chains each of 2000 posterior sample draws (excluding 1000 burn-in draws) to fit the model parameters, ensuring that:
+We use the Hamiltonian Monte Carlo code \textsc{Stan} \citep{standevelopmentteam2018} with four chains each of 2000 posterior sample draws (excluding 1000 burn-in draws) to fit the model parameters, ensuring that:
 \begin{enumerate}
@@
 \subsection{Integral field unit kinematics}
-\drafting{We do some IFU stuff}
+We create mock integral field unit (IFU) observations using \drafting{ reference to Cappellari maybe?}. 
+Prior to performing the analysis, we reorientate each merger remnant so that the $z$-axis coincides with the minor axis of the reduced inertia tensor.
+
+\drafting{Describe voronoi tesselation}.
+For each Voronoi bin, we follow \citet{vandermarel1993} and decompose the line of sight (LOS) velocity into a series of Gauss-Hermite functions, described by the mean radial velocity $V$, the velocity dispersion $\sigma$, asymmetric deviations $h_3$, and symmetric deviations $h_4$.
+The line profile is thus described by:
+\begin{equation}
+    \mathfrak{L} = \frac{1}{\sqrt{2\pi}\sigma} e^{-w^2/2} \left\{ 1 + \sum_{j=3}^4 h_j H_j(w) \right\},
+\end{equation}
+where $w \equiv (v_\mathrm{LOS} - V)/\sigma$, and the Hermite polynomials $H_3$ and $H_4$ are defined:
+\begin{align}
+    H_3 &= \frac{1}{\sqrt{3}} (2w^3 - 3w) \nonumber \\
+    H_4 &= \frac{1}{\sqrt{24}} (4w^4 - 12w^2 + 3).
+\end{align}
+A recent comprehensive study of LOS velocity distribution fitting applied to elliptical galaxies can be found in \citet{mehrgan2023}.
+
 We observe a trend of the Voronoi bins with $h_4<0$ extending to larger radii for remnants with larger recoil kicks than for those remnants with smaller recoil kicks.
@@
 \begin{equation}
-    \langle h_4 \rangle = \frac{\langle R h_4 \rangle}{\langle R \sqrt{V^2+\sigma^2}\rangle},
+    \langle h_4 \rangle = \frac{\langle R h_4 \rangle}{\langle R \rangle},
 \end{equation}
 where the $\langle\rangle$ brackets indicate a mass-weighted average (thus equivalent to a flux weighted average in the case of constant mass-to-light ratio).
+We observe an overall shift to more negative values of $h_4$ for kick velocities up until $\vk\sim400\,\kmps$ \drafting{check this, currently just read off plot}, above which the radial $h_4$ profile is not sensitive to increasing kick velocity.
+Additionally, all $h_4$ profiles are radially decreasing, obtaining a local minimum at $R/\Reff\simeq 0.2$, and a modestly increasing by $R/\Reff\simeq 0.3$.
+
+\drafting{Why is this 
+\begin{enumerate}
+    \item decreasing with increasing kicks,
+    \item saturating beyond some point,
+    \item displaying a dip at 0.2 Re??
+\end{enumerate}
+}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{IFU_v0960}
+    \includegraphics[width=\textwidth]{IFU_v0900}
     \caption{
-        Mock integral field unit kinematic maps of the $\vk=960\,\kmps$ remnant, again plotted out to $0.5\,\Reff$. 
+        Mock integral field unit kinematic maps of the $\vk=900\,\kmps$ remnant, again plotted out to $0.5\,\Reff$. 
         Contrasting to the $\vk=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$ \drafting{check this}. 
     }
-    \label{fig:IFU960}
+    \label{fig:IFU900}
 \end{figure*}
@@
 \appendix
+\section{Triaxiality}\label{sec:app_triax}
+We present the triaxiality of all sixteen merger remnants in the analysis at the time when the SMBH has settled.
+We determine the ratios $b/a$ and $c/a$, where $a$, $b$, and $c$ are the eigenvalues of the reduced inertia tensor $I_\mathrm{red}$ and $c\leq b\leq a$. 
+The tensor $I_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} R_\mathrm{vir}$ to $10^{-1} R_\mathrm{vir}$: i.e., the binning of particles is not cumulative.
+The variation of the ratios $b/a$ and $c/a$ is plotted as a function of radius in \autoref{fig:triax}.
+All remnants display a high degree of symmetry, with the minimum value of $b/a\simeq 0.8$, and the minimum value of $c/a\simeq0.7$.
+With a higher kick velocity, there is a slight tendency to have a higher degree of symmetry at radii $R\lesssim 10\,\mathrm{kpc}$, and a slightly lesser degree of symmetry at radii beyond this distance, compared to lower kick velocity remnants.
+The difference is however minimal between all merger remnants.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{triaxiality}
+    \caption{
+        Triaxiality ratios $b/a$ and $c/a$ as a function of radius for all merger remnants in the analysis.
+        Lines are colour coded by the kick velocity.
+        Note that the $\vk=0\,\kmps$ case is equivalent to the pre-merger remnant. 
+    }
+    \label{fig:triax}
+\end{figure}
+
 \section{Core-Se\'rsic fit parameter estimates}\label{sec:app_fit}
@@
     \caption{
-        Corner plot of the latent parameters in the core-Se\'rsic model for the $\vk=960\kmps$ instance.
+        Corner plot of the latent parameters in the core-Se\'rsic model for the $\vk=960\,\kmps$ instance. \drafting{UPDATE TO 900}
         Contours indicate, from yellow to dark blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs), with sample draws beyond the 99\% HDI shown as individual points. 

### 2024-01-31 code/cm_functions/analysis/analysis_classes/StanModel.py
Minor plot updates

@@ -736,13 +736,13 @@ class _StanModel(ABC):
         num_vars = len(var_names)
         with az.rc_context({"plot.max_subplots":num_vars**2-np.sum(np.arange(num_vars))+1}):
             # first lay down the markers
-            ax = az.plot_pair(self._fit_for_az, group=group, var_names=var_names, kind="scatter", marginals=True, combine_dims=combine_dims, scatter_kwargs={"marker":".", "alpha":0.2}, figsize=figsize, labeller=labeller, textsize=rcParams["font.size"], backend_kwargs=backend_kwargs)
+            ax = az.plot_pair(self._fit_for_az, group=group, var_names=var_names, kind="scatter", marginals=True, combine_dims=combine_dims, scatter_kwargs={"marker":".", "alpha":0.1, "s":10}, figsize=figsize, labeller=labeller, textsize=rcParams["font.size"], backend_kwargs=backend_kwargs)
             # then add the KDE
             try:
-                az.plot_pair(self._fit_for_az, group=group, var_names=var_names, kind="kde", divergences=divergences, combine_dims=combine_dims, ax=ax, figsize=figsize, marginals=True, kde_kwargs={"contour_kwargs":{"linewidths":0.5}, "hdi_probs":levels, "contourf_kwargs":{"cmap":"cividis"}}, point_estimate_marker_kwargs={"marker":""}, labeller=labeller, textsize=rcParams["font.size"], backend_kwargs=backend_kwargs)
+                az.plot_pair(self._fit_for_az, group=group, var_names=var_names, kind="kde", divergences=divergences, combine_dims=combine_dims, ax=ax, figsize=figsize, marginals=True, kde_kwargs={"contour_kwargs":{"linewidths":0.5}, "hdi_probs":levels, "contourf_kwargs":{"cmap":"Blues"}}, point_estimate_marker_kwargs={"marker":""}, labeller=labeller, textsize=rcParams["font.size"], backend_kwargs=backend_kwargs)
             except ValueError:
                 _logger.error(f"HDI interval cannot be determined for corner plots! KDE levels will not correspond to a particular HDI, but follow matplotlib contour defaults")
[…]

### 2024-01-31 code/cm_functions/plotting/specific_plots.py

@@ -202,13 +202,13 @@ def voronoi_plot(vdat, figsize=(7,4.7), clims={}):
     # set up the figure
     fig, ax = plt.subplots(2,2, sharex="all", sharey="all", figsize=figsize)
     for i in range(2):
-        ax[1,i].set_xlabel("x/kpc")
-        ax[i,0].set_ylabel("y/kpc")
+        ax[1,i].set_xlabel(r"$x/\mathrm{kpc}$")
+        ax[i,0].set_ylabel(r"$y/\mathrm{kpc}$")
     ax = np.concatenate(ax).flatten()
     for i, (statkey, cmap, label) in enumerate(zip(
         ("V", "sigma", "h3", "h4"),
-        ("seismic", "plasma", "seismic", "seismic"),
-        (r"$V$ [km/s]", r"$\sigma$ [km/s]", r"$h_3$ [km/s]", r"$h_4$ [km/s]")
+        ("coolwarm", "plasma", "coolwarm", "coolwarm"),
+        (r"$V/\mathrm{km}\,\mathrm{s}^{-1}$", r"$\sigma/\mathrm{km}\,\mathrm{s}^{-1}$", r"$h_3/\mathrm{km}\,\mathrm{s}^{-1}$", r"$h_4/\mathrm{km}\,\mathrm{s}^{-1}$")
     )):
         # plot the statistic
         stat = vdat[f"img_{statkey}"]

### 2024-01-31 papers/paper-corekick/scripts/ifu.py

@@ -1,5 +1,6 @@
 import argparse
 import os.path
+import numpy as np
 import matplotlib.pyplot as plt
 from datetime import datetime
 import pygad
@@ -25,6 +26,12 @@ for k, v in snapfiles["snap_nums"].items():
 
 # dict to store radial h4 profiles
 h4_vals = {}
+
+# determine colour limits
+Vlim = -np.inf
+sigmalim = [np.inf, -np.inf]
+h3lim = -np.inf
+h4lim = -np.inf
 
 for k, v in snapshots.items():
     if float(k[1:]) > 900: break
@@ -58,7 +65,23 @@ for k, v in snapshots.items():
     del snap
     pygad.gc_full_collect()
 
-    ax = cmf.plotting.voronoi_plot(voronoi_stats)
+    # determine colour limits
+    _Vlim =  np.max(np.abs(voronoi_stats["img_V"]))
+    Vlim = _Vlim if _Vlim > Vlim else Vlim
+
+    _sigmalim = [np.min(voronoi_stats["img_sigma"]), np.max(voronoi_stats["img_sigma"])]
+    sigmalim[0] = _sigmalim[0] if _sigmalim[0] < sigmalim[0] else sigmalim[0]
+    sigmalim[1] = _sigmalim[1] if _sigmalim[1] > sigmalim[1] else sigmalim[1]
+
+    _h3lim =  np.max(np.abs(voronoi_stats["img_h3"]))
+    h3lim = _h3lim if _h3lim > h3lim else h3lim
+
+    _h4lim =  np.max(np.abs(voronoi_stats["img_h4"]))
+    h4lim = _h4lim if _h4lim > h4lim else h4lim
+
+    # have to set colour limits by hand
+    ax = cmf.plotting.voronoi_plot(voronoi_stats,
+                                   clims={"V":[25], "sigma":[155,351], "h3":[0.04], "h4":[0.05]})
     fig = ax[0].get_figure()
 
     SL.info(f"Total time: {datetime.now() - tstart}")
@@ -74,7 +97,7 @@ for k, v in snapshots.items():
 fig, ax = plt.subplots(1,1)
 get_kick_val = lambda k: float(k.lstrip("v"))
 kick_vels = [get_kick_val(k) for k in h4_vals.keys()]
[…]

### 2024-02-02 papers/paper-corekick/diagrams/dag.tex
More writing, add figures to paper

@@
     % density error
-    \node[latent, right=of dens_hat, yshift=-3em, xshift=-3em] (dens_err) {$\sigma_{\log_{10}\Sigma}$};
+    \node[latent, right=of dens_hat, yshift=-3em, xshift=-3em] (dens_err) {$\tau_{\log_{10}\Sigma}$};
 
@@
     \factor[above=of dens_err, yshift=4em] {dens_err_D} {right:$\mathcal{N}_\mathrm{T}$} {} {};
-    \node[hyper, above=of dens_err_D, yshift=6em] (dens_err_err) {$\sigma_{\sigma_{\log_{10}\Sigma}}$};
-    \node[hyper, left=of dens_err_err] (dens_err_mu) {$\mu_{\sigma_{\log_{10}\Sigma}}$};
+    \node[hyper, above=of dens_err_D, yshift=6em] (dens_err_err) {$\sigma_{\tau_{\log_{10}\Sigma}}$};
+    \node[hyper, left=of dens_err_err] (dens_err_mu) {$\mu_{\tau_{\log_{10}\Sigma}}$};
     \factoredge{dens_err_mu, dens_err_err} {dens_err_D} {dens_err};

### 2024-02-02 papers/paper-corekick/paper-corekick.tex
Minor plot updates
More writing, add figures to paper

@@
 
-\subsection{Density Profiles}
-We fit the six-parameter core-Se\'rsic profile to the projection of each simulated merger remnants:
+\subsection{Infall time}
+\drafting{
+    \textbf{Atte's section}. Would be good to have:
+    \begin{enumerate}
+        \item Plot of infall time (i.e., time for stability criterion to be reached) as a function of kick velocity
+        \item Number of oscillations before settling?
+        \item Plot of maximum displacement of kicked BH as a function of kick velocity. How does this compare to core size? Which simulations have BH kicked beyond the core?
+    \end{enumerate}
+}
+
+\subsection{Density profiles}
+We fit the six-parameter core-S\'ersic profile to the projection of each simulated merger remnants:
 \begin{equation}\label{eq:cs}
@@
 \end{equation}
-Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius, $\alpha$ is the profile transition index, and $n$ is the Se\'rsic index.
+Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius, $\alpha$ is the profile transition index, and $n$ is the S\'ersic index.
 We refer to the collective vector of these parameters as $\vb{\theta}$.
@@
 The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
-We use the Hamiltonian Monte Carlo code \textsc{Stan} \citep{standevelopmentteam2018} with four chains each of 2000 posterior sample draws (excluding 1000 burn-in draws) to fit the model parameters, ensuring that:
+We use the Hamiltonian Monte Carlo code \textsc{Stan} \citep{standevelopmentteam2018} with four chains each of 2000 posterior sample draws (excluding an additional 2000 burn-in draws) to fit the model parameters, ensuring that:
 \begin{enumerate}
-    \item the number of diverging draws is less than \drafting{5\%},
+    \item the number of diverging draws is less than 5\%,
     \item the effective draws per transition exceeds 0.001,
     \item the sampler transitions of HMC potential energy is in excess of 0.3, and
-    \item the value $\hat{R}$ (comparison of between chain and within chain estimates) is less than 1.05.
+    \item the value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.
 \end{enumerate}
-Example corner plots of the hyperparameter and latent parameters can be found in \autoref{sec:app_fit}.
+Example corner plots of the latent parameters can be found in \autoref{sec:app_fit}.
 \drafting{Do we need specifics of MC parameters and exit conditions?}
 
+In agreement with previous studies \citep{nasim2021b}, we find an increase in the core radius $\rb$ with kick velocity.
+\drafting{Need to check if others discuss this:} We find that the core radius saturates to a value of $\sim2$ times that of the merger remnant with no kick velocity by $\vk = 600\,\kmps$, corresponding to half of the escape velocity of the merger remnant core.
+As shown in \autoref{fig:csparams}, the density at the core radius $\Sigma_\mathrm{b}$ decreases monotonically with kick velocity, consistent with the picture of the kicked SMBH removing more mass the more number of oscillations it is able to undergo before settling to Brownian motion limits.
+Interestingly, we observe a similar decrease in S\'ersic index $n$ with increasing kick velocity, although no correlation is seen between $n$ and $\Sigma_\mathrm{b}$ in the latent parameter corner plots (see for example \autoref{fig:csparams}).
+
 \begin{table}
-    \caption{Hyperparameter distributions for core-Se\'rsic model.}
+    \caption{Hyperparameter distributions for core-S\'ersic model.}
     \label{tab:hyper}
@@
         $\sigma_{R_\mathrm{e}}$ & $\mathcal{N}(0, 20)$ & $x>0$ \\
-        $\mu_{\sigma_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
-        $\sigma_{\sigma_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
+        $\mu_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
+        $\sigma_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
         \hline
@@
     \includegraphics[width=0.4\textwidth]{dag}
-    \caption{Directed acyclic graph of the core-Se\'rsic model of projected mass density within the Bayesian hierarchical framework.
+    \caption{Directed acyclic graph of the core-S\'ersic model of projected mass density within the Bayesian hierarchical framework.
 Single-line circle nodes represent fit parameters, double-line circles represent measured quantities (the data), and diamond nodes represent deterministic quantities.
@@
     \includegraphics[width=0.4\textwidth]{density}
-    \caption{Density profiles with errors for select representative runs.}
+    \caption{
+        Surface density profiles with 25\% Bayesian HDI for select representative runs.
+        Increasing the kick velocity induces a shallower density profile in the inner regions.
+        }
     \label{fig:density}
@@
         Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{b,0}$, as a function of kick velocity $\vk$.
-        The core size distributions are shown as violin plots, with the median core size indicated by the central mark.
+        The core size distributions are shown as box plots, with the median core size indicated by the central mark.
         Larger kick velocities are correlated with larger core sizes.
@@
 We observe an overall shift to more negative values of $h_4$ for kick velocities up until $\vk\sim400\,\kmps$ \drafting{check this, currently just read off plot}, above which the radial $h_4$ profile is not sensitive to increasing kick velocity.
-Additionally, all $h_4$ profiles are radially decreasing, obtaining a local minimum at $R/\Reff\simeq 0.2$, and a modestly increasing by $R/\Reff\simeq 0.3$.
+Additionally, all $h_4$ profiles are radially decreasing, obtaining a local minimum at $R\simeq 2\,\mathrm{kpc}$ ($R/\Reff\simeq 0.2$), and a modestly increasing for $R \gtrsim 2\,\mathrm{kpc}$.
 
@@
         Mock integral field unit kinematic maps of the $\vk=0\,\kmps$ remnant, plotted out to $0.5\,\Reff$. 
-        Note the unordered rotation $V$, and the centrally-concentrated velocity dispersion $\sigma$.
+        Note the small magnitude of the unordered rotation $V$, and the centrally-concentrated velocity dispersion $\sigma$.
         The majority of the merger remnant within this aperture has $h_4>0$, with a small concentration of $h_4<0$ in an annulus about the central regions.
@@
     \caption{
-        Mock integral field unit kinematic maps of the $\vk=900\,\kmps$ remnant, again plotted out to $0.5\,\Reff$. 
-        Contrasting to the $\vk=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$ \drafting{check this}. 
+        Mock integral field unit kinematic maps of the $\vk=900\,\kmps$ remnant, again plotted out to $0.5\,\Reff$, with the same colour scaling as \autoref{fig:IFU0}. 
+        Contrasting to the $\vk=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$. 
     }
@@
     \caption{
-        Plot of the flux-weighted value of $h_4$ as a function of mean bin distance $R$ (scaled to the half mass radius) for each merger remnant simulated.
+        Plot of the flux-weighted value of $h_4$ as a function of mean bin distance $R$for each merger remnant simulated.
         The lines are coloured by the kick velocity. 
         Immediately apparent is the decrease in $\langle h_4 \rangle$ with increasing $\vk$.
-        All merger remnants have a global minimum in $\langle h_4 \rangle$ at $R/\Reff\simeq0.2$.
+        All merger remnants have a global minimum in $\langle h_4 \rangle$ at $R\simeq 2.0$, corresponding to $R/\Reff\simeq0.2$.
     }
@@
 \end{figure}
+
+\subsection{Orbit analysis}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{orbits}
+    \caption{
+        Orbit analysis of all sixteen merger remnants.
+        Stellar particles \drafting{currently it's just a subset} are assigned to one of seven different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 100$.
+        At radii greater than $1\,\mathrm{kpc}$, there is very little to distinguish between the different kick velocities for boxlet, $z$-tube, Keplerian, irregular, and unclassified orbits.
+        Conversely, for radii $1 \leq R/\mathrm{kpc} \leq 10$, there is a clear gradient for $\pi$-box and $x$-tube families.
+        For higher kick velocities, there is a tendency to a higher fraction of $\pi$-box orbits compared to lower kick velocities.
+        An inverse trend is visible for $x$-tube orbits. 
+    }
+    \label{fig:orbits}
+\end{figure*}
+
+We perform an orbit analysis on the sixteen merger remnants in our sample following \citet{frigo2021}.
+\drafting{Discuss how the orbit analysis is done, in particular the different orbital families}.
+First, the merger remnant is rotated so that the $z$-axis coincides with the minor axis of the reduced inertia tensor (and hence has the same orientation as the IFU maps in \autoref{fig:IFU0} and \autoref{fig:IFU900}).
+The potential of the merger remnant is fit using a \drafting{self consistent field (SCF) potential, describe more?} method, providing an analytical potential in which the orbits of individual stellar particles can be integrated.
+The potential of the merger remnant is checked for stability, and each \drafting{currently a subset} stellar particle is integrated for fifty orbits to determine which, if any, orbital resonances exist.
+
+As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box orbits, and a lower fraction of $x$-tube orbits, than their lower kick velocity counterparts.
+The other five orbital families are generally consistent between all merger remnants. 
+\drafting{
+    Might move this to discussion, just put here for now to get the idea down.
+    z-tubes not primary contributor to LOS velocity (rotation about z axis).
+    Higher kick correlated with reduction in x-tubes (orbits about the x-axis), increase in pi-box.
+    Pi-box show no resonant motion (no net ang. mom.)\dots here's the leap: reduced resonant motion seen as a reduction in `symmetric deviations' in line profile?
+    Also look directly at kinematic maps: 0 kick case. Dominated by x tubes at radii > 1kpc, small region of negative h4 extends to here before becoming positive. 
+}
 
@@
 
-\section{Core-Se\'rsic fit parameter estimates}\label{sec:app_fit}
+\section{Core-S\'ersic fit parameter estimates}\label{sec:app_fit}
 
@@
     \caption{
-        Violin plots of the six parameters in the core-Se\'sic model (from top left to bottom right) as a function of the kick velocity: the core radius normalised to the pre-merger core radius $\rb/r_{\mathrm{b},0}$, the density at the core radius $\log_{10}\Sigma_\mathrm{b}$, the core slope $\gamma$, the effective radius $\Reff$, the profile transition index $\alpha$, and the Se\'rsic index $n$.
+        Box plots of the six parameters in the core-S\'ersic model (from top left to bottom right) as a function of the kick velocity: the core radius normalised to the pre-merger core radius $\rb/r_{\mathrm{b},0}$, the density at the core radius $\log_{10}\Sigma_\mathrm{b}$, the core slope $\gamma$, the effective radius $\Reff$, the profile transition index $\alpha$, and the S\'ersic index $n$.
     }
@@
     \centering
-    \includegraphics[width=\textwidth]{corner_latent_960}
+    \includegraphics[width=\textwidth]{corner_latent_900}
     \caption{
-        Corner plot of the latent parameters in the core-Se\'rsic model for the $\vk=960\,\kmps$ instance. \drafting{UPDATE TO 900}
+        Corner plot of the latent parameters in the core-S\'ersic model for the $\vk=900\,\kmps$ instance.
         Contours indicate, from yellow to dark blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs), with sample draws beyond the 99\% HDI shown as individual points. 
@@
     }
+\end{figure*}
+
+\section{Additional orbital plots}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{orbits2}
+    \caption{
+        Same as \autoref{fig:orbits}, but each merger remnant is shown in its own panel, with the different orbital families depicted by colour, to better demonstrate the radial fraction of orbits for a given merger.
+        \drafting{TODO: maybe change km/s to $\kmps$}
+    }
+    \label{fig:orbits2}
 \end{figure*}

### 2024-02-07 papers/paper-corekick/paper-corekick.tex
Writing on orbital stuff

@@
 
+Following \citet{nasim2021b}, we fit a power law of the form
+\begin{equation}\label{eq:vkrb}
+    \frac{\rb}{r_{\mathrm{b},0}} = K \left( \frac{\vk}{v_\mathrm{esc}} \right)^\beta + 1
+\end{equation}
+to our simulations, which is thus valid for $0\leq\vk/\kmps\leq900$, or equivalently $0\leq\vk/v_\mathrm{esc}\leq 0.75$.
+\drafting{We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=$ and $\beta=$.}
+
+\drafting{Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal alignment (their Fig. 15, right panel.) 
+This allows us to obtain a distribution of core radius values predicted by a given kick velocity model.
+We find\dots}
+
 \begin{table}
@@
 \begin{equation}
-    \langle h_4 \rangle = \frac{\langle R h_4 \rangle}{\langle R \rangle},
+    \langle h_4(R') \rangle = \frac{\langle R h_4 \rangle}{\langle R \rangle} = \frac{\sum_i^N M_i h_{4,i} R_i}{\sum_i^N M_i R_i} \;\forall R_i \leq R',
 \end{equation}
@@
 The other five orbital families are generally consistent between all merger remnants. 
+As discussed in \citet{frigo2021}, negative values of $h_4$ indicate regions with weaker tails of the LOS velocity distribution due to the superposition of radial orbits with different orientations.
+This explanation supports our observation of an increase in $\pi$-box orbits (which have no net angular momentum are typically very radial orbits) with kick velocity, and a corresponding decrease in the value of $h_4$ (comparing the IFU maps in \autoref{fig:IFU0} to \autoref{fig:IFU900}).
+Equivalently, $x$-tube orbits (which rotate about the major axis of the inertia tensor) appear as a peaked LOS velocity distribution in the projection we use, leading to extended tails, and hence a more positive value of $h_4$ than for a weak tail distribution.
+All together, we can infer that as a SMBH ploughs through the surrounding stellar environment, regular $x$-tube orbits are disrupted to non-resonant $\pi$-box orbits.
+With a higher kick velocity, the SMBH is able to cumulatively disturb orbits at a larger radial extent, converting a greater fraction of $x$-tube orbits to $\pi$-box orbits, resulting in the increased expanses of negative $h_4$ regions in the kinematic maps.
+This also explains the overall downward shift in the $\langle h_4 \rangle$ profiles in \autoref{fig:h4} for higher kick velocities. 
+\drafting{Why are the $\langle h_4 \rangle$ profiles consistent in shape though??}
+
 \drafting{

### 2024-02-17 papers/paper-corekick/paper-corekick.tex
Small textual changes to reflect new figures

@@
 We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.
-Additionally, the merger remnant displays almost no triaxiality, as seen in the $\vk=0\,\kmps$ case in \autoref{sec:app_triax}.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+Additionally, the merger remnant displays almost no triaxiality, as seen in the $\vk=0\,\kmps$ case in \autoref{sec:app_triax}.
+We check for this by directing the recoil kick along the global $y$-axis, and final identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
 The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant.
@@
 Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
-
-\drafting{Atte's runs of mergers of unequal mass ratio?}
-
-\drafting{Kicks randomly orientated?}
 
@@
     \includegraphics[width=0.4\textwidth]{rb_pdf}
-    \caption{Probability density function of core radius, where kick velocity is MC sampled from Zlochower relation, and pumped through our fitted model in \autoref{fig:vkrb}.}
+    \caption{Probability density function of core radius, where kick velocity is MC sampled from Zlochower relation, and pumped through our fitted model in \autoref{eq:vkrb}.}
     \label{fig:rb_pdf}
@@
         Orbit analysis of all sixteen merger remnants.
-        Stellar particles \drafting{currently it's just a subset} are assigned to one of seven different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 100$.
+        Stellar particles are assigned to one of seven different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 30.0$.
         At radii greater than $1\,\mathrm{kpc}$, there is very little to distinguish between the different kick velocities for boxlet, $z$-tube, Keplerian, irregular, and unclassified orbits.
@@
 The potential of the merger remnant is fit using a \drafting{self consistent field (SCF) potential, describe more?} method, providing an analytical potential in which the orbits of individual stellar particles can be integrated.
-The potential of the merger remnant is checked for stability, and each \drafting{currently a subset} stellar particle is integrated for fifty orbits to determine which, if any, orbital resonances exist.
+The potential of the merger remnant is checked for stability, and each stellar particle within $30\,\mathrm{kpc}$ of the centre is integrated for fifty orbits to determine which, if any, orbital resonances exist.
 
@@
         Same as \autoref{fig:orbits}, but each merger remnant is shown in its own panel, with the different orbital families depicted by colour, to better demonstrate the radial fraction of orbits for a given merger.
-        \drafting{TODO: maybe change km/s to $\kmps$}
     }

### 2024-02-27 papers/paper-corekick/paper-corekick.tex
Update figures, small edits to text

@@
 \section{Results}\label{sec:results}
-\drafting{
-    Things to show:
-    \begin{enumerate}
-        \item DAG for hierarchical model
-        \item Density profiles 
-        \item Scatter plot $v_k\mathrm{-}r_b$ with best fit function
-        \item Relation $v_k\mathrm{-}r_b$ model comparison (elpd?)
-        \item MC-sampled distribution of $r_b$
-        \item $h_4$ maps
-    \end{enumerate}
-}
 
@@
     \begin{enumerate}
-        \item Parameters associated with the core (most importantly the core size) have a dependence on the kick velocity, whereas large-scale parameters (e.g. effective radius) do not.
+        \item Core radius, break density, sersic index have a dependence on the kick velocity, whereas other parameters (e.g. effective radius) do not.
         \item Increasing kick velocity induces a decrease in the overall $\langle h_4 \rangle$ profile
+        \item Explanation of h4 trends through orbit decompostion.
     \end{enumerate}
@@
 }
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{infall_time}
+    \caption{Plot of infall time (time to stability criterion) as a function of kick velocity.}
+    \label{fig:infall}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{max_disp}
+    \caption{Plot of the maximum BH displacement as a function of kick velocity. Where do the displacement (red) and core size (green) lines intersect?}
+    \label{fig:max_disp}
+\end{figure}
 
@@
 to our simulations, which is thus valid for $0\leq\vk/\kmps\leq900$, or equivalently $0\leq\vk/v_\mathrm{esc}\leq 0.75$.
-\drafting{We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=$ and $\beta=$.}
+We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=2.830$ and $\beta=0.558$.
 
-\drafting{Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal alignment (their Fig. 15, right panel.) 
-This allows us to obtain a distribution of core radius values predicted by a given kick velocity model.
-We find\dots}
+Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig. 15, right panel).
+The dimensionless spin parameter for the dry mergers follows a beta distribution, namely:
+\begin{equation}
+    P_\mathrm{Zlochower}(\alpha) \propto (1-\alpha)^{4.66884-1} \alpha^{10.5868-1}
+\end{equation}
+This allows us to obtain a distribution of core radius values predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}.
+We find that the mode of the forward-folded core radius distribution is $1.61\,\mathrm{kpc}$ ($3.06\,r_{\mathrm{b},0}$).
+Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, most massive elliptical galaxies should have a non-negligible core \drafting{add references}.
 
@@
 
-\drafting{Why is this 
+\drafting{Jens: maybe you have some thoughts\dots Why is this 
 \begin{enumerate}
-    \item decreasing with increasing kicks,
     \item saturating beyond some point,
@@
 
-We perform an orbit analysis on the sixteen merger remnants in our sample following \citet{frigo2021}.
+To understand the kinematic maps, we perform an orbit analysis on the sixteen merger remnants in our sample following \citet{frigo2021}.
 \drafting{Discuss how the orbit analysis is done, in particular the different orbital families}.
@@
 
-As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box orbits, and a lower fraction of $x$-tube orbits, than their lower kick velocity counterparts.
-The other five orbital families are generally consistent between all merger remnants. 
+As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box and boxlet orbits, and a lower fraction of $x$-tube orbits, and to some extent Keplerian orbits, than their lower kick velocity counterparts.
+The other two orbital families are generally consistent between all merger remnants. 
 As discussed in \citet{frigo2021}, negative values of $h_4$ indicate regions with weaker tails of the LOS velocity distribution due to the superposition of radial orbits with different orientations.
@@
 This also explains the overall downward shift in the $\langle h_4 \rangle$ profiles in \autoref{fig:h4} for higher kick velocities. 
-\drafting{Why are the $\langle h_4 \rangle$ profiles consistent in shape though??}
-
-\drafting{
-    Might move this to discussion, just put here for now to get the idea down.
-    z-tubes not primary contributor to LOS velocity (rotation about z axis).
-    Higher kick correlated with reduction in x-tubes (orbits about the x-axis), increase in pi-box.
-    Pi-box show no resonant motion (no net ang. mom.)\dots here's the leap: reduced resonant motion seen as a reduction in `symmetric deviations' in line profile?
-    Also look directly at kinematic maps: 0 kick case. Dominated by x tubes at radii > 1kpc, small region of negative h4 extends to here before becoming positive. 
-}
+A peculiar feature of \autoref{fig:h4} is the consistency in the $\langle h_4 \rangle$ profile shape.
+This can be explained by the derivative of the orbital fraction distribution being consistent across all kick velocities for $x$-tube and boxlet orbits out to $\sim3\,\mathrm{kpc}$, as shown in \autoref{fig:orbits2}.
+\drafting{Thorsten: I would really appreciate your thoughts on this section, and its relation to the kinematic maps. }
 

### 2024-02-27 papers/paper-corekick/scripts/core-dist.py

@@ -9,6 +9,7 @@ except ImportError:
     use("Agg")
     import matplotlib.pyplot as plt
 import cm_functions as cmf
+import figure_config # noqa
 
 parser = argparse.ArgumentParser(
     "Determine core - kick relation",

### 2024-03-01 papers/paper-corekick/paper-corekick.tex
add Peter comments

@@
 
+\subsection{Simulation Code}
 \drafting{Taken from eccentricity paper, leave here as a guide for a similar description about Ketju}
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017}.
+We construct a number of idealised galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017} coupled with \gadget{}.
 The dynamics of
@@
 
+\subsection{Initial Conditions}
 \drafting{Rantala 2018 ICs}
@@
 
-We create six independently Monte Carlo sampled galaxy ICs, and merge each combination of galaxies on a radial orbit.
+\subsection{Merger Simulations}
+We merge two independently Monte Carlo sampled galaxy ICs, creating an equal mass galaxy merger with a near-radial orbit. 
 The merger orbit has an initial separation of $D=30\,\mathrm{kpc}$, a first pericentre distance of $r_\mathrm{peri}=2\,\mathrm{kpc}$, and an initial eccentricity of $e_0=0.97$.
-We select a merger combination that results in a rapid coalescence of the BH binary, noting that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
-We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.
+The merger configuration results in a rapid coalescence of the BH binary, however we note that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
+We then select the snapshot just prior ($\sim8\,\mathrm{Myr}$) to the GW-driven SMBH merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.
 Additionally, the merger remnant displays almost no triaxiality, as seen in the $\vk=0\,\kmps$ case in \autoref{sec:app_triax}.
 We check for this by directing the recoil kick along the global $y$-axis, and final identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
-The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant.
-We additionally run one simulation with a recoil kick above the escape velocity, with $\vk=2000\,\kmps$.
+The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
+We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$.
 Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
+The time to this point, from the time of SMBH binary coalescence, we term the `infall time'.
 Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
@@
 In agreement with previous studies \citep{nasim2021b}, we find an increase in the core radius $\rb$ with kick velocity.
-\drafting{Need to check if others discuss this:} We find that the core radius saturates to a value of $\sim2$ times that of the merger remnant with no kick velocity by $\vk = 600\,\kmps$, corresponding to half of the escape velocity of the merger remnant core.
+\drafting{Need to check if others discuss this:} We find that the core radius saturates to a value of $\sim2$ times that of the merger remnant with no kick velocity by $\vk = 600\,\kmps$, corresponding to $0.3 v_\mathrm{esc}$.
 As shown in \autoref{fig:csparams}, the density at the core radius $\Sigma_\mathrm{b}$ decreases monotonically with kick velocity, consistent with the picture of the kicked SMBH removing more mass the more number of oscillations it is able to undergo before settling to Brownian motion limits.
@@
 
+\subsection{Predicted core size distribution}
 Following \citet{nasim2021b}, we fit a power law of the form
@@
 \end{equation}
-to our simulations, which is thus valid for $0\leq\vk/\kmps\leq900$, or equivalently $0\leq\vk/v_\mathrm{esc}\leq 0.75$.
-We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=2.830$ and $\beta=0.558$.
+to our simulations, which is thus valid for $0\leq\vk/\kmps\leq900$, or equivalently $0\leq\vk/v_\mathrm{esc}\leq 0.5$.
+We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=3.70$ and $\beta=0.562$.
 
 Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig. 15, right panel).
-The dimensionless spin parameter for the dry mergers follows a beta distribution, namely:
+The dimensionless spin parameter $\alpha_\bullet$ for the dry mergers follows a beta distribution, namely:
 \begin{equation}
-    P_\mathrm{Zlochower}(\alpha) \propto (1-\alpha)^{4.66884-1} \alpha^{10.5868-1}
+    P_\mathrm{Zlochower}(\alpha_\bullet) \propto (1-\alpha_\bullet)^{4.66884-1} \alpha_\bullet^{10.5868-1}.
 \end{equation}
-This allows us to obtain a distribution of core radius values predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}.
-We find that the mode of the forward-folded core radius distribution is $1.61\,\mathrm{kpc}$ ($3.06\,r_{\mathrm{b},0}$).
+The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$ \drafting{check this}.
+Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}.
+We find that the mode of the forward-folded core radius distribution is $1.64\,\mathrm{kpc}$ ($3.12\,r_{\mathrm{b},0}$).
 Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, most massive elliptical galaxies should have a non-negligible core \drafting{add references}.
@@
 \begin{table}
-    \caption{Hyperparameter distributions for core-S\'ersic model.}
+    \caption{Hyperparameter distributions for core-S\'ersic model. $\mathcal{N}$ indicates a normal distribution, $\mathrm{Exp}(\lambda)$ indicates an exponential distribution.}
     \label{tab:hyper}
@@
 The $R$ box indicates variables fit for each radial bin, and the projection box indicates variables specific to each projection realisation.
-Note that we additionally use $\hat{\Sigma}$ to distinguish the calculated value of surface density from the measured value $\Sigma$.}
+Note that we additionally use $\hat{\Sigma}$ to distinguish the calculated value of surface density from the measured value $\Sigma$.
+The various distributions are normal ($\mathcal{N}$), exponential ($\mathrm{Exp}$), and Rayleigh ($\mathrm{Ray}$).}
     \label{fig:dag}
@@
     \includegraphics[width=0.4\textwidth]{rb_pdf}
-    \caption{Probability density function of core radius, where kick velocity is MC sampled from Zlochower relation, and pumped through our fitted model in \autoref{eq:vkrb}.}
+    \caption{Probability density function of transformation sampled core radius.
+The SMBH kick velocity is Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins, and pushed through the fitted model in \autoref{eq:vkrb}.
+The predicted kick velocities range from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
+\drafting{check this}}
     \label{fig:rb_pdf}

### 2024-03-25 papers/paper-corekick/paper-corekick.tex
Add new description of when snapshots are analysed following Thorsten discussions

@@
 \\%
-$^{2}$
-Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1, D-85748 Garching, Germany
+%$^{2}$
+%Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1, D-85748 Garching, Germany
 }
@@
 We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$.
-Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
-The time to this point, from the time of SMBH binary coalescence, we term the `infall time'.
-Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
-In practice, this limits our sample of analysed simulations to those with $\vk\leq 900\,\kmps$, thus totalling sixteen simulations in our analysis.
-Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
+As discussed in \drafting{section}, higher recoil velocities typically result in larger excursions of the remnant SMBH from the isophotal centre.
+To provide a fair comparison between simulations, we choose to analyse each child simulation at a fixed time after the recoil kick, namely $t\simeq 0.075\,\mathrm{Gyr}$.
+This value is chosen so as to minimise the effects of relaxation in the central regions of the galaxy remnant.
+We estimate the relaxation time $t_\mathrm{relax}$ at a radius $r$ as \citep{binney2008}:
+\begin{equation}
+    t_\mathrm{relax} \simeq 2.1 \frac{\sigma r^2}{G \bar{m} \ln \Lambda},
+\end{equation}
+where $\sigma$ is the particle velocity dispersion within $r$, $\bar{m}$ is the mean particle mass (hence the mean of the stellar and DM particle masses), and $\Lambda$ is the argument of the Coulomb logarithm, given by:
+\begin{equation}
+    \Lambda = \frac{r \langle v^2 \rangle}{2 G \bar{m}},
+\end{equation}
+where $\langle v^2 \rangle$ is the mean squared particle velocity within $r$. 
+The relaxation time increases exponentially with increasing $r$, and we find for a radius of $r=0.1\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 0.1 \, \mathrm{Gyr}$, thus motivating our choice of $0.075\,\mathrm{Gyr}$ for the analysis time.
+
+For all simulations, interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, DM with a softening length of $\varepsilon = 300\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
 
@@
     \end{enumerate}
+}
+\drafting{
+    Include this text here, this is how we identify `settled' systems, but we're not really interested in when the SMBH has settled for the rest of the paper (refer to Nasim where the majority of the GW-induced scouring occurs during the first excursion of the SMBH).
+    Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
+    The time to this point, from the time of SMBH binary coalescence, we term the `infall time'.
+    Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
 }

### 2024-04-08 papers/paper-corekick/paper-corekick.tex
Update some methods in the text to reflect changes done during analysis
- new figures yet to be added

@@
 \drafting{Rantala 2018 ICs}
-We model the merger of two gas-poor elliptical galaxies using collisionless merger simulations.
+We model the merger of two gas-free elliptical galaxies using collisionless merger simulations.
 Our galaxy initial conditions are chosen to match the model IC-3 presented in \citet{rantala2019}. 
@@
 \begin{equation}\label{eq:dehnen}
-    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
+    \rho(r) = \frac{(3-\gamma)M}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
 \end{equation}
@@
 We then select the snapshot just prior ($\sim8\,\mathrm{Myr}$) to the GW-driven SMBH merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.
-Additionally, the merger remnant displays almost no triaxiality, as seen in the $\vk=0\,\kmps$ case in \autoref{sec:app_triax}.
-We check for this by directing the recoil kick along the global $y$-axis, and final identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
 The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
@@
 We estimate the relaxation time $t_\mathrm{relax}$ at a radius $r$ as \citep{binney2008}:
-\begin{equation}
+\begin{equation}\label{eq:relax}
     t_\mathrm{relax} \simeq 2.1 \frac{\sigma r^2}{G \bar{m} \ln \Lambda},
@@
 where $\langle v^2 \rangle$ is the mean squared particle velocity within $r$. 
-The relaxation time increases exponentially with increasing $r$, and we find for a radius of $r=0.1\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 0.1 \, \mathrm{Gyr}$, thus motivating our choice of $0.075\,\mathrm{Gyr}$ for the analysis time.
+The relaxation time increases quadratically with increasing $r$, and we find for a radius of $r=0.1\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 0.1 \, \mathrm{Gyr}$, thus motivating our choice of $0.075\,\mathrm{Gyr}$ for the analysis time.
+We  note that \autoref{eq:relax} does not take into account gravitational softening, and thus provides a conservative lower bound for the relaxation time.
+This is critical as the \ketju{} region does not include gravitational softening between stellar particles and the SMBH, so is a not fully-softened simulation.
 
@@
 \end{equation}
-Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius, $\alpha$ is the profile transition index, and $n$ is the S\'ersic index.
+Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $\alpha$ is the profile transition index, and $n$ is the S\'ersic index.
 We refer to the collective vector of these parameters as $\vb{\theta}$.
-We view the simulated merger remnant from fifteen different angles, and use a Bayesian hierarchical model (HM) to fit the model parameters.
+We view the simulated merger remnant from fifteen random angles, and use a Bayesian hierarchical model (HM) to fit the model parameters.
 As each of the projections are unordered and exchangeable (i.e., there is no distinguishing label for the projections), and each projection is of the same merger remnant, a HM naturally lends itself.
-The idea behind the HM is straightforward: for each projection, we infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible vectors $\vb{\theta}$.
+The idea behind the HM is as such: for each projection, we infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible $\vb{\theta}$ vectors.
 Hence, we are effectively inferring the global hyperparameters $\vb{\theta}^\mathrm{hyp}$ that describe the latent parameters $\vb{\theta}$ that we are interested in.
@@
         \hline
-        $\mu_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(10, 1)$ & None \\
-        $\sigma_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(0, 0.05)$ & $x>0$ \\
+        $\mu_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(10, 2)$ & None \\
+        $\sigma_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
         $\lambda_\gamma$ & $\mathrm{Exp}(10)$ & None \\
-        $\mu_n$ & $\mathcal{N}(4,2)$ & $0 < x \leq 15$ \\
-        $\sigma_n$ & $\mathcal{N}(0,2)$ & $x>0$ \\
-        $\sigma_\alpha$ & $\mathcal{N}(0, 20)$ & $x>0$ \\
-        $\sigma_{r_\mathrm{b}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
-        $\sigma_{R_\mathrm{e}}$ & $\mathcal{N}(0, 20)$ & $x>0$ \\
+        $\mu_n$ & $\mathcal{N}(8,4)$ & $0 < x \leq 15$ \\
+        $\sigma_n$ & $\mathcal{N}(0,4)$ & $x>0$ \\
+        $\sigma_\alpha$ & $\mathrm{Gamma}(2, 0.2)$ & None \\
+        $\sigma_{r_\mathrm{b}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
+        $\sigma_{R_\mathrm{e}}$ & $\mathcal{N}(0, 12)$ & $x>0$ \\
         $\mu_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
@@
 \subsection{Integral field unit kinematics}
-We create mock integral field unit (IFU) observations using \drafting{ reference to Cappellari maybe?}. 
-Prior to performing the analysis, we reorientate each merger remnant so that the $z$-axis coincides with the minor axis of the reduced inertia tensor.
+We create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the BH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the BH kick axis (`orthogonal').
+Testing these two special projections, we investigate the limiting cases of how the galaxy kinematics behave when all or none of the BH motion is directed along the LOS to the observer.
 
-\drafting{Describe voronoi tesselation}.
-For each Voronoi bin, we follow \citet{vandermarel1993} and decompose the line of sight (LOS) velocity into a series of Gauss-Hermite functions, described by the mean radial velocity $V$, the velocity dispersion $\sigma$, asymmetric deviations $h_3$, and symmetric deviations $h_4$.
+We create our mock IFU observations following \citet{naab2014}.
+Our observations are centred on the stellar centre of mass using the shrinking sphere method, and encompass a circular aperture of $0.25R_{1/2}$, where $R_{1/2}$ is the three-dimensional half mass radius.
+For each particle within this aperture, we generate 25 pseudo-particles with identical LOS velocities as the original particle, but are spatially displaced from the original particle with a Gaussian distribution in the projected $x$ and $y$ direction with a standard deviation of $0.3\,\mathrm{kpc}$, to mimic seeing effects.
+The pseudo-particles are then binned onto a spatial grid, where the grid is assumed to have a resolution of $0.2''$, similar to the observations of \citet{neureiter2023}, which then corresponds to a physical resolution\footnote{https://cosmocalc.icrar.org} of $\sim0.04\,\mathrm{kpc}$ at redshift $z=0.01$.
+Following the method in \citet{cappellari2003}, we then group adjacent bins into larger Voronoi bins to achieve a particle number of $\sim 50000$ particles per Voronoi bin.
+
+For each Voronoi bin, we follow \citet{vandermarel1993} and decompose the LOS velocity into a series of Gauss-Hermite functions, described by the mean radial velocity $V$, the velocity dispersion $\sigma$, asymmetric deviations $h_3$, and symmetric deviations $h_4$.
 The line profile is thus described by:
@@
 We observe a trend of the Voronoi bins with $h_4<0$ extending to larger radii for remnants with larger recoil kicks than for those remnants with smaller recoil kicks.
-To quantify this recoil velocity dependence, we calculate a radially-varying mass-weighted $h_4$ parameter analogous to the observational spin parameter $\lambda$:
+To quantify this recoil velocity dependence, we calculate the cumulative $h_4$ parameter:
 \begin{equation}
-    \langle h_4(R') \rangle = \frac{\langle R h_4 \rangle}{\langle R \rangle} = \frac{\sum_i^N M_i h_{4,i} R_i}{\sum_i^N M_i R_i} \;\forall R_i \leq R',
+    \langle h_4 \rangle = \sum_i^N h_4(R_i)
 \end{equation}
-where the $\langle\rangle$ brackets indicate a mass-weighted average (thus equivalent to a flux weighted average in the case of constant mass-to-light ratio).
 We observe an overall shift to more negative values of $h_4$ for kick velocities up until $\vk\sim400\,\kmps$ \drafting{check this, currently just read off plot}, above which the radial $h_4$ profile is not sensitive to increasing kick velocity.
@@
 
-To understand the kinematic maps, we perform an orbit analysis on the sixteen merger remnants in our sample following \citet{frigo2021}.
+To understand the kinematic maps, we perform an orbit analysis of the merger remnants in our sample following \citet{frigo2021}, and is briefly described here.
+
 \drafting{Discuss how the orbit analysis is done, in particular the different orbital families}.
-First, the merger remnant is rotated so that the $z$-axis coincides with the minor axis of the reduced inertia tensor (and hence has the same orientation as the IFU maps in \autoref{fig:IFU0} and \autoref{fig:IFU900}).
-The potential of the merger remnant is fit using a \drafting{self consistent field (SCF) potential, describe more?} method, providing an analytical potential in which the orbits of individual stellar particles can be integrated.
-The potential of the merger remnant is checked for stability, and each stellar particle within $30\,\mathrm{kpc}$ of the centre is integrated for fifty orbits to determine which, if any, orbital resonances exist.
+First, the merger remnant is rotated so that the $z$-axis coincides with the minor axis of the reduced inertia tensor as measured using the top 50\% most bound stellar particles to the stellar centre of mass.
+We cannot centre on the recoiling SMBH for this analysis, as for high kick velocities the SMBH is significantly displaced from the centre.
+The stellar potential of the merger remnant is fit using a self consistent field (SCF) potential \citep{hernquist1992}, using the \citet{hernquist1990} profile as a zeroth-order basis and limiting the expansion to $n_\mathrm{max}=18$ and $l_\mathrm{max}=7$ following \citet{frigo2021}.
+The potential from the SMBH is then added as a point mass potential to the SCF potential.
+Collectively, this represents the potential in an analytical form in which the orbits of individual stellar particles can be integrated.
+The potential of the merger remnant is checked for stability by comparing the potential from the SCF method to the potential computed from the particle data, and ensuring the ratio of the two is $\sim1$ at all radii.
+Each stellar particle within $30\,\mathrm{kpc}$ of the centre is integrated for fifty orbits to determine which, if any, orbital resonances exist. 
+The orbital resonances define the different families of orbits, as given in \citet{frigo2021}.
+
+\begin{table}
+    \caption{Orbital families}
+    \label{tab:orbits}
+    \begin{tabular*}{0.48\textwidth}{p{0.1\textwidth}p{0.35\textwidth}}
+        \hline
+        Family & Description \\
+        \hline
+        $x$-tube & Rotate about the major axis of the galaxy \\
+        $z$-tube & Rotate about the minor axis of the galaxy \\
+        $\pi$-box & Non-resonant motion with no net angular momentum (radial motion) \\
+        boxlet & Resonant motion with angular momentum \\
+        Keplerian & Typical orbit in a point-mass dominated spherically-symmetric potential \\
+        irregular & No integrals of motion \\
+        unclassified & Orbits unable to be classified \\
+        \hline
+    \end{tabular*}
+\end{table}
 

### 2024-04-08 papers/paper-corekick/paper-corekick.tex
Implement Noa's comments

@@
 
-We use weakly-constrained priors for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these priors are not necessarily Gaussian (given that distributions of positive-constrained quantities, like radius, should not be described by a distribution with support on the full real space).
+We use weakly-constrained prior distributions for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these prior distributions are chosen to reflect reasonable values that the parameters may take (e.g.
+distributions of a radial quantity are chosen so as to be positively-constrained).
 The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
-We use the Hamiltonian Monte Carlo code \textsc{Stan} \citep{standevelopmentteam2018} with four chains each of 2000 posterior sample draws (excluding an additional 2000 burn-in draws) to fit the model parameters, ensuring that:
-\begin{enumerate}
-    \item the number of diverging draws is less than 5\%,
-    \item the effective draws per transition exceeds 0.001,
-    \item the sampler transitions of HMC potential energy is in excess of 0.3, and
-    \item the value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.
-\end{enumerate}
-Example corner plots of the latent parameters can be found in \autoref{sec:app_fit}.
-\drafting{Do we need specifics of MC parameters and exit conditions?}
+
+We fit the model with \textsc{Stan} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
+This resulted in 2000 draws from the posterior distribution of $\vb{\theta}$, whereby we ensure the diagnostic value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.
+We perform additional prior sensitivity tests using the power scaling method described in \citet{kallioinen21}, ensuring that the cumulative Jensen-Shannon (CJS) divergence is less than 0.05 for the latent parameters $\vb{\theta}$ of the model.
+This recommended threshold indicates that the sampled posterior distributions of the latent parameters are not sensitive to the particular prior distributions used.
 
@@
 Single-line circle nodes represent fit parameters, double-line circles represent measured quantities (the data), and diamond nodes represent deterministic quantities.
-The particular distribution connecting nodes is written below the corresponding black square, with a subscript `T' indicating a truncated distribution, and the subscript `likelihood' indicating the likelihood function.
+The particular distribution connecting nodes is written below the corresponding black square, with a subscript `$\mathrm{T}[l,u]$' indicating a distribution $f(\lambda)$ truncated to $l<\lambda<u$, and the subscript `likelihood' indicating the likelihood function.
 The $R$ box indicates variables fit for each radial bin, and the projection box indicates variables specific to each projection realisation.
@@
 Matplotlib \citep{hunter2007},
-pygad \citep{rottgers2020}.
+pygad \citep{rottgers2020},
+\textsc{Stan} \citep{standevelopmentteam2018},
+CmdStanPy \citep{standevelopmentteam2018},
+Arviz \citep{kumar2019}.
 

### 2024-04-11 papers/paper-corekick/paper-corekick.tex
Begin working on intro

@@
 \section{Introduction}
-\drafting{Some general intro here}
+There is much evidence for the presence of supermassive black holes (SMBHs) at the centres of massive galaxies, with the shadow of one such SMBH being directly imaged.
+The dynamical interactions of two or more galaxies, during a galaxy merger, are thus also expected to induce an interaction of the residing SMBHs.
+The interaction of two SMBHs during the merger of two galaxies is well understood to occur in a three stage process \citep{begelman1980}.
+First, as the galaxies begin to dynamically interact and the SMBHs traverse through a dense stellar field, the overdensity of stars in the wakes of the SMBHs provides a restoring force which brings the SMBHs to the centre of the two interacting galaxies, generally torquing the SMBHs to highly radial orbits in the process.
+Second, after the SMBHs have formed a bound binary, iterative interactions with the surrounding stars ejects stars with high velocity, thus removing orbital energy and angular momentum from the SMBH binary orbit.
+Finally, at very small, sub-parsec, separations the SMBH binary loses its remaining orbital energy and angular momentum through gravitational wave (GW) emission, driving the SMBH binary to coalescence \citep{peters1963,peters1964}.
+
+Due to the anisotropic emission of the final GW emission \drafting{during? following?} coalescence, the remnant SMBH recoils with some velocity, termed the kick velocity. 
+The kick velocity is dependent on the mass ratio between the SMBHs prior to coalescence, and the magnitude and direction of angular momentum (spin) vector of each SMBH. 
+Previous studies \drafting{reference} indicate that while the majority of recoil kicks are of the order of a few hundred kilometres per second, in special configurations the SMBH may have an imparted kick velocity in excess of $4000\,\kmps$ \drafting{references}.
+Taking the central escape velocity of a typical early type galaxy (ETG) as $\sim2000\,\kmps$, there may be a non-negligible fraction of ETGs with a SMBH that has escaped the galaxy.
+In cases where the kick velocity is less than the escape velocity of the inner regions of the host galaxy, the radial oscillations of the remnant SMBH as it returns to the centre of the merger remnant is expected to influence the surrounding stellar environment.
+
+In this work, we first aim to determine how the velocity with which the SMBH is kicked affects the surrounding stellar environment.
+We then aim to determine if there is an observational signature of the SMBH recoil kick that may be used to infer the occurrence of a kicked SMBH. 
+\drafting{Previous work by \dots indicates \dots}
+
+\drafting{This paper is divided as follows:}
 
@@
 
-\subsection{Density profiles}
-We fit the six-parameter core-S\'ersic profile to the projection of each simulated merger remnants:
+\subsection{Mass density profiles in three dimensions}
+Before considering the projected, 2D stellar mass density profiles of the merger remnants, we investigate the 3D stellar mass distribution.
+This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
+We observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $2\,\mathrm{kpc}$ than those mergers with $\vk>540\,\kmps$.
+All merger remnants have a consistent mass density distribution at $r>2\,\mathrm{kpc}$. 
+The reasons for the dichotomy are discussed further in \drafting{section}, however we continue our analysis using the projected stellar density (as an observer would see) with the knowledge that there is \textit{a} dependence of the removal of central stellar mass on the SMBH kick velocity.
+
+\drafting{INCLUDE FIGURE HERE}
+
+\subsection{Projected mass density profiles}
+To facilitate comparisons with observations \drafting{some references}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projection of each simulated merger remnants:
 \begin{equation}\label{eq:cs}
@@
 \end{equation}
-Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $\alpha$ is the profile transition index, and $n$ is the S\'ersic index.
+Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $n$ is the S\'ersic index, $b$ is estimated as $b\simeq 2n - 1/3 + 0.009876/n$ \citep{prugniel1997}, and $\alpha$ is the profile transition index.
 We refer to the collective vector of these parameters as $\vb{\theta}$.
@@
 
-We fit the model with \textsc{Stan} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
+We fit the model with \textsc{Stan} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo (HMC) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
 This resulted in 2000 draws from the posterior distribution of $\vb{\theta}$, whereby we ensure the diagnostic value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.

### 2024-04-13 papers/paper-corekick/paper-corekick.tex
Update intro, methods sections

@@
 \newcommand{\dv}[2]{\ensuremath{\frac{\dd{#1}}{\dd{#2}}}}     % derivative
+\newcommand{\dnv}[3]{\ensuremath{\frac{\mathrm{d}^#1#2}{\dd{#3}^#1}}}  % nth derivative
 \newcommand{\vb}[1]{\ensuremath{\bm{#1}}}                     % vectors
@@
 \section{Introduction}
-There is much evidence for the presence of supermassive black holes (SMBHs) at the centres of massive galaxies, with the shadow of one such SMBH being directly imaged.
+There is much evidence for the presence of supermassive black holes (SMBHs) at the centres of massive galaxies, with the shadow of one such SMBH being directly imaged \citep{eht2019}.
 The dynamical interactions of two or more galaxies, during a galaxy merger, are thus also expected to induce an interaction of the residing SMBHs.
 The interaction of two SMBHs during the merger of two galaxies is well understood to occur in a three stage process \citep{begelman1980}.
-First, as the galaxies begin to dynamically interact and the SMBHs traverse through a dense stellar field, the overdensity of stars in the wakes of the SMBHs provides a restoring force which brings the SMBHs to the centre of the two interacting galaxies, generally torquing the SMBHs to highly radial orbits in the process.
-Second, after the SMBHs have formed a bound binary, iterative interactions with the surrounding stars ejects stars with high velocity, thus removing orbital energy and angular momentum from the SMBH binary orbit.
+First, as the galaxies begin to dynamically interact and the SMBHs traverse through a dense stellar field, the overdensity of stars in the wakes of the SMBHs provides a restoring force which brings the SMBHs to the centre of the two interacting galaxies \citep{chandrasekhar1943}, generally torquing the SMBHs to highly radial orbits in the process.
+Second, after the SMBHs have formed a bound binary, iterative interactions with the surrounding stars ejects stars with high velocity \citep[e.g.][]{hills1980,hills1983,quinlan1996}, thus removing orbital energy and angular momentum from the SMBH binary orbit.
+The removal of stars from the central regions by this slingshot mechanism produces a `core' \citep[e.g.][]{lauer1983,lauer1985,kormendy1984}, or depletion, in the luminosity profile of the remnant galaxy \citep[e.g.][]{begelman1980,hills1983,quinlan1996,rantala2018}, and can extend up to some kiloparsecs \citep{postman2012}.
 Finally, at very small, sub-parsec, separations the SMBH binary loses its remaining orbital energy and angular momentum through gravitational wave (GW) emission, driving the SMBH binary to coalescence \citep{peters1963,peters1964}.
 
-Due to the anisotropic emission of the final GW emission \drafting{during? following?} coalescence, the remnant SMBH recoils with some velocity, termed the kick velocity. 
+The final GW emission from the coalescing SMBH binary occurs anisotropically, and carries with it linear momentum from the system \citep[e.g.][]{gonzalez2007}.
+This results in the remnant SMBH recoiling with some velocity, termed the kick velocity. 
 The kick velocity is dependent on the mass ratio between the SMBHs prior to coalescence, and the magnitude and direction of angular momentum (spin) vector of each SMBH. 
-Previous studies \drafting{reference} indicate that while the majority of recoil kicks are of the order of a few hundred kilometres per second, in special configurations the SMBH may have an imparted kick velocity in excess of $4000\,\kmps$ \drafting{references}.
-Taking the central escape velocity of a typical early type galaxy (ETG) as $\sim2000\,\kmps$, there may be a non-negligible fraction of ETGs with a SMBH that has escaped the galaxy.
+As reported by \citet{campanelli2007}, symmetries in the masses or spins of the SMBHs suppresses the resulting recoil kick imparted to the coalesced SMBH.
+In particular, it is found that asymmetry in the spins of the SMBHs has a larger impact on the magnitude of the recoil kick than asymmetry in the masses \citep{campanelli2007}, with antialigned spins in the orbital plane producing the largest recoil kick velocities \citep{gonzalez2007b,tichy2007}.
+Numerical relativity studies indicate that while the majority of recoil kicks are of the order of a few hundred kilometres per second, in special configurations the SMBH may be imparted a kick velocity in excess of a $2000\,\kmps$, even up to $4000\,\kmps$ \citep{campanelli2007,gonzalez2007b,tichy2007}.
+Taking the central escape velocity of a typical early type galaxy (ETG) as $\sim2000\,\kmps$, there may be a non-negligible fraction of ETGs with a SMBH that has escaped the galaxy \citep{madau2004}, though not so many so as to introduce considerable scatter into the observed relation between SMBH mass and stellar velocity dispersion \citep{volonteri2007}.
 In cases where the kick velocity is less than the escape velocity of the inner regions of the host galaxy, the radial oscillations of the remnant SMBH as it returns to the centre of the merger remnant is expected to influence the surrounding stellar environment.
+\drafting{What sort of influences?
+References?}
+
+Observations by \citet{komossa2008} point to a recoiling SMBH \drafting{need to read\dots}
 
@@
 \subsection{Simulation Code}
-\drafting{Taken from eccentricity paper, leave here as a guide for a similar description about Ketju}
-We construct a number of idealised galaxy merger simulations, which we evolve with our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017} coupled with \gadget{}.
-The dynamics of
-SMBHs, and stars in a small region around them, are integrated with an algorithmically regularised integrator \citep{rantala2020}, whereas the dynamics of the remaining particles
-is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles. Together with hierarchical time integration this allows for
-symmetric interactions and manifest momentum conservation. \ketju{} also includes post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{blanchet2014}.
+\drafting{Taken from eccentricity paper, edited a little bit but need to ensure it is different enough.}
+To investigate the effect of the recoiling SMBH on the host galaxy, and potential observational signatures of this, we run a number of numerical simulations of a galaxy merger setting using our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017} coupled with \gadget{}.
+\ketju{} integrates the dynamics of SMBHs, and stars in a small region thrice the stellar softening length around them, to high accuracy using the algorithmically regularised integrator \textsc{mstar} \citep{rantala2020}.
+The dynamics of stellar particles beyond this small region of high integration accuracy, and all dark matter (DM) particles, is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles.
+Additionally, we use hierarchical time integration, which allows for symmetric interactions and manifest momentum conservation. 
+\ketju{} also includes post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{blanchet2014} and the fitting formula of \citet{zlochower2015} for recoil kick velocity, making \ketju{} a particularly well-suited code for investigating the consequence of SMBH recoil self-consistently in a galaxy merger environment.
 
@@
 \drafting{Rantala 2018 ICs}
-We model the merger of two gas-free elliptical galaxies using collisionless merger simulations.
-Our galaxy initial conditions are chosen to match the model IC-3 presented in \citet{rantala2019}. 
-The galaxy is thus represented as a multicomponent sphere, including a stellar component of total mass $M_\star\sim 1.38\times10^{11}\,\Msun$ embedded within a DM \drafting{ensure abbreviation stated} component of total mass $M_\mathrm{DM}=2.5\times10^{13}\,\Msun$, and at the centre a SMBH with mass $M_\bullet=2.93\times10^{9}\,\Msun$.
+We model the merger of two gas-free elliptical galaxies using gas-free merger simulations.
+Our galaxy initial conditions are chosen similarly to the model IC-3 presented in \citet{rantala2019}, however with different stellar density slope. 
+The galaxy is thus represented as an isotropic multicomponent sphere, and consists of a stellar component of total mass $M_\star\sim 1.38\times10^{11}\,\Msun$ embedded within a DM component of total mass $M_\mathrm{DM}=2.5\times10^{13}\,\Msun$.
+At the centre a SMBH with mass $M_\bullet=2.93\times10^{9}\,\Msun$ is placed with zero velocity.
 The stellar and DM components each follow a \citet{dehnen1993} profile with shape parameter $\gamma=1$, and scale radius $a_\star=3.9\,\mathrm{kpc}$ and $a_\mathrm{DM}=245\,\mathrm{kpc}$, respectively\footnote{A Dehnen profile with $\gamma=1$ is commonly referred to as a Hernquist profile.}.
-The density profile is given by:
+The density profile $\rho_i$ for a given component $i$ is given by:
 \begin{equation}\label{eq:dehnen}
-    \rho(r) = \frac{(3-\gamma)M}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
+    \rho_i(r) = \frac{(3-\gamma)M}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
 \end{equation}
-The mass of a stellar particle is set to $m_\star=5\times10^4\,\Msun$, and the mass of a DM particle is set to $m_\mathrm{DM}=5\times10^6\,\Msun$.
+We generate the ICs using the distribution function method following \citet{hilz2012}, where for each component (stellar and DM) the distribution function $f_i$ is computed using Eddington's formula \citep{binney2008} for each density profile $\rho_i$:
+\begin{equation}
+    f_i(\mathcal{E}) = \frac{1}{2\sqrt{2}\pi^2} \int_{\Phi_\mathrm{T}=0}^{\Phi_\mathrm{T}=\mathcal{E}} \dnv{2}{\rho_i}{\Phi_\mathrm{T}} \frac{\dd{\Phi_\mathrm{T}}}{\sqrt{\mathcal{E}-\Phi_\mathrm{T}}}.
+\end{equation}
+Here $\mathcal{E}$ is the relative energy, and $\Phi_\mathrm{T}$ is the total gravitational potential. 
+The distribution functions are then sampled with discrete particles, where the mass of a stellar particle is set to $m_\star=5\times10^4\,\Msun$, and the mass of a DM particle is set to $m_\mathrm{DM}=5\times10^6\,\Msun$.
+The radial velocity profiles of both stellar and DM particles is ergodic.
 
-\subsection{Merger Simulations}
+As we are interested in the evolution of the galaxy merger \textit{remnant} following SMBH coalescence, the initial conditions such that that at the time of SMBH coalescence (\autoref{ssec:mergers}) the galaxy remnant agrees with observational data.
+Specifically, we ensure that our remnant agrees with the half-light -- stellar mass data presented in \citet{sahu2020}, and lies within the $1\sigma$ predictive interval of the SMBH mass -- stellar velocity dispersion data given in \citet{vandenbosch2016}.
+
+\subsection{Merger Simulations}\label{ssec:mergers}
 We merge two independently Monte Carlo sampled galaxy ICs, creating an equal mass galaxy merger with a near-radial orbit. 

### 2024-04-18 papers/paper-corekick/paper-corekick.tex
- small addition for new core-kick relation

@@
 
-As we are interested in the evolution of the galaxy merger \textit{remnant} following SMBH coalescence, the initial conditions such that that at the time of SMBH coalescence (\autoref{ssec:mergers}) the galaxy remnant agrees with observational data.
+As we are interested in the evolution of the galaxy merger \textit{remnant} following SMBH coalescence, the initial conditions are constructed such that at the time of SMBH coalescence (\autoref{ssec:mergers}) the galaxy remnant agrees with observational data.
 Specifically, we ensure that our remnant agrees with the half-light -- stellar mass data presented in \citet{sahu2020}, and lies within the $1\sigma$ predictive interval of the SMBH mass -- stellar velocity dispersion data given in \citet{vandenbosch2016}.
@@
 To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+At the moment of the merger, the centre of mass (CoM) velocity of the SMBH binary is $v_{\mathrm{bin},\mathrm{CoM}}\sim[4.2, 1.1, 1.6]\,\kmps$ in the global coordinate frame.
 The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
@@
 \subsection{Predicted core size distribution}
-Following \citet{nasim2021b}, we fit a power law of the form
+To predict the distribution of core sizes, we require a mapping from the kick velocity to the core radius. 
+Previous work by \citet{nasim2021b} fit a power law of the form:
+\begin{equation}\label{eq:vkrb_nasim}
+    \frac{\rb}{r_{\mathrm{b},0}} = K (v')^\beta + 1
+\end{equation}
+where $v'=\vk/v_\mathrm{esc}$, and $K$ and $\beta$ are free parameters.
+We see immediately that \autoref{eq:vkrb_nasim} does not capture the peak in the data at \drafting{$600\kmps$}, and so opt for an alternative form.
+We fit a function that is the sum of a modified $t$-distribution with a sigmoid function, and takes the form:
 \begin{equation}\label{eq:vkrb}
-    \frac{\rb}{r_{\mathrm{b},0}} = K \left( \frac{\vk}{v_\mathrm{esc}} \right)^\beta + 1
+    \frac{\rb}{r_{\mathrm{b},0}} = \frac{A(v'; \mu,\nu,\sigma)+B(v'; \mu,\nu,\sigma)}{2\pi\sigma} - S(v';k,b,\mu) + c,
 \end{equation}
-to our simulations, which is thus valid for $0\leq\vk/\kmps\leq900$, or equivalently $0\leq\vk/v_\mathrm{esc}\leq 0.5$.
-We estimate the parameters $K$ and $\beta$ using HMC, and find median values of $K=3.70$ and $\beta=0.562$.
+where
+\begin{equation}
+    A(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu-2}{\nu}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu-2}{\nu}}} \right)^2 \right)^{-(\nu + 1)/2},
+\end{equation}
+and
+\begin{equation}
+    B(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu}{\nu-2}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu}{\nu-2}}} \right)^2 \right)^{-(\nu + 1)/2},
+\end{equation}
+and
+\begin{equation}
+    S(v';k,b,\mu) = \frac{k}{1 + \exp(-b (v'-\mu))}.
+\end{equation}
+Hence \autoref{eq:vkrb} takes six free parameters ($\mu$, $\nu$, $\sigma$, $k$, $b$, $c$), which we constrain using the median trend of the $vk/v_\mathrm{esc}$--$\rb$ relation in \autoref{fig:vkrb}, and plot with an orange line.
 
@@
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
-A.R., acknowledge the support
+P.H.J., acknowledge the support
 by the European Research Council via ERC Consolidator Grant KETJU (no. 818930) and the support of the Academy of Finland grant 339127.

### 2024-04-19 papers/paper-corekick/paper-corekick.tex
Update text to reflect fixed time analysis

@@
 \subsection{Initial Conditions}
-\drafting{Rantala 2018 ICs}
 We model the merger of two gas-free elliptical galaxies using gas-free merger simulations.
@@
 
-\drafting{
-    Main points to hammer home:
-    \begin{enumerate}
-        \item Core radius, break density, sersic index have a dependence on the kick velocity, whereas other parameters (e.g. effective radius) do not.
-        \item Increasing kick velocity induces a decrease in the overall $\langle h_4 \rangle$ profile
-        \item Explanation of h4 trends through orbit decompostion.
-    \end{enumerate}
-}
-
 \subsection{Infall time}
@@
 \subsection{Mass density profiles in three dimensions}
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{density_3d}
+    \caption{
+        Three dimensional stellar mass density, with the line colour corresponding to the kick velocity of the SMBH. 
+        All simulations with $\vk \geq 0.5v_\mathrm{esc}=900\,\kmps$ share the same colour, as there is little variation in the profiles for these kick velocities. 
+        A clear dichotomy is revealed in the mass profiles, whereby high kick velocoties induce a core, and low kick velocities do not, with a sudden transition at \drafting{$\vk=600\,\kmps$}.
+        }
+    \label{fig:density3d}
+\end{figure}
 Before considering the projected, 2D stellar mass density profiles of the merger remnants, we investigate the 3D stellar mass distribution.
 This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
-We observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $2\,\mathrm{kpc}$ than those mergers with $\vk>540\,\kmps$.
+In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $2\,\mathrm{kpc}$ than those mergers with $\vk>540\,\kmps$.
 All merger remnants have a consistent mass density distribution at $r>2\,\mathrm{kpc}$. 
+Additionally, for kick velocities $\vk<540\,\kmps$, we observe a gradient in decreasing central stellar density with increasing kick velocity.
+
 The reasons for the dichotomy are discussed further in \drafting{section}, however we continue our analysis using the projected stellar density (as an observer would see) with the knowledge that there is \textit{a} dependence of the removal of central stellar mass on the SMBH kick velocity.
-
-\drafting{INCLUDE FIGURE HERE}
 
@@
 
-In agreement with previous studies \citep{nasim2021b}, we find an increase in the core radius $\rb$ with kick velocity.
-\drafting{Need to check if others discuss this:} We find that the core radius saturates to a value of $\sim2$ times that of the merger remnant with no kick velocity by $\vk = 600\,\kmps$, corresponding to $0.3 v_\mathrm{esc}$.
-As shown in \autoref{fig:csparams}, the density at the core radius $\Sigma_\mathrm{b}$ decreases monotonically with kick velocity, consistent with the picture of the kicked SMBH removing more mass the more number of oscillations it is able to undergo before settling to Brownian motion limits.
-Interestingly, we observe a similar decrease in S\'ersic index $n$ with increasing kick velocity, although no correlation is seen between $n$ and $\Sigma_\mathrm{b}$ in the latent parameter corner plots (see for example \autoref{fig:csparams}).
+In agreement with previous studies \citep{nasim2021b} \drafting{others?}, we find an increase in the core radius $\rb$ with kick velocity, however only up to a given kick velocity.
+We find that unlike the results of \citet{nasim2021b}, where core radius monotonically increases with kick velocity, we observe a peak in core radius occurring at $480\,\kmps$, with the core radii measured from simulations with kick velocities in excess of $480\,\kmps$ plateauing at $\rb\sim1.75 r_{\mathrm{b},0}$.
+We caution however that the systems in the study of \citet{nasim2021b} differ from ours in both structure, and the time of the analysis.
+Whereas we use a fixed time following the SMBH binary coalescence for our analysis, \citet{nasim2021b} wait until either the ejected SMBH settles back to the CoM, or $2\,\mathrm{Gyr}$ has passed.
+\drafting{As discussed in section \textbf{Atte}, we find relaxation increasingly dominates our merger remnants with increasing time.}
+As a result, a direct comparison between this work and that of \citet{nasim2021b} cannot be made.
 
-\subsection{Predicted core size distribution}
-To predict the distribution of core sizes, we require a mapping from the kick velocity to the core radius. 
-Previous work by \citet{nasim2021b} fit a power law of the form:
-\begin{equation}\label{eq:vkrb_nasim}
-    \frac{\rb}{r_{\mathrm{b},0}} = K (v')^\beta + 1
-\end{equation}
-where $v'=\vk/v_\mathrm{esc}$, and $K$ and $\beta$ are free parameters.
-We see immediately that \autoref{eq:vkrb_nasim} does not capture the peak in the data at \drafting{$600\kmps$}, and so opt for an alternative form.
-We fit a function that is the sum of a modified $t$-distribution with a sigmoid function, and takes the form:
-\begin{equation}\label{eq:vkrb}
-    \frac{\rb}{r_{\mathrm{b},0}} = \frac{A(v'; \mu,\nu,\sigma)+B(v'; \mu,\nu,\sigma)}{2\pi\sigma} - S(v';k,b,\mu) + c,
-\end{equation}
-where
-\begin{equation}
-    A(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu-2}{\nu}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu-2}{\nu}}} \right)^2 \right)^{-(\nu + 1)/2},
-\end{equation}
-and
-\begin{equation}
-    B(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu}{\nu-2}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu}{\nu-2}}} \right)^2 \right)^{-(\nu + 1)/2},
-\end{equation}
-and
-\begin{equation}
-    S(v';k,b,\mu) = \frac{k}{1 + \exp(-b (v'-\mu))}.
-\end{equation}
-Hence \autoref{eq:vkrb} takes six free parameters ($\mu$, $\nu$, $\sigma$, $k$, $b$, $c$), which we constrain using the median trend of the $vk/v_\mathrm{esc}$--$\rb$ relation in \autoref{fig:vkrb}, and plot with an orange line.
-
-Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig. 15, right panel).
-The dimensionless spin parameter $\alpha_\bullet$ for the dry mergers follows a beta distribution, namely:
-\begin{equation}
-    P_\mathrm{Zlochower}(\alpha_\bullet) \propto (1-\alpha_\bullet)^{4.66884-1} \alpha_\bullet^{10.5868-1}.
-\end{equation}
-The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$ \drafting{check this}.
-Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}.
-We find that the mode of the forward-folded core radius distribution is $1.64\,\mathrm{kpc}$ ($3.12\,r_{\mathrm{b},0}$).
-Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, most massive elliptical galaxies should have a non-negligible core \drafting{add references}.
+Complimentary to the peak in core radius at $480\,\kmps$, we observe a dip in the projected density at the core radius $\Sigma_\mathrm{b}$ at this same kick velocity in \autoref{fig:csparams}.
+The effective radius $\Reff$ and core density slope $\gamma$ are relatively robust to the SMBH recoil velocity.
 
@@
     \caption{
-        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{b,0}$, as a function of kick velocity $\vk$.
+        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{\mathrm{b},0}$, as a function of kick velocity $\vk$.
         The core size distributions are shown as box plots, with the median core size indicated by the central mark.
@@
 \end{figure}
+
+\subsection{Predicted core size distribution}
+\drafting{Would a Gaussian process fit be better motivated here?}
+To predict the distribution of core sizes, we require a mapping from the kick velocity to the core radius. 
+Previous work by \citet{nasim2021b} fit a power law of the form:
+\begin{equation}\label{eq:vkrb_nasim}
+    \frac{\rb}{r_{\mathrm{b},0}} = K v'^\beta + 1
+\end{equation}
+where $v'=\vk/v_\mathrm{esc}$, and $K$ and $\beta$ are free parameters.
+We see immediately that \autoref{eq:vkrb_nasim} does not capture the peak in the data at $480\kmps$, and so opt for an alternative form.
+We fit a function that is the sum of a modified $t$-distribution with a sigmoid function, and takes the form:
+\begin{equation}\label{eq:vkrb}
+    \frac{\rb}{r_{\mathrm{b},0}} = \frac{A(v'; \mu,\nu,\sigma)+B(v'; \mu,\nu,\sigma)}{2\pi\sigma} - S(v';k,b,\mu) + c,
+\end{equation}
+where
+\begin{equation}
+    A(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu-2}{\nu}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu-2}{\nu}}} \right)^2 \right)^{-(\nu + 1)/2},
+\end{equation}
+and
+\begin{equation}
+    B(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu}{\nu-2}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu}{\nu-2}}} \right)^2 \right)^{-(\nu + 1)/2},
+\end{equation}
+and
+\begin{equation}
+    S(v';k,b,\mu) = \frac{k}{1 + \exp(-b (v'-\mu))}.
+\end{equation}
+Hence \autoref{eq:vkrb} takes six free parameters ($\mu$, $\nu$, $\sigma$, $k$, $b$, $c$), which we constrain using the median trend of the $vk/v_\mathrm{esc}$--$\rb$ relation in \autoref{fig:vkrb}, and assume a normally-distributed scatter $\tau$.
+The fit is plotted with an orange line in \autoref{fig:vkrb}, and the values of the fitted parameters are given in \autoref{tab:vkrb}.
+
+\begin{table}
+    \caption{Values of the fitted parameters in the core radius -- kick velocity relation in \autoref{eq:vkrb}.}
+    \label{tab:vkrb}
+    \begin{tabular}{lccc}
+        \hline
+        Parameter & Median & IQR & Constraints \\
+        \hline
+        $\mu$ & 0.182 & 0.0427 & $\geq 0$ \\
+        $\nu$ & 21.5 & 11.9 & $\geq 0$ \\
+        $\sigma$ & 0.601 & 0.298 & $\geq 0$ \\
+        $k$ & 0.894 & 0.189 & $\geq 0$ \\
+        $b$ & -34.5 & 30.9 & $\leq 0$ \\
+        $c$ & 1.64 & 0.0661 & $\geq 0$ \\
+        $\tau$ & 0.170 & 0.0336 & $\geq 0$ \\
+        \hline
+    \end{tabular}
+\end{table}
+
+Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig. 15, right panel).
+The dimensionless spin parameter $\alpha_\bullet$ for the dry mergers follows a beta distribution, namely:
+\begin{equation}
+    P_\mathrm{Zlochower}(\alpha_\bullet) \propto (1-\alpha_\bullet)^{4.66884-1} \alpha_\bullet^{10.5868-1}.
+\end{equation}
+The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$ \drafting{check this}.
+Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}.
+We find that the mode of the forward-folded core radius distribution is $1.64\,\mathrm{kpc}$ ($3.12\,r_{\mathrm{b},0}$).
+Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, most massive elliptical galaxies should have a non-negligible core \drafting{add references}.
 
@@
 The predicted kick velocities range from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
-\drafting{check this}}
+\drafting{check this}
+    The asymmetry at $\rb/r_{\mathrm{b},0}\simeq 1$ is due to the low kick velocities producing a modest core, as seen in \autoref{fig:vkrb}.}
     \label{fig:rb_pdf}
@@
 
-We observe a trend of the Voronoi bins with $h_4<0$ extending to larger radii for remnants with larger recoil kicks than for those remnants with smaller recoil kicks.
-To quantify this recoil velocity dependence, we calculate the cumulative $h_4$ parameter:
-\begin{equation}
+For the parallel projections, we do not observe a systematic variation in $h_4$ with increasing kick velocity.
+To quantify the radial dependence of $h_4$, we calculate the cumulative $h_4$ parameter:
+\begin{equation}\label{eq:h4}
     \langle h_4 \rangle = \sum_i^N h_4(R_i)
 \end{equation}
-We observe an overall shift to more negative values of $h_4$ for kick velocities up until $\vk\sim400\,\kmps$ \drafting{check this, currently just read off plot}, above which the radial $h_4$ profile is not sensitive to increasing kick velocity.
-Additionally, all $h_4$ profiles are radially decreasing, obtaining a local minimum at $R\simeq 2\,\mathrm{kpc}$ ($R/\Reff\simeq 0.2$), and a modestly increasing for $R \gtrsim 2\,\mathrm{kpc}$.
-
-\drafting{Jens: maybe you have some thoughts\dots Why is this 
-\begin{enumerate}
-    \item saturating beyond some point,
-    \item displaying a dip at 0.2 Re??
-\end{enumerate}
-}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_v0000}
-    \caption{
-        Mock integral field unit kinematic maps of the $\vk=0\,\kmps$ remnant, plotted out to $0.5\,\Reff$. 
-        Note the small magnitude of the unordered rotation $V$, and the centrally-concentrated velocity dispersion $\sigma$.
-        The majority of the merger remnant within this aperture has $h_4>0$, with a small concentration of $h_4<0$ in an annulus about the central regions.
-    }
-    \label{fig:IFU0}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_v0900}
-    \caption{
-        Mock integral field unit kinematic maps of the $\vk=900\,\kmps$ remnant, again plotted out to $0.5\,\Reff$, with the same colour scaling as \autoref{fig:IFU0}. 
-        Contrasting to the $\vk=0\,\kmps$ case in \autoref{fig:IFU0}, this IFU map has a much greater extent of $h_4<0$ that also reaches more extreme values of $h_4$. 
-    }
-    \label{fig:IFU900}
-\end{figure*}
+As shown in \autoref{fig:h4}, for the parallel projections, we do not observe a systematic variation in $h_4$ with increasing kick velocity.
+Conversely, for the orthogonal projections, we observe a dichotomy similar to that seen in \autoref{fig:density3d}, with the bifurcation occurring at a similar kick velocity, here $600\,\kmps$. 
+Within the two partitions of the bifurcation, there is no apparent dependence on kick velocity.
 
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{h4}
+    \includegraphics[width=0.5\textwidth]{h4}
     \caption{
-        Plot of the flux-weighted value of $h_4$ as a function of mean bin distance $R$for each merger remnant simulated.
-        The lines are coloured by the kick velocity. 
-        Immediately apparent is the decrease in $\langle h_4 \rangle$ with increasing $\vk$.
-        All merger remnants have a global minimum in $\langle h_4 \rangle$ at $R\simeq 2.0$, corresponding to $R/\Reff\simeq0.2$.
+        Cumulative $h_4$ parameter, as defined in \autoref{eq:h4}, for the parallel (top panel) and orthogonal (bottom panel) projections.
+        Similar to \autoref{fig:density3d}, simulations with $\vk\geq 900\,\kmps$ have a consistent colour.
+        Lines terminating with a circle point indicate simulations with $\vk/\kmps < 600$, those with square points indicate simulations with $600 \leq \vk/\kmps < 1200$, those with triangle points $1200 \leq \vk/\kmps < 1800$, and those with diamond points $\vk/\kmps > 1800$.
+        For the parallel case, no systematic trend between $h_4$ and $\vk$ appears to exist.
+        Conversely, for the orthogonal projection, a dichotomy similar to the bifurcation seen in \autoref{fig:density3d} is observed. 
     }
@@
         Stellar particles are assigned to one of seven different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 30.0$.
-        At radii greater than $1\,\mathrm{kpc}$, there is very little to distinguish between the different kick velocities for boxlet, $z$-tube, Keplerian, irregular, and unclassified orbits.
-        Conversely, for radii $1 \leq R/\mathrm{kpc} \leq 10$, there is a clear gradient for $\pi$-box and $x$-tube families.
-        For higher kick velocities, there is a tendency to a higher fraction of $\pi$-box orbits compared to lower kick velocities.
-        An inverse trend is visible for $x$-tube orbits. 
+        Note the colour scheme is consistent with that in \autoref{fig:density3d}, with kick velocities exceeding $900\,\kmps$ represented using the same colour.
     }
@@
 
-To understand the kinematic maps, we perform an orbit analysis of the merger remnants in our sample following \citet{frigo2021}, and is briefly described here.
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{orbits2}
+    \caption{
+        Same as \autoref{fig:orbits}, but each merger remnant is shown in its own panel, with the different orbital families depicted by colour, to better demonstrate the radial fraction of orbits for a given merger.
+    }
+    \label{fig:orbits2}
+\end{figure*}
 
-\drafting{Discuss how the orbit analysis is done, in particular the different orbital families}.
+To understand the kinematic maps, we perform an orbit analysis of the merger remnants in our sample following \citet{frigo2021}, which is briefly described here.
+
 First, the merger remnant is rotated so that the $z$-axis coincides with the minor axis of the reduced inertia tensor as measured using the top 50\% most bound stellar particles to the stellar centre of mass.
 We cannot centre on the recoiling SMBH for this analysis, as for high kick velocities the SMBH is significantly displaced from the centre.
+Thus, we use the shrinking sphere method on stellar particles to determine the centre of mass of the system.
 The stellar potential of the merger remnant is fit using a self consistent field (SCF) potential \citep{hernquist1992}, using the \citet{hernquist1990} profile as a zeroth-order basis and limiting the expansion to $n_\mathrm{max}=18$ and $l_\mathrm{max}=7$ following \citet{frigo2021}.
@@
 
+Three distinct trends are made apparent in \autoref{fig:orbits}.
+The first and most important, is the complete absence of Keplerian orbits for all simulations with $\vk>600\,\kmps$.
+Conversely, all simulations with $\vk<600\,\kmps$ (except for $\vk=540\,\kmps$) have more than $\sim25\%$ of the orbits in the innermost radial bin being identified as Keplerian. 
+The second and third trend are complimentary, and are an increase in the fraction of $\pi$-box orbits with increasing kick velocity, and a similar decrease in the fraction of $x$-tube orbits with increasing kick velocity.
+Finally, there is a weak tendency for simulations with $\vk<900\,\kmps$ to have slightly higher fractions of boxlet orbits at small radii, and higher fractions of $z$-tube orbits, than those simulations with $\vk>900\,\kmps$.
+
+% DISCUSSION
+\section{Discussion} \label{sec:discussion}
+\drafting{
+    Ok, we're here. Let's recap:
+    We see a bifurcation in simulations, where this seems to happen at $\vk=600\,\kmps$.
+    This is manifested in:
+    \begin{itemize}
+        \item The 3D density
+        \item the transition from increasing to decreasing/constant core radius
+        \item the orthogonal projection of the IFU h4 quantity
+        \item The presence of Keplerian orbits. 
+    \end{itemize}
+    What is going on? 
+    I believe the driver of all this is the orbits. The other items in the above list are ways we can \textit{observe} these differences in orbits. 
+    I suspect, and need to think about how we determine this, that below $600\,\kmps$, there is insufficient energy released from the recoiling SMBH to significantly alter the orbital structure, i.e.
+little core heating.
+    Conversely, above $600\,\kmps$, the SMBH is so rapidly removed from the centre (and in this analysis doesn't return yet) that the system is unable to respond to the rapid change in potential quick enough.
+    $600\,\kmps$ therefore represents a sweetspot, where we have a lot of energy that is then able to be given to the surrounding stars in a reasonable time.
+
+    Some incongruencies:
+    \begin{itemize}
+        \item In 3D density, there is a smooth decrease with increasing kick velocity, until it saturates. How do we then explain a peak in the core radius in the projected density maps?
+    \end{itemize}
+}
+
+\drafting{Need to update, leave this here for now
 As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box and boxlet orbits, and a lower fraction of $x$-tube orbits, and to some extent Keplerian orbits, than their lower kick velocity counterparts.
@@
 This can be explained by the derivative of the orbital fraction distribution being consistent across all kick velocities for $x$-tube and boxlet orbits out to $\sim3\,\mathrm{kpc}$, as shown in \autoref{fig:orbits2}.
-\drafting{Thorsten: I would really appreciate your thoughts on this section, and its relation to the kinematic maps. }
-
-% DISCUSSION
-\section{Discussion}
-\label{sec:discussion}
+}
 
@@
 \section{Triaxiality}\label{sec:app_triax}
+\drafting{Need to update}
 We present the triaxiality of all sixteen merger remnants in the analysis at the time when the SMBH has settled.
@@
         Note that the $\vk=0\,\kmps$ case is equivalent to the pre-merger remnant. 
+        \drafting{Update colour bar to match other plots}
     }
@@
     \centering
-    \includegraphics[width=\textwidth]{corner_latent_900}
+    \includegraphics[width=\textwidth]{corner_latent_600}
     \caption{
-        Corner plot of the latent parameters in the core-S\'ersic model for the $\vk=900\,\kmps$ instance.
-        Contours indicate, from yellow to dark blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs), with sample draws beyond the 99\% HDI shown as individual points. 
+        Corner plot of the latent parameters in the core-S\'ersic model for the $\vk=600\,\kmps$ instance.
+        Contours indicate, from dark to light blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs), with sample draws beyond the 99\% HDI shown as individual points. 
         All distributions are unimodal with little cross-correlation between the latent variables.
@@
 
-\section{Additional orbital plots}
+\section{Specific IFU maps}
+\drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, $\vk=600\,\kmps$, and $\vk=1800\,\kmps$.}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{orbits2}
+    \includegraphics[width=\textwidth]{IFU_para_v0000}
     \caption{
-        Same as \autoref{fig:orbits}, but each merger remnant is shown in its own panel, with the different orbital families depicted by colour, to better demonstrate the radial fraction of orbits for a given merger.
+        Mock integral field unit kinematic maps of the $\vk=0\,\kmps$ remnant in the parallel projection, plotted out to $0.25\,\Reff$ in a circular aperture. 
+        The small ordered rotation $V$ varies little in magnitude ($20\,\kmps$), and the velocity dispersion $\sigma$ is centrally-concentrated, peaking at around $\sim260\,\kmps$.
+        The central regions of the merger remnant have a $h_4$ value that is predominantly $0\,\kmps$, and positive $h_4$ at larger radial extents.
     }
-    \label{fig:orbits2}
+    \label{fig:IFU0p}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{IFU_ortho_v0000}
+    \caption{
+        The same as \autoref{fig:IFU0p}, but in the orthogonal projection. 
+        Note the colour scaling is consistent.
+        Qualitatively, there is little difference between the orthogonal and parallel projections for the $\vk=0\,\kmps$ case.
+    }
+    \label{fig:IFU0o}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{IFU_para_v0600}
+    \caption{
+        Mock integral field unit kinematic maps of the $\vk=600\,\kmps$ remnant in the parallel projection, plotted out to $0.25\,\Reff$ in a circular aperture. 
+        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
+        Most striking is the concentration of negative ordered velocity $V$ slightly offset from the remnant centre.
+        The other kinematic moments display a similar spatial offset.
+    }
+    \label{fig:IFU600p}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{IFU_ortho_v0600}
+    \caption{
+        The same as \autoref{fig:IFU600p}, but in the orthogonal projection. 
+        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
+        Compared to the parallel case, the ordered velocity $V$ and $h_3$ moments appear consistent with the $\vk=0,\kmps$ case, however a spatial offset from the centre persists in $\sigma$ and $h_4$.
+    }
+    \label{fig:IFU600o}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{IFU_para_v1800}
+    \caption{
+        Mock integral field unit kinematic maps of the $\vk=1800\,\kmps$ remnant in the parallel projection, plotted out to $0.25\,\Reff$ in a circular aperture. 
+        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
+        This projection displays clear ordered rotation, however there is no clear peak in velocity dispersion.
+        The other kinematic moments are quantitatively similar to the $\vk=0,\kmps$ case. 
+    }
+    \label{fig:IFU1800p}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=\textwidth]{IFU_ortho_v1800}
+    \caption{
+        The same as \autoref{fig:IFU1800p}, but in the orthogonal projection. 
+        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
+        Aside from the lack of ordered rotation, there is little difference between the orthogonal and parallel projections at this kick velocity. 
+    }
+    \label{fig:IFU180o}
 \end{figure*}

### 2024-04-25 papers/paper-corekick/paper-corekick.tex
Text updates

@@
 The final GW emission from the coalescing SMBH binary occurs anisotropically, and carries with it linear momentum from the system \citep[e.g.][]{gonzalez2007}.
-This results in the remnant SMBH recoiling with some velocity, termed the kick velocity. 
+This results in the remnant SMBH recoiling with some velocity directed opposite to the linear momentum of the GW emission \citep{bekenstein1973}, and is termed the kick velocity. 
 The kick velocity is dependent on the mass ratio between the SMBHs prior to coalescence, and the magnitude and direction of angular momentum (spin) vector of each SMBH. 
@@
 
-Observations by \citet{komossa2008} point to a recoiling SMBH \drafting{need to read\dots}
+Numerous observational studies have also hinted at the existence of recoiling SMBHs, generally through either an offset from the centre of the host galaxy in velocity, in position, or both.
+One of the earliest observations that pointed to a recoiling SMBH was of the quasar SDSSJ0927+2943, reported by \citet{komossa2008}, where an observed velocity offset suggested a recoil kick in excess of $2600\,\kmps$.
+Another recoiling SMBH candidate was identified by an offset in velocity by \citet{steinhardt2012}, with an estimated recoil velocity greater than $4000\,\kmps$.
+Other observations by \citet{comerford2014} and \citet{pesce2018} have also reported potential recoiling SMBHs, albeit at much lower kick velocities (some few hundreds of $\kmps$).
+\drafting{In \citet{pesce2018}, they talk about a peculiar velocity of 70ish km/s, whereas in later work, they stated velocity is 4000 km/s.
+Not sure which is correct, that difference seems suspicious\dots}
+Recoiling SMBHs may also be identified through a spatial offset from the host centre, where the offset can range from some parsecs \citep{batcheldor2010,lena2014,barrows2016} to over a kiloparsec \citep{koss2014,skipper2018}.
+Finally, some candidate recoiling SMBHs have both spatial and velocity offsets, such as CXO J101527.2+625911 \citet{kim2017} and 2MASX J00423991+3017515 \citet{hogg2021}.
 
@@
 Additionally, for kick velocities $\vk<540\,\kmps$, we observe a gradient in decreasing central stellar density with increasing kick velocity.
+We check the time dependence of this dichotomy by constructing the 3D stellar mass distributions at additional times $t=0.056\,\mathrm{Gyr}$, $t=0.094\,\mathrm{Gyr}$, and $t=0.12\,\mathrm{Gyr}$, and find the dichotomy persists, with the same kick velocity split.
 
@@
 
-\subsection{Projected mass density profiles}
-To facilitate comparisons with observations \drafting{some references}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projection of each simulated merger remnants:
+\subsection{Projected mass density profiles}\label{ssec:projdens}
+To facilitate comparisons with observations \citep[e.g.][]{graham2003,rusli2013}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projection of each simulated merger remnants:
 \begin{equation}\label{eq:cs}
@@
 
-In agreement with previous studies \citep{nasim2021b} \drafting{others?}, we find an increase in the core radius $\rb$ with kick velocity, however only up to a given kick velocity.
-We find that unlike the results of \citet{nasim2021b}, where core radius monotonically increases with kick velocity, we observe a peak in core radius occurring at $480\,\kmps$, with the core radii measured from simulations with kick velocities in excess of $480\,\kmps$ plateauing at $\rb\sim1.75 r_{\mathrm{b},0}$.
+Our results indicate that a kick non-zero kick velocity always serves to enlarge the core region relative to the core formed during the SMBH binary scouring phase.
+In agreement with previous studies \citep{gualandris2008,nasim2021b}, we find an increase in the core radius $\rb$ with kick velocity, however only up to a given kick velocity, in agreement with the results of \citet{boylan2004}.
+We find that unlike the results of \citet{nasim2021b}, where core radius monotonically increases with kick velocity, we observe a peak in core radius occurring at $\vk=480\,\kmps=0.27\,v_\mathrm{esc}$, with the core radii measured from simulations with kick velocities in excess of $480\,\kmps$ plateauing at $\rb\sim1.7 r_{\mathrm{b},0}$.
 We caution however that the systems in the study of \citet{nasim2021b} differ from ours in both structure, and the time of the analysis.
@@
 We see immediately that \autoref{eq:vkrb_nasim} does not capture the peak in the data at $480\kmps$, and so opt for an alternative form.
-We fit a function that is the sum of a modified $t$-distribution with a sigmoid function, and takes the form:
+We fit a function that is the sum of a modified Student's $t$-distribution with a sigmoid function, and takes the form:
 \begin{equation}\label{eq:vkrb}
@@
 Hence \autoref{eq:vkrb} takes six free parameters ($\mu$, $\nu$, $\sigma$, $k$, $b$, $c$), which we constrain using the median trend of the $vk/v_\mathrm{esc}$--$\rb$ relation in \autoref{fig:vkrb}, and assume a normally-distributed scatter $\tau$.
-The fit is plotted with an orange line in \autoref{fig:vkrb}, and the values of the fitted parameters are given in \autoref{tab:vkrb}.
+The fit is found using the same Bayesian methods as in \autoref{ssec:projdens} (including posterior checking), and is plotted with an orange line in \autoref{fig:vkrb}, with the values of the fitted parameters given in \autoref{tab:vkrb}.
 
@@
 \end{equation}
-The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$ \drafting{check this}.
-Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}.
-We find that the mode of the forward-folded core radius distribution is $1.64\,\mathrm{kpc}$ ($3.12\,r_{\mathrm{b},0}$).
-Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, most massive elliptical galaxies should have a non-negligible core \drafting{add references}.
+The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
+Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}, whilst simultaneously accounting for uncertainty in the fit described by \autoref{eq:vkrb}.
+We find that the mode of the forward-folded core radius distribution is $1.07\,\mathrm{kpc}$ ($1.69\,r_{\mathrm{b},0}$).
+Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, a large fraction massive elliptical galaxies should have a non-negligible core \citep{ferrarese1994,savorgnan2016,sahu2019}.
 
@@
 The predicted kick velocities range from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
-\drafting{check this}
     The asymmetry at $\rb/r_{\mathrm{b},0}\simeq 1$ is due to the low kick velocities producing a modest core, as seen in \autoref{fig:vkrb}.}

### 2024-05-02 papers/paper-corekick/paper-corekick.tex
Continue discussion section

@@
 This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
-In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $2\,\mathrm{kpc}$ than those mergers with $\vk>540\,\kmps$.
-All merger remnants have a consistent mass density distribution at $r>2\,\mathrm{kpc}$. 
+For the $\vk=0\,\kmps$ case, we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$, and is insensitive to the time of measurement.
+In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $3\,\mathrm{kpc}$ ($2r_\mathrm{infl}$) than those mergers with $\vk>540\,\kmps$.
+All merger remnants have a consistent mass density distribution at $r>3\,\mathrm{kpc}$. 
 Additionally, for kick velocities $\vk<540\,\kmps$, we observe a gradient in decreasing central stellar density with increasing kick velocity.
@@
 
-\subsection{Orbit analysis}
+\subsection{Orbit analysis}\label{ssec:orbits}
 
@@
         boxlet & Resonant motion with angular momentum \\
-        Keplerian & Typical orbit in a point-mass dominated spherically-symmetric potential \\
+        rosette & Typical orbit in a point-mass dominated spherically-symmetric potential \\
         irregular & No integrals of motion \\
@@
 Three distinct trends are made apparent in \autoref{fig:orbits}.
-The first and most important, is the complete absence of Keplerian orbits for all simulations with $\vk>600\,\kmps$.
-Conversely, all simulations with $\vk<600\,\kmps$ (except for $\vk=540\,\kmps$) have more than $\sim25\%$ of the orbits in the innermost radial bin being identified as Keplerian. 
+The first and most important, is the complete absence of rosette orbits for all simulations with $\vk>600\,\kmps$.
+Conversely, all simulations with $\vk<600\,\kmps$ (except for $\vk=540\,\kmps$) have more than $\sim25\%$ of the orbits in the innermost radial bin being identified as rosette. 
 The second and third trend are complimentary, and are an increase in the fraction of $\pi$-box orbits with increasing kick velocity, and a similar decrease in the fraction of $x$-tube orbits with increasing kick velocity.
@@
         \item the orthogonal projection of the IFU h4 quantity
-        \item The presence of Keplerian orbits. 
+        \item The presence of rosette orbits. 
     \end{itemize}
@@
 
+Using 32 simulations of gas-free post-merger elliptical galaxy remnants, we have investigated the effect of a recoiling SMBH on the stellar environment. 
+Most importantly, we report a bifurcation in the simulations at $\vk \simeq 0.3 v_\mathrm{esc}$.
+This bifurcation manifests in the three-dimensional stellar density profiles and the orthogonal projection of the kinematic $h_4$ moment.
+Additionally, the peak projected core radius $\rb$ occurs approximately at the bifurcation, where for $\vk \lesssim 0.3 v_\mathrm{esc}$ the core radius $\rb$ increases with $\vk$, and then decreases after the peak to reach a plateau.
+
+These observations, whilst helpful in understanding the presence of a general trend, are not the cause of the bifurcation.
+We argue that the bifurcation is a result of the changing orbital structure in the inner regions of the merger remnant due to the displaced SMBH. 
+As discussed in \autoref{ssec:orbits}, all simulations in which the velocity of the recoiling SMBH exceeds $\vk=600\,\kmps$ do not contain any rosette orbits. 
+Rosette orbits are bounded by distinct radial bounds \citep[pericentre and apocentre, e.g.][]{binney2008}, and produce a distinctive ring pattern in the $h_4$ kinematic moment, whereby a torus of negative $h_4$ separates an inner and outer region of positive $h_4$, as seen in fig.
+14 of \citet{frigo2021}.
+We see evidence of a faint ring structure of negative $h_4$ in both the parallel and orthogonal IFU projections of the $\vk=0\,\kmps$ simulation (\autoref{fig:IFU0p} and \autoref{fig:IFU0o}), which agrees with the dominance of rosette orbits at the inner radii of the galaxy merger remnant (\autoref{fig:orbits}).
+Additionally, as rosette orbits are constrained to distinct radial bounds, these orbits contribute a relatively constant mass to the inner stellar density profile. 
+For all simulations with $\vk\geq660\,\kmps$, the majority of stellar particles have $\pi$-box orbits, i.e.
+those orbits with no net angular momentum, and thus move radially and can come arbitrarily close to the galactic potential centre \citep{frigo2021}.
+These orbits also display no resonant motion, unlike their boxlet counterparts; boxlet orbits are the second most dominant orbital family at small radii for $\vk \geq 660\,\kmps$, and are also radially-biased.
+We can gain insight into the probability of finding a stellar particle on a radial orbit by approximating the motion as a simple harmonic oscillator normalised by some dispersion measure $\xi$. 
+The radial motion is described by a Gaussian function of $r/\xi$ about the centre of the potential, and thus the probability is proportional to $\int_r ^{r+\dd{r}} |\exp(-(r/\xi)^2)|^2\dd{r}$.
+The integral of a Gaussian function results in the error function, and thus the probability of finding a stellar particle in the interval $r+\dd{r}$ is near zero for small $r/\xi$, and rapidly grows as $r/\xi \rightarrow 1$, before plateauing as $r/\xi \rightarrow \infty$.
+From this simple argument, we expect a deficit of stellar mass in the central regions of the galaxy merger remnant when the majority of the stellar mass belongs to a radially-biased orbit family ($\pi$-box or boxlet).
+We see the manifestation of an increase in radially-biased orbits in the three dimensional stellar density profile, where the bifurcation in the density profiles occur at the same kick velocity as when $\pi$-box and boxlet orbits start to dominate the inner orbital structure.
+
 \drafting{Need to update, leave this here for now
-As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box and boxlet orbits, and a lower fraction of $x$-tube orbits, and to some extent Keplerian orbits, than their lower kick velocity counterparts.
+As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box and boxlet orbits, and a lower fraction of $x$-tube orbits, and to some extent rosette orbits, than their lower kick velocity counterparts.
 The other two orbital families are generally consistent between all merger remnants. 
@@
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
-P.H.J., acknowledge the support
-by the European Research Council via ERC Consolidator Grant KETJU (no. 818930) and the support of the Academy of Finland grant 339127.
+P.H.J., acknowledge the support by the European Research Council via ERC Consolidator Grant KETJU (no. 818930) and the support of the Academy of Finland grant 339127.
 

### 2024-05-06 papers/paper-corekick/paper-corekick.tex
Edit discussion

@@
 }
+
+\drafting{
+    Also maybe we can discuss the number of bound particles to the BH here. I've already looked at this, but we can talk about combining that information with the apocentre plots. 
+    It seems that increasing $\vk$ decreases the amount of bound mass, in agreement with previous studies, but also for each $\vk$, the maximum amount of stellar mass bound to the BH is maximal at apocentre.
+}
 \begin{figure}
@@
 
-\subsection{Integral field unit kinematics}
+\subsection{Integral field unit kinematics}\label{ssec:ifu}
 We create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the BH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the BH kick axis (`orthogonal').
@@
 A recent comprehensive study of LOS velocity distribution fitting applied to elliptical galaxies can be found in \citet{mehrgan2023}.
+Relative to a standard Gaussian (which has $h_4=0$), a positive value of $h_4$ corresponds to extended high velocity tails in the LOS velocity distribution, whereas a negative value of $h_4$ corresponds to weaker tails in the LOS velocity distribution.
 
-For the parallel projections, we do not observe a systematic variation in $h_4$ with increasing kick velocity.
 To quantify the radial dependence of $h_4$, we calculate the cumulative $h_4$ parameter:
@@
 \end{equation}
-As shown in \autoref{fig:h4}, for the parallel projections, we do not observe a systematic variation in $h_4$ with increasing kick velocity.
-Conversely, for the orthogonal projections, we observe a dichotomy similar to that seen in \autoref{fig:density3d}, with the bifurcation occurring at a similar kick velocity, here $600\,\kmps$. 
+As shown in \autoref{fig:h4}, for the parallel projections, simulations with $\vk \geq 600\,\kmps$ tend to positive values of $\langle h_4 \rangle$, asymptotically reaching $\langle h_4 \rangle \sim 1$.
+Conversely, when the recoiling SMBH has a kick velocity $\vk < 600\,\kmps$, the final value of $\langle h_4 \rangle$ is evenly distributed about $\langle h_4 \rangle = 0$: there is no systematic preference to heaver or weaker tails in the LOS velocity distribution.
+
+Conversely, for the orthogonal projections, we observe a dichotomy similar to that seen in \autoref{fig:density3d}, with the bifurcation occurring at a similar kick velocity, here $600\,\kmps$; both partitions tend to a positive value of $\langle h_4 \rangle$.
 Within the two partitions of the bifurcation, there is no apparent dependence on kick velocity.
+The gradient between the two sets differs most noticeably in the radial range $2  \leq R/\mathrm{kpc} < 3$.
+This can be understood as those recoiling SMBHs with $\vk \geq 600\,\kmps$ having a larger $h_4$ magnitude per Voronoi bin than the $\vk < 600\,\kmps$ simulations.
 
@@
 As discussed in \autoref{ssec:orbits}, all simulations in which the velocity of the recoiling SMBH exceeds $\vk=600\,\kmps$ do not contain any rosette orbits. 
-Rosette orbits are bounded by distinct radial bounds \citep[pericentre and apocentre, e.g.][]{binney2008}, and produce a distinctive ring pattern in the $h_4$ kinematic moment, whereby a torus of negative $h_4$ separates an inner and outer region of positive $h_4$, as seen in fig.
+Rosette orbits are bounded by distinct radial limits \citep[pericentre and apocentre, e.g.][]{binney2008}, and produce a distinctive ring pattern in the $h_4$ kinematic moment, whereby a torus of negative $h_4$ separates an inner and outer region of positive $h_4$, as seen in fig.
 14 of \citet{frigo2021}.
@@
 These orbits also display no resonant motion, unlike their boxlet counterparts; boxlet orbits are the second most dominant orbital family at small radii for $\vk \geq 660\,\kmps$, and are also radially-biased.
-We can gain insight into the probability of finding a stellar particle on a radial orbit by approximating the motion as a simple harmonic oscillator normalised by some dispersion measure $\xi$. 
-The radial motion is described by a Gaussian function of $r/\xi$ about the centre of the potential, and thus the probability is proportional to $\int_r ^{r+\dd{r}} |\exp(-(r/\xi)^2)|^2\dd{r}$.
-The integral of a Gaussian function results in the error function, and thus the probability of finding a stellar particle in the interval $r+\dd{r}$ is near zero for small $r/\xi$, and rapidly grows as $r/\xi \rightarrow 1$, before plateauing as $r/\xi \rightarrow \infty$.
+We can gain insight into the probability of finding a radially-moving stellar particle at a position $r+\dd{r}$ by approximating the motion as a classical simple harmonic oscillator. 
+The differential probability $\dd{P}$ that the particle is within the interval $r+\dd{r}$ is proportional to $1/(\xi^2-r^2)$, where $\xi$ is the maximum radial displacement.
+Hence, the likelihood that the particle is at $r\ll\xi$ is much less compared to the likelihood that the particle is at $r\sim\xi$.
 From this simple argument, we expect a deficit of stellar mass in the central regions of the galaxy merger remnant when the majority of the stellar mass belongs to a radially-biased orbit family ($\pi$-box or boxlet).
-We see the manifestation of an increase in radially-biased orbits in the three dimensional stellar density profile, where the bifurcation in the density profiles occur at the same kick velocity as when $\pi$-box and boxlet orbits start to dominate the inner orbital structure.
+We see the manifestation of an increase in radially-biased orbits in the three dimensional stellar density profile (\autoref{fig:density3d}), where the bifurcation in the density profiles occur at the same kick velocity as when $\pi$-box and boxlet orbits start to dominate the inner orbital structure, and a sudden deficit of stellar mass in the central regions of the merger is apparent.
+
+To explain the presence of the bifurcation in the IFU maps, we must investigate how the stellar density is behaving in the wake of the recoiling SMBH.
+We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (approximately matching the aperture of the mock IFU maps), recalling that the SMBH is ejected along the positive $x$ axis.
+We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane \drafting{see Appendix}.
+For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
+This is contrasted to the $\vk=1500\,\kmps$ case, where there is no clear alignment of the stellar velocity flow with either the $x$- or $y$- axes.
+We interpret this feature as stellar particles diffusing into the potential wake carved by the recoiling SMBH in the $\vk=1500\,\kmps$ case, where the diffusion is neither parallel nor orthogonal to the SMBH motion, but rather at a variety of angles to it.
+
+The IFU maps describe the line of sight velocity distribution, thus the motion along the $x$-axis in \drafting{figure} is equivalent to the parallel projection, and the $y$-axis equivalent to the orthogonal projection previously described in \autoref{ssec:ifu}.
+When the SMBH recoil velocity is small, there is no potential wake created by the SMBH, and thus stellar diffusion to fill that wake does not occur, or occurs weakly in the case of a small but non-zero kick velocity.
+As a result, stellar velocities are in general aligned with one of the two observation axes, leading to a narrow distribution of stellar velocities.
+A result of this is the $h_4$ kinematic moment, which describes excess distribution tails, is close to zero: there are few high velocity outliers to the distribution.
+Conversely, when a large SMBH recoil velocity has propelled the SMBH through the galaxy, and stellar particles are diffusing into the potential wake, the variety of incoming angles, coupled with projection effects, results in a wide velocity distribution.
+A result of this is more stellar particles with high velocities, causing high velocity tails to be present in the LOSVD, and hence an increase in the $h_4$ kinematic moment.
 
@@
 \section{Triaxiality}\label{sec:app_triax}
-\drafting{Need to update}
-We present the triaxiality of all sixteen merger remnants in the analysis at the time when the SMBH has settled.
-We determine the ratios $b/a$ and $c/a$, where $a$, $b$, and $c$ are the eigenvalues of the reduced inertia tensor $I_\mathrm{red}$ and $c\leq b\leq a$. 
-The tensor $I_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} R_\mathrm{vir}$ to $10^{-1} R_\mathrm{vir}$: i.e., the binning of particles is not cumulative.
+We present the triaxiality of all sixteen merger remnants at the time $t \simeq 0.075\,\mathrm{Gyr}$. 
+We determine the ratios $b/a$ and $c/a$, where $a$, $b$, and $c$ are the eigenvalues of the reduced inertia tensor \citep{gerhard1983,bailin2005}
+\begin{equation}
+    \tilde{I}_\mathrm{red} = \sum_k \frac{r_{k,i} r_{k,j}}{r_k^2}
+\end{equation}
+and $c\leq b\leq a$. 
+The tensor $\tilde{I}_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} R_\mathrm{vir}$ to $10^{-1} R_\mathrm{vir}$: i.e., the binning of particles is not cumulative.
 The variation of the ratios $b/a$ and $c/a$ is plotted as a function of radius in \autoref{fig:triax}.
-All remnants display a high degree of symmetry, with the minimum value of $b/a\simeq 0.8$, and the minimum value of $c/a\simeq0.7$.
-With a higher kick velocity, there is a slight tendency to have a higher degree of symmetry at radii $R\lesssim 10\,\mathrm{kpc}$, and a slightly lesser degree of symmetry at radii beyond this distance, compared to lower kick velocity remnants.
-The difference is however minimal between all merger remnants.
+The merger remnants display triaxial cores, albeit without dependence on the recoiling SMBH velocity. 
+At radii greater than $4\,\mathrm{kpc}$, the triaxiality of the merger remnants is almost identical.
 
@@
         Triaxiality ratios $b/a$ and $c/a$ as a function of radius for all merger remnants in the analysis.
-        Lines are colour coded by the kick velocity.
+        Lines are colour coded by the kick velocity, with the same scheme as in \autoref{fig:density3d}.
         Note that the $\vk=0\,\kmps$ case is equivalent to the pre-merger remnant.
-        \drafting{Update colour bar to match other plots}
     }
@@
 \label{lastpage}
+
+\section{Stellar velocity fields}
+We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
+The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \drafting{figure}.
+
 \end{document}

### 2024-05-06 papers/paper-corekick/paper-corekick.tex
Add wake figure to paper

@@
 We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (approximately matching the aperture of the mock IFU maps), recalling that the SMBH is ejected along the positive $x$ axis.
-We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane \drafting{see Appendix}.
+We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane \autoref{fig:wake}.
 For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
@@
 
-The IFU maps describe the line of sight velocity distribution, thus the motion along the $x$-axis in \drafting{figure} is equivalent to the parallel projection, and the $y$-axis equivalent to the orthogonal projection previously described in \autoref{ssec:ifu}.
+The IFU maps describe the line of sight velocity distribution, thus the motion along the $x$-axis in \autoref{fig:wake} is equivalent to the parallel projection, and the $y$-axis equivalent to the orthogonal projection previously described in \autoref{ssec:ifu}.
 When the SMBH recoil velocity is small, there is no potential wake created by the SMBH, and thus stellar diffusion to fill that wake does not occur, or occurs weakly in the case of a small but non-zero kick velocity.
@@
 
+\section{Stellar velocity fields}
+We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
+The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \drafting{figure}.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{wake}
+    \caption{
+        Stellar velocity field for $\vk=0\,\kmps$ (left panel) and $\vk=1500\,\kmps$ (right panel), for an aperture that is approximately consistent with the IFU maps presented in e.g.
+\autoref{fig:IFU0p}.
+        The right panel demonstrates the stellar particles streaming into the potential wake caused by the recoiling SMBH (the SMBH is beyond the extent of the plot), whereas the same motion is not present in the left panel.
+        This causes a difference in the LOSVDs of the two systems, noticeably the $h_4$ kinematic moment.
+    }
+    \label{fig:wake}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\section{Stellar velocity fields}
-We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
-The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \drafting{figure}.
-
 \end{document}

### 2024-05-08 papers/paper-corekick/paper-corekick.tex
small edits

@@
         All simulations with $\vk \geq 0.5v_\mathrm{esc}=900\,\kmps$ share the same colour, as there is little variation in the profiles for these kick velocities. 
-        A clear dichotomy is revealed in the mass profiles, whereby high kick velocoties induce a core, and low kick velocities do not, with a sudden transition at \drafting{$\vk=600\,\kmps$}.
+        A clear dichotomy is revealed in the mass profiles, whereby high kick velocities induce a core, and low kick velocities do not, with a sudden transition at $\vk=540\,\kmps$.
         }
@@
 To explain the presence of the bifurcation in the IFU maps, we must investigate how the stellar density is behaving in the wake of the recoiling SMBH.
-We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (approximately matching the aperture of the mock IFU maps), recalling that the SMBH is ejected along the positive $x$ axis.
-We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane \autoref{fig:wake}.
+We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (the aperture of the mock IFU maps in \autoref{sec:app_ifu} extended to approximately $5\,\mathrm{kpc}$), recalling that the SMBH is ejected along the positive $x$ axis.
+We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane (see \autoref{fig:wake}).
 For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
@@
 
-\section{Specific IFU maps}
+\section{Specific IFU maps}\label{sec:app_ifu}
 \drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, $\vk=600\,\kmps$, and $\vk=1800\,\kmps$.}
@@
 We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
-The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \drafting{figure}.
+The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \autoref{fig:wake}.
 

### 2024-05-14 papers/paper-corekick/paper-corekick.tex
Minor comments and thoughts

@@
 This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
+\drafting{Maybe move the following two sentences to Atte's section
+To centre the data herein, we use the shrinking sphere method \citep{power2003} to determine the stellar centre of mass, and find the time evolution of the CoM does not systematically vary with kick velocity.
+Consequently, we are confident that our results are not a result of malign centring.}
 For the $\vk=0\,\kmps$ case, we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$, and is insensitive to the time of measurement.
@@
     \end{itemize}
+
+    For simulation with vk 480 and less, the BH and the CoM coincide (on approx. scales). 
+    Keplerian orbits only present in those sims where the BH and CoM are close (also confirmed with early analysis of snapshot 8 instead of 16) and also where the BH velocity is less than 300km per s.
+
+    We can use results from this work to identify systems where the SMBH is settling, difficult (impossible?) to determine exact history of system once BH settled. 
 }

### 2024-05-22 papers/paper-corekick/paper-corekick.tex
Update figures and captions

@@
 % Keep the title short and informative.
-\title[Cores from kicks]{Cores from kicks: constraints for observations of gravitational recoil in elliptical galaxies}
+\title[Cores from kicks]{Cores from kicks: inferring the gravitational recoil history of SMBHs in elliptical galaxies.}
 
@@
 \begin{abstract}
-\drafting{Investigate the relationship between core size and kick velocity}
+\drafting{We answer the questions:
+\begin{enumerate}
+    \item How does a GW kick influence the stellar mass distribution?
+    \item Can we infer if a GW kick has occurred in a ETG?
+\end{enumerate}
+}
 \end{abstract}
@@
         All simulations with $\vk \geq 0.5v_\mathrm{esc}=900\,\kmps$ share the same colour, as there is little variation in the profiles for these kick velocities.
-        A clear dichotomy is revealed in the mass profiles, whereby high kick velocities induce a core, and low kick velocities do not, with a sudden transition at $\vk=540\,\kmps$.
+\drafting{Should probably update this if max is 1020, maybe add special colour for 2000?}.
+        At radii $R\lesssim3\,\mathrm{kpc}$, the projected stellar mass density decreases smoothly with increasing kick velocity, with the exception of the $\vk=2000\,\kmps$ simulation, in which the SMBH has not returned to the centre of the merger remnant.
+        At radii $R>3\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity only affects the inner mass distribution of the galaxy merger remnant.
         }
@@
 Consequently, we are confident that our results are not a result of malign centring.}
-For the $\vk=0\,\kmps$ case, we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$, and is insensitive to the time of measurement.
+For the $\vk=0\,\kmps$ case, \drafting{check} we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$.
+
 In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $3\,\mathrm{kpc}$ ($2r_\mathrm{infl}$) than those mergers with $\vk>540\,\kmps$.
@@
         Surface density profiles with 25\% Bayesian HDI for select representative runs.
-        Increasing the kick velocity induces a shallower density profile in the inner regions.
+        Increasing the kick velocity induces a shallower density profile in the inner regions, consistent with the three dimensional density maps in \autoref{fig:density3d}.
         }
@@
     \caption{
-        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{\mathrm{b},0}$, as a function of kick velocity $\vk$.
+        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{\mathrm{b},0}$ (left axis) and in physical units (right axis), as a function of kick velocity $\vk$.
         The core size distributions are shown as box plots, with the median core size indicated by the central mark.
         Larger kick velocities are correlated with larger core sizes.
-Additionally, a greater spread in the distribution of core sizes over different viewing projections of the merger remnant is associated with larger kick velocities.}
+Additionally, a greater spread in the distribution of core sizes over different viewing projections of the merger remnant is associated with larger kick velocities.
+        Three models, fit to the median $\vk$--$\rb$ trend, are overplotted.
+        For the models, we do not show the parameter uncertainty for visual clarity.
+        }
     \label{fig:vkrb}
@@
     \includegraphics[width=0.4\textwidth]{rb_pdf}
-    \caption{Probability density function of transformation sampled core radius.
-The SMBH kick velocity is Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins, and pushed through the fitted model in \autoref{eq:vkrb}.
+    \caption{Probability density functions of transformation sampled core radius assuming a given model relation. 
+    The SMBH kick velocity is Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins, and pushed through each of the three fitted model in \autoref{eq:vkrb} \drafting{add other models}.
+    The sigmoid model is the preferred model, and has its mode at a larger value of $\rb$ than either the exponential or linear models, at $\rb/r_{\mathrm{b},0}\sim2.8$ ($\rb/\mathrm{kpc}\sim1.6$).
     The predicted kick velocities range from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
-    The asymmetry at $\rb/r_{\mathrm{b},0}\simeq 1$ is due to the low kick velocities producing a modest core, as seen in \autoref{fig:vkrb}.}
+    }
     \label{fig:rb_pdf}
@@
         Similar to \autoref{fig:density3d}, simulations with $\vk\geq 900\,\kmps$ have a consistent colour.
-        Lines terminating with a circle point indicate simulations with $\vk/\kmps < 600$, those with square points indicate simulations with $600 \leq \vk/\kmps < 1200$, those with triangle points $1200 \leq \vk/\kmps < 1800$, and those with diamond points $\vk/\kmps > 1800$.
-        For the parallel case, no systematic trend between $h_4$ and $\vk$ appears to exist.
-        Conversely, for the orthogonal projection, a dichotomy similar to the bifurcation seen in \autoref{fig:density3d} is observed. 
+        Lines terminating with a circle point indicate simulations with $\vk/\kmps < 600$, those with square points indicate simulations with $600 \leq \vk/\kmps < 1200$, and the diamond point indicates the $\vk=2000\,\kmps$ simulation.
+        For both the parallel and orthogonal projections, for those simulations with $\vk \leq 1020\,\kmps$ (i.e.
+have their SMBHs settled), there is a systematic decrease in the cumulative $h_4$ parameter with increasing $\vk$ for $\vk\lesssim600\,\kmps$.
+        For $600\lesssim\vk/\kmps\leq1020$, the cumulative $h_4$ parameter is indistinguishable from each other. 
+        For the $\vk=2000\,\kmps$ case, the cumulative $h_4$ is both positive and significantly larger than the $600\lesssim\vk/\kmps\leq1020$ simulations, for both parallel and orthogonal projections.
     }
@@
     \caption{
-        Orbit analysis of all sixteen merger remnants.
+        Orbit analysis of the merger remnants, with a consistent colour scheme to \autoref{fig:density3d}.
         Stellar particles are assigned to one of seven different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 30.0$.
-        Note the colour scheme is consistent with that in \autoref{fig:density3d}, with kick velocities exceeding $900\,\kmps$ represented using the same colour.
+        Noticeably, increasing recoil velocity induces a higher fraction of $\pi$-box orbits at all radii, and a complimentary decrease in rosette orbits at radii $r\lesssim1\,\mathrm{kpc}$.
     }
@@
 We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (the aperture of the mock IFU maps in \autoref{sec:app_ifu} extended to approximately $5\,\mathrm{kpc}$), recalling that the SMBH is ejected along the positive $x$ axis.
-We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane (see \autoref{fig:wake}).
+We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane.
 For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
@@
 
-The IFU maps describe the line of sight velocity distribution, thus the motion along the $x$-axis in \autoref{fig:wake} is equivalent to the parallel projection, and the $y$-axis equivalent to the orthogonal projection previously described in \autoref{ssec:ifu}.
+The IFU maps describe the line of sight velocity distribution, thus the motion along the $x$-axis in \drafting{removed figure} is equivalent to the parallel projection, and the $y$-axis equivalent to the orthogonal projection previously described in \autoref{ssec:ifu}.
 When the SMBH recoil velocity is small, there is no potential wake created by the SMBH, and thus stellar diffusion to fill that wake does not occur, or occurs weakly in the case of a small but non-zero kick velocity.
@@
     \caption{
+        \drafting{NEED TO UPDATE}
         Corner plot of the latent parameters in the core-S\'ersic model for the $\vk=600\,\kmps$ instance.
@@
 \section{Specific IFU maps}\label{sec:app_ifu}
-\drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, $\vk=600\,\kmps$, and $\vk=1800\,\kmps$.}
+\drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, and $\vk=660\,\kmps$. Add others?}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{IFU_para_v0600}
+    \includegraphics[width=\textwidth]{IFU_para_v0660}
     \caption{
@@
         Note the colour scaling is consistent with \autoref{fig:IFU0p}.
-        Most striking is the concentration of negative ordered velocity $V$ slightly offset from the remnant centre.
-        The other kinematic moments display a similar spatial offset.
+        Most striking is the slightly enhanced ordered velocity $V$ (some $20\,\kmps$) and the ring of negative $h_4$.
     }
@@
     \centering
-    \includegraphics[width=\textwidth]{IFU_ortho_v0600}
+    \includegraphics[width=\textwidth]{IFU_ortho_v0660}
     \caption{
@@
         Note the colour scaling is consistent with \autoref{fig:IFU0p}.
-        Compared to the parallel case, the ordered velocity $V$ and $h_3$ moments appear consistent with the $\vk=0,\kmps$ case, however a spatial offset from the centre persists in $\sigma$ and $h_4$.
+        Compared to the parallel case, the ordered velocity $V$ and $h_3$ moments appear consistent with the $\vk=0,\kmps$ case, however the ring of negative $h_4$ persists.
     }
@@
 \end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_para_v1800}
-    \caption{
-        Mock integral field unit kinematic maps of the $\vk=1800\,\kmps$ remnant in the parallel projection, plotted out to $0.25\,\Reff$ in a circular aperture. 
-        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
-        This projection displays clear ordered rotation, however there is no clear peak in velocity dispersion.
-        The other kinematic moments are quantitatively similar to the $\vk=0,\kmps$ case. 
-    }
-    \label{fig:IFU1800p}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_ortho_v1800}
-    \caption{
-        The same as \autoref{fig:IFU1800p}, but in the orthogonal projection. 
-        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
-        Aside from the lack of ordered rotation, there is little difference between the orthogonal and parallel projections at this kick velocity. 
-    }
-    \label{fig:IFU180o}
-\end{figure*}
-
-\section{Stellar velocity fields}
-We create vector fields of stellar particles for two representative merger remnants, namely $\vk=0\,\kmps$ and $\vk=1500\,\kmps$.
-The velocity fields are created for particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$, as described in \autoref{sec:discussion}, and are shown in \autoref{fig:wake}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{wake}
-    \caption{
-        Stellar velocity field for $\vk=0\,\kmps$ (left panel) and $\vk=1500\,\kmps$ (right panel), for an aperture that is approximately consistent with the IFU maps presented in e.g.
-\autoref{fig:IFU0p}.
-        The right panel demonstrates the stellar particles streaming into the potential wake caused by the recoiling SMBH (the SMBH is beyond the extent of the plot), whereas the same motion is not present in the left panel.
-        This causes a difference in the LOSVDs of the two systems, noticeably the $h_4$ kinematic moment.
-    }
-    \label{fig:wake}
-\end{figure}
 

### 2024-05-24 papers/paper-corekick/paper-corekick.tex
Update results section to reflect the change to a varied time analysis
- discussion section still to be edited

@@
 \drafting{Should probably update this if max is 1020, maybe add special colour for 2000?}.
-        At radii $R\lesssim3\,\mathrm{kpc}$, the projected stellar mass density decreases smoothly with increasing kick velocity, with the exception of the $\vk=2000\,\kmps$ simulation, in which the SMBH has not returned to the centre of the merger remnant.
-        At radii $R>3\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity only affects the inner mass distribution of the galaxy merger remnant.
+        At radii $r\lesssim3\,\mathrm{kpc}$, the projected stellar mass density decreases smoothly with increasing kick velocity, with the exception of the $\vk=2000\,\kmps$ simulation, in which the SMBH has not returned to the centre of the merger remnant.
+        At radii $r>3\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity only affects the inner mass distribution of the galaxy merger remnant.
         }
@@
 
-In \autoref{fig:density3d}, we observe a clear dichotomy in the 3D spatial distribution of stellar mass, with mergers that had an SMBH with a kick velocity $\vk<540\,\kmps$ having more mass in the central $3\,\mathrm{kpc}$ ($2r_\mathrm{infl}$) than those mergers with $\vk>540\,\kmps$.
-All merger remnants have a consistent mass density distribution at $r>3\,\mathrm{kpc}$. 
-Additionally, for kick velocities $\vk<540\,\kmps$, we observe a gradient in decreasing central stellar density with increasing kick velocity.
-We check the time dependence of this dichotomy by constructing the 3D stellar mass distributions at additional times $t=0.056\,\mathrm{Gyr}$, $t=0.094\,\mathrm{Gyr}$, and $t=0.12\,\mathrm{Gyr}$, and find the dichotomy persists, with the same kick velocity split.
+In \autoref{fig:density3d}, we observe a clear decrease in stellar mass density at radii $r\lesssim3\,\mathrm{kpc}$ with increasing kick velocity.
+An exception to this trend is the highest kick velocity we simulated, $\vk=2000\,\kmps$, which shows an excess of stellar mass between $0.3 \lesssim r/\mathrm{kpc} \lesssim 2$ relative to the $vk=1020\,\kmps$ simulation.
+Recalling that the time of the analysis is performed when the remnant SMBH has settled, and thus occurs at increasingly later times with increasing kick velocity, SMBHs kicked with a larger $\vk$ have undergone more oscillations to return the SMBH to an equilibrium state compared to SMBHs with a lower $\vk$.
+An increased number of oscillations presents an enhanced ability to remove stellar mass from the centre, as reflected in the density profiles in \autoref{fig:density3d}.
+Similarly, the $\vk=2000\,\kmps$ case, in which the SMBH has not undergone any oscillations in the centre of the galaxy merger remnant, has not had an opportunity to evacuate as much mass from the central regions as, for example, the $\vk=1020\,\kmps$ has.
+Conversely, the mass deficit in the $\vk=2000\,\kmps$ is greater than that exhibited by all kick velocities up to and including $\vk=480\,\kmps$.
+This observation is in agreement with \citet{nasim2021b}, in that the major contributor to a central mass deficit is the initial excursion of the recoiling SMBH, and the subsequent oscillations playing a subdominant role in the enhancement of the stellar core.
 
-The reasons for the dichotomy are discussed further in \drafting{section}, however we continue our analysis using the projected stellar density (as an observer would see) with the knowledge that there is \textit{a} dependence of the removal of central stellar mass on the SMBH kick velocity.
+The primary result of the three dimensional density profile is the observation of a recoil velocity dependent impact on the stellar density structure, which we can take as a ground truth when we consider the galaxy merger remnants from a more observationally-motivated perspective.
+We hypothesise that the difference between merger remnants in three dimensional space will manifest in other methods of observing the galaxy.
 
 \subsection{Projected mass density profiles}\label{ssec:projdens}
-To facilitate comparisons with observations \citep[e.g.][]{graham2003,rusli2013}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projection of each simulated merger remnants:
-\begin{equation}\label{eq:cs}
-    \Sigma(R) = \Sigma' \left[1 + \left(\frac{\rb}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[-b \left(\frac{R^\alpha + \rb^\alpha}{\Reff^\alpha}\right)^{(1/\alpha n)}\right]
-\end{equation}
-where
-\begin{equation}
-    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[b\left(2^{1/a} \frac{\rb}{\Reff}\right)^{1/n}\right].
-\end{equation}
-Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $n$ is the S\'ersic index, $b$ is estimated as $b\simeq 2n - 1/3 + 0.009876/n$ \citep{prugniel1997}, and $\alpha$ is the profile transition index.
-We refer to the collective vector of these parameters as $\vb{\theta}$.
-We view the simulated merger remnant from fifteen random angles, and use a Bayesian hierarchical model (HM) to fit the model parameters.
-As each of the projections are unordered and exchangeable (i.e., there is no distinguishing label for the projections), and each projection is of the same merger remnant, a HM naturally lends itself.
-The idea behind the HM is as such: for each projection, we infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible $\vb{\theta}$ vectors.
-Hence, we are effectively inferring the global hyperparameters $\vb{\theta}^\mathrm{hyp}$ that describe the latent parameters $\vb{\theta}$ that we are interested in.
-This idea is demonstrated in the directed acyclic graph (DAG) for the model, shown in \autoref{fig:dag}.
-
-We use weakly-constrained prior distributions for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these prior distributions are chosen to reflect reasonable values that the parameters may take (e.g.
-distributions of a radial quantity are chosen so as to be positively-constrained).
-The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
-
-We fit the model with \textsc{Stan} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo (HMC) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
-This resulted in 2000 draws from the posterior distribution of $\vb{\theta}$, whereby we ensure the diagnostic value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.
-We perform additional prior sensitivity tests using the power scaling method described in \citet{kallioinen21}, ensuring that the cumulative Jensen-Shannon (CJS) divergence is less than 0.05 for the latent parameters $\vb{\theta}$ of the model.
-This recommended threshold indicates that the sampled posterior distributions of the latent parameters are not sensitive to the particular prior distributions used.
-
-Our results indicate that a kick non-zero kick velocity always serves to enlarge the core region relative to the core formed during the SMBH binary scouring phase.
-In agreement with previous studies \citep{gualandris2008,nasim2021b}, we find an increase in the core radius $\rb$ with kick velocity, however only up to a given kick velocity, in agreement with the results of \citet{boylan2004}.
-We find that unlike the results of \citet{nasim2021b}, where core radius monotonically increases with kick velocity, we observe a peak in core radius occurring at $\vk=480\,\kmps=0.27\,v_\mathrm{esc}$, with the core radii measured from simulations with kick velocities in excess of $480\,\kmps$ plateauing at $\rb\sim1.7 r_{\mathrm{b},0}$.
-We caution however that the systems in the study of \citet{nasim2021b} differ from ours in both structure, and the time of the analysis.
-Whereas we use a fixed time following the SMBH binary coalescence for our analysis, \citet{nasim2021b} wait until either the ejected SMBH settles back to the CoM, or $2\,\mathrm{Gyr}$ has passed.
-\drafting{As discussed in section \textbf{Atte}, we find relaxation increasingly dominates our merger remnants with increasing time.}
-As a result, a direct comparison between this work and that of \citet{nasim2021b} cannot be made.
-
-Complimentary to the peak in core radius at $480\,\kmps$, we observe a dip in the projected density at the core radius $\Sigma_\mathrm{b}$ at this same kick velocity in \autoref{fig:csparams}.
-The effective radius $\Reff$ and core density slope $\gamma$ are relatively robust to the SMBH recoil velocity.
-
-\begin{table}
-    \caption{Hyperparameter distributions for core-S\'ersic model. $\mathcal{N}$ indicates a normal distribution, $\mathrm{Exp}(\lambda)$ indicates an exponential distribution.}
-    \label{tab:hyper}
-    \begin{tabular}{llc}
-        \hline
-        Hyperparameter & Distribution & Truncation \\
-        \hline
-        $\mu_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(10, 2)$ & None \\
-        $\sigma_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
-        $\lambda_\gamma$ & $\mathrm{Exp}(10)$ & None \\
-        $\mu_n$ & $\mathcal{N}(8,4)$ & $0 < x \leq 15$ \\
-        $\sigma_n$ & $\mathcal{N}(0,4)$ & $x>0$ \\
-        $\sigma_\alpha$ & $\mathrm{Gamma}(2, 0.2)$ & None \\
-        $\sigma_{r_\mathrm{b}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
-        $\sigma_{R_\mathrm{e}}$ & $\mathcal{N}(0, 12)$ & $x>0$ \\
-        $\mu_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
-        $\sigma_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
-        \hline
-    \end{tabular}
-\end{table}
 
@@
 
-\subsection{Predicted core size distribution}
-\drafting{Would a Gaussian process fit be better motivated here?}
-To predict the distribution of core sizes, we require a mapping from the kick velocity to the core radius. 
-Previous work by \citet{nasim2021b} fit a power law of the form:
-\begin{equation}\label{eq:vkrb_nasim}
-    \frac{\rb}{r_{\mathrm{b},0}} = K v'^\beta + 1
-\end{equation}
-where $v'=\vk/v_\mathrm{esc}$, and $K$ and $\beta$ are free parameters.
-We see immediately that \autoref{eq:vkrb_nasim} does not capture the peak in the data at $480\kmps$, and so opt for an alternative form.
-We fit a function that is the sum of a modified Student's $t$-distribution with a sigmoid function, and takes the form:
-\begin{equation}\label{eq:vkrb}
-    \frac{\rb}{r_{\mathrm{b},0}} = \frac{A(v'; \mu,\nu,\sigma)+B(v'; \mu,\nu,\sigma)}{2\pi\sigma} - S(v';k,b,\mu) + c,
+To facilitate comparisons with observations \citep[e.g.][]{graham2003,rusli2013}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projection of each simulated merger remnants:
+\begin{equation}\label{eq:cs}
+    \Sigma(R) = \Sigma' \left[1 + \left(\frac{\rb}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[-b \left(\frac{R^\alpha + \rb^\alpha}{\Reff^\alpha}\right)^{(1/\alpha n)}\right]
 \end{equation}
@@
 \begin{equation}
-    A(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu-2}{\nu}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu-2}{\nu}}} \right)^2 \right)^{-(\nu + 1)/2},
+    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[b\left(2^{1/a} \frac{\rb}{\Reff}\right)^{1/n}\right].
 \end{equation}
-and
-\begin{equation}
-    B(v'; \mu,\nu,\sigma) = \sqrt{\frac{\nu}{\nu-2}} \left(1 + \left(\frac{v' - \mu}{\sigma \sqrt{\frac{\nu}{\nu-2}}} \right)^2 \right)^{-(\nu + 1)/2},
-\end{equation}
-and
-\begin{equation}
-    S(v';k,b,\mu) = \frac{k}{1 + \exp(-b (v'-\mu))}.
-\end{equation}
-Hence \autoref{eq:vkrb} takes six free parameters ($\mu$, $\nu$, $\sigma$, $k$, $b$, $c$), which we constrain using the median trend of the $vk/v_\mathrm{esc}$--$\rb$ relation in \autoref{fig:vkrb}, and assume a normally-distributed scatter $\tau$.
-The fit is found using the same Bayesian methods as in \autoref{ssec:projdens} (including posterior checking), and is plotted with an orange line in \autoref{fig:vkrb}, with the values of the fitted parameters given in \autoref{tab:vkrb}.
+Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $n$ is the S\'ersic index, $b$ is estimated as $b\simeq 2n - 1/3 + 0.009876/n$ \citep{prugniel1997}, and $\alpha$ is the profile transition index.
+We refer to the collective vector of these parameters as $\vb{\theta}\equiv[\rb, \Sigma_\mathrm{b}, \gamma, \Reff, n, \alpha]$.
+We view the simulated merger remnant from fifteen random angles, and use a Bayesian hierarchical model (HM) to fit the model parameters.
+As each of the projections are unordered and exchangeable (i.e., there is no distinguishing label for the projections), and each projection is of the same merger remnant, a HM naturally lends itself.
+The idea behind the HM is as such: for each projection, we infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible $\vb{\theta}$ vectors.
+Hence, we are effectively inferring the global hyperparameters $\vb{\theta}^\mathrm{hyp}$ that describe the latent parameters $\vb{\theta}$ that we are interested in.
+This idea is demonstrated in the directed acyclic graph (DAG) for the model, shown in \autoref{fig:dag}.
+
+We use weakly-constrained prior distributions for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these prior distributions are chosen to reflect reasonable values that the parameters may take (e.g.
+distributions of a radial quantity are chosen so as to be positively-constrained).
+The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
+
+We fit the model with \textsc{Stan} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo (HMC) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
+This resulted in 2000 draws from the posterior distribution of $\vb{\theta}$, whereby we ensure the diagnostic value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.
+We perform additional prior sensitivity tests using the power scaling method described in \citet{kallioinen21}, ensuring that the cumulative Jensen-Shannon (CJS) divergence is less than 0.05 for the latent parameters $\vb{\theta}$ of the model.
+This recommended threshold indicates that the sampled posterior distributions of the latent parameters are not sensitive to the particular prior distributions used.
+
+In \autoref{fig:density}, we see a clear decrease in the projected stellar mass density with increasing kick velocity for four representative values of $\vk$, consistent with the results in \autoref{fig:density3d}.
+The shaded regions in \autoref{fig:density} correspond to the $25\%$ Bayesian highest density interval (HDI), where the uncertainty is a direct result of choosing different projection angles when constructing the projected densities.
+
+To discuss the difference in the evacuated core region of the merger remnants more quantitatively, we show the median (blue dashes) and interquartile range (IQR, blue boxes) of the core radius parameter $\rb$ from \autoref{eq:cs} for kick velocities $\vk\leq 1020\,\kmps$ in \autoref{fig:vkrb}.
+Our results indicate that a non-zero kick velocity always serves to enlarge the core region relative to the core formed during the SMBH binary scouring phase.
+In agreement with previous studies \citep{gualandris2008,nasim2021b}, we find a monotonic increase in the core radius $\rb$ with kick velocity.
+Additionally, we see that the IQR of $\rb$ increases with increasing $\vk$, indicating that with an increased core size, the uncertainty introduced due to projection effects also increases.
+
+As shown in the upper central panel of \autoref{fig:csparams}, the projected density at the core radius decreases with increasing kick velocity, in agreement with the decreasing core density observed in the three dimensional density profiles in \autoref{fig:density3d}.
+Similarly, the S\'ersic index also decreases with increasing kick velocity (bottom right panel of \autoref{fig:csparams}), although the overall decrease is from $n\simeq2.1$ at $\vk=0\,\kmps$ to $n\simeq1.6$ at $\vk=1020\,\kmps$.
+The other three parameters: $\Reff$, $\alpha$, and $\gamma$, all appear insensitive to the kick velocity.
+The fact that $\Reff$ and $\alpha$ are agnostic to $\vk$ is not surprising, given that these parameters control more of the larger-scale profile of \autoref{eq:cs}.
+\drafting{A constant $\gamma$ however implies that irrespective of the core size, the steepness of the core density is \dots what does it mean? Mass is iteratively removed in oscillations?}
 
 \begin{table}
-    \caption{Values of the fitted parameters in the core radius -- kick velocity relation in \autoref{eq:vkrb}.}
-    \label{tab:vkrb}
-    \begin{tabular}{lccc}
+    \caption{Hyperparameter distributions for core-S\'ersic model. $\mathcal{N}$ indicates a normal distribution, $\mathrm{Exp}(\lambda)$ indicates an exponential distribution.}
+    \label{tab:hyper}
+    \begin{tabular}{llc}
         \hline
-        Parameter & Median & IQR & Constraints \\
+        Hyperparameter & Distribution & Truncation \\
         \hline
-        $\mu$ & 0.182 & 0.0427 & $\geq 0$ \\
-        $\nu$ & 21.5 & 11.9 & $\geq 0$ \\
-        $\sigma$ & 0.601 & 0.298 & $\geq 0$ \\
-        $k$ & 0.894 & 0.189 & $\geq 0$ \\
-        $b$ & -34.5 & 30.9 & $\leq 0$ \\
-        $c$ & 1.64 & 0.0661 & $\geq 0$ \\
-        $\tau$ & 0.170 & 0.0336 & $\geq 0$ \\
+        $\mu_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(10, 2)$ & None \\
+        $\sigma_{\log_{10}\Sigma_\mathrm{b}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
+        $\lambda_\gamma$ & $\mathrm{Exp}(10)$ & None \\
+        $\mu_n$ & $\mathcal{N}(8,4)$ & $0 < x \leq 15$ \\
+        $\sigma_n$ & $\mathcal{N}(0,4)$ & $x>0$ \\
+        $\sigma_\alpha$ & $\mathrm{Gamma}(2, 0.2)$ & None \\
+        $\sigma_{r_\mathrm{b}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
+        $\sigma_{R_\mathrm{e}}$ & $\mathcal{N}(0, 12)$ & $x>0$ \\
+        $\mu_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 1)$ & $x>0$ \\
+        $\sigma_{\tau_{\log_{10}\Sigma}}$ & $\mathcal{N}(0, 0.2)$ & $x>0$ \\
         \hline
@@
 
-Using \autoref{eq:vkrb}, we push through the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig. 15, right panel).
-The dimensionless spin parameter $\alpha_\bullet$ for the dry mergers follows a beta distribution, namely:
+\subsection{Predicted core size distribution}
+To predict the distribution of core sizes given our merger remnant of two massive ETGs, we require a mapping from the kick velocity to the core radius. 
+Let us define a scaled kick velocity $v'=\vk/v_\mathrm{esc}$, where $v_\mathrm{esc}=1800\,\kmps$.
+Previous work by \citet{nasim2021b} fit an empirical power law of the form:
+\begin{equation}\label{eq:vkrb_exp}
+    \frac{\rb}{r_{\mathrm{b},0}} = K v'^\beta + 1
+\end{equation}
+where $K$ and $\beta$ are free parameters.
+\autoref{eq:vkrb_exp} gives that the core radius monotonically increases irrespective of kick velocity, implying that as $\vk\rightarrow\infty$, $\rb\rightarrow\infty$.
+
+We test two other empirical relations between $\rb$ and $\vk$.
+The first is a simple linear relation:
+\begin{equation}\label{eq:vkrb_lin}
+    \frac{\rb}{r_{\mathrm{b},0}} = a v' + b,
+\end{equation}
+and the second takes the form of a sigmoid function:
+\begin{equation}\label{eq:vkrb_sig}
+    \frac{\rb}{r_{\mathrm{b},0}} = K \left( 1 - e^{-\beta v'} \right) + c.
+\end{equation}
+
+We fit each of these models to the median trend in \autoref{fig:vkrb}, and display the relation with its best fit parameters in the same figure.
+In the fit, we assume weakly informative positively-constrained normally distributed priors on the parameters, a normally distributed scatter $\tau$ truncated to $\tau>0$, and a Gaussian likelihood function.
+The fit is found using the same Bayesian methods as in \autoref{ssec:projdens} (including posterior checking and prior sensitivity analysis), allowing for uncertainties to be estimated for the model fits.
+The medians, IQRs, and constraints of the relation parameters in \autoref{eq:vkrb_exp}-\autoref{eq:vkrb_sig} are given in \autoref{tab:vkrb}.
+
+\begin{table}
+    \caption{Values of the fitted parameters in the three core radius -- kick velocity relations.}
+    \label{tab:vkrb}
+    \begin{tabular}{llcc}
+        \hline
+        Model & Parameter & Median & IQR \\
+        \hline
+        Exponential & $K$ & 2.90 & 0.37 \\
+        & $\beta$ & 0.78 & 0.13 \\
+        \hline
+        Linear & $a$ & 3.26 & 0.41 \\
+        & $b$ & 1.1 & 0.14 \\
+        \hline
+        Sigmoid & $K$ & 2.47 & 0.59 \\
+        & $\beta$ & 2.62 & 1.31 \\
+        & $c$ & 0.87 & 0.16 \\
+        \hline
+    \end{tabular}
+\end{table}
+
+Using the three $\vk$--$\rb$ relations, we wish to construct the posterior predictive distributions for new data $\tilde{\rb}$ given the observed values of $\rb$:
+\begin{equation}
+    p(\tilde{\rb}|\rb) = \int p(\tilde{\rb} | \vb{\kappa}) p(\vb{\kappa} | \rb) \dd{\vb{\kappa}},
+\end{equation}
+where the integral marginalises out the respective $\vk$--$\rb$ relation parameters, collectively referred to as $\vb{\kappa}$ \citep{gelman2015}.
+The new data $\tilde{\rb}$ is generated by transformation sampling the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig.
+15, right panel), and `pushing' the values through the desired $\vk$--$\rb$ relation.
+Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}, whilst simultaneously accounting for uncertainty in the fit described by the $\vk$--$\rb$ relations.
+
+For the \citet{zlochower2015} dry merger distribution, the dimensionless spin parameter $\alpha_\bullet$ follows a beta distribution, namely:
 \begin{equation}
@@
 The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
-Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}, whilst simultaneously accounting for uncertainty in the fit described by \autoref{eq:vkrb}.
-We find that the mode of the forward-folded core radius distribution is $1.07\,\mathrm{kpc}$ ($1.69\,r_{\mathrm{b},0}$).
-Thus, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, a large fraction massive elliptical galaxies should have a non-negligible core \citep{ferrarese1994,savorgnan2016,sahu2019}.
+An immediately apparent caveat is that the $\vk$--$\rb$ relations are constrained using kick velocities $\vk\leq1020\,\kmps$: extrapolating from these relations introduces an inherent risk when we do not know how the $\vk$--$\rb$ relation holds for very large $\vk$, which we discuss further below.
+
+We find that the mode of the forward-folded core radius distribution is similar for the exponential and linear models, with $1.04\,\mathrm{kpc}$ ($1.84\,r_{\mathrm{b},0}$) and $0.97\,\mathrm{kpc}$ ($1.71\,r_{\mathrm{b},0}$), respectively.
+Conversely, the mode for the sigmoid relation to occur at a larger value of $\rb$, namely $1.60\,\mathrm{kpc}$ ($2.83\,r_{\mathrm{b},0}$).
+Looking at the shapes of the three distribution in \autoref{fig:rb_pdf}, we see a similarity in shapes between the exponential and linear models, which are both right-skewed. 
+This is due to the ever-increasing value of $\rb$ predicted for an increasing $\vk$: higher recoil velocities produce larger cores, but as higher recoil velocities become increasingly unlikely (assuming the \citet{zlochower2015} distribution), we observe the skew in the $\rb$ probability density function (PDF).
+Conversely, the sigmoid model has a small hump prior to the mode, arising from the high likelihood of small kick velocities being pushed through the sigmoid function (\autoref{eq:vkrb_sig}) prior to the upper plateau.
+The peak in the $\rb$ distribution from the sigmoid model is a result of a large range of kick velocities (albeit with decreasing frequency) being mapped to the upper plateau of the sigmoid function, hence the correspondence between the median of the scale factor $K=2.47$ in \autoref{eq:vkrb_sig} and the distribution mode for the model, $2.83\,r_{\mathrm{b},0}$.
+
+To determine which model of the three (exponential, linear, or sigmoid) best describes the data, we use leave one out cross validation (LOO-CV) with Pareto smoothed importance sampling \citep[PSIS][]{vehtari2017}.
+PSIS LOO-CV estimates the expected log pointwise predictive density ($\widehat{\mathrm{ELPD}}$), defined
+\begin{equation}
+    \widehat{\mathrm{ELPD}} = \sum_{i=1}^N \log \int p(y_i | \theta) p(\theta | y_{-i}) \dd{\theta},
+\end{equation}
+to determine the out-of-sample predictive accuracy of a Bayesian model, i.e., the ability to predict new data $\tilde{y}$ conditioned on the fitted data $y$. 
+A larger value of $\widehat{\mathrm{ELPD}}$ indicates better predictive ability of a given model compared to some other model \citep{vehtari2017, riha2024}.
+We report the preferred model to be the sigmoid model (\autoref{eq:vkrb_sig}), with an $\widehat{\mathrm{ELPD}}$ of 4.06, compared to 2.68 and 0.54 for the exponential and linear models, respectively.
+
+All in all, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, a large fraction massive elliptical galaxies should have a non-negligible core, which is in agreement with observations \citep{ferrarese1994,savorgnan2016,sahu2019}, even if the recoil velocity is modest.
 
@@
     \caption{Probability density functions of transformation sampled core radius assuming a given model relation. 
-    The SMBH kick velocity is Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins, and pushed through each of the three fitted model in \autoref{eq:vkrb} \drafting{add other models}.
+    The SMBH kick velocity is Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins, and pushed through each of the three fitted model in \autoref{eq:vkrb_exp}, \autoref{eq:vkrb_lin}, and \autoref{eq:vkrb_sig}
     The sigmoid model is the preferred model, and has its mode at a larger value of $\rb$ than either the exponential or linear models, at $\rb/r_{\mathrm{b},0}\sim2.8$ ($\rb/\mathrm{kpc}\sim1.6$).
@@
 \subsection{Integral field unit kinematics}\label{ssec:ifu}
-We create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the BH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the BH kick axis (`orthogonal').
+Having established from a forward-modelling point of view that a changing SMBH recoil velocity induces a systematic affect in the mass distribution of the galaxy remnant, we turn our attention to inferring the occurrence of an SMBH that has experienced a recoil velocity, based on the snapshot when the SMBH is seen to be settled.
+
+We first create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the BH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the BH kick axis (`orthogonal').
 Testing these two special projections, we investigate the limiting cases of how the galaxy kinematics behave when all or none of the BH motion is directed along the LOS to the observer.
@@
 \begin{equation}\label{eq:h4}
-    \langle h_4 \rangle = \sum_i^N h_4(R_i)
+    \langle h_4 \rangle = \sum_i^N h_4(R_i),
 \end{equation}
-As shown in \autoref{fig:h4}, for the parallel projections, simulations with $\vk \geq 600\,\kmps$ tend to positive values of $\langle h_4 \rangle$, asymptotically reaching $\langle h_4 \rangle \sim 1$.
-Conversely, when the recoiling SMBH has a kick velocity $\vk < 600\,\kmps$, the final value of $\langle h_4 \rangle$ is evenly distributed about $\langle h_4 \rangle = 0$: there is no systematic preference to heaver or weaker tails in the LOS velocity distribution.
+which improves legibility of the radial trends of the $h_4$ parameter, which can oscillate about zero.
 
-Conversely, for the orthogonal projections, we observe a dichotomy similar to that seen in \autoref{fig:density3d}, with the bifurcation occurring at a similar kick velocity, here $600\,\kmps$; both partitions tend to a positive value of $\langle h_4 \rangle$.
-Within the two partitions of the bifurcation, there is no apparent dependence on kick velocity.
-The gradient between the two sets differs most noticeably in the radial range $2  \leq R/\mathrm{kpc} < 3$.
-This can be understood as those recoiling SMBHs with $\vk \geq 600\,\kmps$ having a larger $h_4$ magnitude per Voronoi bin than the $\vk < 600\,\kmps$ simulations.
+The radial dependence of $\langle h_4 \rangle$ is shown in \autoref{fig:h4}, where simulations with $\vk/\kmps < 600$ are terminated with a circle point, simulations with $600 \leq \vk/\kmps < 1020$ are terminated with a square point, and the $\vk=2000\,\kmps$ simulation is terminated with a diamond point.
+Aside from the $\vk=2000\,\kmps$ simulation, increasing the kick velocity results in progressively more negative $\langle h_4 \rangle$ in both the parallel and orthogonal projections. 
+The value of $\langle h_4 \rangle$ at the largest radii saturates for simulations with $\vk\gtrsim 600\,\kmps$, at a value of $\sim -0.5$ for the parallel projection, and $\sim -0.1$ for the orthogonal projection.
+The case is markedly different though for the $\vk=2000\,\kmps$ simulation, which appears more consistent with the $\vk=0\,\kmps$ simulation than the simulations with $\vk\gg0\,\kmps$.
+The value of $\langle h_4 \rangle$ increases at all radii for both projections, exceeding a maximum value of 1.0 for the parallel projection, and 1.5 for the orthogonal projection. 
+\drafting{can we relate this to the Langrangian radii perhaps?}
 
@@
         Similar to \autoref{fig:density3d}, simulations with $\vk\geq 900\,\kmps$ have a consistent colour.
-        Lines terminating with a circle point indicate simulations with $\vk/\kmps < 600$, those with square points indicate simulations with $600 \leq \vk/\kmps < 1200$, and the diamond point indicates the $\vk=2000\,\kmps$ simulation.
+        Lines terminating with a circle point indicate simulations with $\vk/\kmps < 600$, those with square points indicate simulations with $600 \leq \vk/\kmps < 1020$, and the diamond point indicates the $\vk=2000\,\kmps$ simulation.
         For both the parallel and orthogonal projections, for those simulations with $\vk \leq 1020\,\kmps$ (i.e.
@@
 Three distinct trends are made apparent in \autoref{fig:orbits}.
-The first and most important, is the complete absence of rosette orbits for all simulations with $\vk>600\,\kmps$.
-Conversely, all simulations with $\vk<600\,\kmps$ (except for $\vk=540\,\kmps$) have more than $\sim25\%$ of the orbits in the innermost radial bin being identified as rosette. 
-The second and third trend are complimentary, and are an increase in the fraction of $\pi$-box orbits with increasing kick velocity, and a similar decrease in the fraction of $x$-tube orbits with increasing kick velocity.
-Finally, there is a weak tendency for simulations with $\vk<900\,\kmps$ to have slightly higher fractions of boxlet orbits at small radii, and higher fractions of $z$-tube orbits, than those simulations with $\vk>900\,\kmps$.
+The first is the increase in the fraction of $\pi$-box orbits with increasing recoil velocity, most notable at radii $r\lesssim5\,\mathrm{kpc}$.
+The simulation with $\vk=2000\,\kmps$ noticeably stands out, with a fraction of $\pi$-box orbits being $\sim0.6$ at the inner most radii. 
+Complimentary but converse to the first trend, increasing the SMBH recoil velocity decreases the fraction of $x$-tube orbits at inner radii from $\sim 0.4$ to $\sim 0.2$. 
+Again the $\vk=2000\,\kmps$ is a noticeable outlier to the other simulations, however remains consistent with the general trend.
+Finally, increasing the SMBH recoil velocity decreases the fraction of rosette orbits at small radii $r<1\,\mathrm{kpc}$, with no rosette orbits for the $\vk=2000\,\kmps$ simulation.
+There is evidence for a minor decrease in $z$-tube orbits with increasing recoil velocity, however this trend is not as dramatic as the three previously discussed.
+
+We can naturally explain the decrease in rosette orbits with increasing recoil velocity by considering a rosette orbit requires a point-mass like potential to orbit in.
+With increasing recoil velocity, we are displacing the SMBH increasingly further from the stellar centre of mass, thus disrupting the conditions required to maintain such an orbit.
+A consequence of disrupting these regular, rosette orbits, is the inducement of more radial orbits with increasing recoil velocity, manifested as an increase in non-resonant $\pi$-box orbits.
+
+We show the orbit families as a function of radius organised by recoil velocity in \autoref{fig:orbits2}, instead of family class.
+At large radii ($r>10\,\mathrm{kpc}$), $x$-tube orbits dominate over all other families for all simulations, \drafting{a known result from major mergers, Naab??}.
+Interestingly, increasing the recoil velocity pushes the intersection of boxlet and $z$-tube orbits to increasingly larger radii.
+Finally, for the $\vk=2000\,\kmps$ simulation, the dominance of $\pi$-box orbits over all other families is apparent for radii $r\lesssim5\,\mathrm{kpc}$.
+This is the only recoil velocity for which $\pi$-box orbits are the dominant orbital family; in the simulations with $\vk\leq1020\,\kmps$, $\pi$-box orbits are always subdominant to boxlet orbits.
 

### 2024-05-25 papers/paper-corekick/paper-corekick.tex
New sketch figures for Atte

@@
 As discussed in \drafting{section}, higher recoil velocities typically result in larger excursions of the remnant SMBH from the isophotal centre.
+\drafting{Need to edit all this, \textbf{Atte} could you do this seeing as you've chosen the snapshots?
+
 To provide a fair comparison between simulations, we choose to analyse each child simulation at a fixed time after the recoil kick, namely $t\simeq 0.075\,\mathrm{Gyr}$.
@@
 This is critical as the \ketju{} region does not include gravitational softening between stellar particles and the SMBH, so is a not fully-softened simulation.
+}
 
@@
         \item Plot of infall time (i.e., time for stability criterion to be reached) as a function of kick velocity
-        \item Number of oscillations before settling?
-        \item Plot of maximum displacement of kicked BH as a function of kick velocity. How does this compare to core size? Which simulations have BH kicked beyond the core?
+        \item Number of oscillations before settling, probably on the same plot but as a second y axis?
+        \item Plot of apocentre of kicked BH as a function of kick velocity. How does this compare to core size? Which simulations have BH kicked beyond the core? Can fit a function to data maybe
+        \item Langrangian radii: how is the stellar mass moving about? 
     \end{enumerate}
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{infall_time}
-    \caption{Plot of infall time (time to stability criterion) as a function of kick velocity.}
+    \includegraphics[width=0.4\textwidth]{infall_apo}
+    \caption{Plot of infall time (time to stability criterion) as a function of kick velocity, and number of oscillations. Maybe can do apocentre plot too: make this a 2x1 figure??}
     \label{fig:infall}
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{max_disp}
-    \caption{Plot of the maximum BH displacement as a function of kick velocity. Where do the displacement (red) and core size (green) lines intersect?}
+    \includegraphics[width=0.4\textwidth]{lagrangian}
+    \caption{Plot of Lagragian radii: comment on matter movement}
     \label{fig:max_disp}

### 2024-06-04 papers/paper-corekick/paper-corekick.tex
Add Max beta plot to paper

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{anisotropy_profiles_std_log}
+    \caption{
+        \drafting{\textbf{Max contribution} Put here for now, a better spot elsewhere? Also need to update caption and include in text.}
+        Stellar velocity anisotropy parameter $\beta$ as a function of radius for the galaxy mergers with kick velocities between $0\,\kmps$ and $2000\,\kmps$. 
+        The stellar particles are divided into 20 logarithmic bins between $0.1\,\mathrm{kpc}$ and $20\,\mathrm{kpc}$.
+        The dotted horizontal line indicates $\beta=0$. 
+    }
+    \label{fig:beta}
+\end{figure}
+
 \subsection{Orbit analysis}\label{ssec:orbits}

### 2024-07-04 papers/paper-corekick/diagrams/dag.tex

@@
     % break radius hypers
-    \factor[left=of dens_b] {dens_b_D} {below:$\mathcal{N}$} {} {};
+    \factor[left=of dens_b, xshift=-0.7em] {dens_b_D} {below:$\mathcal{N}_\mathrm{T[-5,15]}$} {} {};
     \node[hyper, left=of dens_b_D, yshift=2em] (dens_b_mu) {$\mu_{\log_{10}\Sigma_\mathrm{b}}$};
@@
     % inner slope hypers
-    \factor[left=of g, xshift=-1.5em] {g_D} {below:$\mathrm{Exp}_\mathrm{T}$} {} {};
+    \factor[left=of g, xshift=-1.5em] {g_D} {below:$\mathrm{Exp}_\mathrm{T[0,2]}$} {} {};
     \node[hyper, left=of g_D] (g_lam) {$\lambda_\gamma$};
@@
     % sersic index hypers
-    \factor[left=of n, xshift=-1.5em] {n_D} {below:$\mathcal{N}_\mathrm{T}$} {} {};
+    \factor[left=of n, xshift=-1.5em] {n_D} {below:$\mathcal{N}_\mathrm{T[0,20]}$} {} {};
     \node[hyper, left=of n_D, yshift=1em] (n_mu) {$\mu_{n}$};
@@
     % transition index hypers
-    \factor[left=of a, xshift=-1.5em] {a_D} {below:$\mathrm{Ray}$} {} {};
+    \factor[left=of a, xshift=-1.5em] {a_D} {below:$\mathrm{Ray}_\mathrm{T[0,15]}$} {} {};
     \node[hyper, left=of a_D] (a_sig) {$\sigma_\alpha$};
@@
     % core radius hypers
-    \factor[left=of rb, xshift=-1.5em] {rb_D} {below:$\mathrm{Ray}_\mathrm{T}$} {} {};
+    \factor[left=of rb, xshift=-1.5em] {rb_D} {below:$\mathrm{Ray}_\mathrm{T[0,5]}$} {} {};
     \node[hyper, left=of rb_D] (rb_sig) {$\sigma_{r_\mathrm{b}}$};
@@
     % effective radius hypers
-    \factor[left=of Re, xshift=-1.5em] {Re_D} {below:$\mathrm{Ray}_\mathrm{T}$} {} {};
+    \factor[left=of Re, xshift=-1.5em] {Re_D} {below:$\mathrm{Ray}_\mathrm{T[0,20]}$} {} {};
     \node[hyper, left=of Re_D] (Re_sig) {$\sigma_{R_\mathrm{e}}$};
@@
     % surface density error hypers
-    \factor[above=of dens_err, yshift=4em] {dens_err_D} {right:$\mathcal{N}_\mathrm{T}$} {} {};
+    \factor[above=of dens_err, yshift=4em] {dens_err_D} {left:$\mathcal{N}_\mathrm{T[0,]}$} {} {};
     \node[hyper, above=of dens_err_D, yshift=6em] (dens_err_err) {$\sigma_{\tau_{\log_{10}\Sigma}}$};

### 2024-10-10 papers/paper-corekick/paper-corekick.tex
Bring changes from Overleaf to git

--- 2024-05-24 papers/paper-corekick/paper-corekick.tex
@@
 \usepackage{bm}         % Bold math fonts
+\usepackage{pifont}
 
@@
 \usepackage{xcolor}
-\newcommand{\drafting}[1]{
-{\leavevmode\color[RGB]{224, 77, 24}#1}
-}
+\newcommand{\alex}[1]{
+{\leavevmode\color[RGB]{224, 77, 24}Alex: #1}
+}
+\newcommand{\atte}[1]{{\color{teal}Atte: #1}}
+\newcommand{\peter}[1]{{\color{purple}Peter: #1}}
+\newcommand{\maxx}[1]{{\color{orange}Max: #1}}
+\newcommand{\ruby}[1]{{\color{blue}Ruby: #1}}
+\newcommand{\sonja}[1]{{\color{violet}Sonja: #1}}
 
@@
 \newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}            % Solar Mass
+\newcommand{\Lsun}{\ensuremath{\mathrm{L}_{\sun}}}            % Solar luminosity
 \newcommand{\kmps}{\ensuremath{\mathrm{km}\,\mathrm{s}^{-1}} }% km/s formatting
@@
 \newcommand{\rb}{\ensuremath{r_\mathrm{b}}}                   % core radius
+\newcommand{\rbbin}{\ensuremath{r_\mathrm{b,0}}}                   % initial core radius
 \newcommand{\vk}{\ensuremath{v_\mathrm{kick}}}                % kick velocity
+\newcommand{\vkmax}{\ensuremath{v_\mathrm{kick, max}}}        % max kick velocity
 
@@
 \newcommand{\vb}[1]{\ensuremath{\bm{#1}}}                     % vectors
+\newcommand{\ordinal}[1]{\ensuremath{#1^\mathrm{th}}}                     % ordinal th
+\newcommand{\ordinalthird}[1]{\ensuremath{#1^\mathrm{rd}}}                     % ordinal rd
 
@@
 % Keep the title short and informative.
-\title[Cores from kicks]{Cores from kicks: inferring the gravitational recoil history of SMBHs in elliptical galaxies.}
+\title[Identifying SMBH recoil]{
+Identifying supermassive black hole recoil in elliptical galaxies
+}
 
@@
 % If you need two or more lines of authors, add an extra line using \newauthor
-\author[A. Rawlings et al.]{
-Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
-et al.
-\vspace*{0.1cm}\\%
+\author[A. Rawlings et al.]{Alexander Rawlings,$^{1}$\thanks{E-mail: alexander.rawlings@helsinki.fi}
+Atte Keitaanranta,$^{1}$
+Max Mattero,$^{1}$
+Sonja Soininen,$^{1}$
+Ruby J. Wright,$^{2,1}$
+\newauthor
+Noa Kallioinen,$^{3}$
+Shihong Liao,$^{4}$
+Antti Rantala,$^{5}$
+Peter H. Johansson,$^{1}$
+Thorsten Naab,$^{5}$ and
+\newauthor Dimitrios Irodotou$^{6}$
+\vspace*{0.1cm}\\
 % List of institutions
@@
 Gustaf H\"allstr\"omin katu 2, FI-00014, University of Helsinki, Finland
-\\%
-%$^{2}$
-%Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1, D-85748 Garching, Germany
+\\
+$^{2}$
+International Centre for Radio Astronomy Research (ICRAR), University of Western Australia, Crawley, WA 6009, Australia
+\\
+$^{3}$
+Department of Computer Science, Aalto University, Espoo, Finland
+\\
+$^{4}$
+Key Laboratory for Computational Astrophysics, National Astronomical Observatories, Chinese Academy of Sciences, Beijing 100101, China
+\\
+$^{5}$
+Max-Planck-Institut f\"ur Astrophysik, Karl-Schwarzchild-Str 1, D-85748 Garching, Germany
+\\
+$^{6}$
+The Institute of Cancer Research, London SW7 3RP, UK
+\\
 }
@@
 % Enter the current year, for the copyright statements etc.
-\pubyear{2023}
+\pubyear{2024}
 
@@
 \begin{abstract}
-\drafting{We answer the questions:
-\begin{enumerate}
-    \item How does a GW kick influence the stellar mass distribution?
-    \item Can we infer if a GW kick has occurred in a ETG?
-\end{enumerate}
-}
+\alex{We have 250 words to use here. Currently 247 excluding the 3 math inlines.}
+
+We study the contribution to stellar core growth in simulations of merging massive ($M_\star>10^9\,\Msun$) elliptical galaxies by a coalesced supermassive black hole (SMBH) displaced by gravitational wave (GW) induced recoil velocity.
+With controlled, dense sampling of the SMBH recoil velocity, we find the stellar core originally formed by SMBH binary scouring can grow by a factor of 2-3 with increasing recoil velocity, and a similar growth in the central mass deficit.
+Using Bayesian inference we predict the distribution of stellar core sizes formed through this process to peak at $\sim1\,\mathrm{kpc}$.
+An orbital decomposition of stellar particles within the core reveals that when the recoil velocity exceeds the velocity dispersion of the core, radial orbits dominate over tube orbits, providing a mass 10\% of the coalesced SMBH mass, whereas tube orbits dominate for the lowest recoil kicks.
+A change in orbital structure is reflected in the anisotropy parameter, with a tangential bias present only for kick velocities less than the local stellar velocity dispersion.
+With mock integral field unit observations of the stellar line of sight velocity distribution emulating the capabilities of current detectors, we observe a distinct signature in the Gauss-Hermite symmetric deviation coefficient $h_4$ that allows for an estimate of the core size due to binary scouring, irrespective of later evolution due to SMBH recoil.
+Our results provide a method to estimate the SMBH recoil magnitude from observations of local elliptical galaxies, and implies these galaxies primarily experienced recoil velocities less than the stellar velocity dispersion of the core.
 \end{abstract}
@@
 \section{Introduction}
-There is much evidence for the presence of supermassive black holes (SMBHs) at the centres of massive galaxies, with the shadow of one such SMBH being directly imaged \citep{eht2019}.
-The dynamical interactions of two or more galaxies, during a galaxy merger, are thus also expected to induce an interaction of the residing SMBHs.
-The interaction of two SMBHs during the merger of two galaxies is well understood to occur in a three stage process \citep{begelman1980}.
-First, as the galaxies begin to dynamically interact and the SMBHs traverse through a dense stellar field, the overdensity of stars in the wakes of the SMBHs provides a restoring force which brings the SMBHs to the centre of the two interacting galaxies \citep{chandrasekhar1943}, generally torquing the SMBHs to highly radial orbits in the process.
-Second, after the SMBHs have formed a bound binary, iterative interactions with the surrounding stars ejects stars with high velocity \citep[e.g.][]{hills1980,hills1983,quinlan1996}, thus removing orbital energy and angular momentum from the SMBH binary orbit.
-The removal of stars from the central regions by this slingshot mechanism produces a `core' \citep[e.g.][]{lauer1983,lauer1985,kormendy1984}, or depletion, in the luminosity profile of the remnant galaxy \citep[e.g.][]{begelman1980,hills1983,quinlan1996,rantala2018}, and can extend up to some kiloparsecs \citep{postman2012}.
-Finally, at very small, sub-parsec, separations the SMBH binary loses its remaining orbital energy and angular momentum through gravitational wave (GW) emission, driving the SMBH binary to coalescence \citep{peters1963,peters1964}.
+%There is much evidence for the presence of supermassive black holes (SMBHs) at the centres of massive galaxies, with the shadow of one such SMBH being directly imaged \citep{eht2019}.
+%As two galaxies dynamically interact during a galaxy merger, so too is it expected that the residing SMBHs will dynamically interact.
+Observations have revealed a clear bimodality in the properties of massive elliptical galaxies, with brighter ellipticals $(M_{B}\lesssim -20.5)$ typically having `cored' flat central surface brightness profiles, boxy isophotes and showing little or no rotation (slow rotators).
+Fainter intermediate-mass ellipticals $(-20.5 \lesssim M_{B} \lesssim -18.5)$ on the other hand typically show steeper `cuspy' central brightness profiles, more disc-like isophotes and evidence for more rotational support (fast rotators) (e.g \citealt{Kormendy1996,Faber1997,emsellem2007,Kormendy2009,Thomas2014,dullo2014,dullo2019,Quenneville2024}).
+
+The observed bimodality can be linked to the distinct formation paths of elliptical galaxies.
+Intermediate-mass ellipticals have likely formed predominantly through in-situ star formation and in mergers of gas-rich disc-dominated galaxies accompanied by merger-induced starbursts resulting in cuspy central stellar profiles (e.g.
+\citealt{Barnes1996,Cappellari2007,hopkins2009a,hopkins2009b,Johansson2009,Krajnovic2011,Cappellari2016,Lahen2018,Forster2020}).
+More massive elliptical galaxies are on the other hand believed to have assembled in a two-stage process, in which the early assembly at redshifts of $z \gtrsim 1.5$ is dominated by in-situ star formation fuelled by cold gas flows in massive dark matter (DM) haloes, and the accretion of multiple star-bursting progenitors.
+The later evolution of the galaxies then proceeds mainly through gas-poor (dry) merging in which the galaxy predominantly accretes stars formed  outside the main galaxy (e.g.
+\citealt{Naab2009,Bezanson2009,Oser2010,Johansson2012,Wellons2015,Rodriguez-Gomez2016,Qu2017,lagos2022,Cannarozzo2023}, see also \citealt{naab2017} for a review).
+However, if massive galaxies have formed by merging of fainter cuspy galaxies they should also possess central cusps, as simulations have shown that mergers tend to preserve the density cusps of merging galaxies \citep{boylan2004,Dehnen2005}.
+
+In the case of two massive elliptical galaxies undergoing a dry merger, the leading mechanism for stellar core formation is through the interaction of binary supermassive black holes (SMBHs) in the galaxy merger remnant.
+The merging of SMBHs
+is generally understood in the context of a three stage process \citep{begelman1980}.
+First, the dynamical friction from the mass overdensity in the wakes of the SMBHs provides a restoring force which brings the SMBHs to the centre of the two interacting galaxies \citep{chandrasekhar1943}, generally torquing the SMBHs to highly radial orbits in the process.
+Second, after the mass enclosed within the orbit of the two SMBHs is comparable to the sum of the SMBH masses, the SMBHs form a bound binary \citep{merritt2005}.
+Iterative interactions of the SMBHs with surrounding stars on low angular momentum orbits ejects these stars with high velocity \citep[e.g.][]{hills1980,hills1983,quinlan1996}, thus removing orbital energy and angular momentum from the SMBH binary orbit.
+The removal of stars from the central regions by this slingshot mechanism produces a `core' \citep[e.g.][]{lauer1983,lauer1985,kormendy1984}, or depletion, in the luminosity profile of the remnant galaxy \citep[e.g.][]{begelman1980,hills1983,quinlan1996,rantala2018,rantala2024,nasim2021b,khonji2024,partmann2024}, that can in extreme cases extend up to $3\,\mathrm{kpc}$ \citep{postman2012}.
+Finally, at very small, sub-parsec separations the SMBH binary loses its remaining orbital energy and angular momentum through gravitational wave (GW) emission, driving the SMBH binary to coalescence \citep{peters1963,peters1964}.
+The GW emission from such SMBH coalescence events are prime observational targets for ongoing pulsar timing array (PTA) detectors for SMBHs with masses of $M_\bullet \gtrsim 10^9\,\Msun$ \citep{agazie2023, 2023Antoniadis, 2023Xu, zic2023}, and also the forthcoming Laser Interferometer Space Antenna (LISA) mission for SMBHs with slightly lower masses of $M_\bullet \lesssim 10^8\,\Msun$ \citep{amaro-seoane2023}.
 
 The final GW emission from the coalescing SMBH binary occurs anisotropically, and carries with it linear momentum from the system \citep[e.g.][]{gonzalez2007}.
-This results in the remnant SMBH recoiling with some velocity directed opposite to the linear momentum of the GW emission \citep{bekenstein1973}, and is termed the kick velocity. 
-The kick velocity is dependent on the mass ratio between the SMBHs prior to coalescence, and the magnitude and direction of angular momentum (spin) vector of each SMBH. 
+This results in the remnant SMBH recoiling with some velocity directed opposite to the linear momentum of the GW emission \citep{bekenstein1973}, and is termed the kick, or recoil, velocity\footnote{We use the terms `kick' and `recoil' interchangeably in this work.}.
+The kick velocity is dependent on the mass ratio between the SMBHs prior to coalescence, and the magnitude and direction of the angular momentum (spin) vector of each SMBH. 
 As reported by \citet{campanelli2007}, symmetries in the masses or spins of the SMBHs suppresses the resulting recoil kick imparted to the coalesced SMBH.
-In particular, it is found that asymmetry in the spins of the SMBHs has a larger impact on the magnitude of the recoil kick than asymmetry in the masses \citep{campanelli2007}, with antialigned spins in the orbital plane producing the largest recoil kick velocities \citep{gonzalez2007b,tichy2007}.
-Numerical relativity studies indicate that while the majority of recoil kicks are of the order of a few hundred kilometres per second, in special configurations the SMBH may be imparted a kick velocity in excess of a $2000\,\kmps$, even up to $4000\,\kmps$ \citep{campanelli2007,gonzalez2007b,tichy2007}.
-Taking the central escape velocity of a typical early type galaxy (ETG) as $\sim2000\,\kmps$, there may be a non-negligible fraction of ETGs with a SMBH that has escaped the galaxy \citep{madau2004}, though not so many so as to introduce considerable scatter into the observed relation between SMBH mass and stellar velocity dispersion \citep{volonteri2007}.
-In cases where the kick velocity is less than the escape velocity of the inner regions of the host galaxy, the radial oscillations of the remnant SMBH as it returns to the centre of the merger remnant is expected to influence the surrounding stellar environment.
-\drafting{What sort of influences?
-References?}
+In particular, it is found that asymmetry in the spins of the SMBHs has a larger impact on the magnitude of the recoil kick than asymmetry in the masses \citep{campanelli2007}, with spins which are anti-aligned in the orbital plane and maximal producing the largest recoil kick velocities \citep{gonzalez2007b,tichy2007}.
+Numerical relativity studies indicate that while the majority of recoil kicks are of the order of a few hundred kilometres per second, in special configurations the SMBH may be imparted a kick velocity in excess of $2000\,\kmps$, even up to $4000\,\kmps$ \citep{campanelli2007,gonzalez2007b,tichy2007}.
+Taking the central escape velocity of a typical massive elliptical galaxy as $\sim2000\,\kmps$, there may be a non-negligible fraction of elliptical galaxies with an SMBH that has escaped the galaxy (e.g.
+\citealt{madau2004,mannerkoski2022}), though not so many so as to introduce considerable scatter into the observed relation between SMBH mass and stellar velocity dispersion \citep{volonteri2007}.
+This is in clear tension with all current observations indicating that no massive elliptical galaxy in the local Universe is without an SMBH, and demands further investigation.
 
@@
 Another recoiling SMBH candidate was identified by an offset in velocity by \citet{steinhardt2012}, with an estimated recoil velocity greater than $4000\,\kmps$.
-Other observations by \citet{comerford2014} and \citet{pesce2018} have also reported potential recoiling SMBHs, albeit at much lower kick velocities (some few hundreds of $\kmps$).
-\drafting{In \citet{pesce2018}, they talk about a peculiar velocity of 70ish km/s, whereas in later work, they stated velocity is 4000 km/s.
-Not sure which is correct, that difference seems suspicious\dots}
+Other observations by \citet{comerford2014} and \citet{pesce2018, pesce2021} have also reported potential recoiling SMBHs, albeit at much lower kick velocities (some tens to few hundreds of $\kmps$).
 Recoiling SMBHs may also be identified through a spatial offset from the host centre, where the offset can range from some parsecs \citep{batcheldor2010,lena2014,barrows2016} to over a kiloparsec \citep{koss2014,skipper2018}.
-Finally, some candidate recoiling SMBHs have both spatial and velocity offsets, such as CXO J101527.2+625911 \citet{kim2017} and 2MASX J00423991+3017515 \citet{hogg2021}.
-
-In this work, we first aim to determine how the velocity with which the SMBH is kicked affects the surrounding stellar environment.
-We then aim to determine if there is an observational signature of the SMBH recoil kick that may be used to infer the occurrence of a kicked SMBH. 
-\drafting{Previous work by \dots indicates \dots}
-
-\drafting{This paper is divided as follows:}
+Additionally, some candidate recoiling SMBHs have both spatial and velocity offsets, such as CXO J101527.2+625911 \citep{kim2017} and 2MASX J00423991+3017515 \citep{hogg2021}.
+Attributing an escaping SMBH to GW recoil as opposed to three body interactions between an SMBH binary and a third SMBH is also not trivial, as demonstrated in the recent observations presented in \citet{vandokkum2023}.
+
+Numerically, the effect of recoiling SMBHs on the formation of cores in massive elliptical galaxies has been 
+primarily studied using collisionless simulations. Early work by \citet{boylan2004}
+found that a recoiling SMBH induced a core in the stellar density profile, particularly when the SMBH remained bound to the bulge of the host galaxy and had a recoil velocity comparable to the bulge velocity dispersion.
+In another early study
+\citet{merritt2004} also found that recoiling SMBHs facilitated the formation of cores
+and found that ejected SMBHs settled roughly in one orbital period in spherical potentials, with this `infall time' increasing when more realistic triaxial potentials were considered. 
+In \citet{gualandris2008} the motion of a settling SMBH in a stellar potential representative of a galaxy merger remnant was found to occur in three stages: first, the SMBH oscillates with decreasing amplitude due to energy dissipation by dynamical friction.
+Next, both the SMBH and stellar core oscillate about their common centre of mass when the radial excursion of the SMBH is less than the core radius of the stellar system.
+Finally, the SMBH settles into thermal equilibrium with the surrounding stars.
+During these three stages, the stellar density core is enhanced by the motion of the SMBH, producing a mass deficit in some cases up to five times the SMBH mass (see also \citealt{merritt2004,boylan2004,partmann2024}).
+However, a more recent study by \citet{nasim2021b} on the contribution to stellar core expansion through GW recoil of an SMBH found intriguingly that the greatest contribution to the stellar mass deficit is due to the first excursion of the SMBH following its ejection, with subsequent oscillations only providing a minor contribution to the growth of the stellar core.
+%This result is in tension with the work of \citet{gualandris2008}.
+
+%The effect of recoiling SMBHs have been studied previously, with some of the first explorations presented in \citet{boylan2004} and \citet{merritt2004}.
+%In the work of \citet{boylan2004}, a recoiling SMBH was found to induce a core in the stellar density profile, particularly when the SMBH remained bound to the bulge of the host galaxy and had a recoil velocity comparable to the bulge velocity dispersion.
+%A recoiling SMBH was also found to induce a stellar density core in the work of \citet{merritt2004}.
+%Additionally, the time for a recoiling SMBH to settle after being ejected in a spherical potential was found to be roughly one orbital period when the recoil velocity was less than the velocity dispersion of the bulge.
+%This, which are more representative of a galaxy merger remnant than the spherical case.
+
+%More recently, the contribution of stellar core expansion through GW recoil of an SMBH has been considered by \citet{nasim2021b}.
+%Intriguingly, the authors find that greatest contribution to the stellar mass deficit is due to the first excursion of the SMBH following its ejection, with subsequent oscillations only providing a minor contribution to the growth of the stellar core.
+%This result is in tension with the work of \citet{gualandris2008}.
+
+%The effect of recoiling SMBHs have been studied previously, with some of the first explorations presented in \citet{boylan2004} and \citet{merritt2004}.
+%In the work of \citet{boylan2004}, a recoiling SMBH was found to induce a core in the stellar density profile, particularly when the SMBH remained bound to the bulge of the host galaxy and had a recoil velocity comparable to the bulge velocity dispersion.
+%A recoiling SMBH was also found to induce a stellar density core in the work of \citet{merritt2004}.
+%Additionally, the time for a recoiling SMBH to settle after being ejected in a spherical potential was found to be roughly one orbital period when the recoil velocity was less than the velocity dispersion of the bulge.
+%This `infall time' was found to increase when considering triaxial potentials, which are more representative of a galaxy merger remnant than the spherical case.
+
+%In \citet{gualandris2008}, a stellar potential representative of a galaxy merger remnant with SMBH binary core scouring is used to investigate the motion of an SMBH ejected through GW recoil.
+%The motion of a settling SMBH was found to occur in three stages: first, the SMBH oscillates with decreasing amplitude due to energy dissipation by dynamical friction.
+%Next, both the SMBH and stellar core oscillate about their common centre of mass when the radial excursion of the SMBH is less than the core radius of the stellar system.
+%Finally, the SMBH settles into thermal equilibrium with the surrounding stars.
+%During these three stages, the stellar density core is enhanced by the motion of the SMBH (in agreement with \citet{merritt2004} and \citet{boylan2004}), with the mass deficit in some cases reaching up to five times the SMBH mass being produced.
+
+%Work by \citet{blecha2008} considered the effect of recoiling SMBHs in an analytical model of an isolated galaxy comprised of a multicomponent system of a stellar bulge and gaseous disc.
+%An interesting result was the mass growth of the ejected SMBH as it settled at the centre of the potential, with the mass growth accumulating to some 10\% of the initial SMBH mass, irrespective of the recoil velocity.
+%Recoiling SMBHs in gas-rich systems was further explored in \citet{blecha2011} using hydrodynamic simulations of isolated galaxy mergers.
+%In this work, seed SMBHs were allowed to grow through gas accretion prior to merging, with the majority of final SMBH masses at the time of merger sampling the $10^7\,\Msun \lesssim M_\bullet \lesssim 10^8\,\Msun$ mass regime.
+
+%The impact of recoiling SMBHs on the growth and dynamics of SMBHs in hydrodynamical simulations have also been studied using both isolated galaxy models consisting of a stellar bulge and a gaseous disc \citep{blecha2008}, as well as in isolated gas-rich galaxy mergers \citep{blecha2011}.
+%The study of the isolated galaxy model found that
+%the mass growth of the ejected SMBH as it settled at the centre of the potential, with the mass growth accumulating to some 10\% of the initial SMBH mass, irrespective of the recoil velocity. 
+%In the merger simulations, the seed SMBHs were allowed to grow through gas accretion prior to merging, with the majority of the final SMBH masses at the time of merger sampling the $10^7\,\Msun \lesssim M_\bullet \lesssim 10^8\,\Msun$ mass regime.
+The simulations also showed that recoil events typically reduce the lifetimes of bright active galactic nuclei (AGNs), with most kicked AGNs confined to the central few kpc of the galaxy.
+Whilst the results of \citet{blecha2011} provide valuable insight into the occurrence of variable AGN variability due to an oscillating SMBH plunging through a gaseous disc, we stress that the results are applicable to a different mass regime of galaxies and SMBHs than what we focus on in this work.
+In addition, the formation of stellar cores which is primarily driven by complex three-body interactions between the SMBHs and the stellar component remains largely unresolved in simulations that include gravitational softening.
+
+Whilst previous studies have convincingly shown that there is a clear case for a relation between SMBH recoil velocity and the final core size carved out by its motion, little attention has been devoted to understanding the statistical impact of SMBH recoil on the properties of the merger remnant galaxy.
+Invoking large SMBH recoils as a mechanism to grow the most massive cores in elliptical galaxies \citep[e.g.][]{nasim2021b,khonji2024} has contributed an important piece to the overall puzzle of explaining current observations, however recoil velocities approaching the escape velocity of the host galaxy are unlikely to be the norm, given that all observed elliptical galaxies host a central SMBH \citep{ferrarese1994,haring2004,kormendy2013,vandenbosch2016}.
+The role of mild SMBH recoil velocities remains largely unexplored.
+In addition, how a recoiling SMBH impacts the surrounding stellar orbital distribution has remained largely unanswered. 
+Orbital modelling techniques, in particular orbit-superposition Schwarzschild modelling, offer a unique bridging between observations and simulations, often revealing a rich understanding of a galaxy's history, and the effect of SMBH interactions \citep[e.g.][]{rottgers2014,neureiter2021,neureiter2023,santucci2022,santucci2023,santucci2024}.
+
+In this study we use the \gadget{} based regularised tree code \ketju{} \citep{rantala2017,rantala2020,mannerkoski2023} to study the effect of recoiling SMBHs on the structure of massive elliptical galaxies.
+Previously, \ketju{} has been used to study the formation of cores through SMBH binary scouring in gas-free simulations, but without the contribution of SMBH recoils \citep{rantala2018,rantala2019,rantala2024}.
+Here, we run a large set of gas-free simulations with systematically varying SMBH kick velocities.  
+The aim of this work is twofold.
+We first aim to determine how the velocity with which the SMBH is kicked affects the mass distribution of the surrounding stellar environment, and the likelihood of the magnitude of these effects.
+We then aim to determine if a recoiling SMBH imparts a signature on the stellar kinematics of the remnant galaxy, and if such a signature could potentially be observed with current detectors.
+
+%In this study, we wish to consider SMBH recoil by GW emission in massive elliptical galaxies.
+%The aim of this work is twofold.
+%We first aim to determine how the velocity with which the SMBH is kicked affects the mass distribution of the surrounding stellar environment, and the likelihood of the magnitude of these effects.
+%We then aim to determine if a recoiling SMBH imparts a signature on the stellar kinematics of the remnant galaxy, and if such a signature could be observed with current detectors.
+
+This paper is divided as follows: we present the numerical simulations used in our investigation in \autoref{sec:num_sims}.
+We discuss the motion of the recoiling SMBHs in \autoref{sec:motion}, and the resulting stellar density profiles of the remnant galaxy in \autoref{sec:density}.
+We investigate the distribution of stellar orbits as a function of the kick velocity in \autoref{sec:orbits} and construct mock stellar kinematic observations to be compared with observational data in \autoref{sec:kinematics}.
+Finally, we discuss of our results in \autoref{sec:discussion} and present 
+our conclusions in \autoref{sec:conclusions}, respectively.
 
@@
 \subsection{Simulation Code}
-\drafting{Taken from eccentricity paper, edited a little bit but need to ensure it is different enough.}
-To investigate the effect of the recoiling SMBH on the host galaxy, and potential observational signatures of this, we run a number of numerical simulations of a galaxy merger setting using our new version of the \ketju{} code \citep{mannerkoski2023,rantala2017} coupled with \gadget{}.
-\ketju{} integrates the dynamics of SMBHs, and stars in a small region thrice the stellar softening length around them, to high accuracy using the algorithmically regularised integrator \textsc{mstar} \citep{rantala2020}.
-The dynamics of stellar particles beyond this small region of high integration accuracy, and all dark matter (DM) particles, is computed with the \gadget{} \citep{springel2021} fast multiple method (FMM) with second order multipoles.
-Additionally, we use hierarchical time integration, which allows for symmetric interactions and manifest momentum conservation. 
+To investigate the effect of the recoiling SMBH on the host galaxy, and potential observational signatures of this, we run a number of numerical simulations of an idealised galaxy merger setting using the new public version of the \ketju{}\footnote{\url{https://www.mv.helsinki.fi/home/phjohans/group-website/research/ketju/}} code \citep{rantala2017,rantala2020,mannerkoski2023} coupled with \gadget{} \citep{springel2021}.
+\ketju{} integrates the dynamics of SMBHs, and stellar particles in a small region thrice the stellar softening length around them (i.e.
+the \ketju{} region), to high accuracy using the algorithmically regularised integrator \textsc{mstar} \citep{rantala2020}.
+The dynamics of stellar particles beyond this small region of high integration accuracy, and all DM particles, is computed with the \gadget{} fast multiple method (FMM) with second order multipoles. 
+Additionally, we use hierarchical time integration, which allows for mutually symmetric interactions and manifest momentum conservation. 
+In this study, we do not use comoving integration, as we are investigating an idealised system.
 \ketju{} also includes post-Newtonian (PN) correction terms up to order 3.5 between each pair of SMBHs \citep{blanchet2014} and the fitting formula of \citet{zlochower2015} for recoil kick velocity, making \ketju{} a particularly well-suited code for investigating the consequence of SMBH recoil self-consistently in a galaxy merger environment.
@@
 \subsection{Initial Conditions}
-We model the merger of two gas-free elliptical galaxies using gas-free merger simulations.
-Our galaxy initial conditions are chosen similarly to the model IC-3 presented in \citet{rantala2019}, however with different stellar density slope. 
+We model the merger of two massive elliptical galaxies using gas-free merger simulations.
+Our galaxy initial conditions follow that of IC $\gamma$-1.0-BH-6 presented in \citet{rantala2018} but scaled down by a factor of $\sim 3$ in mass, consistent with the method presented in \citet{rantala2019}.
 The galaxy is thus represented as an isotropic multicomponent sphere, and consists of a stellar component of total mass $M_\star\sim 1.38\times10^{11}\,\Msun$ embedded within a DM component of total mass $M_\mathrm{DM}=2.5\times10^{13}\,\Msun$.
-At the centre a SMBH with mass $M_\bullet=2.93\times10^{9}\,\Msun$ is placed with zero velocity.
-The stellar and DM components each follow a \citet{dehnen1993} profile with shape parameter $\gamma=1$, and scale radius $a_\star=3.9\,\mathrm{kpc}$ and $a_\mathrm{DM}=245\,\mathrm{kpc}$, respectively\footnote{A Dehnen profile with $\gamma=1$ is commonly referred to as a Hernquist profile.}.
+At the centre an SMBH with mass $M_{\bullet,0}=2.93\times10^{9}\,\Msun$ is placed with zero velocity.
+The stellar and DM components each follow a \citet{hernquist1990} profile with a scale radius of $a_\star=3.9\,\mathrm{kpc}$ and $a_\mathrm{DM}=245\,\mathrm{kpc}$, respectively.
 The density profile $\rho_i$ for a given component $i$ is given by:
 \begin{equation}\label{eq:dehnen}
-    \rho_i(r) = \frac{(3-\gamma)M}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
-\end{equation}
-We generate the ICs using the distribution function method following \citet{hilz2012}, where for each component (stellar and DM) the distribution function $f_i$ is computed using Eddington's formula \citep{binney2008} for each density profile $\rho_i$:
+    \rho_i(r) = \frac{M_i}{2\pi} \frac{a_i}{r (r+a_i)^{3}}.
+\end{equation}
+We generate the ICs using the distribution function method following \citet{hilz2012} and \citet{rantala2017}, where for each component (stellar and DM) the distribution function $f_i$ is computed using Eddington's formula \citep{binney2008} for each density profile $\rho_i$:
 \begin{equation}
@@
 The distribution functions are then sampled with discrete particles, where the mass of a stellar particle is set to $m_\star=5\times10^4\,\Msun$, and the mass of a DM particle is set to $m_\mathrm{DM}=5\times10^6\,\Msun$.
-The radial velocity profiles of both stellar and DM particles is ergodic.
-
-As we are interested in the evolution of the galaxy merger \textit{remnant} following SMBH coalescence, the initial conditions are constructed such that at the time of SMBH coalescence (\autoref{ssec:mergers}) the galaxy remnant agrees with observational data.
-Specifically, we ensure that our remnant agrees with the half-light -- stellar mass data presented in \citet{sahu2020}, and lies within the $1\sigma$ predictive interval of the SMBH mass -- stellar velocity dispersion data given in \citet{vandenbosch2016}.
+The radial velocity profiles of both stellar and DM particles are ergodic.
+
+As we are interested in the evolution of the galaxy merger \textit{remnant} following SMBH coalescence, the initial conditions are constructed such that at the time of SMBH coalescence (\autoref{ssec:mergers}) the galaxy remnant agrees with local scaling relations.
+Specifically, we ensure that our remnant agrees with the half-light -- stellar mass data presented in \citet{sahu2020}, and lies within the $1\sigma$ prediction interval of the SMBH mass -- stellar velocity dispersion data given in \citet{vandenbosch2016}.
 
@@
 We merge two independently Monte Carlo sampled galaxy ICs, creating an equal mass galaxy merger with a near-radial orbit. 
-The merger orbit has an initial separation of $D=30\,\mathrm{kpc}$, a first pericentre distance of $r_\mathrm{peri}=2\,\mathrm{kpc}$, and an initial eccentricity of $e_0=0.97$.
-The merger configuration results in a rapid coalescence of the BH binary, however we note that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
+The merger orbit has an initial separation consistent with \citet{rantala2018} of $D=30\,\mathrm{kpc}$, however the initial eccentricity is set to $e_0=0.97$, and the first pericentre distance adjusted accordingly to $r_\mathrm{peri}=D(1-e_0^2) \simeq 2\,\mathrm{kpc}$ \citep{khochfar2006,rantala2017}.
+The merger configuration results in a rapid coalescence of the SMBH binary, however we note that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
+The resulting coalesced SMBH has a mass $M_\bullet = 2M_{\bullet,0} = 5.86\times10^9\,\Msun$.
 We then select the snapshot just prior ($\sim8\,\mathrm{Myr}$) to the GW-driven SMBH merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $\vk$ prescribed along the $x$-axis\footnote{As the $x$-axis is in the global coordinate frame, the direction of the kick is essentially random with respect to the angular momentum and inertia tensors.
-To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find identical evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
-At the moment of the merger, the centre of mass (CoM) velocity of the SMBH binary is $v_{\mathrm{bin},\mathrm{CoM}}\sim[4.2, 1.1, 1.6]\,\kmps$ in the global coordinate frame.
+To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find very similar evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
+%At the moment of the merger, the centre of mass (CoM) velocity of the SMBH binary is $v_{\mathrm{bin},\mathrm{CoM}}\sim[4.2, 1.1, 1.6]\,\kmps$ in the global coordinate frame.
 The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
-We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$.
-As discussed in \drafting{section}, higher recoil velocities typically result in larger excursions of the remnant SMBH from the isophotal centre.
-To provide a fair comparison between simulations, we choose to analyse each child simulation at a fixed time after the recoil kick, namely $t\simeq 0.075\,\mathrm{Gyr}$.
-This value is chosen so as to minimise the effects of relaxation in the central regions of the galaxy remnant.
-We estimate the relaxation time $t_\mathrm{relax}$ at a radius $r$ as \citep{binney2008}:
+We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$, to test the effect of rapidly removing an SMBH.
+
+For all simulations, snapshots are produced every $5\,\mathrm{Myr}$.
+Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, DM with a softening length of $\varepsilon = 300\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
+Within the \ketju{} region, dynamical interactions between stellar particles and SMBHs are unsoftened.
+For the \gadget{} component of the simulation, the integration error tolerance is set to 0.002, whereas the Gragg-Bulirsch-Stoer (GBS) tolerance for \ketju{} is set to $10^{-8}$.
+
+We impose a maximum integration time on all simulations, corresponding to three dynamical times $t_\mathrm{dyn}$, defined in terms of the virial radius $r_{200}$ and virial velocity $v_{200}$ of the galaxy merger remnant as:
+\begin{equation}
+    t_\mathrm{dyn} = \frac{r_{200}}{v_{200}} \simeq 1\,\mathrm{Gyr}.
+\end{equation}
+Thus, our simulations are run for a maximum of $3\,\mathrm{Gyr}$.
+
+% RESULTS
+\section{Settling SMBHs}\label{sec:motion}
+
+\subsection{Motion of the kicked SMBH}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figures/trajectory.pdf}
+    \caption{
+    The trajectory of the kicked SMBH in the $\vk=780\,\kmps$ simulation, with its final position shown by the marked circle.
+    The line is coloured by the time following merger, and an arrow displays the initial outward radial motion of the SMBH for clarity.
+    The inset panel depicts the motion of the SMBH as it is settling within the binary-scoured core radius, marked by the dotted circle.
+    The final settling of the SMBH is chaotic once its velocity falls below the local stellar velocity dispersion: it is unable to leave the core again from this moment onwards.
+    }
+    \label{fig:trajectory}
+\end{figure}
+
+Before SMBH binary coalescence, the binary scours a stellar core with a radius of $\rbbin\sim0.58\,\mathrm{kpc}$ in the centre of the galaxy (see \autoref{ssec:projdens}).
+In the interior of the core, the stellar one-dimensional velocity dispersion is $\sigma_{\star,0} \equiv \sigma_\star(r<\rbbin)\sim 270\,\kmps$.
+Immediately following the coalescence of the SMBH binary and the subsequent kick, the trajectory of the remnant SMBH is predominantly radial in the direction of the GW recoil kick, as shown in \autoref{fig:trajectory}.
+We can thus estimate the crossing time of the binary-scoured core based on the 1D stellar velocity dispersion as:
+\begin{equation}
+    t_\mathrm{cross, core} \simeq \frac{\rbbin}{\sigma_{\star,0}} \approx 2\,\mathrm{Myr},
+\end{equation}
+whereas the actual time the recoiling SMBH traverses the core is:
+\begin{equation}
+    t_\mathrm{cross, \bullet} \simeq \frac{\rbbin}{\vk}.
+\end{equation}
+If the velocity of the recoil kick is not enough to completely eject the SMBH from the galaxy ($\vk < v_\mathrm{esc}$), the SMBH will oscillate about the centre of the galaxy. 
+The amplitude of the oscillations, termed the apocentre distance $r_\mathrm{apo}$, decreases with each oscillation due to dynamical friction, which is automatically resolved with \ketju{} \citep[e.g.][]{karl2015,rantala2017,genina2024}.
+Ultimately the SMBH `settles' to the centre of the galaxy.
+If $\vk < \sigma_{\star,0}$, the SMBH has insufficient kinetic energy to escape the stellar core and instead remains within $100\,\mathrm{pc}$ from the centre, shown in the top panel of \autoref{fig:infall}.
+Conversely, if $\vk > \sigma_{\star,0}$ the SMBH can exit the core, with $r_\mathrm{apo}$ increasing as $\vk$ is increased.
+If $r_\mathrm{apo}$ exceeds $\sim 10\,\mathrm{kpc}$, mild triaxiality in the outer regions of the galaxy merger remnant (shown in \autoref{fig:triax}) can torque the SMBH into an orbit with some degree of tangential motion, an example of which is given in \autoref{fig:trajectory} for the $\vk=780\,\kmps$ simulation.
+For the first few orbits, the SMBH may not necessarily pass through the stellar core, though it loses energy to the surrounding medium with each oscillation.
+During the final passages however, the SMBH passes through the core, before its velocity decreases below the velocity dispersion of the core.
+The predominantly-radial motion of the SMBH rapidly degrades to a random walk as it settles to the Brownian limit from this point, and does not leave the core again (shown in the inset panel of \autoref{fig:trajectory}).
+
+\subsection{Identifying a settled SMBH}\label{ssec:settle}
+As we focus on galaxy merger remnants and the possible observational signatures caused by recoil kicks in this work, we analyse herein times after the SMBH has settled back to the centre of the galaxy.
+Therefore we first need to determine the time (the infall time henceforth) where this occurs. 
+The stellar centre of the merger remnant is found using the shrinking sphere method \citep{power2003}: we use this method herein for our analysis.
+We begin by estimating the expected Brownian velocity of an SMBH in a distribution of stellar and DM particles of lower unit mass ($m_\star, m_\mathrm{DM} \ll M_\bullet$) following \citet{merritt2007}.
+We determine the root mean square velocity of the massive particle as
+\begin{equation}
+    V_\mathrm{RMS} = \sqrt{3 \xi} \tilde{\sigma},
+\end{equation}
+where $\tilde{\sigma}$ is the average velocity dispersion of the stellar and DM particles, and $\xi$ is the mass ratio to the SMBH, determined:
+\begin{equation}
+    \xi = \frac{\int n(m)m^2\dd{m}}{\int n(m)m\dd{m}} \Big / M_\bullet,
+\end{equation}
+where $n(m)$ is the number density of particles with masses $m$ to $m+\dd{m}$.
+Our calculated value of $V_\mathrm{RMS}$ that we use to define a settled velocity is $V_\mathrm{RMS} \sim 25\,\kmps$.
+
+Next, we go through the simulations output in `windows' of $100\,\mathrm{Myr}$. 
+If the SMBH stays within $0.1\,\mathrm{kpc}$ from the centre and has a median velocity less than $V_\mathrm{RMS}$ relative to the centre within the $100\,\mathrm{Myr}$ window, we consider the SMBH to be settled at the time when the velocity first falls below $V_\mathrm{RMS}$ within the time interval.
+The chosen separation from the centre is smaller than the large offsets observed by \citet{barrows2016} and slightly larger than the $\lesssim 10\,\mathrm{pc}$ offsets in \citet{lena2014}.
+
+The settling times are shown in the lower panel of \autoref{fig:infall}.
+For recoil velocities $\vk \leq 240\,\kmps$ the kick is not enough to displace the SMBH beyond the binary-scoured core and the SMBH is considered settled in less than $100\,\mathrm{Myr}$ after SMBH binary coalescence.
+As the kick velocity is increased further, the settling time begins to increase exponentially. 
+The slight deviations from an exponential increase are caused by numerical noise, which prevents the median velocity being below $V_\mathrm{RMS}$.
+
+The SMBHs which received a kick velocity larger than $1020\,\kmps$ did not settle back to the centre of the galaxy within the maximum simulation time of $3t_\mathrm{dyn} = 3\,\mathrm{Gyr}$, and already reach maximum displacements in excess of $>60\,\mathrm{kpc}$ during their first excursion from the centre.
+We expect that any SMBH with such a large displacement from its host galaxy would not be observationally identified with that galaxy, and thus we do not include simulations with $\vk>1020\,\kmps$ in this analysis.
+The used limit for maximum displacement is a few times larger than the largest suggested offset of $19.4\,\mathrm{kpc}$ from observations \citep{barrows2016}.
+We do however analyse the $\vk=2000\,\kmps$ simulation, which is above the escape velocity of the galaxy, as a proxy for an SMBH that is instantaneously removed from the simulation.
+Importantly, an escaping SMBH never completes a single oscillation about the centre of the galaxy.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figures/infall_times_and_apocenters.pdf}
+    \caption{ \textit{Top panel}: the distance of first apocentre passage from the centre of the galaxy for each kick velocity that resulted in a settled SMBH. 
+    The horizontal dashed line, $\rbbin$, marks the size of the stellar core the SMBH binary scours before coalescence, and the vertical dashed line, $\sigma_\star(r<\rbbin)$, is the one-dimensional velocity dispersion of the stellar particles interior to the binary-scoured core radius.
+\textit{Bottom panel}: the settling times for each kick velocity with $\vk>0\,\kmps$.
+The time increases exponentially as  $\vk$ increases.
+    }
+    \label{fig:infall}
+\end{figure}
+
+\subsection{Movement of stellar mass by a settling SMBH}\label{ssec:movement}
+We consider the amount of stellar mass that is bound to the recoiling SMBH, $M_{\star,\mathrm{bound}}$, at different times following coalescence.
+A stellar particle is defined to be bound to the SMBH if it is within the influence radius $r_\mathrm{infl}$ of the SMBH, which satisfies \citep[e.g.][]{merritt2013}:
+\begin{equation}
+    M_\star(r < r_\mathrm{infl}) = 2M_\bullet,
+\end{equation}
+and the binding energy of the stellar particle $K-U$, where $K$ is the kinetic energy and $U$ is the potential energy (in the SMBH frame of reference), is less than zero.
+We determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$ at the time of SMBH coalescence.
+At each pericentre passage of the oscillating SMBH (where the SMBH velocity is maximal), $M_{\star,\mathrm{bound}}$ is minimal; conversely, $M_{\star,\mathrm{bound}}$ is maximal at apocentre, irrespective of $\vk$.
+Increasing $\vk$ however reduces the magnitude of $M_{\star,\mathrm{bound}}$ compared to lower values of $\vk$.
+Following coalescence for the $\vk=60\,\kmps$ case (the lowest, non-zero recoil velocity simulated), $M_{\star,\mathrm{bound}} \sim 2.18\times10^9\,\Msun$, or roughly 37 per cent of the coalesced SMBH mass and less than 2 per cent of the total stellar mass.
+By $\vk=1020\,\kmps$, $M_{\star,\mathrm{bound}}$ has decreased to 0.2 per cent of the coalesced SMBH mass.
+These results are in close quantitative agreement with the findings of \citet{gualandris2008} and agree well with those of \citet{nasim2021b}, despite different initial conditions being used between our study and those in the literature.
+We hypothesise that the movement of stellar mass bound to the SMBH is not the dominant mechanism by which the stellar mass distribution changes, and investigate this further in \autoref{ssec:missmass}.
+
+The core region of the galaxy constantly evolves while the SMBH moves through the galaxy. 
+%Furthermore, it is important to ascertain that the distribution of stars does not drastically change after the times chosen as the focus of the analysis. 
+We inspect how the Lagrangian radius relative to the centre of mass and enclosing a stellar mass equal to the SMBH mass, $r(M_\star = M_\bullet)$, evolves with time in \autoref{fig:lagrangian} until the SMBH is settled.
+In addition to kick velocities $\vk \leq1020\,\kmps$, the kick velocity $\vk=2000\,\kmps$ is included. 
+As the SMBH with $\vk=2000\,\kmps$ is completely removed from the galaxy and does not settle back to the centre, the evolution is shown for $\sim 2.0\,\mathrm{Gyr}$ following SMBH binary coalescence.
+
+During the first $\sim 30\,\mathrm{Myr}$, the initial evolution of the Lagrangian radii is very similar for kick velocities $\vk> 300\,\kmps$.
+We can understand this from the core-crossing time argument, where if
+\begin{equation}
+    \frac{\vk}{\sigma_{\star,0}} > 1 \implies \frac{t_\mathrm{cross, \bullet}}{t_\mathrm{cross, core}} < 1.
+\end{equation}
+Consequently, for $\vk>\sigma_{\star,0}$, the SMBH is moving faster than the stellar core is able to react to the rapid change in potential caused by the recoiling SMBH.
+For these simulations, the Lagrangian radii initially increases sharply to $\sim1.4\,\mathrm{kpc}$ and the SMBH leaves the core almost instantly. 
+This large rise occurs during the first oscillation. 
+The trend which the Lagrangian radius follows at this phase is caused by the reaction of the core region to the changed gravitational potential \citep{boylan2004}.
+Each simulation shifts away from the trend when the SMBH returns to the centre.
+For $\vk<\sigma_{\star,0}$, the SMBH settles back to within 100 pc from the centre before the Lagrangian radius can reach $\sim1.4\,\mathrm{kpc}$.
+
+The higher recoil kick velocities in \autoref{fig:lagrangian} show that the Lagrangian radii can further increase at later times in addition to the initial increase, as long as the SMBH is kicked beyond the core.
+The increase is larger for higher kick velocities, with the Lagrangian radius in the simulation with $\vk=1020\,\kmps$ reaching nearly $1.8\,\mathrm{kpc}$.
+As was seen from \autoref{fig:trajectory}, the SMBH can miss the core region entirely during the first few orbits.
+The Lagrangian radius is only affected when the SMBH passes through the core, or very near to it. 
+For each of the simulations with $\sigma_{\star,0} < \vk \leq 1020\,\kmps$, the final few orbits will always include a passage through the core region.
+The SMBHs which received larger kick velocities possess more kinetic energy compared to those SMBHs which received smaller kick velocities and will cross the core region more times.
+Since the SMBH struggles to leave the core region during the final oscillations, the kinetic energy heats up the stellar particles within the central region, causing the additional growth of the Lagrangian radius.
+
+Importantly, the Lagrangian radius for each $\vk$ stays largely constant for times after $t_\mathrm{settle}$.
+Thus, the core region of the galaxy has had enough time to react to the effect the kicked SMBH has caused at the moment when the SMBH settles at the centre.
+Based on this, the state of the galaxy merger remnant  at $t=t_\mathrm{settle}$ captures the full effect which the kicked SMBH has on the galaxy.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/lagrangian_radii.pdf}
+    \caption{ 
+    The radius from the centre of the galaxy which encloses a stellar mass equivalent to the mass of the SMBH as a function of time for each kick velocity. 
+    Recoil velocities from $0\,\kmps$ to $1020\,\kmps$ are shown with a continuous colour scale, and the $2000\,\kmps$ simulation is shown in red.
+For each recoil velocity, the Lagrangian radius is shown until the SMBH has settled.
+The evolution is shown until $t\sim2.0\,\mathrm{Gyr}$ for the SMBH ejected from the galaxy with a $2000\, \kmps$ kick.
+    During the first $\sim 30\,\mathrm{Myr}$, the Lagrangian radius increases with larger kick velocities resulting in larger growth of the core.
+For the kick velocities that manage to generate multiple oscillations from the centre, the Lagrangian radius increases further when the SMBH travels through the core region during its final few orbits.
+    }
+    \label{fig:lagrangian}
+\end{figure}
+
+\subsection{Relaxation time}\label{ssec:relax}
+As already discussed, the time required for the kicked SMBH to return and settle at the centre of the merger remnant in general increases with increasing recoil velocity.
+To ensure that our results are robust to the variable ending time of the simulations, and are not a manifestation of relaxation processes, we determine the relaxation time of the $\vk=0\,\kmps$ system at different radii.
+We conservatively estimate the relaxation time $t_\mathrm{relax}$ at a radius $r$ as \citep{binney2008}:
 \begin{equation}\label{eq:relax}
@@
 where $\langle v^2 \rangle$ is the mean squared particle velocity within $r$. 
-The relaxation time increases quadratically with increasing $r$, and we find for a radius of $r=0.1\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 0.1 \, \mathrm{Gyr}$, thus motivating our choice of $0.075\,\mathrm{Gyr}$ for the analysis time.
+The relaxation time increases quadratically with increasing $r$, and we find for a radius of $r=1.0\,\mathrm{kpc}$ the relaxation time to be $t_\mathrm{relax} \sim 5.0 \, \mathrm{Gyr}$; this increases to $\sim 100\,\mathrm{Gyr}$ by $r=5\,\mathrm{kpc}$.
+Of the simulations where the SMBH was determined to have settled, the maximum time for this to occur was $\sim 2.5\,\mathrm{Gyr}$, well below the relaxation time of the system. 
 We  note that \autoref{eq:relax} does not take into account gravitational softening, and thus provides a conservative lower bound for the relaxation time.
-This is critical as the \ketju{} region does not include gravitational softening between stellar particles and the SMBH, so is a not fully-softened simulation.
-
-For all simulations, interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, DM with a softening length of $\varepsilon = 300\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
-
-% RESULTS
-\section{Results}\label{sec:results}
-
-\subsection{Infall time}
-\drafting{
-    \textbf{Atte's section}. Would be good to have:
-    \begin{enumerate}
-        \item Plot of infall time (i.e., time for stability criterion to be reached) as a function of kick velocity
-        \item Number of oscillations before settling?
-        \item Plot of maximum displacement of kicked BH as a function of kick velocity. How does this compare to core size? Which simulations have BH kicked beyond the core?
-    \end{enumerate}
-}
-\drafting{
-    Include this text here, this is how we identify `settled' systems, but we're not really interested in when the SMBH has settled for the rest of the paper (refer to Nasim where the majority of the GW-induced scouring occurs during the first excursion of the SMBH).
-    Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
-    The time to this point, from the time of SMBH binary coalescence, we term the `infall time'.
-    Those simulations where the maximum displacement of the kicked SMBH exceeds $30\,\mathrm{kpc}$ are not included in the analysis, as we expect that observationally a SMBH further than this distance from the centre of its host galaxy would not be identified with that galaxy.
-}
-
-\drafting{
-    Also maybe we can discuss the number of bound particles to the BH here. I've already looked at this, but we can talk about combining that information with the apocentre plots. 
-    It seems that increasing $\vk$ decreases the amount of bound mass, in agreement with previous studies, but also for each $\vk$, the maximum amount of stellar mass bound to the BH is maximal at apocentre.
-}
+This is critical as the \ketju{} region does not include gravitational softening between stellar particles and the SMBH, so it is not a fully-softened simulation.
+In practice, we expect the relaxation time to be even longer than reported due to the inclusion of gravitational softening in the simulation for particles outside of the \ketju{} region.
+Thus, for the analysis herein, we are confident that the results we observe are directly caused by the influence of the recoiling SMBH on the galaxy, and are not due to relaxation effects.
+
+\section{Stellar mass density profiles}\label{sec:density}
+\subsection{Three dimensional mass density profiles}
+
 \begin{figure}
     \centering
-    \includegraphics[width=0.4\textwidth]{infall_time}
-    \caption{Plot of infall time (time to stability criterion) as a function of kick velocity.}
-    \label{fig:infall}
-\end{figure}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\textwidth]{max_disp}
-    \caption{Plot of the maximum BH displacement as a function of kick velocity. Where do the displacement (red) and core size (green) lines intersect?}
-    \label{fig:max_disp}
-\end{figure}
-
-\subsection{Mass density profiles in three dimensions}
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\textwidth]{density_3d}
+    \includegraphics[width=0.5\textwidth]{density_3d}
     \caption{
         Three dimensional stellar mass density, with the line colour corresponding to the kick velocity of the SMBH. 
-        All simulations with $\vk \geq 0.5v_\mathrm{esc}=900\,\kmps$ share the same colour, as there is little variation in the profiles for these kick velocities.
-\drafting{Should probably update this if max is 1020, maybe add special colour for 2000?}.
-        At radii $r\lesssim3\,\mathrm{kpc}$, the projected stellar mass density decreases smoothly with increasing kick velocity, with the exception of the $\vk=2000\,\kmps$ simulation, in which the SMBH has not returned to the centre of the merger remnant.
-        At radii $r>3\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity only affects the inner mass distribution of the galaxy merger remnant.
+        At radii $r\lesssim3\,\mathrm{kpc}$, the stellar mass density decreases smoothly with increasing kick velocity, with the exception of the $\vk=2000\,\kmps$ simulation, in which the SMBH has not returned to the centre of the merger remnant.
+        A density cusp is apparent for all mergers where the SMBH settled at the centre.
+        At radii $r>3\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity has a larger impact on the inner mass distribution of the galaxy merger remnant.
         }
@@
 \end{figure}
+
 Before considering the projected, 2D stellar mass density profiles of the merger remnants, we investigate the 3D stellar mass distribution.
-This alleviates the influence of projection effects on inferring if a core has formed during the BH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
-\drafting{Maybe move the following two sentences to Atte's section
-To centre the data herein, we use the shrinking sphere method \citep{power2003} to determine the stellar centre of mass, and find the time evolution of the CoM does not systematically vary with kick velocity.
-Consequently, we are confident that our results are not a result of malign centring.}
-For the $\vk=0\,\kmps$ case, \drafting{check} we determine the SMBH influence radius to be $r_\mathrm{infl}\simeq 1.5\,\mathrm{kpc}$.
-
-In \autoref{fig:density3d}, we observe a clear decrease in stellar mass density at radii $r\lesssim3\,\mathrm{kpc}$ with increasing kick velocity.
-An exception to this trend is the highest kick velocity we simulated, $\vk=2000\,\kmps$, which shows an excess of stellar mass between $0.3 \lesssim r/\mathrm{kpc} \lesssim 2$ relative to the $vk=1020\,\kmps$ simulation.
-Recalling that the time of the analysis is performed when the remnant SMBH has settled, and thus occurs at increasingly later times with increasing kick velocity, SMBHs kicked with a larger $\vk$ have undergone more oscillations to return the SMBH to an equilibrium state compared to SMBHs with a lower $\vk$.
-An increased number of oscillations presents an enhanced ability to remove stellar mass from the centre, as reflected in the density profiles in \autoref{fig:density3d}.
-Similarly, the $\vk=2000\,\kmps$ case, in which the SMBH has not undergone any oscillations in the centre of the galaxy merger remnant, has not had an opportunity to evacuate as much mass from the central regions as, for example, the $\vk=1020\,\kmps$ has.
-Conversely, the mass deficit in the $\vk=2000\,\kmps$ is greater than that exhibited by all kick velocities up to and including $\vk=480\,\kmps$.
-This observation is in agreement with \citet{nasim2021b}, in that the major contributor to a central mass deficit is the initial excursion of the recoiling SMBH, and the subsequent oscillations playing a subdominant role in the enhancement of the stellar core.
-
-The primary result of the three dimensional density profile is the observation of a recoil velocity dependent impact on the stellar density structure, which we can take as a ground truth when we consider the galaxy merger remnants from a more observationally-motivated perspective.
-We hypothesise that the difference between merger remnants in three dimensional space will manifest in other methods of observing the galaxy.
+This alleviates the influence of projection effects on inferring if a core has formed during the SMBH binary merger process, and gives insight into the dependence of how matter is spatially distributed as a function of the recoiling SMBH kick velocity.
+
+In \autoref{fig:density3d}, we observe a clear decrease in stellar mass density at radii $r\lesssim2\,\mathrm{kpc}$ with increasing kick velocity.
+For $\vk=0\,\kmps$, the central stellar density is $\sim10^{10}\,\Msun\,\mathrm{kpc}^{-3}$, whereas for $\vk=1020\,\kmps$ it is $\sim10^{9}\,\Msun\,\mathrm{kpc}^{-3}$.
+An exception to this trend is the highest kick velocity we simulated, $\vk=2000\,\kmps$, which shows an excess of stellar mass between $0.3 \lesssim r/\mathrm{kpc} \lesssim 2$ relative to the simulations with $\vk \gtrsim 300\,\kmps$.
+This is readily explained by the evolution of the Lagrangian radii in \autoref{fig:lagrangian}.
+The final few orbits of the settling SMBH in simulations with $\vk>300\,\kmps$ take the SMBH through the stellar core, enlarging it further with each oscillation. 
+As a result, the central stellar density decreases as the number of oscillations increase. 
+The $\vk=2000\,\kmps$ simulation, in which the SMBH does not oscillate through the core, has a central stellar density that decreases only due to the initial response of the stellar particles to the suddenly-changed potential.
+The simulations with a settled SMBH have a density profile within $r<1\,\mathrm{kpc}$ that follows a power law $\rho \propto r^{-\nu}$ with $0<\nu<1$.
+Conversely, the $\vk=2000\,\kmps$ simulation has a flat density profile with $\nu=0$ in the centre, in agreement with the recent work by \citet{khonji2024}.
+
+In the outer regions $r\gtrsim 4\,\mathrm{kpc}$, the stellar density falls with increasing radius consistently between all simulations, indicating that the dynamical effect of a recoiling SMBH on the stellar mass distribution is predominantly confined to the central regions of the galaxy.
 
@@
 Single-line circle nodes represent fit parameters, double-line circles represent measured quantities (the data), and diamond nodes represent deterministic quantities.
-The particular distribution connecting nodes is written below the corresponding black square, with a subscript `$\mathrm{T}[l,u]$' indicating a distribution $f(\lambda)$ truncated to $l<\lambda<u$, and the subscript `likelihood' indicating the likelihood function.
-The $R$ box indicates variables fit for each radial bin, and the projection box indicates variables specific to each projection realisation.
+The particular distribution connecting nodes is written below the corresponding black square, with a subscript `$\mathrm{T}[l,u]$' denoting a distribution $f(\lambda)$ truncated to $l<\lambda<u$, and the subscript `likelihood' indicating the likelihood function.
+The $R$ box indicates variables fit for each radial bin, and the projection box distinguishes variables specific to each projection realisation.
 Note that we additionally use $\hat{\Sigma}$ to distinguish the calculated value of surface density from the measured value $\Sigma$.
@@
     \caption{
-        Surface density profiles with 25\% Bayesian HDI for select representative runs.
+        Surface density profiles with 25 per cent Bayesian HDI for select representative runs.
         Increasing the kick velocity induces a shallower density profile in the inner regions, consistent with the three dimensional density maps in \autoref{fig:density3d}.
@@
     \caption{
-        Bayesian estimate of the merger remnant core size $r_b$, scaled to the core size of the pre-merger remnant $r_{\mathrm{b},0}$ (left axis) and in physical units (right axis), as a function of kick velocity $\vk$.
-        The core size distributions are shown as box plots, with the median core size indicated by the central mark.
+        \textit{Top panel}: Bayesian estimate of the merger remnant core size $r_\mathrm{b}$, scaled to the core size scoured solely by the SMBH binary prior to coalescence $\rbbin$ (left axis) and in physical units (right axis), as a function of kick velocity $\vk$.
+        The core size distributions are shown as box plots, with the median core size indicated by the central mark and shaded regions corresponding to the IQR.
         Larger kick velocities are correlated with larger core sizes.
 Additionally, a greater spread in the distribution of core sizes over different viewing projections of the merger remnant is associated with larger kick velocities.
-        Three models, fit to the median $\vk$--$\rb$ trend, are overplotted.
+        Three models, fit to the median $\rb$--$\vk$ trend, are overplotted.
         For the models, we do not show the parameter uncertainty for visual clarity.
+        \textit{Bottom panel}: Mass deficit $M_\mathrm{def}$ between S\'ersic and core-S\'ersic fits as a function of the recoil velocity scaled to the mass deficit due to SMBH binary scouring ($M_{\mathrm{def},0}$, left axis) and physical units (right axis).
         }
@@
 
-To facilitate comparisons with observations \citep[e.g.][]{graham2003,rusli2013}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projection of each simulated merger remnants:
+To facilitate comparisons with observations \citep[e.g.][]{graham2003,rusli2013}, we fit the six-parameter core-S\'ersic profile \citep{graham2003} to the projected stellar mass density of each simulated merger remnant:
 \begin{equation}\label{eq:cs}
-    \Sigma(R) = \Sigma' \left[1 + \left(\frac{\rb}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[-b \left(\frac{R^\alpha + \rb^\alpha}{\Reff^\alpha}\right)^{(1/\alpha n)}\right]
+    \Sigma(R) = \Sigma' \left[1 + \left(\frac{\rb}{R}\right)^\alpha \right]^{\gamma/\alpha} \exp\left[-b_n \left(\frac{R^\alpha + \rb^\alpha}{\Reff^\alpha}\right)^{(1/\alpha n)}\right]
 \end{equation}
 where
-\begin{equation}
-    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[b\left(2^{1/a} \frac{\rb}{\Reff}\right)^{1/n}\right].
-\end{equation}
-Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $n$ is the S\'ersic index, $b$ is estimated as $b\simeq 2n - 1/3 + 0.009876/n$ \citep{prugniel1997}, and $\alpha$ is the profile transition index.
+\begin{equation}\label{eq:cs_prefactor}
+    \Sigma' = \Sigma_\mathrm{b} 2^{-\gamma/\alpha} \exp\left[b_n \left(2^{1/\alpha} \frac{\rb}{\Reff}\right)^{1/n}\right].
+\end{equation}
+Here $\rb$ is the core radius, $\Sigma_\mathrm{b}$ is the projected surface density at the core radius, $\gamma$ is the core slope, $\Reff$ is the effective (half-light) radius\footnote{As we assume a constant mass-to-light ratio, the half-light radius is equivalent to the projected half-mass radius.}, $n$ is the S\'ersic index, $b_n$ is estimated as $b_n\simeq 2n - 1/3 + 0.009876/n$ \citep{prugniel1997}, and $\alpha$ is the profile transition index.
 We refer to the collective vector of these parameters as $\vb{\theta}\equiv[\rb, \Sigma_\mathrm{b}, \gamma, \Reff, n, \alpha]$.
-We view the simulated merger remnant from fifteen random angles, and use a Bayesian hierarchical model (HM) to fit the model parameters.
-As each of the projections are unordered and exchangeable (i.e., there is no distinguishing label for the projections), and each projection is of the same merger remnant, a HM naturally lends itself.
-The idea behind the HM is as such: for each projection, we infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible $\vb{\theta}$ vectors.
-Hence, we are effectively inferring the global hyperparameters $\vb{\theta}^\mathrm{hyp}$ that describe the latent parameters $\vb{\theta}$ that we are interested in.
-This idea is demonstrated in the directed acyclic graph (DAG) for the model, shown in \autoref{fig:dag}.
-
-We use weakly-constrained prior distributions for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these prior distributions are chosen to reflect reasonable values that the parameters may take (e.g.
-distributions of a radial quantity are chosen so as to be positively-constrained).
-The distributions\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$, as opposed to the variance $\sigma^2$ as is common in many statistics references.} of the hyperparameters are given in \autoref{tab:hyper}.
-
-We fit the model with \textsc{Stan} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo (HMC) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
-This resulted in 2000 draws from the posterior distribution of $\vb{\theta}$, whereby we ensure the diagnostic value $\hat{R}$ (comparison of between-chain and within-chain estimates) is less than 1.05.
-We perform additional prior sensitivity tests using the power scaling method described in \citet{kallioinen21}, ensuring that the cumulative Jensen-Shannon (CJS) divergence is less than 0.05 for the latent parameters $\vb{\theta}$ of the model.
-This recommended threshold indicates that the sampled posterior distributions of the latent parameters are not sensitive to the particular prior distributions used.
-
-In \autoref{fig:density}, we see a clear decrease in the projected stellar mass density with increasing kick velocity for four representative values of $\vk$, consistent with the results in \autoref{fig:density3d}.
-The shaded regions in \autoref{fig:density} correspond to the $25\%$ Bayesian highest density interval (HDI), where the uncertainty is a direct result of choosing different projection angles when constructing the projected densities.
-
-To discuss the difference in the evacuated core region of the merger remnants more quantitatively, we show the median (blue dashes) and interquartile range (IQR, blue boxes) of the core radius parameter $\rb$ from \autoref{eq:cs} for kick velocities $\vk\leq 1020\,\kmps$ in \autoref{fig:vkrb}.
-Our results indicate that a non-zero kick velocity always serves to enlarge the core region relative to the core formed during the SMBH binary scouring phase.
-In agreement with previous studies \citep{gualandris2008,nasim2021b}, we find a monotonic increase in the core radius $\rb$ with kick velocity.
-Additionally, we see that the IQR of $\rb$ increases with increasing $\vk$, indicating that with an increased core size, the uncertainty introduced due to projection effects also increases.
-
-As shown in the upper central panel of \autoref{fig:csparams}, the projected density at the core radius decreases with increasing kick velocity, in agreement with the decreasing core density observed in the three dimensional density profiles in \autoref{fig:density3d}.
-Similarly, the S\'ersic index also decreases with increasing kick velocity (bottom right panel of \autoref{fig:csparams}), although the overall decrease is from $n\simeq2.1$ at $\vk=0\,\kmps$ to $n\simeq1.6$ at $\vk=1020\,\kmps$.
-The other three parameters: $\Reff$, $\alpha$, and $\gamma$, all appear insensitive to the kick velocity.
-The fact that $\Reff$ and $\alpha$ are agnostic to $\vk$ is not surprising, given that these parameters control more of the larger-scale profile of \autoref{eq:cs}.
-\drafting{A constant $\gamma$ however implies that irrespective of the core size, the steepness of the core density is \dots what does it mean? Mass is iteratively removed in oscillations?}
+We view the simulated merger remnant from $N_\mathrm{proj}=15$ random angles in $N_R=21$ logarithmically-spaced bins from $0.1\,\mathrm{kpc}$ to $30\,\mathrm{kpc}$, and use a Bayesian hierarchical model to fit the model parameters, as shown in \autoref{fig:dag}.
+An excellent introduction to Bayesian data analysis, and in particular hierarchical modelling, can be found in \citet{gelman2015}, which we summarise in \autoref{sec:app_fit} of the Appendix in the context of our work.
+As part of the Bayesian inference workflow, we perform rigorous prior sensitivity analysis and posterior checking, also detailed in \autoref{sec:app_fit} of the Appendix.
 
 \begin{table}
-    \caption{Hyperparameter distributions for core-S\'ersic model. $\mathcal{N}$ indicates a normal distribution, $\mathrm{Exp}(\lambda)$ indicates an exponential distribution.}
+    \caption{Hyperparameter distributions for core-S\'ersic model.
+The distributions are either a normal ($\mathcal{N}$), exponential ($\mathrm{Exp}$), or gamma ($\mathrm{Gamma}$) distribution.
+The rationale for each distribution is discussed in \autoref{sec:app_fit} of the Appendix.}
     \label{tab:hyper}
@@
 
-\subsection{Predicted core size distribution}
-To predict the distribution of core sizes given our merger remnant of two massive ETGs, we require a mapping from the kick velocity to the core radius. 
+In \autoref{fig:density}, we see a clear decrease in the projected stellar mass density with increasing kick velocity for four representative values of $\vk$, consistent with the results in \autoref{fig:density3d}.
+The shaded regions in \autoref{fig:density} correspond to the $25$ per cent Bayesian highest density interval (HDI), where the uncertainty is a direct result of choosing different projection angles when constructing the projected densities.
+
+To discuss the difference in the evacuated core region of the merger remnants more quantitatively, we show the median (blue dashes) and interquartile range (IQR, blue boxes) of the core radius parameter $\rb$ from \autoref{eq:cs} for kick velocities $\vk\leq 1020\,\kmps$ in \autoref{fig:vkrb}, in units of both kpc and scaled to the $\vk=0\,\kmps$ core radius $\rbbin$, which corresponds to the pre-merger core radius of the galaxy.
+Our results indicate that a non-zero kick velocity where the SMBH returns to the centre of the merger remnant serves to enlarge the core region relative to the core formed during the SMBH binary scouring phase.
+In agreement with previous studies, we find a monotonic increase in the core radius $\rb$ with kick velocity.
+Additionally, we see that the IQR of $\rb$ increases with increasing $\vk$, indicating that with an increased core size, the uncertainty introduced due to projection effects also increases.
+To test the significance of the difference in distributions of $\rb$ between the $\vk=0\,\kmps$ case and the $\vk=1020\,\kmps$ case, we calculate the probability of superiority \citep[PS,][]{grissom2001} as:
+\begin{equation}
+    \mathrm{PS} = \frac{1}{n} \sum_i \mathbb{I}_{\rbbin^{(i)} < r_{\mathrm{b},1020}^{(i)}}(i),
+\end{equation}
+where the index $i$ runs over the $n$ posterior samples, and the indicator function $\mathbb{I}_A$ maps the $i^\mathrm{th}$ posterior draw to 1 if it belongs to the subset $A$, and to 0 otherwise.
+With this measure we find that in more than 87 per cent of the posterior draws, $\rb$ from the $\vk=1020\,\kmps$ case is greater than $\rb$ from the $\vk=0\,\kmps$ case; the distributions are distinct.
+
+As shown in the top right panel of \autoref{fig:csparams} in the Appendix, the projected density at the core radius decreases with increasing kick velocity, in agreement with the decreasing core density observed in the three dimensional density profiles in \autoref{fig:density3d}.
+Similarly, the S\'ersic index $n$ also decreases with increasing kick velocity (bottom right panel of \autoref{fig:csparams}).
+For recoil velocities $\vk\leq 300\,\kmps$, $n\simeq2.1$, and for $\vk\geq 660\,\kmps$, $n\simeq 1.6$.
+For recoil velocities $360\,\kmps \leq \vk \leq 600\,\kmps$, the decrease in $n$ appears linear.
+We note that the recoil velocity for which $n$ begins to decrease coincides closely with the stellar velocity dispersion of the SMBH binary-scoured core, $\sigma_{\star,0} \sim 270\,\kmps$.
+The variation in $n$ can be understood in the context of the SMBH leaving the core, and thus affecting the stellar mass distribution beyond the core as it oscillates back to its equilibrium state.
+Conversely, the effective radius $\Reff$ appears insensitive to the SMBH recoil velocity.
+
+For the inner slope parameter $\gamma$, a weak trend to lower values at higher recoil velocities is apparent, after peaking at $\vk\sim 300\,\kmps \sim \sigma_{\star,0}$, though the distributions for $\gamma$ are broad.
+Recently, \citet{khonji2024} demonstrated that increasing SMBH recoil velocity resulted in a suppressed $\gamma$, however the authors did not capture the $\vk \sim \sigma_{\star,0}$ regime of recoil velocities.
+We understand the broad distributions that we observe for $\gamma$ as arising from the collapse of a three-dimensional density distribution to a two-dimensional representation, and then enforcing a spherically-symmetric one-dimensional profile to describe the two-dimensional surface density.
+This is supported by the IQR of the $\gamma$ parameter remaining relatively broad with irrespective of kick velocity, unlike the core radius $\rb$ for example.
+
+To investigate this further, for a subset of simulations we take particles in the centre of the merger remnant within a spherical annulus defined by $0.9\,\mathrm{kpc} \leq r \leq 1.0\,\mathrm{kpc}$ and consider sixty different, random projections into the two-dimensional plane.
+We perform a singular value decomposition (SVD) to obtain an ellipse representation of the projected annulus, and determine the ellipticity $\varepsilon$ of the projection
+\begin{equation}
+    \varepsilon = \frac{a-b}{a},
+\end{equation}
+where $a$ and $b$ are the semimajor and semiminor axes of the ellipse, respectively.
+A value of $\varepsilon \neq 0$ indicates a departure from circular symmetry.
+We find the distribution of $\varepsilon$ is positively-constrained and peaks at  $\varepsilon \simeq 0$ but is right-skewed, with the $\ordinal{75}$ percentile of $\varepsilon$ typically about $\sim0.05$.
+This indicates that asymmetries are present in the projected stellar mass density profiles.
+Non-circularity of the core region complicates using a radially-dependent density profile parameterised by a constant exponent $\gamma$: a range of possible values for $\gamma$ could describe the core region comparatively well.
+The difficulty in constraining $\gamma$ provides good motivation for determining distributions for the parameters in \autoref{eq:cs} with Bayesian techniques, as opposed to point estimates achieved through typical frequentist methods.
+
+\subsection{Estimating the amount of missing stellar mass}\label{ssec:missmass}
+
+We continue the quantitative analysis of the core region by calculating the missing stellar mass of the merger remnants. 
+As well as the core radius $\rb$, the mass deficit $M_{\mathrm{def}}$ can also be used to measure the core size.
+Past studies have found the mass deficit to correlate with the combined mass of the SMBH binary and the number of major `dry' mergers \citep[e.g.][]{milosavljevic2001, graham2003, merritt2006,rantala2024}.
+%Past studies have found the mass deficit to be proportional to the mass of the binary approximately $M_{\mathrm{def}}\sim 0.5 -- 5 M_{BH}$ for a single merger. 
+In order to determine the mass displaced by the gravitational interactions and recoil, we adopt an observational approach by comparing the stellar mass density profiles of the inward extrapolated S\'ersic $\Sigma_{\mathrm{s}}(R)$ \citep{Sersic1963,Sersic1968} and the core-S\'ersic $\Sigma(R)$ fits  \citep[e.g.][]{dullo2014,bonfini2016, dullo2019, nasim2021b} as:
+\begin{equation}
+    M_{\mathrm{def}}=2\pi \Upsilon \int_0^\infty \left[ \Sigma_{\mathrm{s}}(R)-\Sigma(R)\right] R \dd{R},
+\end{equation}
+where the mass-to-light ratio is chosen to be $\Upsilon=3.5 \, \Msun / \Lsun$ \citep{bonfini2016}.
+The core-S\'ersic profile is calculated using \autoref{eq:cs} and \autoref{eq:cs_prefactor}, while the S\'ersic profile can be obtained using the same set of equations with a slight modification.
+By setting $\rb = 0$ in \autoref{eq:cs}, the profile becomes the standard S\'ersic profile:
+\begin{equation}
+    \Sigma_{\mathrm{s}} = \Sigma ' \exp \left( -b_n (R/R_{\mathrm{e}})^{1/n} \right).
+\end{equation}
+To achieve the right normalisation, it is important to let $\rb$ retain its value in \autoref{eq:cs_prefactor} when constructing the S\'ersic profile \citep{graham2003}. 
+In fitting the luminosity profiles, we perform for each kick velocity $10^4$ Monte Carlo samplings of the posterior distributions of the parameters $\vb{\theta}\equiv[\rb, \Sigma_\mathrm{b}, \gamma, \Reff, n, \alpha]$ described earlier in \autoref{ssec:projdens}, hence giving us $10^4$ realisations of the quantity $M_\mathrm{def}$.
+
+The mass deficits as a function of kick velocities $\vk\leq 1020\,\kmps$ are shown in the lower panel of \autoref{fig:vkrb}.
+At low kick velocities, $\vk\leq 240\,\kmps$, we see a slow monotonic increase of mass deficit as a function of the kick velocity.
+Although the scatter grows at higher kick velocities, an increase of the median trend can be observed over the whole kick velocity range.
+For high $\vk$ we see a median trend of $M_\mathrm{def} \sim 3 M_{\mathrm{def},0}$, where $M_{\mathrm{def},0}$ corresponds to the mass deficit from binary scouring alone.
+We find that the median mass deficit due to SMBH binary scouring prior to the coalescence of the SMBHs is $M_{\mathrm{def},0}=5.11\times10^9\,\Msun$, roughly $0.87M_\bullet$.
+As a consequence of the mass deficit growing to $M_\mathrm{def} > M_\bullet$ for non-zero recoil velocities, as well as higher recoil velocities resulting in larger $M_\mathrm{def}$, we confirm our earlier hypothesis in \autoref{ssec:movement} that stellar mass bound to the SMBH (at most $0.37M_\bullet$) is not the primary mechanism by which the core grows.
+
+Although \citet{merritt2006} and \citet{nasim2021b} found smaller mass deficit values from binary scouring alone for single mergers ($M_{\mathrm{def},0}\simeq 0.5\, M_\bullet$ and $M_{\mathrm{def},0}\simeq 0.26\,M_\bullet$, respectively), our value is in agreement with other previous studies.
+\citet{gualandris2008} reported a $M_{\mathrm{def},0}\sim M_\bullet$ relation, which is comparable to our value.
+In addition, similar mass deficits have also been found by \citet{rantala2024}.
+Slight discrepancies can be attributed to the differences in initial conditions and to varying approaches to computing mass deficits.
+Additionally, \citet{gualandris2008} use 3D density profiles, while we utilise projected profiles to make our results more comparable to observational data.
+
+%\citet{merritt2006} found
+%Although slightly smaller, our value is in good agreement with 26 per cent from the first generation equal mass merger of \citet{nasim2021b} (see merger `A1' in their Table 4).
+Slight differences can be attributed to the differences in initial conditions.
+
+%\citet{merritt2006} and \citet{gualandris2008} found larger mass deficit values from binary scouring alone for single mergers: $M_{\mathrm{def},0}\simeq 0.5\, M_\bullet$ and $M_{\mathrm{def},0}\sim M_\bullet$, respectively) \sonja{Should I add Rantala2024 $M_{\mathrm{def},0}\sim (0.75-2)\, M_\bullet$ to this list?
+Their approach to Mdef is very different.}.
+Although differing initial conditions already cause discrepancies of the missing mass values, varying approaches to computing mass deficits may also contribute to variations.
+Additionally, \citet{gualandris2008} use 3D density profiles, while we utilize projected profiles to make our results more comparable to observational data.
+
+As for core sizes presented in \autoref{fig:vkrb}, the interquartile ranges of the missing mass broaden with increasing kick velocities.
+This is due to growing uncertainties in the projection effects, even though the parameters $\vb{\theta}$ from the core-S\'ersic fits are Monte Carlo sampled as discussed at the end of \autoref{ssec:projdens}.
+Also, we have found a positive correlation between mass deficit and the core radius.
+Thus, the increasingly large range of $\rb$ values introduce a wider range of mass deficits as a function of kick velocity.
+Additionally, we note that the choice of the value of the mass-to-light ratio impacts the resulting mass deficits.
+For example, by using $\Upsilon=4.5 \, \Msun / \Lsun$ (instead of $\Upsilon=3.5 \, \Msun / \Lsun$) the value of the binary scouring alone increases to $\sim 112$ per cent of the total SMBH binary mass.
+
+\subsection{Predicted core size distribution}\label{ssec:predcore}
+To predict the distribution of expected stellar core sizes given our merger remnant of two massive elliptical galaxies, we require a mapping from the kick velocity to the core radius. 
 Let us define a scaled kick velocity $v'=\vk/v_\mathrm{esc}$, where $v_\mathrm{esc}=1800\,\kmps$.
@@
 \begin{equation}\label{eq:vkrb_exp}
-    \frac{\rb}{r_{\mathrm{b},0}} = K v'^\beta + 1
-\end{equation}
-where $K$ and $\beta$ are free parameters.
-\autoref{eq:vkrb_exp} gives that the core radius monotonically increases irrespective of kick velocity, implying that as $\vk\rightarrow\infty$, $\rb\rightarrow\infty$.
+    \frac{\rb}{\rbbin} = K v'^q + 1,
+\end{equation}
+where $K$ and $q$ are free parameters.
+\autoref{eq:vkrb_exp} gives that the core radius monotonically increases irrespective of kick velocity, na\"ively implying that as $\vk\rightarrow\infty$, $\rb\rightarrow\infty$.
 
@@
 \begin{equation}\label{eq:vkrb_lin}
-    \frac{\rb}{r_{\mathrm{b},0}} = a v' + b,
+    \frac{\rb}{\rbbin} = K v' + 1,
 \end{equation}
@@
 \begin{equation}\label{eq:vkrb_sig}
-    \frac{\rb}{r_{\mathrm{b},0}} = K \left( 1 - e^{-\beta v'} \right) + c.
-\end{equation}
-
-We fit each of these models to the median trend in \autoref{fig:vkrb}, and display the relation with its best fit parameters in the same figure.
+    \frac{\rb}{\rbbin} = K \left( 1 - e^{-q v'} \right) + 1.
+\end{equation}
+
+We fit each of the exponential, linear, and sigmoid models to the median $\rb$--$\vk$ trend in \autoref{fig:vkrb}, and display the relation with its best fit parameters in the same figure.
 In the fit, we assume weakly informative positively-constrained normally distributed priors on the parameters, a normally distributed scatter $\tau$ truncated to $\tau>0$, and a Gaussian likelihood function.
 The fit is found using the same Bayesian methods as in \autoref{ssec:projdens} (including posterior checking and prior sensitivity analysis), allowing for uncertainties to be estimated for the model fits.
-The medians, IQRs, and constraints of the relation parameters in \autoref{eq:vkrb_exp}-\autoref{eq:vkrb_sig} are given in \autoref{tab:vkrb}.
+The medians and 68 per cent HDI interval of the relation parameters in \autoref{eq:vkrb_exp}-\autoref{eq:vkrb_sig} are given in \autoref{tab:vkrb}.
 
@@
         \hline
-        Model & Parameter & Median & IQR \\
+        Model & Parameter & Median & 68\% HDI \\
         \hline
-        Exponential & $K$ & 2.90 & 0.37 \\
-        & $\beta$ & 0.78 & 0.13 \\
+        Exponential & $K$ & 2.72 & [2.19, 3.34] \\
+        & $q$ & 0.78 & [0.57, 0.99] \\
         \hline
-        Linear & $a$ & 3.26 & 0.41 \\
-        & $b$ & 1.1 & 0.14 \\
+        Linear & $K$ & 3.03 & [2.39, 3.64] \\
         \hline
-        Sigmoid & $K$ & 2.47 & 0.59 \\
-        & $\beta$ & 2.62 & 1.31 \\
-        & $c$ & 0.87 & 0.16 \\
+        Sigmoid & $K$ & 3.04 & [1.66, 5.73] \\
+        & $q$ & 1.56 & [0.51, 3.07] \\
         \hline
@@
 
-Using the three $\vk$--$\rb$ relations, we wish to construct the posterior predictive distributions for new data $\tilde{\rb}$ given the observed values of $\rb$:
-\begin{equation}
+Using the three $\rb$--$\vk$ relations, we wish to construct the posterior predictive distributions for new data $\tilde{\rb}$ given the observed values of $\rb$:
+\begin{equation}\label{eq:ppd}
     p(\tilde{\rb}|\rb) = \int p(\tilde{\rb} | \vb{\kappa}) p(\vb{\kappa} | \rb) \dd{\vb{\kappa}},
 \end{equation}
-where the integral marginalises out the respective $\vk$--$\rb$ relation parameters, collectively referred to as $\vb{\kappa}$ \citep{gelman2015}.
-The new data $\tilde{\rb}$ is generated by transformation sampling the distribution of kick velocities from \citet{zlochower2015} for dry mergers assuming random azimuthal spin alignment (their Fig.
-15, right panel), and `pushing' the values through the desired $\vk$--$\rb$ relation.
-Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}, whilst simultaneously accounting for uncertainty in the fit described by the $\vk$--$\rb$ relations.
-
-For the \citet{zlochower2015} dry merger distribution, the dimensionless spin parameter $\alpha_\bullet$ follows a beta distribution, namely:
-\begin{equation}
-    P_\mathrm{Zlochower}(\alpha_\bullet) \propto (1-\alpha_\bullet)^{4.66884-1} \alpha_\bullet^{10.5868-1}.
-\end{equation}
-The range of kick velocities predicted by the model varies from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
-An immediately apparent caveat is that the $\vk$--$\rb$ relations are constrained using kick velocities $\vk\leq1020\,\kmps$: extrapolating from these relations introduces an inherent risk when we do not know how the $\vk$--$\rb$ relation holds for very large $\vk$, which we discuss further below.
-
-We find that the mode of the forward-folded core radius distribution is similar for the exponential and linear models, with $1.04\,\mathrm{kpc}$ ($1.84\,r_{\mathrm{b},0}$) and $0.97\,\mathrm{kpc}$ ($1.71\,r_{\mathrm{b},0}$), respectively.
-Conversely, the mode for the sigmoid relation to occur at a larger value of $\rb$, namely $1.60\,\mathrm{kpc}$ ($2.83\,r_{\mathrm{b},0}$).
-Looking at the shapes of the three distribution in \autoref{fig:rb_pdf}, we see a similarity in shapes between the exponential and linear models, which are both right-skewed. 
-This is due to the ever-increasing value of $\rb$ predicted for an increasing $\vk$: higher recoil velocities produce larger cores, but as higher recoil velocities become increasingly unlikely (assuming the \citet{zlochower2015} distribution), we observe the skew in the $\rb$ probability density function (PDF).
-Conversely, the sigmoid model has a small hump prior to the mode, arising from the high likelihood of small kick velocities being pushed through the sigmoid function (\autoref{eq:vkrb_sig}) prior to the upper plateau.
-The peak in the $\rb$ distribution from the sigmoid model is a result of a large range of kick velocities (albeit with decreasing frequency) being mapped to the upper plateau of the sigmoid function, hence the correspondence between the median of the scale factor $K=2.47$ in \autoref{eq:vkrb_sig} and the distribution mode for the model, $2.83\,r_{\mathrm{b},0}$.
-
-To determine which model of the three (exponential, linear, or sigmoid) best describes the data, we use leave one out cross validation (LOO-CV) with Pareto smoothed importance sampling \citep[PSIS][]{vehtari2017}.
-PSIS LOO-CV estimates the expected log pointwise predictive density ($\widehat{\mathrm{ELPD}}$), defined
-\begin{equation}
-    \widehat{\mathrm{ELPD}} = \sum_{i=1}^N \log \int p(y_i | \theta) p(\theta | y_{-i}) \dd{\theta},
-\end{equation}
-to determine the out-of-sample predictive accuracy of a Bayesian model, i.e., the ability to predict new data $\tilde{y}$ conditioned on the fitted data $y$. 
-A larger value of $\widehat{\mathrm{ELPD}}$ indicates better predictive ability of a given model compared to some other model \citep{vehtari2017, riha2024}.
-We report the preferred model to be the sigmoid model (\autoref{eq:vkrb_sig}), with an $\widehat{\mathrm{ELPD}}$ of 4.06, compared to 2.68 and 0.54 for the exponential and linear models, respectively.
-
-All in all, assuming the \citet{zlochower2015} model to be a reasonable description of SMBH recoil velocities, a large fraction massive elliptical galaxies should have a non-negligible core, which is in agreement with observations \citep{ferrarese1994,savorgnan2016,sahu2019}, even if the recoil velocity is modest.
+where the integral marginalises out the respective $\rb$--$\vk$ relation parameters, collectively referred to as $\vb{\kappa}$ \citep{gelman2015}.
+To generate the new data $\tilde{\rb}$, we first generate a distribution of kick velocities for our merger configuration by randomly sampling SMBH spin values, as we assume this is the only unconstrained variable in our simulated SMBH merger.
+For the dimensionless SMBH spin parameter $\alpha_\bullet$ we use the distribution for dry mergers given in \citet{zlochower2015}, which takes the form of a beta distribution:
+\begin{equation}
+    P_\mathrm{Zlochower}(\alpha_\bullet) \propto (1-\alpha_\bullet)^{4.66884-1} \alpha_\bullet^{10.5868-1}
+\end{equation}
+and assume the direction of each SMBH spin vector is randomly distributed uniformly on the sphere.
+With the distribution of $\alpha_\bullet$, and the state of the SMBH binary just prior to coalescence, we use the relations presented in \citet{zlochower2015} to obtain a distribution of SMBH recoil velocities.
+
+The new data $\tilde{\rb}$ is then generated by transformation sampling the distribution of kick velocities, and `pushing' the values through the desired $\rb$--$\vk$ relation.
+Using transformation sampling allows us to obtain a distribution of core radii predicted by a given kick velocity model, as shown in \autoref{fig:rb_pdf}, whilst simultaneously accounting for uncertainty in the fit described by the $\rb$--$\vk$ relations.
+Using $10^4$ Monte Carlo sampled values for $\vk$, where each value is used across 4 chains each of length 2000 samples, provides a total sample size of $8\times10^7$ values of $\tilde{\rb}$.
+
+By only varying the SMBH spin vectors prior to merger, the range of kick velocities predicted by the model for the particular merger configuration in our study varies from $\vk=0\,\kmps$ to $\vkmax \equiv \vk\sim4000\,\kmps$, with a peak at $\vk \simeq 250\,\kmps$.
+The maximum kick velocity that resulted in a settled SMBH in our sample, $\vk=1020\,\kmps$, corresponds to the $\sim\ordinalthird{63}$ percentile of the empirical cumulative distribution function (CDF) of the kick velocity distribution.
+Additionally, from the empirical CDF we find that approximately 10 per cent of recoil velocities exceed the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}=1800\,\kmps$.
+
+The distribution of recoil velocities becomes slightly more right-skewed if the masses of the SMBHs are allowed to vary prior to merger.
+By randomly sampling each SMBH mass as $M_{\bullet,\mathrm{new}} = 10^{\mathcal{N}(0, 0.3)} M_{\bullet,0}$, where the dispersion of 0.3 is the intrinsic scatter in the $M_\bullet$--$M_\mathrm{bulge}$ relation from \citet{haring2004}, we find that a recoil velocity of $1020\,\kmps$ corresponds to the $\sim\ordinal{76}$ percentile, and the central escape velocity of $v_\mathrm{esc}=1800\,\kmps$ corresponds to the $\sim\ordinal{95}$ percentile.
+Thus, whilst a non-equal mass SMBH binary merger results in a slightly different recoil velocity distribution, large recoil velocities are still allowed, and are frequent, within the expected SMBH mass variation.
+Herein we consider our fiducial case of equal mass SMBH binaries, but do not expect the conclusions to drastically differ in the more general case.
+
+As we cannot be guaranteed that the proposed $\rb$--$\vk$ relations hold beyond our fitted data range of $0\,\kmps$--$1020\,\kmps$, we discuss the distribution of core radii in two parts: the first limiting our analysis to $\vk \leq 1020\kmps$, and the second to $\vk \leq \vkmax$.
 
@@
     \includegraphics[width=0.4\textwidth]{rb_pdf}
-    \caption{Probability density functions of transformation sampled core radius assuming a given model relation. 
-    The SMBH kick velocity is Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins, and pushed through each of the three fitted model in \autoref{eq:vkrb_exp}, \autoref{eq:vkrb_lin}, and \autoref{eq:vkrb_sig}
-    The sigmoid model is the preferred model, and has its mode at a larger value of $\rb$ than either the exponential or linear models, at $\rb/r_{\mathrm{b},0}\sim2.8$ ($\rb/\mathrm{kpc}\sim1.6$).
-    The predicted kick velocities range from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
+    \caption{Probability density functions of transformation sampled core radii assuming a given model relation. 
+    Solid lines depict a truncated recoil velocity distribution, whereas dash-dot lines depict a non-truncated recoil velocity distribution.
+    The SMBH spin vectors are Monte Carlo sampled from the \citet{zlochower2015} relation assuming randomly-aligned azimuthal spins to generate a distribution of recoil velocities, which are then pushed through each of the three fitted models in \autoref{eq:vkrb_exp}, \autoref{eq:vkrb_lin}, and \autoref{eq:vkrb_sig}.
+    The sigmoid model is the preferred model, and has its mode at a larger value of $\rb$ than either the exponential or linear models, at $\rb/\rbbin\sim2.29$ ($\rb/\mathrm{kpc}\sim1.27$) in the truncated case.
+    The (non-truncated) predicted kick velocities range from $\vk=0\,\kmps$ to $\vk\sim4000\,\kmps$.
     }
@@
 
-\subsection{Integral field unit kinematics}\label{ssec:ifu}
-Having established from a forward-modelling point of view that a changing SMBH recoil velocity induces a systematic affect in the mass distribution of the galaxy remnant, we turn our attention to inferring the occurrence of an SMBH that has experienced a recoil velocity, based on the snapshot when the SMBH is seen to be settled.
-
-We first create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the BH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the BH kick axis (`orthogonal').
-Testing these two special projections, we investigate the limiting cases of how the galaxy kinematics behave when all or none of the BH motion is directed along the LOS to the observer.
-
-We create our mock IFU observations following \citet{naab2014}.
-Our observations are centred on the stellar centre of mass using the shrinking sphere method, and encompass a circular aperture of $0.25R_{1/2}$, where $R_{1/2}$ is the three-dimensional half mass radius.
-For each particle within this aperture, we generate 25 pseudo-particles with identical LOS velocities as the original particle, but are spatially displaced from the original particle with a Gaussian distribution in the projected $x$ and $y$ direction with a standard deviation of $0.3\,\mathrm{kpc}$, to mimic seeing effects.
-The pseudo-particles are then binned onto a spatial grid, where the grid is assumed to have a resolution of $0.2''$, similar to the observations of \citet{neureiter2023}, which then corresponds to a physical resolution\footnote{https://cosmocalc.icrar.org} of $\sim0.04\,\mathrm{kpc}$ at redshift $z=0.01$.
-Following the method in \citet{cappellari2003}, we then group adjacent bins into larger Voronoi bins to achieve a particle number of $\sim 50000$ particles per Voronoi bin.
-
-For each Voronoi bin, we follow \citet{vandermarel1993} and decompose the LOS velocity into a series of Gauss-Hermite functions, described by the mean radial velocity $V$, the velocity dispersion $\sigma$, asymmetric deviations $h_3$, and symmetric deviations $h_4$.
-The line profile is thus described by:
-\begin{equation}
-    \mathfrak{L} = \frac{1}{\sqrt{2\pi}\sigma} e^{-w^2/2} \left\{ 1 + \sum_{j=3}^4 h_j H_j(w) \right\},
-\end{equation}
-where $w \equiv (v_\mathrm{LOS} - V)/\sigma$, and the Hermite polynomials $H_3$ and $H_4$ are defined:
-\begin{align}
-    H_3 &= \frac{1}{\sqrt{3}} (2w^3 - 3w) \nonumber \\
-    H_4 &= \frac{1}{\sqrt{24}} (4w^4 - 12w^2 + 3).
-\end{align}
-A recent comprehensive study of LOS velocity distribution fitting applied to elliptical galaxies can be found in \citet{mehrgan2023}.
-Relative to a standard Gaussian (which has $h_4=0$), a positive value of $h_4$ corresponds to extended high velocity tails in the LOS velocity distribution, whereas a negative value of $h_4$ corresponds to weaker tails in the LOS velocity distribution.
-
-To quantify the radial dependence of $h_4$, we calculate the cumulative $h_4$ parameter:
-\begin{equation}\label{eq:h4}
-    \langle h_4 \rangle = \sum_i^N h_4(R_i),
-\end{equation}
-which improves legibility of the radial trends of the $h_4$ parameter, which can oscillate about zero.
-
-The radial dependence of $\langle h_4 \rangle$ is shown in \autoref{fig:h4}, where simulations with $\vk/\kmps < 600$ are terminated with a circle point, simulations with $600 \leq \vk/\kmps < 1020$ are terminated with a square point, and the $\vk=2000\,\kmps$ simulation is terminated with a diamond point.
-Aside from the $\vk=2000\,\kmps$ simulation, increasing the kick velocity results in progressively more negative $\langle h_4 \rangle$ in both the parallel and orthogonal projections. 
-The value of $\langle h_4 \rangle$ at the largest radii saturates for simulations with $\vk\gtrsim 600\,\kmps$, at a value of $\sim -0.5$ for the parallel projection, and $\sim -0.1$ for the orthogonal projection.
-The case is markedly different though for the $\vk=2000\,\kmps$ simulation, which appears more consistent with the $\vk=0\,\kmps$ simulation than the simulations with $\vk\gg0\,\kmps$.
-The value of $\langle h_4 \rangle$ increases at all radii for both projections, exceeding a maximum value of 1.0 for the parallel projection, and 1.5 for the orthogonal projection. 
-\drafting{can we relate this to the Langrangian radii perhaps?}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.5\textwidth]{h4}
-    \caption{
-        Cumulative $h_4$ parameter, as defined in \autoref{eq:h4}, for the parallel (top panel) and orthogonal (bottom panel) projections.
-        Similar to \autoref{fig:density3d}, simulations with $\vk\geq 900\,\kmps$ have a consistent colour.
-        Lines terminating with a circle point indicate simulations with $\vk/\kmps < 600$, those with square points indicate simulations with $600 \leq \vk/\kmps < 1020$, and the diamond point indicates the $\vk=2000\,\kmps$ simulation.
-        For both the parallel and orthogonal projections, for those simulations with $\vk \leq 1020\,\kmps$ (i.e.
-have their SMBHs settled), there is a systematic decrease in the cumulative $h_4$ parameter with increasing $\vk$ for $\vk\lesssim600\,\kmps$.
-        For $600\lesssim\vk/\kmps\leq1020$, the cumulative $h_4$ parameter is indistinguishable from each other. 
-        For the $\vk=2000\,\kmps$ case, the cumulative $h_4$ is both positive and significantly larger than the $600\lesssim\vk/\kmps\leq1020$ simulations, for both parallel and orthogonal projections.
-    }
-    \label{fig:h4}
-\end{figure}
-
-\subsection{Orbit analysis}\label{ssec:orbits}
+\subsubsection{Core size distribution for $\vk \leq 1020\kmps$}\label{sssec:coreT}
+To truncate the kick velocity distribution with an upper bound of $1020\,\kmps$, we take only those values of $\vk$ which satisfy the velocity criterion, noting that as $1020\,\kmps$ corresponds to the $\sim\ordinalthird{63}$ percentile, we still have in excess of $4.5\times10^7$ samples of the core radius distribution.
+
+We show the core radius distribution for the three different models with solid lines in \autoref{fig:rb_pdf}.
+We find that the mode of the forward-folded core radius distribution is similar for the exponential and linear models, with $1.01\,\mathrm{kpc}$ ($1.82\,\rbbin$) and $0.88\,\mathrm{kpc}$ ($1.59\,\rbbin$), respectively.
+Conversely, the mode for the sigmoid relation occurs at a larger value of $\rb$, namely $1.28\,\mathrm{kpc}$ ($2.31\,\rbbin$).
+
+Looking at the shapes of the three distributions in \autoref{fig:rb_pdf}, we see a similarity in shapes between the exponential and linear models, which are both slightly right-skewed. 
+This is due to the ever-increasing value of $\rb$ predicted for an increasing $\vk$: higher recoil velocities produce larger cores, but as higher recoil velocities become increasingly unlikely (assuming the \citet{zlochower2015} distribution), we observe the skew in the $\rb$ probability density function (PDF).
+Conversely, the sigmoid model has a small hump prior to the mode, arising from the high likelihood of small kick velocities being pushed through the sigmoid function (\autoref{eq:vkrb_sig}) prior to the upper plateau of the function.
+The peak in the $\rb$ distribution from the sigmoid model is a result of a large range of kick velocities (albeit with decreasing frequency) being mapped to the upper plateau of the sigmoid function, hence the correspondence between the median of the scale factor $K=3.04$ in \autoref{eq:vkrb_sig} and the distribution mode for the model, $2.31\,\rbbin$.
+
+To determine which model of the three (exponential, linear, or sigmoid) best describes the data, we use approximate leave-one-out cross validation (LOO-CV) with Pareto smoothed importance sampling \citep[PSIS,][]{vehtari2017,vehtari2024}.
+PSIS LOO-CV estimates the expected log pointwise predictive density ($\widehat{\mathrm{ELPD}}$), defined
+\begin{equation}\label{eq:elpd}
+    \widehat{\mathrm{ELPD}} = \sum_{i=1}^N \log \int p(r_{\mathrm{b},i} | \vb{\kappa}) p(\vb{\kappa} | r_{\mathrm{b},-i}) \dd{\vb{\kappa}},
+\end{equation}
+to determine the out-of-sample predictive accuracy of each model.
+The integral in \autoref{eq:elpd} is analogous to \autoref{eq:ppd}, and gives the ability to predict the $i^\mathrm{th}$ data point $r_{\mathrm{b},i}$ when that same data point is excluded from the estimate of $\vb{\kappa}$, namely $p(\vb{\kappa} | r_{\mathrm{b},-i})$ \citep{gelman2014}.
+A larger value of $\widehat{\mathrm{ELPD}}$ indicates better predictive ability of a given model compared to some other model \citep{vehtari2017,riha2024}.
+When performing the LOO-CV, we ensure that the Pareto-$k$ values in the PSIS method are below 0.5, indicating reliable importance sampling estimates \citep{vehtari2024}.
+
+We report the preferred model to be the sigmoid model (\autoref{eq:vkrb_sig}), with an $\widehat{\mathrm{ELPD}}$ of 1.70, compared to 1.58 and 0.61 for the exponential and linear models, respectively.
+However, we note that the difference in $\widehat{\mathrm{ELPD}}$ between the sigmoid and exponential models is not large, and the standard error of the difference in $\widehat{\mathrm{ELPD}}$, denoted $\widehat{\mathrm{SE}}\left( \Delta\widehat{\mathrm{ELPD}} \right)$, is 2.62 when comparing the sigmoid model to the exponential model, and 2.71 when comparing the sigmoid model to the linear model.
+This indicates that the difference between the models is likely not substantial in this range of recoil velocities.
+
+\subsubsection{Core size distribution for $\vk \leq \vkmax$}\label{sssec:coreNT}
+We show the core radius distribution, without a truncation in the recoil velocity distribution, for the three different models with dash-dot lines in \autoref{fig:rb_pdf}.
+For the exponential and linear models, the distributions of $\rb$ are significantly right-skewed and broadened compared to when a truncation is applied to the recoil velocities.
+The right-skew arises due to the derivatives of these models being positive and greater than unity, resulting in increasingly-higher recoil velocities being mapped to increasingly-larger core radii.
+It is important to note that the mode of the $\rb$ distributions for the exponential and linear models with and without recoil velocity truncation are consistent, despite the differing shape of the distributions.
+This is a result of low recoil velocities being more likely to occur, and hence contributing more to the transformation sampling of $\vk$ to $\rb$, than the high recoil velocity samples.
+
+Similarly, the sigmoid model demonstrates a broader distribution of core radii when no truncation is applied to the recoil velocities compared to when a truncation is applied.
+However the overall shape is more consistent with the truncated case than in the instance of the exponential or linear models.
+The mode of the $\rb$ distribution is shifted to $1.49\,\mathrm{kpc}$ ($2.70\,\rbbin$) from $1.28\,\mathrm{kpc}$.
+By construct, the sigmoid function is not as sensitive to large values of $\vk$ (compared to the exponential and linear models) if the function plateau is constrained to recoil velocities within our sampled range, as $\vk \rightarrow \rb$ and $\vk+\delta_{v_\mathrm{kick}} \rightarrow \sim \rb$.
+
+Despite the three models mapping SMBH recoil velocities $0\leq \vk/\kmps \leq 1020$ to $\rb$ with a similar level of accuracy, as revealed by the $\widehat{\mathrm{ELPD}}$ measure, we argue that the sigmoid model is the overall preferred model when considering an unconstrained SMBH recoil velocity distribution.
+The response of the stellar system to SMBH recoil velocities greater than the escape velocity of the system should be consistent, irrespective of how much greater $\vk$ is than $v_\mathrm{esc}$.
+This behaviour is not captured with the exponential nor linear models.
+For example, in the case of the sigmoid model, a normalised recoil velocity of $v'=2 \rightarrow \rb=3.9$, whereas $v'=4 \rightarrow \rb=4.0$.
+Conversely, for the exponential model, $v'=2 \rightarrow \rb=5.7$, whereas $v'=4 \rightarrow \rb=9.0$.
+From this argument, we suggest the sigmoid model to be the preferred model when considering the core radius produced by arbitrary SMBH recoil velocity.
+
+Irrespective of using the full distribution of SMBH recoil velocities predicted by the \citet{zlochower2015} model, or a truncated distribution thereof, a recoiling SMBH serves to enhance the stellar core formed by the SMBH binary prior to merger typically by a factor of 2-3.
+This leads to core sizes in excess of a kiloparsec, which is also seen in observations \citep[e.g.][]{thomas2016,dullo2019}.
+
+\section{Orbit analysis}\label{sec:orbits}
 
@@
     \caption{
-        Orbit analysis of the merger remnants, with a consistent colour scheme to \autoref{fig:density3d}.
-        Stellar particles are assigned to one of seven different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 30.0$.
+        Orbit analysis of the merger remnants, with a consistent colour scheme to \autoref{fig:lagrangian}.
+        Stellar particles are assigned to one of eight different orbital families, and binned into ten logarithmically-spaced radial shells such that $0.2 \leq R/\mathrm{kpc} \leq 11.0$.
         Noticeably, increasing recoil velocity induces a higher fraction of $\pi$-box orbits at all radii, and a complimentary decrease in rosette orbits at radii $r\lesssim1\,\mathrm{kpc}$.
@@
         Same as \autoref{fig:orbits}, but each merger remnant is shown in its own panel, with the different orbital families depicted by colour, to better demonstrate the radial fraction of orbits for a given merger.
+        Note that the orbits that were unable to be classified are not shown in this plot.
     }
@@
 
-To understand the kinematic maps, we perform an orbit analysis of the merger remnants in our sample following \citet{frigo2021}, which is briefly described here.
-
-First, the merger remnant is rotated so that the $z$-axis coincides with the minor axis of the reduced inertia tensor as measured using the top 50\% most bound stellar particles to the stellar centre of mass.
-We cannot centre on the recoiling SMBH for this analysis, as for high kick velocities the SMBH is significantly displaced from the centre.
-Thus, we use the shrinking sphere method on stellar particles to determine the centre of mass of the system.
-The stellar potential of the merger remnant is fit using a self consistent field (SCF) potential \citep{hernquist1992}, using the \citet{hernquist1990} profile as a zeroth-order basis and limiting the expansion to $n_\mathrm{max}=18$ and $l_\mathrm{max}=7$ following \citet{frigo2021}.
-The potential from the SMBH is then added as a point mass potential to the SCF potential.
-Collectively, this represents the potential in an analytical form in which the orbits of individual stellar particles can be integrated.
-The potential of the merger remnant is checked for stability by comparing the potential from the SCF method to the potential computed from the particle data, and ensuring the ratio of the two is $\sim1$ at all radii.
-Each stellar particle within $30\,\mathrm{kpc}$ of the centre is integrated for fifty orbits to determine which, if any, orbital resonances exist. 
-The orbital resonances define the different families of orbits, as given in \citet{frigo2021}.
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{orbit_bt}
+    \caption{
+        Amount of stellar mass classified as box orbits (boxlet or $\pi$-box) compared to tube orbits ($x$-tubes, $z$-tube, or rosette) within the median core radius $\rb$ for each simulation.
+        Points are coloured by kick velocity.
+        Simulations with low $\vk$ predominantly contain tube orbits within $\rb$, whereas larger recoil velocities have a tendency for box orbits to dominate. 
+    }
+    \label{fig:orbitsbt}
+\end{figure}
+
+Having established from a forward-modelling point of view that a changing SMBH recoil velocity induces a systematic effect in the mass distribution of the galaxy remnant, we turn our attention to inferring the occurrence of an SMBH that has experienced a recoil velocity, based on the snapshot when the SMBH is seen to be settled.
+To this end, we perform an orbit analysis of the merger remnants in our sample following \citet{frigo2021}, which is briefly described here.
+
+First, the merger remnant is rotated so that the $z$-axis coincides with the minor axis of the reduced inertia tensor (\autoref{eq:inertia}) as measured using the top 50 per cent most bound stellar particles to the stellar centre of mass.
+We then need to fit the gravitational potential of the system, which we do so by using a self-consistent field (SCF) potential \citep{hernquist1992,jesseit2005,rottgers2014}, 
+representing the potential $\Phi(\vb{x})$ generated by the particles in the simulation as a combination of harmonics with expansion coefficients $A$:
+\begin{equation}
+    \Phi(\vb{x}) = \sum_{nlm} A_{nlm} \Phi_{nlm}(\vb{x}),
+\end{equation}
+where each individual harmonic independently satisfies the Poisson equation.
+The coefficients $A_{nlm}$ describe the order $n$ the radial component of the potential is expanded to, and to which order $l$, $m$ the angular components are expanded; the angular dependence is a (somewhat complicated) function of the spherical harmonics $Y_{lm}(\theta, \phi)$, where $|m|<l$ as usual.
+The zeroth order of the potential expansion $\Phi_{000}(\vb{x})$ is the model upon which we wish to construct our expansion, and a common choice for this is the \citet{hernquist1990} profile.
+Following \citet{frigo2021}, the expansion is limited to $n_\mathrm{max}=18$ and $l_\mathrm{max}=7$, as this sufficiently captures the potential of the galactic merger remnant.
+
+The chosen basis functions are unable to describe a point-mass potential, thus complicating the simulations for which the SMBH has settled at the centre. 
+To overcome this, we centre the snapshot on the SMBH position, but exclude the SMBH from the SCF routine.
+This provides us with a potential field without the SMBH contribution; the potential from the SMBH is added as a point-mass potential following the SCF potential construction.
+In the case where the SMBH has not settled at the centre (i.e.
+the $\vk=2000\,\kmps$ simulation) we simply align the snapshot with the centre of mass determined using the shrinking sphere method on stellar particles, and neglect the contribution from the SMBH (as it is very far from the centre and thus no longer influences the orbits of the stellar particles).
+We now have a representation of the potential in an analytical form in which the orbits of individual stellar particles can be integrated.
+The potential of the merger remnant is checked for stability by comparing the potential from the SCF method to the potential computed from the particle data by \gadget{} during the simulation, and ensuring the ratio of the two is $\sim1$ at all radii.
+
+Each stellar particle within $2\Reff \simeq 11\,\mathrm{kpc}$ of the centre is integrated for fifty orbits to determine which, if any, orbital resonances exist\footnote{As each stellar particle is integrated in a fixed potential, the effect of two-body interactions are not captured by this method.}.
+The orbital resonances define the different families of orbits, as given in \citet{frigo2021}, and are listed in \autoref{tab:orbits}.
+We then determine the fraction of each orbital family in ten logarithmically-spaced radial bins extending from $0.2\,\mathrm{kpc}$ to $11\,\mathrm{kpc}$.
 
 \begin{table}
-    \caption{Orbital families}
+    \caption{Classification of orbital families}
     \label{tab:orbits}
@@
         \hline
-        $x$-tube & Rotate about the major axis of the galaxy \\
+        inner $x$-tube & Rotate about the major axis of the galaxy, but move radially when far from centre (concave shaped) \\
+        outer $x$-tube & Rotate about the major axis of the galaxy (convex shaped) \\
         $z$-tube & Rotate about the minor axis of the galaxy \\
         $\pi$-box & Non-resonant motion with no net angular momentum (radial motion) \\
-        boxlet & Resonant motion with angular momentum \\
+        boxlet & Resonant motion with no net angular momentum \\
         rosette & Typical orbit in a point-mass dominated spherically-symmetric potential \\
@@
 The first is the increase in the fraction of $\pi$-box orbits with increasing recoil velocity, most notable at radii $r\lesssim5\,\mathrm{kpc}$.
-The simulation with $\vk=2000\,\kmps$ noticeably stands out, with a fraction of $\pi$-box orbits being $\sim0.6$ at the inner most radii. 
-Complimentary but converse to the first trend, increasing the SMBH recoil velocity decreases the fraction of $x$-tube orbits at inner radii from $\sim 0.4$ to $\sim 0.2$. 
-Again the $\vk=2000\,\kmps$ is a noticeable outlier to the other simulations, however remains consistent with the general trend.
+The simulation with $\vk=2000\,\kmps$ noticeably stands out, with the fraction of $\pi$-box orbits being $\sim0.6$ at the innermost radii. 
+Complimentary but converse to the first trend, increasing the SMBH recoil velocity marginally decreases the fraction of $z$-tube orbits from $\sim 0.2$ to $\lesssim 0.1$ within $1\,\mathrm{kpc}$.
+Again the $\vk=2000\,\kmps$ is noticeably offset from the other simulations, however remains consistent with the general trend.
 Finally, increasing the SMBH recoil velocity decreases the fraction of rosette orbits at small radii $r<1\,\mathrm{kpc}$, with no rosette orbits for the $\vk=2000\,\kmps$ simulation.
-There is evidence for a minor decrease in $z$-tube orbits with increasing recoil velocity, however this trend is not as dramatic as the three previously discussed.
-
-We can naturally explain the decrease in rosette orbits with increasing recoil velocity by considering a rosette orbit requires a point-mass like potential to orbit in.
-With increasing recoil velocity, we are displacing the SMBH increasingly further from the stellar centre of mass, thus disrupting the conditions required to maintain such an orbit.
-A consequence of disrupting these regular, rosette orbits, is the inducement of more radial orbits with increasing recoil velocity, manifested as an increase in non-resonant $\pi$-box orbits.
+There is evidence for a minor decrease in outer $x$-tube orbits with increasing recoil velocity at radii, however this trend is not as dramatic as the three previously discussed.
+In particular, the $\vk=2000\,\kmps$ simulation has almost no outer $x$-tube orbits until $r=2\,\mathrm{kpc}$, and only comes to dominate the orbit fraction at $r\sim10\,\mathrm{kpc}$.
+
+We can naturally explain the decrease in rosette orbits with increasing recoil velocity by considering that a rosette orbit requires a point-mass like potential to orbit in.
+With a larger recoil velocity, we are displacing the SMBH increasingly further from the stellar centre of mass, thus disrupting the conditions required to maintain such an orbit.
+A consequence of disrupting these regular, rosette orbits, is the inducement of more radial orbits as recoil velocity grows, manifested as an increase in non-resonant $\pi$-box orbits. 
+With a settled SMBH dominating the potential at the centre of the merger remnant, stellar orbits may slowly diffuse through relaxation processes to form rosette orbits at times $t>t_\mathrm{settle}$.
+However, the timescale required for boxlet and $\pi$-box orbits to acquire sufficient angular momentum to become rosette orbits is very long; as discussed in \autoref{ssec:relax}, relaxation processes do not dominate the centre of the galaxy over the timescales relevant for these simulations, and violent relaxation is not applicable in the instance of a single SMBH.
 
 We show the orbit families as a function of radius organised by recoil velocity in \autoref{fig:orbits2}, instead of family class.
-At large radii ($r>10\,\mathrm{kpc}$), $x$-tube orbits dominate over all other families for all simulations, \drafting{a known result from major mergers, Naab??}.
-Interestingly, increasing the recoil velocity pushes the intersection of boxlet and $z$-tube orbits to increasingly larger radii.
-Finally, for the $\vk=2000\,\kmps$ simulation, the dominance of $\pi$-box orbits over all other families is apparent for radii $r\lesssim5\,\mathrm{kpc}$.
+At intermediate radii ($r>2\,\mathrm{kpc}$), outer $x$-tube orbits dominate over all other families for all simulations, a distinct feature of prolate systems.
+The merger remnants in this study are predominantly prolate rotators at the time of analysis (see bottom panel of \autoref{fig:triax}), in agreement with results from previous simulation studies of major mergers \citep[e.g.][]{naab2003, gonzalez2009,ebrova2017}.
+Interestingly, increasing the recoil velocity pushes the intersection of boxlet and $z$-tube orbits to increasingly larger radii, from $\sim 1.5\,\mathrm{kpc}$ for $\vk=0\,\kmps$ to $\sim 3\,\mathrm{kpc}$ for $\vk=1020\,\kmps$.
+Finally, for the $\vk=2000\,\kmps$ simulation, the dominance of $\pi$-box orbits over all other families is apparent for radii $r\lesssim8\,\mathrm{kpc}$.
 This is the only recoil velocity for which $\pi$-box orbits are the dominant orbital family; in the simulations with $\vk\leq1020\,\kmps$, $\pi$-box orbits are always subdominant to boxlet orbits.
+
+Further insight is provided by grouping the orbital families into two broad classes, `boxes' and `tubes'.
+Those orbits which do not have a net angular momentum we collectively refer to as boxes, and include boxlet and $\pi$-box orbits.
+Similarly, those orbits which do have some net angular momentum we collectively refer to as tubes, and include the two types of $x$-tubes, $z$-tubes, and rosette orbits.
+The amount of stellar mass within the median core radius $\rb$ for these two broad classes is shown in \autoref{fig:orbitsbt}, where the points are coloured by the SMBH recoil velocity.
+
+Immediately we see that for low kick velocities, $\vk\leq 180\,\kmps$, tube orbits dominate the merger remnant core.
+In particular, the $\vk=0\,\kmps$ case has the greatest ratio of tube to box orbits within the core, $M_{\star,\mathrm{tube}}/M_{\star,\mathrm{box}} = 2.0$. 
+This is in agreement with an SMBH binary preferentially ejecting box orbits which lie within the loss cone during the three-body scattering phase.
+
+Increasing the SMBH recoil velocity, there is a strong tendency for box orbits to dominate the core region, in some cases contributing more than $6.0\times10^8\,\Msun$ of stellar mass ($\sim0.1 M_\bullet$), compared to some $3.5\times10^8\,\Msun$ provided by tube orbits.
+The recoil velocity where the transition from a tube orbit-dominated core to a box orbit-dominated core occurs is around $\vk\sim240\,\kmps$, approximately the same recoil velocity that allows the kicked SMBH to exit the SMBH binary-scoured core (see \autoref{fig:infall}, upper panel).
+This suggests that as the kicked SMBH ploughs through the stellar core whilst it is settling but before its velocity falls below the velocity dispersion of the core, the rapid change in potential disrupts the conditions for angular momentum conserving orbits, altering the stellar particles to radial orbits instead.
+For the highest kick velocities, $\vk > 900\,\kmps$, the amount of stellar mass in box orbits is almost twice that in tube orbits, as shown in \autoref{fig:orbitsbt}.
+
+In summary, a recoiling SMBH exiting the core leaves an imprint in the stellar orbits as it settles.
+Consequently, the stellar kinematics are also affected by the SMBH motion.
+
+\section{Stellar kinematics}\label{sec:kinematics}
+\subsection{Integral field unit kinematics}\label{ssec:ifu}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.99\textwidth]{IFU_bt_0000}
+    \caption{
+        Mock integral field unit kinematic maps of the $\vk=0\,\kmps$ remnant stellar particles in the parallel projection at the time the SMBH has settled.
+        The IFU maps extend out to $0.25\,r_{1/2}$ in a square aperture.
+        The top row shows the IFU maps for each of the ordered rotation $V$, velocity dispersion $\sigma$, asymmetric deviation coefficient $h_3$, and symmetric deviation coefficient $h_4$, for all stellar particles within the square aperture.
+        The second row shows the IFU maps for those stellar particles within the aperture classified as box orbits, and the third row shows the complementary IFU maps for those stellar particles within the aperture classified as tube orbits.
+        The dashed circle indicates the core radius $\rb$ of the merger remnant.
+        For the box and tube orbit rows, the colour scale is consistent for each column, however the colour scale is different for the top row.
+        The ordered rotation $V$ is generally correlated with $h_3$.
+        A centrally-peaked velocity dispersion is only visible in box orbits and not tube orbits, as is an annulus of negative $h_4$.
+        Almost no negative $h_4$ is present in the IFU map for all stellar particles.
+        From \autoref{fig:orbitsbt}, this simulation has $1.4\times10^8\,\Msun$ stellar mass in box orbits, and $3.0\times10^8\,\Msun$ stellar mass in tube orbits, within $\rb$.
+    }
+    \label{fig:IFU0000}
+\end{figure*}
+
+\begin{figure*}
+    \centering
+    \includegraphics[width=0.99\textwidth]{IFU_bt_0600}
+    \caption{
+        Similar to \autoref{fig:IFU0000}, mock integral field unit kinematic maps of the $\vk=600\,\kmps$ remnant stellar particles in the parallel projection at the time the SMBH has settled.
+        From \autoref{fig:orbitsbt}, this simulation has $5.9\times10^8\,\Msun$ stellar mass in box orbits, and $4.1\times10^8\,\Msun$ stellar mass in tube orbits, within $\rb$.
+        Similar to the $\vk=0\,\kmps$ case, a peaked velocity dispersion is visible for all stellar particles, but this contribution comes solely from the box orbits.
+        An annulus of negative $h_4$ is seen to extend to the core radius in the IFU map for all stellar particles, and for box orbits.
+    }
+    \label{fig:IFU0600}
+\end{figure*}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{h4}
+    \caption{
+        Annuli-averaged $h_4$ parameter for the parallel (top panel) and orthogonal (bottom panel) projections.
+        For both the parallel and orthogonal projections, for those simulations with $\vk \leq 1020\,\kmps$ (i.e.
+have their SMBHs settled), there is a systematic decrease in the $h_4$ parameter with increasing $\vk$ for $\vk\lesssim 720 \,\kmps$.
+        For $720 \lesssim\vk/\kmps\leq 1020$, the $h_4$ parameter is indistinguishable from each other. 
+        For the $\vk=2000\,\kmps$ case, the $h_4$ is both positive and significantly larger than the $720\lesssim\vk/\kmps\leq1020$ simulations, for both parallel and orthogonal projections. 
+    }
+    \label{fig:h4}
+\end{figure}
+
+We next create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the SMBH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the SMBH kick axis (`orthogonal').
+Testing these two special projections, we investigate the limiting cases of how the galaxy kinematics behave when all or none of the SMBH motion is directed along the LOS to the observer.
+
+We create our mock IFU observations following \citet{naab2014}, assuming our merger remnants represent local elliptical galaxies that lie at $z=0.01$.
+Our observations encompass a square aperture of side length $0.25r_{1/2}$, where $r_{1/2}$ is the half mass radius. 
+For each particle within this aperture, we generate 25 pseudo-particles with identical LOS velocities as the original particle, but are spatially displaced from the original particle with a Gaussian distribution in the projected $x$ and $y$ direction with a standard deviation of $6.8''$ (corresponding to a physical scale\footnote{https://cosmocalc.icrar.org} of $0.3\,\mathrm{kpc}$) to mimic seeing effects.
+The pseudo-particles are then binned onto a spatial grid, where the grid is assumed to have a resolution of $0.2''$, similar to the observations of \citet{neureiter2023}, which then corresponds to a physical resolution of $\sim0.04\,\mathrm{kpc}$.
+Following the method in \citet{cappellari2003}, we then group adjacent bins into larger Voronoi bins to achieve a particle number of $\sim 50000$ particles per Voronoi bin.
+
+For each Voronoi bin, we follow \citet{vandermarel1993} and decompose the LOS velocity into a series of Gauss-Hermite functions, described by the mean radial velocity $V$, the velocity dispersion $\sigma$, asymmetric deviation $h_3$, and symmetric deviation $h_4$.
+The line profile is thus described by:
+\begin{equation}
+    \mathscr{L} = \frac{1}{\sqrt{2\pi}\sigma} e^{-w^2/2} \left\{ 1 + \sum_{j=3}^4 h_j H_j(w) \right\},
+\end{equation}
+where $w \equiv (v_\mathrm{LOS} - V)/\sigma$, and the Hermite polynomials $H_3$ and $H_4$ are defined:
+\begin{align}
+    H_3 &= \frac{1}{\sqrt{3}} (2w^3 - 3w) \nonumber \\
+    H_4 &= \frac{1}{\sqrt{24}} (4w^4 - 12w^2 + 3).
+\end{align}
+Relative to a standard Gaussian (which has $h_4=0$), a positive value of $h_4$ corresponds to extended high velocity tails in the LOS velocity distribution and a more peaked distribution about the mean value $V$.
+Conversely, a negative value of $h_4$ corresponds to weaker tails in the LOS velocity distribution and a broader distribution of LOS velocities about the mean value $V$.
+A recent comprehensive study of LOS velocity distribution fitting applied to elliptical galaxies can be found in \citet{mehrgan2023}.
+
+We show example IFU maps for the parallel projections of the $\vk=0\,\kmps$ case in \autoref{fig:IFU0000}, and of the $\vk=600\,\kmps$ case in \autoref{fig:IFU0600}.
+In both figures, the first row corresponds to mock observations of all stellar particles within the square aperture, the second row corresponds to the subset of particles classified as box orbits, and the third row corresponds to the subset of particles classified as tube orbits.
+The first column of each figure depicts the ordered rotation $V$ of the stellar particles, the second column shows the velocity dispersion $\sigma$, the third column shows the coefficient of asymmetric deviation $h_3$, and the fourth column shows the coefficient of symmetric deviation $h_4$.
+The core radius of the merger remnant, as determined in \autoref{ssec:projdens}, is depicted with a dashed ring in each panel.
+
+We first note that all our simulated merger remnants have almost negligible rotation, as immediately seen in the ordered rotation maps of \autoref{fig:IFU0000} and \autoref{fig:IFU0600}.
+Quantitatively, the dimensionless spin parameter is a useful measure to assess the level of rotation \citep{emsellem2007}, and is defined:
+\begin{equation}
+    \lambda_R = \frac{\sum_{i=1}^N F_i R_i |V_i|}{\sum_{i=1}^N F_i R_i \sqrt{V_i^2 + \sigma_i^2}},
+\end{equation}
+where $F_i$ is the flux, $R_i$ is the radial displacement from the centre, $V_i$ is the mean velocity, and $\sigma_i$ is the velocity dispersion, each taken within the $\ordinal{i}$ Voronoi bin.
+We find that our galaxies have $\lambda_R<0.02$ within the IFU aperture, and thus we conclude very little rotation is present.
+
+For the $\vk=0\,\kmps$ case in \autoref{fig:IFU0000}, we first note an interesting structure in the ordered rotation $V$ of all stellar particles. 
+There is a hint of rotation occurring about the $z=0$ axis, however a counter-rotating structure is present close to the $y=0$ axis (keeping in mind that as previously discussed, overall the magnitude of this rotation is very minimal).
+This feature is also seen in the mock IFU images of the tube orbits.
+Additionally, creating mock IFU maps for two other projections (one with the line of sight aligned with the major axis of the inertia tensor, and the other with the line of sight aligned with the vector normal to the merger plane), this feature is not present.
+We thus conclude that immediately following the merger of an SMBH binary, complex structure in the LOSVD may be apparent for tube orbits.
+
+The kinematic features for the second, third, and fourth moments of the LOSVD are not as sensitive to the viewing angle, with box orbits displaying a centrally-peaked velocity dispersion compared to a spatially homogeneous velocity dispersion for tube orbits.
+The spatial extent of the centrally peaked velocity dispersion extends beyond the core radius when viewing all stellar particles, however is overall reduced in magnitude compared to the box orbits owing to the inclusion of the tube orbits.
+A peaked central velocity dispersion can be naturally explained for box orbits owing to the radial motion of these particles increasing in velocity closer to the centre of the potential, when these particles are at their pericentre.
+
+Another curious feature is an annulus of negative $h_4$ that is clearly visible for box orbits, but is absent in the $h_4$ maps for tube orbits and all stellar particles.
+At the centre of the annulus, $h_4\sim0$, with the inner ring of the annulus coinciding with the core radius.
+The region of negative $h_4$ in the IFU maps of the box orbits can be understood as the population of box orbits which did not interact with the SMBH binary prior to coalescence. 
+The positive $h_4$ beyond the annulus is a result of the box orbits reaching apocentre, where the velocity becomes zero (irrespective of orientation), thus resulting in an overly peaked LOSVD.
+From \autoref{fig:orbitsbt}, for $\vk=0\,\kmps$ there is more stellar mass in tube orbits than box orbits within the core (and at radii $r>\rb$, shown by the increased number of Voronoi cells for tube orbits in \autoref{fig:IFU0000} compared to box orbits).
+More tube orbits leads to the $h_4$ annulus being absent in the map of $h_4$ for all stellar particles.
+
+Turning our attention to the IFU maps in \autoref{fig:IFU0600} for the $\vk=600\,\kmps$ case, we notice immediately that ordered rotation about the $z=0$ axis is minimal, and the velocity dispersion for all stellar particles has a centrally-peaked structure that extends to the core radius.
+Converse to the $\vk=0\,\kmps$ case the core radius extends to the same radius as the enhanced velocity dispersion in the box orbits.
+
+Similar to the $\vk=0\,\kmps$ case, an annulus of negative $h_4$ is present for the $\vk=600\,\kmps$ simulation.
+However, the core radius has expanded beyond the central $h_4\sim0$ region to encompass some of the annulus.
+Owing to the box orbit-dominated core (from \autoref{fig:orbitsbt}, $M_{\star,\mathrm{tube}} / M_{\star,\mathrm{box}} \sim 0.75$), the annulus of negative $h_4$ is now present in the $h_4$ kinematic map for all stellar particles.
+Critically however, a central region of $h_4\sim0$ is present in the map for all stellar particles, and corresponds to the same spatial extent of the $h_4\sim0$ region in the $\vk=0\,\kmps$ simulation.
+We propose that this central region of $h_4 \sim 0$ is indicative of the core radius due solely to SMBH binary scouring $\rbbin$, prior to any core enhancement by the recoiling SMBH, and set about identifying the radial extent of this region.
+
+\subsection{Radial trends in $h_4$}
+To quantify the radial dependence of $h_4$, we bin the Voronoi cells into ten concentric annuli out to $5 \rbbin$, and take the median $h_4$ value within each annulus, denoting this as $\langle h_4 \rangle$.
+
+The radial dependence of $\langle h_4 \rangle$ is shown in \autoref{fig:h4}.
+Aside from the $\vk=2000\,\kmps$ simulation, increasing the kick velocity results in progressively more negative $\langle h_4 \rangle$ in both the parallel and orthogonal projections at radii $R\gtrsim1\,\mathrm{kpc}$, seen as an overall downward shift in \autoref{fig:h4}.
+The value of $\langle h_4 \rangle$ saturates for simulations with $\vk\gtrsim 720\,\kmps$, with a minimum of $\langle h_4 \rangle < 0$ at $R \sim 1\,\mathrm{kpc}$ for both the parallel and orthogonal projections.
+
+As discussed in \autoref{ssec:ifu}, identifying the radius at which $\langle h_4 \rangle$ starts to become negative can give insight to the core size due solely to SMBH binary scouring. 
+For the parallel projection, $\langle h_4 \rangle$ starts to become negative at $R \sim 0.5\,\mathrm{kpc}$, and $R \sim 0.6\,\mathrm{kpc}$ for the orthogonal projection, irrespective of the recoil velocity $\vk$.
+Both of these values are in good agreement with the known core size due to binary scouring of $0.58\,\mathrm{kpc}$ discussed in \autoref{ssec:settle}, as indicated by the vertical dotted line in \autoref{fig:h4}.
+
+To understand the overall trends in \autoref{fig:h4}, we must refer to the orbit analysis of \autoref{sec:orbits}.
+Within $R \sim 1\,\mathrm{kpc}$, the $\langle h_4 \rangle$ profiles are largely consistent between kick velocities.
+The dominant orbital family within this radius for simulations with recoil velocities $60 \leq \vk/\kmps \leq 1020$ is the boxlet family: radial orbits which can reach arbitrarily close to the centre.
+As these orbits approach their pericentre at small radii, the different projections of these boxlet orbits result in a LOSVD that is relatively broad around the mean, resulting in a value of $\langle h_4 \rangle$ that is either 0 or slightly negative, irrespective of the SMBH recoil velocity.
+This also explains the consistency in $\langle h_4 \rangle$ between the parallel and orthogonal projections at small radii.
+
+At radii beyond $R\sim1\,\mathrm{kpc}$, there is a strong tendency, at a given radius, for simulations with a low $\vk$ to have a higher value of $\langle h_4 \rangle$ than simulations with a high $\vk$.
+From \autoref{fig:orbits2}, this is also the radius at which the dominant orbital family transitions from boxlet orbits to $x$-tube orbits for simulations with $60 \leq \vk/\kmps \leq 1020$, though boxlet orbits are by no means negligible.
+Additionally, when considering the apocentre distribution of stellar particles with a mean position within $\sim0.25r_{1/2}$ (\autoref{fig:apo_dist}), the distributions are all left-skewed, and become increasingly left-skewed with increasing $\vk$.
+At apocentre, a stellar particle on a radial orbit will have a zero velocity vector, which is then viewed as a zero velocity vector when projected to an observer.
+Similarly, for stellar particles on non-radial orbits, the apocentre represents a minimum, but not necessarily zero, velocity\footnote{Except in the case where the motion is completely tangential to the observer, in which case there is no projected component of the velocity vector.}.
+In either case, the LOSVD peaks towards the mean velocity at apocentre, thus inducing a mildly-positive $\langle h_4 \rangle$.
+As the peak of the apocentre distribution shifts to larger radii with increasing $\vk$, we expect to find more stellar particles at apocentre at small radii for low $\vk$ simulations, than for high $\vk$ simulations.
+This in turn leads to an elevated value of $\langle h_4 \rangle$ at small radii for simulations with low $\vk$ compared to those with high $\vk$.
+
+For the $\vk=2000\,\kmps$ simulation, where the SMBH did not return to the centre of the merger remnant, the radial $\langle h_4 \rangle$ profile is clearly different to those simulations in which the SMBH managed to return to the centre.
+Here, the $\langle h_4 \rangle$ profile is elevated compared to the lower $\vk$ simulations at all radii in both the parallel and orthogonal projections.
+Again, we can readily understand this phenomenon with the orbital analysis in \autoref{sec:orbits}.
+In \autoref{fig:orbits2}, we clearly see that $\pi$-box and boxlet orbits dominate over all other orbital families in the radial range covered by the mock IFU maps.
+Additionally, the apocentre distribution of the $\vk=2000\,\kmps$ simulation peaks at smaller radii than the majority of the simulations with a settled SMBH: it is less left-skewed, indicating that more particles have their apocentre at radii covered by the mock IFU maps.
+By the same reasoning as above, a higher concentration of apocentres at small radii narrows the LOSVD profile, resulting in an elevated value of $\langle h_4 \rangle$, and is seen in \autoref{fig:h4}.
+
+In the analysis of the kinematic maps, a complication arises from choosing the snapshot once the SMBH has settled, which as discussed in \autoref{ssec:settle} varies as a function of recoil velocity.
+To explore the effect of our choice of snapshot, we redo the IFU analysis for various times following when the SMBH kick occurs.
+A representative example is shown in \autoref{fig:h4900} for the case of $\vk=900\,\kmps$.
+We take snapshots corresponding to times between $0.18\,\mathrm{Gyr}$ and $1.70\,\mathrm{Gyr}$, in addition to the fiducial snapshot at $1.28\,\mathrm{Gyr}$.
+Prior to the SMBH settling, the value of $\langle h_4 \rangle$ fluctuates as the settling SMBH influences the surrounding stellar kinematics.
+The $\langle h_4 \rangle$ profile is consistent for all times following the SMBH settling, and thus gives us confidence that, provided the SMBH has settled, the $\langle h_4 \rangle$ profile is insensitive to the precise time of measurement.
+As a result, the differences in $\langle h_4 \rangle$ profile presented in \autoref{fig:h4}
+ are a result of the stellar mass distribution responding to the change in potential caused by the oscillating SMBH, and not due to general relaxation processes (refer also to \autoref{ssec:relax}).
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{figures/beta_profiles_vkick.pdf}
+    \caption{
+        Stellar velocity anisotropy parameter $\beta$ as a function of the kick velocity $\vk$. 
+        The stellar particles are divided into two radial bins with $r \leq \rb$ and $r>\rb$, where the core radius $\rb$ is sampled $N=10^4$ times for each simulation from its marginal $\rb$ posterior distribution.
+        The final value for $\beta$ and the corresponding error are computed as the median and interquartile range across the entire sample (note for the orange points, the error is smaller than the marker).
+        The blue and orange points correspond to the inner and outer bins, respectively, while the dotted horizontal line indicates an ergodic system with $\beta=0$. 
+        Only simulations with $\vk \leq 180 \, \kmps$ display tangential bias ($\beta < 0$) for $r \leq \rb$, whereas all simulations are radially-biased for $r> \rb$.
+    }
+    \label{fig:beta}
+\end{figure}
+
+\subsection{Velocity anisotropy}
+As described in \autoref{sec:orbits}, SMBHs can induce a significant change in the radial distribution of the stellar orbital families.
+On a more global scale, the change in orbit distribution manifests as a stellar velocity anisotropy.
+A widely used measure for the velocity anisotropy of a given stellar system is the anisotropy parameter $\beta$ \citep{binney2008} given by
+\begin{equation}
+    \beta = 1 - \frac{\sigma^2_{\theta} + \sigma^2_{\phi}}{2 \sigma^2_r},
+\end{equation}
+where $\sigma_{\theta}$, $\sigma_{\phi}$, and $\sigma_{r}$ are the components of the stellar velocity dispersion in spherical coordinates. 
+Stellar systems with $\beta < 0$ are referred to as `tangentially-biased' systems, whereas $\beta > 0$  corresponds to `radially-biased' systems. 
+A stellar system with purely radial orbits ($\sigma_{\theta}=\sigma_{\phi}=0$) corresponds to $\beta = 1$, while purely circular orbits ($\sigma_{r}=0$) would yield $\beta = -\infty$.
+An ergodic system is one where $\sigma_{\theta}=\sigma_{\phi}=\sigma_r \iff \beta=0$, which thus describes a system where there is a balance between box and tube orbits.
+
+Prior to their eventual merger, SMBH binaries can interact strongly with the stellar particles in their vicinity via three-body scattering. 
+However, in order to undergo a strong interaction, the stellar particle must come sufficiently close to the SMBH binary. 
+Such stellar particles are predominantly on radial orbits with small pericentre distances and low angular momenta, which enables them to reach the centre of the galaxy: they belong to the loss cone of the SMBH binary.
+As a consequence, SMBH binaries are expected to induce tangentially-biased cores in the central regions of their host galaxies by primarily ejecting stellar particles on radial orbits \citep[e.g.][]{quinlan1997,milosavljevic2001,Thomas2014,rantala2018,rantala2019}.
+Recoiling SMBHs, on the other hand, can induce velocity anisotropy by changing the shape of the local gravitational potential as described in \autoref{sec:orbits}.
+Rosette orbits, for instance, are easily disturbed by a recoiling SMBH: as the potential in the initial position of the SMBH becomes non-spherical, stars initially on rosette orbits can move onto $\pi$-box orbits, which manifests as additional tangential bias in the central regions of the galaxy.
+
+\autoref{fig:beta} shows the anisotropy parameter $\beta$ as a function of $\vk$ for the merger simulations once the SMBH has settled, for $\vk$ between $0\,\kmps$ and $1020\,\kmps$.
+For each simulation, the stellar particles are divided into two radial bins corresponding to $r \leq \rb$ (blue points) and $r>\rb$ (orange points), where the core radius $\rb$ is Monte Carlo sampled $N=10^4$ times for each simulation from its marginal $\rb$ posterior distribution (\autoref{ssec:projdens}).
+The final $\beta$ and the corresponding error are obtained as the median and interquartile range across the entire sample for each merger remnant.
+
+As is evident from \autoref{fig:beta}, the merger remnants with $\vk \leq 180 \, \kmps$ have tangentially-biased ($\beta<0$) stellar populations within $\rb$.
+This is consistent with \autoref{fig:orbitsbt}, in which increasing the kick velocity is shown to induce a lower fraction of angular momentum-conserving orbits ($x$-tube, $z$-tube, and rosette) and a higher fraction of radial orbits ($\pi$-box and boxlet) in the inner regions.
+For larger kick velocities ($\vk\gtrsim\sigma_{\star,0}$), the $\beta$ parameter for stellar particles within $\rb$ is consistent with an ergodic system. 
+This results from the growth of the core radius with increasing $\vk$ shown in \autoref{fig:vkrb}: as $\rb$ increases, the enlarged core region encompasses stellar particles on box orbits which were not ejected by the SMBH binary prior to coalescence, as demonstrated by the IFU maps in \autoref{fig:IFU0600}.
+These stellar particles, found at radii $\rbbin < r < \rb$ (recalling that $\rbbin$ is the core radius due to SMBH binary scouring only), have apocentres that in general increase with increasing kick velocity.
+As the said stellar particles pass radially within the core region, they increase the range of radial velocities within $\rb$, leading to a larger radial velocity dispersion so that $2 \sigma_r^2 \simeq \sigma_\theta^2 + \sigma_\phi^2$ and $\beta \approx 0$ for the ergodic systems seen in \autoref{fig:beta}.
+In this way, the anisotropy parameter $\beta$ of the core region of an elliptical galaxy might give insight into whether or not the recoiling SMBH was ejected with a $\vk$ that removed it from the SMBH binary-scoured core.
+
+For $r>\rb$, all merger remnants display a strong radial bias of $\beta \gtrsim 0.3$.
+This might initially seem contradictory with \autoref{fig:orbits}, which shows that the tangentially-biased outer x-tube and z-tube orbits are the dominant orbital families at $r\sim10$ kpc.
+However, the stellar mass density decreases rapidly towards $r\sim10$ kpc as shown in \autoref{fig:density3d} and hence the outer regions contribute only weakly to the overall $\beta$ for $r>\rb$. 
+Furthermore, the tangential bias of the aforementioned orbital families is weak at large radii, as shown in Figure 9 of \citet{frigo2021}, whereas the non-negligible $\pi$-box and boxlet orbits display strong radial bias at large radii.
+Finally, it is worth noting that the radial bias towards large radii is not induced by the recoiling SMBH: as is evident, the outer $\beta$ bins of all merger remnants with $\vk>0\,\kmps$ are consistent with the $\vk=0 \, \kmps$ remnant, which likewise displays a strong radial bias for $r > \rb$.
+Instead, the radial bias is primarily induced by the (radial) merger orbit, which enables stellar particles to be deposited on radially-biased orbits during the merger.
+A radially-biased outer region is also predicted by cosmological galaxy formation models, in which the late-stage growth of elliptical galaxies is attributed to numerous minor mergers \citep[e.g.][]{Naab2009}.
+As discussed in \citet{rantala2018}, further mergers would likely induce even stronger radial bias towards the outer regions and, assuming the inclusion of new SMBHs, also stronger tangential bias in the central regions.
 
@@
 \section{Discussion} \label{sec:discussion}
-\drafting{
-    Ok, we're here. Let's recap:
-    We see a bifurcation in simulations, where this seems to happen at $\vk=600\,\kmps$.
-    This is manifested in:
-    \begin{itemize}
-        \item The 3D density
-        \item the transition from increasing to decreasing/constant core radius
-        \item the orthogonal projection of the IFU h4 quantity
-        \item The presence of rosette orbits. 
-    \end{itemize}
-    What is going on? 
-    I believe the driver of all this is the orbits. The other items in the above list are ways we can \textit{observe} these differences in orbits. 
-    I suspect, and need to think about how we determine this, that below $600\,\kmps$, there is insufficient energy released from the recoiling SMBH to significantly alter the orbital structure, i.e.
-little core heating.
-    Conversely, above $600\,\kmps$, the SMBH is so rapidly removed from the centre (and in this analysis doesn't return yet) that the system is unable to respond to the rapid change in potential quick enough.
-    $600\,\kmps$ therefore represents a sweetspot, where we have a lot of energy that is then able to be given to the surrounding stars in a reasonable time.
-
-    Some incongruencies:
-    \begin{itemize}
-        \item In 3D density, there is a smooth decrease with increasing kick velocity, until it saturates. How do we then explain a peak in the core radius in the projected density maps?
-    \end{itemize}
-
-    For simulation with vk 480 and less, the BH and the CoM coincide (on approx. scales). 
-    Keplerian orbits only present in those sims where the BH and CoM are close (also confirmed with early analysis of snapshot 8 instead of 16) and also where the BH velocity is less than 300km per s.
-
-    We can use results from this work to identify systems where the SMBH is settling, difficult (impossible?) to determine exact history of system once BH settled. 
-}
-
-Using 32 simulations of gas-free post-merger elliptical galaxy remnants, we have investigated the effect of a recoiling SMBH on the stellar environment. 
-Most importantly, we report a bifurcation in the simulations at $\vk \simeq 0.3 v_\mathrm{esc}$.
-This bifurcation manifests in the three-dimensional stellar density profiles and the orthogonal projection of the kinematic $h_4$ moment.
-Additionally, the peak projected core radius $\rb$ occurs approximately at the bifurcation, where for $\vk \lesssim 0.3 v_\mathrm{esc}$ the core radius $\rb$ increases with $\vk$, and then decreases after the peak to reach a plateau.
-
-These observations, whilst helpful in understanding the presence of a general trend, are not the cause of the bifurcation.
-We argue that the bifurcation is a result of the changing orbital structure in the inner regions of the merger remnant due to the displaced SMBH. 
-As discussed in \autoref{ssec:orbits}, all simulations in which the velocity of the recoiling SMBH exceeds $\vk=600\,\kmps$ do not contain any rosette orbits. 
-Rosette orbits are bounded by distinct radial limits \citep[pericentre and apocentre, e.g.][]{binney2008}, and produce a distinctive ring pattern in the $h_4$ kinematic moment, whereby a torus of negative $h_4$ separates an inner and outer region of positive $h_4$, as seen in fig.
-14 of \citet{frigo2021}.
-We see evidence of a faint ring structure of negative $h_4$ in both the parallel and orthogonal IFU projections of the $\vk=0\,\kmps$ simulation (\autoref{fig:IFU0p} and \autoref{fig:IFU0o}), which agrees with the dominance of rosette orbits at the inner radii of the galaxy merger remnant (\autoref{fig:orbits}).
-Additionally, as rosette orbits are constrained to distinct radial bounds, these orbits contribute a relatively constant mass to the inner stellar density profile. 
-For all simulations with $\vk\geq660\,\kmps$, the majority of stellar particles have $\pi$-box orbits, i.e.
-those orbits with no net angular momentum, and thus move radially and can come arbitrarily close to the galactic potential centre \citep{frigo2021}.
-These orbits also display no resonant motion, unlike their boxlet counterparts; boxlet orbits are the second most dominant orbital family at small radii for $\vk \geq 660\,\kmps$, and are also radially-biased.
-We can gain insight into the probability of finding a radially-moving stellar particle at a position $r+\dd{r}$ by approximating the motion as a classical simple harmonic oscillator. 
-The differential probability $\dd{P}$ that the particle is within the interval $r+\dd{r}$ is proportional to $1/(\xi^2-r^2)$, where $\xi$ is the maximum radial displacement.
-Hence, the likelihood that the particle is at $r\ll\xi$ is much less compared to the likelihood that the particle is at $r\sim\xi$.
-From this simple argument, we expect a deficit of stellar mass in the central regions of the galaxy merger remnant when the majority of the stellar mass belongs to a radially-biased orbit family ($\pi$-box or boxlet).
-We see the manifestation of an increase in radially-biased orbits in the three dimensional stellar density profile (\autoref{fig:density3d}), where the bifurcation in the density profiles occur at the same kick velocity as when $\pi$-box and boxlet orbits start to dominate the inner orbital structure, and a sudden deficit of stellar mass in the central regions of the merger is apparent.
-
-To explain the presence of the bifurcation in the IFU maps, we must investigate how the stellar density is behaving in the wake of the recoiling SMBH.
-We take two representative recoil velocities, $\vk=0\,\kmps$ and $\vk=1500\,\kmps$, and limit our investigation to stellar particles in a rectangular box defined by $0\leq x/\mathrm{kpc}$, $-5\leq y/\mathrm{kpc}\leq 5$, and $-5\leq z/\mathrm{kpc}\leq 5$ (the aperture of the mock IFU maps in \autoref{sec:app_ifu} extended to approximately $5\,\mathrm{kpc}$), recalling that the SMBH is ejected along the positive $x$ axis.
-We create velocity vector fields of the stellar particle motion in the $x$-$y$ plane.
-For the $\vk=0\,\kmps$ case, there is an excess of stellar motion parallel to the $x$-axis, with some orthogonal motion present at small radii. 
-This is contrasted to the $\vk=1500\,\kmps$ case, where there is no clear alignment of the stellar velocity flow with either the $x$- or $y$- axes.
-We interpret this feature as stellar particles diffusing into the potential wake carved by the recoiling SMBH in the $\vk=1500\,\kmps$ case, where the diffusion is neither parallel nor orthogonal to the SMBH motion, but rather at a variety of angles to it.
-
-The IFU maps describe the line of sight velocity distribution, thus the motion along the $x$-axis in \drafting{removed figure} is equivalent to the parallel projection, and the $y$-axis equivalent to the orthogonal projection previously described in \autoref{ssec:ifu}.
-When the SMBH recoil velocity is small, there is no potential wake created by the SMBH, and thus stellar diffusion to fill that wake does not occur, or occurs weakly in the case of a small but non-zero kick velocity.
-As a result, stellar velocities are in general aligned with one of the two observation axes, leading to a narrow distribution of stellar velocities.
-A result of this is the $h_4$ kinematic moment, which describes excess distribution tails, is close to zero: there are few high velocity outliers to the distribution.
-Conversely, when a large SMBH recoil velocity has propelled the SMBH through the galaxy, and stellar particles are diffusing into the potential wake, the variety of incoming angles, coupled with projection effects, results in a wide velocity distribution.
-A result of this is more stellar particles with high velocities, causing high velocity tails to be present in the LOSVD, and hence an increase in the $h_4$ kinematic moment.
-
-\drafting{Need to update, leave this here for now
-As seen in \autoref{fig:orbits}, we observe a clear trend of higher kick velocity remnants having a higher fraction of $\pi$-box and boxlet orbits, and a lower fraction of $x$-tube orbits, and to some extent rosette orbits, than their lower kick velocity counterparts.
-The other two orbital families are generally consistent between all merger remnants. 
-As discussed in \citet{frigo2021}, negative values of $h_4$ indicate regions with weaker tails of the LOS velocity distribution due to the superposition of radial orbits with different orientations.
-This explanation supports our observation of an increase in $\pi$-box orbits (which have no net angular momentum are typically very radial orbits) with kick velocity, and a corresponding decrease in the value of $h_4$ (comparing the IFU maps in \autoref{fig:IFU0} to \autoref{fig:IFU900}).
-Equivalently, $x$-tube orbits (which rotate about the major axis of the inertia tensor) appear as a peaked LOS velocity distribution in the projection we use, leading to extended tails, and hence a more positive value of $h_4$ than for a weak tail distribution.
-All together, we can infer that as a SMBH ploughs through the surrounding stellar environment, regular $x$-tube orbits are disrupted to non-resonant $\pi$-box orbits.
-With a higher kick velocity, the SMBH is able to cumulatively disturb orbits at a larger radial extent, converting a greater fraction of $x$-tube orbits to $\pi$-box orbits, resulting in the increased expanses of negative $h_4$ regions in the kinematic maps.
-This also explains the overall downward shift in the $\langle h_4 \rangle$ profiles in \autoref{fig:h4} for higher kick velocities. 
-A peculiar feature of \autoref{fig:h4} is the consistency in the $\langle h_4 \rangle$ profile shape.
-This can be explained by the derivative of the orbital fraction distribution being consistent across all kick velocities for $x$-tube and boxlet orbits out to $\sim3\,\mathrm{kpc}$, as shown in \autoref{fig:orbits2}.
-}
+\subsection{Comparison to literature}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{core-obs}
+    \caption{
+        Comparison of the median measured core radii for each merger remnant to the observational data and the best fit constant-exponent power law presented in \citet{thomas2016}.
+        The shaded region shows the reported intrinsic scatter of the relation.
+        Only the merger remnants with very low $\vk$ lie within the intrinsic scatter of the relation, whereas remnants with large $\vk$ are significantly offset from the relation.
+        }
+    \label{fig:core-obs}
+\end{figure}
+
+In agreement with previous work \citep{boylan2004,merritt2004}, we find that an SMBH displaced due to gravitational recoil decreases the stellar density in the central regions of the merger remnant as it settles to an equilibrium state.
+Additionally, we find that this oscillatory motion, by the time the SMBH has settled, increases the stellar density core beyond that induced by the scouring of the SMBH binary prior to coalescence, confirming the results of \citet{gualandris2008}, \citet{nasim2021b}, and \citet{khonji2024}.
+We assess the magnitude of this scouring with time using the Lagrangian radii presented in \autoref{fig:lagrangian}.
+We see a sharp increase in the Lagrangian radii immediately after the SMBH binary coalesces.
+If the SMBH is kicked beyond the core, the SMBH generally does not pass through the core during the next few orbits due to tangential motion induced by the stellar potential torquing the SMBH.
+When the SMBH does pass through the core region in the last orbits when the SMBH velocity $v_\bullet$ is only marginally greater than the stellar velocity dispersion within the core, the Lagrangian radius rapidly increases further.
+\citet{gualandris2008} found that each oscillation the SMBH enlarges the core, whereas \citet{nasim2021b} found that only the initial oscillation due to the recoil kick affects the core. 
+Our result suggest that differences in the torquing due to the specific distribution of stellar mass could alleviate some of the tension between the two studies.
+
+The three-dimensional stellar density profile (\autoref{fig:density3d}) appears centrally-flat only in the instance where the SMBH escaped the galaxy entirely ($\vk=2000\,\kmps$).
+Conversely, a mild density cusp $\rho \propto r^{-\nu}$ with $0 < \nu < 1$ is present in simulations where the SMBH settled, $\vk<1020\,\kmps$, with the exponent $\nu$ decreasing with increasing $\vk$, in agreement with previous work \citep[e.g.][]{nasim2021b,khonji2024}.
+All simulations with $\vk<1020\,\kmps$ display a core in projection however, in alignment with previous observational and theoretical work \citep[e.g.][]{ferrarese1994,nakano1999}.
+
+The presence of a core in the three-dimensional stellar density profile for $\vk=2000\,\kmps$ can be understood due to the lack of an SMBH at the centre of the remnant.
+The SMBH, which would ordinarily contribute a point-mass potential that acts as an attractor to stellar particles, is not present, and thus angular momentum-conserving rosette orbits are unable to form or be sustained, thus lowering the central stellar density.
+Instead, we observe $\pi$-box orbits to dominate at all radii for the $\vk=2000\,\kmps$ simulation, in tension with the understanding of box orbits being the dominant orbit family to interact with the SMBH binary prior to coalescence.
+However, we find that the excess of $\pi$-box orbits in the $\vk=2000\,\kmps$ simulation primarily originate from angular momentum conserving orbits (63\%) which have reacted to a sudden change in potential due to the escaping SMBH.
+
+We caution that the claims made in \citet{nasim2021b}, that cores produced by gravitational recoil of the coalesced SMBH are present in the three-dimensional stellar density profile, has the caveat that the recoil velocity must remove the SMBH entirely from the centre of the galaxy remnant, where no oscillations passing through the core are possible.
+We find instead that arbitrary recoil kicks, which have a distribution generally peaking well below the escape velocity of the centre of the merger remnant, do not necessarily result in a truly flat three-dimensional stellar density profile, however do produce a flatter density profile compared to the density profile produced by SMBH binary scouring alone.
+
+\subsection{Determining SMBH recoil velocity in observations}
+In \autoref{ssec:predcore}, we determined the distribution of possible core radii $\rb$ for our merger remnant by Monte Carlo sampling different SMBH spin vectors prior to SMBH binary coalescence, and using the \citet{zlochower2015} relations to determine the expected recoil velocity.
+Our results indicate that GW recoil typically enhances the core region by a factor of 2-3 compared to the core formed through SMBH binary scouring.
+However, this information alone cannot be directly used to determine if GW recoil core formation has occurred in a randomly selected galaxy in the real universe, due to the reference radius, $\rbbin$, being in general unknown.
+
+If the core radius of a randomly selected galaxy is significantly enhanced compared to core radii for other galaxies with a similar stellar or SMBH mass, then this galaxy would be an exciting target for assessing the history of GW recoil.
+In \autoref{fig:core-obs} we plot, as a function of the remnant SMBH mass, the median core radii for each of our merger remnants and compare to the observational data presented in \citet{thomas2016}.
+%What we can do though is assess how our simulated core sizes compare to observed core sizes.
+The inferred median core sizes range from $\rb\sim0.5\,\mathrm{kpc}$ for the $\vk=0\,\kmps$ simulation, to $\rb\sim1.4\,\mathrm{kpc}$ for the $\vk=1020\,\kmps$ simulation.
+Comparing, in \autoref{fig:core-obs}, to the catalogue of core radii in elliptical galaxies presented in \citet{thomas2016} as a function of SMBH mass, we find that the core radii for simulations with $\vk\lesssim240\,\kmps \simeq \sigma_{\star,0}$ are consistent with the observed data to within the intrinsic scatter.
+For recoil velocities in excess of $240\,\kmps$, the core radii start to become significantly offset from the data in \citet{thomas2016}.
+A similar conclusion is reached when comparing to the core radii data presented in \citet{dullo2014}.
+
+Considering that for our initial conditions, the peak in the transformation-sampled core radius distribution is in the range of $1.0\,\mathrm{kpc}$ -- $1.5\,\mathrm{kpc}$ (\autoref{ssec:predcore}, depending on the assumed relation between recoil velocity and core radius), cores in the surface density profile in excess of $1.0\,\mathrm{kpc}$ are viable candidates for GW recoil induced core formation.
+Certainly galaxies hosting cores much greater than $2.0\,\mathrm{kpc}$, as for example in the central galaxy A2261-BCG \citep{bonfini2016, dullo2019}, offer prime targets for SMBH recoil kicks to have been a contributing mechanism to the core formation.
+
+Presuming that a galaxy with a large core radius has been found, assessing the magnitude of the SMBH recoil kick can be achieved with the inclusion of IFU observations of the central regions of the galaxy.
+As we demonstrated in \autoref{ssec:ifu}, IFU maps uniquely allow for an observational connection to the underlying orbital structure of a galaxy, with imprints of the SMBH binary dynamics apparent in particular in the spatial distribution of $h_4$.
+Due to the removal of box orbits from the central regions of the galaxy by the SMBH binary prior to coalescence, a concentration of $h_4\sim0$ appears surrounded by an annulus of $h_4 < 0$.
+We propose that identifying the inner radius of the annulus of $h_4<0$ provides a tight estimate on the core radius due to binary scouring, $\rbbin$.
+
+The appearance of $\rbbin$ is particularly apparent in IFU maps restricted to just box orbits. 
+Whilst orbital structure cannot be directly observed, but rather inferred through Schwarzschild modelling \citep[e.g.][] {rix1997, cappellari2002, thomas2004} which introduces a new source of uncertainties, generating a model of the observed galaxy to produce mock IFU images with, analogous to the technique presented in this work, would provide a means of determining the radial extent of the annulus of $h_4 < 0$ if it were not apparent in the observed IFU maps.
+
+With an estimate of $\rbbin$, we are able to ascertain the relative size of the observed core radius, $\rb/\rbbin$.
+Assuming that a major merger has taken place to produce the observed galaxy, the GW recoil velocity can be inferred by inverting the relation between core radius and recoil velocity, as in our \autoref{eq:vkrb_sig}.
+Presumably this relation between $\vk$ and $\rb$ would depend on the mass ratio of the progenitor SMBH binary, so we caution that the relations we provide are only applicable to a major merger involving two equal mass SMBHs of $M_{\bullet,0}\sim 10^9\,\Msun$.
+However equivalent relations may be found through a targeted parameter scan of SMBH recoil velocities, following our methods presented in \autoref{sec:num_sims}.
+
+In the event of a non-monotonic relation between recoil velocity and core radius (unlike our case, where the relation is monotonic), distinguishing the most likely $\vk$ predicted by a single value of $\rb$ may be achieved through assessing the fraction of rosette orbits present in the Schwarzschild modelling of the observed galaxy.
+As demonstrated in \autoref{fig:orbits}, a larger $\vk$ results in less rosette orbits at the centre of the galaxy.
+
+In this way, the SMBH recoil velocity can be well estimated with a synergy of photometric observations (to obtain $\rb$), IFU observations (to determine the spatial distribution of $h_4$), and potentially Schwarzschild modelling (to assess relative orbit fractions).
+
+\subsection{Implications for local cored galaxies}
+Our work has important implications for the population of local cored elliptical galaxies.
+For elliptical galaxies with similar stellar masses to our simulated mergers ($M_\star \sim 3\times10^{11}\,\Msun$), the observed core radii are generally less than $1\,\mathrm{kpc}$ \citep[e.g.][]{thomas2016,dullo2019}, and the central regions of the galaxy display a tangential bias in the anisotropy parameter $\beta$, seen for example in the SINFONI black hole survey \citep{Thomas2014,saglia2016}.
+These observations suggest that following SMBH coalescence, the typical SMBH recoil velocity is low: specifically, $\vk$ is generally less than the stellar velocity dispersion of the SMBH binary-scoured core $\sigma_{\star,0}$.
+In our work, whilst the SMBH recoil velocity is set \textit{ab initio}, the only physical mechanism by which it could vary is the orientation and magnitude of the SMBH binary spin vectors.
+Zero recoil velocity is achieved through symmetry of the SMBH binary system, and small deviations from symmetry (whether in mass or spin orientations) give rise to small recoil velocities.
+Consequently, our work suggests a bias towards near-aligned SMBH spin vectors, rather than SMBH spin vectors orientated randomly on the sphere.
+The spin vectors of an SMBH binary can evolve as a result of gas accretion and general relativistic effects.
+
+In the case of gas rich mergers, \citet{miller2013} showed that the spin vectors of the SMBH binary are expected to align when gas accretion onto the SMBHs is prograde, and counteralign when the gas accretion is retrograde.
+Similar results have also been found recently by \citet{steinle2023} and \citet{bourne2024}.
+In the case of gas-poor mergers, SMBH binary spins likely remain isotropic until entering the GW-dominated hardening phase \citep[e.g.][]{spadaro2024}, with alignment later occurring through general relativistic effects, particularly in the case of unequal mass SMBHs where the more massive SMBH spin is closely aligned with the orbital angular momentum of the binary \citep{berti2012}.
+However, \citet{sayeb2021} found that whilst general relativistic effects can cause precession of the SMBH binary spin vectors, the overall distribution of resulting recoil velocities remains relatively unaffected, owing to the dominant mechanism for SMBH binary spin alignment being the Bardeen-Peterson effect.
+The alignment of SMBH binary spin vectors remains an open challenge in modern astrophysics, both observationally and theoretically.
+
+\subsection{Modelling caveats and future work}
+It is worthwhile to consider the model limitations of our work, and their potential implications on our conclusions.
+As a first limitation, our models do not incorporate gas physics.
+Work by \citet{blecha2008} and \citet{blecha2011} have suggested that higher gas fractions in merger remnants can act to bring recoiling SMBHs to an equilibrium state more rapidly than in the collisionless case.
+For instance, following the $1:1$ merger of two disc galaxies (leading to a total remnant baryonic mass of $\approx 2\times10^{11}\,\Msun$), \citet{blecha2011} find that for a fixed $v_{\rm kick}$, increasing progenitor disk gas fractions from $0.1$ to $0.3$ can decrease the maximum radius reached by the recoiling SMBH by a factor of $2$-$3$.
+The inclusion of gas in these systems leads to a significant steepening of the central stellar density profile in the remnant due to merger-driven central star formation, as well as additional dynamical friction from the gas itself.
+While \citet{blecha2011} consider systems with similar stellar mass to those studied here, disc-like progenitors are used as opposed to the elliptical progenitors in this work, which for a given stellar mass typically exhibit much lower gas fractions than their disc-like counterparts (e.g.
+\citealt{Cook2019}).
+\citet{liao2024b} directly show that the differences in central stellar density induced by the inclusion of gas in elliptical-elliptical mergers is negligible compared to the disc-disc case, where the authors specifically match the properties of the disc and elliptical progenitors to observed galaxies.
+As such, for the systems considered in this work, we do not expect gas to play a dominant role in the decay of the kicked SMBH orbit.
+
+The inclusion of hydrodynamics may also serve to alter the stellar environment \textit{prior} to SMBH binary coalescence, affecting when precisely the SMBH binary coalesces.
+Higher central stellar densities generally lead to more rapid coalescence of the SMBH binary through three-body interactions and a constantly-refilled loss cone \citep[e.g.][]{chapon2013,souzalima2017,liao2024a}.
+We thus have two timescales to consider: the timescale for the galaxy merger remnant to reach an equilibrium state, and the timescale for the SMBH binary to coalesce.
+Most importantly, due to uncertainty in the SMBH binary merger timescale driven by eccentricity \citep{nasim2020,rawlings2023} and stellar density \citep{liao2024b}, the triaxiality of the merger remnant at the time of SMBH coalescence is not known a priori.
+Additionally, AGN feedback can also alter the central triaxiality of the galaxy merger remnant, though this is a minor contribution in elliptical galaxies \citep{liao2024b}.
+A varying triaxiality may impact the asymmetric torquing of the recoiling SMBH, affecting the frequency and timing of the SMBH passages through the stellar core.
+Consequently, this might in principle alter the size of the core radius of the merger remnant once the SMBH has settled.
+However, as demonstrated in \autoref{fig:trajectory} and discussed in \autoref{ssec:movement}, the passages of the SMBH through the stellar core occur when the SMBH trajectory is predominantly radial, despite a low degree of triaxiality being present in the central regions of the merger remnant at the time of coalescence.
+Thus, we do not expect the number of oscillations required for the kicked SMBH to settle to depend strongly on the triaxiality, however the time required for the SMBH to settle may show some variation.
+From the analysis of the Lagrangian radii, the later stages of the stellar core growth occur during the passages of the SMBH that take it through the core, and not the total time for the SMBH to settle, thus we expect our results to be robust against triaxiality variation.
+
+In this work, we have only explored one set of initial conditions, and not considered differing mass ratios of progenitor galaxies or SMBHs, nor different stellar density profiles.
+Additionally, we consider only a single generation of galaxy mergers, whereas in the cosmological context multiple generations of mergers are expected to have occurred, particularly for massive systems such as the ones we discuss \citep[e.g.][]{Bezanson2009, Naab2009, feldmann2010, vandokkum2015, rantala2024}.
+Whilst exploring the full parameter space is beyond the scope of this work, using the methodology developed here and applying it to a statistically significant sample of galaxy mergers from cosmological simulations would be an interesting future avenue of investigation.
+In particular, a statistically significant sample would require both a large number of galaxy mergers, and a representative, diverse sample of galaxy mergers with an occurrence frequency consistent with cosmology.
+Such an investigation would not only help address the first limitation raised above, but also answer a new question: what is the overall distribution of core radii in the Universe?
+
+Finally, we make use of the \citet{zlochower2015} relations to determine the SMBH recoil velocity from the pre-merger SMBH binary configuration, which are derived from numerical relativity simulations.
+As part of the modelling in this work, we assume that the distribution of spin magnitude $\alpha_\bullet$ at the separation of a binary merger in \ketju{} (some $12\,R_\mathrm{s}$, where $R_\mathrm{s}$ is the Schwarzschild radius of the combined SMBH binary mass) is consistent with the distribution arrived at from numerical relativity, which is necessarily resolved to much smaller scales than our simulations.
+Whilst using a different relation than that presented in \citet{zlochower2015} would not affect the kinematic analysis nor the derived relation between recoil velocity and core radius, the distribution of core radii (\autoref{fig:rb_pdf}) would change.
+A distribution of core radii derived from the method proposed in the third limitation discussed would also be sensitive to this.
 
 % CONCLUSION
-\section{Conclusions}
-\label{sec:conclusions}
+\section{Conclusions}\label{sec:conclusions}
+In conclusion, we find that the GW-induced recoil of an SMBH in a massive elliptical galaxy leaves distinct signatures that would be observable in the real Universe.
+Specifically, we find:
+\begin{enumerate}
+    \item Asymmetric torques on the recoiling SMBH can alter the SMBH trajectory, potentially delaying stellar core growth by repeated passages of the settling SMBH until the SMBH passes through the core.
+    \item A core is not present in the three-dimensional stellar density profile unless the SMBH has been ejected entirely from the galaxy, corresponding to the case of zero oscillations of the kicked SMBH about the galaxy centre.
+    \item An increase in the recoil velocity induces a larger core in the projected stellar mass density of the merger remnant than low recoil velocities, with our models indicating that core radii exceeding $1\,\mathrm{kpc}$ are not only possible, but likely.
+    As a consequence, the mass deficit due to SMBH recoil increases with larger recoil velocity, up to $\sim90$ per cent of the coalesced SMBH mass.
+    \item For merger remnants with $\vk \lesssim \sigma_{\star,0}$, the majority of stellar mass within the core has angular momentum conserving orbits. 
+    For recoil velocities above this threshold, box orbits which do not conserve angular momentum dominate the mass budget.
+    \item Similarly, merger remnants with $\vk \gtrsim \sigma_{\star,0}$ are ergodic within the core, with $\beta\sim 0$.
+    Only for the lowest kick velocities are tangentially-biased cores observed, suggesting that the majority of local cored galaxies do not have an SMBH which has experienced a large recoil velocity.
+    \item In kinematic observations of the inner region of the merger remnant, the core due solely to SMBH binary scouring, without a contribution from SMBH recoil, can be estimated by identifying the inner radius when the LOSVD symmetric deviation parameter $h_4$ decreases to $h_4 < 0$.
+    This in principle allows for an estimate of $\vk$ from the observed core radius.
+\end{enumerate}
+
+Taken together, this work establishes a causal link between forward-modelled observations of a galaxy merger remnant -- both photometric and kinematic -- with the complex physics underlying SMBH binary mergers, and their associated GW-induced recoil kicks.
+In the future, this will allow us to quantify the recoil velocity of SMBHs in observations of major mergers, opening the way for tighter constraints to be placed on the SMBH binary merger process, particularly for those galaxies in the local Universe.
 
@@
 \section*{acknowledgments}
+We thank Jens Thomas for useful discussions on IFU observations.
+\alex{\textbf{Everyone: please add your acknowledgements here}}
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
-P.H.J., acknowledge the support by the European Research Council via ERC Consolidator Grant KETJU (no. 818930) and the support of the Academy of Finland grant 339127.
+M.M., P.H.J., and D.I. acknowledge the support by the European Research Council via ERC Consolidator Grant KETJU (no. 818930).
+N.K. acknowledges support of the Research Council of Finland Flagship programme: Finnish Center for Artificial Intelligence and the Finnish Foundation for Technology Promotion.
+S.L. acknowledges the supports by the National Natural Science Foundation of China (NSFC) grant (no. 12473015, 11988101) and the K. C. Wong Education Foundation.
+P.H.J acknowledges the support of the Academy of Finland grant 339127.
+TN acknowledges support from the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany’s Excellence Strategy - EXC-2094 - 390783311 from the DFG Cluster of Excellence ``ORIGINS''.
 
@@
 We list here the roles and contributions of the authors according to the Contributor Roles Taxonomy (\href{https://credit.niso.org}{CRediT}). 
-\textbf{AR}:
+\textbf{AR}: conceptualisation, methodology, formal analysis, investigation, data curation, writing: original.
+\textbf{AK}: formal analysis, writing: original.
+\textbf{MM}: formal analysis, writing: original.
+\textbf{SS}: formal analysis, writing: original.
+\textbf{RJW}: investigation, writing: original.
+\textbf{NK}: methodology, writing: original.
+\textbf{SL}: conceptualisation, writing: review.
+\textbf{AR}: writing: review.
+\textbf{PHJ}: writing: original, supervision, funding acquisition.
+\textbf{TN}: conceptualisation, writing: review.
+\textbf{DI}: writing: review.
 
@@
 CmdStanPy \citep{standevelopmentteam2018},
-Arviz \citep{kumar2019}.
+Arviz \citep{kumar2019},
+Seaborn \citep{waskom2021}.
 
@@
 \section{Triaxiality}\label{sec:app_triax}
-We present the triaxiality of all sixteen merger remnants at the time $t \simeq 0.075\,\mathrm{Gyr}$. 
-We determine the ratios $b/a$ and $c/a$, where $a$, $b$, and $c$ are the eigenvalues of the reduced inertia tensor \citep{gerhard1983,bailin2005}
-\begin{equation}
-    \tilde{I}_\mathrm{red} = \sum_k \frac{r_{k,i} r_{k,j}}{r_k^2}
+To assess the impact of our (arbitrarily) chosen direction in which the SMBH recoil velocity is applied, we calculate the triaxiality of all merger remnants at the time when the SMBH has settled.
+
+We take all stellar particles within a fraction of the virial radius $R_\mathrm{vir}$, defined as the radius at which the mean density is equal to 200 times greater than the critical density; the virial radius is $\sim450\,\mathrm{kpc}$ for the merger remnants in this study.
+
+We determine the ratios $b/a$ and $c/a$, where $a$, $b$, and $c$ are the eigenvalues of the reduced inertia tensor \citep{gerhard1983,bailin2005}:
+\begin{equation}\label{eq:inertia}
+    \tilde{I}_\mathrm{red} = \sum_k \frac{r_{k,i} r_{k,j}}{r_k^2},
 \end{equation}
 and $c\leq b\leq a$. 
-The tensor $\tilde{I}_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} R_\mathrm{vir}$ to $10^{-1} R_\mathrm{vir}$: i.e., the binning of particles is not cumulative.
-The variation of the ratios $b/a$ and $c/a$ is plotted as a function of radius in \autoref{fig:triax}.
-The merger remnants display triaxial cores, albeit without dependence on the recoiling SMBH velocity. 
-At radii greater than $4\,\mathrm{kpc}$, the triaxiality of the merger remnants is almost identical.
+The tensor $\tilde{I}_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} R_\mathrm{vir}$ to $10^{-1} R_\mathrm{vir}$: the binning of particles is not cumulative.
+We apply a uniform filter to reduce Poisson noise due to low particle numbers at small radii.
+The radial variations of the ratios $b/a$ and $c/a$ are plotted in \autoref{fig:triax} in the top and middle panels, respectively.
+The merger remnants display mildly triaxial cores, albeit without dependence on the recoiling SMBH velocity. 
+At radii greater than $4\,\mathrm{kpc}$, the triaxiality of the merger remnants is consistent.
+Importantly, the $\vk=0,\kmps$ case also displays only minor variation in $b/a$ and $c/a$ with radius: as this is a good proxy for the pre-merger galaxy remnant, we do not expect triaxiality to dominate over the recoil velocity in the motion of the settling SMBH.
+
+To quantify the nature of the merger remnant triaxiality as a function of radius, we also compute the triaxiality parameter $T$:
+\begin{equation}
+    T = \frac{1 - (b/a)^2}{1 - (c/a)^2}
+\end{equation}
+following \citet[][]{binney2008}.
+Values of $T$ approaching 1 indicate a prolate system, and values of $T$ approaching 0 indicate an oblate system.
+As seen in the bottom panel of \autoref{fig:triax}, at radii $r\gtrsim 5\,\mathrm{kpc}$ all merger remnants are prolate systems.
+For radii $r\lesssim 5\,\mathrm{kpc}$, only the $\vk=1020\,\kmps$ merger remnant tends to an oblate system.
 
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{triaxiality}
-    \caption{
-        Triaxiality ratios $b/a$ and $c/a$ as a function of radius for all merger remnants in the analysis.
+    \includegraphics[width=0.48\textwidth]{triaxiality}
+    \caption{
+        Triaxiality ratios $b/a$ and $c/a$ as a function of radius for all merger remnants in the analysis in the top panel and middle panel, respectively.
         Lines are colour coded by the kick velocity, with the same scheme as in \autoref{fig:density3d}.
         Note that the $\vk=0\,\kmps$ case is equivalent to the pre-merger remnant.
+        In the bottom panel, the variation of the triaxiality parameter $T$ is shown as a function of radius: the merger remnants in this study are predominantly prolate systems.
     }
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.48\textwidth]{all-kick}
+    \caption{
+        Box plots of the six parameters in the core-S\'ersic model (from top left to bottom right) as a function of the kick velocity: the core radius normalised to the pre-merger core radius $\rb/\rbbin$, the density at the core radius $\log_{10}\Sigma_\mathrm{b}$, the core slope $\gamma$, the effective radius $\Reff$, the profile transition index $\alpha$, and the S\'ersic index $n$.
+    }
+    \label{fig:csparams}
+\end{figure}
+
+We use a Bayesian hierarchical model to determine the posterior distributions for the core-S\'ersic projected mass density profile \autoref{eq:cs}.
+Bayesian inference, in its simplest form, aims to infer the distribution of the parameter $\theta$ in some model that describe data $y$: this is done with Bayes' formula:
+\begin{equation}\label{eq:bayes}
+    p(\theta | y) \propto p(\theta) \mathcal{L}(y | \theta),
+\end{equation}
+where $p(\theta | y)$ is the posterior distribution, $p(\theta)$ is the prior distribution that encodes belief about how the parameter $\theta$ is distributed, and $\mathcal{L}(y | \theta)$ is the likelihood of observing the data $y$ given the model parameter $\theta$.
+In the case that there are groups of data that are subsets of some global population, and these subsets are exchangeable in the sense that there is no distinguishing order to them, we can model each subset of data $y_j$ as having its own value of the model parameter $\theta_j$.
+The distribution of all values of $\theta_j$ can then be modelled as samples from a global distribution, termed a hyperdistribution, with hyperparameter(s) $\phi$: this is where the hierarchical nature of the model is introduced.
+A prior distribution can then be assigned to the hyperparameter $\phi$, $p(\phi)$.
+As such, \autoref{eq:bayes} then takes the form:
+\begin{equation}
+    p(\theta, \phi | y) \propto p(\phi) p(\theta|\phi) \mathcal{L}(y | \theta, \phi).
+\end{equation}
+
+In our application of inferring the parameter vector $\vb{\theta}\equiv[\rb, \Sigma_\mathrm{b}, \gamma, \Reff, n, \alpha]$ that describes the projected stellar mass density profile of the galaxy merger remnant, each projection from which the merger remnant is viewed from constitutes a subset of the surface density data $\Sigma_j$.
+As each of the projections are unordered and exchangeable (i.e., there is no distinguishing label for the projections), and each projection is of the same merger remnant, the hierarchical model naturally lends itself.
+For each projection, we thus infer the six parameters of $\vb{\theta}$, and assume that these values are draws from a common, global distribution of all possible $\vb{\theta}$ vectors.
+Hence, we are effectively inferring the global hyperparameters $\vb{\theta}^\mathrm{hyp}$ that describe the latent parameters $\vb{\theta}$ that we are interested in.
+This idea is demonstrated in the directed acyclic graph (DAG) for the model, shown in \autoref{fig:dag}.
+The posterior density can thus be constructed for each projection- and radially-dependent surface density element $\Sigma_{i,j}^\mathrm{L} \equiv \log_{10}(\Sigma_{i,j})$ as:
+\begin{multline}
+    p(\vb{\theta}, \vb{\theta}^\mathrm{hyp} | \Sigma_{i,j}') \propto p\left(\vb{\theta}^\mathrm{hyp}\right) p\left(\vb{\theta} | \vb{\theta}^\mathrm{hyp}\right) \\ \times \prod_j^{N_\mathrm{proj}} \prod_i^{N_R} \mathcal{N}\left(\Sigma_{i,j}^\mathrm{L} | \hat{\Sigma}^\mathrm{L}\left(R_i | \vb{\theta}_j, \vb{\theta}^\mathrm{hyp}\right), \tau_i\right),
+\end{multline}
+where we use a normal distribution\footnote{Note that the normal distributions $\mathcal{N}$ are parameterised using the standard deviation $\sigma$ in this work, as opposed to the variance $\sigma^2$ as is common in many statistics references.} as the likelihood function, and the error in the fit depends on the radial bin $i$.
+
+We use weakly-constrained prior distributions for each of the hyperparameters in $\vb{\theta}^\mathrm{hyp}$, where these prior distributions are chosen to reflect reasonable values \citep{gabry2019} that the parameters may take (e.g.
+distributions of a radial quantity are chosen so as to be positively-constrained).
+The distributions of the hyperparameters are given in \autoref{tab:hyper}.
+
+We fit the model with \textsc{Stan} \citep{standevelopmentteam2018} using the No U-turn sampler (NUTS, a variant of Hamiltonian Monte Carlo (HMC)) with four chains each of 4000 iterations, of which the first 2000 are discarded as warmup.
+This resulted in 8000 draws from the posterior distribution of $\vb{\theta}$, whereby we ensure the quality of the HMC fit by enforcing a diagnostic value $\hat{R}$ (comparison of between-chain and within-chain variance) less than 1.05, the effective sample size is high, and that the number of divergent transitions is less than 5 per cent \citep[for details see][]{vehtari2021,standevelopmentteam2018}.
+We perform additional prior sensitivity tests using the power-scaling method described in \citet{kallioinen2024}, ensuring that the sensitivity diagnostic for prior power-scaling is less than 0.05 for the latent parameters $\vb{\theta}$ of the model.
+This recommended threshold indicates that the sampled posterior distributions of the latent parameters are not sensitive to the particular prior distributions used.
+
+In \autoref{fig:csparams}, we show as box plots the distributions of the six latent parameters in the core-S\'ersic profile, \autoref{eq:cs}.
+The dark bars indicate the median of each distribution, and the boxes are bounded at the bottom by the $\ordinal{25}$ percentile, and capped at the $\ordinal{75}$ percentile.
+As discussed in the main text, the median core radius $\rb$ (top left) increases in both magnitude and variance with increasing kick velocity.
+The density at the core radius (top left) decreases with increasing kick velocity, as well as having a variance which is largely independent of the kick velocity.
+The slope of the inner density profile, $\gamma$ (centre left) is largely independent of the kick velocity, as is the effective radius of the galaxy (centre right) and transition index (bottom left).
+The S\'ersic index (bottom right) is constant for $\vk \leq 300\,\kmps$, and is constant, albeit with a smaller value, for $\vk \geq 660\,\kmps$.
+For the intermediate kick velocities, there appears to be a linear decrease in the S\'ersic index $n$.
+
+The joint relation between the latent parameters in \autoref{eq:cs} can be seen in the representative example of $\vk=600\,\kmps$ in \autoref{fig:cscorner}.
+The leading diagonal shows the marginal distributions of each latent parameter; each distribution is unimodal.
+The contours in the lower diagonal triangle show the joint posterior distribution between each of the latent parameters in \autoref{eq:cs}, with the contour colouring indicating various highest density intervals (HDIs).
+In no pairing of parameters is a strong covariance observed; the marginal distributions are orthogonal to each other.
+One case worth a special mention is the width of the 25\% HDI for the parameter $\alpha$.
+The wide HDI indicates that a large range of $\alpha$ values maximise the posterior distribution.
+However, as discussed in \citet{graham2003}, values of $\alpha\gtrsim 3$ have a similar effect to induce a sharp transition between the inner and outer surface density profiles.
+The variation of $\alpha$, particularly to large values greater than 10, are not expected to have a strong effect on the overall surface density profile.
+
 \begin{figure*}
     \centering
-    \includegraphics[width=\textwidth]{all-kick}
-    \caption{
-        Box plots of the six parameters in the core-S\'ersic model (from top left to bottom right) as a function of the kick velocity: the core radius normalised to the pre-merger core radius $\rb/r_{\mathrm{b},0}$, the density at the core radius $\log_{10}\Sigma_\mathrm{b}$, the core slope $\gamma$, the effective radius $\Reff$, the profile transition index $\alpha$, and the S\'ersic index $n$.
-    }
-    \label{fig:csparams}
+    \includegraphics[width=\textwidth]{corner_latent_600}
+    \caption{
+        Corner plot of the latent parameters in the core-S\'ersic model for a representative example, here the $\vk=600\,\kmps$ instance.
+        Contours indicate, from dark to light blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs).
+        All distributions are unimodal with little cross-correlation between the latent variables.
+    }
+    \label{fig:cscorner}
 \end{figure*}
 
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{corner_latent_600}
-    \caption{
-        \drafting{NEED TO UPDATE}
-        Corner plot of the latent parameters in the core-S\'ersic model for the $\vk=600\,\kmps$ instance.
-        Contours indicate, from dark to light blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs), with sample draws beyond the 99\% HDI shown as individual points. 
-        All distributions are unimodal with little cross-correlation between the latent variables.
-    }
-\end{figure*}
-
-\section{Specific IFU maps}\label{sec:app_ifu}
-\drafting{Here we include both paralle and orthogonal projections of the IFU maps for three representative kick velocities: $\vk=0\,\kmps$, and $\vk=660\,\kmps$. Add others?}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_para_v0000}
-    \caption{
-        Mock integral field unit kinematic maps of the $\vk=0\,\kmps$ remnant in the parallel projection, plotted out to $0.25\,\Reff$ in a circular aperture. 
-        The small ordered rotation $V$ varies little in magnitude ($20\,\kmps$), and the velocity dispersion $\sigma$ is centrally-concentrated, peaking at around $\sim260\,\kmps$.
-        The central regions of the merger remnant have a $h_4$ value that is predominantly $0\,\kmps$, and positive $h_4$ at larger radial extents.
-    }
-    \label{fig:IFU0p}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_ortho_v0000}
-    \caption{
-        The same as \autoref{fig:IFU0p}, but in the orthogonal projection. 
-        Note the colour scaling is consistent.
-        Qualitatively, there is little difference between the orthogonal and parallel projections for the $\vk=0\,\kmps$ case.
-    }
-    \label{fig:IFU0o}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_para_v0660}
-    \caption{
-        Mock integral field unit kinematic maps of the $\vk=600\,\kmps$ remnant in the parallel projection, plotted out to $0.25\,\Reff$ in a circular aperture. 
-        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
-        Most striking is the slightly enhanced ordered velocity $V$ (some $20\,\kmps$) and the ring of negative $h_4$.
-    }
-    \label{fig:IFU600p}
-\end{figure*}
-
-\begin{figure*}
-    \centering
-    \includegraphics[width=\textwidth]{IFU_ortho_v0660}
-    \caption{
-        The same as \autoref{fig:IFU600p}, but in the orthogonal projection. 
-        Note the colour scaling is consistent with \autoref{fig:IFU0p}.
-        Compared to the parallel case, the ordered velocity $V$ and $h_3$ moments appear consistent with the $\vk=0,\kmps$ case, however the ring of negative $h_4$ persists.
-    }
-    \label{fig:IFU600o}
-\end{figure*}
+\section{Supporting information for IFU analysis}\label{sec:app_ifu}
+In \autoref{fig:apo_dist}, we show, for each recoil velocity,  kernel density estimates of the apocentre distributions for stellar particles within a mean position of $0.25r_{1/2}$ from the merger remnant centre.
+The distributions are clearly left-skewed, and becoming increasingly skewed as the recoil velocity increases, with the exception of the $\vk=2000\,\kmps$ simulation, as discussed in the main text.
+
+In \autoref{fig:h4900}, we show how the radial dependence of $\langle h_4 \rangle$ varies with regard to the time when the SMBH has settled, defined $t-t_\mathrm{settle}=0\,\mathrm{Gyr}$, for a representative example ($\vk=900\,\kmps$).
+At the time the SMBH settled, and times following this point, the radial dependence of $\langle h_4 \rangle$ is time-invariant, for both the parallel and orthogonal projections.
+This ensures that our conclusions in the main text are not sensitive to the exact moment the LOSVD is constructed.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{apo.pdf}
+    \caption{
+        The distribution of apocentres for stellar orbits with a mean position within $0.25r_{1/2}$ (matching the mock IFU maps), determined using the orbit analysis routine described in \autoref{sec:orbits}.
+        All the distributions are left-skewed, where simulations with higher $\vk$ are more left-skewed than those with lower $\vk$.
+    }
+    \label{fig:apo_dist}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\textwidth]{h4_900.pdf}
+    \caption{
+        The dependence of the $\langle h_4 \rangle$ profile on the time of analysis for a representative example $\vk=900\,\kmps$.
+        At early times, when the SMBH is still oscillating through the stellar medium, the $\langle h_4 \rangle$ profile is overall elevated, and decreases as the SMBH begins to settle towards an equilibrium point.
+        For times after the SMBH has settled (according to the definition presented in \autoref{ssec:settle}), the profiles are consistent in both the parallel and orthogonal projections of $\langle h_4 \rangle$.
+        The profile corresponding to $t-t_\mathrm{settle}=0\,\mathrm{Gyr}$ is accentuated with circle markers.
+    }
+    \label{fig:h4900}
+\end{figure}
 

### 2024-10-21 papers/paper-corekick/paper-corekick.tex
Paper version as submitted to arXiv

@@
 $^{3}$
-Department of Computer Science, Aalto University, Espoo, Finland
+Department of Computer Science, Konemiehentie 2, 02150, Aalto University, Espoo, Finland
 \\
@@
 $^{6}$
-The Institute of Cancer Research, London SW7 3RP, UK
+The Institute of Cancer Research, 123 Old Brompton Road, London SW7 3RP, United Kingdom
 \\
@@
 \begin{abstract}
-\alex{We have 250 words to use here. Currently 247 excluding the 3 math inlines.}
-
-We study the contribution to stellar core growth in simulations of merging massive ($M_\star>10^9\,\Msun$) elliptical galaxies by a coalesced supermassive black hole (SMBH) displaced by gravitational wave (GW) induced recoil velocity.
-With controlled, dense sampling of the SMBH recoil velocity, we find the stellar core originally formed by SMBH binary scouring can grow by a factor of 2-3 with increasing recoil velocity, and a similar growth in the central mass deficit.
+We study stellar core growth in simulations of merging massive ($M_\star>10^{11}\,\Msun$) elliptical galaxies by a supermassive black hole (SMBH) displaced by gravitational wave induced recoil velocity.
+With controlled, dense sampling of the SMBH recoil velocity, we find the core radius originally formed by SMBH binary scouring can grow by a factor of 2-3 when the recoil velocity exceeds $\sim50$ per cent of the central escape velocity, and the mass deficit grows by up to a factor of $\sim4$.
 Using Bayesian inference we predict the distribution of stellar core sizes formed through this process to peak at $\sim1\,\mathrm{kpc}$.
-An orbital decomposition of stellar particles within the core reveals that when the recoil velocity exceeds the velocity dispersion of the core, radial orbits dominate over tube orbits, providing a mass 10\% of the coalesced SMBH mass, whereas tube orbits dominate for the lowest recoil kicks.
-A change in orbital structure is reflected in the anisotropy parameter, with a tangential bias present only for kick velocities less than the local stellar velocity dispersion.
-With mock integral field unit observations of the stellar line of sight velocity distribution emulating the capabilities of current detectors, we observe a distinct signature in the Gauss-Hermite symmetric deviation coefficient $h_4$ that allows for an estimate of the core size due to binary scouring, irrespective of later evolution due to SMBH recoil.
-Our results provide a method to estimate the SMBH recoil magnitude from observations of local elliptical galaxies, and implies these galaxies primarily experienced recoil velocities less than the stellar velocity dispersion of the core.
+An orbital decomposition of stellar particles within the core reveals that radial orbits dominate over tube orbits when the recoil velocity exceeds the velocity dispersion of the core, whereas tube orbits dominate for the lowest recoil kicks.
+A change in orbital structure is reflected in the anisotropy parameter, with a central tangential bias present only for recoil velocities less than the local stellar velocity dispersion.
+Emulating current integral field unit observations of the stellar line-of-sight velocity distribution, we uncover a distinct signature in the Gauss-Hermite symmetric deviation coefficient $h_4$ that uniquely constrains the core size due to binary scouring.
+This signature is insensitive to the later evolution of the stellar mass distribution due to SMBH recoil.
+Our results provide a novel method to estimate the SMBH recoil magnitude from observations of local elliptical galaxies, and implies these galaxies primarily experienced recoil velocities less than the stellar velocity dispersion of the core.
 \end{abstract}
@@
 \section{Introduction}
-%There is much evidence for the presence of supermassive black holes (SMBHs) at the centres of massive galaxies, with the shadow of one such SMBH being directly imaged \citep{eht2019}.
-%As two galaxies dynamically interact during a galaxy merger, so too is it expected that the residing SMBHs will dynamically interact.
 Observations have revealed a clear bimodality in the properties of massive elliptical galaxies, with brighter ellipticals $(M_{B}\lesssim -20.5)$ typically having `cored' flat central surface brightness profiles, boxy isophotes and showing little or no rotation (slow rotators).
-Fainter intermediate-mass ellipticals $(-20.5 \lesssim M_{B} \lesssim -18.5)$ on the other hand typically show steeper `cuspy' central brightness profiles, more disc-like isophotes and evidence for more rotational support (fast rotators) (e.g \citealt{Kormendy1996,Faber1997,emsellem2007,Kormendy2009,Thomas2014,dullo2014,dullo2019,Quenneville2024}).
+Fainter intermediate-mass ellipticals $(-20.5 \lesssim M_{B} \lesssim -18.5)$ on the other hand typically show steeper `cuspy' central brightness profiles, more disc-like isophotes and evidence for more rotational support (fast rotators) (e.g.
+\citealt{Kormendy1996,Faber1997,emsellem2007,Kormendy2009,Thomas2014,dullo2014,dullo2019,Quenneville2024}).
 
@@
 Second, after the mass enclosed within the orbit of the two SMBHs is comparable to the sum of the SMBH masses, the SMBHs form a bound binary \citep{merritt2005}.
-Iterative interactions of the SMBHs with surrounding stars on low angular momentum orbits ejects these stars with high velocity \citep[e.g.][]{hills1980,hills1983,quinlan1996}, thus removing orbital energy and angular momentum from the SMBH binary orbit.
+Iterative interactions of the SMBHs with surrounding stars on low angular momentum orbits eject these stars with high velocity \citep[e.g.][]{hills1980,hills1983,quinlan1996}, thus removing orbital energy and angular momentum from the SMBH binary orbit.
 The removal of stars from the central regions by this slingshot mechanism produces a `core' \citep[e.g.][]{lauer1983,lauer1985,kormendy1984}, or depletion, in the luminosity profile of the remnant galaxy \citep[e.g.][]{begelman1980,hills1983,quinlan1996,rantala2018,rantala2024,nasim2021b,khonji2024,partmann2024}, that can in extreme cases extend up to $3\,\mathrm{kpc}$ \citep{postman2012}.
 Finally, at very small, sub-parsec separations the SMBH binary loses its remaining orbital energy and angular momentum through gravitational wave (GW) emission, driving the SMBH binary to coalescence \citep{peters1963,peters1964}.
-The GW emission from such SMBH coalescence events are prime observational targets for ongoing pulsar timing array (PTA) detectors for SMBHs with masses of $M_\bullet \gtrsim 10^9\,\Msun$ \citep{agazie2023, 2023Antoniadis, 2023Xu, zic2023}, and also the forthcoming Laser Interferometer Space Antenna (LISA) mission for SMBHs with slightly lower masses of $M_\bullet \lesssim 10^8\,\Msun$ \citep{amaro-seoane2023}.
+The GW emission from such SMBH coalescence events are prime observational targets for ongoing pulsar timing array (PTA) detectors for SMBHs with masses of $M_\bullet \gtrsim 10^9\,\Msun$ \citep{agazie2023, antoniadis2024, 2023Xu, zic2023}, and also the forthcoming Laser Interferometer Space Antenna (LISA) mission for SMBHs with slightly lower masses of $M_\bullet \lesssim 10^8\,\Msun$ \citep{amaro-seoane2023}.
 
@@
 \citealt{madau2004,mannerkoski2022}), though not so many so as to introduce considerable scatter into the observed relation between SMBH mass and stellar velocity dispersion \citep{volonteri2007}.
-This is in clear tension with all current observations indicating that no massive elliptical galaxy in the local Universe is without an SMBH, and demands further investigation.
 
@@
 During these three stages, the stellar density core is enhanced by the motion of the SMBH, producing a mass deficit in some cases up to five times the SMBH mass (see also \citealt{merritt2004,boylan2004,partmann2024}).
-However, a more recent study by \citet{nasim2021b} on the contribution to stellar core expansion through GW recoil of an SMBH found intriguingly that the greatest contribution to the stellar mass deficit is due to the first excursion of the SMBH following its ejection, with subsequent oscillations only providing a minor contribution to the growth of the stellar core.
-%This result is in tension with the work of \citet{gualandris2008}.
-
-%The effect of recoiling SMBHs have been studied previously, with some of the first explorations presented in \citet{boylan2004} and \citet{merritt2004}.
-%In the work of \citet{boylan2004}, a recoiling SMBH was found to induce a core in the stellar density profile, particularly when the SMBH remained bound to the bulge of the host galaxy and had a recoil velocity comparable to the bulge velocity dispersion.
-%A recoiling SMBH was also found to induce a stellar density core in the work of \citet{merritt2004}.
-%Additionally, the time for a recoiling SMBH to settle after being ejected in a spherical potential was found to be roughly one orbital period when the recoil velocity was less than the velocity dispersion of the bulge.
-%This, which are more representative of a galaxy merger remnant than the spherical case.
-
-%More recently, the contribution of stellar core expansion through GW recoil of an SMBH has been considered by \citet{nasim2021b}.
-%Intriguingly, the authors find that greatest contribution to the stellar mass deficit is due to the first excursion of the SMBH following its ejection, with subsequent oscillations only providing a minor contribution to the growth of the stellar core.
-%This result is in tension with the work of \citet{gualandris2008}.
-
-%The effect of recoiling SMBHs have been studied previously, with some of the first explorations presented in \citet{boylan2004} and \citet{merritt2004}.
-%In the work of \citet{boylan2004}, a recoiling SMBH was found to induce a core in the stellar density profile, particularly when the SMBH remained bound to the bulge of the host galaxy and had a recoil velocity comparable to the bulge velocity dispersion.
-%A recoiling SMBH was also found to induce a stellar density core in the work of \citet{merritt2004}.
-%Additionally, the time for a recoiling SMBH to settle after being ejected in a spherical potential was found to be roughly one orbital period when the recoil velocity was less than the velocity dispersion of the bulge.
-%This `infall time' was found to increase when considering triaxial potentials, which are more representative of a galaxy merger remnant than the spherical case.
-
-%In \citet{gualandris2008}, a stellar potential representative of a galaxy merger remnant with SMBH binary core scouring is used to investigate the motion of an SMBH ejected through GW recoil.
-%The motion of a settling SMBH was found to occur in three stages: first, the SMBH oscillates with decreasing amplitude due to energy dissipation by dynamical friction.
-%Next, both the SMBH and stellar core oscillate about their common centre of mass when the radial excursion of the SMBH is less than the core radius of the stellar system.
-%Finally, the SMBH settles into thermal equilibrium with the surrounding stars.
-%During these three stages, the stellar density core is enhanced by the motion of the SMBH (in agreement with \citet{merritt2004} and \citet{boylan2004}), with the mass deficit in some cases reaching up to five times the SMBH mass being produced.
-
-%Work by \citet{blecha2008} considered the effect of recoiling SMBHs in an analytical model of an isolated galaxy comprised of a multicomponent system of a stellar bulge and gaseous disc.
-%An interesting result was the mass growth of the ejected SMBH as it settled at the centre of the potential, with the mass growth accumulating to some 10\% of the initial SMBH mass, irrespective of the recoil velocity.
-%Recoiling SMBHs in gas-rich systems was further explored in \citet{blecha2011} using hydrodynamic simulations of isolated galaxy mergers.
-%In this work, seed SMBHs were allowed to grow through gas accretion prior to merging, with the majority of final SMBH masses at the time of merger sampling the $10^7\,\Msun \lesssim M_\bullet \lesssim 10^8\,\Msun$ mass regime.
-
-%The impact of recoiling SMBHs on the growth and dynamics of SMBHs in hydrodynamical simulations have also been studied using both isolated galaxy models consisting of a stellar bulge and a gaseous disc \citep{blecha2008}, as well as in isolated gas-rich galaxy mergers \citep{blecha2011}.
-%The study of the isolated galaxy model found that
-%the mass growth of the ejected SMBH as it settled at the centre of the potential, with the mass growth accumulating to some 10\% of the initial SMBH mass, irrespective of the recoil velocity. 
-%In the merger simulations, the seed SMBHs were allowed to grow through gas accretion prior to merging, with the majority of the final SMBH masses at the time of merger sampling the $10^7\,\Msun \lesssim M_\bullet \lesssim 10^8\,\Msun$ mass regime.
-The simulations also showed that recoil events typically reduce the lifetimes of bright active galactic nuclei (AGNs), with most kicked AGNs confined to the central few kpc of the galaxy.
-Whilst the results of \citet{blecha2011} provide valuable insight into the occurrence of variable AGN variability due to an oscillating SMBH plunging through a gaseous disc, we stress that the results are applicable to a different mass regime of galaxies and SMBHs than what we focus on in this work.
-In addition, the formation of stellar cores which is primarily driven by complex three-body interactions between the SMBHs and the stellar component remains largely unresolved in simulations that include gravitational softening.
+However, a recent study by \citet{nasim2021b} on the contribution to stellar core expansion through GW recoil of an SMBH found intriguingly that the greatest contribution to the stellar mass deficit is due to the first excursion of the SMBH following its ejection, with subsequent oscillations only providing a minor contribution to the growth of the stellar core.
 
@@
 
-%In this study, we wish to consider SMBH recoil by GW emission in massive elliptical galaxies.
-%The aim of this work is twofold.
-%We first aim to determine how the velocity with which the SMBH is kicked affects the mass distribution of the surrounding stellar environment, and the likelihood of the magnitude of these effects.
-%We then aim to determine if a recoiling SMBH imparts a signature on the stellar kinematics of the remnant galaxy, and if such a signature could be observed with current detectors.
-
 This paper is divided as follows: we present the numerical simulations used in our investigation in \autoref{sec:num_sims}.
@@
 We investigate the distribution of stellar orbits as a function of the kick velocity in \autoref{sec:orbits} and construct mock stellar kinematic observations to be compared with observational data in \autoref{sec:kinematics}.
-Finally, we discuss of our results in \autoref{sec:discussion} and present 
+Finally, we discuss our results in \autoref{sec:discussion} and present 
 our conclusions in \autoref{sec:conclusions}, respectively.
@@
 To investigate the effect of the recoiling SMBH on the host galaxy, and potential observational signatures of this, we run a number of numerical simulations of an idealised galaxy merger setting using the new public version of the \ketju{}\footnote{\url{https://www.mv.helsinki.fi/home/phjohans/group-website/research/ketju/}} code \citep{rantala2017,rantala2020,mannerkoski2023} coupled with \gadget{} \citep{springel2021}.
-\ketju{} integrates the dynamics of SMBHs, and stellar particles in a small region thrice the stellar softening length around them (i.e.
+\ketju{} integrates the dynamics of SMBHs, and stellar particles in a small region three times the stellar softening length around them (i.e.
 the \ketju{} region), to high accuracy using the algorithmically regularised integrator \textsc{mstar} \citep{rantala2020}.
@@
 As we are interested in the evolution of the galaxy merger \textit{remnant} following SMBH coalescence, the initial conditions are constructed such that at the time of SMBH coalescence (\autoref{ssec:mergers}) the galaxy remnant agrees with local scaling relations.
-Specifically, we ensure that our remnant agrees with the half-light -- stellar mass data presented in \citet{sahu2020}, and lies within the $1\sigma$ prediction interval of the SMBH mass -- stellar velocity dispersion data given in \citet{vandenbosch2016}.
+Specifically, we ensure that our remnant agrees with the half-light -- stellar mass data presented in \citet{sahu2020}, and lies within the $1\sigma$ prediction interval of the SMBH mass -- stellar velocity dispersion data as given in \citet{vandenbosch2016}.
 
 \subsection{Merger Simulations}\label{ssec:mergers}
-We merge two independently Monte Carlo sampled galaxy ICs, creating an equal mass galaxy merger with a near-radial orbit. 
+We merge two independently Monte Carlo sampled galaxy ICs, creating an equal-mass galaxy merger with a near-radial orbit. 
 The merger orbit has an initial separation consistent with \citet{rantala2018} of $D=30\,\mathrm{kpc}$, however the initial eccentricity is set to $e_0=0.97$, and the first pericentre distance adjusted accordingly to $r_\mathrm{peri}=D(1-e_0^2) \simeq 2\,\mathrm{kpc}$ \citep{khochfar2006,rantala2017}.
@@
 To test the effect of differing kick directions on the merger remnant, we test directing the recoil kick along the global $y$-axis, and find very similar evolution to the $x$-axis case.}, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
-%At the moment of the merger, the centre of mass (CoM) velocity of the SMBH binary is $v_{\mathrm{bin},\mathrm{CoM}}\sim[4.2, 1.1, 1.6]\,\kmps$ in the global coordinate frame.
-The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}$.
+The upper limit on $\vk$ is chosen to match the escape velocity of the centre of the merger remnant, $v_\mathrm{esc}=1800\,\kmps$.
 We additionally run one simulation with a recoil kick above $v_\mathrm{esc}$, with $\vk=2000\,\kmps$, to test the effect of rapidly removing an SMBH.
@@
 For all simulations, snapshots are produced every $5\,\mathrm{Myr}$.
-Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, DM with a softening length of $\varepsilon = 300\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.
-Within the \ketju{} region, dynamical interactions between stellar particles and SMBHs are unsoftened.
+Interactions between stellar particles are softened with a softening length of $\varepsilon_\star = 2.5\,\mathrm{pc}$, DM is softened with a softening length of $\varepsilon_\mathrm{DM} = 300\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon_\star = 7.5\,\mathrm{pc}$.
+Within the \ketju{} region, dynamical interactions between stellar particles and SMBHs, and SMBHs-SMBHs are unsoftened.
 For the \gadget{} component of the simulation, the integration error tolerance is set to 0.002, whereas the Gragg-Bulirsch-Stoer (GBS) tolerance for \ketju{} is set to $10^{-8}$.
@@
 
-Before SMBH binary coalescence, the binary scours a stellar core with a radius of $\rbbin\sim0.58\,\mathrm{kpc}$ in the centre of the galaxy (see \autoref{ssec:projdens}).
+Before the SMBH binary coalescence, the binary scours a stellar core with a radius of $\rbbin\sim0.58\,\mathrm{kpc}$ in the centre of the galaxy (see \autoref{ssec:projdens}).
 In the interior of the core, the stellar one-dimensional velocity dispersion is $\sigma_{\star,0} \equiv \sigma_\star(r<\rbbin)\sim 270\,\kmps$.
@@
 \end{equation}
-If the velocity of the recoil kick is not enough to completely eject the SMBH from the galaxy ($\vk < v_\mathrm{esc}$), the SMBH will oscillate about the centre of the galaxy. 
+If the velocity of the recoil kick is not high enough to completely eject the SMBH from the galaxy ($\vk < v_\mathrm{esc}$), the SMBH will oscillate about the centre of the galaxy. 
 The amplitude of the oscillations, termed the apocentre distance $r_\mathrm{apo}$, decreases with each oscillation due to dynamical friction, which is automatically resolved with \ketju{} \citep[e.g.][]{karl2015,rantala2017,genina2024}.
@@
 If $r_\mathrm{apo}$ exceeds $\sim 10\,\mathrm{kpc}$, mild triaxiality in the outer regions of the galaxy merger remnant (shown in \autoref{fig:triax}) can torque the SMBH into an orbit with some degree of tangential motion, an example of which is given in \autoref{fig:trajectory} for the $\vk=780\,\kmps$ simulation.
-For the first few orbits, the SMBH may not necessarily pass through the stellar core, though it loses energy to the surrounding medium with each oscillation.
+For the first few orbits, the SMBH may thus not necessarily pass through the stellar core, though it loses energy to the surrounding medium with each oscillation.
 During the final passages however, the SMBH passes through the core, before its velocity decreases below the velocity dispersion of the core.
@@
 \end{equation}
-where $\tilde{\sigma}$ is the average velocity dispersion of the stellar and DM particles, and $\xi$ is the mass ratio to the SMBH, determined:
+where $\tilde{\sigma}$ is the average velocity dispersion of the stellar and DM particles, and $\xi$ is the mass ratio to the SMBH, calculated as:
 \begin{equation}
@@
 The settling times are shown in the lower panel of \autoref{fig:infall}.
-For recoil velocities $\vk \leq 240\,\kmps$ the kick is not enough to displace the SMBH beyond the binary-scoured core and the SMBH is considered settled in less than $100\,\mathrm{Myr}$ after SMBH binary coalescence.
+For recoil velocities $\vk \leq 240\,\kmps$ the kick is not large enough to displace the SMBH beyond the binary-scoured core and the SMBH is considered settled in less than $100\,\mathrm{Myr}$ after SMBH binary coalescence.
 As the kick velocity is increased further, the settling time begins to increase exponentially. 
@@
 The used limit for maximum displacement is a few times larger than the largest suggested offset of $19.4\,\mathrm{kpc}$ from observations \citep{barrows2016}.
-We do however analyse the $\vk=2000\,\kmps$ simulation, which is above the escape velocity of the galaxy, as a proxy for an SMBH that is instantaneously removed from the simulation.
+We do however analyse the $\vk=2000\,\kmps$ simulation, which is above the escape velocity of the galaxy, as a proxy for an SMBH that is almost instantaneously removed from the simulation.
 Importantly, an escaping SMBH never completes a single oscillation about the centre of the galaxy.
@@
     \includegraphics[width=0.4\textwidth]{figures/infall_times_and_apocenters.pdf}
-    \caption{ \textit{Top panel}: the distance of first apocentre passage from the centre of the galaxy for each kick velocity that resulted in a settled SMBH. 
+    \caption{ \textit{Top panel}: the distance of the first apocentre passage from the centre of the galaxy for each kick velocity that resulted in a settled SMBH. 
     The horizontal dashed line, $\rbbin$, marks the size of the stellar core the SMBH binary scours before coalescence, and the vertical dashed line, $\sigma_\star(r<\rbbin)$, is the one-dimensional velocity dispersion of the stellar particles interior to the binary-scoured core radius.
@@
 The core region of the galaxy constantly evolves while the SMBH moves through the galaxy.
-%Furthermore, it is important to ascertain that the distribution of stars does not drastically change after the times chosen as the focus of the analysis. 
 We inspect how the Lagrangian radius relative to the centre of mass and enclosing a stellar mass equal to the SMBH mass, $r(M_\star = M_\bullet)$, evolves with time in \autoref{fig:lagrangian} until the SMBH is settled.
@@
         A density cusp is apparent for all mergers where the SMBH settled at the centre.
-        At radii $r>3\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity has a larger impact on the inner mass distribution of the galaxy merger remnant.
+        At radii $r>2\,\mathrm{kpc}$ all the density profiles are consistent with each other, indicating that a changing SMBH recoil velocity has a larger impact on the inner mass distribution of the galaxy merger remnant.
         }
@@
 As a result, the central stellar density decreases as the number of oscillations increase. 
-The $\vk=2000\,\kmps$ simulation, in which the SMBH does not oscillate through the core, has a central stellar density that decreases only due to the initial response of the stellar particles to the suddenly-changed potential.
+The $\vk=2000\,\kmps$ simulation, in which the SMBH does not oscillate through the core, has a central stellar density that decreases only due to the initial response of the stellar particles to the sudden change of the potential.
 The simulations with a settled SMBH have a density profile within $r<1\,\mathrm{kpc}$ that follows a power law $\rho \propto r^{-\nu}$ with $0<\nu<1$.
@@
         For the models, we do not show the parameter uncertainty for visual clarity.
-        \textit{Bottom panel}: Mass deficit $M_\mathrm{def}$ between S\'ersic and core-S\'ersic fits as a function of the recoil velocity scaled to the mass deficit due to SMBH binary scouring ($M_{\mathrm{def},0}$, left axis) and physical units (right axis).
+        \textit{Bottom panel}: Box plots of the Monte Carlo sampled mass deficit $M_\mathrm{def}$ between S\'ersic and core-S\'ersic fits as a function of the recoil velocity. 
+        $M_\mathrm{def}$ is scaled to the mass deficit due to SMBH binary scouring ($M_{\mathrm{def},0}$, left axis) and shown in physical units (right axis). 
         }
@@
 Past studies have found the mass deficit to correlate with the combined mass of the SMBH binary and the number of major `dry' mergers \citep[e.g.][]{milosavljevic2001, graham2003, merritt2006,rantala2024}.
-%Past studies have found the mass deficit to be proportional to the mass of the binary approximately $M_{\mathrm{def}}\sim 0.5 -- 5 M_{BH}$ for a single merger. 
 In order to determine the mass displaced by the gravitational interactions and recoil, we adopt an observational approach by comparing the stellar mass density profiles of the inward extrapolated S\'ersic $\Sigma_{\mathrm{s}}(R)$ \citep{Sersic1963,Sersic1968} and the core-S\'ersic $\Sigma(R)$ fits  \citep[e.g.][]{dullo2014,bonfini2016, dullo2019, nasim2021b} as:
 \begin{equation}
-    M_{\mathrm{def}}=2\pi \Upsilon \int_0^\infty \left[ \Sigma_{\mathrm{s}}(R)-\Sigma(R)\right] R \dd{R},
-\end{equation}
-where the mass-to-light ratio is chosen to be $\Upsilon=3.5 \, \Msun / \Lsun$ \citep{bonfini2016}.
+    M_{\mathrm{def}}=2\pi \int_0^\infty \left[ \Sigma_{\mathrm{s}}(R)-\Sigma(R)\right] R \dd{R},
+\end{equation}
+assuming a constant mass-to-light ratio \citep{bonfini2016}. 
 The core-S\'ersic profile is calculated using \autoref{eq:cs} and \autoref{eq:cs_prefactor}, while the S\'ersic profile can be obtained using the same set of equations with a slight modification. 
@@
 \begin{equation}
-    \Sigma_{\mathrm{s}} = \Sigma ' \exp \left( -b_n (R/R_{\mathrm{e}})^{1/n} \right).
+    \Sigma_{\mathrm{s}} = \Sigma ' \exp \left( -b_n \left(\frac{R}{R_{\mathrm{e}}} \right)^{1/n} \right).
 \end{equation}
@@
 
-The mass deficits as a function of kick velocities $\vk\leq 1020\,\kmps$ are shown in the lower panel of \autoref{fig:vkrb}.
+The resulting mass deficits as a function of kick velocities $\vk\leq 1020\,\kmps$ are shown in the lower panel of \autoref{fig:vkrb}.
 At low kick velocities, $\vk\leq 240\,\kmps$, we see a slow monotonic increase of mass deficit as a function of the kick velocity.
 Although the scatter grows at higher kick velocities, an increase of the median trend can be observed over the whole kick velocity range.
-For high $\vk$ we see a median trend of $M_\mathrm{def} \sim 3 M_{\mathrm{def},0}$, where $M_{\mathrm{def},0}$ corresponds to the mass deficit from binary scouring alone.
-We find that the median mass deficit due to SMBH binary scouring prior to the coalescence of the SMBHs is $M_{\mathrm{def},0}=5.11\times10^9\,\Msun$, roughly $0.87M_\bullet$.
-As a consequence of the mass deficit growing to $M_\mathrm{def} > M_\bullet$ for non-zero recoil velocities, as well as higher recoil velocities resulting in larger $M_\mathrm{def}$, we confirm our earlier hypothesis in \autoref{ssec:movement} that stellar mass bound to the SMBH (at most $0.37M_\bullet$) is not the primary mechanism by which the core grows.
-
-Although \citet{merritt2006} and \citet{nasim2021b} found smaller mass deficit values from binary scouring alone for single mergers ($M_{\mathrm{def},0}\simeq 0.5\, M_\bullet$ and $M_{\mathrm{def},0}\simeq 0.26\,M_\bullet$, respectively), our value is in agreement with other previous studies.
-\citet{gualandris2008} reported a $M_{\mathrm{def},0}\sim M_\bullet$ relation, which is comparable to our value.
-In addition, similar mass deficits have also been found by \citet{rantala2024}.
-Slight discrepancies can be attributed to the differences in initial conditions and to varying approaches to computing mass deficits.
-Additionally, \citet{gualandris2008} use 3D density profiles, while we utilise projected profiles to make our results more comparable to observational data.
-
-%\citet{merritt2006} found
-%Although slightly smaller, our value is in good agreement with 26 per cent from the first generation equal mass merger of \citet{nasim2021b} (see merger `A1' in their Table 4).
-Slight differences can be attributed to the differences in initial conditions.
-
-%\citet{merritt2006} and \citet{gualandris2008} found larger mass deficit values from binary scouring alone for single mergers: $M_{\mathrm{def},0}\simeq 0.5\, M_\bullet$ and $M_{\mathrm{def},0}\sim M_\bullet$, respectively) \sonja{Should I add Rantala2024 $M_{\mathrm{def},0}\sim (0.75-2)\, M_\bullet$ to this list?
-Their approach to Mdef is very different.}.
-Although differing initial conditions already cause discrepancies of the missing mass values, varying approaches to computing mass deficits may also contribute to variations.
-Additionally, \citet{gualandris2008} use 3D density profiles, while we utilize projected profiles to make our results more comparable to observational data.
+For high $\vk$ we see a median trend of $M_\mathrm{def} \sim 4 M_{\mathrm{def},0}$, where $M_{\mathrm{def},0}$ corresponds to the mass deficit from binary scouring alone. 
+We find that the median mass deficit due to SMBH binary scouring prior to the coalescence of the SMBHs is $M_{\mathrm{def},0}=1.46\times10^9\,\Msun$, roughly $0.25M_\bullet$.
+As a consequence of the mass deficit growing to $M_\mathrm{def} \sim M_\bullet$ for non-zero recoil velocities, as well as $M_\mathrm{def}$ correlating positively with recoil velocity, we confirm our earlier hypothesis in \autoref{ssec:movement} that stellar mass bound to the SMBH (at most $0.37M_\bullet$) is not the primary mechanism by which the core grows.
+
+There are numerous ways to calculate the mass deficit in simulations, and differences between methods along with varying initial conditions may contribute to discrepancies across studies.
+A common approach to calculating the mass deficit, which we also adopt, is to compare the S\'ersic and the extrapolated core-S\'ersic profiles \citep{graham2003, Dosopoulou2021, nasim2021b}.
+This technique has been used by \citet{dullo2014} for observations of cored galaxies, resulting in typical mass deficits of $0.5 \lesssim M_{\mathrm{def}} /M_\bullet \lesssim 4$; these mass deficits may not be due solely to SMBH binary scouring however.
+Other methods, including calculating the difference in densities between the initial and final states of the galaxy merger \citep[e.g.][]{merritt2006, gualandris2008}, and running the simulations with and without SMBHs \citep[e.g.][]{partmann2024, rantala2024}, typically find greater mass deficits than we do, but of the same magnitude.
 
@@
 Thus, the increasingly large range of $\rb$ values introduce a wider range of mass deficits as a function of kick velocity.
-Additionally, we note that the choice of the value of the mass-to-light ratio impacts the resulting mass deficits.
-For example, by using $\Upsilon=4.5 \, \Msun / \Lsun$ (instead of $\Upsilon=3.5 \, \Msun / \Lsun$) the value of the binary scouring alone increases to $\sim 112$ per cent of the total SMBH binary mass.
 
@@
 To predict the distribution of expected stellar core sizes given our merger remnant of two massive elliptical galaxies, we require a mapping from the kick velocity to the core radius. 
-Let us define a scaled kick velocity $v'=\vk/v_\mathrm{esc}$, where $v_\mathrm{esc}=1800\,\kmps$.
+Let us define a scaled kick velocity $v'=\vk/v_\mathrm{esc}$.
 Previous work by \citet{nasim2021b} fit an empirical power law of the form:
@@
 Thus, whilst a non-equal mass SMBH binary merger results in a slightly different recoil velocity distribution, large recoil velocities are still allowed, and are frequent, within the expected SMBH mass variation.
-Herein we consider our fiducial case of equal mass SMBH binaries, but do not expect the conclusions to drastically differ in the more general case.
+Herein we consider our fiducial case of equal-mass SMBH binaries, but do not expect the conclusions to drastically differ in the more general case.
 
@@
 This behaviour is not captured with the exponential nor linear models.
-For example, in the case of the sigmoid model, a normalised recoil velocity of $v'=2 \rightarrow \rb=3.9$, whereas $v'=4 \rightarrow \rb=4.0$.
+For example, in the case of the sigmoid model, a scaled recoil velocity ($v'=v/v_\mathrm{esc}$) of $v'=2 \rightarrow \rb=3.9$, whereas $v'=4 \rightarrow \rb=4.0$.
 Conversely, for the exponential model, $v'=2 \rightarrow \rb=5.7$, whereas $v'=4 \rightarrow \rb=9.0$.
@@
 The zeroth order of the potential expansion $\Phi_{000}(\vb{x})$ is the model upon which we wish to construct our expansion, and a common choice for this is the \citet{hernquist1990} profile.
-Following \citet{frigo2021}, the expansion is limited to $n_\mathrm{max}=18$ and $l_\mathrm{max}=7$, as this sufficiently captures the potential of the galactic merger remnant.
+Following \citet{frigo2021}, the expansion is limited to $n_\mathrm{max}=18$ and $l_\mathrm{max}=7$, as this captures the potential of the galactic merger remnant with sufficient accuracy.
 
@@
 To overcome this, we centre the snapshot on the SMBH position, but exclude the SMBH from the SCF routine.
-This provides us with a potential field without the SMBH contribution; the potential from the SMBH is added as a point-mass potential following the SCF potential construction.
+This provides us with a potential field without the SMBH contribution; the potential from the SMBH is then added as a point-mass potential following the SCF potential construction.
 In the case where the SMBH has not settled at the centre (i.e.
@@
 Finally, increasing the SMBH recoil velocity decreases the fraction of rosette orbits at small radii $r<1\,\mathrm{kpc}$, with no rosette orbits for the $\vk=2000\,\kmps$ simulation.
-There is evidence for a minor decrease in outer $x$-tube orbits with increasing recoil velocity at radii, however this trend is not as dramatic as the three previously discussed.
+There is evidence for a minor decrease in outer $x$-tube orbits with increasing recoil velocity at radii $r<3\,\mathrm{kpc}$, however this trend is not as dramatic as the three previously discussed.
 In particular, the $\vk=2000\,\kmps$ simulation has almost no outer $x$-tube orbits until $r=2\,\mathrm{kpc}$, and only comes to dominate the orbit fraction at $r\sim10\,\mathrm{kpc}$.
@@
 The recoil velocity where the transition from a tube orbit-dominated core to a box orbit-dominated core occurs is around $\vk\sim240\,\kmps$, approximately the same recoil velocity that allows the kicked SMBH to exit the SMBH binary-scoured core (see \autoref{fig:infall}, upper panel).
-This suggests that as the kicked SMBH ploughs through the stellar core whilst it is settling but before its velocity falls below the velocity dispersion of the core, the rapid change in potential disrupts the conditions for angular momentum conserving orbits, altering the stellar particles to radial orbits instead.
+This suggests that as the kicked SMBH ploughs through the stellar core whilst it is settling but before its velocity falls below the velocity dispersion of the core, the rapid change in potential disrupts the conditions for angular momentum-conserving orbits, altering the stellar particles to radial orbits instead.
 For the highest kick velocities, $\vk > 900\,\kmps$, the amount of stellar mass in box orbits is almost twice that in tube orbits, as shown in \autoref{fig:orbitsbt}.
@@
 
-We next create mock integral field unit (IFU) observations using two different projections for the line of sight (LOS) velocity distribution: one where the LOS is along the SMBH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the SMBH kick axis (`orthogonal').
+We next create mock integral field unit (IFU) observations using two different projections for the line-of-sight (LOS) velocity distribution: one where the LOS is along the SMBH kick axis (`parallel'), and another where the LOS is along an axis orthogonal to the SMBH kick axis (`orthogonal').
 Testing these two special projections, we investigate the limiting cases of how the galaxy kinematics behave when all or none of the SMBH motion is directed along the LOS to the observer.
@@
 We create our mock IFU observations following \citet{naab2014}, assuming our merger remnants represent local elliptical galaxies that lie at $z=0.01$.
-Our observations encompass a square aperture of side length $0.25r_{1/2}$, where $r_{1/2}$ is the half mass radius. 
-For each particle within this aperture, we generate 25 pseudo-particles with identical LOS velocities as the original particle, but are spatially displaced from the original particle with a Gaussian distribution in the projected $x$ and $y$ direction with a standard deviation of $6.8''$ (corresponding to a physical scale\footnote{https://cosmocalc.icrar.org} of $0.3\,\mathrm{kpc}$) to mimic seeing effects.
+Our observations encompass a square aperture of side length $0.25r_{1/2}$, where $r_{1/2}$ is the 3D half mass radius. 
+For each particle within this aperture, we generate 25 pseudo-particles with identical LOS velocities as the original particle, but are spatially displaced from the original particle with a Gaussian distribution in the projected $x$ and $y$ direction with a standard deviation of $6.8''$ (corresponding to a physical scale\footnote{\url{https://cosmocalc.icrar.org} with $\Lambda\mathrm{CDM}$ cosmological parameters $H_0=70\,\kmps\,\mathrm{Mpc}^{-1}$, $\Omega_\mathrm{M}=0.3$, and $\Omega_\Lambda=0.7$.} of $0.3\,\mathrm{kpc}$ at $z=0.01$) to mimic seeing effects.
 The pseudo-particles are then binned onto a spatial grid, where the grid is assumed to have a resolution of $0.2''$, similar to the observations of \citet{neureiter2023}, which then corresponds to a physical resolution of $\sim0.04\,\mathrm{kpc}$.
@@
 where $F_i$ is the flux, $R_i$ is the radial displacement from the centre, $V_i$ is the mean velocity, and $\sigma_i$ is the velocity dispersion, each taken within the $\ordinal{i}$ Voronoi bin.
-We find that our galaxies have $\lambda_R<0.02$ within the IFU aperture, and thus we conclude very little rotation is present.
+We find that our galaxies have $\lambda_R<0.02$ within the IFU aperture, and thus we conclude that very little rotation is present.
 
@@
 This feature is also seen in the mock IFU images of the tube orbits.
-Additionally, creating mock IFU maps for two other projections (one with the line of sight aligned with the major axis of the inertia tensor, and the other with the line of sight aligned with the vector normal to the merger plane), this feature is not present.
+Additionally, creating mock IFU maps for two other projections (one with the line-of-sight aligned with the major axis of the inertia tensor, and the other with the line-of-sight aligned with the vector normal to the merger plane), this feature is not present.
 We thus conclude that immediately following the merger of an SMBH binary, complex structure in the LOSVD may be apparent for tube orbits.
@@
 A stellar system with purely radial orbits ($\sigma_{\theta}=\sigma_{\phi}=0$) corresponds to $\beta = 1$, while purely circular orbits ($\sigma_{r}=0$) would yield $\beta = -\infty$.
-An ergodic system is one where $\sigma_{\theta}=\sigma_{\phi}=\sigma_r \iff \beta=0$, which thus describes a system where there is a balance between box and tube orbits.
+An ergodic system is one where $\sigma_{\theta}=\sigma_{\phi}=\sigma_r \iff \beta=0$, and hence corresponds to a balance between box and tube orbits.
 
@@
 In agreement with previous work \citep{boylan2004,merritt2004}, we find that an SMBH displaced due to gravitational recoil decreases the stellar density in the central regions of the merger remnant as it settles to an equilibrium state.
-Additionally, we find that this oscillatory motion, by the time the SMBH has settled, increases the stellar density core beyond that induced by the scouring of the SMBH binary prior to coalescence, confirming the results of \citet{gualandris2008}, \citet{nasim2021b}, and \citet{khonji2024}.
+Additionally, we find that this oscillatory motion, by the time the SMBH has settled, increases the size of the stellar density core beyond that induced by the scouring of the SMBH binary prior to coalescence, confirming the results of \citet{gualandris2008}, \citet{nasim2021b}, and \citet{khonji2024}.
 We assess the magnitude of this scouring with time using the Lagrangian radii presented in \autoref{fig:lagrangian}.
@@
 When the SMBH does pass through the core region in the last orbits when the SMBH velocity $v_\bullet$ is only marginally greater than the stellar velocity dispersion within the core, the Lagrangian radius rapidly increases further.
-\citet{gualandris2008} found that each oscillation the SMBH enlarges the core, whereas \citet{nasim2021b} found that only the initial oscillation due to the recoil kick affects the core. 
+\citet{gualandris2008} found that each oscillation of the SMBH enlarges the core, whereas \citet{nasim2021b} found that only the initial oscillation due to the recoil kick affects the core. 
 Our result suggest that differences in the torquing due to the specific distribution of stellar mass could alleviate some of the tension between the two studies.
@@
 Conversely, a mild density cusp $\rho \propto r^{-\nu}$ with $0 < \nu < 1$ is present in simulations where the SMBH settled, $\vk<1020\,\kmps$, with the exponent $\nu$ decreasing with increasing $\vk$, in agreement with previous work \citep[e.g.][]{nasim2021b,khonji2024}.
-All simulations with $\vk<1020\,\kmps$ display a core in projection however, in alignment with previous observational and theoretical work \citep[e.g.][]{ferrarese1994,nakano1999}.
+All simulations with $\vk<1020\,\kmps$ display a core in projection however, in agreement with previous observational and theoretical work \citep[e.g.][]{ferrarese1994,nakano1999}.
 
@@
 Instead, we observe $\pi$-box orbits to dominate at all radii for the $\vk=2000\,\kmps$ simulation, in tension with the understanding of box orbits being the dominant orbit family to interact with the SMBH binary prior to coalescence.
-However, we find that the excess of $\pi$-box orbits in the $\vk=2000\,\kmps$ simulation primarily originate from angular momentum conserving orbits (63\%) which have reacted to a sudden change in potential due to the escaping SMBH.
+However, we find that the excess of $\pi$-box orbits in the $\vk=2000\,\kmps$ simulation primarily originate from angular momentum-conserving orbits (63 per cent) which have reacted to a sudden change in the potential due to the escaping SMBH.
 
@@
 In \autoref{fig:core-obs} we plot, as a function of the remnant SMBH mass, the median core radii for each of our merger remnants and compare to the observational data presented in \citet{thomas2016}.
-%What we can do though is assess how our simulated core sizes compare to observed core sizes.
 The inferred median core sizes range from $\rb\sim0.5\,\mathrm{kpc}$ for the $\vk=0\,\kmps$ simulation, to $\rb\sim1.4\,\mathrm{kpc}$ for the $\vk=1020\,\kmps$ simulation.
@@
 
-Considering that for our initial conditions, the peak in the transformation-sampled core radius distribution is in the range of $1.0\,\mathrm{kpc}$ -- $1.5\,\mathrm{kpc}$ (\autoref{ssec:predcore}, depending on the assumed relation between recoil velocity and core radius), cores in the surface density profile in excess of $1.0\,\mathrm{kpc}$ are viable candidates for GW recoil induced core formation.
+Considering that for our initial conditions, the peak in the transformation-sampled core radius distribution is in the range of $1.0\,\mathrm{kpc}$ - $1.5\,\mathrm{kpc}$ (\autoref{ssec:predcore}, depending on the assumed relation between recoil velocity and core radius), cores in the surface density profile in excess of $1.0\,\mathrm{kpc}$ are viable candidates for GW recoil induced core formation.
 Certainly galaxies hosting cores much greater than $2.0\,\mathrm{kpc}$, as for example in the central galaxy A2261-BCG \citep{bonfini2016, dullo2019}, offer prime targets for SMBH recoil kicks to have been a contributing mechanism to the core formation.
@@
 Assuming that a major merger has taken place to produce the observed galaxy, the GW recoil velocity can be inferred by inverting the relation between core radius and recoil velocity, as in our \autoref{eq:vkrb_sig}.
-Presumably this relation between $\vk$ and $\rb$ would depend on the mass ratio of the progenitor SMBH binary, so we caution that the relations we provide are only applicable to a major merger involving two equal mass SMBHs of $M_{\bullet,0}\sim 10^9\,\Msun$.
+Presumably this relation between $\vk$ and $\rb$ would depend on the mass ratio of the progenitor SMBH binary, so we caution that the relations we provide are only applicable to a major merger involving two equal-mass SMBHs of $M_{\bullet,0}\sim 10^9\,\Msun$.
 However equivalent relations may be found through a targeted parameter scan of SMBH recoil velocities, following our methods presented in \autoref{sec:num_sims}.
@@
 Similar results have also been found recently by \citet{steinle2023} and \citet{bourne2024}.
-In the case of gas-poor mergers, SMBH binary spins likely remain isotropic until entering the GW-dominated hardening phase \citep[e.g.][]{spadaro2024}, with alignment later occurring through general relativistic effects, particularly in the case of unequal mass SMBHs where the more massive SMBH spin is closely aligned with the orbital angular momentum of the binary \citep{berti2012}.
+In the case of gas-poor mergers, SMBH binary spins likely remain isotropic until entering the GW-dominated hardening phase \citep[e.g.][]{spadaro2024}, with alignment later occurring through general relativistic effects, particularly in the case of unequal-mass SMBHs where the more massive SMBH spin is closely aligned with the orbital angular momentum of the binary \citep{berti2012}.
 However, \citet{sayeb2021} found that whilst general relativistic effects can cause precession of the SMBH binary spin vectors, the overall distribution of resulting recoil velocities remains relatively unaffected, owing to the dominant mechanism for SMBH binary spin alignment being the Bardeen-Peterson effect.
@@
     \item A core is not present in the three-dimensional stellar density profile unless the SMBH has been ejected entirely from the galaxy, corresponding to the case of zero oscillations of the kicked SMBH about the galaxy centre.
-    \item An increase in the recoil velocity induces a larger core in the projected stellar mass density of the merger remnant than low recoil velocities, with our models indicating that core radii exceeding $1\,\mathrm{kpc}$ are not only possible, but likely.
+    \item An higher recoil velocity induces a larger core in the projected stellar mass density of the merger remnant than lower recoil velocities, with our models indicating that core radii exceeding $1\,\mathrm{kpc}$ are not only possible, but likely.
     As a consequence, the mass deficit due to SMBH recoil increases with larger recoil velocity, up to $\sim90$ per cent of the coalesced SMBH mass.
-    \item For merger remnants with $\vk \lesssim \sigma_{\star,0}$, the majority of stellar mass within the core has angular momentum conserving orbits. 
+    \item For merger remnants with $\vk \lesssim \sigma_{\star,0}$, the majority of stellar mass within the core has angular momentum-conserving orbits. 
     For recoil velocities above this threshold, box orbits which do not conserve angular momentum dominate the mass budget.
@@
 We thank Jens Thomas for useful discussions on IFU observations.
-\alex{\textbf{Everyone: please add your acknowledgements here}}
 A.R. acknowledges the support by the University of Helsinki Research Foundation.
-M.M., P.H.J., and D.I. acknowledge the support by the European Research Council via ERC Consolidator Grant KETJU (no. 818930).
+A.R., A.K., M.M., R.J.W., and P.H.J. acknowledge the support by the European Research Council via ERC Consolidator Grant KETJU (no. 818930).
+M.M. and P.H.J acknowledge the support of the Academy of Finland grant 339127.
+R.J.W. acknowledges the support of the Forrest Research Foundation.
 N.K. acknowledges support of the Research Council of Finland Flagship programme: Finnish Center for Artificial Intelligence and the Finnish Foundation for Technology Promotion.
 S.L. acknowledges the supports by the National Natural Science Foundation of China (NSFC) grant (no. 12473015, 11988101) and the K. C. Wong Education Foundation.
-P.H.J acknowledges the support of the Academy of Finland grant 339127.
-TN acknowledges support from the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany’s Excellence Strategy - EXC-2094 - 390783311 from the DFG Cluster of Excellence ``ORIGINS''.
+T.N.
+acknowledges support from the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany’s Excellence Strategy - EXC-2094 - 390783311 from the DFG Cluster of Excellence ``ORIGINS''.
 
@@
 \textbf{AR}: writing: review.
-\textbf{PHJ}: writing: original, supervision, funding acquisition.
+\textbf{PHJ}: supervision, writing: original, funding acquisition.
 \textbf{TN}: conceptualisation, writing: review.
@@
 
-We take all stellar particles within a fraction of the virial radius $R_\mathrm{vir}$, defined as the radius at which the mean density is equal to 200 times greater than the critical density; the virial radius is $\sim450\,\mathrm{kpc}$ for the merger remnants in this study.
+We take all stellar particles within a fraction of the virial radius $r_{200}$, defined as the radius at which the mean density is equal to 200 times greater than the critical density; the virial radius is $\sim450\,\mathrm{kpc}$ for the merger remnants in this study.
 
@@
 and $c\leq b\leq a$. 
-The tensor $\tilde{I}_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} R_\mathrm{vir}$ to $10^{-1} R_\mathrm{vir}$: the binning of particles is not cumulative.
+The tensor $\tilde{I}_\mathrm{red}$ is determined by binning the stellar component of the remnant into 20 radial shells from $10^{-3} r_{200}$ to $10^{-1} r_{200}$: the binning of particles is not cumulative.
 We apply a uniform filter to reduce Poisson noise due to low particle numbers at small radii.
@@
 \end{equation}
-following \citet[][]{binney2008}.
+following \citet{franx1991}.
 Values of $T$ approaching 1 indicate a prolate system, and values of $T$ approaching 0 indicate an oblate system.
@@
 In no pairing of parameters is a strong covariance observed; the marginal distributions are orthogonal to each other.
-One case worth a special mention is the width of the 25\% HDI for the parameter $\alpha$.
+One case worth a special mention is the width of the 25 per cent HDI for the parameter $\alpha$.
 The wide HDI indicates that a large range of $\alpha$ values maximise the posterior distribution.
@@
         Corner plot of the latent parameters in the core-S\'ersic model for a representative example, here the $\vk=600\,\kmps$ instance.
-        Contours indicate, from dark to light blue, the 25\%, 50\%, 75\%, and 99\% highest density intervals (HDIs).
+        Contours indicate, from dark to light blue, the 25 per cent, 50 per cent, 75 per cent, and 99 per cent highest density intervals (HDIs).
         All distributions are unimodal with little cross-correlation between the latent variables.

### 2024-10-23 baggins/docs/_build/latex/bagginsbayesiananalysisofgalaxy-galaxyinteractionsinsimulations.tex
Move baggins subdirectory to top level

%% Generated by Sphinx.
\def\sphinxdocclass{report}
\documentclass[letterpaper,10pt,english]{sphinxmanual}
\ifdefined\pdfpxdimen
   \let\sphinxpxdimen\pdfpxdimen\else\newdimen\sphinxpxdimen
\fi \sphinxpxdimen=.75bp\relax
\ifdefined\pdfimageresolution
    \pdfimageresolution= \numexpr \dimexpr1in\relax/\sphinxpxdimen\relax
\fi
%% let collapsible pdf bookmarks panel have high depth per default
\PassOptionsToPackage{bookmarksdepth=5}{hyperref}

\PassOptionsToPackage{booktabs}{sphinx}
\PassOptionsToPackage{colorrows}{sphinx}

\PassOptionsToPackage{warn}{textcomp}
\usepackage[utf8]{inputenc}
\ifdefined\DeclareUnicodeCharacter
% support both utf8 and utf8x syntaxes
  \ifdefined\DeclareUnicodeCharacterAsOptional
    \def\sphinxDUC#1{\DeclareUnicodeCharacter{"#1}}
  \else
    \let\sphinxDUC\DeclareUnicodeCharacter
  \fi
  \sphinxDUC{00A0}{\nobreakspace}
  \sphinxDUC{2500}{\sphinxunichar{2500}}
  \sphinxDUC{2502}{\sphinxunichar{2502}}
  \sphinxDUC{2514}{\sphinxunichar{2514}}
  \sphinxDUC{251C}{\sphinxunichar{251C}}
  \sphinxDUC{2572}{\textbackslash}
\fi
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amstext}
\usepackage{babel}

\usepackage{tgtermes}
\usepackage{tgheros}
\renewcommand{\ttdefault}{txtt}

\usepackage[Bjarne]{fncychap}
\usepackage{sphinx}

\fvset{fontsize=auto}
\usepackage{geometry}

% Include hyperref last.
\usepackage{hyperref}
% Fix anchor placement for figures with captions.
\usepackage{hypcap}% it must be loaded after hyperref.
% Set up styles of URL: it should be placed after hyperref.
\urlstyle{same}

\addto\captionsenglish{\renewcommand{\contentsname}{Contents:}}

\usepackage{sphinxmessages}
\setcounter{tocdepth}{3}
\setcounter{secnumdepth}{3}

\title{BAGGInS: Bayesian Analysis of Galaxy\sphinxhyphen{}Galaxy Interactions in Simulations}
\date{Oct 21, 2024}
\release{1.0}
\author{Alex Rawlings}
\newcommand{\sphinxlogo}{\vbox{}}
\renewcommand{\releasename}{Release}
\makeindex
\begin{document}

\ifdefined\shorthandoff
  \ifnum\catcode`\=\string=\active\shorthandoff{=}\fi
  \ifnum\catcode`\"=\active\shorthandoff{"}\fi
\fi

\pagestyle{empty}
\sphinxmaketitle
\pagestyle{plain}
\sphinxtableofcontents
\pagestyle{normal}
\phantomsection\label{\detokenize{index::doc}}

\sphinxstepscope

\chapter{Python API}
\label{\detokenize{references/API:python-api}}\label{\detokenize{references/API:baggins-api}}\label{\detokenize{references/API::doc}}

\section{BAGGInS}
\label{\detokenize{references/API:baggins}}
\sphinxAtStartPar
Here you can find a list of available functions and classes in the BAGGInS package.

\sphinxstepscope

\subsection{\_backend package}
\label{\detokenize{references/_backend:backend-package}}\label{\detokenize{references/_backend::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/_backend:submodules}}

\subsubsection{\_backend.Logging module}
\label{\detokenize{references/_backend:module-_backend.Logging}}\label{\detokenize{references/_backend:backend-logging-module}}\index{module@\spxentry{module}!\_backend.Logging@\spxentry{\_backend.Logging}}\index{\_backend.Logging@\spxentry{\_backend.Logging}!module@\spxentry{module}}\index{setup\_logger() (in module \_backend.Logging)@\spxentry{setup\_logger()}\spxextra{in module \_backend.Logging}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.Logging.setup_logger}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{\_backend.Logging.}}\sphinxbfcode{\sphinxupquote{setup\_logger}}}
{\sphinxparam{\DUrole{n}{name}}\sphinxparamcomma \sphinxparam{\DUrole{n}{console\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{logfile}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{file\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{capture\_warnings}\DUrole{o}{=}\DUrole{default_value}{True}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Set up a logger instance
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{name}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}) \textendash{} name of logger

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{console\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for console printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{logfile}} (\sphinxstyleliteralemphasis{\sphinxupquote{path\sphinxhyphen{}like}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} file to print logs to, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{file\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for file printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{capture\_warnings}} (\sphinxstyleliteralemphasis{\sphinxupquote{bool}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} redirect all warnings to the logger, by default True

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{CustomLogger} \textendash{} logger

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{\_backend.States module}
\label{\detokenize{references/_backend:module-_backend.States}}\label{\detokenize{references/_backend:backend-states-module}}\index{module@\spxentry{module}!\_backend.States@\spxentry{\_backend.States}}\index{\_backend.States@\spxentry{\_backend.States}!module@\spxentry{module}}\index{PublishingState (class in \_backend.States)@\spxentry{PublishingState}\spxextra{class in \_backend.States}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.PublishingState}}
\pysigstartsignatures
\pysigline
{\sphinxbfcode{\sphinxupquote{class\DUrole{w}{ }}}\sphinxcode{\sphinxupquote{\_backend.States.}}\sphinxbfcode{\sphinxupquote{PublishingState}}}
\pysigstopsignatures
\sphinxAtStartPar
Bases: \sphinxcode{\sphinxupquote{\_State}}
\index{\_\_init\_\_() (\_backend.States.PublishingState method)@\spxentry{\_\_init\_\_()}\spxextra{\_backend.States.PublishingState method}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.PublishingState.__init__}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{\_\_init\_\_}}}
{}
{}
\pysigstopsignatures
\sphinxAtStartPar
Simple object to store the state of a plotting\sphinxhyphen{}style setting

\end{fulllineitems}

\end{fulllineitems}

\index{TmpDirRegister (class in \_backend.States)@\spxentry{TmpDirRegister}\spxextra{class in \_backend.States}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{class\DUrole{w}{ }}}\sphinxcode{\sphinxupquote{\_backend.States.}}\sphinxbfcode{\sphinxupquote{TmpDirRegister}}}
{\sphinxparam{\DUrole{n}{basename}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Bases: \sphinxcode{\sphinxupquote{object}}
\index{\_\_init\_\_() (\_backend.States.TmpDirRegister method)@\spxentry{\_\_init\_\_()}\spxextra{\_backend.States.TmpDirRegister method}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister.__init__}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{\_\_init\_\_}}}
{\sphinxparam{\DUrole{n}{basename}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Class to hold a register of temporary directories. This means that
temporary directories are particular to the given invocation of
baggins, and can be cleaned up at exit without affecting the
temporary directories of other invocations running concurrently.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{basename}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{path\sphinxhyphen{}like}}) \textendash{} base temporary directory path to which a timestamp will be appended

\end{description}\end{quote}

\end{fulllineitems}

\index{make\_new\_dir() (\_backend.States.TmpDirRegister method)@\spxentry{make\_new\_dir()}\spxextra{\_backend.States.TmpDirRegister method}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister.make_new_dir}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxbfcode{\sphinxupquote{make\_new\_dir}}}
{}
{}
\pysigstopsignatures
\sphinxAtStartPar
Make a new temporary directory, and add it to the register

\end{fulllineitems}

\index{register (\_backend.States.TmpDirRegister property)@\spxentry{register}\spxextra{\_backend.States.TmpDirRegister property}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.States.TmpDirRegister.register}}
\pysigstartsignatures
\pysigline
{\sphinxbfcode{\sphinxupquote{property\DUrole{w}{ }}}\sphinxbfcode{\sphinxupquote{register}}}
\pysigstopsignatures
\end{fulllineitems}

\end{fulllineitems}

\subsubsection{Module contents}
\label{\detokenize{references/_backend:module-_backend}}\label{\detokenize{references/_backend:module-contents}}\index{module@\spxentry{module}!\_backend@\spxentry{\_backend}}\index{\_backend@\spxentry{\_backend}!module@\spxentry{module}}\index{setup\_logger() (in module \_backend)@\spxentry{setup\_logger()}\spxextra{in module \_backend}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/_backend:backend.setup_logger}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{\_backend.}}\sphinxbfcode{\sphinxupquote{setup\_logger}}}
{\sphinxparam{\DUrole{n}{name}}\sphinxparamcomma \sphinxparam{\DUrole{n}{console\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{logfile}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{file\_level}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}WARNING\textquotesingle{}}}\sphinxparamcomma \sphinxparam{\DUrole{n}{capture\_warnings}\DUrole{o}{=}\DUrole{default_value}{True}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Set up a logger instance
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{name}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}) \textendash{} name of logger

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{console\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for console printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{logfile}} (\sphinxstyleliteralemphasis{\sphinxupquote{path\sphinxhyphen{}like}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} file to print logs to, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{file\_level}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} logging level for file printing, by default “WARNING”

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{capture\_warnings}} (\sphinxstyleliteralemphasis{\sphinxupquote{bool}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} redirect all warnings to the logger, by default True

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{CustomLogger} \textendash{} logger

\end{description}\end{quote}

\end{fulllineitems}

\sphinxstepscope

\subsection{analysis package}
\label{\detokenize{references/analysis:analysis-package}}\label{\detokenize{references/analysis::doc}}

\subsubsection{Subpackages}
\label{\detokenize{references/analysis:subpackages}}
\sphinxstepscope

\paragraph{analysis.analysis\_classes package}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-package}}\label{\detokenize{references/analysis.analysis_classes::doc}}

\subparagraph{Submodules}
\label{\detokenize{references/analysis.analysis_classes:submodules}}

\subparagraph{analysis.analysis\_classes.BHBinary module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-bhbinary-module}}

\subparagraph{analysis.analysis\_classes.BHBinaryData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-bhbinarydata-module}}

\subparagraph{analysis.analysis\_classes.Brownian module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-brownian-module}}

\subparagraph{analysis.analysis\_classes.BrownianData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-browniandata-module}}

\subparagraph{analysis.analysis\_classes.ChildSim module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-childsim-module}}

\subparagraph{analysis.analysis\_classes.ChildSimData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-childsimdata-module}}

\subparagraph{analysis.analysis\_classes.CoreKick module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-corekick-module}}

\subparagraph{analysis.analysis\_classes.GaussianProcesses module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-gaussianprocesses-module}}

\subparagraph{analysis.analysis\_classes.GrahamModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-grahammodels-module}}

\subparagraph{analysis.analysis\_classes.HDF5Base module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hdf5base-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesBinary module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiesbinary-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesBinaryData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiesbinarydata-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesSingle module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiessingle-module}}

\subparagraph{analysis.analysis\_classes.HMQuantitiesSingleData module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-hmquantitiessingledata-module}}

\subparagraph{analysis.analysis\_classes.KeplerModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-keplermodels-module}}

\subparagraph{analysis.analysis\_classes.PQModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-pqmodels-module}}

\subparagraph{analysis.analysis\_classes.QuinlanModels module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-quinlanmodels-module}}

\subparagraph{analysis.analysis\_classes.StanModel module}
\label{\detokenize{references/analysis.analysis_classes:analysis-analysis-classes-stanmodel-module}}

\subparagraph{Module contents}
\label{\detokenize{references/analysis.analysis_classes:module-contents}}

\subsubsection{Submodules}
\label{\detokenize{references/analysis:submodules}}

\subsubsection{analysis.analyse\_ketju module}
\label{\detokenize{references/analysis:analysis-analyse-ketju-module}}

\subsubsection{analysis.analyse\_snap module}
\label{\detokenize{references/analysis:analysis-analyse-snap-module}}

\subsubsection{analysis.general module}
\label{\detokenize{references/analysis:analysis-general-module}}

\subsubsection{analysis.masks module}
\label{\detokenize{references/analysis:module-analysis.masks}}\label{\detokenize{references/analysis:analysis-masks-module}}\index{module@\spxentry{module}!analysis.masks@\spxentry{analysis.masks}}\index{analysis.masks@\spxentry{analysis.masks}!module@\spxentry{module}}\index{get\_all\_id\_masks() (in module analysis.masks)@\spxentry{get\_all\_id\_masks()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_all_id_masks}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_all\_id\_masks}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}stars\textquotesingle{}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Return a dict of masks that mask the chosen particles by ID number to a
specific galaxy, assuming each galaxy has a single SMBH in it.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type we want to mask, by default ‘stars’

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstylestrong{masks} (\sphinxstyleemphasis{dict})

\item {} 
\sphinxAtStartPar
\sphinxstyleemphasis{keys corresponding to the host galaxy SMBH particle ID, values to pygad.}

\item {} 
\sphinxAtStartPar
\sphinxstyleemphasis{snapshot.masks.IDMask objects}

\end{itemize}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_all\_radial\_masks() (in module analysis.masks)@\spxentry{get\_all\_radial\_masks()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_all_radial_masks}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_all\_radial\_masks}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{radius}}\sphinxparamcomma \sphinxparam{\DUrole{n}{centre}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{id\_masks}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}stars\textquotesingle{}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Return a dict of masks that mask the chosen particles by radial distance to
the desired centre. Either a ball mask or a shell mask may be constructed.
Additionally, the option to initially filter particles by ID to constrain
the mask to those particles belonging to a given galaxy is possible.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{radius}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{tuple}}) \textendash{} radius to constrain the particles to \sphinxhyphen{} can be either a number to
construct a ball mask, or a tuple of (inner\_radius, outer\_radius)
to construct a shell mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{centre}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.UnitArr}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} centre from which the radial measurements should be made, by default
None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{id\_mask}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} ID masks to constrain particles to a given galaxy, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type to construct the mask for, by default None (all particle
used)

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstylestrong{masks} (\sphinxstyleemphasis{dict}) \textendash{} pygad mask objects, with keys corresponding to the host galaxy SMBH
particle ID number

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{ValueError}} \textendash{} centre method is invalid

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_binding\_energy\_mask() (in module analysis.masks)@\spxentry{get\_binding\_energy\_mask()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_binding_energy_mask}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_binding\_energy\_mask}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{energy}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{id\_mask}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{None}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Mask particles into bins of binding energy. Assumes that we are in centre of mass coordinates (including velocity): no centring is done.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{energy}} (\sphinxstyleliteralemphasis{\sphinxupquote{array\sphinxhyphen{}like}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{list}}\sphinxstyleliteralemphasis{\sphinxupquote{ of }}\sphinxstyleliteralemphasis{\sphinxupquote{tuples}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} energy bins to mask particles into \sphinxhyphen{}  can either be a regular array, in
which a “ball” mask is created, or a list of tuples, in which a “shell”
mask is created. By default (None) creates a “shell” type mask from the
5\% to 95\% quantile of binding energy in 20 bins evenly spaced in
logscale

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{id\_mask}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.snapshot.masks.IDMask}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} ID mask to constrain particles to a given galaxy, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{\_type\_}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type we want to mask, by default None (all)

\end{itemize}

\sphinxlineitem{Yields}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstylestrong{mask} (\sphinxstyleemphasis{pygad.snapshots.mask.IDMask}) \textendash{} mask for that energy interval

\item {} 
\sphinxAtStartPar
\sphinxstylestrong{energy} (\sphinxstyleemphasis{array\sphinxhyphen{}like, list of tuples}) \textendash{} energy sequence

\item {} 
\sphinxAtStartPar
\sphinxstyleemphasis{str} \textendash{} energy units

\end{itemize}

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{ValueError}} \textendash{} invalid energy input

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_id\_mask() (in module analysis.masks)@\spxentry{get\_id\_mask()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_id_mask}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_id\_mask}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{bhid}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{\textquotesingle{}stars\textquotesingle{}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Obtain a mask that allows filtering of particular particles. Particles are
first filtered to the host galaxy depending on their particle ID number,
where it is necessarily assumed that the particle ID ordering follows:
\begin{quote}

\sphinxAtStartPar
galaxy 1 IDs \textendash{}\textgreater{} galaxy 2 IDs
\end{quote}

\sphinxAtStartPar
To obtain a mask of all particles in a system, the routine can be run
for each particle family, and the masks combined as e.g.:
\begin{quote}

\sphinxAtStartPar
all\_id\_mask = star\_id\_mask \& dm\_id\_mask \& bh\_id\_mask
\end{quote}

\sphinxAtStartPar
\#TODO should this be a separate method??
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{bhid}} (\sphinxstyleliteralemphasis{\sphinxupquote{int}}) \textendash{} SMBH id number we want to find particles around

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type we want to mask, by default ‘stars’

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{pygad.snapshot.masks.IDMask} \textendash{} object to mask future snapshots of the same simulation

\sphinxlineitem{Raises}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{RuntimeError}} \textendash{} ID ordering is not divided by galaxy

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{RuntimeError}} \textendash{} only galaxy in the system

\end{itemize}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_radial\_mask() (in module analysis.masks)@\spxentry{get\_radial\_mask()}\spxextra{in module analysis.masks}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/analysis:analysis.masks.get_radial_mask}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{analysis.masks.}}\sphinxbfcode{\sphinxupquote{get\_radial\_mask}}}
{\sphinxparam{\DUrole{n}{snap}}\sphinxparamcomma \sphinxparam{\DUrole{n}{radius}}\sphinxparamcomma \sphinxparam{\DUrole{n}{centre}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{id\_mask}\DUrole{o}{=}\DUrole{default_value}{None}}\sphinxparamcomma \sphinxparam{\DUrole{n}{family}\DUrole{o}{=}\DUrole{default_value}{None}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Create a radial\sphinxhyphen{}based mask, can be either a ball or a shell. The default
function arguments are designed for compatability with get\_all\_radial\_masks.
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{snap}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.Snapshot}}) \textendash{} snapshot to mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{radius}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{tuple}}) \textendash{} radius to constrain the particles to \sphinxhyphen{} can be either a number to
construct a ball mask, or a tuple of (inner\_radius, outer\_radius)
to construct a shell mask

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{centre}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.UnitArr}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} centre from which the radial measurements should be made, by default
None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{id\_mask}} (\sphinxstyleliteralemphasis{\sphinxupquote{pygad.snapshot.masks.IDMask}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} ID mask to constrain particles to a given galaxy, by default None

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{family}} (\sphinxstyleliteralemphasis{\sphinxupquote{str}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} particle type to construct the mask for, by default None (all particle
used)

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstylestrong{mask} (\sphinxstyleemphasis{pygad.snapshot.masks.BallMask}) \textendash{} radial mask of particles

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{ValueError}} \textendash{} radius input is invalid

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{analysis.orbits module}
\label{\detokenize{references/analysis:analysis-orbits-module}}

\subsubsection{analysis.voronoi module}
\label{\detokenize{references/analysis:analysis-voronoi-module}}

\subsubsection{Module contents}
\label{\detokenize{references/analysis:module-contents}}
\sphinxstepscope

\subsection{cosmology package}
\label{\detokenize{references/cosmology:cosmology-package}}\label{\detokenize{references/cosmology::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/cosmology:submodules}}

\subsubsection{cosmology.conversions module}
\label{\detokenize{references/cosmology:module-cosmology.conversions}}\label{\detokenize{references/cosmology:cosmology-conversions-module}}\index{module@\spxentry{module}!cosmology.conversions@\spxentry{cosmology.conversions}}\index{cosmology.conversions@\spxentry{cosmology.conversions}!module@\spxentry{module}}\index{time2z() (in module cosmology.conversions)@\spxentry{time2z()}\spxextra{in module cosmology.conversions}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.conversions.time2z}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.conversions.}}\sphinxbfcode{\sphinxupquote{time2z}}}
{\sphinxparam{\DUrole{n}{t}}\sphinxparamcomma \sphinxparam{\DUrole{n}{H0}\DUrole{o}{=}\DUrole{default_value}{67.36}}\sphinxparamcomma \sphinxparam{\DUrole{n}{pres}\DUrole{o}{=}\DUrole{default_value}{False}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Estimate the redshift for a corresponding cosmic time
from Carmeli 2008
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{t}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} cosmic time since big bang {[}Gyr{]} \sphinxhyphen{}\textgreater{} pres = False
time before present {[}Gyr{]} \sphinxhyphen{}\textgreater{} pres = True

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{H0}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} Hubble constant in km/s/Mpc, by default 100*cosmology{[}‘h’{]}

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{pres}} (\sphinxstyleliteralemphasis{\sphinxupquote{bool}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} t is given from today not Big Bang?, by default False

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} redshift

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{cosmology.cosmology module}
\label{\detokenize{references/cosmology:module-cosmology.cosmology}}\label{\detokenize{references/cosmology:cosmology-cosmology-module}}\index{module@\spxentry{module}!cosmology.cosmology@\spxentry{cosmology.cosmology}}\index{cosmology.cosmology@\spxentry{cosmology.cosmology}!module@\spxentry{module}}\index{angular\_diameter\_distance() (in module cosmology.cosmology)@\spxentry{angular\_diameter\_distance()}\spxextra{in module cosmology.cosmology}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.cosmology.angular_diameter_distance}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.cosmology.}}\sphinxbfcode{\sphinxupquote{angular\_diameter\_distance}}}
{\sphinxparam{\DUrole{n}{z}}\sphinxparamcomma \sphinxparam{\DUrole{n}{cosmology}\DUrole{o}{=}\DUrole{default_value}{\{\textquotesingle{}h\textquotesingle{}: 0.6736, \textquotesingle{}omega\_L\textquotesingle{}: 0.6847, \textquotesingle{}omega\_M\textquotesingle{}: 0.3153, \textquotesingle{}zeq\textquotesingle{}: 3402\}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Determine the angular diameter distance for a flat universe
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{z}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} redshift

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{cosmology}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} cosmological parameters, by default cosmology

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} angular diameter distance {[}kpc{]}

\end{description}\end{quote}

\end{fulllineitems}

\index{get\_a0r() (in module cosmology.cosmology)@\spxentry{get\_a0r()}\spxextra{in module cosmology.cosmology}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.cosmology.get_a0r}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.cosmology.}}\sphinxbfcode{\sphinxupquote{get\_a0r}}}
{\sphinxparam{\DUrole{n}{z}}\sphinxparamcomma \sphinxparam{\DUrole{n}{cosmology}\DUrole{o}{=}\DUrole{default_value}{\{\textquotesingle{}h\textquotesingle{}: 0.6736, \textquotesingle{}omega\_L\textquotesingle{}: 0.6847, \textquotesingle{}omega\_M\textquotesingle{}: 0.3153, \textquotesingle{}zeq\textquotesingle{}: 3402\}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Determine the value a\_0*r from MBW eq. 3.106, for use in cosmological
distance calculations
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{z}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} redshift

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{cosmology}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} cosmological parameters, by default cosmology

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} a0*r {[}kpc{]}

\sphinxlineitem{Raises}
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{AssertionError}} \textendash{} redshift larger than redshift of matter radiation equality

\end{description}\end{quote}

\end{fulllineitems}

\index{luminosity\_distance() (in module cosmology.cosmology)@\spxentry{luminosity\_distance()}\spxextra{in module cosmology.cosmology}}

\begin{fulllineitems}
\phantomsection\label{\detokenize{references/cosmology:cosmology.cosmology.luminosity_distance}}
\pysigstartsignatures
\pysiglinewithargsret
{\sphinxcode{\sphinxupquote{cosmology.cosmology.}}\sphinxbfcode{\sphinxupquote{luminosity\_distance}}}
{\sphinxparam{\DUrole{n}{z}}\sphinxparamcomma \sphinxparam{\DUrole{n}{cosmology}\DUrole{o}{=}\DUrole{default_value}{\{\textquotesingle{}h\textquotesingle{}: 0.6736, \textquotesingle{}omega\_L\textquotesingle{}: 0.6847, \textquotesingle{}omega\_M\textquotesingle{}: 0.3153, \textquotesingle{}zeq\textquotesingle{}: 3402\}}}}
{}
\pysigstopsignatures
\sphinxAtStartPar
Determine the luminosity distance for a flat universe
\begin{quote}\begin{description}
\sphinxlineitem{Parameters}\begin{itemize}
\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{z}} (\sphinxstyleliteralemphasis{\sphinxupquote{float}}) \textendash{} redshift

\item {} 
\sphinxAtStartPar
\sphinxstyleliteralstrong{\sphinxupquote{cosmology}} (\sphinxstyleliteralemphasis{\sphinxupquote{dict}}\sphinxstyleliteralemphasis{\sphinxupquote{, }}\sphinxstyleliteralemphasis{\sphinxupquote{optional}}) \textendash{} cosmological parameters, by default cosmology

\end{itemize}

\sphinxlineitem{Returns}
\sphinxAtStartPar
\sphinxstyleemphasis{float} \textendash{} luminosity distance {[}kpc{]}

\end{description}\end{quote}

\end{fulllineitems}

\subsubsection{Module contents}
\label{\detokenize{references/cosmology:module-cosmology}}\label{\detokenize{references/cosmology:module-contents}}\index{module@\spxentry{module}!cosmology@\spxentry{cosmology}}\index{cosmology@\spxentry{cosmology}!module@\spxentry{module}}
\sphinxstepscope

\subsection{general package}
\label{\detokenize{references/general:general-package}}\label{\detokenize{references/general::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/general:submodules}}

\subsubsection{general.general module}
\label{\detokenize{references/general:general-general-module}}

\subsubsection{general.pygad\_helper module}
\label{\detokenize{references/general:general-pygad-helper-module}}

\subsubsection{general.units module}
\label{\detokenize{references/general:general-units-module}}

\subsubsection{Module contents}
\label{\detokenize{references/general:module-contents}}
\sphinxstepscope

\subsection{initialise package}
\label{\detokenize{references/initialise:initialise-package}}\label{\detokenize{references/initialise::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/initialise:submodules}}

\subsubsection{initialise.GalaxyIC module}
\label{\detokenize{references/initialise:initialise-galaxyic-module}}

\subsubsection{initialise.MergerIC module}
\label{\detokenize{references/initialise:initialise-mergeric-module}}

\subsubsection{initialise.galaxy\_components module}
\label{\detokenize{references/initialise:initialise-galaxy-components-module}}

\subsubsection{initialise.mergers module}
\label{\detokenize{references/initialise:initialise-mergers-module}}

\subsubsection{Module contents}
\label{\detokenize{references/initialise:module-contents}}
\sphinxstepscope

\subsection{literature package}
\label{\detokenize{references/literature:literature-package}}\label{\detokenize{references/literature::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/literature:submodules}}

\subsubsection{literature.LiteratureTables module}
\label{\detokenize{references/literature:literature-literaturetables-module}}

\subsubsection{literature.bh\_bulge module}
\label{\detokenize{references/literature:literature-bh-bulge-module}}

\subsubsection{literature.density\_profiles module}
\label{\detokenize{references/literature:literature-density-profiles-module}}

\subsubsection{literature.dm\_bulge module}
\label{\detokenize{references/literature:literature-dm-bulge-module}}

\subsubsection{literature.radial\_relations module}
\label{\detokenize{references/literature:literature-radial-relations-module}}

\subsubsection{literature.smbh\_recoil module}
\label{\detokenize{references/literature:literature-smbh-recoil-module}}

\subsubsection{literature.smbh\_spins module}
\label{\detokenize{references/literature:literature-smbh-spins-module}}

\subsubsection{Module contents}
\label{\detokenize{references/literature:module-contents}}
\sphinxstepscope

\subsection{mathematics package}
\label{\detokenize{references/mathematics:mathematics-package}}\label{\detokenize{references/mathematics::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/mathematics:submodules}}

\subsubsection{mathematics.array\_operations module}
\label{\detokenize{references/mathematics:mathematics-array-operations-module}}

\subsubsection{mathematics.coordinates module}
\label{\detokenize{references/mathematics:mathematics-coordinates-module}}

\subsubsection{mathematics.geometry module}
\label{\detokenize{references/mathematics:mathematics-geometry-module}}

\subsubsection{mathematics.statistics module}
\label{\detokenize{references/mathematics:mathematics-statistics-module}}

\subsubsection{Module contents}
\label{\detokenize{references/mathematics:module-contents}}
\sphinxstepscope

\subsection{plotting package}
\label{\detokenize{references/plotting:plotting-package}}\label{\detokenize{references/plotting::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/plotting:submodules}}

\subsubsection{plotting.PlotClasses module}
\label{\detokenize{references/plotting:plotting-plotclasses-module}}

\subsubsection{plotting.custom\_cmaps module}
\label{\detokenize{references/plotting:plotting-custom-cmaps-module}}

\subsubsection{plotting.general module}
\label{\detokenize{references/plotting:plotting-general-module}}

\subsubsection{plotting.plot\_utils module}
\label{\detokenize{references/plotting:plotting-plot-utils-module}}

\subsubsection{plotting.specific\_plots module}
\label{\detokenize{references/plotting:plotting-specific-plots-module}}

\subsubsection{Module contents}
\label{\detokenize{references/plotting:module-contents}}
\sphinxstepscope

\subsection{utils package}
\label{\detokenize{references/utils:utils-package}}\label{\detokenize{references/utils::doc}}

\subsubsection{utils.data\_handling module}
\label{\detokenize{references/utils:utils-data-handling-module}}

\subsubsection{utils.dataset\_operations module}
\label{\detokenize{references/utils:utils-dataset-operations-module}}

\subsubsection{utils.get\_cl\_args module}
\label{\detokenize{references/utils:utils-get-cl-args-module}}

\subsubsection{utils.os\_operations module}
\label{\detokenize{references/utils:utils-os-operations-module}}

\subsubsection{utils.param\_io module}
\label{\detokenize{references/utils:utils-param-io-module}}

\subsubsection{Module contents}
\label{\detokenize{references/utils:module-contents}}
\sphinxstepscope

\subsection{visualisation package}
\label{\detokenize{references/visualisation:visualisation-package}}\label{\detokenize{references/visualisation::doc}}

\subsubsection{Submodules}
\label{\detokenize{references/visualisation:submodules}}

\subsubsection{visualisation.animation module}
\label{\detokenize{references/visualisation:visualisation-animation-module}}

\subsubsection{Module contents}
\label{\detokenize{references/visualisation:module-contents}}

\renewcommand{\indexname}{Python Module Index}
\begin{sphinxtheindex}
\let\bigletter\sphinxstyleindexlettergroup
\bigletter{\_}
\item\relax\sphinxstyleindexentry{\_backend}\sphinxstyleindexpageref{references/_backend:\detokenize{module-_backend}}
\item\relax\sphinxstyleindexentry{\_backend.Logging}\sphinxstyleindexpageref{references/_backend:\detokenize{module-_backend.Logging}}
\item\relax\sphinxstyleindexentry{\_backend.States}\sphinxstyleindexpageref{references/_backend:\detokenize{module-_backend.States}}
\indexspace
\bigletter{a}
\item\relax\sphinxstyleindexentry{analysis.masks}\sphinxstyleindexpageref{references/analysis:\detokenize{module-analysis.masks}}
\indexspace
\bigletter{c}
\item\relax\sphinxstyleindexentry{cosmology}\sphinxstyleindexpageref{references/cosmology:\detokenize{module-cosmology}}
\item\relax\sphinxstyleindexentry{cosmology.conversions}\sphinxstyleindexpageref{references/cosmology:\detokenize{module-cosmology.conversions}}
\item\relax\sphinxstyleindexentry{cosmology.cosmology}\sphinxstyleindexpageref{references/cosmology:\detokenize{module-cosmology.cosmology}}
\end{sphinxtheindex}

\renewcommand{\indexname}{Index}
\printindex
\end{document}
